### 2019-06-13 manuscript/manuscript.Rmd
Added ignores, moved ms to manuscript folder using pakillo's template with adjustment for citr

---
title: "Tightly integrated human and machine intelligence within an active learning approach to cropland mapping"

author:

- name: Lyndon D. Estes
  affilnum: '1'

# - name: Francisco Rodriguez-Sanchez 
#   affilnum: 3
#   email: f.rodriguez.sanc@gmail.com

affiliation:

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA

#   
# - affilnum: 2
#   affil: Big Name University, Department of R, City, BN, 01020, USA
#   
# - affilnum: 3
#   affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: no
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library(knitr)
library(citr)

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

\singlespace

\vspace{2mm}\hrule

Write your abstract here.

\vspace{3mm}\hrule

*Keywords*: rmarkdown, reproducible science

\doublespace

\bleft

# Introduction

[@Esteslargeareaspatiallycontinuous2018]

- The problem in general 
    - Africa in particular
    - Types of analyses that depend on cropland maps
- Why cropland mapping is hard
    - 3 problems
        - The target itself and its spectral variability and the spectral variability of its background
            - Spectral and spatial resolution of sensors
            - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also
    - Training data dependence issues
        - Also interact with image resolution issues
        - Increasingly rely on high amounts of training data
        - Human interpretation error (training data quality)
          - Temporal mismatch with classification feedstock (Lesiv paper)
          - Errors in interpreting these sources of imagery
- Review of who does what to solve the problem (probably integrated with problem statement) 
    - croplands.org
- Our approach that tries to solve these three problems
- Within a framework that tightly integrates tightly training data collection
- And handles large-scale computation
- Segmented boundaries

# Methods

- Imagery and image processing methods
- Sample grid
- The system
    - Crowdsourcing platform
          - Training and validation data
          - Accuracy assessment
    - Machine learning
          - Feature engineering
          - ML pipeline Apache Spark
    - Iteration cycle
    - Segmentation

# Results

- Image quality and density
- Labeling results/quality
- Impact of features on accuracy
- Accuracy scores by iteration
- Accuracy by label quality
- Segmentation results 
  - field size class distributions
  - Improvement of accuracy

# Discussion

- Left unquantified is how much training on base maps matters

# Acknowledgement

# References

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file="knitcitations.bib") -->
<!-- ``` -->

<div id = "refs"></div>

\eleft

\clearpage

<!-- \listoftables -->

\newpage

<!-- ```{r Table1, results='asis', echo=FALSE, cache=FALSE} -->
<!-- kable(head(iris), caption = "A glimpse of the famous *Iris* dataset.") -->
<!-- ``` -->

\newpage

<!-- ```{r Table2, results='asis', echo=FALSE, cache=FALSE} -->
<!-- kable(mtcars[10:16, ], caption = "Now a subset of mtcars dataset.") -->
<!-- ``` -->

\clearpage

<!-- \listoffigures -->

\newpage

<!-- ```{r Fig1, echo=FALSE, fig.cap="Just my first figure with a very fantastic caption.", cache=FALSE} -->
<!-- x <- rnorm(100) -->
<!-- y <- jitter(x, 1000) -->
<!-- plot(x, y) -->
<!-- ``` -->

\newpage

\blandscape

<!-- ```{r Fig2, echo=FALSE, fig.cap="Second figure in landscape format.", cache=FALSE} -->
<!-- a <- sort(rnorm(100)) -->
<!-- b <- c(rep("Group Small", 35), rep("Group Big", 65)) -->
<!-- boxplot(a ~ b) -->
<!-- ``` -->

\elandscape

\clearpage

```{r sessioninfo, echo = FALSE, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```

### 2019-06-14 manuscript/manuscript.Rmd
First few sections of introduction

@@
 ---
-title: "Tightly integrated human and machine intelligence within an active learning approach to cropland mapping"
+title: "Improving cropland maps through tight integration of human and machine intelligence"
 
@@
 footerdate: yes
-abstract: no
+abstract: False
 bibliography: 
@@
 \vspace{2mm}\hrule
-
-Write your abstract here.
+\vspace{-20pt}
+# Abstract
+The abstract can go either here or in the YAML
 
@@
 # Introduction
+How much cropland is on the planet, and how much more will be needed to meet humanity's rapidly growing food demands?
+The first part of this question is unclear, as existing estimates tend to vary widely [e.g.
+@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], which makes it substantially harder to answer the second part.
+The reason for these difficulties is because cropland estimates rely heavily on remote sensing-derived landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for an accurate baseline is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015].
 
-[@Esteslargeareaspatiallycontinuous2018]
+<!-- - The problem in general  -->
+<!--     - Africa in particular -->
+<!--     - Types of analyses that depend on cropland maps -->
 
-- The problem in general 
-    - Africa in particular
-    - Types of analyses that depend on cropland maps
-- Why cropland mapping is hard
-    - 3 problems
-        - The target itself and its spectral variability and the spectral variability of its background
-            - Spectral and spatial resolution of sensors
-            - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also
+Cropland mapping over a region such as Africa is difficult for several reasons.
+The primary reason relates to the characteristics of characteristics of the region's smallholder-dominated croplands, where field size averages 1-2 ha [@Debatsgeneralizedcomputervision2016; @FritzMappingglobalcropland2015].
+This size is small relative to the 30-250 m resolution of the sensors typically used for most landcover mapping efforts, which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979], in this case the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
+On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons.
+These latter two aspects poses challenges for the classification algorithms that are applied to the imagery.
+
+These problems arising from cropland characteristics are increasingly being overcome due to technological advances.
+Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
+This spatial and temporal resolution addresses the sensor-scale mismatch, and can capture the intra-seasonal dynamics of cropland more effectively, which helps classifiers more effectively distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017a], the opening of extensive image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are being leveraged to create improved and higher resolution (20-30 m) of cropland and landcover for Africa and globally [@ChenGloballandcover2015; @ESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017]
+
+<!-- , alongside other methodological improvements leading to fusion or synergy maps, which further reduce map error   -->
+
+<!-- - Why cropland mapping is hard -->
+<!--     - 3 problems -->
+<!--         - The target itself and its spectral variability and the spectral variability of its background -->
+<!--             - Spectral and spatial resolution of sensors -->
+<!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->
     - Training data dependence issues
@@
 - Image quality and density
+- Computational time/resources, including interaction time
 - Labeling results/quality
@@
 
-# Acknowledgement
+# Acknowledgements
 
 # References
+\singlespace
 

### 2019-06-17 manuscript/manuscript.Rmd
More paragraphs added to introduction

@@
 Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
-This spatial and temporal resolution addresses the sensor-scale mismatch, and can capture the intra-seasonal dynamics of cropland more effectively, which helps classifiers more effectively distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
-On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017a], the opening of extensive image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are being leveraged to create improved and higher resolution (20-30 m) of cropland and landcover for Africa and globally [@ChenGloballandcover2015; @ESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017]
+This spatial and temporal resolution addresses the sensor-scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017a], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale.
+These capabilities are aleady being used to create a new generation of higher resolution (20-30 m) cropland and landcover maps for Africa and other regions [@ChenGloballandcover2015; @ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
+
+Despite these advances, the highest resolution (<5 m) image sources are still not used to map cropland over large extents because they are commercial and relatively high cost to acquire.
+As such, map accuracy still remains below desired levels, as many of these newer maps still have accuracies in the 60-75% range for the cropland class [e.g.
+@LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
+Accuracy may also be compromised because of factors that are new or becoming more pronounced as a consequence of the recent technological gains, particularly with respect to new algorithms and the manner in which these are trained.
+Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly neural network-based "deep-learning" methods (cite).
+To satisfy this need for more training samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g.
+@XiongNominal30mCropland2017; @ChenGloballandcover2015].
+Several platforms have been developed to facilitate such efforts [e.g.
+@BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012], which provide convenient and highly scalable tools for training data collection.
+Visually interpreted training labels present two particular problems.
+The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019].
+These errors are typically not accounted for in reported accuracy metrics (cite Arthur's paper), thus the degree to which such labelling error impacts map accuracy is unclear.
+However, studies of how land cover map error propagates through subsequent analyses [@Esteslargeareaspatiallycontinuous2018] suggests training data quality could have a large impact.
+The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland.
+Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial image sources that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018].
+This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g.
+Landsat, Sentinel) and selected to represent a single epoch.
+If a land change occurs in the interval between the two image sets (e.g.
+a new field was created), this can introduce error into the training data that is then passed on to the classifier.
+This source of error may be elevated in smallholder-dominated systems, where swidden systems are more comment (cite), or in rapidly developing agricultural frontiers.
+
+We describe here a cropland mapping system that seeks to leverage the recent advances that help increase mapping accuracy over smallholder systems, while minimizing those newer sources of error that are associated with these advances.
 

### 2019-06-18 manuscript/manuscript.Rmd
Introduction rough completed

@@
 On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons.
-These latter two aspects poses challenges for the classification algorithms that are applied to the imagery.
+These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.
 
@@
 
-Despite these advances, the highest resolution (<5 m) image sources are still not used to map cropland over large extents because they are commercial and relatively high cost to acquire.
-As such, map accuracy still remains below desired levels, as many of these newer maps still have accuracies in the 60-75% range for the cropland class [e.g.
+Despite these advances, the highest resolution (<5 m) image sources are still not used to map cropland over very large extents, presumably because they are commercial and relatively high cost to acquire, in addition to the greater computational demands.
+As such, map accuracy can still be a challenge, particular for user's accuracy, which ranges between 46 and 76% for for the cropland class [e.g.
 @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
-Accuracy may also be compromised because of factors that are new or becoming more pronounced as a consequence of the recent technological gains, particularly with respect to new algorithms and the manner in which these are trained.
+
+Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms.
 Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly neural network-based "deep-learning" methods (cite).
-To satisfy this need for more training samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g.
-@XiongNominal30mCropland2017; @ChenGloballandcover2015].
-Several platforms have been developed to facilitate such efforts [e.g.
-@BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012], which provide convenient and highly scalable tools for training data collection.
+To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g.
+@XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019].
+Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g.
+@BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012].
 Visually interpreted training labels present two particular problems.
@@
 These errors are typically not accounted for in reported accuracy metrics (cite Arthur's paper), thus the degree to which such labelling error impacts map accuracy is unclear.
-However, studies of how land cover map error propagates through subsequent analyses [@Esteslargeareaspatiallycontinuous2018] suggests training data quality could have a large impact.
+However, a study of how land cover map error propagates through subsequent analyses [@Esteslargeareaspatiallycontinuous2018] suggests training data quality could have a large impact.
 The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland.
@@
 This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g.
-Landsat, Sentinel) and selected to represent a single epoch.
+Landsat, Sentinel; @XiongNominal30mCropland2017) and often represent a different year.
 If a land change occurs in the interval between the two image sets (e.g.
 a new field was created), this can introduce error into the training data that is then passed on to the classifier.
-This source of error may be elevated in smallholder-dominated systems, where swidden systems are more comment (cite), or in rapidly developing agricultural frontiers.
-
-We describe here a cropland mapping system that seeks to leverage the recent advances that help increase mapping accuracy over smallholder systems, while minimizing those newer sources of error that are associated with these advances.
-
-<!-- , alongside other methodological improvements leading to fusion or synergy maps, which further reduce map error   -->
-
+This source of error may be elevated in smallholder-dominated systems, where swidden practices are common (cite), or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
 <!-- - Why cropland mapping is hard -->
@@
 <!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->
-    - Training data dependence issues
-        - Also interact with image resolution issues
-        - Increasingly rely on high amounts of training data
-        - Human interpretation error (training data quality)
-          - Temporal mismatch with classification feedstock (Lesiv paper)
-          - Errors in interpreting these sources of imagery
-- Review of who does what to solve the problem (probably integrated with problem statement) 
-    - croplands.org
-- Our approach that tries to solve these three problems
-- Within a framework that tightly integrates tightly training data collection
-- And handles large-scale computation
-- Segmented boundaries
+
+To improve the accuracy of cropland maps over smallholder-dominated systems therefore requires an approach that meets three requirements.
+First, it should be based on high spatial and temporal resolution imagery, to be able to capture the fine grain and temporal variability of these fields.
+Second, an algorithm with suitable skill for classifying these images must be selected, and combined with the computational resources needed to process large imagery volumes.
+Third, a method for collecting large volumes of high quality training and validation data based on image interpretation is essential.
+This method should quantify and minimize the errors associated with image interpretation.
+It should also ensure that labels are collected either from the same imagery that is being classified, or from contemporanous imagery, in order to reduce errors introduced by land change processes.
+
+<!-- - Training data dependence issues -->
+<!--     - Also interact with image resolution issues -->
+<!--     - Increasingly rely on high amounts of training data -->
+<!--     - Human interpretation error (training data quality) -->
+<!--       - Temporal mismatch with classification feedstock (Lesiv paper) -->
+<!--       - Errors in interpreting these sources of imagery -->
+<!-- - Review of who does what to solve the problem (probably integrated with problem statement) -->
+<!--   - croplands.org -->
+
+We describe here a cropland mapping system that addresses these requirements.
+The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
+Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016].
+Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is fine enough for humans to be able to discern fields with sufficient confidence [e.g.
+see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009].
+This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3).
+The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016] and was re-engineered to run on high performance, cloud-based computing clusters.
+This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error.
+The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled area after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017].
+Our framework automates this interactive process, which has been shown to be important for improving accuracy in previous studies [e.g.
+@XiongNominal30mCropland2017; others add].
+Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017].
+
+We demonstrate this approach to map cropland in Ghana, a country where agriculture is extensive and dominated by smallholder production.
+
+<!-- - Our approach that tries to solve these three problems -->
+<!-- - Within a framework that tightly integrates tightly training data collection -->
+<!-- - And handles large-scale computation -->
+<!-- - Segmented boundaries -->
 

### 2019-06-21 manuscript/manuscript.Rmd
Paper drafted through imagery section

@@
 
+- name: \*Provisional author list\*
+  affilnum: '1'
+
 - name: Lyndon D. Estes
   affilnum: '1'
+  email: lestes@clarku.edu
 
-# - name: Francisco Rodriguez-Sanchez 
-#   affilnum: 3
-#   email: f.rodriguez.sanc@gmail.com
+- name: Su Ye
+  affilnum: '1'
+
+- name: Lei Song
+  affilnum: '1'
+
+- name: Ron Eastman
+  affilnum: '1'
+
+- name: Sitian Xiong
+  affilnum: '1'
+
+- name: Tammy Woodard
+  affilnum: '1'
+
+- name: Boka Luo
+  affilnum: '1'
+
+- name: Dennis McRitchie
+  affilnum: '2'
+
+- name: Ryan Avery
+  affilnum: '3'
+
+- name: Kelly Caylor
+  affilnum: '3'
+
+- name: Stephanie Debats
+  affilnum: '4'
+
+- name: SpatialCollective
+  affilnum: '5'
+
+- name: Meridia
+  affilnum: '6'
+
+- name: Azavea
+  affilnum: '7'
 
@@
 
-#   
-# - affilnum: 2
-#   affil: Big Name University, Department of R, City, BN, 01020, USA
-#   
-# - affilnum: 3
-#   affil: Estacion Biologica de Donana (EBD-CSIC), E-41092 Sevilla, Spain
+- affilnum: 2
+  affil: Dennis and Sons, Tucson, AZ, USA
+
+- affilnum: 3
+  affil: UCSB
+
+- affilnum: 4
+  affil: Uber
+
+- affilnum: 5
+  affil: SpatialCollective
+
+- affilnum: 5
+  affil: Meridia
+
+- affilnum: 5
+  affil: Azavea
 
@@
 # Introduction
-How much cropland is on the planet, and how much more will be needed to meet humanity's rapidly growing food demands?
-The first part of this question is unclear, as existing estimates tend to vary widely [e.g.
-@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], which makes it substantially harder to answer the second part.
-The reason for these difficulties is because cropland estimates rely heavily on remote sensing-derived landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for an accurate baseline is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015].
+How much cropland is on the planet, and where is it located?
+How much more will be needed to meet humanity's rapidly growing food demands?
+The answer to the first question is unclear, as existing estimates tend to vary widely and cropland maps show substantial spatial disagreements [e.g.
+@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], which makes it difficult to answer the second.
+The first question can't be clearly answered because cropland estimates rely heavily on remote sensing-derived landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for an accurate baseline is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015].
+*[Fill in more reason why cropland maps are important]*
 
@@
 
-We describe here a cropland mapping system that addresses these requirements.
-The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
+We describe here a cropland mapping system that addresses these requirements on the way to creating an object-based map of individual fields The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
 Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016].
-Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is fine enough for humans to be able to discern fields with sufficient confidence [e.g.
+Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g.
 see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009].
@@
 This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error.
-The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled area after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017].
-Our framework automates this interactive process, which has been shown to be important for improving accuracy in previous studies [e.g.
+The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017].
+Our framework automates and formalizes this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g.
 @XiongNominal30mCropland2017; others add].
 Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017].
+Finally, an unsupervised  segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, **resulting in a further sharpening of accuracy** and resulting in a final vector-based field boundary map.
 
-We demonstrate this approach to map cropland in Ghana, a country where agriculture is extensive and dominated by smallholder production.
+We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
+These factors make Ghana emblematic of the substantial challenges facing cropland classification over the broader region.
 
@@
 # Methods
+## System overview 
+The mapping system is centered around four major components: 1) an image acquisition and pre-processing component, 2) a training data collection platform, 3) a machine learning component, and a 4) segmentation component.
+The first component is applied as a once-off step to generate an image catalog necessary for the mapping geography, while the second component initiates the the interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig.
+1).
+The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in the final vectorized fields boundary maps.
 
-- Imagery and image processing methods
-- Sample grid
-- The system
-    - Crowdsourcing platform
+## System components
+### Imagery
+The primary image source used in our system is PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although can be readily adapted to apply to Sentinel and Landsat images.
+PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency.
+Although these images are substantially pre-preprocessed and corrected for atmospheric effects, there are substantial residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], substantial variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016].
+To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.
+
+```{r, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
+The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
+knitr::include_graphics('figures/figure1.png')
+```
+
+PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3).
+Imagery that don't having certain quality flags are excluded (e.g.
+failing to meet minimum spatial precision).
+
+We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a reference grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
+Each 1 degree cell intersecting the mapping geography (Fig.
+2A) is divided into 400 0.05 degree cells (Fig.
+2B), which are the reference cells for image compositing.
+The ARD process then clips the imagery intersecting each 0.05 cell.
+<!-- and then applies the usable data mask (UDM) supplied with the imagery to remove low quality and cloudy pixels.  -->
+
+```{r, echo = FALSE, out.width="100%", fig.cap="Figure scales", fig.align='center'}
+knitr::include_graphics('figures/figure2.png')
+```
+The temporal compositing process is then applied to the ARD imagery, and is done for two time periods.
+In Ghana, a composite is made to cover the dry season period (December-February) and to span the entire wet season (May - September).
+The longer period is necessary for the wet season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area.
+For each image in the compositing process, we pass a 3X3 median filter over the image to remove speckling, and then calculate a weight value based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:
+
+$$
+HOT_{weight} = \frac{1}{(blue - 0.5red)^2}
+$$
+This assigns a low weight to pixels that are cloudy, hazy, or smoky pixels on a particular date.
+We then use the weights to calculate a weight mean for each pixel and each band for the final composite for each time period.
+The composites are then added to the S3 store (Fig.
+1), and a "slippy map" rendering is created for each composite using RasterFoundry (cite).
+This web-based rendering of the image is used in the training data platform (see section X).
+
+### Training data platform
 - Training and validation data
 - Accuracy assessment
-    - Machine learning
+
+### Classification pipeline
 - Feature engineering
 - ML pipeline Apache Spark
-    - Iteration cycle
-    - Segmentation
+
+### Segmentation
+
+## The active learning loop
 

### 2019-07-05 manuscript/manuscript.Rmd
Added section in SI on merging, as well as related text in main manuscript. Added climatology figure for Ghana.

@@
     keep_tex: yes
-    number_sections: no
+    number_sections: yes
     template: manuscript.latex
@@
 footerdate: yes
-abstract: False
+abstract: The abstract can go either here or below
+keywords: rmarkdown, reproducible science
 bibliography: 
@@
 
-\singlespace
+<!-- \singlespace -->
 
-\vspace{2mm}\hrule
-\vspace{-20pt}
-# Abstract
-The abstract can go either here or in the YAML
+<!-- \vspace{2mm}\hrule -->
+<!-- \vspace{-20pt} -->
+<!-- <!-- # Abstract -->
 
-\vspace{3mm}\hrule
+<!-- \vspace{3mm}\hrule -->
 
-*Keywords*: rmarkdown, reproducible science
+<!-- *Keywords*: rmarkdown, reproducible science -->
 
@@
 The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in the final vectorized fields boundary maps.
+Each system component comprises an individual module software hosted on a GitHub repository (see data and software availability section for details).
 
@@
 ### Imagery
+**[Note: This will require updating to reflect changes to compositing methods]**
 The primary image source used in our system is PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although can be readily adapted to apply to Sentinel and Landsat images.
@@
 We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a reference grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
-Each 1 degree cell intersecting the mapping geography (Fig.
+Each 1 degree cell defines the minimum mapping area of interest (Fig.
 2A) is divided into 400 0.05 degree cells (Fig.
-2B), which are the reference cells for image compositing.
-The ARD process then clips the imagery intersecting each 0.05 cell.
+2B), which provides the scale for creating image composite tiles.
+The ARD process simply clips all available imagery intersecting each 0.05 cell to the extent of that cell, regardless of cloud cover.
+
 <!-- and then applies the usable data mask (UDM) supplied with the imagery to remove low quality and cloudy pixels.  -->
 
-```{r, echo = FALSE, out.width="100%", fig.cap="Figure scales", fig.align='center'}
+```{r, echo = FALSE, out.width="100%", fig.cap="An overview of analytical scales used in the mapping platform, including the 1 degree areas of interest defining the minimum mapping geography (A), the 0.05 degree grid used for creating image composite tiles (B), and the 0.005 degree resolution reference grid used for collecting training data and distributed computing (C).", fig.align='center'}
 knitr::include_graphics('figures/figure2.png')
 ```
-The temporal compositing process is then applied to the ARD imagery, and is done for two time periods.
-In Ghana, a composite is made to cover the dry season period (December-February) and to span the entire wet season (May - September).
+The temporal compositing process is applied to the ARD imagery for two time periods, the dry season period (December-February) and the entire wet season (May - September).
 The longer period is necessary for the wet season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area.
-For each image in the compositing process, we pass a 3X3 median filter over the image to remove speckling, and then calculate a weight value based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:
+For each pixel in each image in each ARD temporal stack **we pass a 3X3 median filter over the image to remove speckling**, we calculate two weight values.
+The first is based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:
 
 $$
-HOT_{weight} = \frac{1}{(blue - 0.5red)^2}
+\mathrm{HOT_{weight}} = \frac{1}{(\mathrm{B_1} - 0.5\mathrm{B_3})^2}
 $$
-This assigns a low weight to pixels that are cloudy, hazy, or smoky pixels on a particular date.
-We then use the weights to calculate a weight mean for each pixel and each band for the final composite for each time period.
+Where B$_1$ is the blue band and B$_3$ is the red band.
+This equation results in low values for pixels that are cloudy, hazy, or smoky pixels on a particular date, and high values for clear images.
+The second weight is designed to detect cloud shadow:
+
+$$
+\mathrm{Shadow_{weight}} = \frac{1}{(some values)^4}
+$$
+
+Where X is the median value across the full time series.
+This formula assigns a low weight to pixels falling within cloud shadows.
+Weight values lower than the X for a given date and pixel are assigned to that pixel, replacing the $HOT_{weight}$.
+We then use the resulting to calculate a weighted mean for each pixel for each band for the final composite for each time period.
 The composites are then added to the S3 store (Fig.
@@
 
-### Training data platform
+### Labelling platform
+Both training and reference data are collected by a custom crowdsourcing platform, which refer to here for simplicity's sake as *mapper*.
+The original version of *mapper* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
+The basic structure of the system remains the same, but we have re-engineered it to be a standalone platform that 1) allows us to register and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process.
+The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process.
+Each *mapper* instance is focused on a specific mapping geography (e.g.
+one to several 1 degree cells in Fig.
+2A), thus several instances are set up to run in parallel to cover a larger mapping geography such as Ghana.
+
+Here we provide an overview of *mapper*'s architecture within the context of the labeling workflow within a single instance.
+
+#### Mapping workflow
+##### Selecting training and reference sites
+The labelling process begins with a python script that randomly selects a subset (e.g.
+100) of the 0.005 degree grid (Fig.
+2C) cells that intersect the *mapper* instance's specific area of interest (e.g.
+shaded grey cell in Fig.
+2A).
+These site names are stored in a Postgres table that represents the mapping queue.
+Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.
+
+##### Mapping assignments
+Workers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Fig.
+3).
+There the worker is presented with a white target box representing the 0.005 degree cell that was randomly selected to label, a set of digitizing tools, and different image backdrops, which include false and true color renderings of both the growing season and off-growing season PlanetScope images covering that cell, as well as a basemap provided by Bing maps.
+The worker uses the polygon drawing tool to digitize the boundaries in crop fields visible within the PlanetScope overlays that intersect the grid target, following a set of pre-defined digitizing rules.
+To aid with interpretation, the worker toggles between the various seasonal renderings and the satellite base map to help form a judgement about what constitutes a field within the PlanetScope imagery.
+The worker assigns each completed polygon a label that describes the field class (e.g.
+annual cropland, tree crop, fallow), and after completing all visible fields saves the full set of polygons to the server.
+The worker is then automatically  taken to their next assignment at a different randomly selected site.
+
+```{r, echo = FALSE, out.width="100%", fig.cap="An overview of mapper's interface", fig.align='center'}
+knitr::include_graphics('figures/figure3.png')
+```
+
+##### Processing completed assignments
+Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table.
+The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning.
+If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score as based on the following formula:
+
+$$\mathrm{score_i}=\beta0\mathrm{I}+\beta1\mathrm{O}+\beta2\mathrm{F}+\beta3\mathrm{E}+\beta4\mathrm{C}$$
+
+Where *i* represents the particular assignment, and $\beta_{0-5}$ represent varying weights that sum to 1.
+*I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the worker's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the worker delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the worker's thematic labels (see SI for individual formulae).
+
+Over time, workers are assessed multiple times across a range of accuracy tasks that are selected to represent the variability of the agricultural system being mapped.
+Each worker's score history is averaged to provide an overall assessment of accuracy, and this information is used for creating labels, the second task.
+
+If the worker's completed assignment was a labelling tasks, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers.
+Once complete, an R-based consensus labelling routine is invoked, which uses combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:
+
+$$
+P(\theta|\mathrm{D})=\sum_{i=1}^{n}\mathrm{P}(\mathrm{W_i}|\mathrm{D})\mathrm{P}(\mathrm{\theta}|\mathrm{D}, \mathrm{W_i})
+$$
+Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and *W_$i$* is an individual worker.
+P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) represents the average producer's accuracy for a given worker, and P(W$\theta$|D, *W$_i$*) is the individual worker's label for that pixel.
+This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details of equation).
+
 - Training and validation data

### 2019-07-05 manuscript/si.Rmd
Added section in SI on merging, as well as related text in main manuscript. Added climatology figure for Ghana.

---
title: "Supporting Information"

author:

- name: \*Provisional author list\*
  affilnum: '1'

- name: Lyndon D. Estes
  affilnum: '1'
  email: lestes@clarku.edu

- name: Su Ye
  affilnum: '1'

- name: Lei Song
  affilnum: '1'

- name: Ron Eastman
  affilnum: '1'

- name: Sitian Xiong
  affilnum: '1'

- name: Tammy Woodard
  affilnum: '1'

- name: Boka Luo
  affilnum: '1'

- name: Dennis McRitchie
  affilnum: '2'

- name: Ryan Avery
  affilnum: '3'

- name: Kelly Caylor
  affilnum: '3'

- name: Stephanie Debats
  affilnum: '4'

- name: SpatialCollective
  affilnum: '5'

- name: Meridia
  affilnum: '6'

- name: Azavea
  affilnum: '7'

affiliation:

- affilnum: 1
  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA

- affilnum: 2
  affil: Dennis and Sons, Tucson, AZ, USA

- affilnum: 3
  affil: UCSB

- affilnum: 4
  affil: Uber

- affilnum: 5
  affil: SpatialCollective

- affilnum: 5
  affil: Meridia

- affilnum: 5
  affil: Azavea

output:

  pdf_document:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    keep_tex: yes
    number_sections: yes
    template: manuscript.latex
    includes:
      in_header: header.tex

  html_document: null

  word_document: null

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
  - knitcitations.bib
csl: ecology.csl

---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library(knitr)
library(citr)

#opts_knit$set(root.dir=normalizePath('../'))

### Chunk options: see http://yihui.name/knitr/options/ ###

## Text results
opts_chunk$set(echo = TRUE, warning = TRUE, message = TRUE, include = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "output/figures/")

```

```{r knitcitations, echo=FALSE, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

\singlespace

\bleft
# Methods
## Mapping platform

### Digitizing tools
To minimize the risk of topological errors, *mapper*'s polygon digitizing tools prevent drawing that results in self-intersections and overlaps between adjacent polygons. Upon submission, the PostGIS [`ST_MakeValid`](https://www.postgis.net/docs/ST_MakeValid.html) function is applied to each polygon's geometry to clean remaining topological errors upon insertion into the database.

### Accuracy assessment

### Consensus labelling
As described in the main text, the formula used for creating a consensus label is:

\begin{equation} \label{eq:main}
\mathrm{P(\theta|D)=\sum_{i=1}^{n}P(W_i|D)P(\theta|D, W_i)}
\end{equation}

Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and *W$_i$* is an individual worker.  Looking in greater details at this equation, the first half of the righthand side of the equation, P(W$_i$|D), is the "prior" for worker *i* for the current site based on their history of scores from prior accuracy assessment assignments. The second term, P($\theta$|D, W$_i$), is the probability that the actual class of the pixel in the current assignment is the class that worker *i*'s says that it is, which is either 0 or 1. There are four possible values for this second term:

\begin{equation} \label{eq:tp}
P(\theta = field|D_i = field) = 1
\end{equation}
\begin{equation} \label{eq:fp}
P(\theta = no field|D_i = field) = 0
\end{equation}
\begin{equation} \label{eq:tn}
P(\theta = no field|D_i = no field) = 1
\end{equation}
\begin{equation} \label{eq:fn}
P(\theta = field|D_i = no field) = 0
\end{equation}

Where equations \ref{eq:tp} and \ref{eq:tn} represent true positives and negatives, respectively, and equation \ref{eq:fp} is a false positive, and equation \ref{eq:fn} is a false negative.

Coming back to the first term, the calculation of prior probability can be re-expressed as:

\begin{equation} \label{eq:prior}
\mathrm{P(W_i|D) \approx P(D|W_i)P(W_i)}
\end{equation}

Where:

\begin{equation} 
\mathrm{P(D|W_i) \propto exp\left(-\frac{1}{2}BIC_i\right)}
\end{equation}

With BIC being the Bayesian information criterion:

\begin{equation}
\mathrm{BIC = ln(n)k - 2ln(\hat{L})}
\end{equation}

In which *n* is the sample size, *k* is the number of parameters to estimate, and $\hat{L}$ is the maximum likeihood function. In this case, we are only interested in one parameter (the label that maximizes the likelihood function), thus the BIC becomes:

\begin{equation}
\mathrm{BIC \approx -2ln(\hat{L}) = 2ln(p(D|\hat{\theta}, W))}
\end{equation}

After rearranging, we have:

\begin{equation}
\mathrm{P(D|W_i) \propto p(D|\hat{\theta}, W_i))}
\end{equation}

Which is the worker maximum likelihood, which can be computed as:

\begin{equation} \label{eq:pacc1}
\mathrm{P(\theta = field|\hat{\theta}, M_I) = P(D = field|\theta = field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tp_j}{tp_j + fn_j} \right)}
\end{equation}

\begin{equation} \label{eq:pacc2}
\mathrm{P(\theta = no field|\hat{\theta}, M_I) = P(D = no field|\theta = no field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tn_j}{tn_j + fp_j} \right)}
\end{equation}

Equations \ref{eq:pacc1} and \ref{eq:pacc2} are producer's accuracies, thus the maximum worker likelihood is equivalent to the worker's average producer's accuracy.

<!-- \ref{eq:prior} represents "mapper likelihood" for the current map: -->

The other component of equation \ref{eq:prior}, P(W$_i$), is the worker's average score over *m* accuracy assessment assignment:  
\begin{equation}
\mathrm{P(W_i) \propto \frac{1}{m}\sum_{j=1}^{m}score_j}
\end{equation}

Thus equation \ref{eq:prior} uses two measures of worker accuracy, 1) their overall average accuracy score multipled by 2) their average producer's accuracy to create a *weight* for their individual maps for the given site. Equation \ref{eq:main} becomes:

\begin{equation} 
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(\theta|D, W_i)}{\sum_{i=1}^{n}weight_i}}
\end{equation}

With $\mathrm{P(\theta|D, W_i)}$ being either 0 or 1. In labelling, if the consensus result for a pixel is: $\mathrm{P(\theta = field|D)}$ > 0.5, then we assign that pixel to the field category, otherwise to the no field category.

#### Example

To provide an example of this approach in practice, we'll imagine two workers A and B, each with the following histories:

```{r, echo = FALSE}
library(magrittr)
sc <- tibble::tibble(Worker = LETTERS[1:2], `Prod. Acc. (Field)` = c(0.8, 0.62),
                     `Prod. Acc (no field)` = c(0.81, 0.61), 
                     Score = c(0.75, 0.60))
wtA <- sc[1, ]$Score * sc[1, ]$`Prod. Acc. (Field)`
wtB <- sc[2, ]$Score * sc[2, ]$`Prod. Acc (no field)`

# sc
knitr::kable(sc, align = c("r", "c", "c"), format = "pandoc")
```

In this scenario, worker A thinks that the given pixel falls within a field, and worker B thinks it is not a field.  First, we calculate the weights for each worker:

$$
Weight_A = score_A * PA_A(field) = P(W_A)P(D=field|W_A) = `r sc[1, 2]` * `r sc[1, 4]` = `r wtA`
$$

$$
Weight_B = score_B * PA_B(field) = P(W_B)P(D=no field|W_B) = `r sc[2, 3]` * `r sc[2, 4]` = `r wtB`
$$

And then we plug these weights into the full equation:

$$
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(L = field|D, W_i)}{\sum_{i=1}^{n}weight_i}} = \frac{`r wtA` * 1 + `r wtB` * 0}{`r wtA` + `r wtB`} = `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)`
$$

Since `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)` > 0.5, we label the particular pixel a field.

# References
\singlespace

<div id = "refs"></div>

\eleft

\clearpage

```{r sessioninfo, echo = FALSE, eval = FALSE}
# set eval = FALSE if you don't want this info (useful for reproducibility) to appear 
sessionInfo()
```

### 2019-07-28 manuscript/manuscript.Rmd
Completed draft through methods (but has gaps)

@@
 The longer period is necessary for the wet season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area.
-For each pixel in each image in each ARD temporal stack **we pass a 3X3 median filter over the image to remove speckling**, we calculate two weight values.
-The first is based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:
+For each pixel in each image in each ARD temporal stack, we calculate two weights, with the first being:
 
-$$
-\mathrm{HOT_{weight}} = \frac{1}{(\mathrm{B_1} - 0.5\mathrm{B_3})^2}
-$$
-Where B$_1$ is the blue band and B$_3$ is the red band.
-This equation results in low values for pixels that are cloudy, hazy, or smoky pixels on a particular date, and high values for clear images.
-The second weight is designed to detect cloud shadow:
+<!-- values. The first is based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:  -->
 
-$$
-\mathrm{Shadow_{weight}} = \frac{1}{(some values)^4}
-$$
+\begin{equation} \label{eq:cloud}
+\mathrm{W1_t} = \frac{1}{\mathrm{blue_t}^2}
+\end{equation}
 
-Where X is the median value across the full time series.
-This formula assigns a low weight to pixels falling within cloud shadows.
-Weight values lower than the X for a given date and pixel are assigned to that pixel, replacing the $HOT_{weight}$.
-We then use the resulting to calculate a weighted mean for each pixel for each band for the final composite for each time period.
-The composites are then added to the S3 store (Fig.
-1), and a "slippy map" rendering is created for each composite using RasterFoundry (cite).
+And the second:
+
+<!-- Equation solution from here: https://stackoverflow.com/questions/4027363/two-statements-next-to-curly-brace-in-an-equation -->
+
+\begin{equation} \label{eq:shadow}  
+\mathrm{W2_t} =\begin{cases}
+    1, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_1}$, $\mathrm{NIR_2}$, ..., $\mathrm{NIR_i}$\}}.\\
+    \frac{1}{\mathrm{NIR_t}^4}, & \text{otherwise}.
+  \end{cases}
+\end{equation}
+
+Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
+Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels, while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow.
+
+Once these two weights are calculated, the final composited pixel value for each of the four PlanetScope bands is calculated as:
+
+\begin{equation}
+\mathrm{\bar{B} = \frac{\sum_{t=1}^{T}B_t * W1_t * W2_t}{\sum_{t=1}^{T}W1_t * W2_t}}
+\end{equation}
+
+Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
+The composited tiles are then added to the S3 store (Fig.
+1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using RasterFoundry (cite).
 This web-based rendering of the image is used in the training data platform (see section X).
@@
 
-$$\mathrm{score_i}=\beta0\mathrm{I}+\beta1\mathrm{O}+\beta2\mathrm{F}+\beta3\mathrm{E}+\beta4\mathrm{C}$$
+\begin{equation}
+\mathrm{score_i}=\beta0\mathrm{I}+\beta1\mathrm{O}+\beta2\mathrm{F}+\beta3\mathrm{E}+\beta4\mathrm{C}
+\end{equation}
 
@@
 
-If the worker's completed assignment was a labelling tasks, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers.
-Once complete, an R-based consensus labelling routine is invoked, which uses combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:
+If the worker's completed assignment was a labelling task, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers.
+Once complete, an R-based consensus labelling routine is invoked, which combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:
 
-$$
+\begin{equation}
 P(\theta|\mathrm{D})=\sum_{i=1}^{n}\mathrm{P}(\mathrm{W_i}|\mathrm{D})\mathrm{P}(\mathrm{\theta}|\mathrm{D}, \mathrm{W_i})
-$$
-Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and *W_$i$* is an individual worker.
-P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) represents the average producer's accuracy for a given worker, and P(W$\theta$|D, *W$_i$*) is the individual worker's label for that pixel.
+\end{equation}
+
+Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and $W_i$ is an individual worker.
+P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel.
 This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details of equation).
 
-- Training and validation data
-- Accuracy assessment
+<!-- - Training and validation data -->
+<!-- - Accuracy assessment -->
+
+Once the full set of training sites has been completed, two random draws of pre-determined size (typically **X%** each of the overall number of labels) are made from the randomly selected consensus labels.
+The first set is designated as a model test set, and the final set is set aside as final map validation sample.
+The remaining **Y%** of the labels are used to train the classifier.
 
 ### Classification pipeline
-- Feature engineering
-- ML pipeline Apache Spark
+Upon completion of the labels, *mapper* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of memory-optimized instances (e.g.
+m4.xlarge), depending on the size of the mapping geography.
+
+#### Feature extraction
+The first step in the pipeline is the extraction of additional features from the growing season/non-growing season PlanetScope imagery.
+Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
+We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of a mapping geography orders of magnitude larger than that initial work.
+Here, we used a set of X features, comprised of the eight original composite bands (four from each season), a 5X5 and YXY mean of each band, as well as the slope of each image calculated within the 5X5 and YXY windows (Table X).
+
+\begin{center}Table X. List of features.\end{center}
+
+Feature         Window Size     N Features
+-------------- ---------------- --------------------------------
+RGB-NIR        1X1              8 
+Low-pass       3X3              8
+Slope          3X3              8
+Other          YXY              X
+-------------- ---------------- --------------------------------
+
+Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`.
+These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table X), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data.
+Features are extracted on the fly for each cell in the training and test sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.
+
+#### Classification
+Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of RandomForest [@BreimanRandomForests2001].
+\textcolor{red}{The model is trained with a balanced sample and a tree depth of 16 and total tree number of 30, which initial testing showed to provide a reasonable balance between computational time and performance.}
+
+After fitting, the model is applied to the features of the model test set, and a set of accuracy metrics are calculated, including ordinary binary accuracy and the true skill statistic [@AlloucheAssessingaccuracyspecies2006], which is considered to be more effective than the Kappa statistic in that it controls for class prevalence.
+
+### The active learning loop
+After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process.
+The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training, test, and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand.
+The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:
+
+\begin{equation}
+\mathrm{Q_I = \sum_{I(x, y) \epsilon I} (p(x, y) - 0.5)^2}
+\end{equation}
+
+```{r, echo = FALSE, eval = FALSE}
+uncertainty <- function(x) sum((x - 0.5)^2)
+a <- runif(10, min = 0.4, max = 0.6)
+b <- c(runif(5, min = 0, max = 0.2), runif(5, min = 0.8, max = 1))
+
+# sum(a - 0.5)^2
+# 1 - sum(a - 0.5)^2
+# 1 - sum(b - 0.5)^2
+# ass <- (a - 0.5)^2
+# bss <- (b - 0.5)^2
+# 1 - sum(ass)
+# 1 - sum(bss)
+# (1 - sum(b - 0.5)^2)
+uncertainty(a)
+uncertainty(b)
+```
+Where Q is the uncertainty for cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
+Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus images having the lowest values of $Q_I$ represent sites posing the most difficulty for the classifier.
+
+After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to *mapper*, where workers then digitize new labels.
+These new consensus labels are then added to *learner*'s initial training pool upon completion, and a new machine learning instance is launched and completed, and uncertainty scores for the remaining unlabelled sites are calculaetd, and the most uncertain sites are sent again to the workers for labeling.
+This loop repeats until the mapping skill as assessed with the model test set reaches an asymptote.
 
 ### Segmentation
+After the final iteration, the segmentation algorithm is invoked, which entails several steps.
+In the first step, the meanshift algorithm [@YizongChengMeanshiftmode1995a] is applied to the original bands of the growing season/off-growing season PlanetScope image pair.
 
-## The active learning loop
+In the second step, a Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands [**I assume**], and weighted edge image is computed using weights of 1, 1, and 2 to each band for each season [**for both growing season on and off?**].
+A watershed algorithm (**cite**) is then run on the weighted edge image, with a high level of segmentation specified (**2400** segements per tile).
+
+In the third step, a region adjacency graph is constructed for each tile, in which each node represents all pixels within each polygon created in the previous step.
+The edge between two adjacent regions (polygons) is calculated as the difference between the means of normalized colors of all bands.
+Hierarchical merging is then applied, in which the most similar pairs of adjacent nodes are merged until there are no edges remaining below a predetermined threshold of **X**.
+
+In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (typically 0.5, but is possible to vary locally) are retained as field polygons.
+
+In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Douglas-Peucker algorithm (cite).
+Neighboring polygon that touch along tile boundaries are merged.
 

### 2019-07-29 manuscript/manuscript.Rmd
Added to results. Placeholder Figure 4.

@@
 
+To assess the accuracy of the final segmented boundaries, we used a two-step approach.
+First, we assessed the overall thematic accuracy of the resulting classification against our reserved validation set within each AOI.
+Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to map the given AOI.
+[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy (**cite**), because both an automated and  human-interpreted segmentation worker interpreting imagery cue in on--abrupt, physically detectable breaks within the imagery--are the same.
+On the other hand, no matter how well these boundaries align the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate such an approach against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).
+
 # Results
+We developed a cropland map for **X different [replace with whole country]**, agro-ecologically regions of Ghana (X km$^2$ total) (or whole of Ghana), ranging from the primarily grain and vegetable crop producing regions in the drier northern savanna regions to tree crop dominated regions in the forested southwestern corner of the country, where cocoa and oil palm are among the dominant crops.
+For these latter regions, we did not attempt to classify tree crops, and instead mapping clearing that potentially contain field crops or recently felled or newly replanted tree crops.
+Our rationale for this decision was that the resolution of PlanetScope imagery is insufficient to distinguish such crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.
 
-- Image quality and density
+## Image catalog
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+tiles <- sf::read_sf(system.file("extdata/ghana_tiles.geojson", 
+                                 package = "activemapper"))
+km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
+```
+We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$.
+This entailed downloading all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
+The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile (**Figure Xa**).
+
+We evaluated the overall quality of the resulting composites across the region using a random selection of 100 tiles, which we graded according to an ordinal measure of image accuracy that ranged from 1 (lowest quality: significant residual haze/cloud and image edge artifacts that rendered visual interpretation difficult) to 5 (highest quality, no detectable haze/cloud and no image edge effects; see SI for full interpretation of categories.
+The results showed that the majority of the sampled composites (**X%; Fig Xb**) were rated 4 or higher, with small proportions falling within the lowest two quality categories (**Y and Z%**).
+These lower quality images were confined primarily to the growing season period and the Southwest corner, where the substantially greater cloud cover resulting in a much lower density of available PlanetScope imagery for each time period (**Fig Xc**).
+Overall, the available image density and quality of composites were closely related (see SI for full details).
+
+```{r, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D) [This a hypothetical placeholder figure].", fig.align='center'}
+knitr::include_graphics('figures/figure4.png')
+```
+
 - Computational time/resources, including interaction time
@@
 - Left unquantified is how much training on base maps matters
+- Image quality:
+    - why mean-weighting not median composite (artifacts) (maybe methods)
+    - impact on accuracy
+
+- Assessment of segmentation quality
 

### 2019-08-15 manuscript/manuscript.Rmd
Placeholder figure and text for overall map accuracy.

@@
 
+## The active learning process
+We ran 16 instances to create cropland probability maps for the **X test regions/whole of Ghana**.
+Each instance covered a distinct mapping geography comprised of one or more adjacent AOIs (Figure 5), ranging in extent from 12,160 to 23,535 km$^2$ (average = 15,457 km$^2$), and represented a separate active learning process.
+The average number of iterations per active learning process was **X (range = X-X)**, and the average overall accuracy was **X, with a range of X-X** (Figure 5).
+Overall user's accuracy was slightly **lower/higher** at **X, (range X-X)**, while producer's accuracy averaged **X, (range X-X)**.
+Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southerwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.
+
+<!-- and an accuracy surface interpolated from the average cell-wise accuracies calculated between the predicted map and validation labels.  -->
+
+```{r, echo = FALSE, out.width="80%", fig.cap="The overall, user's, and producer's accuracies (indicated by fill color) for each of the 16 mapping zones (zones numbers are shown on maps) [This a hypothetical placeholder figure].", fig.align='center'}
+knitr::include_graphics('figures/figure5.png')
+```
+
+## Labelling accuracy
+The accuracy patterns generally align with the difficulty of the labelling task, that is, **the southwestern regions were hardest for workers to interpret, as shown by Figure 6.**
+
 - Computational time/resources, including interaction time

### 2019-10-04 manuscript/manuscript.Rmd
Updated manuscript and analyses (reference site selection)

@@
 Cropland mapping over a region such as Africa is difficult for several reasons.
-The primary reason relates to the characteristics of characteristics of the region's smallholder-dominated croplands, where field size averages 1-2 ha [@Debatsgeneralizedcomputervision2016; @FritzMappingglobalcropland2015].
+The primary reason relates to the characteristics of the region's smallholder-dominated croplands, where field size averages 1-2 ha [@Debatsgeneralizedcomputervision2016; @FritzMappingglobalcropland2015].
 This size is small relative to the 30-250 m resolution of the sensors typically used for most landcover mapping efforts, which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979], in this case the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
@@
 PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3).
-Imagery that don't having certain quality flags are excluded (e.g.
-failing to meet minimum spatial precision).
 
-We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a reference grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
+We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
 Each 1 degree cell defines the minimum mapping area of interest (Fig.
@@
 ### Labelling platform
-Both training and reference data are collected by a custom crowdsourcing platform, which refer to here for simplicity's sake as *mapper*.
-The original version of *mapper* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
+Both training and reference data are collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*.
+The original version of *labeller* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
 The basic structure of the system remains the same, but we have re-engineered it to be a standalone platform that 1) allows us to register and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process.
 The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process.
-Each *mapper* instance is focused on a specific mapping geography (e.g.
+Each *labeller* instance is focused on a specific mapping geography (e.g.
 one to several 1 degree cells in Fig.
@@
 
-Here we provide an overview of *mapper*'s architecture within the context of the labeling workflow within a single instance.
+Here we provide an overview of *labeller*'s architecture within the context of the labeling workflow within a single instance.
 
@@
 100) of the 0.005 degree grid (Fig.
-2C) cells that intersect the *mapper* instance's specific area of interest (e.g.
+2C) cells that intersect the *labeller* instance's specific area of interest (e.g.
 shaded grey cell in Fig.
@@
 ##### Processing completed assignments
-Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table.
+Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table.
 The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning.
@@
 
+Once a the full set of sites has been completed, they are randomly partitioned into training and validation sets according to predetermined proportions.
+
 <!-- - Training and validation data -->
@@
 
-Once the full set of training sites has been completed, two random draws of pre-determined size (typically **X%** each of the overall number of labels) are made from the randomly selected consensus labels.
-The first set is designated as a model test set, and the final set is set aside as final map validation sample.
-The remaining **Y%** of the labels are used to train the classifier.
-
 ### Classification pipeline
-Upon completion of the labels, *mapper* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of memory-optimized instances (e.g.
+Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of instances (e.g.
 m4.xlarge), depending on the size of the mapping geography.
@@
 #### Feature extraction
-The first step in the pipeline is the extraction of additional features from the growing season/non-growing season PlanetScope imagery.
+The first step is the extraction of additional features from the growing season/non-growing season PlanetScope imagery.
 Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
 We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of a mapping geography orders of magnitude larger than that initial work.
-Here, we used a set of X features, comprised of the eight original composite bands (four from each season), a 5X5 and YXY mean of each band, as well as the slope of each image calculated within the 5X5 and YXY windows (Table X).
+Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table X).
 
@@
 Feature             Window Size  N Features
--------------- ---------------- --------------------------------
+------------------- ------------ ------------
 RGB-NIR                      1X1            8 
-Low-pass       3X3              8
-Slope          3X3              8
-Other          YXY              X
--------------- ---------------- --------------------------------
+Mean                       11X11            8
+Standard deviation           5X5            8
+------------------ ------------- ------------
 
@@
 These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table X), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data.
-Features are extracted on the fly for each cell in the training and test sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.
+Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.
 
@@
 Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of RandomForest [@BreimanRandomForests2001].
-\textcolor{red}{The model is trained with a balanced sample and a tree depth of 16 and total tree number of 30, which initial testing showed to provide a reasonable balance between computational time and performance.}
+\textcolor{red}{The model is trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.}
 
-After fitting, the model is applied to the features of the model test set, and a set of accuracy metrics are calculated, including ordinary binary accuracy and the true skill statistic [@AlloucheAssessingaccuracyspecies2006], which is considered to be more effective than the Kappa statistic in that it controls for class prevalence.
+After fitting, the model is applied to the features of the model validation set, and a set of accuracy metrics are calculated, including ordinary binary accuracy and the true skill statistic [@AlloucheAssessingaccuracyspecies2006], which is considered to be more effective than the Kappa statistic in that it controls for class prevalence.
 
@@
 After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process.
-The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training, test, and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand.
+The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand.
 The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:
@@
 After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to *mapper*, where workers then digitize new labels.
-These new consensus labels are then added to *learner*'s initial training pool upon completion, and a new machine learning instance is launched and completed, and uncertainty scores for the remaining unlabelled sites are calculaetd, and the most uncertain sites are sent again to the workers for labeling.
-This loop repeats until the mapping skill as assessed with the model test set reaches an asymptote.
+These new consensus labels are then added to *learner*'s initial training pool upon completion, and a new machine learning cluster is launched, resulting in a reassessment of uncertainty for unlabelled sites, a selection from the most uncertain sites, and a labelling of those sites by the workers.
+This loop repeats until the map accuracy metrics, as assessed against the validation set, show no further gains.
 
@@
 To assess the accuracy of the final segmented boundaries, we used a two-step approach.
-First, we assessed the overall thematic accuracy of the resulting classification against our reserved validation set within each AOI.
-Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to map the given AOI.
-[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy (**cite**), because both an automated and  human-interpreted segmentation worker interpreting imagery cue in on--abrupt, physically detectable breaks within the imagery--are the same.
-On the other hand, no matter how well these boundaries align the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate such an approach against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).
+First, we assessed the overall thematic accuracy of the resulting classification against our map reference data.
+Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map.
+[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy (**cite**), because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
+On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).
+
+### Accuracy assessment
+**[Put in somewhere above text about training reference samples]**
+The accuracy assessment procedure described in section 2.2 is used strictly to assess performance gains in the active learning loop, according to suggested best practice guidelines [@ElmesAccountingtrainingdata2019].
+To assess the accuracy of final map products, following these same guidelines we used an independent map reference sample.
+To develop this sample, we first made a random selection of grid cells within each mapping AOI, with the selected number scaled relative to the AOI.
+Next, a primary supervisor examined the image composites within each randomly selected cell and placed three rectangular polygons of ~0.1 ha area in areas that were fully contained within in one of three classes: cropland, non-cropland, potentially cropland or non-cropland.
+In cases where less than three classes were present within the cell, the supervisor placed polygons in proportion to the dominant class.
+For example, in the case where the entire cell was taken up by non-cropland, the supervisor placed all three polygons in non-cropland.
+If two-thirds of the cell was non-cropland, and one-third ambiguous, two polygons would be placed in non-cropland, and one in the ambiguous class.
+After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery.
+The four sets of labels were then combined and the overall agreement between them was calculated.
+This agreement provided the upper bound on knowable map accuracy.
+
+## Application and evaluation of the system
+[Description of map creation process]
+To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each 1X1 degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent 1X1 cells that overlapped with neighboring countries.
+The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.
+
+Primary to initiating active learning in each AOI, we grouped the AOIs into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI).
+Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group.
+At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations resulted in the
+
+To more fully evaluate the capacities of different components of the system, we performed several experiments.
+In the first experiment, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, in addition to those trained with the consensus labels.
+This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy, as recommended by @ElmesAccountingtrainingdata2019.
+Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites.
+Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.
+Finally, we evaluated production performance in terms of speed of labelling and accuracy of results by comparing the generation time (label creation, training, and then classification) and accuracy of maps trained by 1) our core team of 10-12 experienced workers, and 2) a team of 100 novice workers from three different countries.
 
@@
 ## The active learning process
-We ran 16 instances to create cropland probability maps for the **X test regions/whole of Ghana**.
-Each instance covered a distinct mapping geography comprised of one or more adjacent AOIs (Figure 5), ranging in extent from 12,160 to 23,535 km$^2$ (average = 15,457 km$^2$), and represented a separate active learning process.
+**[Fill in with intro]**.
 The average number of iterations per active learning process was **X (range = X-X)**, and the average overall accuracy was **X, with a range of X-X** (Figure 5).

### 2019-10-15 manuscript/manuscript.Rmd
Added more to results section (placeholder text for label strategy accuracy results)

@@
 
-We then transform all images into an "analysis ready data" (ARD) product (cite) that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
+We then transform all images into an "analysis ready data" (ARD) product [@DwyerAnalysisReadyData2018] that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
 Each 1 degree cell defines the minimum mapping area of interest (Fig.
@@
 Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group.
-At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations resulted in the
+At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations targeted new training sites for collection in each AOI based on classification uncertainty.
 
 To more fully evaluate the capacities of different components of the system, we performed several experiments.
-In the first experiment, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, in addition to those trained with the consensus labels.
+In the first experiments, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, in addition to those trained with the consensus labels.
 This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy, as recommended by @ElmesAccountingtrainingdata2019.
@@
 # Results
-We developed a cropland map for **X different [replace with whole country]**, agro-ecologically regions of Ghana (X km$^2$ total) (or whole of Ghana), ranging from the primarily grain and vegetable crop producing regions in the drier northern savanna regions to tree crop dominated regions in the forested southwestern corner of the country, where cocoa and oil palm are among the dominant crops.
+We developed a cropland map covering **X [replace with whole country]**, agro-ecologically regions of Ghana (X km$^2$ total) (or whole of Ghana), ranging from the primarily grain and vegetable crop producing regions in the drier northern savanna regions to tree crop dominated regions in the forested southwestern corner of the country, where cocoa and oil palm are among the dominant crops.
 For these latter regions, we did not attempt to classify tree crops, and instead mapping clearing that potentially contain field crops or recently felled or newly replanted tree crops.
@@
 
-We evaluated the overall quality of the resulting composites across the region using a random selection of 100 tiles, which we graded according to an ordinal measure of image accuracy that ranged from 1 (lowest quality: significant residual haze/cloud and image edge artifacts that rendered visual interpretation difficult) to 5 (highest quality, no detectable haze/cloud and no image edge effects; see SI for full interpretation of categories.
-The results showed that the majority of the sampled composites (**X%; Fig Xb**) were rated 4 or higher, with small proportions falling within the lowest two quality categories (**Y and Z%**).
-These lower quality images were confined primarily to the growing season period and the Southwest corner, where the substantially greater cloud cover resulting in a much lower density of available PlanetScope imagery for each time period (**Fig Xc**).
+We evaluated the overall quality of the resulting composites across the region using a random selection of 100 tiles, which we graded according to a four category quality that evaluated the degree of residual 1) cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image in which resolution was degraded below the 3-4 m PlanetScope resolution, due to between-date image mis-registrations.
+Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), thus the highest possible score was 12.
+The results showed that the majority of the sampled composites (**X%; Fig Xb**) received a score of **X** or higher (**Y and Z%**), as rated by four independent observers (mean standard deviation of observer scores was **XX**).
+Image quality varied spatially and temporally, with the lower quality images found during in the growing season period and the Southwest corner, where the substantially greater cloud cover resulting in a much lower density of available PlanetScope imagery for each time period (**Fig Xc**).
 Overall, the available image density and quality of composites were closely related (see SI for full details).
@@
 **[Fill in with intro]**.
-The average number of iterations per active learning process was **X (range = X-X)**, and the average overall accuracy was **X, with a range of X-X** (Figure 5).
+The average number of iterations per active learning process was **3**, as overall performance gains, as assessed against validation sets, saturated quickly (see **Figure X** in SI).
+
+The average computational time per iteration per AOI was *6 hours*. **Training time. Interaction time**
+
+## Map accuracy
+To assess the overall accuracy of the final maps, we used our reference set to evaluate the accuracy of classifications derived from thresholding the posterior probabilities produced by the RandomForest model, as well as the segmented field boundary maps.
+First, we evaluated the inter-rater agreement between reference supervisors in labeling reference polygons.
+The average inter-rater agreement was **X%, with a standard deviation of Y% (see Figure X in SI)**, which provides the upper bound on achievable map accuracy.
+
+Based on this reference sample, we found that the average accuracy of the pixel-wise classifications was **X, with a range of X-X** (Figure 5).
 Overall user's accuracy was slightly **lower/higher** at **X, (range X-X)**, while producer's accuracy averaged **X, (range X-X)**.
-Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southerwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.
+Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.
+The segmentation process **increased** accuracy slightly, with a mean overall, user's and producer's accuracy of **X (range X-X), Y (range X-X), and Z (range X-X) percent**.
 
@@
 
-## Labelling accuracy
-The accuracy patterns generally align with the difficulty of the labelling task, that is, **the southwestern regions were hardest for workers to interpret, as shown by Figure 6.**
+The accuracy patterns generally align with the difficulty of the labelling task, that is, **the southwestern regions were hardest for workers to interpret, as shown by Figure 5.**
 
-- Computational time/resources, including interaction time
+## System evaluation/experiment results
+### Label quality and map accuracy
+We used the reference labels to evaluate the accuracy of maps generated with the least and most accurate workers' labels, and then compared the resulting scores to accuracy assessed using the consensus labels.
+On average, there was a substantial difference between maps trained with the least and most accurate workers' labels, with the former being on average **X percentage points (range Y-Y%)** more accurate (**Figure 6**).
+Maps generated with consensus labels were on average **X percentage points (range Y-Y%)** more accurate than those created with the most accurate workers' labels.
+
+```{r, echo = FALSE, out.height = "3in", out.width="5in", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.
+[This a hypothetical placeholder figure].", fig.align='center', message=FALSE}
+# knitr::include_graphics('figures/figure6.png')
+library(tidyverse)
+library(cowplot)
+set.seed(1)
+acc_tb <- tibble(aoi = c(1:9, 12, 15), 
+                 Consensus = runif(11, min = 0.75, max = 0.85),
+                 Highest = runif(11, min = 0.72, max = 0.8),
+                 Least = runif(11, min = 0.65, max = 0.75))
+acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi)
+ggplot(acc_tb) + 
+  geom_boxplot(aes(x = type, y = score), fill = "grey") + 
+  xlab("Labelling strategy") + ylab("Accuracy") + 
+  theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3))
+```
+
+To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team.
+Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an **X%** agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average **X% smaller/larger** than those mapped on the ground.
+**However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but maybe instead be at least partially an indication of how different the perspectives as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
+
+###
+
 - Labeling results/quality
-- Impact of features on accuracy
-- Accuracy scores by iteration
-- Accuracy by label quality
 - Segmentation results 

### 2020-01-07 manuscript/manuscript.Rmd
Added labelling strategy text to manuscript.

@@
 
-Primary to initiating active learning in each AOI, we grouped the AOIs into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI).
+Prior to initiating active learning in each AOI, we grouped the AOIs into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI).
 Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group.
@@
 Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.
-Finally, we evaluated production performance in terms of speed of labelling and accuracy of results by comparing the generation time (label creation, training, and then classification) and accuracy of maps trained by 1) our core team of 10-12 experienced workers, and 2) a team of 100 novice workers from three different countries.
+
+<!-- Finally, we evaluated production performance in terms of speed of labelling and accuracy of results by comparing the generation time (label creation, training, and then classification) and accuracy of maps trained by 1) our core team of 10-12 experienced workers, and 2) a team of 100 novice workers from three different countries.
+-->
 
 # Results
-We developed a cropland map covering **X [replace with whole country]**, agro-ecologically regions of Ghana (X km$^2$ total) (or whole of Ghana), ranging from the primarily grain and vegetable crop producing regions in the drier northern savanna regions to tree crop dominated regions in the forested southwestern corner of the country, where cocoa and oil palm are among the dominant crops.
-For these latter regions, we did not attempt to classify tree crops, and instead mapping clearing that potentially contain field crops or recently felled or newly replanted tree crops.
-Our rationale for this decision was that the resolution of PlanetScope imagery is insufficient to distinguish such crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.
-
-## Image catalog
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
+We developed maps for all of Ghana (`r km2` km$^2$).
+Ghana's has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the drier northern savanna regions to tree crop dominated regions in the forested southwestern corner of the country, where cocoa and oil palm are among the dominant crops.
+For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or recently felled or newly replanted tree crops.
+Our rationale for this decision was that the resolution of PlanetScope imagery is insufficient to distinguish such crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.
+
+## Image catalog
+
 We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$.
@@
 **[Fill in with intro]**.
-The average number of iterations per active learning process was **3**, as overall performance gains, as assessed against validation sets, saturated quickly (see **Figure X** in SI).
+The average number of iterations per active learning process was **4**, as overall performance gains, as assessed against validation sets, saturated quickly (see **Figure X** in SI).
 
@@
 Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.
-The segmentation process **increased** accuracy slightly, with a mean overall, user's and producer's accuracy of **X (range X-X), Y (range X-X), and Z (range X-X) percent**.
+The segmentation process **increased/decreased** accuracy slightly, with a mean overall, user's and producer's accuracy of **X (range X-X), Y (range X-X), and Z (range X-X) percent**.
 
@@
 ### Label quality and map accuracy
-We used the reference labels to evaluate the accuracy of maps generated with the least and most accurate workers' labels, and then compared the resulting scores to accuracy assessed using the consensus labels.
-On average, there was a substantial difference between maps trained with the least and most accurate workers' labels, with the former being on average **X percentage points (range Y-Y%)** more accurate (**Figure 6**).
-Maps generated with consensus labels were on average **X percentage points (range Y-Y%)** more accurate than those created with the most accurate workers' labels.
+```{r, echo=FALSE}
+load(here::here("external/data/label_experiment_metrics.rda"))
+consensus <- scores %>% filter(Label == "Consensus") %>% 
+  select(AUC, Accuracy, TSS) 
+high <- scores %>% filter(Label == "High") %>% select(AUC, Accuracy, TSS) 
+low <- scores %>% filter(Label == "Low") %>% select(AUC, Accuracy, TSS)
+auc <- high$AUC / low$AUC - 1
+tss <- (high$TSS + 1) / (low$TSS + 1) - 1
+acc <- high$Accuracy / low$Accuracy - 1
+cauc <- consensus$AUC / high$AUC - 1
+ctss <- (consensus$TSS + 1) / (high$TSS + 1) - 1
+cacc <- consensus$Accuracy / high$Accuracy - 1
+```
 
-```{r, echo = FALSE, out.height = "3in", out.width="5in", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.
+We used the validation labels to evaluate the accuracy of maps generated in three AOIs (1, 8, and 15; **Figure X**) with the least and most accurate workers' labels, and then compared the resulting scores to accuracy assessed using the consensus labels (**Figure 6**).
+On average, there was a substantial difference between maps trained with the most and least accurate workers' labels, with the former being on average `r round(mean(acc) * 100, 1)` percent higher (range `r round(min(acc) * 100, 1)`-`r round(max(acc) * 100, 1)`%) in terms of accuracy, `r round(mean(tss) * 100, 1)` percent higher (range `r round(min(tss) * 100, 1)`-`r round(max(tss) * 100, 1)`%) in skill (measured with TSS, which ranges from -1 to 1).
+Maps generated with consensus labels were on average `r round(mean(cacc) * 100, 1)` percent more accurate (range `r round(min(cacc) * 100, 1)`-`r round(max(cacc) * 100, 1)`%) and `r round(mean(ctss) * 100, 1)` (range `r round(min(ctss) * 100, 1)`-`r round(max(ctss) * 100, 1)`% more skillful.
+AUC scores showed similar average differences to TSS with respect to both comparisons; on average `r round(mean(auc) * 100, 1)` higher when comparing maps generated with the most and least accurate labels, and `r round(mean(cauc) * 100, 1)` percent more skillful when comparing maps genereated with consensus labels compared to those with the most accurate worker's labels.
+Although these comparisons were assessed against consensus-generated validation labels, a visual assessment of probability images generated from each labelling strategy shows markedly greater uncertainty between consensus and high accuracy label-generated maps relative to those generated with low accuracy (**see Figure SX in SI**).
+
+```{r, echo = FALSE, out.height = "4in", out.width="7in", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.
 [This a hypothetical placeholder figure].", fig.align='center', message=FALSE}
-# knitr::include_graphics('figures/figure6.png')
-library(tidyverse)
-library(cowplot)
-set.seed(1)
-acc_tb <- tibble(aoi = c(1:9, 12, 15), 
-                 Consensus = runif(11, min = 0.75, max = 0.85),
-                 Highest = runif(11, min = 0.72, max = 0.8),
-                 Least = runif(11, min = 0.65, max = 0.75))
-acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi)
-ggplot(acc_tb) + 
-  geom_boxplot(aes(x = type, y = score), fill = "grey") + 
-  xlab("Labelling strategy") + ylab("Accuracy") + 
-  theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3))
+knitr::include_graphics('figures/figure6.png')
+# library(tidyverse)
+# library(cowplot)
+# set.seed(1)
+# acc_tb <- tibble(aoi = c(1:9, 12, 15), 
+#                  Consensus = runif(11, min = 0.75, max = 0.85),
+#                  Highest = runif(11, min = 0.72, max = 0.8),
+#                  Least = runif(11, min = 0.65, max = 0.75))
+# acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi)
+# ggplot(acc_tb) + 
+#   geom_boxplot(aes(x = type, y = score), fill = "grey") + 
+#   xlab("Labelling strategy") + ylab("Accuracy") + 
+#   theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3))
 ```
@@
 Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an **X%** agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average **X% smaller/larger** than those mapped on the ground.
-**However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but maybe instead be at least partially an indication of how different the perspectives as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
+**However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but may instead be at least partially an indication of how different the perspectives are as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
 

### 2020-01-09 manuscript/manuscript.Rmd
Added results to main ms for labeling strategy; added figure for corresponding probability images to SI

@@
 This spatial and temporal resolution addresses the sensor-scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
-On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017a], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale.
+On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale.
 These capabilities are aleady being used to create a new generation of higher resolution (20-30 m) cropland and landcover maps for Africa and other regions [@ChenGloballandcover2015; @ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
@@
 ### Classification pipeline
-Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of instances (e.g.
-m4.xlarge), depending on the size of the mapping geography.
+Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of instances, depending on the size of the mapping geography.
+We identifed an optimal cluster of 50-60 instances (m4.2xlarge) each having 8 virtual CPUs and 30 GB RAM, which ensured minimal idle time for any single node.
 
@@
 Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites.
-Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.
+
+<!-- Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.  -->
 
@@
 km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
+km2 <- format(km2, big.mark = ",", scientific = FALSE)
 ```
 We developed maps for all of Ghana (`r km2` km$^2$).
-Ghana's has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the drier northern savanna regions to tree crop dominated regions in the forested southwestern corner of the country, where cocoa and oil palm are among the dominant crops.
+Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
 For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or recently felled or newly replanted tree crops.
@@
 
-## System evaluation/experiment results
+## System evaluation
 ### Label quality and map accuracy
@@
 
-```{r, echo = FALSE, out.height = "4in", out.width="7in", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.
-[This a hypothetical placeholder figure].", fig.align='center', message=FALSE}
+```{r, echo = FALSE, out.width="100%", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/figure6.png')
@@
 
-###
+### Is active learning any better than random? 
+To evaluate whether active learning is more effective than a purely random approach to training site selection, we retrained the RandomForest model on the same three AOIs (1, 8, 15) using sites that were selected purely at random for all iterations after the initial training draw.
+After the mapping team labelled these sites, we subsequently retrained the models with each increment of randomly selected sites, and evaluated prediction performance against the original validation set.
 

### 2020-01-09 manuscript/si.Rmd
Added results to main ms for labeling strategy; added figure for corresponding probability images to SI

@@
 
+\clearpage
+# Results
+
+```{r, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_hml_probs.png')
+```
+
 # References

### 2020-03-05 manuscript/manuscript.Rmd
Recent commits

@@
 - Assessment of segmentation quality
+- Relevant literature: 
+    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values).
+Semi-automated approach to generating reference data from imagery (based on phenology)
+    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
+    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.
 

### 2020-06-25 manuscript/manuscript.Rmd
Updates to manuscript, addition of map reference sample framework to SI, and selection of map reference sample in analyses

@@
 library(citr)
+library(dplyr)
 
@@
 
-We describe here a cropland mapping system that addresses these requirements on the way to creating an object-based map of individual fields The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
+We describe here a cropland mapping system that addresses these requirements on the way to creating an object-based map of individual fields.
+The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
 Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016].
@@
 The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning.
-If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score as based on the following formula:
+If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score based on the following formula:
 
@@
 We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$.
-This entailed downloading all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
+This entailed processing all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
+For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
 The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile (**Figure Xa**).
 
-We evaluated the overall quality of the resulting composites across the region using a random selection of 100 tiles, which we graded according to a four category quality that evaluated the degree of residual 1) cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image in which resolution was degraded below the 3-4 m PlanetScope resolution, due to between-date image mis-registrations.
+We evaluated the overall quality of the resulting composites across the region using a random selection of 100 tiles, which we graded according to a four category quality score that evaluated the degree of residual 1) cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image in which resolution was degraded below the 3-4 m PlanetScope resolution, due to between-date image mis-registrations.
 Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), thus the highest possible score was 12.
@@
 
-- Labeling results/quality
-- Segmentation results 
+### Segmentation quality
 - field size class distributions
@@
 
+## Ghana's croplands
+- Average field size
+- Cropland percentage
+- Comparison against other products (croplands.org; Sentinel Landcover)
+
 # Discussion
-
 - Left unquantified is how much training on base maps matters
@@
     - impact on accuracy
-
 - Assessment of segmentation quality
+- Next steps: 
+    - Annually updated cropland maps
 - Relevant literature: 

### 2020-06-25 manuscript/si.Rmd
Updates to manuscript, addition of map reference sample framework to SI, and selection of map reference sample in analyses

@@
 
+#### Protocol
+Following Stehman and Foody (2019)
+
+- Sample design
+  - Randomization approach in sample selection protocol
+  - "Specify inclusion probabilities or the information needed to derive the inclusion probabilities"
+  - If stratified sampling design chosen, provide:
+    - details of how strata were created
+    - details of sampling design within each strata
+      - Random within each
+    - what sample size was in each stratum
+      - Equal samples in agricultural and non-agricultural classes
+  - Register sample: 
+    - specify locations originally selected
+    - if actual sample site moved, give rationale (find areas fully containing class within sampling unit)
+    - specify which sites weren't sampled due to problems (e.g. missing data)
+  - Desired Confidence Interval (user's accuracy only)
+
+$$n = \frac{z^2p(1-p)}{d^2}$$
+
+      - p = estimated probability
+      - d = desired half-width
+
+- Response design
+  - Definition of spatial assessment
+  - Definition of classes
+  - Reference source
+  - Specific information collected from each source
+  - Rules for reference class labels
+  - How agreement between map and reference classes is defined
+
 ### Consensus labelling

### 2020-07-13 manuscript/manuscript.Rmd
Updated map reference sample, some changes to SI, moving map reference sample selection out of analyses to own vignette

@@
     - Annually updated cropland maps
+    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
 - Relevant literature: 

### 2020-07-13 manuscript/si.Rmd
Updated map reference sample, some changes to SI, moving map reference sample selection out of analyses to own vignette

--- 2020-01-09 manuscript/si.Rmd
@@
 geometry: margin=1in
-linenumbers: yes
+linenumbers: false
 spacing: doublespacing
@@
 ```{r setup, include=FALSE, cache=FALSE, message = FALSE}
-
 library(knitr)
 library(citr)
+library(tidyverse)
+library(sf)
 
@@
 Upon submission, the PostGIS [`ST_MakeValid`](https://www.postgis.net/docs/ST_MakeValid.html) function is applied to each polygon's geometry to clean remaining topological errors upon insertion into the database.
-
-### Accuracy assessment
 
@@
 
+### Accuracy assessment
+We designed and implemented a map accuracy assessment protocol following procedures summarized by @StehmanKeyissuesrigorous2019.
+This entailed the creation of a map reference sample, which first entailed designing a sample, and then designing how the sample response would be collected.
+
+#### Map reference sample design
+A reference sample should ideally have some randomization involved in the sample selection protocol, and the probabilities for a sample being included should be provided.
+There are multiple sample designs that can be chosen, but we employed a stratified sampling design, using the segmented field boundaries to define the strata for field/non-field.
+To create the sample, we first extracted the centroids for each of the sample grid cells in Ghana.
+We then intersected the centroid points with the field segments, assigned a class of 1 (cropland) where points intersected a field, and 0 where they didn't (non-cropland).
+We then removed from this set of points all those that corresponded to model training, validation, or training reference sites.
+We then extracted a random sample from both the cropland and non-cropland points.
+To determine the sample size of each, we specified a desired confidence interval using the following formula [@StehmanKeyissuesrigorous2019]:
+
+$$n = \frac{z^2p(1-p)}{d^2}$$
+
+Where p is the estimated probability (or mapped class accuracy), and d is the size of the margin of error (1/2 the confidence interval).
+We selected a d value of 0.03 and assumed that the user's accuracy of the field class would be 0.75 and that of the non-cropland class would be 0.8, returning sample sizes of 800 and 683, respectively.
+The distribution of the resulting map reference sample is shown in Figure 1.
+
+```{r, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
+africa <- st_read(
+  system.file("extdata/africa.geojson", package = "activemapper"), quiet = TRUE,
+)
+aois <- st_read(system.file("extdata/aois.geojson", package = "activemapper"), 
+                quiet = TRUE)
+africa_cont <- africa %>% st_union %>% st_sf
+ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)
+
+ref_sample <- readr::read_csv(
+  here::here("external/reference/map_reference_sample.csv")
+)
+ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)
+
+p <- ggplot() + 
+  # geom_sf(data = ghana) +
+  geom_sf(data = aois) + 
+  geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) +
+  scale_color_manual(values = c("red", "blue"), 
+                     labels = c("Non-cropland", "Cropland"), name = "") + 
+  theme_light()
+p
+```
+
+#### Response design
+##### Stage 1
+We collected the map reference sample on a separate instance of `labeller` set up for the purpose (`validator.crowdmapper.org`).
+For the sampling unit, we selected a rectangular polygon of ~0.1 ha (0.0002866$^\circ$ resolution), centered on the centroid of each grid cell selected for map reference sample.
+Four classes were established for the validation: **cropland**, **non-cropland**, **uncertain but likely cropland**, and **uncertain but likely non-cropland**.
+The latter two classes were designed to capture information related to swidden dynamics, following the rationale that uncertainty and time since last cropping are likely to be positively correlated.
+This uncertainty also captures information about the inherent difficulty of the mapping task.
+Samples were collected by visually interpreting the overlays of PlanetScope composites presented in `labeller`, following the same interpretation protocols used by the labelling team, with the exception that the polygons placed were square and of 0.0002866$^\circ$ resolution.
+
+The map reference sample was first placed and defined in an initial stage.
+In this stage, the class of the initial square polygon was determined, and then the polygon was adjusted if it overlapped more than one class.
+In sites where the initial map reference grid fell on one of the two uncertain classes, a second polygon was placed in the nearest location where a definite cropland or non-cropland interpretation could be made, with the choice determined by whether the uncertain class was more likely cropland or not cropland.
+The uncertain class was only used for providing insight into fallow cycles, while the second, more certain polygon was used in a second stage of the map reference sample, which was given to multiple independent interpreters (see below).
+
+The following set of rules were followed in this first, sample placing phase (undertaken by Lyndon Estes):
+
+1. When determining the class corresponding with the initial location of the target grid, IF:
+  - More than half of the target falls within what appears to be a clear arable crop field, then assign it to **cropland**. 
+  - More than half falls in what is clearly not a field, then assign it to **non-cropland**. 
+  - More than half falls in a location where it is harder to tell whether it is cropland or non-cropland, determine whether it is more likely a crop field or not a crop field, and then assign either **uncertain but likely cropland**, or **uncertain but likely non-cropland**.
+
+2. After determining the class, IF:
+  a. The target polygon is contained entirely within a single clear class AND
+      - The class is either **cropland** or **non-cropland** THEN:
+        - Digitize a point within the center of the target box, assign appropriate class, and complete the assignment
+      - The class belongs to one of the two **uncertain** classes THEN:
+
+        - Digitize a square aligned exactly with the initial target polygon, and choose the class **cropland** or **non-cropland** (following the most likely interpretation) and move the new polygon to the nearest location where it can be contained entirely within that clear **cropland** or **non-cropland** class;
+
+        - Digitize a point in the middle of the target polygon, and assign it the appropriate label: **cropland** or **non-cropland**.
+
+  b. The target polygon partially overlaps a second class AND:
+      - That class is either **cropland** or **non-cropland** THEN:
+
+        - Digitize a square aligned exactly with the target polygon, assign the appropriate class, and move the box to the nearest location (shortest possible distance) where it can be contained entirely within that same class.
+
+      - That class is **uncertain** THEN:
+
+        - Digitize a square aligned exactly with the target polygon, assign the appropriate **uncertain** class, then move the box to the nearest location (shortest possible distance) where it can be contained entirely within that same **uncertain** class.
+        - Digitize a second square aligned exactly with the initial target polygon, and choose the class **cropland** or **non-cropland** (following the most likely interpretation) and move the new polygon to the nearest location where it can be contained entirely within that clear **cropland** or **non-cropland** class.
+
+We then 
+  - Definition of spatial assessment
+  - Definition of classes
+  - Reference source
+  - Specific information collected from each source
+  - Rules for reference class labels
+  - How agreement between map and reference classes is defined
+
+##### Stage 2
+
 \clearpage

### 2020-07-17 manuscript/si.Rmd
Updates to analyses, SI, etc. Starting to reorganize analytical code.

@@
 # Methods
+## Assessing quality of composites
+The rubric presented in Table 1 was used to assess the quality of the seasonal image composites.
+The imagery was evaluated by examining their Raster Foundry overlays within a `labeller` instance set up for the purpose.
+
+```{r, echo=FALSE, message=FALSE}
+library(kableExtra)
+image_qual_criteria <- tibble(
+  `Quality dimension` = c(
+    "Percent affected by residual cloud", "Percent affected by cloud shadow", 
+    "Number of visible scene boundaries", "Percent blurred"
+  ),
+  `3 pts` = c("<1%", "<1%", "None", "None"), 
+  `2 pts` = c("1-5%", "1-5%", "1", "<20%"), 
+  `1 pts` = c("5-10%", "5-10%", "2-3", "20-50%"),
+  `0 pts` = c(">10%", ">10%", ">3", ">50%")
+) 
+cap <- glue::glue("Four dimensions used to assess the assess the quality
+                  of the temporally composited image tiles, including the
+                  criteria used to award points for scoring each dimension.")
+kable(image_qual_criteria, format = "latex", booktabs = "T", caption = cap) %>% 
+  kable_styling(full_width = TRUE, latex_options = c("hold_position")) %>% 
+  column_spec(1, width = "8cm")
+```
+
 ## Mapping platform
@@
 # Results
+## Image quality
 

### 2020-07-24 inst/manuscript/analyses-mapreference-sample.Rmd
Substantial updates to analyses and organization. Manuscript, figures notebook, and analyses moved to inst/manuscript. Added scripts to data-raw/ for processing inputs into figures, headed to instances, etc (much of this is moving code out of analyses* notebooks).

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(sf)
library(dplyr)
library(ggplot2)
library(aws.s3)
```

### Base Datasets

The pre-preparation of base datasets that install with the `activemapper` package is shown in `external/scripts/data_setup.R`. These are loaded in here.

```{r, eval=FALSE}
aois <- st_read(
  system.file("extdata/aois.geojson", package = "activemapper")
)
# ggplot(data = tiles) + 
#   geom_sf(aes(fill = as.factor(aoi1)), col = "transparent")
```

## Map reference site selection

Use formula from Stehman and Foody (2019) to estimate desired sample size for different classes, using range of expected user's accuracies per class
```{r, fig.height=4, fig.width=4, fig.align="center"}
fdensity <- raster(here("external/data/field_density005.tif")) 
fdensity[fdensity > 1] <- 1

fpoints <- data.table::fread(here("external/data/field_points005.csv")) %>% 
  as_tibble %>% filter(aoi != "3N")

# proportion of Ghana that is cropland, based on density
dropna <- function(x) x[!is.na(x)]
v <- dropna(values(fdensity))
fcropland <- sum(v) / length(v)  # 21.5%
fcropland2 <- (fpoints %>% filter(class == 1) %>% nrow()) / nrow(fpoints)

# sample size function
ref_sample_size <- function(p, z, d) {
  n <- (z^2 * p * (1 - p)) / d^2
  data.frame(p = p, n = n)
}

# sample size
sample_size <- ref_sample_size(seq(0.6, 0.9, 0.05), 1.96, 0.03)

p1 <- ggplot(sample_size) + geom_point(aes(p, n)) + 
  geom_text(aes(p, n, label = round(n)), nudge_x = 0.017) + theme_bw()
p1
```

Remove all points that were used for training, training referenec, or  validation. Contained in CSV of sites prepared in `data_prep.R`
```{r, eval = FALSE}
train_val <- readr::read_csv(here("external/data/train_val_sites.csv"))
fpointsr <- fpoints %>% filter(!id %in% unique(train_val$id))
```

Going with sample size of 800 for cropland class, and 683 for non-cropland, per the above.  MOE of 0.03 and 95% confidence interval on assumed cropland 
```{r, eval=FALSE}
# Take weighted average
set.seed(111)
fsample <- fpointsr %>% filter(class == 1) %>% sample_n(size = 800)
set.seed(111)
nofsample <- fpointsr %>% filter(class == 0) %>% sample_n(size = 683)

ref_sample <- bind_rows(fsample, nofsample) %>% 
  mutate(class = as.character(class))

# process into format for labeller, needing name and id
mgrid_tb <- data.table::fread(
  system.file("extdata/ghana_grid.csv", package = "activemapper")
) %>% as_tibble()

ref_sample <- left_join(ref_sample %>% select(id, aoi, class), mgrid_tb, 
                        by = "id") 
data.table::fwrite(
  ref_sample, file = here("external/reference/map_reference_sample.csv")
)

ref_sample_labeller <- ref_sample %>% select(name) %>% 
  mutate(run = 0, iteration = 0, processed = TRUE, usage = "train",
         label = TRUE)

# write to s3 bucket as input to new validation instance
s3write_using(
  ref_sample_labeller,
  FUN = readr::write_csv, bucket = "activemapper",
  object = glue::glue("planet/incoming_names_static_validator.csv")
)
# chk <-  s3read_using(
#   readr::read_csv, bucket = "activemapper",
#   object = "planet/incoming_names_static_validator.csv"
# )
# all(ref_sample_labeller == chk)

# ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)
# ggplot() + geom_sf(data = ghana) +
#   geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) + 
#   theme_light()
```

Set up new instance for performing accuracy assessment, using set up functions provided in `labeller` utilities, following doc `setting-up-new-labeller-instance.md` 
```bash
./common/tools/create_ami.sh labeller8 labeller8-29June2020
./common/tools/create_instance.sh ami-0ee381fa2e7c5b33d t2.small labeller-security
./common/tools/create_elasticip.sh validator crowdmapper.org
```
And then on board modification of `validator`.

Steps:

- Switched to reformat branch
- Ran `revoke_qualifications.sh`  
- Ran `fire_up_labeller.py`

```python
python common/fire_up_labeller.py --initial 2 --ec2_instance  "validator" --run_id 0 --aoi_index 1 --aoi_name "validator" --github_branch "reformat" --incoming_names_static_path "planet/incoming_names_static_validator.csv"  --security_group_id "sg-0a8bbc91697d6a76b" --secret_key <key> --db_user <user> --db_pwd <pwd> --github_token <token> --aws_access <access> --aws_secret <secret> --aws_region "us-east-1"
```

Had to change variables manually in `config.yaml` after this to point to correct catalogs.

- Then run `Rscript spatial/update_db.R`
- Update DB configurations: Set parameters for running validation sample, remembering that the size of the cell for the target will be about 1/10 of a hectare (0.0001433 degrees)

```{r, eval=FALSE}
gcs <- unname(unlist(st_crs(ghana))[2])
ref_ss <- sample(1:nrow(ref_sample), 10)
ref_dt <- data.table::data.table(ref_sample %>% select(-aoi, -fwts))[ref_ss, ]
point_to_gridpoly(ref_dt, 0.0001433, gcs, gcs) %>% 
  st_area() %>% units::set_units("ha")
```

Updating
```{r, eval = FALSE}
update_tbl <- tibble(
  key = c("Hit_AvailTarget", "Hit_FqaqcPercentage", "Hit_MaxAssignmentsF", 
    "Hit_QaqcPercentage", "KMLdlat", "KMLdlon", "Hit_StandAlone"), 
  value = c(5, 100, 1, 0, 0.0001433, 0.0001433, "true")
)

host <- "validator.crowdmapper.org"
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
                      dbname = "Africa", user = dinfo$db_username,
                      password = dinfo$db_password)
for(i in 1:nrow(update_tbl)) {
  update_sql <- glue::glue(
    "UPDATE configuration SET value={update_tbl[i, 'value']} ",
    "WHERE key='{update_tbl[i, 'key']}'"
  )
  print(update_sql)
  dbExecute(con, update_sql)
}

fs <- tbl(con, "kml_data") %>% collect() %>% filter(kml_type == "F")
# all(fs$name %in% ref_sample$name)

DBI::dbDisconnect(con)
```

And made three categories for validation: annualcropland, noncropland, unsure, entering those manually in database.

Next, start the system running to create hits. Use an R chunk to execute from chunk using a system call
```{r, eval=FALSE}
cmd <- glue::glue("ssh mapper@validator.crowdmapper.org ", 
                  "python labeller/common/crontab_runner.py --hits T")
system(cmd)

# check whether daemons are running and creating hits
system("ssh mapper@validator.crowdmapper.org ps -ef | grep mapper")
system(glue::glue("ssh mapper@validator.crowdmapper.org ",
                  "labeller/common/tools/list_hits.py"))
```

Quick check to see if size of mapping target is correct--map box on qualification test and then check size
```{r, eval=FALSE}
user_sql <- "SELECT name, try, category, geom_clean FROM qual_user_maps"
user_polys <- suppressWarnings(st_read(con, query = user_sql))
user_polys %>% st_area()  # 0.1 ha

# compare to grid
grid_dt <- data.table::data.table(
  mgrid_tb %>% filter(name == gsub("_1", "", user_polys$name))
)
grid_poly <- point_to_gridpoly(grid_dt, 0.005 / 2, gcs, gcs)
grid_poly %>% st_area() / user_polys %>% st_area()

plot(st_geometry(grid_poly))
plot(st_geometry(user_polys), add = TRUE)
```

Then grant a qualification
```{r, eval=FALSE}
system(glue::glue("ssh mapper@validator.crowdmapper.org ",
                  "labeller/common/tools/grantQualification.py ", 
                  "<email>"))
```

### Distribution of points 
Map reference sample so far
```{r, eval=FALSE}
host <- "validator.crowdmapper.org"
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
                      dbname = "Africa", user = dinfo$db_username,
                      password = dinfo$db_password)
user_sql <- "SELECT name, category, geom_clean FROM user_maps"
user_polys <- suppressWarnings(st_read(con, query = user_sql))
pt_nms <- unique(gsub("_.*", "", user_polys$name))

pts <- tbl(con, "master_grid") %>% filter(name %in% pt_nms) %>% 
  select(name, x, y) %>% collect()
pts <- pts %>% st_as_sf(coords = c("x", "y"), crs = 4326)

ggplot() + geom_sf(data = aois) + 
  geom_sf(data = pts, size = 0.2) + theme_linedraw()
```

#### Old below here
## Map reference site selection

Now read in and assess accuracy. Grab probability images for AOI1
```{r, eval=FALSE}
library(aws.s3)
library(raster)

# create master_grid extent, get list of probability images, download fpool 
# (for checking)
aois <- st_read(here("external/data/aois.geojson"))
mgrid_ext <- raster(extent(-17.541, 51.459, -35.46, 37.54), res = 0.05)
pimgs <- get_bucket_df(bucket = "activemapper", 
                       prefix = "classified-images/1_whole_test") %>% as_tibble
fpool1 <- s3read_using(FUN = readr::read_csv, 
                       object = "planet/f_pool_1.csv", bucket = "activemapper")
host <- "labeller1.crowdmapper.org"
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host, 
                      dbname = "Africa", user = "postgis", password = pw)
mgrid <- tbl(con, "master_grid") %>% filter(name %in% local(fpool1$name)) %>% 
  collect()
kml_data <- tbl(con, "kml_data") %>% filter(kml_type == "F") %>% collect()
# mgrid %>% st_as_sf(coords = c("x", "y")) %>% st_geometry %>% plot

# strip out row, col from names of probability images, and convert them to x, y
# from dummy master_grid
sstr <- "classified-images/1_whole_test/image_"
pimgs <- pimgs %>% select(Key) %>% 
  mutate(row = gsub("c*.*_r|.tif", "", gsub(sstr, "", Key))) %>% 
  mutate(col = gsub("c|_r*.*.tif", "", gsub(sstr, "", Key))) %>% 
  select(Key, row, col)
pimgs_xy <- bind_cols(pimgs, 
                      x = xFromCol(mgrid_ext, as.numeric(pimgs$col) + 1),
                      y = yFromRow(mgrid_ext, as.numeric(pimgs$row) + 1)) %>%
  st_as_sf(coords = c("x", "y"), crs = 4326)

# use xy to extract tile ids
tiles <- read_sf(here("inst/extdata/ghana_tiles.geojson"))
tiles_xy <- st_intersection(tiles, pimgs_xy)
tiles_xy <- st_intersection(tiles_xy, aois %>% select(aois)) %>% as_tibble() %>%
  select(tile, row, col, aois)
tiles_rc <- inner_join(tiles_xy, tiles %>% select(tile)) %>% st_as_sf() %>% 
  select(aois, tile, row, col)
# ggplot(aois) + geom_sf() + geom_sf(data = tiles_xy) +
#   theme_void()

# select grid cell coordinates for reference cells, and then use those to 
# intersect with tiles to get row, col for probability images
ref_xys <- mgrid %>% filter(name %in% ref_aoi1$name) %>% 
  distinct(id, name, x, y) %>% st_as_sf(coords = c("x", "y"), crs = 4326)
# which(ref_xys$name %in% kml_data$name)  # confirm no reference sites intersect
ref_xys_tiles <- st_intersection(ref_xys, tiles_rc) %>% as_tibble

# Then use to get probability image values
# ref_xys_tiles %>% distinct(row, col)
dup_tiles <- ref_xys_tiles %>% filter(duplicated(tile)) %>% pull(tile)
# ref_xys_tiles %>% filter(tile %in% dup_tiles)
unique_ref_tiles <- ref_xys_tiles %>% distinct(tile) %>% pull %>% sort
ref_pvals <- lapply(unique_ref_tiles, function(x) {  # x <- unique_ref_tiles[6]
  print(x)
  tile <- ref_xys_tiles %>% filter(tile == x) 
  ref <- ref_aoi1 %>% filter(name %in% tile$name)
  prob_tile <- pimgs %>% 
    filter(row == unique(tile$row) & col == unique(tile$col))
  prob_tiler <- raster(paste0("/vsis3/activemapper/", prob_tile$Key))
  pvals <- extract(prob_tiler, ref)
  v <- do.call(rbind, lapply(pvals, function(x) {
    tibble(mu = mean(x), sd = sd(x), n = length(x))
  }))
  bind_cols(ref, v) %>% select(name, rid, cls, mu, sd, n)
})
ref_pvals <- do.call(rbind, ref_pvals)
ref_pvals_tb <- ref_pvals %>% as_tibble
ref_pvals_tb %>% ggplot() + geom_boxplot(aes(x = cls, y = mu))

tp <- ref_pvals_tb %>% filter(cls == "f" & mu > 50) %>% count
fn <- ref_pvals_tb %>% filter(cls == "f" & mu < 50) %>% count
fp <- ref_pvals_tb %>% filter(cls == "n" & mu > 50) %>% count
tn <- ref_pvals_tb %>% filter(cls == "n" & mu < 50) %>% count

# ref_pvals_tb %>% filter(cls == "f" & mu > 40) %>% count
# ref_pvals_tb %>% filter(cls == "n" & mu > 40) %>% count
ref_pvals_tb %>% filter(cls == "n" & mu < 40) %>% count
laref_pvals_tb %>% filter(cls == "u" & between(mu, 30, 70))
ref_pvals_tb %>% filter(cls == "u" & (mu < 30 | mu > 70))

(tp + tn) / (tp + tn + fp + fn)

ref_pvals_tb %>% filter(cls == "n" & mu > 50)

```

### 2020-07-24 inst/manuscript/analyses.Rmd
Substantial updates to analyses and organization. Manuscript, figures notebook, and analyses moved to inst/manuscript. Added scripts to data-raw/ for processing inputs into figures, headed to instances, etc (much of this is moving code out of analyses* notebooks).

# Analyses and figures
## Datasets

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(raster)
library(sf)
library(dplyr)
library(ggplot2)
library(purrr)
library(rmapaccuracy)
library(here)
library(doMC)
library(activemapper)
library(aws.s3)
```

### Base Datasets

The pre-preparation of base datasets that install with the `activemapper` package is shown in `data-raw/data_for_instances.R`. These are loaded in here from `ext_data`.

```{r, eval=FALSE}
res <- 0.005
res_supergrid <- 1
num_supergrid <- res_supergrid / res

mgrid <- raster(
  system.file("extdata/ghana_grid.tif", package = "activemapper")
)
africa <- st_read(
  system.file("extdata/africa.geojson", package = "activemapper")
)
africa_cont <- africa %>% st_union %>% st_sf
ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)

degrees <- st_read(
  system.file("extdata/degrees.geojson", package = "activemapper")
)
tiles <- st_read(
  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
) 
aois <- st_read(
  system.file("extdata/aois.geojson", package = "activemapper")
)
# ggplot(data = tiles) + 
#   geom_sf(aes(fill = as.factor(aoi1)), col = "transparent")
```

### Map results
```{r, eval=FALSE}
## segments
segment_files <- dir(here("external/data/results/segments/merged"), 
                     pattern = "boundarymerge", full.names = TRUE)

iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
```

## Intermediate processing
### Grid for stratifying model results
```{r, eval=FALSE}
r <- raster(extent(tiles), res = 0.05)
crs(r) <- crs(tiles)

# for field areas
tilesr <- fasterize::fasterize(tiles, r, field = "aoi1")
tilesr005 <- disaggregate(tilesr, fact = 10)
tilesr005_sf <- stars::st_as_stars(tilesr005) %>% st_as_sf() %>% 
  rename(aoi1 = layer)

# for point intersection
# mask main grid and convert to points
mgrid <- mask(mgrid, tiles)
mgrid_xy <- as.data.frame(mgrid, xy = TRUE) %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% rename(id = ghana_grid) %>% 
  tidyr::drop_na(id)

# join with tiles
# system.time(extract(tilesr, master_grid_xy))  # 11.81
system.time(mgrid_xytiles <- st_join(mgrid_xy, tilesr005_sf))  # 26

# ggplot(tilesr005_sf) + geom_sf(aes(fill = layer), col = "transparent")
# tilesr005_sf %>% filter(layer == 1) %>% plot()
# tiles_ss <- tiles %>% filter(aoi1 == 1) %>% slice(1:6, 10:15)
# tiles_ss <- tiles %>% filter(aoi1 == 1) %>% slice(1)
# plot(st_geometry(tiles_ss))
# plot(crop(tilesr, tiles_ss), add = TRUE)
# plot(crop(tilesr005, tiles_ss), add = TRUE)
# plot(crop(master_grid, tiles_ss), add = TRUE)
# plot(master_grid)

```

### Field characteristics
#### Field density

Process the field segments to calculate cropland coverage per 0.005 degree pixel. Run these in loop.

```{r, eval=FALSE}
logf <- here("external/logs/gridding_005.log")
registerDoMC(cores = 7)
fareas <- foreach(i = segment_files$order, .combine = stack) %dopar% {  
  # i <- segment_files[11]
  aoi_id <- gsub("[[:alpha:]]|_|\\.", "", basename(i))

  cat(glue::glue("Starting {aoi_id} at {Sys.time()}"), file = logf, 
      sep = "\n", append = TRUE)
  print(aoi_id)

  # read in fields
  cat(glue::glue("...aoi {aoi_id} reading in fields and calculating tile area"),
      file = logf, sep = "\n", append = TRUE)
  flds <- read_sf(i)
  # select out AOI grid and 
  aoi_tiles <- tilesr005_sf %>% filter(aoi1 == aoi_id) %>% 
    mutate(cid = 1:n(), area = as.numeric(st_area(.) / 10000))

  # intersect fields
  cat(glue::glue("...aoi {aoi_id} intersecting fields and calculating area"),
      file = logf, sep = "\n", append = TRUE)
  # fld_int <- st_intersection(aoi_tiles, flds)
  fld_int <- st_join(aoi_tiles_xy, flds)
  fld_int %>% mutate(class = )

  drops <- fld_int %>%
    st_is(c("POINT", "MULTIPOINT", "LINESTRING", "MULTILINESTRING")) %>%
    which()
  polys <- fld_int %>% st_is("POLYGON") %>% which()
  fld_int <- fld_int %>% slice(-drops)
  fld_int <- fld_int %>% mutate(farea = as.numeric(st_area(.) / 10000))
  fld_areas <- fld_int %>% as_tibble() %>% group_by(cid) %>%
    summarize(farea = sum(farea))

  cat(glue::glue("...aoi {aoi_id} joining areas with tiles"), file = logf,
      sep = "\n", append = TRUE)
  aoi_tiles <- left_join(aoi_tiles, fld_areas) %>%
    mutate(farea = ifelse(is.na(farea), 0, farea)) %>%
    mutate(fprop = round(farea / area, 4))

  # crop down to just 
  cat(glue::glue("...aoi {aoi_id} rasterizing"), file = logf, sep = "\n",
      append = TRUE)
  aoir <- tilesr005 == as.numeric(aoi_id)
  arear <- area(aoir) * 100
  # aoir[aoir == 0] <- NA
  farea <- fasterize::fasterize(aoi_tiles, aoir, field = "fprop",
                                fun = "sum", background = 0)

  cat(glue::glue("Finished aoi {aoi_id} at {Sys.time()}"), file = logf,
      sep = "\n", append = TRUE)
  cat("", file = logf, sep = "\n", append = TRUE)
  return(farea)
}

# plot(fdensity)

# writeRaster(fdensity, filename = here("external/data/field_density005.tif"))
# fdensity <- raster(here("external/data/field_density005.tif"))
```

#### Field points
For reference site selection and stratification
```{r, eval=FALSE}
logf <- here("external/logs/field_points_005.log")
registerDoMC(cores = 7)
fpoints <- foreach(i = segment_files$file) %dopar% {  
  # i <- segment_files$file[1]
  aoi_id <- gsub("[[:alpha:]]|_|\\.", "", basename(i))

  cat(glue::glue("Starting {aoi_id} at {Sys.time()}"), file = logf, 
      sep = "\n", append = TRUE)

  # read in fields
  cat(glue::glue("...aoi {aoi_id} reading in fields and calculating tile area"),
      file = logf, sep = "\n", append = TRUE)
  flds <- read_sf(i) %>% rename(fid = id)  # rename id field

  # intersect fields with main grid
  cat(glue::glue("...aoi {aoi_id} intersecting fields and calculating area"),
      file = logf, sep = "\n", append = TRUE)
  mgrid_ss <- mgrid_xytiles %>% filter(aoi1 == aoi_id)  # subset mgrid
  fld_int <- st_join(mgrid_ss, flds)
  fld_int <- fld_int %>% mutate(class = ifelse(is.na(fid), 0, 1)) %>% 
    select(id, fid, aoi1, class) %>% rename(aoi = aoi1)

  cat(glue::glue("aoi {aoi_id} complete"),
      file = logf, sep = "\n", append = TRUE)
  return(fld_int)
}

# Rename one of two AOI3 results, and then bind
fpoints[[3]] <- fpoints[[3]] %>% mutate(aoi = "3N")
fpoints <- do.call(rbind, fpoints)
xy <- fpoints %>% st_coordinates() %>% as_tibble() %>% rename(x = X, y = Y)
fpoints_tb <- bind_cols(fpoints %>% as_tibble() %>% select(-geometry), xy)
# data.table::fwrite(
#   fpoints_tb, file = here("external/data/field_points005.csv")
# )

# readr::read_csv(here("external/data/field_points005.csv"))
# fpoints[[3]] %>% filter(class == 1) %>% st_geometry %>% plot(cex = 0.1, pch = 16)
```

## Climatology (supplemental)
Grab CHIRPs monthly data from `mapper`
```{r, eval = FALSE}
library(stars)
dir(file.path(data_path, "external/climatology/chirps"), full.names = TRUE) %>% 
  stack(.) %>% crop(., ghana) -> chirps
rftot <- calc(chirps, sum)
monthly_pct <- chirps / rftot * 100
names(monthly_pct) <- lubridate::month(1:12, label = TRUE)
monthly_pct <- setZ(monthly_pct, lubridate::month(1:12, label = TRUE))
monthly_pct_stars <- stars::st_as_stars(monthly_pct)

# plot(monthly_pct_stars, breaks = seq(0, 35), ) + 
#   geom_sf(data = ghana, fill = "transparent", size = 0.3)
p <- ggplot() + geom_stars(data = monthly_pct_stars) + 
  facet_wrap("band", nrow = 2) + 
  geom_sf(data = ghana, fill = "transparent", size = 0.3) +
  scale_fill_viridis_c(breaks = seq(0, 35, 5), "%", na.value = "grey90") +
  xlab("") + ylab("") + ggtitle("Percent of annual rainfall by month") + 
  theme(panel.background = element_rect(fill = NA), 
        axis.text = element_blank(), axis.ticks = element_blank()) 
ggsave("manuscript/figures/si_rainfall_pct.png", height = 4, width = 6, 
       units = "in", dpi = 300)
```

### Mapping AOIs

These were defined in `data_prep.R`.

## Average quality of image tiles
Visually analyzed on instance `labellertest`.

### Accuracy within blocks

Placeholder. Use imagined accuracy based on rainfall, using dummy approach as with Figure 3.  This will be replaced by metrics from tables for each instance. 
```{r, eval=FALSE}
rf <- aois %>% st_centroid %>% extract(rftot, .)
aois <- aois %>% mutate(rf = rf)

# Rainfall-accuracy relationship
rflm <- coef(lm(c(0.9, 0.75) ~ range(aois$rf)))  # 75-90%
set.seed(7) 
aois <- aois %>% 
  mutate(Accuracy = (rflm[1] + rflm[2] * rf) * 
           runif(n(), min = 0.95, max = 1.05) * 100)

# Users
rflm <- coef(lm(c(0.85, 0.75) ~ range(aois$rf)))  # 75-90%
set.seed(1) 
aois <- aois %>% 
  mutate(User = (rflm[1] + rflm[2] * rf) * 
           runif(n(), min = 0.95, max = 1.05) * 100)

# Producers
rflm <- coef(lm(c(0.9, 0.75) ~ range(aois$rf)))  # 75-90%
set.seed(2) 
aois <- aois %>% 
  mutate(Producer = (rflm[1] + rflm[2] * rf) * 
           runif(n(), min = 0.95, max = 1.05) * 100) %>% 
  select(grp, aois, rf, Accuracy, User, Producer)

```

## Figure 5
Plot of Zones and average accuracy
```{r, eval = FALSE}
# ghana_tilesr %>% group_by(aoi) %>% summarize(aois = mean(aoi)) %>% ungroup %>% 
#   ggplot() + geom_sf(col = "grey70") +
#   geom_sf_text(aes(label = aoi), size = 2) + 
#   geom_sf(data = ghana, fill = "transparent", lwd = 0.3) + 
#   theme_void()

cols <- RColorBrewer::brewer.pal(n = 9, "RdYlGn") %>% #[c(1, 9)] %>% 
  colorRampPalette(colors = .)
# RColorBrewer::display.brewer.all()

# ggsave("manuscript/figures/si_mapping_blocks.png", width = 5, height = 5, 
#        units = "in", dpi = 300)

col_grad <- scale_fill_gradientn(colors = cols(10), limits = c(70, 100), 
                                 breaks = seq(70, 100, 5)) 
p <- map_blocks %>% 
  ggplot() + geom_sf(aes(fill = Accuracy)) + #, col = "grey70") +
  geom_sf_text(aes(label = grp), size = 2) + 
  geom_sf(data = ghana, fill = "transparent", lwd = 0.3) + col_grad + 
  theme_void()
p2 <- map_blocks %>% 
  ggplot() + geom_sf(aes(fill = User)) + #, col = "grey70") +
  geom_sf_text(aes(label = grp), size = 2) + 
  geom_sf(data = ghana, fill = "transparent", lwd = 0.3) + col_grad + 
  theme_void()
p3 <- map_blocks %>% 
  ggplot() + geom_sf(aes(fill = Producer)) + #, col = "grey70") +
  geom_sf_text(aes(label = grp), size = 2) + 
  geom_sf(data = ghana, fill = "transparent", lwd = 0.3) + col_grad + 
  theme_void()

leg <- cowplot::get_legend(p + theme(legend.position = c(0.6, 1)))
gleg <- cowplot::plot_grid(NULL, leg, ncol = 1, align = "h")
gp <- cowplot::plot_grid(p + theme(legend.position = 'none'), 
                         p2 + theme(legend.position = 'none'), 
                         p3 + theme(legend.position = 'none'), nrow = 1, 
                         labels = c("Overall", "User's", "Producer's"), 
                         label_x = 0.0)
gall <- cowplot::plot_grid(gp, gleg, rel_widths = c(1, 0.1), label_x = 0)
# gall

ggsave("manuscript/figures/figure5.png", width = 9, 
       height = 4, units = "in", dpi = 300)

```

## Figure 6: High, low, medium comparisons

Statistics from high, medium, low experiment results from Boka. 
```{r, eval = FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(cowplot)

# metrics
scores <- tibble(
  # AOI = as.character(c(rep(1, 3), rep(8, 3), rep(15, 3))), 
  AOI = c(rep(1, 3), rep(8, 3), rep(15, 3)), 
  Label = c("Low", "High", "Consensus", "Low", "High", "Consensus", 
            "Low", "High", "Consensus"), 
  AUC = c(0.6987, 0.7803, 0.7741, 0.7333, 0.8461, 0.8563, 0.6916, 0.7628, 
          0.7714),
  Accuracy = c(0.6575, 0.7919, 0.7897, 0.5313, 0.8361, 0.8421, 0.6279, 0.6916,
               0.7245),
  TSS = c(0.2940, 0.4194, 0.4138, 0.2568, 0.5390, 0.5639, 0.2718, 0.3835, 
          0.3916)
)
p <- scores %>% pivot_longer(cols = c("AUC", "Accuracy", "TSS")) %>% 
  arrange(AOI) %>%
  group_by(AOI) %>% 
  ggplot(.) + geom_bar(aes(x = name, y = value, fill = Label), 
                       stat = "identity", position = "dodge") + 
  scale_fill_manual(values = c("blue", "purple", "orange"), 
                    name = "Labelling strategy", 
                    labels = c("Consensus", "Highest scorer", 
                               "Lowest scorer")) + 
  xlab("Metric") + ylab("Score") + ggtitle("AOI") +
  facet_grid(. ~ AOI) + theme_cowplot() 
ggsave(p, 
       filename = here::here("manuscript/figures/figure6.png"),
       height = 4, width = 8, dpi = 300)
save(scores, file = here::here("external/data/label_experiment_metrics.rda"))
```

### Figure SX: SI probability images
```{r, eval=FALSE}
library(raster)
probpal <- viridis::viridis
cprob <- raster(glue::glue("/vsis3/activemapper/classified-images/", 
                           "1_whole/image_c294_r532.tif"))
hprob <- raster(glue::glue("/vsis3/activemapper/classified-images/", 
                           "test_byScore/aoi1/high/image_c294_r532.tif"))
lprob <- raster(glue::glue("/vsis3/activemapper/classified-images/", 
                           "test_byScore/aoi1/low/image_c294_r532.tif"))
png(here::here("manuscript/figures/si_hml_probs.png"), height = 2.65,
    width = 9.2, res = 300, units = "in", bg = "transparent")
par(mfrow = c(1, 3), mar = c(0, 0, 0, 1), oma = c(0, 0, 0, 2))
plot(as(extent(cprob), "SpatialPolygons"), lty = 0, #main = "Consensus labels", 
     cex.main = 2)
plot(cprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
     col = probpal(20))
mtext("A", side = 3, line = -2, adj = 0.00)
plot(as(extent(cprob), "SpatialPolygons"), lty = 0, #main = "Highest scorer", 
     cex.main = 2)
mtext("B", side = 3, line = -2, adj = 0.0)
plot(hprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
     col = probpal(20))
plot(as(extent(cprob), "SpatialPolygons"), lty = 0, #main = "Lowest scorer", 
     cex.main = 2)
mtext("C", side = 3, line = -2, adj = 0.0)
plot(lprob, lty = 0, add = TRUE, zlim = c(0, 100),
     legend.width = 1, legend.shrink = 0.5, col = probpal(20),
     legend.args = list(text = '%', side = 3, cex = 0.8))
dev.off()

```

## Consensus maps
```{r, eval=FALSE}
library(rmapaccuracy)
library(leaflet)

instance <- "mapper0"
dbase <- "Africa"
host <- paste0(instance, ".crowdmapper.org")
upw <- list("user" = "postgis", "password" = "P0stG1S")
coninfo <- mapper_connect(host)
con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
                      dbname = "Africa", user = upw$user,
                      password = upw$password)

# coninfo <- mapper_connect(host = host)
coninfo$con <- con

common_path <- "~/Dropbox/projects/activelearning/mapper"
common_path <- file.path(common_path, "common")
params <- yaml::yaml.load_file(file.path(common_path, 'config.yaml'))

name <- "GH0431618"#"GH0448854"
sqls <- paste0("select id from master_grid where name='", name, "'")
cid <- DBI::dbGetQuery(con, sqls)[[1]]

user.sql <- paste0("select name, category, worker_id, geom_clean ",
                   "FROM user_maps INNER JOIN categories USING (category)",
                   "INNER JOIN assignment_data USING (assignment_id)",
                   " where name like '",  name, "%' ",
                   "AND categ_group ='field'")

user.polys <- DBI::dbGetQuery(con, gsub(", geom_clean", "", user.sql))
user.polys <- suppressWarnings(st_read(con, query = user.sql))
user.polys <- user.polys %>% filter(st_is(. , "POLYGON"))

# # calling consensus map manually, line by line in script
kmlid <- name
qsite <- FALSE
mode <- "consensus"

# execute consensus script
source(file.path(here::here(), "external/scripts/consensus_labels.R"))
l <- list(user.polys["worker_id"], bayesoutput[[2]], bayesoutput[[1]])
save(l, file = file.path(here::here(), "data/consensus.rda"))

scenes <- tbl(con, "scenes_data") %>% filter(cell_id == cid) %>% 
  dplyr::select(scene_id, cell_id, url, season, tms_url) %>% collect()
# scenes$scene_id

gcs <- "+proj=longlat +datum=WGS84 +no_defs"
name <- tbl(coninfo$con, "master_grid") %>% filter(id == cid) %>%
  dplyr::select(id, name, x, y) %>% collect()
gpoly <- rmapaccuracy::point_to_gridpoly(data.table::data.table(name), 
                                         w = 0.005 / 2, gcs, gcs)

s3url <- scenes %>% filter(season == "GS") %>% pull(url)
planet_path <- file.path("/vsis3", gsub("s3://", "", s3url)) 
planet <- brick(planet_path)

crfeat <- as(extend(extent(as_Spatial(ejura_sites[1, ])), 0.005),
             "SpatialPolygons")
zoomfeat <- as(extend(extent(as_Spatial(ejura_sites[1, ])), 0.002),
               "SpatialPolygons")
proj4string(crfeat) <- proj4string(as_Spatial(user.polys))
proj4string(zoomfeat) <- proj4string(as_Spatial(user.polys))
crfeat_utm <- spTransform(crfeat, crs(planet))
zoomfeat_utm <- spTransform(zoomfeat, crs(planet))
planet_cr <- crop(planet, crfeat_utm)
planet_cr2 <- crop(planet2, crfeat_utm)

# set up display
os <- (scenes %>% filter(season == "OS") %>% filter(row_number() == 1) %>% 
  dplyr::select(tms_url))$tms_url
gs <- (scenes %>% filter(season == "GS") %>% filter(row_number() == 1) %>% 
         dplyr::select(tms_url))$tms_url

m <- leaflet() %>% addTiles() %>% setView(name$x, name$y, zoom = 14) %>% 
  addTiles(os, group = "OS") %>% addTiles(gs, group = "GS") %>% 
  addPolygons(data = gpoly, fill = FALSE, group = "Cell") %>% 
  addLayersControl(overlayGroups = c("Cell", "GS", "OS"),
    options = layersControlOptions(collapsed = FALSE, autoZIndex = FALSE))
m

```

### 2020-07-24 inst/manuscript/manuscript.Rmd
Substantial updates to analyses and organization. Manuscript, figures notebook, and analyses moved to inst/manuscript. Added scripts to data-raw/ for processing inputs into figures, headed to instances, etc (much of this is moving code out of analyses* notebooks).

--- 2020-06-25 manuscript/manuscript.Rmd
@@
 The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019].
-These errors are typically not accounted for in reported accuracy metrics (cite Arthur's paper), thus the degree to which such labelling error impacts map accuracy is unclear.
-However, a study of how land cover map error propagates through subsequent analyses [@Esteslargeareaspatiallycontinuous2018] suggests training data quality could have a large impact.
+These errors are typically not accounted for in reported accuracy metrics, despite the fact that they can introduce substantial error into the resulting maps and subsequent analyses [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018].
 The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland.
@@
 ## System overview 
-The mapping system is centered around four major components: 1) an image acquisition and pre-processing component, 2) a training data collection platform, 3) a machine learning component, and a 4) segmentation component.
-The first component is applied as a once-off step to generate an image catalog necessary for the mapping geography, while the second component initiates the the interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig.
+The mapping system is centered around four components: 1) an image acquisition and compositing component, 2) a training data collection platform, 3) a machine learning component, and 4) a segmentation component.
+The first component is applied as a once-off step to generate an image catalog covering the mapping geography, while the second component initiates an interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig.
 1).
-The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in the final vectorized fields boundary maps.
-Each system component comprises an individual module software hosted on a GitHub repository (see data and software availability section for details).
+The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in vectorized field boundaries.
+Each system component comprises an individual software module designed to be implemented on cloud computing architecture, and available on a GitHub repository (see data and software availability section for details).
 
@@
 **[Note: This will require updating to reflect changes to compositing methods]**
-The primary image source used in our system is PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although can be readily adapted to apply to Sentinel and Landsat images.
+The primary image source used in our system is the PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although it can be readily adapted to apply to other images (e.g.
+Sentinel and Landsat).
 PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency.
-Although these images are substantially pre-preprocessed and corrected for atmospheric effects, there are substantial residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], substantial variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016].
+Although these images are already pre-preprocessed and corrected for atmospheric effects, there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016].
 To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.
@@
 
-PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (hosted on Amazon Web Services S3).
+PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (Amazon Web Services S3).
 
@@
 2B), which provides the scale for creating image composite tiles.
-The ARD process simply clips all available imagery intersecting each 0.05 cell to the extent of that cell, regardless of cloud cover.
+The ARD process simply clips all available imagery to the extent of each 0.05 cell it intersects, regardless of cloud cover.
 
@@
 
-```{r, echo = FALSE, out.width="100%", fig.cap="An overview of analytical scales used in the mapping platform, including the 1 degree areas of interest defining the minimum mapping geography (A), the 0.05 degree grid used for creating image composite tiles (B), and the 0.005 degree resolution reference grid used for collecting training data and distributed computing (C).", fig.align='center'}
+```{r, echo = FALSE, out.width="100%", fig.cap="An overview of analytical scales used in the mapping platform, including A) the areas of interest defining the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), the B) 0.05 degree grid used for creating image composite tiles, and the C) 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
 knitr::include_graphics('figures/figure2.png')
 ```
-The temporal compositing process is applied to the ARD imagery for two time periods, the dry season period (December-February) and the entire wet season (May - September).
-The longer period is necessary for the wet season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area.
-For each pixel in each image in each ARD temporal stack, we calculate two weights, with the first being:
+The temporal compositing process is applied to the ARD imagery for the time period of interest, which in the case of this study is one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year.
+Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in same year helps to minimize differences caused by land change.
+For each pixel in each image in each ARD temporal stack for a given season, we calculate two weights, with the first being:
 
@@
 
-Once these two weights are calculated, the final composited pixel value for each of the four PlanetScope bands is calculated as:
+Once these two weights were calculated, the final composited pixel value for each of the four PlanetScope bands was calculated as:
 
@@
 Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
-The composited tiles are then added to the S3 store (Fig.
+The composited tiles were then added to the S3 store (Fig.
 1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using RasterFoundry (cite).
-This web-based rendering of the image is used in the training data platform (see section X).
+This web-based rendering of the image is used in the training data platform (see **section X**).
 
 ### Labelling platform
-Both training and reference data are collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*.
+Both training and reference data were collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*.
 The original version of *labeller* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
-The basic structure of the system remains the same, but we have re-engineered it to be a standalone platform that 1) allows us to register and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process.
+The basic structure of the system remains the same, but we have re-engineered it into a standalone platform that 1) allows us to enroll and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process.
 The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process.
@@
 one to several 1 degree cells in Fig.
-2A), thus several instances are set up to run in parallel to cover a larger mapping geography such as Ghana.
+2A)
 
-Here we provide an overview of *labeller*'s architecture within the context of the labeling workflow within a single instance.
+Here we provide an overview of *labeller*'s architecture.
 
@@
 2A).
-These site names are stored in a Postgres table that represents the mapping queue.
+These site names are stored in a Postgres table with the PostGIS extension that represents the mapping queue.
 Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.
@@
 3).
-There the worker is presented with a white target box representing the 0.005 degree cell that was randomly selected to label, a set of digitizing tools, and different image backdrops, which include false and true color renderings of both the growing season and off-growing season PlanetScope images covering that cell, as well as a basemap provided by Bing maps.
+There the worker is presented with a white target box representing the 0.005 degree cell that was randomly selected to label, a set of digitizing tools, and different image backdrops, which include false and true color "slippy map" renderings (generated by Raster Foundry) of both the growing season and off-growing season PlanetScope images covering that cell, as well as basemap imagery provided by Bing and ESRI.
 The worker uses the polygon drawing tool to digitize the boundaries in crop fields visible within the PlanetScope overlays that intersect the grid target, following a set of pre-defined digitizing rules.
@@
 Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table.
-The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training machine learning.
-If is was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score based on the following formula:
+The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training the machine learning algorithm.
+If it was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score based on the following formula:
 
@@
 If the worker's completed assignment was a labelling task, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers.
-Once complete, an R-based consensus labelling routine is invoked, which combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:
+Once complete, a `R`-based routine is invoked, which combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:
 
@@
 P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel.
-This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details of equation).
+This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details).
+The Bayesian risk of the completed final is then assessed, which measures the average level of consensus for the label assigned to each pixel in a given site, with values closer to 0 indicating complete consensus, and thus the highest confidence, while values approaching 1 reveal progressively increasing disagreement (see SI).
 
-Once a the full set of sites has been completed, they are randomly partitioned into training and validation sets according to predetermined proportions.
+Once all training sites are completed, they are randomly partitioned into training and validation sets according to predetermined proportions.
 
@@
 ### Classification pipeline
-Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens to hundreds of instances, depending on the size of the mapping geography.
-We identifed an optimal cluster of 50-60 instances (m4.2xlarge) each having 8 virtual CPUs and 30 GB RAM, which ensured minimal idle time for any single node.
+Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens of instances, depending on the size of the mapping geography.
 
@@
 Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
-We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of a mapping geography orders of magnitude larger than that initial work.
-Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table X).
+We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of mapping geography that were several orders of magnitude larger.
+Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table 1).
 
-\begin{center}Table X. List of features.\end{center}
+\begin{center}Table 1. List of features.\end{center}
 
@@
 Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`.
-These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table X), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data.
+These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table 1), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data.
 Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.
@@
 After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process.
-The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computation demand.
+The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computational demand.
 The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:
@@
 
-In the second step, a Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands [**I assume**], and weighted edge image is computed using weights of 1, 1, and 2 to each band for each season [**for both growing season on and off?**].
-A watershed algorithm (**cite**) is then run on the weighted edge image, with a high level of segmentation specified (**2400** segements per tile).
+In the second step, a Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands [**I assume**], and weighted edge image is computed using weights of 1, 1, and 2 to each band for the dry season only [**or growing season?**].
+A watershed algorithm (**cite**) is then run on the weighted edge image, with a high level of segmentation specified.
+In this case, we specified **2400** segments per tile.
 
@@
 
-In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (typically 0.5, but is possible to vary locally) are retained as field polygons.
+In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (typically 0.5, but can vary locally) are retained as field polygons.
 
@@
 
+## Application and evaluation of the system
+To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each one degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent one degree cells that overlapped with neighboring countries.
+The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.
+
+Prior to initiating active learning in each AOI, we grouped them into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI).
+Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group.
+At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations targeted new training sites for collection in each AOI based on classification uncertainty.
+
+To evaluate the results, we undertook a number of analyses.
+
 ### Accuracy assessment
-**[Put in somewhere above text about training reference samples]**
-The accuracy assessment procedure described in section 2.2 is used strictly to assess performance gains in the active learning loop, according to suggested best practice guidelines [@ElmesAccountingtrainingdata2019].
-To assess the accuracy of final map products, following these same guidelines we used an independent map reference sample.
-To develop this sample, we first made a random selection of grid cells within each mapping AOI, with the selected number scaled relative to the AOI.
-Next, a primary supervisor examined the image composites within each randomly selected cell and placed three rectangular polygons of ~0.1 ha area in areas that were fully contained within in one of three classes: cropland, non-cropland, potentially cropland or non-cropland.
-In cases where less than three classes were present within the cell, the supervisor placed polygons in proportion to the dominant class.
-For example, in the case where the entire cell was taken up by non-cropland, the supervisor placed all three polygons in non-cropland.
-If two-thirds of the cell was non-cropland, and one-third ambiguous, two polygons would be placed in non-cropland, and one in the ambiguous class.
+
+The accuracy assessment procedure described in section 2.2 was used strictly to assess performance gains in the active learning loop, according to suggested  best practices [@ElmesAccountingtrainingdata2020].
+To independently assess the categorical accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a map reference sample.
+To develop this sample, we first made a stratified random sample of grid cells across the entire country, using the mapped field boundaries to define the two strata.
+Next, a primary supervisor examined the image composites in each randomly selected cell, and placed a square polygon of ~0.1 ha on the primary class intersecting the centerpoint of the cell.
+The square was shifted so that it was contained entirety with the intersected class: cropland or non-cropland.
+Where the supervisor was uncertain about the class label, one of two additional classes were selected: likely cropland or likely non-cropland.
 After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery.
@@
 This agreement provided the upper bound on knowable map accuracy.
+See the SI for further details on the creation and analysis of the map reference sample.
 
-## Application and evaluation of the system
-[Description of map creation process]
-To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each 1X1 degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent 1X1 cells that overlapped with neighboring countries.
-The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.
+To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against those of the  workers' digitized polygons at map validation sites.
+We calculated the differences in the measures of central tendency (mean, median, mode) between the two datasets within each validation site, and at the scale of the AOI.
 
-Prior to initiating active learning in each AOI, we grouped the AOIs into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI).
-Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group.
-At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations targeted new training sites for collection in each AOI based on classification uncertainty.
-
-To more fully evaluate the capacities of different components of the system, we performed several experiments.
-In the first experiments, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, in addition to those trained with the consensus labels.
-This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy, as recommended by @ElmesAccountingtrainingdata2019.
+To further evaluate the capacities of different components of the system, we performed several further experiments.
+In the first experiment, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, as well as those trained with the consensus labels.
+This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy [@ElmesAccountingtrainingdata2020].
 Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites.
+Lastly, we repeated the entire map development and map assessment process for one AOI for the 2019 agricultural year, in order to assess the ability of the system to create annually updated maps and detect inter-annual change.
 
@@
 -->
+
+## The characteristics of Ghanaian cropland
+Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the characteristics of field sizes through the country.
+We further used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.
 
@@
 ## Image catalog
-
 We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$.
 This entailed processing all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
+The longer period was necessary for the growing season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area.
 For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
 The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile (**Figure Xa**).
+We also collected and composited imagery for one of the 16 AOIs for the two seasons of the 2019 agricultural year.
+All images were converted into Raster Foundry slippy maps for presentation in *labeller*.
 
@@
 
-## The active learning process
+## Active learning
+### Training data collection
+```{r, echo=FALSE, message=FALSE}
+initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
+  count()
+active <- train_val_sites$active %>% as_tibble %>% filter(usage == "train") %>% 
+  group_by(aoi) %>% count()
+typicaln <- raster::modal(initial$n) + raster::modal(active$n)
+aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
+  (active %>% filter(aoi == "labeller10") %>% pull(n))
+aoi14n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
+  (active %>% filter(aoi == "labeller14") %>% pull(n))
+aoi15n <- (initial %>% filter(Cluster %in% 2) %>% pull(n)) + 
+  (active %>% filter(aoi == "labeller15") %>% pull(n))
+aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
+aoi3n <- aoi3init + (active %>% filter(aoi == "labeller3") %>% pull(n))
+
+initial_total <- train_val_sites$initial %>% as_tibble() %>% distinct(name) %>% 
+  count()
+active_total <- sum(active$n)
+validate_total <- train_val_sites$active %>% as_tibble() %>% 
+  filter(usage == "validate") %>% distinct(name) %>% count()
+```
+
+After initial training, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (Figure S2A).
+AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement, thus were trained with `r aoi10n` -  `r aoi14n` samples.
+AOI 15 was run for 4 iterations and had a training sample size of `r aoi15n`, while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
+In this second run, `r aoi3init`  initial training sites randomly selected within the AOI were used (Figure S2A), followed by 2 subsequent active learning subsequent iterations, resulting in a training sample of `r aoi3n`.
+The total number of unique training and validation sites across the country were `r initial_total + active_total` and `r validate_total`, respectively.
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("label_summary")  
+tsites <- label_summary$nlabels %>% group_by(kml_type) %>% 
+  summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
+            stdev = round(sd(n)), lower = min(n), upper = max(n))
+
+# training rate
+train_rate <- label_summary$nlabels %>% 
+  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
+  rename_all(funs(c("ID", "tr", "tref"))) %>% 
+  mutate(trate = tref / tr)
+```
+
+The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core team of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each.
+As each training site was mapped by 4 separate labellers, `r tsites$tot[1]` sets of vector labels were made.
+Each labeller digitized an average of `r tsites$mu[1]` (see Figure S3A for more details on labelling effort).
+
+This accuracy of the labellers was assessed `r tsites$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
+The average of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S3B for detailed score distributions).
+
+After each site was mapped by four labellers, consensus labels were generated.
+
+### Model performance
+
 **[Fill in with intro]**.
@@
 ## Map accuracy
-To assess the overall accuracy of the final maps, we used our reference set to evaluate the accuracy of classifications derived from thresholding the posterior probabilities produced by the RandomForest model, as well as the segmented field boundary maps.
+### Categorical accuracy
+To assess the overall accuracy of the final maps, we used our map reference sample to evaluate the accuracy of classifications derived from thresholding the posterior probabilities produced by the RandomForest model, as well as the segmented field boundary maps.
 First, we evaluated the inter-rater agreement between reference supervisors in labeling reference polygons.
@@
 The accuracy patterns generally align with the difficulty of the labelling task, that is, **the southwestern regions were hardest for workers to interpret, as shown by Figure 5.**
+
+###
 
@@
 # Discussion
+- Label quality
+  - Accuracy asssessment at training reference sites maybe over-inflated. Each worker saw training reference site multiple times. Need larger number of reference samples
+  - Training reference to training ratio was artificially inflated. Workers not always notified when training labels complete, and thus only served training reference sites.
+
 - Left unquantified is how much training on base maps matters

### 2020-07-24 inst/manuscript/si.Rmd
Substantial updates to analyses and organization. Manuscript, figures notebook, and analyses moved to inst/manuscript. Added scripts to data-raw/ for processing inputs into figures, headed to instances, etc (much of this is moving code out of analyses* notebooks).

--- 2020-07-17 manuscript/si.Rmd
@@
 ---
-title: "Supporting Information"
-
-author:
-
-- name: \*Provisional author list\*
-  affilnum: '1'
-
-- name: Lyndon D. Estes
-  affilnum: '1'
-  email: lestes@clarku.edu
-
-- name: Su Ye
-  affilnum: '1'
-
-- name: Lei Song
-  affilnum: '1'
-
-- name: Ron Eastman
-  affilnum: '1'
-
-- name: Sitian Xiong
-  affilnum: '1'
-
-- name: Tammy Woodard
-  affilnum: '1'
-
-- name: Boka Luo
-  affilnum: '1'
-
-- name: Dennis McRitchie
-  affilnum: '2'
-
-- name: Ryan Avery
-  affilnum: '3'
-
-- name: Kelly Caylor
-  affilnum: '3'
-
-- name: Stephanie Debats
-  affilnum: '4'
-
-- name: SpatialCollective
-  affilnum: '5'
-
-- name: Meridia
-  affilnum: '6'
-
-- name: Azavea
-  affilnum: '7'
-
-affiliation:
-
-- affilnum: 1
-  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA
-
-- affilnum: 2
-  affil: Dennis and Sons, Tucson, AZ, USA
-
-- affilnum: 3
-  affil: UCSB
-
-- affilnum: 4
-  affil: Uber
-
-- affilnum: 5
-  affil: SpatialCollective
-
-- affilnum: 5
-  affil: Meridia
-
-- affilnum: 5
-  affil: Azavea
+title: "Supporting Information: Improving cropland maps through tight integration of human and machine intelligence"
 
 output:
-
   pdf_document:
@@
 footerdate: yes
-abstract: The abstract can go either here or below
+# abstract: The abstract can go either here or below
 keywords: rmarkdown, reproducible science
@@
 
+<!-- Provides custom captions for figures in supplement -->
+\newcommand{\beginsupplement}{%
+        \setcounter{table}{0}
+        \renewcommand{\thetable}{S\arabic{table}}%
+        \setcounter{figure}{0}
+        \renewcommand{\thefigure}{S\arabic{figure}}%
+     }
+
 ```{r setup, include=FALSE, cache=FALSE, message = FALSE}
@@
 library(sf)
+library(activemapper)
 
@@
 
+\beginsupplement
 \singlespace
@@
 \bleft
+
 # Methods
@@
 
+After creating the consensus label, the degree of confidence in the resulting label value can be expressed in terms of Bayesian risk. Risk for a given training sample site *i* is calculated as:
+
+\begin{equation}
+\mathrm{R_i=\frac{1}{n}\sum_{j=1}^{n} C_j(1 - L_j) + (1 - C_j)L_j}
+\end{equation}
+
+Where *C* is the consensus probability that a given pixel *j* is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e.
+non-field if *C* < 0.5, field if *C* > 0.5) for that pixel.
+*R* is thus the average risk across all pixels at a given site, and the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}).
+The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling.
+
+```{r riskcurve, echo = FALSE, fig.align='center', out.width="70%", message=FALSE, warning=FALSE, fig.cap="Bayesian risk values (Y-axis) for consensus values (X axis) ranging from 0 to 1 (0 indicates no consensus that a pixel falls into the field class, 1 means complete consensus) for field and non-field consensus labels."}
+knitr::include_graphics('figures/si_label_risk.png')
+```
+
 #### Example
@@
 
-Since `r round((wtA * 1 + wtB * 0) / (wtA + wtB), 3)` > 0.5, we label the particular pixel a field.
+Since `r rj <- round((wtA * 1 + wtB * 0) / (wtA + wtB), 3); rj` > 0.5, we label the particular pixel a field.
+
+Using the relevant part of equation 15, T=the corresponding risk associated with this particular pixel's label $\mathrm{R_j}$ is thus `r rj * (1 - 1) + (1 - rj) * 1`.
 
@@
 We selected a d value of 0.03 and assumed that the user's accuracy of the field class would be 0.75 and that of the non-cropland class would be 0.8, returning sample sizes of 800 and 683, respectively.
-The distribution of the resulting map reference sample is shown in Figure 1.
+The distribution of the resulting map reference sample is shown in Figure \ref{fig:refsample}.
 
-```{r, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
-africa <- st_read(
-  system.file("extdata/africa.geojson", package = "activemapper"), quiet = TRUE,
-)
-aois <- st_read(system.file("extdata/aois.geojson", package = "activemapper"), 
-                quiet = TRUE)
-africa_cont <- africa %>% st_union %>% st_sf
-ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)
-
-ref_sample <- readr::read_csv(
-  here::here("external/reference/map_reference_sample.csv")
-)
-ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)
-
-p <- ggplot() + 
-  # geom_sf(data = ghana) +
-  geom_sf(data = aois) + 
-  geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) +
-  scale_color_manual(values = c("red", "blue"), 
-                     labels = c("Non-cropland", "Cropland"), name = "") + 
-  theme_light()
-p
+```{r refsample, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
+knitr::include_graphics('figures/si_map_reference_sample.png')
 ```
@@
 
+## Active learning process
+### Labelling
+The distributions of the initial 500 sites used to train the RandomForest model for each AOI are shown in Figure \ref{fig:trainval}.
+In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season.
+Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
+```{r trainval, echo = FALSE, out.width="80%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected within AOI 3 to retrain a second active learning run for this region, as well as the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_training_validation_pts.png')
+```
+
+The distribution of data collection effort and scores received at training reference sites by labelling team (Figure \ref{fig:trainqual}).
+
+```{r trainqual, echo = FALSE, out.width="60%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
+Labellers' identities are anonymized.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_training_quality_n.png')
+```
+
+## Experiments
+### Label quality
+
 ```{r, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
@@
 
+\clearpage
 # References

### 2020-07-24 data-raw/data_for_instances.R

@@ -1,0 +1,155 @@
+# Data set up
+
+# Clip down labeller's master grid to just Ghana tiles extent and ID layer for
+# shipping with activemapper package data
+library(raster)
+library(sf)
+library(dplyr)
+
+# get data from labeller
+data_path <- file.path("/Users", unname(Sys.info()["user"]),
+                       "Dropbox/projects/activelearning/mapper/spatial/data")
+master_grid <- brick(file.path(data_path, "processed/master_grid.tif"))
+v <- values(master_grid[[1]])
+# length(v[!is.na(v)])
+# length(unique(v[!is.na(v)]))
+
+# crop master_grid to ghana_grid
+tiles <- st_read("inst/extdata/ghana_tiles.geojson")
+ghana_grid <- crop(master_grid[[1]], extent(tiles))
+# length(ghana_grid[(!is.na(ghana_grid))])
+# length(ghana_grid[!is.na(ghana_grid)])
+# length(unique(ghana_grid[!is.na(ghana_grid)]))
+# which(duplicated(ghana_grid[!is.na(ghana_grid)]))
+
+writeRaster(ghana_grid , file = "inst/extdata/ghana_grid.tif",
+            overwrite = TRUE, datatype = "INT4U")  # careful for datatype
+r <- raster("inst/extdata/ghana_grid.tif")
+# length(r[!is.na(r)])
+# length(unique(r[!is.na(r)]))
+
+# slim down Africa data
+africa <- st_read(file.path(data_path, "external/africa_noisl_gcs.sqlite")) %>%
+  st_sf(crs = 4326) %>% select(name, iso3) %>%
+  st_write("inst/extdata/africa.geojson")
+
+## Set up mapping AOIs
+prod_groupings1 <- list(c(1216, 1217, 1283, 1284), c(1218, 1285), c(1219, 1286),
+                        c(1348, 1349), 1350, c(1351, 1352), c(1413, 1414), 1415,
+                        c(1416, 1417), c(1477, 1478), 1479, c(1480, 1481),
+                        c(1539, 1540), 1541, c(1542, 1543), 1599:1602)
+purrr::map(1:length(prod_groupings1), function(x) {
+  ind <- prod_groupings1[[x]]
+  tiles %>% filter(aoi %in% ind) %>% mutate(grp = x) %>% select(aoi, tile, grp)
+}) %>% purrr::reduce(rbind) -> aois
+
[…]

### 2020-07-24 data-raw/data_from_instances.R

@@ -1,0 +1,387 @@
+##  Preparing data outputs from instances
+
+# Lightweight data go into data/
+# Heavier data go into inst/extdata
+# Intermediate (untracked) inputs go into external/..
+
+# usethis::use_data_raw("data_from_instances")
+
+# Packages
+library(aws.s3)
+library(dplyr)
+
+#------------------------------------------------------------------------------#
+# Train/validation/training reference sites: collected from csvs on S3, as these
+# included initial 500 sites used to train each model that were not on the
+# instance databases
+
+# read in incoming and outgoing names files
+# get names of csv files holding sample sites
+# also new Q sites from labeller 8
+str <- glue::glue("aws s3 ls s3://activemapper/planet/")
+planet_bucket <- system(str, intern = TRUE)
+fnames <- sapply(strsplit(planet_bucket, split = " "), function(x) x[length(x)])
+sample_files <- as_tibble(fnames) %>%
+  filter(grepl("incoming_names|outgoing_names|q_sites", value)) %>%
+  filter(!grepl("congo|BF|GH|tanzania|empty|test", value)) %>% pull()
+
+# read in in loop
+calval_pts <- lapply(sample_files, function(x) {  # x <- sample_files[1]
+  print(x)
+  sites <- s3read_using(readr::read_csv, bucket = "activemapper",
+                        object = glue::glue("planet/{x}"))
+  sites %>% mutate(file = gsub("\\.csv", "", x)) %>% select(file, !!names(.))
+})
+
+# new Q sites from labeller8 (after turning on labeller8). These were sites used
+# after initial maps were completed, and created from best labels made during
+# that process
+params <- yaml::yaml.load_file(here::here("common/config.yaml"))
+dinfo <- params$mapper
+host <- "labeller8.crowdmapper.org"
+con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
+                      dbname = "Africa", user = dinfo$db_username,
+                      password = dinfo$db_password)
+newqsites <- tbl(con, "kml_data_static") %>% collect() %>%
+  mutate(file = "qsites_new", name) %>% select(file, name)
+
+# combine with calval_pts
+calval_pts <- append(calval_pts, list(newqsites))
+
+# combine into single tibble
[…]

### 2020-07-24 data-raw/data_for_instances.R
Updated data-raw scripts, added two new ones: map_reference_sample.R and data_from_segmenter.R

@@ -1,12 +1,14 @@
 # Data set up
 
-# Clip down labeller's master grid to just Ghana tiles extent and ID layer for
+#------------------------------------------------------------------------------#
+# Prepare main grid for Ghana
+# Clip down labeller's primary grid to just Ghana tiles extent and ID layer for
 # shipping with activemapper package data
 library(raster)
 library(sf)
 library(dplyr)
 
-# get data from labeller
+# get data from labeller (get from S3)
 data_path <- file.path("/Users", unname(Sys.info()["user"]),
                        "Dropbox/projects/activelearning/mapper/spatial/data")
 master_grid <- brick(file.path(data_path, "processed/master_grid.tif"))
@@ -22,12 +24,14 @@ ghana_grid <- crop(master_grid[[1]], extent(tiles))
 # length(unique(ghana_grid[!is.na(ghana_grid)]))
 # which(duplicated(ghana_grid[!is.na(ghana_grid)]))
 
-writeRaster(ghana_grid , file = "inst/extdata/ghana_grid.tif",
+writeRaster(ghana_grid, file = "inst/extdata/ghana_grid.tif",
             overwrite = TRUE, datatype = "INT4U")  # careful for datatype
 r <- raster("inst/extdata/ghana_grid.tif")
 # length(r[!is.na(r)])
 # length(unique(r[!is.na(r)]))
 
+#------------------------------------------------------------------------------#
+# Create AOIs, tiles, degrees
 # slim down Africa data
 africa <- st_read(file.path(data_path, "external/africa_noisl_gcs.sqlite")) %>%
   st_sf(crs = 4326) %>% select(name, iso3) %>%
@@ -80,76 +84,3 @@ con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
 mgrid <- tbl(con, "master_grid") %>% collect()
 mgrid %>% select(id, name, x, y, fwts) %>%
   readr::write_csv(here::here("inst/extdata/ghana_grid.csv"))
-
-################################################################################
-# Train/validation/training reference sites
-
-# read in incoming and outgoing names files
-# get names of csv files holding sample sites
-# also new Q sites from labeller 8
-str <- glue::glue("aws s3 ls s3://activemapper/planet/")
-planet_bucket <- system(str, intern = TRUE)
[…]

### 2020-07-24 data-raw/data_from_segmenter.R

@@ -1,0 +1,149 @@
+## Processing field segment for:
+## 1. Stratifying validation sample
+## 2. Analyzing field statistics
+
+# Datasets
+# grid, tiles, etc
+mgrid <- raster(
+  system.file("extdata/ghana_grid.tif", package = "activemapper")
+)
+tiles <- st_read(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+
+# Read in segments
+segment_files <- dir(here("external/data/results/segments/merged"),
+                     pattern = "boundarymerge", full.names = TRUE)
+iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
+segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
+
+r <- raster(extent(tiles), res = 0.05)
+crs(r) <- crs(tiles)
+
+# for field areas
+tilesr <- fasterize::fasterize(tiles, r, field = "aoi1")
+tilesr005 <- disaggregate(tilesr, fact = 10)
+tilesr005_sf <- stars::st_as_stars(tilesr005) %>% st_as_sf() %>%
+  rename(aoi1 = layer)
+
+
+#------------------------------------------------------------------------------#
+# Point based stratification for map reference sample selection
+
+##!! Replace with csv grid
+# Create point grid for stratifying model results
+# mask main grid and convert to points
+mgrid <- mask(mgrid, tiles)
+mgrid_xy <- as.data.frame(mgrid, xy = TRUE) %>%
+  st_as_sf(coords = c("x", "y"), crs = 4326) %>% rename(id = ghana_grid) %>%
+  tidyr::drop_na(id)
+
+# join with tiles
+# system.time(extract(tilesr, master_grid_xy))  # 11.81
+system.time(mgrid_xytiles <- st_join(mgrid_xy, tilesr005_sf))  # 26
+
+#------------------------------------------------------------------------------#
+# Intersect points with grids
+logf <- here("external/logs/field_points_005.log")
+registerDoMC(cores = 7)
+fpoints <- foreach(i = segment_files$file) %dopar% {
+  # i <- segment_files$file[1]
+  aoi_id <- gsub("[[:alpha:]]|_|\\.", "", basename(i))
+
[…]

### 2020-07-24 data-raw/map_reference_sample.R

@@ -1,0 +1,55 @@
+## Create map reference sample
+## Reads in intermediate data from data_from_segmenter.R
+
+library(dplyr)
+library(here)
+
+train_val <- readr::read_csv(
+  system.file("extdata", "train_val_sites.csv", package = "activemapper")
+)
+mgrid_tb <- data.table::fread(
+  system.file("extdata/ghana_grid.csv", package = "activemapper")
+) %>% as_tibble()
+
+# intermediate results: stratified points
+fpoints <- data.table::fread(here("external/data/field_points005.csv")) %>%
+  as_tibble %>% filter(aoi != "3N")
+fpointsr <- fpoints %>% filter(!id %in% unique(train_val$id))
+
+#!!!!!!!!! Add sample size analysis to SI
+# Going with sample size of 800 for cropland class, and 683 for non-cropland,
+# per SI.  MOE of 0.03 and 95% confidence interval on assumed cropland
+
+# Take sample from each stratum and combine
+set.seed(111)
+fsample <- fpointsr %>% filter(class == 1) %>% sample_n(size = 800)
+set.seed(111)
+nofsample <- fpointsr %>% filter(class == 0) %>% sample_n(size = 683)
+ref_sample <- bind_rows(fsample, nofsample) %>%
+  mutate(class = as.character(class))
+
+# Join with main grid to get name
+# !!! Note: this step probably not need if fix lines 33-39 in
+# !!! data_from_segmenter.R)
+ref_sample <- left_join(ref_sample %>% select(id, aoi, class), mgrid_tb,
+                        by = "id") %>% select(id, name, aoi, class, x, y)
+# data.table::fwrite(
+#   ref_sample, file = here("external/reference/map_reference_sample.csv")
+# )
+data.table::fwrite(
+  ref_sample, file = here("inst/extdata/map_reference_sample.csv")
+)
+
+# convert to format for incoming_names* to set up new validator instance
+ref_sample_labeller <- ref_sample %>% select(name) %>%
+  mutate(run = 0, iteration = 0, processed = TRUE, usage = "train",
+         label = TRUE)
+
+# write to s3 bucket as input to new validation instance
+aws.s3::s3write_using(
+  ref_sample_labeller,
+  FUN = readr::write_csv, bucket = "activemapper",
+  object = glue::glue("planet/incoming_names_static_validator.csv")
+)
+
+# usethis::use_data("map_reference_sample")

### 2020-07-26 inst/manuscript/analyses-mapreference-sample.Rmd
Update through analysis and reshaping of iteration_metrics

@@
 
+###! Move this to SI
 # sample size function
@@
 
-Remove all points that were used for training, training referenec, or  validation. Contained in CSV of sites prepared in `data_prep.R`
+Remove all points that were used for training, training reference, or  validation. Contained in CSV of sites prepared in `data_prep.R`
 ```{r, eval = FALSE}

### 2020-07-26 inst/manuscript/analyses.Rmd
Update through analysis and reshaping of iteration_metrics

@@
 fpoints_tb <- bind_cols(fpoints %>% as_tibble() %>% select(-geometry), xy)
+
 # data.table::fwrite(

### 2020-07-26 inst/manuscript/manuscript.Rmd
Update through analysis and reshaping of iteration_metrics

@@
 This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details).
-The Bayesian risk of the completed final is then assessed, which measures the average level of consensus for the label assigned to each pixel in a given site, with values closer to 0 indicating complete consensus, and thus the highest confidence, while values approaching 1 reveal progressively increasing disagreement (see SI).
+**The "risk" of the final consensus label is then assessed, which measures the proportion of pixels in the label exceeding a Bayesian risk threshold (see SI), and thus provides an additional measure of label quality.**
 
@@
 data("label_summary")  
-tsites <- label_summary$nlabels %>% group_by(kml_type) %>% 
+toofew <- label_summary$nlabels %>% filter(n < 10) %>% pull(ID)
+tsites <- label_summary$nlabels %>% filter(!ID %in% toofew) %>% 
+  group_by(kml_type) %>% 
   summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
@@
   rename_all(funs(c("ID", "tr", "tref"))) %>% 
-  mutate(trate = tref / tr)
+  mutate(trate = tref / tr) %>% filter(tr > 10)
 ```
@@
 
-After each site was mapped by four labellers, consensus labels were generated.
+After each site was mapped by four labellers, consensus labels were generated. **The average risk for each consensus label was *X* (see SI)**.
 
 ### Model performance
+The average performance associated
 

### 2020-07-26 inst/manuscript/si.Rmd
Update through analysis and reshaping of iteration_metrics

@@
 
-After creating the consensus label, the degree of confidence in the resulting label value can be expressed in terms of Bayesian risk. Risk for a given training sample site *i* is calculated as:
+After creating the consensus label, the degree of confidence in the resulting label value can be expressed in terms of Bayesian risk. Risk is calculated as:
 
-\begin{equation}
-\mathrm{R_i=\frac{1}{n}\sum_{j=1}^{n} C_j(1 - L_j) + (1 - C_j)L_j}
+\begin{equation} \label{eq:pixelrisk}
+\mathrm{r=C(1 - L) + (1 - C)L}
 \end{equation}
+
+Where *C* is the consensus probability that a given pixel is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e.
+non-field if *C* < 0.5, field if *C* > 0.5) for that pixel.
+The risk values across the entire sample site can be processed in two ways to provide useful information about the confidence in the consensus label for that site.
+The first is a simple average of all risk values in the site, where the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}).
+The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling.
+A second approach is to calculate the proportion of pixels having risk values that exceed a certain threshold.
+This second approach is currently implemented in our system, using a threshold value of 0.4.
 
@@
 
+\clearpage
 $$
@@
 
-Using the relevant part of equation 15, T=the corresponding risk associated with this particular pixel's label $\mathrm{R_j}$ is thus `r rj * (1 - 1) + (1 - rj) * 1`.
+Using equation \ref{eq:pixelrisk}, the corresponding risk associated with this particular pixel's label is thus `r rj * (1 - 1) + (1 - rj) * 1`.
 
@@
 The distribution of data collection effort and scores received at training reference sites by labelling team (Figure \ref{fig:trainqual}).
+These results include those from a repeated labelling of the initial training in Cluster 2.
+This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
+We replaced the image overlays in the instance for Cluster 2 \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
+The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.
 
@@
 Labellers' identities are anonymized.", fig.align='center', message=FALSE, results='hold'}
-knitr::include_graphics('figures/si_training_quality_n.png')
+knitr::include_graphics('figures/si_training_qual_n.png')
 ```

### 2020-07-26 data-raw/data_from_instances.R

@@ -384,4 +384,43 @@ label_consensus <- instance_tbls$kml_data %>% filter(aoi %in% aoinms) %>%
 #             nonzeros = sum(ifelse(Risk > 0, 1, 0)))
 usethis::use_data(label_consensus, overwrite = TRUE)
 
+#------------------------------------------------------------------------------#
+# Iteration metrics
+# Analyze and reshape for figure making
+
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+# function to calculate change over time
+delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)
+
+# process individual AOIs
+metrics <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "TPR",
+             "FPR")
+aoinms <- paste0("labeller", 1:16)
+iteration_metrics <- instance_tbls$iteration_metrics %>%
+  filter(aoi %in% aoinms) %>%
+  filter(iteration > 0) %>%
+  filter(!(aoi == "labeller3" & run == 0))  %>% # remove run 0 labeller3
+  mutate(iteration = iteration - 1) %>% # rebase iteration counter on 0
+  mutate(F1 = 2 * (precision * recall) / (precision + recall)) %>%
+  rename(AOI = aoi, Iteration = iteration, Accuracy = accuracy, AUC = auc,
+         TSS = tss, Precision = precision, Recall = recall, TPR = tpr,
+         FPR = fpr) %>%
+  select(AOI, Iteration, !!metrics) %>%
+  group_by(AOI) %>%  # calculate deltas
+  mutate_at(metrics, funs("change" = delta_func)) %>%
+  ungroup() %>%
+  tidyr::pivot_longer(cols = Accuracy:FPR_change, names_to = "Metric",
+                      values_to = "Score")
+
+# process across AOI means and combine with AOIs
+iteration_metrics %>%
+  filter(Iteration < 4) %>%
+  group_by(Iteration, Metric) %>% summarize(Score = mean(Score)) %>%
+  mutate(AOI = "All") %>%
+  ungroup %>%
+  select(AOI, !!names(.)) %>%
+  bind_rows(iteration_metrics, .) -> iteration_metrics
+
+usethis::use_data(iteration_metrics)
 

### 2020-07-26 data-raw/map_reference_sample.R

@@ -33,9 +33,11 @@ ref_sample <- bind_rows(fsample, nofsample) %>%
 # !!! data_from_segmenter.R)
 ref_sample <- left_join(ref_sample %>% select(id, aoi, class), mgrid_tb,
                         by = "id") %>% select(id, name, aoi, class, x, y)
+
+# Write to package
 # data.table::fwrite(
 #   ref_sample, file = here("external/reference/map_reference_sample.csv")
-# )
+# ) # original location
 data.table::fwrite(
   ref_sample, file = here("inst/extdata/map_reference_sample.csv")
 )
@@ -45,7 +47,7 @@ ref_sample_labeller <- ref_sample %>% select(name) %>%
   mutate(run = 0, iteration = 0, processed = TRUE, usage = "train",
          label = TRUE)
 
-# write to s3 bucket as input to new validation instance
+# write to s3 bucket as input for new validation instance
 aws.s3::s3write_using(
   ref_sample_labeller,
   FUN = readr::write_csv, bucket = "activemapper",

### 2020-07-27 inst/manuscript/figures.Rmd
Add figures Rmd, manuscript updated through Figure 5

---
title: Figures
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    math: katex
    toc_depth: 4
    toc: yes
    css: css/vignette.css
---

```{r setup, include=FALSE, cache=FALSE, message = FALSE}
library(raster)
library(sf)
library(dplyr)
library(ggplot2)
library(here)
library(RColorBrewer)
library(activemapper)
```

Creation of figures for main manuscript and the supplementary information.

## Base Datasets

The preparation of datasets used here and which install with the `activemapper` package, is shown in `data-raw/data_[from|to]_instaces.R`. These are loaded in here.

```{r, eval=FALSE}
pth <- function(x) file.path(here("inst/manuscript"), x)

res <- 0.005
res_supergrid <- 1
num_supergrid <- res_supergrid / res

mgridr <- raster(
  system.file("extdata/ghana_grid.tif", package = "activemapper")
)
mgrid <- readr::read_csv(
  system.file("extdata", "ghana_grid.csv", package = "activemapper")
) %>% select(-fwts)
africa <- st_read(
  system.file("extdata/africa.geojson", package = "activemapper")
)
africa_cont <- africa %>% st_union %>% st_sf
ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)

degrees <- st_read(
  system.file("extdata/degrees.geojson", package = "activemapper")
)
tiles <- st_read(
  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
) 
aois <- st_read(
  system.file("extdata/aois.geojson", package = "activemapper")
)
# ggplot(data = tiles) + 
#   geom_sf(aes(fill = as.factor(aoi1)), col = "transparent")

# database results
load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
```

## Methods

### Figure 1: System architecture
Created in keynote

### Figure 2: Grid system and AOIs

```{r, eval = FALSE}
ghana_aois <- tiles %>% pull(aoi1) %>% unique
ghana_degrees <- tiles %>% pull(aoi2) %>% unique

ind <- 8  # index of selected AOI for plotting (from ghana_aois)
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

p1 <- ggplot() + 
  geom_sf(data = africa, fill = "gray90", color = "grey50", size = 0.2) +
  geom_sf(data = ghana, fill = "grey85", col = "transparent", size = 0.3) +
  geom_sf(data = degrees %>% filter(grid %in% ghana_degrees),
          fill = "transparent", col = "grey70", size = 0.4, lty = 3) +
  geom_sf(data = aois, fill = "transparent", col = "black", size = 0.4) +
  geom_sf(data = aois %>% filter(aois == ind), col = "black", fill = "gray70", 
          size = 0.4) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  theme_void()

deg_sel <- tiles %>% as_tibble %>% filter(aoi1 %in% 4:12) %>% 
  distinct(aoi1, aoi2) %>% pull()
deg_sel <- c(1349, 1350, 1351, 1414, 1415, 1416, 1478, 1479, 1480)

xylims <- degrees %>% filter(grid %in% deg_sel) %>% st_coordinates(.)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

tile_sel <- tiles %>% filter(aoi1 == ind)  # selected tile
p2 <- ggplot(ghana) + geom_sf(fill = "grey85", col = "grey") +
  geom_sf(data = degrees %>% filter(grid %in% ghana_degrees),
          fill = "transparent", col = "grey70", size = 0.4, lty = 3) +
  geom_sf(data = aois, fill = "transparent", size = 0.3) +
  geom_sf(data = tile_sel, fill = "gray70", size = 0.3) +
  geom_sf(data = tile_sel %>% slice(190), fill = "red3", size = 0.3) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  theme_void()
# p2

gpols <- crop(mgridr, tile_sel[190, ]) %>% rasterToPolygons(.) %>%
  st_as_sf %>% st_sf(crs = 4326)
p3 <- ggplot(tile_sel %>% slice(169:171, 189:191, 209:211)) +
  geom_sf(fill = "grey80", size = 0.3) +
  geom_sf(data = gpols, fill = "red3", size = 0.3) +
  theme_void()
# p3

gp <- cowplot::plot_grid(p1, p2, p3, nrow = 1, scale = c(1, 1, 0.7), 
                         labels = LETTERS[1:3], label_x = 0.05)
ggsave(pth("figures/figure2.png"), width = 7, height = 3,
       units = "in", dpi = 300)
```

### Figure 3: Labeller interface
Taken as a screenshot.

### Figure S1: Bayesian Risk

A plot of the Bayesian risk curves associated with different levels of consensus, for both a cropland and non-cropland consensus label.
```{r, eval=FALSE}
# Heat map function
consensus <- seq(0, 1, 0.05)
risk <- function(r1, r2) {
  r1 * (1 - r2) + (1 - r1) * r2
}

risk_tbl <- tibble(
  Consensus = consensus, `Non-field` = risk(consensus, 0), 
  Field = risk(consensus, 1)
) %>% pivot_longer(cols = contains("field"), names_to = "Label", 
                   values_to = "Risk")
p <- ggplot(risk_tbl) + geom_line(aes(x = Consensus, y = Risk, color = Label)) +   scale_color_manual(values = c("blue", "red")) + 
  theme_linedraw()

ggsave(pth("figures/si_label_risk.png"), width = 5, 
       height = 4, units = "in", dpi = 300, bg = "transparent")
```

### Figure S2: Map reference sample
```{r, eval = FALSE}
ref_sample <- readr::read_csv(
  system.file("extdata", "map_reference_sample.csv", package = "activemapper")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = ref_sample, aes(color = factor(class)), size = 0.1) +
  scale_color_manual(values = c("red", "blue"), 
                     labels = c("Non-cropland", "Cropland"), name = "") + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(colour = guide_legend(override.aes = list(size = 1, fill = NA))) + 
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))

ggsave(pth("figures/si_map_reference_sample.png"), width = 4, 
       height = 4, units = "in", dpi = 300, bg = "transparent")
```

## Results 
### Figure 4: Composite quality
```{r, eval=FALSE, fig.cap="Placeholder for composite quality image."}
data("image_quality")
tile_quality <- image_quality %>% filter(worker_id == 98) %>% 
  group_by(name, season) %>% 
  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() %>% 
  inner_join(tiles, .)

# nms <- image_quality %>% group_by(name, worker_id) %>% count() %>% ungroup() %>% 
#   filter(worker_id == 98) %>% pull(name)
# image_quality %>% filter(worker_id != 98 & !name %in% nms)
# image_quality %>% filter(name, worker_id)
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- colorRampPalette(rev(brewer.pal(9, "RdBu")))
# sc <- scale_colour_gradientn(colours = myPalette(100), limits=c(1, 8))

p1 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = tiles, fill = "grey85", col = "grey", size = 0.3) +
  geom_sf(data = ghana, fill = "transparent", col = "grey", size = 0.3) +
  geom_sf(data = tile_quality, mapping = aes(fill = Score, color = Score)) +  
  scale_fill_gradientn(colours = pal(20)) + 
  scale_color_gradientn(colours = pal(20)) + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void()
p1
```

### Figure 5: Active learning gain
```{r, eval=FALSE}
# iteration metrics
data("iteration_metrics")

# Plots
cols <- c(rep("grey", length(unique(iteration_metrics$AOI)) -1), "black")
lsize <- c(rep(0.3, length(unique(iteration_metrics$AOI)) -1), 1)
V1 <- c("Accuracy", "AUC", "F1")

# prepare data
# remove labeller from AOI
imetrics <- iteration_metrics %>% 
  filter(Metric %in% V1) %>% 
  mutate(AOI = gsub("labeller", "", AOI))

# select out final iteration for each AOI
mx_metrics <- imetrics %>% group_by(AOI) %>% 
  filter(Iteration == max(Iteration)) %>% filter(Metric %in% V1)

# plot
p <- ggplot(imetrics) + 
  geom_line(aes(Iteration, Score, color = AOI, size = AOI)) +
  geom_point(data =  mx_metrics, aes(Iteration, Score, color = AOI), 
             size = 1) +
  ggrepel::geom_text_repel(
    data = mx_metrics %>% filter(AOI != "All"), 
    aes(label = AOI, x = Iteration, y = Score), col = "grey50", 
    segment.color = "grey80", point.padding = 0.05 
  ) +
  scale_color_manual(values = cols, guide = FALSE) +
  scale_size_manual(values = lsize, guide = FALSE) + 
  scale_y_continuous(breaks = seq(0.2, 1, 0.1), expand = c(0, 0), 
                     limits = c(0.2, 1)) + 
  facet_wrap(vars(Metric), ncol = 3) + 
  theme_linedraw() + 
  theme(strip.background = element_rect(fill = "white"))+
  theme(strip.text = element_text(colour = "black"))

ggsave(pth("figures/figure5.png"), 
       width = 7, height = 4, units = "in", dpi = 300)
```

### Figure S4: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

qsites <- instance_tbls$kml_data %>% 
  filter(aoi %in% paste0("labeller", 1:16)) %>% 
  filter(kml_type == "Q") %>% select(name, x, y) %>% distinct() %>%
  mutate(Cluster = "Training reference") %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326) %>% select(Cluster, name)

train_val_q <- rbind(train_val_sites$initial %>% select(-iteration), qsites)

# initial train cluster panel
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- RColorBrewer::brewer.pal(n = 9, "Spectral")[c(1, 2, 3, 4, 9)]

p1 <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = train_val_q, aes(color = Cluster, fill = Cluster), 
          size = 0.05) +   
  scale_color_manual(values = pal) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(colour = guide_legend(override.aes = list(size = 1, fill = NA))) + 
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))

pal <- c(RColorBrewer::brewer.pal(n = 4, "Spectral"), "white")
p2 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) + 
  geom_sf(data = train_val_sites$active, aes(color = Type), size = 0.05) +
  scale_color_manual(values = pal) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(colour = guide_legend(override.aes = list(size = 1, fill = NA))) + 
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))

# arrange and save
# cp_theme1 <- theme(legend.box.just = "left")
cp_theme1 <- theme(legend.margin = margin(0, 0, 0, 0, "cm"), 
                   legend.text.align = 0, legend.title.align = 0)
gp <- cowplot::plot_grid(p1 + cp_theme1, 
                         p2 + cp_theme1, 
                         nrow = 2,
                         align = "hv", #axis = "l",
                         labels = LETTERS[1:2], label_x = 0.05)
ggsave(pth("figures/si_training_validation_pts.png"), width = 4, 
       height = 6, units = "in", dpi = 300, bg = "transparent")

```

### Figure S5: Label count and quality
```{r labelquality, eval=FALSE}
data("label_summary")

# combine with label_quality
label_quality2 <- left_join(
  label_summary$scores, label_summary$score_stats, by = "uuid"
) %>% select(uuid, score, N, Mean, StDev) %>% 
  mutate(uuid = substr(uuid, 1, 5)) %>% rename(ID = uuid)

# find Q counts to get IDs infrequent workers
toofewq <- label_quality2 %>% group_by(ID) %>% count() %>% filter(n < 10)
label_quality2 <- label_quality2 %>% filter(!ID %in% toofewq$ID)

# Q site accuracy
p1 <- ggplot(label_quality2) + 
  geom_boxplot(aes(x = reorder(ID, -score), y = score, fill = ID), 
               col = "grey50", varwidth = TRUE, show.legend = FALSE,
               outlier.size = 1) + 
  geom_point(aes(x = ID, y = Mean), shape = 4, size = 2) +
  scale_y_continuous(breaks = seq(0, 1, 0.1), limits = c(0, 1), 
                     expand = c(0, 0.01)) + 
  scale_fill_manual(values = rep("grey", length(unique(label_quality2$ID)))) + 
  xlab("ID") + ylab("Score") + 
  theme_linedraw() + 
  theme(axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

# N sites by type
yl <- c(0, ceiling(max(label_summary$nlabels$n) * 1.01))
p2 <- label_summary$nlabels %>% filter(!ID %in% toofewq$ID) %>% 
  group_by(ID, kml_type) %>% 
  ggplot() + 
  geom_bar(aes(x = reorder(ID, -n), y = n, fill = kml_type), 
           position = "dodge", stat = "identity") +
  xlab("ID") + ylab("N") + 
  labs(fill = "Map Type") + 
  scale_fill_manual(values = c("grey40", "grey70")) + 
  scale_y_continuous(breaks = seq(yl[1], yl[2], by = 200), 
                     expand = c(0, 0.01), limits = yl) + 
  theme_linedraw() + 
  theme(legend.position = c(0.7, 0.85), axis.text = element_text(size = 11), 
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

cp_theme <- theme(axis.text = element_text(size = 11),  
                  legend.text = element_text(size = 11),
                  legend.title = element_text(size = 11),
                  axis.title = element_text(size = 12))
gp <- cowplot::plot_grid(p2 + cp_theme, p1 + cp_theme, nrow = 2, 
                         labels = LETTERS[1:2], label_x = 0.0) 
                         #rel_widths = c(0.65, 0.45))
ggsave(pth("figures/si_training_qual_n.png"), width = 4, 
       height = 8, units = "in", dpi = 300, bg = "transparent")
```

### Figure S6: Consenus label (Bayesian) risk

[Not complete: checking values]
```{r, eval=FALSE}
data("label_consensus")
label_consensus_pts <- label_consensus %>% 
  rename(Risk = consensus_conflict) %>% 
  st_as_sf(coords = c("x", "y"), crs = 4326)
label_consensus %>% distinct(aoi)

xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
pal <- colorRampPalette(rev(brewer.pal(9, "RdBu")))

p <- ggplot(africa) + 
  geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
  geom_sf(data = label_consensus_pts, aes(color = Risk, fill = Risk), 
          size = 0.1) +
  scale_color_gradientn(colours = pal(20)) + 
  scale_fill_gradientn(colours = pal(20)) +
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
  guides(fill = guide_colorbar(nbin = 20), color = FALSE) +
  theme_void() + 
  theme(legend.key = element_rect(fill = "grey85", color = "transparent"))
p
label_consensus %>% rename(Risk = consensus_conflict) %>% 
  ggplot() + geom_histogram(aes(x = Risk)) + 
  facet_wrap(vars(aoi))

ggsave("manuscript/figures/si_map_reference_sample.png", width = 4, 
       height = 4, units = "in", dpi = 300, bg = "transparent")

```

```{r, eval=FALSE}
# randomly select cells, as if those used to assess image quality
set.seed(1)
tiles_ss <- tiles %>% sample_n(100)

# set up dummy values of quality, using rainfall as example and regression with
# some random noise added
rftiles <- unlist(extract(rftot, tiles_ss))  # rainfall
tiles_ss <- tiles_ss %>% mutate(rf = rftiles) %>% select(aoi, tile, rf)
rflm <- coef(lm(c(8, 1) ~ range(tiles_ss$rf)))  # regression 1 to 8
qual_preds <- rflm[1] + rflm[2] * tiles_ss$rf  # predict
set.seed(7) 
qual_preds <- round(qual_preds *  # random noise
                      runif(length(qual_preds), min = 0.7, max = 1.3))
qual_preds <- ifelse(qual_preds > 5, 5, qual_preds)  # force to 5
tiles_ss <- tiles_ss %>% mutate(Quality = factor(qual_preds)) %>% 
  select(aoi, tile, rf, Quality)

# random assignment of 1-5 by rainfall
p1 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
  geom_sf(data = tiles, fill = "grey85", col = "grey", size = 0.3) +
  geom_sf(data = ghana, fill = "transparent", col = "grey", size = 0.3) +
  geom_sf(data = tiles_ss, 
          mapping = aes(fill = Quality, color = Quality)) +   
  scale_fill_brewer(palette = "RdBu") + scale_color_brewer(palette = "RdBu") + 
  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
  theme_void()
p2 <- ggplot(tiles_ss) + 
  geom_histogram(aes(x = Quality), bins = 5, stat = "count", col = "grey", 
                 fill = "blue") + 
  xlab("Quality of composite") + ylab("N tiles") + 
  cowplot::theme_cowplot(font_size = 12)
g <- cowplot::plot_grid(p1, p1, p2, p2, nrow = 2, scale = c(1, 1, 0.9), 
                        labels = LETTERS[1:4], label_x = 0.05)
ggsave("manuscript/figures/figure4.png", width = 5, height = 5, 
       units = "in", dpi = 300)
```

### 2020-07-27 inst/manuscript/manuscript.Rmd
Add figures Rmd, manuscript updated through Figure 5

@@
 ### Model performance
+The gain in model performance was assessed for each AOI and for each iteration using a variety of metrics, including overall accuracy, area under the curve of the Receiver Operating Characteristic (AUC), and the F1 score (Figure \ref{fig:alperformance}).
+
+```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting for the Random Forest model after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/figure5.png')
+```
+
 The average performance associated

### 2020-07-30 inst/manuscript/figures.Rmd
Added figures and results from image quality assessment. Added function that uses row, col information from probability information to get x y from reference grid.

@@
           size = 0.4) +
+  geom_sf_text(data = aois, aes(label = aois)) + 
   coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
@@
 ### Figure 4: Composite quality
-```{r, eval=FALSE, fig.cap="Placeholder for composite quality image."}
+```{r, eval=FALSE}
 data("image_quality")
-tile_quality <- image_quality %>% filter(worker_id == 98) %>% 
-  group_by(name, season) %>% 
-  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() %>% 
-  inner_join(tiles, .)
 
-# nms <- image_quality %>% group_by(name, worker_id) %>% count() %>% ungroup() %>% 
-#   filter(worker_id == 98) %>% pull(name)
-# image_quality %>% filter(worker_id != 98 & !name %in% nms)
-# image_quality %>% filter(name, worker_id)
+# plot parameters
 xylims <- st_coordinates(ghana)
@@
 
+leglabs <- seq(0.4, 1, 0.1)
 p1 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
@@
   geom_sf(data = ghana, fill = "transparent", col = "grey", size = 0.3) +
-  geom_sf(data = tile_quality, mapping = aes(fill = Score, color = Score)) +  
-  scale_fill_gradientn(colours = pal(20)) + 
-  scale_color_gradientn(colours = pal(20)) + 
+  geom_sf(data = image_quality$tile, 
+          mapping = aes(fill = Score, color = Score)) +
+  facet_wrap(vars(season), ncol = 2) +
+  scale_fill_gradientn(colors = pal(7), 
+                       breaks = seq(0.4, 1, 0.1), limits = c(0.4, 1),
+                       labels = leglabs) + 
+  scale_color_gradientn(colors = pal(7), 
+                        breaks = seq(0.4, 1, 0.1), limits = c(0.4, 1), 
+                        labels = leglabs) + 
   coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
-  theme_void()
-p1
+  theme_linedraw() + 
+  theme(legend.position="bottom", 
+        strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(colour = "black"), 
+        panel.grid = element_blank(),
+        axis.ticks = element_blank(),
+        axis.text = element_blank())
+
+p2 <- ggplot(as_tibble(image_quality$tile)) +
+  geom_histogram(aes(x = Score), fill = "blue", col = "grey", #bins = 10, 
+                 binwidth = 0.05, boundary = 0.1) + 
+  scale_x_continuous(breaks = leglabs, labels = leglabs, expand = c(0, 0)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
+  facet_wrap(vars(season), ncol = 2) + 
+  ylab("Count") + 
+  theme_linedraw() + 
+  theme(strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(colour = "black"))
+
+g <- cowplot::plot_grid(p1, p2, nrow = 2, align = "hv", axis = "l",
+                        labels = LETTERS[1:2], label_x = 0, 
+                        label_y = c(0.97, 1),   
+                        scale = c(1, 1), rel_heights = c(1, 0.7))
+ggsave(pth("figures/figure4.png"), 
+       width = 4, height = 6.5, units = "in", dpi = 300)
+
 ```
@@
   theme_linedraw() + 
-  theme(strip.background = element_rect(fill = "white"))+
+  theme(strip.background = element_rect(fill = "white"), 
         theme(strip.text = element_text(colour = "black"))

### 2020-07-30 inst/manuscript/manuscript.Rmd
Added figures and results from image quality assessment. Added function that uses row, col information from probability information to get x y from reference grid.

@@
     number_sections: yes
+    extra_dependencies: ["flafter"]
     template: manuscript.latex
@@
 
-```{r, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
+```{r aois, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
 The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
@@
 
-```{r, echo = FALSE, out.width="100%", fig.cap="An overview of analytical scales used in the mapping platform, including A) the areas of interest defining the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), the B) 0.05 degree grid used for creating image composite tiles, and the C) 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
+```{r, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
 knitr::include_graphics('figures/figure2.png')
@@
 
-After fitting, the model is applied to the features of the model validation set, and a set of accuracy metrics are calculated, including ordinary binary accuracy and the true skill statistic [@AlloucheAssessingaccuracyspecies2006], which is considered to be more effective than the Kappa statistic in that it controls for class prevalence.
+After fitting, the model is applied to the features of the model validation set, and a set of accuracy metrics are calculated, including ordinary binary accuracy, the F1 score, and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
 
@@
 
+### Image quality assessment
+```{r, echo=FALSE}
+data("image_quality")
+tile_quality <- image_quality %>% filter(kml_type == "F") %>% 
+  group_by(name, season, worker_id) %>% 
+  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() 
+tilen <- tile_quality %>% distinct(tile) %>% count() %>% pull(n)
+```
+
+We evaluated the overall quality of the resulting seasonal image composites across the region using a random selection of `r tilen`, which we graded according to a four category quality score that evaluated the degree of residual 1) cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image in which resolution was degraded below the 3-4 m PlanetScope resolution (e.g.
+because of between-date image mis-registrations).
+Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), making the highest possible score 12.
+We rescaled scores to fall between 0 and 1.
+
 ### Accuracy assessment
@@
 ## Image catalog
+```{r, echo=FALSE}
+tqual_mu <- image_quality$tile %>%
+  group_by(season, tile) %>% 
+  summarize(score = mean(Score)) %>% 
+  ungroup()
+tqual_sum <- tqual_mu %>% 
+  group_by(season) %>% 
+  summarize(mu = mean(score), sd = sd(score))
+tqmu <- tqual_sum %>% pull(mu) %>% round(2)
+names(tqmu) <- tqual_sum %>% pull(season)
+
+tqual_ct <- tqual_mu %>% group_by(season) %>% 
+  mutate(gt85 = ifelse(score > 0.85, 1, 0)) %>% 
+  summarize(n = n(), n85 = sum(gt85)) %>% 
+  mutate(prop = n85 / n)
+tqct <- tqual_ct %>% pull(prop) * 100
+names(tqct) <- tqual_ct %>% pull(season)
+```
+
 We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$.
@@
 For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
-The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile (**Figure Xa**).
-We also collected and composited imagery for one of the 16 AOIs for the two seasons of the 2019 agricultural year.
+The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile.
+We also composited imagery for AOI 8 (Figure \ref{fig:aois}A) for the two seasons of the 2019 agricultural year.
 All images were converted into Raster Foundry slippy maps for presentation in *labeller*.
 
-We evaluated the overall quality of the resulting composites across the region using a random selection of 100 tiles, which we graded according to a four category quality score that evaluated the degree of residual 1) cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image in which resolution was degraded below the 3-4 m PlanetScope resolution, due to between-date image mis-registrations.
-Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), thus the highest possible score was 12.
-The results showed that the majority of the sampled composites (**X%; Fig Xb**) received a score of **X** or higher (**Y and Z%**), as rated by four independent observers (mean standard deviation of observer scores was **XX**).
-Image quality varied spatially and temporally, with the lower quality images found during in the growing season period and the Southwest corner, where the substantially greater cloud cover resulting in a much lower density of available PlanetScope imagery for each time period (**Fig Xc**).
-Overall, the available image density and quality of composites were closely related (see SI for full details).
+An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["gs"]`, with `r tqct["gs"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["os"]` (`r tqct["os"]` percent $\geq$ 0.85).
+Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period.
 
-```{r, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D) [This a hypothetical placeholder figure].", fig.align='center'}
+<!-- Overall, the available image density and quality of composites were closely related (see SI for full details).   -->
+
+```{r imqual, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D) [This a hypothetical placeholder figure].", fig.align='center'}
 knitr::include_graphics('figures/figure4.png')
@@
 ### Model performance
-The gain in model performance was assessed for each AOI and for each iteration using a variety of metrics, including overall accuracy, area under the curve of the Receiver Operating Characteristic (AUC), and the F1 score (Figure \ref{fig:alperformance}).
+```{r, echo=FALSE}
+data("iteration_metrics")
+delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)
+
+# Difference iteration 0 to 3
+imet_mu03 <- iteration_metrics %>% 
+  filter(AOI == "All") %>%
+  filter(Iteration %in% c(0, 3)) %>%
+  filter(!grepl("_change", Metric)) %>% 
+  group_by(Metric) %>% 
+  mutate(dscore = delta_func(Score)) %>% 
+  ungroup() %>% 
+  mutate(dscore = round(dscore * 100, 1))
+
+imet_mu0 <- imet_mu03 %>% filter(Iteration == 0) %>% 
+  mutate(Score = round(Score, 3))
+imet_mu3 <- imet_mu03 %>% filter(Iteration == 3) %>% 
+  mutate(Score = round(Score, 3))
+
+# differences per iteration
+imet_dmu <- iteration_metrics %>% 
+  filter(AOI == "All" & Iteration > 0) %>%
+  filter(grepl("_change", Metric)) %>% 
+  mutate(Metric = gsub("_change", "", Metric)) %>% 
+  mutate(Score = round(Score * 100, 1)) %>% 
+  filter(Metric %in% c("Accuracy", "AUC", "F1"))
+
+imet_dmu1 <- imet_dmu %>% filter(Iteration == 1) %>% 
+  mutate(dscore = round(Score, 3))
+imet_dmu2 <- imet_dmu %>% filter(Iteration == 2) %>% 
+  mutate(dscore = round(Score, 3))
+imet_dmu3 <- imet_dmu %>% filter(Iteration == 3) %>% 
+  mutate(dscore = round(Score, 3))
+
+# AOI score ranges
+imet_min <- iteration_metrics %>% group_by(AOI) %>%
+  filter(Iteration == max(Iteration) & !grepl("change", Metric)) %>%
+  filter(Metric %in% c("Accuracy", "AUC", "F1")) %>% 
+  ungroup() %>% 
+  group_by(Metric) %>% 
+  filter(Score == min(Score)) %>% 
+  mutate(AOI = gsub("labeller", "", AOI), Score = round(Score, 3))
+
+imet_max <- iteration_metrics %>% group_by(AOI) %>%
+  filter(Iteration == max(Iteration) & !grepl("change", Metric)) %>%
+  filter(Metric %in% c("Accuracy", "AUC", "F1")) %>% 
+  ungroup() %>% 
+  group_by(Metric) %>% 
+  filter(Score == max(Score)) %>% 
+  mutate(AOI = gsub("labeller", "", AOI), Score = round(Score, 3))
+```
+
+The gain in model performance was assessed for each AOI against several accuracy metrics.
+The average accuracy, AUC, and F1 at Iteration 0 were `r imet_mu0 %>% filter(Metric == "Accuracy") %>% pull(Score)`, `r imet_mu0 %>% filter(Metric == "AUC") %>% pull(Score)`, and `r imet_mu0 %>% filter(Metric == "F1") %>% pull(Score)`, respectively, increasing to `r imet_mu3 %>% filter(Metric == "Accuracy") %>% pull(Score)`, `r imet_mu3 %>% filter(Metric == "AUC") %>% pull(Score)`, and `r imet_mu3 %>% filter(Metric == "F1") %>% pull(Score)` by Iteration 3 (Figure \ref{fig:alperformance}).
+These differences represent respective gains of `r imet_mu3 %>% filter(Metric == "Accuracy") %>% pull(dscore)`, `r imet_mu3 %>% filter(Metric == "AUC") %>% pull(dscore)`, and
+`r imet_mu3 %>% filter(Metric == "F1") %>% pull(dscore)` for the three metrics.
+The largest gains in each metric were registered on Iteration 1, averaging `r imet_dmu1 %>% filter(Metric == "Accuracy") %>% pull(Score)`, `r imet_dmu1 %>% filter(Metric == "AUC") %>% pull(Score)`, and `r imet_dmu1 %>% filter(Metric == "F1") %>% pull(Score)` percent for accuracy, AUC, and F1.
+The lowest gains were received on Iteration 3 for accuracy (`r imet_dmu3 %>% filter(Metric == "Accuracy") %>% pull(Score)`%) and F1 (`r imet_dmu3 %>% filter(Metric == "F1") %>% pull(Score)`%), while Iteration 2 showed the lowest average gain for AUC (`r imet_dmu3 %>% filter(Metric == "AUC") %>% pull(Score)`%).
+The range in scores varied substantially across AOIs and metrics.
+Accuracy ranged between `r imet_min %>% filter(Metric == "Accuracy") %>% pull(Score)` (AOI `r imet_min %>% filter(Metric == "Accuracy") %>% pull(AOI)`) and `r imet_max %>% filter(Metric == "Accuracy") %>% pull(Score)` (AOI `r imet_max %>% filter(Metric == "Accuracy") %>% pull(AOI)`), while AUC varied from `r imet_min %>% filter(Metric == "AUC") %>% pull(Score)` (AOI `r imet_min %>% filter(Metric == "AUC") %>% pull(AOI)`) and `r imet_max %>% filter(Metric == "AUC") %>% pull(Score)` (AOI `r imet_max %>% filter(Metric == "AUC") %>% pull(AOI)`), and F1 from `r imet_min %>% filter(Metric == "F1") %>% pull(Score)` (AOI `r imet_min %>% filter(Metric == "F1") %>% pull(AOI)`) and `r imet_max %>% filter(Metric == "F1") %>% pull(Score)` (AOI `r imet_max %>% filter(Metric == "F1") %>% pull(AOI)`).
 
@@
 ```
-
-The average performance associated
-
-**[Fill in with intro]**.
-The average number of iterations per active learning process was **4**, as overall performance gains, as assessed against validation sets, saturated quickly (see **Figure X** in SI).
-
-The average computational time per iteration per AOI was *6 hours*. **Training time. Interaction time**
 

### 2020-07-30 inst/manuscript/si.Rmd
Added figures and results from image quality assessment. Added function that uses row, col information from probability information to get x y from reference grid.

@@
 ## Assessing quality of composites
-The rubric presented in Table 1 was used to assess the quality of the seasonal image composites.
+The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites.
 The imagery was evaluated by examining their Raster Foundry overlays within a `labeller` instance set up for the purpose.
 
-```{r, echo=FALSE, message=FALSE}
+```{r imqcriteria, echo=FALSE, message=FALSE}
 library(kableExtra)
@@
 #### Response design
+
 ##### Stage 1
@@
 ## Image quality
+```{r, echo=FALSE}
+data("image_quality")
+wqual_diff <- image_quality$img %>% filter(kml_type == "F") %>%
+  group_by(name, season, worker_id) %>%
+  summarize(tile = min(tile), score = sum(score) / 12) %>% 
+  ungroup() %>% 
+  pivot_wider(names_from = worker_id, values_from = score,
+              names_prefix = "w") %>%
+  select(tile, season, w92, w99) %>% 
+  mutate(d = w92 - w99, 
+         absd = abs(d))
+```
+
+The image quality assessment was conducted by two separate observers (L.
+Song and Q.
+Zhang).
+Each observer assessed the composites for each season at all 50 of the randomly selected tiles (see main text) following the defined criteria (Table \ref{tab:imqcriteria}).
+To calculate the score for a given tile for each season, we summed the ranking across the four categories for each observer, rescaled the values, and then calculated the mean tile score across both observers by season.
+The mean difference between the two observers was `r d <- round(mean(wqual_diff$d), 3); d` (sd = `r absd <- round(sd(wqual_diff$d), 3); absd`) and the mean absolute difference was `r round(mean(wqual_diff$absd), 3)` (sd = `r round(sd(wqual_diff$absd), 3)`), thus one observer scored tiles about `r abs(round(d * 100))` percent lower than the other observer, on average.
 

### 2020-07-30 data-raw/data_from_instances.R

@@ -207,38 +207,72 @@ instance_tbls$users <- instance_tbls$users %>% distinct(aoi, id, email) %>%
 save(instance_tbls, file = here::here("inst/extdata/instance_tbls.rda"))
 
 #------------------------------------------------------------------------------#
-# Results from labellertest: image composite quality analysis
-# NOTE: 19/7/2020 - this still needs to be completed
-load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
-
+# Image quality check data from image_checker
 library(tidyr)
-mgrid <- readr::read_csv(
-  system.file("extdata", "ghana_grid.csv", package = "activemapper")
-)
 tiles <- st_read(
   system.file("extdata/ghana_tiles.geojson", package = "activemapper")
 )
 
-image_quality <- left_join(
-  instance_tbls$assignment_data %>% filter(aoi == "labellertest"),
-  instance_tbls$hit_data %>% filter(aoi == "labellertest")
-) %>% left_join(., mgrid, by = "name") %>%
-  select(id, name, x, y, assignment_id, worker_id, comment)
+aws_qry <- glue::glue("aws ec2 describe-instances --filters ",
+                      "'Name=tag:Name,Values=imagechecker' --output text ",
+                      "--query 'Reservations[].Instances[][PublicIpAddress]'")
+host <- system(aws_qry, intern = TRUE)
+# browseURL(glue::glue("https://{host}/phpPgAdmin"))
 
-image_quality <- image_quality %>% select(name, x, y, worker_id, comment) %>%
-  filter(!(is.na(comment) | grepl("can't", comment))) %>%
+# credentials
+params <- yaml::yaml.load_file(here::here("common/config.yaml"))
+dinfo <- params$mapper
+con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
+                      dbname = "Africa", user = dinfo$db_username,
+                      password = dinfo$db_password)
+
+# collect data
+image_quality <- tbl(con, "assignment_data") %>%
+  left_join(., tbl(con, "hit_data"), by = "hit_id") %>%
+  left_join(., tbl(con, "kml_data"), by = "name") %>%
+  left_join(., tbl(con, "master_grid"), by = "name") %>%
+  select(id, name, assignment_id, hit_id, x, y, worker_id, kml_type,
+         comment) %>%
+  collect()
[…]

### 2020-07-31 inst/manuscript/manuscript.Rmd
Extracted and processed fields and segements from validation sites.

@@
 ```{r, echo = FALSE, out.width="80%", fig.cap="The overall, user's, and producer's accuracies (indicated by fill color) for each of the 16 mapping zones (zones numbers are shown on maps) [This a hypothetical placeholder figure].", fig.align='center'}
-knitr::include_graphics('figures/figure5.png')
+knitr::include_graphics('figures/si_mapping_blocks.png')
 ```
@@
 
-###
+### Segmentation accuracy
+To evaluate the accuracy of the segmented field boundaries, we calculate the mean and standard deviation of the most accurate (for the given AOI) labeller's maps for validation sites, calculated the mean and standard deviation of area, and compared these statistics to those derived from the model-generated segments for those sites (Figure X).
 

### 2020-07-31 data-raw/data_from_instances.R

@@ -458,3 +458,83 @@ iteration_metrics %>%
 
 usethis::use_data(iteration_metrics)
 
+#------------------------------------------------------------------------------#
+# Extract maps for validation sites of most accurate labeller for each AOI
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+# main grid
+mgrid <- data.table::fread(
+  system.file("extdata/ghana_grid.csv", package = "activemapper")
+)
+
+aoinms <- paste0("labeller", 1:16)
+
+# calculate worker scores
+worker_scores <- instance_tbls$assignment_data %>%
+  rename(id = worker_id) %>%
+  filter(!is.na(score)) %>%
+  filter(aoi %in% aoinms) %>%
+  group_by(aoi, id) %>%
+  summarize(score = mean(score), n = n()) %>%
+  left_join(., instance_tbls$users) %>%
+  rename(wid = id)
+# worker_scores %>% filter(n < 10)
+
+# identify validation sites
+validation_sites <- instance_tbls$incoming_names %>%
+  filter(aoi %in% aoinms) %>%
+  filter(usage == "validate") %>%
+  filter(!(aoi == "labeller3" & run == 0)) %>%
+  select(aoi, name)
+
+# pull out assignments for those sites
+validation_assignments <- instance_tbls$hit_data %>%
+  left_join(validation_sites, .) %>%
+  select(aoi, name, hit_id, delete_time) %>%
+  left_join(., instance_tbls$assignment_data) %>%
+  select(aoi, name, hit_id, assignment_id, worker_id, status, delete_time)
+
+# which assignment was done by most accurate worker to map that assignment
+top_validation_assignments <- validation_assignments %>%
+  rename(wid = worker_id) %>%
+  filter(status %in% c("Approved")) %>% # only approved assignments
+  group_by(aoi, hit_id) %>% # count()
+  left_join(., worker_scores) %>% #select(aoi, name, hit_id, id, score)
+  filter(score == max(score)) %>%
+  select(aoi, name, hit_id, assignment_id, wid, score) %>%
+  ungroup()
+
+# get the associated maps
+top_validation_maps <- left_join(
+  top_validation_assignments, instance_tbls$user_maps,
+  by = c("aoi", "assignment_id")
+) %>%
+  rename(name = name.x, field = name.y) %>%
[…]

### 2020-07-31 data-raw/data_from_segmenter.R

@@ -2,6 +2,11 @@
 ## 1. Stratifying validation sample
 ## 2. Analyzing field statistics
 
+library(raster)
+library(sf)
+library(dplyr)
+library(here)
+
 # Datasets
 # grid, tiles, etc
 mgrid <- raster(
@@ -25,7 +30,6 @@ tilesr <- fasterize::fasterize(tiles, r, field = "aoi1")
 tilesr005 <- disaggregate(tilesr, fact = 10)
 tilesr005_sf <- stars::st_as_stars(tilesr005) %>% st_as_sf() %>%
   rename(aoi1 = layer)
-
 
 #------------------------------------------------------------------------------#
 # Point based stratification for map reference sample selection
@@ -145,5 +149,97 @@ fareas <- foreach(i = segment_files$order, .combine = stack) %dopar% {
 
 writeRaster(fdensity, filename = here("external/data/field_density005.tif"))
 
-
 # usethis::use_data("data_from_segmenter")
+
+#------------------------------------------------------------------------------#
+# extract and summarize field stats for validation sites
+library(dplyr)
+library(ggplot2)
+library(activemapper)
+library(doMC)
+
+data("top_label_data")
+
+gcs = "+proj=longlat +datum=WGS84 +no_defs"
+top_label_grids <- top_label_data$stats %>% select(name, x, y) %>%
+  data.table::as.data.table() %>%
+  rmapaccuracy::point_to_gridpoly(xy = ., 0.005 / 2, gcs, gcs)
+top_label_grids <- top_label_data$stats %>% select(aoi, name) %>%
+  left_join(., top_label_grids)
+top_label_grids <- top_label_grids %>%
+  mutate(aoi = as.numeric(gsub("labeller", "", aoi))) %>%
+  arrange(aoi)
+
+# Read in segments, drop old AOI3
+segment_files <- dir(here("external/data/results/segments/merged"),
+                     pattern = "boundarymerge", full.names = TRUE)
+iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
+segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
+segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))
+
+
+logf <- here("external/logs/field_stats.log")
+aois <- unique(top_label_grids$aoi)
+
[…]

### 2020-08-03 inst/manuscript/figures.Rmd
Revised validation site analysis

@@
 
+### Figure 6: Categorical map accuracy
+
+### Figure 7: Segmentation statistics
+```{r segstats, eval=FALSE}
+data("top_label_data")
+data("field_validation_stats")
+
+human <- top_label_data$stats %>% 
+  rename(hun = n, humu = Mean, hustd = StDev) %>% 
+  select(-contains("id"))
+machine <- field_validation_stats %>% 
+  rename(man = n, mamu = Mean, mastd = StDev)
+left_join(human, machine) %>% View()
+machine %>% arrange(aoi) %>% View()
+
+```
+
 ### Figure S4: Initial training site distribution

### 2020-08-03 data-raw/data_from_instances.R

@@ -520,7 +520,9 @@ top_validation_maps <- top_validation_maps %>%
   as_tibble()
 
 # and summary statistics
-top_validation_stats <- top_validation_maps %>%
+# for fields
+top_validation_flds <- top_validation_maps %>%
+  filter(!is.na(field)) %>%
   group_by(name) %>%
   summarize(aoi = unique(aoi), wid = unique(wid), hit_id = unique(hit_id),
             assignment_id = unique(assignment_id), n = n(),
@@ -529,12 +531,23 @@ top_validation_stats <- top_validation_maps %>%
   select(-fwts) %>%
   select(aoi, name, x, y, hit_id, assignment_id, wid, n, Mean, StDev)
 
+# non fields
+top_validation_noflds <- top_validation_maps %>%
+  filter(is.na(field)) %>%
+  mutate(n = 0, Mean = 0, StDev = NA) %>%
+  left_join(., mgrid, by = "name") %>%
+  select(-fwts) %>%
+  select(aoi, name, x, y, hit_id, assignment_id, wid, n, Mean, StDev)
+
+# bind
+top_validation_stats <- bind_rows(top_validation_flds, top_validation_noflds)
+
 # combine into list
 top_label_data <- list(
   "maps" = top_validation_maps, "stats" = top_validation_stats,
   "scores" = worker_scores
 )
 
-usethis::use_data(top_label_data)
+usethis::use_data(top_label_data, overwrite = TRUE)
 
 

### 2020-08-03 data-raw/data_from_segmenter.R

@@ -177,13 +177,18 @@ iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
 segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
 segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))
 
+# area and quiet func
+ha <- function(x) as.numeric(units::set_units(st_area(x), "ha"))
+quiet <- function(x) suppressWarnings(suppressMessages(x))
 
+# variables
 logf <- here("external/logs/field_stats.log")
 aois <- unique(top_label_grids$aoi)
 
+# loop over aois to process field/grid intersections
 registerDoMC(cores = 7)
-fld_stats <- foreach(x = 1:length(aois)) %dopar% {
-# fld_stats <- lapply(1:length(aois), function(x) {  # x <- 3
+fld_stats <- foreach(x = 1:length(aois)) %dopar% { # x <- 8 #7 #3 #7 #5
+# fld_stats <- lapply(1:length(aois), function(x) {  # x <- 13
 
   cat(glue::glue("Starting labeller{x} at {Sys.time()}"), file = logf,
       sep = "\n", append = TRUE)
@@ -191,55 +196,118 @@ fld_stats <- foreach(x = 1:length(aois)) %dopar% {
   # pull grids
   grids <- top_label_grids %>% filter(aoi == !!aois[x]) %>% st_as_sf()
   flds <- segment_files %>% filter(order == x) %>% pull(file) %>%
-    read_sf()
+    read_sf() %>% select(-tile) %>%
+    filter(!duplicated(.))
 
   # intersect segments with grids (to get fields for validation sites with
   # fields), and then grids with fields filtering for NA to get empty validation
   # sites
-  fld_int <- st_join(flds, grids, left = FALSE)
-  grid_int <- st_join(grids, flds) %>% filter(is.na(id))
+  # intersecting fields and grids
+  fld_int <- quiet(st_join(flds, grids, left = FALSE)) %>% arrange(name) %>%
+    group_by(name, id) %>% mutate(n = n()) %>%
+    summarize(aoi = unique(aoi), n = mean(n)) %>%  # hack to reduce doubles
+    ungroup()
+
+  # field intersections by grid
+  fld_isect <- lapply(unique(grids$name), function(x) {
+    # x <- unique(grids$name)[1]
[…]

### 2020-08-05 inst/manuscript/figures.Rmd
Completed figure and text for segmentation accuracy assessment. Some reordering of text in methods and results.

--- 2020-07-30 inst/manuscript/figures.Rmd
@@
   theme_linedraw() + 
-  theme(legend.position="bottom", 
+  theme(legend.position = "bottom", 
         strip.background = element_rect(fill = "white"), 
@@
 ```
+
+### Figure 6: Categorical map accuracy
 
@@
 ```
+
+### Figure S7: Segmentation statistics
+```{r segstats, eval=FALSE}
+data("top_label_data")
+data("field_validation_stats")
+
+human <- top_label_data$stats %>%
+  mutate(mu_prop = NA, type = "Validation") %>%
+  select(-contains("id"))
+machine <- field_validation_stats %>%
+  mutate(type = "Segmentation")
+humachine <- bind_rows(human, machine) %>% 
+  mutate(aoi = as.numeric(gsub("labeller", "", aoi))) %>% 
+  arrange(aoi)
+
+# Summarize AOI-wise averages (field size)
+humachine_sum <- humachine %>% group_by(aoi, type) %>% 
+  summarize(Median = median(Mean), Mean = mean(Mean)) %>% 
+  ungroup() %>% 
+  tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
+# field number
+humachine_sum_n <- humachine %>% group_by(aoi, type) %>% 
+  summarize(Median = median(n), Mean = mean(n)) %>% 
+  ungroup() %>% 
+  tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
+
+# plots
+fcols <- c(Segmentation = rgb(0, 0, 1, alpha = 0.5),
+          Validation = rgb(1, 0, 0, alpha = 0.5))
+brks <- seq(0, 30, 5)
+theme_fun <- function(x) {
+  theme_linedraw() + 
+  theme(legend.position = "bottom", legend.text = element_text(size = 11),
+        strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(size = 11, colour = "black"),  
+        axis.text.x = element_text(size = 11, hjust = 1, vjust = 0.5, 
+                                   angle = 90), 
+        axis.text.y = element_text(size = 11))
+}
+
+# Field size
+p <- ggplot(humachine) + 
+  geom_density(aes(x = Mean, y = ..count.., fill = type), col = "transparent", 
+               linetype = "blank") +  
+  geom_vline(data = humachine_sum, 
+             aes(xintercept = value, color = type, linetype = Stat), 
+             size = 0.5) +
+  scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 30)) +
+  scale_y_continuous(expand = c(0, 0)) +
+  scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) + 
+  scale_fill_manual(name = "", values = fcols) +  
+  # scale_linetype_manual(name = "", values = c("dashed", "solid")) +  
+  scale_linetype("") +  
+  facet_wrap(vars(aoi)) + 
+  xlab("Field size (ha)") + ylab("Count") + 
+  theme_fun()
+ggsave(pth("figures/si_validation_stats_fha.png"), 
+       width = 7, height = 7, units = "in", dpi = 300)
+
+# field number
+brks <- seq(0, 20, 5)
+p2 <- ggplot(humachine) + 
+  geom_density(aes(x = n, y = ..count.., fill = type), col = "transparent", 
+               linetype = "blank") +  
+  geom_vline(data = humachine_sum_n, 
+             aes(xintercept = value, color = type, linetype = Stat), 
+             size = 0.5) +
+  scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 20)) +
+  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
+  scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) + 
+  scale_fill_manual(name = "", values = fcols) +  
+  # scale_linetype_manual(name = "", values = c("dashed", "solid")) +  
+  scale_linetype("") +  
+  facet_wrap(vars(aoi)) + 
+  xlab("Number of fields") + ylab("Count") + 
+  theme_fun()
+ggsave(pth("figures/si_validation_stats_fn.png"), 
+       width = 7, height = 7, units = "in", dpi = 300)
+
+```

### 2020-08-05 inst/manuscript/manuscript.Rmd
Completed figure and text for segmentation accuracy assessment. Some reordering of text in methods and results.

@@
 library(dplyr)
+library(activemapper)
 
@@
 The first step is the extraction of additional features from the growing season/non-growing season PlanetScope imagery.
-Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatiation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
+Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
 We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of mapping geography that were several orders of magnitude larger.
@@
 
-After fitting, the model is applied to the features of the model validation set, and a set of accuracy metrics are calculated, including ordinary binary accuracy, the F1 score, and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
+After fitting, the model is applied to the features of the model validation set, and a set of performance metrics are calculated, including ordinary binary accuracy, the F1 score (the geometric mean of user's and producer's accuracy), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
 
@@
 
-### Image quality assessment
+### Image quality
 ```{r, echo=FALSE}
@@
 
-### Accuracy assessment
+### Map assessment
+#### Validation performance
+To assess the gain in RandomForest performance due to active learning loop, we assessed the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
+We then compared the magnitude of these iteration-wise changes to those that occurred when each batch of new training sites was randomly selected [@DebatsIntegratingactivelearning2017].
 
-The accuracy assessment procedure described in section 2.2 was used strictly to assess performance gains in the active learning loop, according to suggested  best practices [@ElmesAccountingtrainingdata2020].
-To independently assess the categorical accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a map reference sample.
-To develop this sample, we first made a stratified random sample of grid cells across the entire country, using the mapped field boundaries to define the two strata.
+#### Independent accuracy assessment
+The accuracy assessment procedure described above was not strictly independent because it was re-used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020].
+To independently assess the accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a separate map reference sample.
+To develop this sample, we first used the segmented field boundaries to define a cropland and non-cropland stratum for Ghana, and then randomly selected a sample of the centroids of grid cells intersecting each stratum.
 Next, a primary supervisor examined the image composites in each randomly selected cell, and placed a square polygon of ~0.1 ha on the primary class intersecting the centerpoint of the cell.
@@
 See the SI for further details on the creation and analysis of the map reference sample.
+These points were then intersected with both the RandomForest probability images and the segmented field boundaries, and confusion matrixes between the map reference sample labels and the extracted map classes was constructed to assess the categorical accuracy of each map product.
 
 To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against those of the  workers' digitized polygons at map validation sites.
-We calculated the differences in the measures of central tendency (mean, median, mode) between the two datasets within each validation site, and at the scale of the AOI.
+We selected the field polygons from the most accurate labeller to digitize each of the 100 validation sites in each AOI, and calculated the site-wise average area and number of polygons.
+We then calculated the same statistics from the segmented boundaries that intersected each validation grid.
+We compared the distributions and proximity of two measures of central tendency (mean and median) calculated from the two datasets for each AOI, and across all AOIs.
 
-To further evaluate the capacities of different components of the system, we performed several further experiments.
-In the first experiment, we evaluated the degree to which training label quality impacts final map accuracy, by comparing maps created by training the RandomForest algorithm with labels created by the least accurate worker, the most accurate worker, as well as those trained with the consensus labels.
-This test aligns with Tier 1 standards for accounting for the impact of training data error on final map accuracy [@ElmesAccountingtrainingdata2020].
-Next, we evaluated the degree to which training sample selection via active learning impacted map accuracy relative to random selection of additional training sites.
-Lastly, we repeated the entire map development and map assessment process for one AOI for the 2019 agricultural year, in order to assess the ability of the system to create annually updated maps and detect inter-annual change.
+## The impact of training data error
+To quantify the potential deleterious impact of training data error on our map results, we compared maps created by training the RandomForest algorithm with three sets of labels: 1) those from the lowest scoring worker to map each training site, 2) those from the most accurate worker, and 3) those trained as usual with the consensus labels.
+We then quantified the prediction performance against the validation sites, and compared the differences in performance metrics for each variant.
+This assessment follows recommended Tier 1 (i.e.
+best practice) standards to account for training data errors in machine learning-based Earth Observation applications [@ElmesAccountingtrainingdata2020].
+
+```{r, echo=FALSE}
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+kml_types <- instance_tbls$kml_data %>% filter(aoi == "labeller8v2") %>% 
+  group_by(kml_type) %>% count() %>% filter(kml_type == "Q") %>% pull(n)
+```
+
+As a final test of training data quality on map accuracy, we retrained the model for AOI 8 for 2018 using new labels that were generated against an updated training reference library (n = `r kml_types`), and compared the performance of this model against the previous version of the model.
 
@@
 Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
-For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or recently felled or newly replanted tree crops.
-Our rationale for this decision was that the resolution of PlanetScope imagery is insufficient to distinguish such crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.
+For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or replanted tree crops.
+We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.
 
-## Image catalog
+## Image catalog and quality
 ```{r, echo=FALSE}
+data("image_quality")
 tqual_mu <- image_quality$tile %>%
@@
 
-### Model performance
+### Validation performance
 ```{r, echo=FALSE}
@@
 
+We compared the gain in performance at AOIs 1, 8, and 15 against models trained for those same locations using the initial random draw of 500, followed by increments of 100 randomly selected points per iteration.
+
 ## Map accuracy
@@
 ### Segmentation accuracy
-To evaluate the accuracy of the segmented field boundaries, we calculate the mean and standard deviation of the most accurate (for the given AOI) labeller's maps for validation sites, calculated the mean and standard deviation of area, and compared these statistics to those derived from the model-generated segments for those sites (Figure X).
+```{r, echo=FALSE}
+# validation statistics
+data("top_label_data")
+data("field_validation_stats")
+
+human <- top_label_data$stats %>%
+  mutate(mu_prop = NA, type = "val") %>%
+  select(-contains("id"))
+machine <- field_validation_stats %>%
+  mutate(type = "seg")
+humachine <- bind_rows(human, machine) %>% 
+  mutate(aoi = as.numeric(gsub("labeller", "", aoi))) %>% 
+  arrange(aoi)
+
+huma_aois <- humachine %>% group_by(aoi, type) %>% 
+  summarize(med = median(Mean), mu = mean(Mean)) %>% 
+  ungroup()
+lrgst <- huma_aois %>% select(-med) %>% filter(aoi %in% c(2, 3, 7, 15)) %>% 
+  tidyr::pivot_wider(names_from = type, values_from = "mu") %>% 
+  filter(seg %in% range(seg))
+lrgstlow <- lrgst %>% filter(seg == min(seg))
+lrgsthigh <- lrgst %>% filter(seg == max(seg))
+
+# overall stats
+huma_sum <- humachine %>% group_by(type) %>% 
+  summarize(med = median(Mean), sd = sd(Mean), min = min(Mean), 
+            max = max(Mean), mu = mean(Mean)) %>% 
+  select(type, mu, med, sd, min, max)
+segha <- huma_sum %>% filter(type == "seg") %>% pull(mu) %>% round(2)
+valha <- huma_sum %>% filter(type == "val") %>% pull(mu) %>% round(2)
+seghamed <- huma_sum %>% filter(type == "seg") %>% pull(med) %>% round(2)
+valhamed <- huma_sum %>% filter(type == "val") %>% pull(med) %>% round(2)
+seghasd <- huma_sum %>% filter(type == "seg") %>% pull(sd) %>% round(2)
+valhasd <- huma_sum %>% filter(type == "val") %>% pull(sd) %>% round(2)
+
+huma_sumn <- humachine %>% group_by(type) %>% 
+  summarize(med = median(n), sd = sd(n), min = min(n), max = max(n), 
+            mu = mean(n)) %>% 
+  select(type, mu, med, sd, min, max)
+segn <- huma_sumn %>% filter(type == "seg") %>% pull(mu) %>% round(2)
+valn <- huma_sumn %>% filter(type == "val") %>% pull(mu) %>% round(2)
+segnmed <- huma_sumn %>% filter(type == "seg") %>% pull(med) %>% round(2)
+valnmed <- huma_sumn %>% filter(type == "val") %>% pull(med) %>% round(2)
+segnsd <- huma_sumn %>% filter(type == "seg") %>% pull(sd) %>% round(2)
+valnsd <- huma_sumn %>% filter(type == "val") %>% pull(sd) %>% round(2)
+```
+The comparisons of digitized versus segmented field boundaries showed that the mean field size across all 1600 validation sites averaged `r round(segha, 2)` ha (StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
+This discrepancy was primarily caused by segmentation results in four AOIs (2, 3, 7, and 15; Figure S5), where segments averaged between `r round(lrgstlow$seg, 2)` and `r round(lrgsthigh$seg, 2)` ha, compared to `r round(lrgstlow$val, 2)` - `r round(lrgsthigh$val, 2)` for the corresponding hand-digitized polygons.
+In contrast, the median segmented field area was closer to that of digitized polygons (`r seghamed` for segments versus `r valhamed` for labeller polygons).
+The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S6).
 

### 2020-08-05 inst/manuscript/si.Rmd
Completed figure and text for segmentation accuracy assessment. Some reordering of text in methods and results.

@@
 
-The distribution of data collection effort and scores received at training reference sites by labelling team (Figure \ref{fig:trainqual}).
-These results include those from a repeated labelling of the initial training in Cluster 2.
+The distribution of data collection effort and scores from training reference sites by labelling team (Figure \ref{fig:trainqual}).
+These results include those from repeat labelling of the initial training in Cluster 2.
 This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
@@
 knitr::include_graphics('figures/si_training_qual_n.png')
+```
+
+## Map accuracy 
+### Categorical
+[To fill]
+
+### Field area and number
+To mean, median, and distributions of the average area of segmented field boundaries over the 100 validation sites in each AOI are compared to the areas of the polygons digitized by the most accurate labeller over the same sites in Figure \ref{fig:areavalidation}.
+The same statistics for average number of segments versus average number of labelled polygons across validation sites in each AOI are shown in Figure \ref{fig:numbervalidation}.
+```{r areavalidation, echo = FALSE, out.width="100%", fig.cap="The distributions of the average areas (in hectares) of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average areas of field boundaries digitized (shown in red) by the most accurate worker to label each site.
+Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_validation_stats_fha.png')
+```
+
+```{r numbervalidation, echo = FALSE, out.width="100%", fig.cap="The distributions of the average number of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average number of digitized polygons (shown in red) by the most accurate worker to label each site.
+The mean and median of each distribution is shown.
+Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_validation_stats_fn.png')
 ```

### 2020-08-07 inst/manuscript/figures.Rmd
Updated analyses and SI, manuscript text to include results from active versus random training

@@
 library(activemapper)
+pth <- function(x) file.path(here("inst/manuscript"), x)
 ```
@@
 ```{r, eval=FALSE}
-pth <- function(x) file.path(here("inst/manuscript"), x)
 
@@
 # Plots
-cols <- c(rep("grey", length(unique(iteration_metrics$AOI)) -1), "black")
-lsize <- c(rep(0.3, length(unique(iteration_metrics$AOI)) -1), 1)
 V1 <- c("Accuracy", "AUC", "F1")
@@
 # prepare data
-# remove labeller from AOI
+# remove labeller from AOI and select AOIs 1-16
+aois <- c(paste0("labeller", 1:16), "All")
 imetrics <- iteration_metrics %>% 
   filter(Metric %in% V1) %>% 
+  filter(AOI %in% aois) %>% 
   mutate(AOI = gsub("labeller", "", AOI))
@@
 # plot
+cols <- c(rep("grey", length(unique(imetrics$AOI)) -1), "black")
+lsize <- c(rep(0.3, length(unique(imetrics$AOI)) -1), 1)
 p <- ggplot(imetrics) + 
@@
   geom_point(data = mx_metrics, aes(Iteration, Score, color = AOI), 
-             size = 1) +
+             size = 1.2) +
   ggrepel::geom_text_repel(
@@
     aes(label = AOI, x = Iteration, y = Score), col = "grey50", 
-    segment.color = "grey80", point.padding = 0.05 
+    segment.color = "grey60", point.padding = 0.1, min.segment.length = 0.2,
+    segment.size = 0.2, size = 3   
   ) +
@@
   scale_size_manual(values = lsize, guide = FALSE) + 
-  scale_y_continuous(breaks = seq(0.2, 1, 0.1), expand = c(0, 0), 
+  scale_x_continuous(expand = c(0.01, 0.01)) + 
+  scale_y_continuous(breaks = seq(0.2, 1, 0.1), expand = c(0.01, 0.01), 
                      limits = c(0.2, 1)) + 
@@
   theme(strip.background = element_rect(fill = "white"), 
-        theme(strip.text = element_text(colour = "black"))
+        strip.text = element_text(colour = "black"))
 
@@
 
-### Figure S6: Consenus label (Bayesian) risk
+### Figure S6: Consensus label (Bayesian) risk
 
@@
 
+### Figure S7: Random versus active sampling
 ```{r, eval=FALSE}
-# randomly select cells, as if those used to assess image quality
-set.seed(1)
-tiles_ss <- tiles %>% sample_n(100)
+data("iteration_metrics")
 
-# set up dummy values of quality, using rainfall as example and regression with
-# some random noise added
-rftiles <- unlist(extract(rftot, tiles_ss))  # rainfall
-tiles_ss <- tiles_ss %>% mutate(rf = rftiles) %>% select(aoi, tile, rf)
-rflm <- coef(lm(c(8, 1) ~ range(tiles_ss$rf)))  # regression 1 to 8
-qual_preds <- rflm[1] + rflm[2] * tiles_ss$rf  # predict
-set.seed(7) 
-qual_preds <- round(qual_preds *  # random noise
-                      runif(length(qual_preds), min = 0.7, max = 1.3))
-qual_preds <- ifelse(qual_preds > 5, 5, qual_preds)  # force to 5
-tiles_ss <- tiles_ss %>% mutate(Quality = factor(qual_preds)) %>% 
-  select(aoi, tile, rf, Quality)
+# Plots
+V1 <- c("Accuracy", "AUC", "F1")
 
-# random assignment of 1-5 by rainfall
-p1 <- ggplot(africa) + geom_sf(fill = "gray97", color = "grey", size = 0.25) +
-  geom_sf(data = tiles, fill = "grey85", col = "grey", size = 0.3) +
-  geom_sf(data = ghana, fill = "transparent", col = "grey", size = 0.3) +
-  geom_sf(data = tiles_ss, 
-          mapping = aes(fill = Quality, color = Quality)) +   
-  scale_fill_brewer(palette = "RdBu") + scale_color_brewer(palette = "RdBu") + 
-  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
-  theme_void()
-p2 <- ggplot(tiles_ss) + 
-  geom_histogram(aes(x = Quality), bins = 5, stat = "count", col = "grey", 
-                 fill = "blue") + 
-  xlab("Quality of composite") + ylab("N tiles") + 
-  cowplot::theme_cowplot(font_size = 12)
-g <- cowplot::plot_grid(p1, p1, p2, p2, nrow = 2, scale = c(1, 1, 0.9), 
-                        labels = LETTERS[1:4], label_x = 0.05)
-ggsave("manuscript/figures/figure4.png", width = 5, height = 5, 
-       units = "in", dpi = 300)
+# prepare data
+# remove labeller from AOI and select AOIs 1-16
+aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
+imetrics <- iteration_metrics %>% 
+  filter(Metric %in% V1 & AOI %in% aois) %>% 
+  mutate(AOI = gsub("labeller", "", AOI)) %>% 
+  mutate(type = case_when(
+    AOI %in% c(1, 8, 15) ~ "active", 
+    AOI %in% c("1r", "8r", "15r") ~ "random"
+  )) %>% 
+  mutate(AOI = gsub("r", "", AOI)) %>% 
+  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
+  mutate(delta = (active - random) / random * 100)
+
+imetrics_all <- imetrics %>% filter(Iteration < 4) %>% 
+  group_by(Iteration, Metric) %>% 
+  summarize(active = mean(active), random = mean(random), 
+            delta = mean(delta)) %>%
+  ungroup() %>% mutate(AOI = "All") %>% 
+  select(AOI, Iteration, Metric, active, random, delta)
+imetrics <- bind_rows(imetrics, imetrics_all)
+
+# select out final iteration for each AOI
+mx_metrics <- imetrics %>% group_by(AOI) %>% 
+  filter(Iteration == max(Iteration)) %>% filter(Metric %in% V1)
+
+# plot
+cols <- c(rep("grey50", length(unique(imetrics$AOI)) -1), "black")
+lsize <- c(rep(0.3, length(unique(imetrics$AOI)) -1), 1)
+p <- ggplot(imetrics) + 
+  geom_hline(yintercept = 0, linetype = "dotted", col = "grey40") +
+  geom_line(aes(Iteration, delta, color = AOI, size = AOI)) +
+  geom_point(data = mx_metrics, aes(Iteration, delta, color = AOI), 
+             size = 1.2) +
+  ggrepel::geom_text_repel(
+    data = mx_metrics %>% filter(AOI != "All"), 
+    aes(label = AOI, x = Iteration, y = delta), col = "grey50", 
+    segment.color = "grey60", point.padding = 0.2, min.segment.length = 0.2,
+    segment.size = 0.2, size = 3, nudge_x = 0 
+  ) +
+  scale_color_manual(values = cols, guide = FALSE) +
+  scale_size_manual(values = lsize, guide = FALSE) + 
+  scale_x_continuous(expand = c(0.01, 0.01)) + 
+  scale_y_continuous(expand = c(0.01, 0.01), breaks = seq(-8, 4, 1)) + 
+  ylab("% difference") +                    
+  facet_wrap(vars(Metric), ncol = 3) +
+  theme_linedraw() + 
+  theme(panel.grid.minor.x = element_blank(), 
+        strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(colour = "black"))
+
+ggsave(pth("figures/si_random_vs_active.png"), 
+       width = 6, height = 4, units = "in", dpi = 300)
+
 ```
 
-### Figure S7: Segmentation statistics
+### Figure S8: Segmentation statistics
 ```{r segstats, eval=FALSE}

### 2020-08-07 inst/manuscript/manuscript.Rmd
Updated analyses and SI, manuscript text to include results from active versus random training

@@
 
-- name: Azavea
-  affilnum: '7'
-
 affiliation:
@@
 
-- affilnum: 5
+- affilnum: 6
   affil: Meridia
-
-- affilnum: 5
-  affil: Azavea
 
@@
 #### Validation performance
-To assess the gain in RandomForest performance due to active learning loop, we assessed the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
-We then compared the magnitude of these iteration-wise changes to those that occurred when each batch of new training sites was randomly selected [@DebatsIntegratingactivelearning2017].
+To assess the gain in RandomForest performance due to active learning, we assessed the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
+To assess whether active learning improved the performance relative to a purely random approach to selecting new training sites, we ran an additional test at a subset of AOIs (1, 8, and 15).
+We first randomly selected and labelled 300-400 sites in each AOI.
+After initial training of AOIs 1, 8, and 15 with their initial pools of 500 training sites (Cluster 1 for AOI 1 and Cluster 2 for AOIs 8 and 15), we then retrained the model for each AOI by progressively adding 100 of the new randomly selected labels, evaluating model performance against the validation set after each retraining.
+We repeated this process so that the number of retrainings and training samples matched those from the active learning process for each AOI.
+We then compared the difference in accuracy, AUC, and F1 between these randomly trained models and those trained with active learning [@DebatsIntegratingactivelearning2017].
 
@@
 
+V1 <- c("Accuracy", "AUC", "F1")
+
 # Difference iteration 0 to 3
@@
   mutate(Score = round(Score * 100, 1)) %>% 
-  filter(Metric %in% c("Accuracy", "AUC", "F1"))
+  filter(Metric %in% V1)
 
@@
 # AOI score ranges
-imet_min <- iteration_metrics %>% group_by(AOI) %>%
+imin <- iteration_metrics %>% 
+  filter(AOI %in% paste0("labeller", 1:16)) %>% group_by(AOI) %>%
   filter(Iteration == max(Iteration) & !grepl("change", Metric)) %>%
-  filter(Metric %in% c("Accuracy", "AUC", "F1")) %>% 
+  filter(Metric %in% V1) %>% 
   ungroup() %>% 
@@
 
-imet_max <- iteration_metrics %>% group_by(AOI) %>%
+imax <- iteration_metrics %>% 
+  filter(AOI %in% paste0("labeller", 1:16)) %>% group_by(AOI) %>%
   filter(Iteration == max(Iteration) & !grepl("change", Metric)) %>%
-  filter(Metric %in% c("Accuracy", "AUC", "F1")) %>% 
+  filter(Metric %in% V1) %>% 
   ungroup() %>% 
@@
   mutate(AOI = gsub("labeller", "", AOI), Score = round(Score, 3))
+
+# summarize and condense for easier in line extraction
+# iteration 0 and 0-3 metrics
+imu0 <- lapply(V1, function(x) {
+  imu <- round(imet_mu0 %>% filter(Metric == x) %>% pull(Score), 3)
+  idf <- round(imet_mu0 %>% filter(Metric == x) %>% pull(dscore), 3)
+  ix <- list("score" = imu, "diff" = idf)
+  ix
+})
+names(imu0) <- V1
+
+imu3 <- lapply(V1, function(x) {
+  imu <- round(imet_mu3 %>% filter(Metric == x) %>% pull(Score), 3)
+  idf <- round(imet_mu3 %>% filter(Metric == x) %>% pull(dscore), 3)
+  ix <- list("score" = imu, "diff" = idf)
+  ix
+})
+names(imu3) <- V1
+
+# iteration-wise diffs
+idmu <- lapply(list(imet_dmu1, imet_dmu2, imet_dmu3), function(x) {
+  l <- lapply(V1, function(y) {
+    imu <- round(x %>% filter(Metric == y) %>% pull(Score), 3)
+    idf <- round(x %>% filter(Metric == y) %>% pull(dscore), 3)
+    ix <- list("score" = imu, "diff" = idf)
+    ix
+  })
+  names(l) <- V1
+  l
+})
+names(idmu) <- paste0("I", 1:3)
 ```
@@
 The gain in model performance was assessed for each AOI against several accuracy metrics.
-The average accuracy, AUC, and F1 at Iteration 0 were `r imet_mu0 %>% filter(Metric == "Accuracy") %>% pull(Score)`, `r imet_mu0 %>% filter(Metric == "AUC") %>% pull(Score)`, and `r imet_mu0 %>% filter(Metric == "F1") %>% pull(Score)`, respectively, increasing to `r imet_mu3 %>% filter(Metric == "Accuracy") %>% pull(Score)`, `r imet_mu3 %>% filter(Metric == "AUC") %>% pull(Score)`, and `r imet_mu3 %>% filter(Metric == "F1") %>% pull(Score)` by Iteration 3 (Figure \ref{fig:alperformance}).
-These differences represent respective gains of `r imet_mu3 %>% filter(Metric == "Accuracy") %>% pull(dscore)`, `r imet_mu3 %>% filter(Metric == "AUC") %>% pull(dscore)`, and
-`r imet_mu3 %>% filter(Metric == "F1") %>% pull(dscore)` for the three metrics.
-The largest gains in each metric were registered on Iteration 1, averaging `r imet_dmu1 %>% filter(Metric == "Accuracy") %>% pull(Score)`, `r imet_dmu1 %>% filter(Metric == "AUC") %>% pull(Score)`, and `r imet_dmu1 %>% filter(Metric == "F1") %>% pull(Score)` percent for accuracy, AUC, and F1.
-The lowest gains were received on Iteration 3 for accuracy (`r imet_dmu3 %>% filter(Metric == "Accuracy") %>% pull(Score)`%) and F1 (`r imet_dmu3 %>% filter(Metric == "F1") %>% pull(Score)`%), while Iteration 2 showed the lowest average gain for AUC (`r imet_dmu3 %>% filter(Metric == "AUC") %>% pull(Score)`%).
+The average accuracy, AUC, and F1 at iteration 0 were `r imu0$Accuracy$score`, `r imu0$AUC$score`, and `r imu0$F1$score`, respectively, increasing to `r imu3$Accuracy$score`, `r imu3$AUC$score`, and `r imu3$F1$score` by iteration 3 (Figure \ref{fig:alperformance}).
+These differences represent respective gains of `r imu3$Accuracy$diff`, `r imu3$AUC$diff`, and `r imu3$F1$diff` percent for the three metrics.
+The largest gains in each metric were registered on iteration 1, averaging `r idmu$I1$Accuracy$score`, `r idmu$I1$AUC$score`, and `r idmu$I1$F1$score` percent for accuracy, AUC, and F1.
+The lowest gains were received on Iteration 3 for accuracy (`r idmu$I3$Accuracy$score`%) and F1 (`r idmu$I3$F1$score`%), while AUC gains were least on iteration 2 (`r idmu$I2$AUC$score`%).
 The range in scores varied substantially across AOIs and metrics.
-Accuracy ranged between `r imet_min %>% filter(Metric == "Accuracy") %>% pull(Score)` (AOI `r imet_min %>% filter(Metric == "Accuracy") %>% pull(AOI)`) and `r imet_max %>% filter(Metric == "Accuracy") %>% pull(Score)` (AOI `r imet_max %>% filter(Metric == "Accuracy") %>% pull(AOI)`), while AUC varied from `r imet_min %>% filter(Metric == "AUC") %>% pull(Score)` (AOI `r imet_min %>% filter(Metric == "AUC") %>% pull(AOI)`) and `r imet_max %>% filter(Metric == "AUC") %>% pull(Score)` (AOI `r imet_max %>% filter(Metric == "AUC") %>% pull(AOI)`), and F1 from `r imet_min %>% filter(Metric == "F1") %>% pull(Score)` (AOI `r imet_min %>% filter(Metric == "F1") %>% pull(AOI)`) and `r imet_max %>% filter(Metric == "F1") %>% pull(Score)` (AOI `r imet_max %>% filter(Metric == "F1") %>% pull(AOI)`).
+Accuracy ranged between `r imin[imin$Metric == "Accuracy", ]$Score` (AOI `r imin[imin$Metric == "Accuracy", ]$AOI`) and `r imax[imax$Metric == "Accuracy", ]$Score` (AOI `r imax[imax$Metric == "Accuracy", ]$AOI`), while AUC varied from `r imin[imin$Metric == "AUC", ]$Score` (AOI `r imin[imin$Metric == "AUC", ]$AOI`) and `r imax[imax$Metric == "AUC", ]$Score` (AOI `r imax[imax$Metric == "AUC", ]$AOI`), and F1 from `r imin[imin$Metric == "F1", ]$Score` (AOI `r imin[imin$Metric == "F1", ]$AOI`) and `r imax[imax$Metric == "F1", ]$Score` (AOI `r imax[imax$Metric == "F1", ]$AOI`).
 
@@
 
-We compared the gain in performance at AOIs 1, 8, and 15 against models trained for those same locations using the initial random draw of 500, followed by increments of 100 randomly selected points per iteration.
+```{r, echo=FALSE}
+# random versus active data
+data("iteration_metrics")
+
+# process data: remove labeller from AOI and select AOIs 1-16
+V1 <- c("Accuracy", "AUC", "F1")
+aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
+imetrics <- iteration_metrics %>% 
+  filter(Metric %in% V1 & AOI %in% aois) %>% 
+  mutate(AOI = gsub("labeller", "", AOI)) %>% 
+  mutate(type = case_when(
+    AOI %in% c(1, 8, 15) ~ "active", 
+    AOI %in% c("1r", "8r", "15r") ~ "random"
+  )) %>% 
+  mutate(AOI = gsub("r", "", AOI)) %>% 
+  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
+  mutate(delta = (active - random) / random * 100) 
+imetrics_all <- imetrics %>% 
+  filter(Iteration < 4) %>% 
+  group_by(Iteration, Metric) %>% 
+  summarize(active = mean(active), random = mean(random), 
+            delta = mean(delta)) %>%
+  ungroup() %>% mutate(AOI = "All") %>% 
+  select(AOI, Iteration, Metric, active, random, delta)
+
+ixl <- lapply(1:3, function(x) {
+  ix <- imetrics_all %>% filter(Iteration == x) %>% pull(delta)
+  ix <- round(ix, 1)
+  names(ix) <- V1
+  ix
+})
+names(ixl) <- as.roman(1:3)
+
+i1acc <- imetrics %>% 
+  filter(Iteration == 1 & Metric == "Accuracy") %>% 
+  pull(delta) %>% round(2)
+names(i1acc) <- imetrics %>% distinct(AOI) %>% pull()
+
+u <- function(x) unname(x)
+```
+
+The training tests at AOIs 1, 8, and 15 showed that active learning produced higher performance than randomized model training after three iterations, with accuracy, AUC, and F1 being respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher after three iterations (Figure S5).
+However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2.
+The lower accuracy under active leaving was caused by results at AOI 15, where accuracy gain was `r abs(u(i1acc["15"]))` percent lower under active training (although accuracy slightly between iterations; see Figure \ref{fig:alperformance}).
+In comparison, active learning accuracies after iteration 1 were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
 

### 2020-08-07 inst/manuscript/si.Rmd
Updated analyses and SI, manuscript text to include results from active versus random training

@@
 
+### Validation
+The differences in accuracy, AUC, and F1 between the active learning process and the random retraining at each iteration for AOIs 1, 8, and 15 (Figure \ref{fig:randomvactive}).
+Small differences due to random variations in the RandomForest models are the reason for the non-zero differences at iteration 0, when both models were trained with the same level set.
+```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites.
+Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_random_vs_active.png')
+```
+
 ## Map accuracy 

### 2020-08-07 data-raw/data_from_instances.R

@@ -430,11 +430,12 @@ delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)
 # process individual AOIs
 metrics <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "TPR",
              "FPR")
-aoinms <- paste0("labeller", 1:16)
+aoinms <- paste0("labeller", c(1:16, "1r", "8r", "15r", "8v2", "82019"))
 iteration_metrics <- instance_tbls$iteration_metrics %>%
   filter(aoi %in% aoinms) %>%
   filter(iteration > 0) %>%
-  filter(!(aoi == "labeller3" & run == 0))  %>% # remove run 0 labeller3
+  filter(!(aoi == "labeller3" & run == 0)) %>% # remove run 0 labeller3
+  filter(!(aoi %in% paste0("labeller", c("1r", "8r")) & run == 1)) %>% # no run1
   mutate(iteration = iteration - 1) %>% # rebase iteration counter on 0
   mutate(F1 = 2 * (precision * recall) / (precision + recall)) %>%
   rename(AOI = aoi, Iteration = iteration, Accuracy = accuracy, AUC = auc,
@@ -442,21 +443,24 @@ iteration_metrics <- instance_tbls$iteration_metrics %>%
          FPR = fpr) %>%
   select(AOI, Iteration, !!metrics) %>%
   group_by(AOI) %>%  # calculate deltas
-  mutate_at(metrics, funs("change" = delta_func)) %>%
+  mutate_at(metrics, list("change" = delta_func)) %>%
   ungroup() %>%
   tidyr::pivot_longer(cols = Accuracy:FPR_change, names_to = "Metric",
                       values_to = "Score")
 
 # process across AOI means and combine with AOIs
+aois <- paste0("labeller", 1:16)  # calculate summary across just AOIs 1-16
 iteration_metrics %>%
+  filter(AOI %in% aois) %>%
   filter(Iteration < 4) %>%
-  group_by(Iteration, Metric) %>% summarize(Score = mean(Score)) %>%
+  group_by(Iteration, Metric) %>%
+  summarize(Score = mean(Score)) %>%
   mutate(AOI = "All") %>%
   ungroup %>%
   select(AOI, !!names(.)) %>%
   bind_rows(iteration_metrics, .) -> iteration_metrics
 
-usethis::use_data(iteration_metrics)
+usethis::use_data(iteration_metrics, overwrite = TRUE)
 
[…]

### 2020-08-10 inst/manuscript/figures.Rmd
Updated data analysis, figures, SI, manuscript through training data error impact section of results

@@
 
-### Figure 6: Categorical map accuracy
+### Figure 6: The impact of training data error
+```{r, eval=FALSE}
+data("consensus_high_low")
+
+V1 <- c("Accuracy", "AUC", "F1")
+
+# calculate mean across AOIs and join. Convert to factor and reorder for
+# plotting
+consensus_high_low_mu <- consensus_high_low %>% 
+  group_by(Strategy, Metric) %>% 
+  summarize(Score = mean(Score)) %>% 
+  ungroup() %>% 
+  mutate(AOI = "Mean") %>% 
+  bind_rows(consensus_high_low, .) %>% 
+  filter(Metric %in% V1) %>% 
+  mutate(AOI = factor(AOI)) %>% 
+  mutate(AOI = forcats::fct_reorder(AOI, ord))
+
+cols <- c("grey50", "grey60", "grey70", "grey80", "black")
+sizes <- c(rep(1.5, 4), 2)
+shapes <- c(3, 4, 8, 10, 16)
+p <- ggplot(consensus_high_low_mu) + 
+  geom_point(aes(x = Strategy, y = Score, color = AOI, size = AOI, 
+                 shape = AOI)) + 
+  ylim(0.4, 1) + xlab("Labelling strategy") + 
+  scale_color_manual(values = cols) +
+  scale_size_manual(values = sizes) + 
+  scale_shape_manual(values = shapes) + 
+  facet_wrap(~Metric) + 
+  theme_linedraw() + 
+  theme(legend.position = "bottom", 
+        strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(colour = "black"))
+ggsave(pth("figures/figure6.png"), height = 4, width = 6, dpi = 300)
+```
+
+### Figure 7: Categorical map accuracy
 
@@
 
-### Figure S8: Segmentation statistics
+### Figure S8: Training data error impact on probability images
+```{r, eval = FALSE}
+probpal <- viridis::viridis
+cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
+hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
+lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))
+
+png(pth("figures/si_label_strategy_probs.png"), height = 2.5,
+    width = 8, res = 300, units = "in", bg = "transparent")
+par(mfrow = c(1, 3), mar = c(0, 0.5, 0, 1.75), oma = c(0, 0, 0, 2))
+plot(as(extent(cprob), "SpatialPolygons"), lty = 0)
+plot(cprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
+     col = probpal(20))
+mtext("A", side = 3, line = -2.5, adj = -0.02, cex = 1)
+plot(as(extent(cprob), "SpatialPolygons"), lty = 0, cex.main = 2)
+plot(hprob, lty = 0, add = TRUE, legend = FALSE, zlim = c(0, 100),
+     col = probpal(20))
+mtext("B", side = 3, line = -2.5, adj = -0.02, cex = 1)
+plot(as(extent(cprob), "SpatialPolygons"), lty = 0)
+plot(lprob, lty = 0, add = TRUE, zlim = c(0, 100),
+     legend.width = 1, legend.shrink = 0.5, col = probpal(20))
+mtext("C", side = 3, line = -2.5, adj = -0.02, cex = 1)
+dev.off()
+
+```
+
+### Figure S9: Segmentation statistics
 ```{r segstats, eval=FALSE}

### 2020-08-10 inst/manuscript/manuscript.Rmd
Updated data analysis, figures, SI, manuscript through training data error impact section of results

@@
 
-#### Independent accuracy assessment
+#### Evaluating the impact of training data error
+To quantify the potential impact of training data error on our map results, we compared maps created by training the RandomForest algorithm with three sets of labels: 1) those from the lowest scoring worker to map each training site, 2) those from the most accurate worker, and 3) those trained as usual with the consensus labels.
+We then quantified the prediction performance against the validation sites, and compared the differences in performance metrics for each variant.
+This assessment follows recommended Tier 1 (i.e.
+best practice) standards to account for training data errors in machine learning-based Earth Observation applications [@ElmesAccountingtrainingdata2020].
+
+```{r, echo=FALSE}
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+kml_types <- instance_tbls$kml_data %>% filter(aoi == "labeller8v2") %>% 
+  group_by(kml_type) %>% count() %>% filter(kml_type == "Q") %>% pull(n)
+```
+
+As a final test of training data quality on map accuracy, we retrained the model for AOI 8 for 2018 using new labels that were generated against an updated training reference library (n = `r kml_types`), and compared the performance of this model against the previous version of the model.
+
+#### Accuracy assessment
 The accuracy assessment procedure described above was not strictly independent because it was re-used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020].
@@
 
-## The impact of training data error
-To quantify the potential deleterious impact of training data error on our map results, we compared maps created by training the RandomForest algorithm with three sets of labels: 1) those from the lowest scoring worker to map each training site, 2) those from the most accurate worker, and 3) those trained as usual with the consensus labels.
-We then quantified the prediction performance against the validation sites, and compared the differences in performance metrics for each variant.
-This assessment follows recommended Tier 1 (i.e.
-best practice) standards to account for training data errors in machine learning-based Earth Observation applications [@ElmesAccountingtrainingdata2020].
-
-```{r, echo=FALSE}
-load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
-kml_types <- instance_tbls$kml_data %>% filter(aoi == "labeller8v2") %>% 
-  group_by(kml_type) %>% count() %>% filter(kml_type == "Q") %>% pull(n)
-```
-
-As a final test of training data quality on map accuracy, we retrained the model for AOI 8 for 2018 using new labels that were generated against an updated training reference library (n = `r kml_types`), and compared the performance of this model against the previous version of the model.
-
 <!-- Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.  -->
@@
 
-## The characteristics of Ghanaian cropland
+## Assessing the characteristics of Ghanaian cropland
 Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the characteristics of field sizes through the country.
-We further used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.
+We also used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.
 
@@
 
+### The impact of training data error
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("consensus_high_low")
+V1 <- c("Accuracy", "AUC", "F1")
+delta_func <- function(x, y, rnd = 1) round((x - y) / y * 100, rnd)
+
+# calculate mean across AOIs and join. Convert to factor and reorder for
+# plotting
+consensus_high_low_mu <- consensus_high_low %>% 
+  group_by(Strategy, Metric) %>% 
+  summarize(Score = mean(Score)) %>% 
+  ungroup()
+
+metdiffs <- lapply(V1, function(x) {  # x <- "F1"
+  cn <- consensus_high_low_mu %>% 
+    filter(Strategy == "Consensus" & Metric == x) %>% pull(Score)
+  h <- consensus_high_low_mu %>% filter(Strategy == "High" & Metric == x) %>% 
+    pull(Score)
+  l <- consensus_high_low_mu %>% filter(Strategy == "Low" & Metric == x) %>% 
+    pull(Score)
+  cnh <- delta_func(cn, h)
+  cnl <- delta_func(cn, l)
+  hl <- delta_func(h, l)
+  list(cons = round(cn, 3), high = round(h, 3), low = round(l, 3), 
+       chdelt = cnh, cldelt = cnl, hldelt = hl) 
+})
+names(metdiffs) <- V1
+```
+
+The potential impact of training data errors on map quality was assessed in four AOIs (1, 2, 8, and 15).
+The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
+Models trained with the most accurate worker's labers had scores that were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than models trained with the most accurate worker's lablers (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate worker's labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
+
+```{r trainingimpact, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from Random Forest models trained with consensus labels, and labels made by the most and least accurate worker to label each site.
+Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/figure6.png')
+```
+
 ## Map accuracy
@@
 
+- The impact of training data error on map quality:
+  - Label quality makes a large difference. 
+  - Although these differences were assessed against the validation sites, which were themselves consensus labels, there are real differences in quality that can be seen in the resulting probability images, where probabilities from the model trained with the least accurate worker's labels are clearly less certain (Figure S6).
+
 - Left unquantified is how much training on base maps matters

### 2020-08-10 inst/manuscript/si.Rmd
Updated data analysis, figures, SI, manuscript through training data error impact section of results

@@
 
+### The impact of training data error
+Probability images resulting from Random Forest models trained with labels generated by 1) consensus and the 2) most and 3) least accurate worker to digitize each training site are illustrated in Figure \ref{fig:labelstrategy}.
+These images were created for a single tile in AOI 1.
+
+```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_strategy_probs.png')
+```
+
 ## Map accuracy 
@@
 
-## Experiments
-### Label quality
-
-```{r, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_hml_probs.png')
-```
-
 \clearpage

### 2020-08-10 data-raw/data_from_instances.R

@@ -554,4 +554,63 @@ top_label_data <- list(
 
 usethis::use_data(top_label_data, overwrite = TRUE)
 
+#------------------------------------------------------------------------------#
+# Metrics for highest, lowest, most accurate worker
+library(aws.s3)
+library(glue)
+
+
+fnames <- unlist(lapply(paste0(c('low', 'high'), ".csv"), function(x) {
+  paste0(glue("planet/incoming_metrics_{c(1, 2, 8, 15)}_"), x)
+}))
+high_low_metrics <- lapply(fnames, function(x) {
+  dat <- s3read_using(FUN = readr::read_csv, object = x,
+                      bucket = "activemapper")
+  highlow <- gsub("planet/incoming_metrics_[0-9]+_|.csv", "", x)
+  dat %>% mutate(aoi = paste0(aoi, "_", highlow)) %>% rename(AOI = aoi) %>%
+    select(AOI, !!names(.))
+}) %>% do.call(rbind, .)
+
+# Process and reshape
+metrics <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "TPR",
+             "FPR")
+high_low_metrics <- high_low_metrics %>%
+  mutate(iteration = iteration - 1) %>% # rebase iteration counter on 0
+  mutate(F1 = 2 * (precision * recall) / (precision + recall)) %>%
+  rename(Iteration = iteration, Accuracy = accuracy, TSS = tss,
+         Precision = precision, Recall = recall, TPR = tpr,
+         FPR = fpr) %>%
+  select(AOI, Iteration, !!metrics) %>%
+  tidyr::pivot_longer(cols = Accuracy:FPR, names_to = "Metric",
+                      values_to = "Score")
+
+# save out for keeping
+usethis::use_data(high_low_metrics, overwrite = TRUE)
+
+# bring in iteration metrics and join with consensus values from those
+consensus <- iteration_metrics %>%
+  filter(AOI %in% paste0("labeller", c(1, 2, 8, 15))) %>%
+  group_by(AOI) %>%
+  filter(Iteration == max(Iteration) & !grepl("change", Metric)) %>%
+  ungroup() %>%
+  mutate(Strategy = "Consensus") %>%
+  mutate(AOI = gsub("labeller", "", AOI))
+
+# join
+consensus_high_low <- high_low_metrics %>%
+  mutate(Strategy = tools::toTitleCase(gsub("[0-9]+_", "", AOI)),
[…]

### 2020-08-12 inst/manuscript/figures.Rmd
Reprocessed all labels from all instances to risk statistics and subset of maps, using consensus_risk.R. Includes refactoring of the consensus code here, added as package functions--these should come from rmapaccuracy instead once tested there.

@@
 library(dplyr)
+library(tidyr)
 library(ggplot2)
@@
 # remove labeller from AOI and select AOIs 1-16
-aois <- c(paste0("labeller", 1:16), "All")
+aoinms <- c(paste0("labeller", 1:16), "All")
 imetrics <- iteration_metrics %>% 
   filter(Metric %in% V1) %>% 
-  filter(AOI %in% aois) %>% 
+  filter(AOI %in% aoinms) %>% 
   mutate(AOI = gsub("labeller", "", AOI))
@@
   st_as_sf(coords = c("x", "y"), crs = 4326)
-label_consensus %>% distinct(aoi)
+# label_consensus %>% distinct(aoi)
 

### 2020-08-12 inst/manuscript/manuscript.Rmd
Reprocessed all labels from all instances to risk statistics and subset of maps, using consensus_risk.R. Includes refactoring of the consensus code here, added as package functions--these should come from rmapaccuracy instead once tested there.

@@
 
+To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team.
+Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an **X%** agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average **X% smaller/larger** than those mapped on the ground.
+
 ### Validation performance
@@
 The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
-Models trained with the most accurate worker's labers had scores that were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than models trained with the most accurate worker's lablers (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate worker's labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
+Models trained with the most accurate worker's labels had scores that were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than models trained with the most accurate worker's lablers (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate worker's labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
 
@@
 ```
+
+A second measure of the impact of training data error can be obtained by comparing the average label risk per AOI against the model performance metrics....
 
@@
 
-## System evaluation
 ### Label quality and map accuracy
-```{r, echo=FALSE}
-load(here::here("external/data/label_experiment_metrics.rda"))
-consensus <- scores %>% filter(Label == "Consensus") %>% 
-  select(AUC, Accuracy, TSS) 
-high <- scores %>% filter(Label == "High") %>% select(AUC, Accuracy, TSS) 
-low <- scores %>% filter(Label == "Low") %>% select(AUC, Accuracy, TSS)
-auc <- high$AUC / low$AUC - 1
-tss <- (high$TSS + 1) / (low$TSS + 1) - 1
-acc <- high$Accuracy / low$Accuracy - 1
-cauc <- consensus$AUC / high$AUC - 1
-ctss <- (consensus$TSS + 1) / (high$TSS + 1) - 1
-cacc <- consensus$Accuracy / high$Accuracy - 1
-```
-
-We used the validation labels to evaluate the accuracy of maps generated in three AOIs (1, 8, and 15; **Figure X**) with the least and most accurate workers' labels, and then compared the resulting scores to accuracy assessed using the consensus labels (**Figure 6**).
-On average, there was a substantial difference between maps trained with the most and least accurate workers' labels, with the former being on average `r round(mean(acc) * 100, 1)` percent higher (range `r round(min(acc) * 100, 1)`-`r round(max(acc) * 100, 1)`%) in terms of accuracy, `r round(mean(tss) * 100, 1)` percent higher (range `r round(min(tss) * 100, 1)`-`r round(max(tss) * 100, 1)`%) in skill (measured with TSS, which ranges from -1 to 1).
-Maps generated with consensus labels were on average `r round(mean(cacc) * 100, 1)` percent more accurate (range `r round(min(cacc) * 100, 1)`-`r round(max(cacc) * 100, 1)`%) and `r round(mean(ctss) * 100, 1)` (range `r round(min(ctss) * 100, 1)`-`r round(max(ctss) * 100, 1)`% more skillful.
-AUC scores showed similar average differences to TSS with respect to both comparisons; on average `r round(mean(auc) * 100, 1)` higher when comparing maps generated with the most and least accurate labels, and `r round(mean(cauc) * 100, 1)` percent more skillful when comparing maps genereated with consensus labels compared to those with the most accurate worker's labels.
-Although these comparisons were assessed against consensus-generated validation labels, a visual assessment of probability images generated from each labelling strategy shows markedly greater uncertainty between consensus and high accuracy label-generated maps relative to those generated with low accuracy (**see Figure SX in SI**).
-
 ```{r, echo = FALSE, out.width="100%", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.", fig.align='center', message=FALSE}
@@
 
-To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team.
-Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an **X%** agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average **X% smaller/larger** than those mapped on the ground.
-**However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but may instead be at least partially an indication of how different the perspectives are as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
-
-### Is active learning any better than random? 
-To evaluate whether active learning is more effective than a purely random approach to training site selection, we retrained the RandomForest model on the same three AOIs (1, 8, 15) using sites that were selected purely at random for all iterations after the initial training draw.
-After the mapping team labelled these sites, we subsequently retrained the models with each increment of randomly selected sites, and evaluated prediction performance against the original validation set.
-
-### Segmentation quality
-- field size class distributions
-- Improvement of accuracy
-
 ## Ghana's croplands
@@
   - Training reference to training ratio was artificially inflated. Workers not always notified when training labels complete, and thus only served training reference sites. 
+  - Image-interpreted versus ground labels: **However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but may instead be at least partially an indication of how different the perspectives are as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
 
@@
   - Label quality makes a large difference. 
-  - Although these differences were assessed against the validation sites, which were themselves consensus labels, there are real differences in quality that can be seen in the resulting probability images, where probabilities from the model trained with the least accurate worker's labels are clearly less certain (Figure S6).
+  - Although these comparisons were assessed against consensus-generated validation labels, a visual assessment of probability images generated from each labelling strategy shows markedly greater uncertainty between consensus and most accurate individual labels maps relative to those generated with the lowest quality labels are clearly less certain (Figure S6).
 

### 2020-08-12 R/consensus_map_functions.R

@@ -1,0 +1,379 @@
+#' Get pointers to key database tables from labeller
+#' @param connection postgresql connection to instance database
+#' @param dbtables Vector of table names to get. Currently "default" brings up
+#' names of tables needed for consensus labelling
+#' @return Named list of lazy connections to labeller database tables
+#' @export
+get_db_tables <- function(connection, dbtables = "default") {
+  if(dbtables == "default") {
+    tblist <- c("kml_data", "hit_data", "assignment_data", "accuracy_data",
+                "qual_assignment_data", "qual_accuracy_data",
+                "master_grid", "categories", "configuration")
+  } else {
+    tblist <- dbtables
+  }
+  dbtbls <- lapply(tblist, function(x) {
+    tab <- dplyr::tbl(connection, x)
+  })
+  names(dbtbls) <- tblist
+  return(dbtbls)
+}
+
+#' Collection assignment history for a given worker
+#' @param assignment_data Assignment data table
+#' @param accuracy_data Accuracy data table
+#' @param workerid Integer ID of worker
+#' @return Tibble of measurements
+#' @importFrom magrittr %>%
+#' @importFrom dplyr filter pull select rename
+#' @keywords internal
+assignment_history <- function(assignment_data, accuracy_data, workerid) {
+  historyassignmentid <- assignment_data %>%
+    filter(worker_id == workerid) %>%
+    filter(status == "Approved" | status == "Rejected") %>%
+    filter(!(is.null(score) | is.na(score))) %>% pull(assignment_id)
+
+  if(length(historyassignmentid) == 0) {
+    measurements <- data.frame(ml.field = NA, ml.nofield = NA,
+                               score.hist = NA)
+  } else {
+    measurements <- accuracy_data %>%
+      filter(assignment_id %in% historyassignmentid) %>%
+      dplyr::select(new_score, field_skill, nofield_skill) %>%
+      rename(ml.field = field_skill, ml.nofield = nofield_skill,
+             score.hist = new_score)
+  }
+  return(measurements)
+}
+
+#' Collection qualification assignment history for a given worker
+#' @param qual_assignment_data Assignment data table
+#' @param qual_accuracy_data Qualification accuracy table
+#' @param workerid Integer ID of worker
+#' @return Tibble of measurements
+#' @importFrom magrittr %>%
+#' @importFrom dplyr filter pull select rename group_by ungroup
+#' @keywords internal
[…]

### 2020-08-12 data-raw/consensus_risk.R

@@ -1,0 +1,198 @@
+# Calculate Bayesian risk metrics for labels
+library(rmapaccuracy)
+library(dplyr)
+library(data.table)
+library(sf)
+library(activemapper)
+library(doMC)
+
+load(here::here("external/data/aois/dbases/instance_dbases.rda"))
+ref_grid <- fread(
+  system.file("extdata", "ghana_grid.csv", package = "activemapper")
+)
+
+# select AOIs and unique names:
+aoinms <- c(
+  "labeller",
+  paste0("labeller", c("trainsouth", "trainsouthnl", "trainsw", 1:16))
+)
+# lapply(aoinms, function(x) nrow(instance_dbases[[x]]$accuracy_data))
+
+logf <- here::here("external/logs/consensus_risk.log")
+if(file.exists(logf)) {
+  file.copy(logf, paste0(logf, "bak"))
+  file.remove(logf)
+}
+starttime <- Sys.time()
+labeldir <- here::here("external/data/aois/labels/")
+
+registerDoMC(cores = 6)
+risk_statsl <- lapply(aoinms, function(x) {  # x <- "labellertrainsw"
+  # x <- "labeller1"
+
+  cat(glue::glue("Starting {x} at {Sys.time()}"), file = logf,
+      sep = "\n", append = TRUE)
+
+  db <- instance_dbases[[x]]  # select dbase
+
+  # ggplot2::ggplot(db$assignment_data) +
+  #   ggplot2::geom_histogram(ggplot2::aes(start_time))
+  #
+  # db$incoming_names %>% filter(run == 1)
+  # db$assignment_data %>%
+
+  # get tables
+  kml_data <- db$kml_data
+  assignment_data <- db$assignment_data
+  accuracy_data <- db$accuracy_data
+  qual_assignment_data <- db$qual_assignment_data
+  qual_accuracy_data <- db$qual_accuracy_data
+  user_maps <- db$user_maps
+  hit_data <- db$hit_data
+  categories <- db$categories
+
+  # filter for labeller3
+  if(x == "labeller3") {
+    # run0nms <- db$incoming_names %>% filter(run == 0 & label == TRUE)
+    # run1nms <- db$incoming_names %>% filter(run == 1 & label == TRUE)
+    # run1nms$name[run0nms$name %in% run1nms$name]
+    # hit_data %>% filter(name %in% run1nms$name)
+    # db$iteration_metrics
+    # hit_data %>%
[…]

### 2020-08-12 data-raw/data_from_instances.R

@@ -113,15 +113,16 @@ dbnames <- paste0(
 )
 
 nonspatial_tables <- c("accuracy_data", "assignment_data", "assignment_history",
-                       "configuration", "hit_data", "incoming_names",
-                       "iteration_metrics", "qual_accuracy_data",
+                       "categories", "configuration", "hit_data",
+                       "incoming_names", "iteration_metrics",
+                       "qual_accuracy_data",
                        "qual_assignment_data", "worker_data", "users")
 
 # credentials
 params <- yaml::yaml.load_file(here::here("common/config.yaml"))
-dinfo <- params$mapper
+dinfo <- params$labeller
 
-db_tables <- lapply(dbnames, function(x) {  # x <- dbnames[7]
+instance_dbases <- lapply(dbnames, function(x) {  # x <- dbnames[7]
   con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
                         dbname = x, user = dinfo$db_username,
                         password = dinfo$db_password)
@@ -169,10 +170,10 @@ db_tables <- lapply(dbnames, function(x) {  # x <- dbnames[7]
   return(tbls)
 })
 # dbnames[dbnames == "labellertrainsouth92019"] <- "labellertrainsouth"
-names(db_tables) <- dbnames  # rename labellertrainsouth92019 to old name
+names(instance_dbases) <- dbnames  # rename labellertrainsouth92019 to old name
 
 # short-term save
-save(db_tables, file = "external/data/aois/dbases/instance_dbases.rda")
+save(instance_dbases, file = "external/data/aois/dbases/instance_dbases.rda")
 # load("external/data/aois/dbases/instance_dbases.rda")
 
 # check for drift in values across dates (check.R)
@@ -182,13 +183,13 @@ save(db_tables, file = "external/data/aois/dbases/instance_dbases.rda")
 
 load("external/data/aois/dbases/instance_dbases.rda")
 
-tbl_names <- names(db_tables$labeller)
+tbl_names <- names(instance_dbases$labeller)
 instance_tbls <- lapply(tbl_names, function(x) {  # x <- tbl_names[11]
   print(x)
-  instance_tbl <- lapply(names(db_tables), function(yy) {
-    # yy <- names(db_tables)[1]
+  instance_tbl <- lapply(names(instance_dbases), function(yy) {
+    # yy <- names(instance_dbases)[1]
[…]

### 2020-08-14 inst/manuscript/figures.Rmd
Label risk assessment completed and integrated into SI and main text.

@@
 ```{r, eval=FALSE}
-
 res <- 0.005
@@
 
-### Figure S4: Initial training site distribution
+### Figure S3: Initial training site distribution
 Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
@@
 
-### Figure S5: Label count and quality
+### Figure S4: Label count and quality
 ```{r labelquality, eval=FALSE}
@@
 
-### Figure S6: Consensus label (Bayesian) risk
+### Figures S5 and S6: Consensus label (Bayesian) risk
 
-[Not complete: checking values]
-```{r, eval=FALSE}
-data("label_consensus")
-label_consensus_pts <- label_consensus %>% 
-  rename(Risk = consensus_conflict) %>% 
+Plots of mean risk in space, divided into training and validation sites
+
+```{r riskpoints, eval=FALSE}
+data("label_risk")
+
+# Training/validation points
+label_risk_pts <- label_risk$pts %>% 
   st_as_sf(coords = c("x", "y"), crs = 4326)
 # label_consensus %>% distinct(aoi)
+# label_risk %>% filter(is.na(meanrisk))
+
+# # prepare for histograms
+# clusters <- paste0("labeller", c("", "trainsouth", "trainsouthnl", "trainsw"))
+# new_cluster_nms <- paste0("Cl", c("1", "2a", "2b", "3"))
+# 
+# label_risk_tb <- as_tibble(label_risk_pts) %>% 
+#   mutate(aoi = case_when(
+#     aoi == clusters[1] ~ "-3",
+#     aoi == clusters[2] ~ "-2",
+#     aoi == clusters[3] ~ "-1",
+#     aoi == clusters[4] ~ "0", 
+#     TRUE ~ gsub("labeller", "", aoi) 
+#   )) %>% 
+#   mutate(ord = as.numeric(aoi)) %>% arrange(aoi) %>% 
+#   mutate(aoi = case_when(
+#     aoi == "-3" ~ new_cluster_nms[1], 
+#     aoi == "-2" ~ new_cluster_nms[2],
+#     aoi == "-1" ~ new_cluster_nms[3],
+#     aoi %in% "0" ~ new_cluster_nms[4], 
+#     TRUE ~ aoi 
+#   )) %>% 
+#   mutate(aoi = forcats::fct_reorder(aoi, ord))
+# 
+# # summarize
+# label_risk_stats <- label_risk_tb %>% group_by(aoi, usage) %>% 
+#   summarize(Risk = mean(Risk, na.rm = TRUE)) %>% 
+#   ungroup()
 
@@
 maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
-pal <- colorRampPalette(rev(brewer.pal(9, "RdBu")))
+pal <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))
+zlim <- c(0, round(max(label_risk_pts$Risk, na.rm = TRUE), 1))
 
@@
   geom_sf(data = aois, fill = "grey85", col = "black", size = 0.3) +
-  geom_sf(data = label_consensus_pts, aes(color = Risk, fill = Risk), 
+  geom_sf(data = label_risk_pts, aes(color = Risk, fill = Risk), 
           size = 0.1) +
-  scale_color_gradientn(colours = pal(20)) + 
-  scale_fill_gradientn(colours = pal(20)) +
+  facet_wrap(~usage) + 
+  scale_color_gradientn(colours = pal(20), limits = zlim) + 
+  scale_fill_gradientn(colours = pal(20), limits = zlim) +
   coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) +
@@
   theme(legend.key = element_rect(fill = "grey85", color = "transparent"))
-p
-label_consensus %>% rename(Risk = consensus_conflict) %>% 
-  ggplot() + geom_histogram(aes(x = Risk)) + 
-  facet_wrap(vars(aoi))
+ggsave(pth("figures/si_label_risk_map.png"), width = 7, 
+       height = 5, units = "in", dpi = 300, bg = "transparent")
+```
 
-ggsave("manuscript/figures/si_map_reference_sample.png", width = 4, 
-       height = 4, units = "in", dpi = 300, bg = "transparent")
-
+Risk histograms. 
+```{r riskhists, eval=FALSE}
+p <- ggplot(label_risk$pts) + 
+  geom_histogram(aes(x = Risk), fill = "grey50", bins = 20) + 
+  geom_vline(data = label_risk$stats, aes(xintercept = Risk), size = 0.5, 
+             linetype = 1) + 
+  scale_y_continuous(expand = c(0, 0), breaks = seq(0, 150, 25), 
+                     labels = c("", 25, "", 75, "", 125, ""), 
+                     limits = c(0, 150)) + 
+  ylab("Number of sites") + 
+  facet_grid(aoi ~ usage) +
+  theme_linedraw() + 
+  theme(strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(size = 9, colour = "black"),  
+        axis.text.x = element_text(size = 9, hjust = 1, vjust = 0.5, 
+                                   angle = 90), 
+        axis.text.y = element_text(size = 7))
+ggsave(pth("figures/si_label_risk_hists.png"), width = 3, 
+       height = 9, units = "in", dpi = 300, bg = "transparent")
 ```

### 2020-08-14 inst/manuscript/manuscript.Rmd
Label risk assessment completed and integrated into SI and main text.

@@
 This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details).
-**The "risk" of the final consensus label is then assessed, which measures the proportion of pixels in the label exceeding a Bayesian risk threshold (see SI), and thus provides an additional measure of label quality.**
+As a further measure of confidence in the final consensus label, its average Bayes Risk is calculated (see SI).
+This measure ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels (n = 40000) in the label, and 1 indicating complete disagreement.
 
@@
 
-After each site was mapped by four labellers, consensus labels were generated. **The average risk for each consensus label was *X* (see SI)**.
-
-To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team.
-Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an **X%** agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average **X% smaller/larger** than those mapped on the ground.
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("label_risk")
+murisk <- label_risk$stats %>% group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3)) %>% 
+  tidyr::pivot_wider(names_from = usage, values_from = Risk)
+mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
+  filter(Risk == min(Risk))
+maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
+  filter(Risk == max(Risk))
+minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
+  filter(Risk == min(Risk))
+maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
+  filter(Risk == max(Risk))
+
+# label clusters
+cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3))
+cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
+  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
+cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
+  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
+
+# training clusters
+cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
+  round(3)
+cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
+  round(3)
+
+```
+
+After each site was mapped by four labellers, consensus labels were generated.
+The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality.
+The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
+Risk was highest in the northern cluster (AOIs 1-6; Figure S5), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figure S6), and lowest in the southwestern cluster (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
+Label risk in the central/eastern cluster (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
+Greater labeller experience also appeared to reduce risk, as the mean risk for the second mapping of Cluster 2's 500 initial training sites was `r cl2brisk`, compared to `r cl2arisk` for the first mapping.
+
+**[To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team.
+Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an *X%* agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average *X% smaller/larger* than those mapped on the ground.]**
 

### 2020-08-14 inst/manuscript/si.Rmd
Label risk assessment completed and integrated into SI and main text.

@@
 spacing: doublespacing
-footerdate: yes
 # abstract: The abstract can go either here or below
@@
 
-After creating the consensus label, the degree of confidence in the resulting label value can be expressed in terms of Bayesian risk. Risk is calculated as:
+After creating the consensus label, the degree of confidence in the resulting label value is measured by Bayes Risk:
 
@@
 A second approach is to calculate the proportion of pixels having risk values that exceed a certain threshold.
-This second approach is currently implemented in our system, using a threshold value of 0.4.
-
-Where *C* is the consensus probability that a given pixel *j* is a field ($\mathrm{P(\theta = field|D)}$), and *L* is the consensus label (i.e.
-non-field if *C* < 0.5, field if *C* > 0.5) for that pixel.
-*R* is thus the average risk across all pixels at a given site, and the slope of the risk varies depending on whether *L* is a field or not a field (Figure \ref{fig:riskcurve}).
-The closer to 0 the lower the risk that the *L* is mislabelled, while values approaching 1 indicate increasing risk of mislabelling.
 
@@
 
+The average Bayes Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}).
+The three initial clusters include the second mapping of Cluster 2.
+
+```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_label_risk_map.png')
+```
+
+```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayes Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
+The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_label_risk_hists.png')
+```
+
 ### Validation

### 2020-08-14 data-raw/data_from_instances.R

@@ -392,12 +392,14 @@ label_summary <- list("scores" = quality_scores, "score_stats" = quality_sum,
 usethis::use_data(label_summary, overwrite = TRUE)
 
 #------------------------------------------------------------------------------#
-# Consensus labelling conflicts (needs updating when consensus conflicts are
-# fixed)
-
-# reading in from separate script
-
+# Consensus labelling conflicts.
+# Processed for all AOIs in consensus_risk.R
+load(system.file("extdata/", "consensus_risk.rda", package = "activemapper"))
 load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+mgrid <- data.table::fread(
+  system.file("extdata/ghana_grid.csv", package = "activemapper")
+)
 
 # select AOIs and unique names:
 aoinms <- c(
@@ -405,21 +407,68 @@ aoinms <- c(
   paste0("labeller", c("trainsouth", "trainsouthnl", "trainsw", 1:16))
 )
 
-nms <- unique(c(train_val_sites$initial$name, train_val_sites$active$name))
-# any(!nms %in% unique(instance_tbls$kml_data$name))
+# get usages
+usages <- instance_tbls$incoming_names %>%
+  filter(!(aoi == "labeller3" & run == 0)) %>%
+  filter(label == TRUE) %>%
+  filter(aoi %in% aoinms) %>%
+  dplyr::select(aoi, name, usage) %>%
+  left_join(., mgrid %>% dplyr::select(name, x, y), by = "name")
 
-# This selects out F sites from the mapped list
-label_consensus <- instance_tbls$kml_data %>% filter(aoi %in% aoinms) %>%
-  filter(name %in% nms) %>%
-  select(-one_of("kml_type", "mapped_count", "mappers_needed")) %>%
-  rename(Risk = consensus_conflict)
+# checks on duplicates
+# dupnms <- usages %>% group_by(name) %>% count() %>% filter(n > 1) %>% pull(name)
+# length(dupnms)  # 503 duplicates
+# dupnms2 <- usages %>% filter(name %in% dupnms) %>% arrange(name) %>%
+#   filter(!aoi %in% c("labellertrainsouth", "labellertrainsouthnl"))
+# nrow(dupnms2) # 7 of those in other AOIs besides 2 runs of labellertrainsouth
[…]

### 2020-08-19 inst/manuscript/manuscript.Rmd
Reorganize into new scripts several of the pre-processing scripts

@@
 ## Ghana's croplands
+Two separate maps of cropland were produced for each AOI, one derived from the cropland probability map, and the other from the segmented field boundaries.
+
 - Average field size

### 2020-08-19 R/consensus_map_functions.R

@@ -336,7 +336,7 @@ consensus_map_creation2 <- function(kml_data, hit_data, assignment_data,
 #' @param results Output list from consensus_map_creation
 #' @param params Parameter file read in from common/
 #' @param s3bucket Name of s3bucket, read from db
-#' @param dupdate Update the table in database with risk results. Default = TRUE
+#' @param dbupdate Update the database with risk results or not. Default = TRUE
 #' @param output_riskmap Save output risk map?
 #' @param coninfo Database connection information from mapper_connect()
 #' @details This function will not save a consensus map to s3 if s3bucket is

### 2020-08-19 data-raw/data_from_instances.R

@@ -208,76 +208,6 @@ instance_tbls$users <- instance_tbls$users %>% distinct(aoi, id, email) %>%
 save(instance_tbls, file = here::here("inst/extdata/instance_tbls.rda"))
 
 #------------------------------------------------------------------------------#
-# Image quality check data from image_checker
-library(tidyr)
-tiles <- st_read(
-  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
-)
-
-aws_qry <- glue::glue("aws ec2 describe-instances --filters ",
-                      "'Name=tag:Name,Values=imagechecker' --output text ",
-                      "--query 'Reservations[].Instances[][PublicIpAddress]'")
-host <- system(aws_qry, intern = TRUE)
-# browseURL(glue::glue("https://{host}/phpPgAdmin"))
-
-# credentials
-params <- yaml::yaml.load_file(here::here("common/config.yaml"))
-dinfo <- params$mapper
-con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
-                      dbname = "Africa", user = dinfo$db_username,
-                      password = dinfo$db_password)
-
-# collect data
-image_quality <- tbl(con, "assignment_data") %>%
-  left_join(., tbl(con, "hit_data"), by = "hit_id") %>%
-  left_join(., tbl(con, "kml_data"), by = "name") %>%
-  left_join(., tbl(con, "master_grid"), by = "name") %>%
-  select(id, name, assignment_id, hit_id, x, y, worker_id, kml_type,
-         comment) %>%
-  collect()
-DBI::dbDisconnect(con)
-save(image_quality, file = "external/data/aois/dbases/image_quality_raw.rda")
-# load("external/data/aois/dbases/image_quality_raw.rda")
-
-# pick up tile information
-image_quality <- image_quality %>%
-  select(name, x, y, worker_id, kml_type, comment) %>%
-  tidyr::separate(comment, into = c(paste0("gs_", 1:4), paste0("os_", 1:4))) %>%
-  st_as_sf(coords = c("x", "y"), crs = 4326) %>%
-  st_join(., tiles) %>% as_tibble() %>% select(-geometry, -aoi2) %>%
-  select(name, tile, aoi1, worker_id, !!names(.))
-# image_quality %>% filter(worker_id == 99) %>% filter(kml_type == "F")
-# image_quality %>% filter(worker_id == 92) %>% filter(kml_type == "F")
-
-# image quality
[…]

### 2020-08-19 data-raw/data_from_segmenter.R

@@ -87,71 +87,6 @@ data.table::fwrite(
 )
 
 #------------------------------------------------------------------------------#
-# Analyze field segments
-
-# Coverage per cell
-logf <- here("external/logs/gridding_005.log")
-registerDoMC(cores = 7)
-fareas <- foreach(i = segment_files$order, .combine = stack) %dopar% {
-  # i <- segment_files[11]
-  aoi_id <- gsub("[[:alpha:]]|_|\\.", "", basename(i))
-
-  cat(glue::glue("Starting {aoi_id} at {Sys.time()}"), file = logf,
-      sep = "\n", append = TRUE)
-  print(aoi_id)
-
-  # read in fields
-  cat(glue::glue("...aoi {aoi_id} reading in fields and calculating tile area"),
-      file = logf, sep = "\n", append = TRUE)
-  flds <- read_sf(i)
-  # select out AOI grid and
-  aoi_tiles <- tilesr005_sf %>% filter(aoi1 == aoi_id) %>%
-    mutate(cid = 1:n(), area = as.numeric(st_area(.) / 10000))
-
-  # intersect fields
-  cat(glue::glue("...aoi {aoi_id} intersecting fields and calculating area"),
-      file = logf, sep = "\n", append = TRUE)
-  # fld_int <- st_intersection(aoi_tiles, flds)
-  fld_int <- st_join(aoi_tiles_xy, flds)
-  fld_int %>% mutate(class = )
-
-  drops <- fld_int %>%
-    st_is(c("POINT", "MULTIPOINT", "LINESTRING", "MULTILINESTRING")) %>%
-    which()
-  polys <- fld_int %>% st_is("POLYGON") %>% which()
-  fld_int <- fld_int %>% slice(-drops)
-  fld_int <- fld_int %>% mutate(farea = as.numeric(st_area(.) / 10000))
-  fld_areas <- fld_int %>% as_tibble() %>% group_by(cid) %>%
-    summarize(farea = sum(farea))
-
-  cat(glue::glue("...aoi {aoi_id} joining areas with tiles"), file = logf,
-      sep = "\n", append = TRUE)
-  aoi_tiles <- left_join(aoi_tiles, fld_areas) %>%
-    mutate(farea = ifelse(is.na(farea), 0, farea)) %>%
-    mutate(fprop = round(farea / area, 4))
-
-  # crop down to just
-  cat(glue::glue("...aoi {aoi_id} rasterizing"), file = logf, sep = "\n",
-      append = TRUE)
[…]

### 2020-08-19 data-raw/top_label_data.R

@@ -1,0 +1,95 @@
+# top label data: maps and statistics from the most accurate labeller to map
+# validation sites
+
+#------------------------------------------------------------------------------#
+# Extract maps for validation sites of most accurate labeller for each AOI
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+# main grid
+mgrid <- data.table::fread(
+  system.file("extdata/ghana_grid.csv", package = "activemapper")
+)
+
+aoinms <- paste0("labeller", 1:16)
+
+# calculate worker scores
+worker_scores <- instance_tbls$assignment_data %>%
+  rename(id = worker_id) %>%
+  filter(!is.na(score)) %>%
+  filter(aoi %in% aoinms) %>%
+  group_by(aoi, id) %>%
+  summarize(score = mean(score), n = n()) %>%
+  left_join(., instance_tbls$users) %>%
+  rename(wid = id)
+# worker_scores %>% filter(n < 10)
+
+# identify validation sites
+validation_sites <- instance_tbls$incoming_names %>%
+  filter(aoi %in% aoinms) %>%
+  filter(usage == "validate") %>%
+  filter(!(aoi == "labeller3" & run == 0)) %>%
+  select(aoi, name)
+
+# pull out assignments for those sites
+validation_assignments <- instance_tbls$hit_data %>%
+  left_join(validation_sites, .) %>%
+  select(aoi, name, hit_id, delete_time) %>%
+  left_join(., instance_tbls$assignment_data) %>%
+  select(aoi, name, hit_id, assignment_id, worker_id, status, delete_time)
+
+# which assignment was done by most accurate worker to map that assignment
+top_validation_assignments <- validation_assignments %>%
+  rename(wid = worker_id) %>%
+  filter(status %in% c("Approved")) %>% # only approved assignments
+  group_by(aoi, hit_id) %>% # count()
+  left_join(., worker_scores) %>% #select(aoi, name, hit_id, id, score)
+  filter(score == max(score)) %>%
+  select(aoi, name, hit_id, assignment_id, wid, score) %>%
+  ungroup()
+
+# get the associated maps
+top_validation_maps <- left_join(
+  top_validation_assignments, instance_tbls$user_maps,
+  by = c("aoi", "assignment_id")
+) %>%
+  rename(name = name.x, field = name.y) %>%
[…]

### 2020-10-09 inst/manuscript/figures.Rmd
Drafted cropland maps panel 1 figure; divided up and renamed data_from_segmenter into different scripts for doing area stats comparisons between validation polygons and segments at grid, tile, and aoi scales

@@
 ### Figure 7: Categorical map accuracy
+
+### Figure 8: Cropland maps
+```{r, eval=FALSE}
+library(stars)
+library(ggplot2)
+library(aws.s3)
+library(RStoolbox)
+library(cowplot)
+
+segment_files <- dir(here("external/data/results/segments/merged"),
+                     pattern = "boundarymerge", full.names = TRUE)
+iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
+segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
+segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))
+
+fdensity <- stars::read_stars(
+  system.file("extdata", "field_density005.tif", package = "activemapper")
+)
+
+# OS images
+osimgs <- get_bucket_df(
+  bucket = "activemapper", prefix = "planet/composite_sr_buf_fix/OS/", max = Inf
+) %>% as_tibble %>% select(Key)
+
+# example tiles
+# aoi 1: tile 486794
+# aoi 5: tile 539730
+# aoi 11: 
+# aoi 15:
+
+flds_1 <- read_sf(segment_files$file[1])
+flds_5 <- read_sf(segment_files$file[5])
+flds_11 <- read_sf(segment_files$file[11])
+flds_15 <- read_sf(segment_files$file[15])
+
+tile_ss <- tiles %>% filter(tile %in% c(486794, 539730, 591243, 616490))
+# ggplot(aois %>% filter(aois == 1)) + geom_sf() + 
+#   geom_sf(data = tile_ss)
+flds_tile1 <- flds_1 %>% filter(st_intersects(., tile_ss[1, ], sparse = FALSE))
+flds_tile2 <- flds_5 %>% filter(st_intersects(., tile_ss[2, ], sparse = FALSE))
+flds_tile3 <- flds_11 %>% filter(st_intersects(., tile_ss[3, ], sparse = FALSE))
+flds_tile4 <- flds_15 %>% filter(st_intersects(., tile_ss[4, ], sparse = FALSE))
+
+impath <- osimgs %>% 
+  filter(grepl(paste0(tile_ss$tile, collapse = "|"), Key)) %>% 
+  mutate(pth = paste0("/vsis3/activemapper/", Key)) %>% pull(pth)
+
+# img <- read_stars(impath)
+# mxval <- st_apply(img, 3, max) %>% pull(max)
+# imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
+
+fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
+isets <- lapply(1:4, function(x) {
+  imgr <- brick(impath[x])[[2:4]]
+  xylims <- st_bbox(imgr)
+  iset <- ggplot() + 
+    ggRGB(imgr, maxpixels = 50000, stretch = "lin", ggLayer = TRUE) +
+    geom_sf(data = fld_tiles[[x]], fill = "transparent", color = "yellow", 
+            size = 0.2) + 
+    geom_sf(data = tile_ss[x, ], color = "red", fill = "transparent") + 
+    scale_x_continuous(expand = c(0, 0)) + 
+    scale_y_continuous(expand = c(0, 0)) + 
+    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
+    theme_void()
+})
+isets[[2]]
+
+# Main map
+xylims <- st_coordinates(ghana)
+# twk <- diff(range(xylims[, 1])) * 0.05
+maprange <-  c(range(xylims[, 1]) + c(-1.25, 1.25), 
+               range(xylims[, 2]))
+
+inset_cols <- c("red", "orange", "pink", "brown")
+p <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = fdensity * 100) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  geom_sf(data = tile_ss, fill = "red", col = "red") +
+  scale_fill_viridis_c(name = "% cover", na.value = "transparent") +
+  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
+  theme_void() + 
+  theme(legend.position = "bottom", 
+        legend.text = element_text(colour = "white"), 
+        legend.title = element_text(colour = "white"))
+
+inset_map <- ggdraw() +
+  draw_plot(p) +
+  draw_plot(isets[[1]], x = 0.07, y = 0.77, width = 0.2, height = 0.2) + 
+  draw_plot(isets[[2]], x = 0.72, y = 0.6, width = 0.2, height = 0.2) + 
+  draw_plot(isets[[3]], x = 0.04, y = 0.37, width = 0.2, height = 0.2) + 
+  draw_plot(isets[[4]], x = 0.76, y = 0.25, width = 0.2, height = 0.2)
+
+ggsave(filename = here("presentations/figures/cropcover_with_insets.png"), 
+       plot = inset_map, bg = "transparent", width = diff(maprange[1:2]), 
+       height = diff(maprange[3:4]), units = "in", dpi = 300)
+
+  # geom_stars(data = imgngb) + scale_fill_identity()
+  # geom_sf(data = flds_tile)
+
+plot(flds_tile[1])
+
+fboundary <- read_sf(segment_files$file[1])
+
+xylims <- st_coordinates(ghana)
+maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
+
+pal <- RColorBrewer::brewer.pal(n = 9, "Spectral")[c(1, 2, 3, 4, 9)]
+
+# Panels, organized top to bottom
+# Panel 1: Cropland probabilities, with 3 m probability side maps (tiles)
+p <- ggplot(africa) + 
+  geom_sf(fill = "gray97", color = "grey", size = 0.25) +  
+  geom_stars(data = fdensity * 100) + 
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") + 
+  scale_fill_viridis_c(name = "% cover", na.value = "transparent") + 
+  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
+  theme_void()
+
+# Panel 2: Cropland cover, with polygon side maps over composites (tiles)
+p <- ggplot(africa) + 
+  geom_sf(fill = "gray97", color = "grey", size = 0.25) +  
+  geom_stars(data = fdensity * 100) + 
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") + 
+  geom_sf(data = tile_ss, fill = "red", col = "red") + 
+  scale_fill_viridis_c(name = "% cover", na.value = "transparent") + 
+  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
+  theme_void()
+p
+
+# Panel 3: Average field size (kriging adjusted?),
+
+# Panel 4: Field number (kriging adjusted)
+
+```
 

### 2020-10-09 data-raw/fboundary_stats_tile_aoi.R

@@ -1,0 +1,160 @@
+# Extract and summarize digitized and segmented field boundary statistics at
+# tile and AOI scales
+library(dplyr)
+library(sf)
+library(ggplot2)
+library(activemapper)
+library(doMC)
+library(here)
+
+data("top_label_data")
+
+tiles <- st_read(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+
+load(here("inst/extdata/segment_validation_data.rda"))
+
+# Read in segments, drop old AOI3
+segment_files <- dir(here("external/data/results/segments/merged"),
+                     pattern = "boundarymerge", full.names = TRUE)
+iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
+segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
+segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file))
+
+# get grids from validation data and intersect with tiles
+segment_val_grids <- lapply(segment_validation_data, function(x) x$grids) %>%
+  do.call(rbind, .)
+segment_val_tiles <- tiles %>% dplyr::select(tile) %>%
+  st_join(segment_val_grids, ., left = FALSE, largest = TRUE)
+
+# get tile ids for validation grids
+val_tiles <- top_label_data$stats %>%
+  st_as_sf(coords = c("x", "y"), crs = 4326) %>%
+  st_join(., tiles %>% select(tile)) %>%
+  as_tibble() %>%
+  select(aoi, tile, name) %>%
+  mutate(aoi = gsub("labeller", "", aoi)) %>%
+  arrange(aoi, tile)
+
+# use validation tiles to get segment and validation field stats
+logf <- here("external/logs/tile_aoi_stats.log")
+aois <- unique(tiles$aoi1)
+
+registerDoMC(cores = 7)
+val_stats <- foreach(x = 1:length(aois)) %dopar% { # x <- 8 #7 #3 #7 #5
+# tile_aoi_fld_stats <- lapply(1:length(aois), function(x) { # x <- 7
+
+  cat(glue::glue("Starting labeller{x} at {Sys.time()}"), file = logf,
+      sep = "\n", append = TRUE)
+
+  # read in segments for aoi
+  aoi_segments <- segment_files %>% filter(order == x) %>% pull(file) %>%
+    read_sf() %>% filter(!duplicated(.))
+  aoi_segments <- aoi_segments %>%
[…]

### 2020-10-09 data-raw/fboundary_stats_valsites.R

@@ -1,93 +1,6 @@
-## Processing field segment for:
-## 1. Stratifying validation sample
-## 2. Analyzing field statistics
+# Extract and summarize digitized and segmented field boundary statistics over
+# validation sites
 
-library(raster)
-library(sf)
-library(dplyr)
-library(here)
-
-# Datasets
-# grid, tiles, etc
-mgrid <- raster(
-  system.file("extdata/ghana_grid.tif", package = "activemapper")
-)
-tiles <- st_read(
-  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
-)
-
-# Read in segments
-segment_files <- dir(here("external/data/results/segments/merged"),
-                     pattern = "boundarymerge", full.names = TRUE)
-iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
-segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
-
-r <- raster(extent(tiles), res = 0.05)
-crs(r) <- crs(tiles)
-
-# for field areas
-tilesr <- fasterize::fasterize(tiles, r, field = "aoi1")
-tilesr005 <- disaggregate(tilesr, fact = 10)
-tilesr005_sf <- stars::st_as_stars(tilesr005) %>% st_as_sf() %>%
-  rename(aoi1 = layer)
-
-#------------------------------------------------------------------------------#
-# Point based stratification for map reference sample selection
-
-##!! Replace with csv grid
-# Create point grid for stratifying model results
-# mask main grid and convert to points
-mgrid <- mask(mgrid, tiles)
-mgrid_xy <- as.data.frame(mgrid, xy = TRUE) %>%
-  st_as_sf(coords = c("x", "y"), crs = 4326) %>% rename(id = ghana_grid) %>%
-  tidyr::drop_na(id)
-
-# join with tiles
-# system.time(extract(tilesr, master_grid_xy))  # 11.81
-system.time(mgrid_xytiles <- st_join(mgrid_xy, tilesr005_sf))  # 26
-
-#------------------------------------------------------------------------------#
-# Intersect points with grids
-logf <- here("external/logs/field_points_005.log")
-registerDoMC(cores = 7)
-fpoints <- foreach(i = segment_files$file) %dopar% {
-  # i <- segment_files$file[1]
[…]

### 2020-12-21 inst/manuscript/figures.Rmd
Revised segmentation stats--decided to just stick with original approach (compare against validation grids, not multiple scales)

@@
 
-### Figure S9: Segmentation statistics
+### Figure S9: Segmentation statistics (grid scale)
+
+[Check whether the means need to be weighted or not]
+[Check whether need to redo median]
 ```{r segstats, eval=FALSE}
-data("top_label_data")
-data("field_validation_stats")
+# data("top_label_data")
+# data("field_validation_stats")
+data("segment_quality_stats")
+# data
+datnms <- c("top_label_data", "field_validation_stats",
+            "tile_aoi_validation_stats")
+for(i in datnms) {
+  load(system.file(glue::glue("extdata/{i}.rda"), package = "activemapper"))
+}
 
+# extract distributions
 human <- top_label_data$stats %>%
@@
 
-# Summarize AOI-wise averages (field size)
-humachine_sum <- humachine %>% group_by(aoi, type) %>% 
-  summarize(Median = median(Mean), Mean = mean(Mean)) %>% 
-  ungroup() %>% 
-  tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
-# field number
-humachine_sum_n <- humachine %>% group_by(aoi, type) %>% 
-  summarize(Median = median(n), Mean = mean(n)) %>% 
-  ungroup() %>% 
-  tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
+squal <- segment_quality_stats %>% 
+  filter(scale == "grid" & aoi != "All") %>% 
+  mutate(aoi = as.numeric(aoi))
+# 
+# # Summarize AOI-wise averages (field size)
+# humachine_sum <- humachine %>% group_by(aoi, type) %>% 
+#   summarize(Median = median(Mean), Mean = mean(Mean)) %>% 
+#   ungroup() %>% 
+#   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
+# # field number
+# humachine_sum_n <- humachine %>% group_by(aoi, type) %>% 
+#   summarize(Median = median(n), Mean = mean(n)) %>% 
+#   ungroup() %>% 
+#   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat")
 
@@
                linetype = "blank") +  
-  geom_vline(data = humachine_sum, 
+  geom_vline(data = squal %>% filter(measure == "area"), 
              aes(xintercept = value, color = type, linetype = Stat), 
@@
   theme_fun()
-ggsave(pth("figures/si_validation_stats_fha.png"), 
+ggsave(pth("figures/si_validation_stats_fha2.png"), 
        width = 7, height = 7, units = "in", dpi = 300)
@@
                linetype = "blank") +  
-  geom_vline(data = humachine_sum_n, 
+  geom_vline(data = squal %>% filter(measure == "n"), 
              aes(xintercept = value, color = type, linetype = Stat), 
@@
   theme_fun()
-ggsave(pth("figures/si_validation_stats_fn.png"), 
+ggsave(pth("figures/si_validation_stats_fn2.png"), 
        width = 7, height = 7, units = "in", dpi = 300)
@@
 ```
+
+<!-- ### Figure S10: Segmentation statistics, tile scale -->
+<!-- ```{r, eval=FALSE} -->
+<!-- data("tile_aoi_validation_stats") -->
+
+<!-- humachine <- tile_aoi_validation_stats$tile %>%  -->
+<!--   mutate(type = ifelse(grepl("segments", type), "Segmentation", type), -->
+<!--          type = ifelse(grepl("validation", type), "Validation", type)) %>%  -->
+<!--   select(type, aoi, tile, area_mu, area_med, n_mu) %>%  -->
+<!--   rename(Mean = area_mu, Median = area_med) -->
+
+<!-- humachine_sum <- humachine %>% group_by(aoi, type) %>%  -->
+<!--   summarize(Median = median(Mean), Mean = mean(Mean)) %>%  -->
+<!--   ungroup() %>%  -->
+<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") -->
+
+<!-- humachine_sum_n <- humachine %>% group_by(aoi, type) %>%  -->
+<!--   summarize(Median = median(n_mu), Mean = mean(n_mu)) %>%  -->
+<!--   ungroup() %>%  -->
+<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") -->
+
+<!-- # Field size -->
+<!-- brks <- seq(0, 20, 5) -->
+<!-- p <- ggplot(humachine) +  -->
+<!--   geom_density(aes(x = Mean, y = ..count.., fill = type), col = "transparent",  -->
+<!--                linetype = "blank") +   -->
+<!--   geom_vline(data = humachine_sum,  -->
+<!--              aes(xintercept = value, color = type, linetype = Stat),  -->
+<!--              size = 0.5) + -->
+<!--   scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 20)) + -->
+<!--   scale_y_continuous(expand = c(0, 0)) + -->
+<!--   scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) +  -->
+<!--   scale_fill_manual(name = "", values = fcols) +   -->
+<!--   # scale_linetype_manual(name = "", values = c("dashed", "solid")) +   -->
+<!--   scale_linetype("") +   -->
+<!--   facet_wrap(vars(aoi)) +  -->
+<!--   xlab("Field size (ha)") + ylab("Count") +  -->
+<!--   theme_fun() -->
+<!-- ggsave(pth("figures/si_tile_validation_stats_fha.png"),  -->
+<!--        width = 7, height = 7, units = "in", dpi = 300) -->
+
+<!-- # field number -->
+<!-- brks <- seq(0, 20, 5) -->
+<!-- p2 <- ggplot(humachine) +  -->
+<!--   geom_density(aes(x = n_mu, y = ..count.., fill = type), col = "transparent",  -->
+<!--                linetype = "blank") +   -->
+<!--   geom_vline(data = humachine_sum_n,  -->
+<!--              aes(xintercept = value, color = type, linetype = Stat),  -->
+<!--              size = 0.5) + -->
+<!--   scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 20)) + -->
+<!--   scale_y_continuous(expand = c(0, 0)) + #, limits = c(0, 40)) + -->
+<!--   scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) +  -->
+<!--   scale_fill_manual(name = "", values = fcols) +   -->
+<!--   # scale_linetype_manual(name = "", values = c("dashed", "solid")) +   -->
+<!--   scale_linetype("") +   -->
+<!--   facet_wrap(vars(aoi)) +  -->
+<!--   xlab("Number of fields") + ylab("Count") +  -->
+<!--   theme_fun() -->
+<!-- ggsave(pth("figures/si_tile_validation_stats_fn.png"),  -->
+<!--        width = 7, height = 7, units = "in", dpi = 300) -->
+<!-- ``` -->
+
+<!-- ### Figure S11: Segmentation statistics: grid, tile, and AOI scales -->
+
+<!-- Read back in segments to get distributions -->
+<!-- ```{r, eval = FALSE} -->
+<!-- data("tile_aoi_validation_stats") -->
+
+<!-- # # Read in segments, drop old AOI3 -->
+<!-- # segment_files <- dir(here("external/data/results/segments/merged"), -->
+<!-- #                      pattern = "boundarymerge", full.names = TRUE) -->
+<!-- # iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files))) -->
+<!-- # segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid) -->
+<!-- # segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file)) -->
+<!-- #  -->
+<!-- # # area and quiet func -->
+<!-- # ha <- function(x) as.numeric(units::set_units(st_area(x), "ha")) -->
+<!-- # quiet <- function(x) suppressWarnings(suppressMessages(x)) -->
+
+<!-- humachine_sum <- tile_aoi_validation_stats$aoi %>%  -->
+<!--   select(type, aoi, area_mu, area_med) %>%  -->
+<!--   rename(Mean = area_mu, Median = area_med) %>%  -->
+<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") %>%  -->
+<!--   tidyr::pivot_wider(names_from = type, values_from = value) -->
+
+<!-- aoi_wide <- tile_aoi_validation_stats$aoi %>% group_by(type) %>%  -->
+<!--   summarize(Mean = weighted.mean(area_mu, w = n),  -->
+<!--             Median = weighted.mean(area_med, w = n)) %>%  -->
+<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") %>%  -->
+<!--   tidyr::pivot_wider(names_from = type, values_from = value) %>%  -->
+<!--   mutate(aoi = "All") %>% select(aoi, !!names(.)) -->
+
+<!-- p <- ggplot(humachine_sum) + -->
+<!--   geom_point(aes(x = segments, y = validation)) +  -->
+<!--   geom_point(data = aoi_wide, aes(x = segments, y = validation), color = "red") +  -->
+<!--   scale_x_continuous(limits = c(0, 8)) +  -->
+<!--   scale_y_continuous(limits = c(0, 8)) +  -->
+<!--   ggrepel::geom_text_repel( -->
+<!--     aes(label = aoi, x = segments, y = validation), col = "grey50",  -->
+<!--     segment.color = "grey60", point.padding = 0.1, min.segment.length = 0.2, -->
+<!--     segment.size = 0.2, size = 3    -->
+<!--   ) +  -->
+<!--   xlab("Segmentation field size (ha)") + ylab("Validation field size (ha)") +  -->
+<!--   geom_abline(slope = 1) + facet_wrap(vars(Stat)) +  -->
+<!--   theme_linedraw() +  -->
+<!--   theme(strip.background = element_rect(fill = "white"),  -->
+<!--         strip.text = element_text(colour = "black")) -->
+
+<!-- ``` -->

### 2020-12-21 data-raw/label_v_segments_summarize.R

@@ -114,7 +114,7 @@ humachine_tile_sum <- humachine_tile %>% group_by(aoi, type) %>%
   select(scale, aoi, type, measure, !!names(.))
 
 # combine with tile-wise averages
-humachine_tile_sum <- sum_func(humachine_grid_sum, "tile", "area")
+humachine_tile_sum <- sum_func(humachine_tile_sum, "tile", "area")
 
 # field count
 humachine_tile_sum_n <- humachine_tile %>% group_by(aoi, type) %>%
@@ -125,7 +125,7 @@ humachine_tile_sum_n <- humachine_tile %>% group_by(aoi, type) %>%
   select(scale, aoi, type, measure, !!names(.))
 
 # combine with tile-wise averages
-humachine_tile_sum_n <- sum_func(humachine_grid_sum, "tile", "n")
+humachine_tile_sum_n <- sum_func(humachine_tile_sum_n, "tile", "n")
 
 # # median and means across AOIs - evaluated against validation tiles
 # # area

### 2020-12-24 inst/manuscript/figures.Rmd
Updates, through placement of risk-metric correlation in SI.

@@
 # data
-datnms <- c("top_label_data", "field_validation_stats",
-            "tile_aoi_validation_stats")
+datnms <- c("top_label_data", "field_validation_stats")#,
+            # "tile_aoi_validation_stats")
 for(i in datnms) {

### 2020-12-24 inst/manuscript/manuscript.Rmd
Updates, through placement of risk-metric correlation in SI.

@@
 
-- name: Sitian Xiong
-  affilnum: '1'
-
-- name: Tammy Woodard
+- name: Qi Zhang
   affilnum: '1'
@@
 
-```{r, echo = FALSE, out.width="80%", fig.cap="The overall, user's, and producer's accuracies (indicated by fill color) for each of the 16 mapping zones (zones numbers are shown on maps) [This a hypothetical placeholder figure].", fig.align='center'}
-knitr::include_graphics('figures/si_mapping_blocks.png')
-```
+<!-- ```{r, echo = FALSE, out.width="80%", fig.cap="The overall, user's, and producer's accuracies (indicated by fill color) for each of the 16 mapping zones (zones numbers are shown on maps) [This a hypothetical placeholder figure].", fig.align='center'} -->
+<!-- knitr::include_graphics('figures/si_mapping_blocks.png') -->
+<!-- ``` -->
 
@@
 # validation statistics
-data("top_label_data")
-data("field_validation_stats")
-
+data("segment_quality_stats")
+
+datnms <- c("top_label_data", "field_validation_stats")#,
+            # "tile_aoi_validation_stats")
+for(i in datnms) {
+  load(system.file(glue::glue("extdata/{i}.rda"), package = "activemapper"))
+}
+
+# extract distributions
 human <- top_label_data$stats %>%
-  mutate(mu_prop = NA, type = "val") %>%
+  mutate(mu_prop = NA, type = "Validation") %>%
   select(-contains("id"))
 machine <- field_validation_stats %>%
-  mutate(type = "seg")
-humachine <- bind_rows(human, machine) %>% 
-  mutate(aoi = as.numeric(gsub("labeller", "", aoi))) %>% 
-  arrange(aoi)
-
-huma_aois <- humachine %>% group_by(aoi, type) %>% 
-  summarize(med = median(Mean), mu = mean(Mean)) %>% 
-  ungroup()
-lrgst <- huma_aois %>% select(-med) %>% filter(aoi %in% c(2, 3, 7, 15)) %>% 
-  tidyr::pivot_wider(names_from = type, values_from = "mu") %>% 
-  filter(seg %in% range(seg))
-lrgstlow <- lrgst %>% filter(seg == min(seg))
-lrgsthigh <- lrgst %>% filter(seg == max(seg))
-
-# overall stats
-huma_sum <- humachine %>% group_by(type) %>% 
-  summarize(med = median(Mean), sd = sd(Mean), min = min(Mean), 
-            max = max(Mean), mu = mean(Mean)) %>% 
-  select(type, mu, med, sd, min, max)
-segha <- huma_sum %>% filter(type == "seg") %>% pull(mu) %>% round(2)
-valha <- huma_sum %>% filter(type == "val") %>% pull(mu) %>% round(2)
-seghamed <- huma_sum %>% filter(type == "seg") %>% pull(med) %>% round(2)
-valhamed <- huma_sum %>% filter(type == "val") %>% pull(med) %>% round(2)
-seghasd <- huma_sum %>% filter(type == "seg") %>% pull(sd) %>% round(2)
-valhasd <- huma_sum %>% filter(type == "val") %>% pull(sd) %>% round(2)
-
-huma_sumn <- humachine %>% group_by(type) %>% 
-  summarize(med = median(n), sd = sd(n), min = min(n), max = max(n), 
-            mu = mean(n)) %>% 
-  select(type, mu, med, sd, min, max)
-segn <- huma_sumn %>% filter(type == "seg") %>% pull(mu) %>% round(2)
-valn <- huma_sumn %>% filter(type == "val") %>% pull(mu) %>% round(2)
-segnmed <- huma_sumn %>% filter(type == "seg") %>% pull(med) %>% round(2)
-valnmed <- huma_sumn %>% filter(type == "val") %>% pull(med) %>% round(2)
-segnsd <- huma_sumn %>% filter(type == "seg") %>% pull(sd) %>% round(2)
-valnsd <- huma_sumn %>% filter(type == "val") %>% pull(sd) %>% round(2)
-```
-The comparisons of digitized versus segmented field boundaries showed that the mean field size across all 1600 validation sites averaged `r round(segha, 2)` ha (StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
-This discrepancy was primarily caused by segmentation results in four AOIs (2, 3, 7, and 15; Figure S5), where segments averaged between `r round(lrgstlow$seg, 2)` and `r round(lrgsthigh$seg, 2)` ha, compared to `r round(lrgstlow$val, 2)` - `r round(lrgsthigh$val, 2)` for the corresponding hand-digitized polygons.
-In contrast, the median segmented field area was closer to that of digitized polygons (`r seghamed` for segments versus `r valhamed` for labeller polygons).
+  mutate(type = "Segmentation")
+
+# calculate SDs
+humac_sds <- bind_rows(
+  human %>% summarize(value = sd(Mean)) %>% 
+    mutate(type = "v", stat = "sd_ha_mu"), 
+  human %>% summarize(value = sd(Median)) %>% 
+    mutate(type = "v", stat = "sd_ha_med"), 
+  human %>% summarize(value = sd(n)) %>% 
+    mutate(type = "v", stat = "sd_n_mu"), 
+  machine %>% summarize(value = sd(Mean)) %>% 
+    mutate(type = "s", stat = "sd_ha_mu"), 
+  machine %>% summarize(value = sd(Median)) %>% 
+    mutate(type = "s", stat = "sd_ha_med"), 
+  machine %>% summarize(value = sd(n)) %>% 
+    mutate(type = "s", stat = "sd_n_mu") 
+)
+
+# extract distributions
+
+# human <- top_label_data$stats %>%
+#   mutate(mu_prop = NA, type = "val") %>%
+#   select(-contains("id"))
+# machine <- field_validation_stats %>%
+#   mutate(type = "seg")
+# humachine <- bind_rows(human, machine) %>% 
+#   mutate(aoi = as.numeric(gsub("labeller", "", aoi))) %>% 
+#   arrange(aoi)
+
+# huma_aois <- humachine %>% group_by(aoi, type) %>% 
+#   summarize(med = median(Mean), mu = mean(Mean)) %>% 
+#   ungroup()
+# lrgst <- huma_aois %>% select(-med) %>% filter(aoi %in% c(2, 3, 7, 15)) %>% 
+#   tidyr::pivot_wider(names_from = type, values_from = "mu") %>% 
+#   filter(seg %in% range(seg))
+seg_stats_grid <- segment_quality_stats %>% filter(scale == "grid")
+lrgst <- seg_stats_grid %>% 
+  filter(aoi %in% c(2, 3, 7, 15)) %>% 
+  tidyr::pivot_wider(names_from = type, values_from = value) %>%
+  filter(Stat == "Mean" & measure == "area")
+
+lrgst_seg <- lrgst %>% summarize(rng = range(Segmentation))
+lrgst_val <- lrgst %>% summarize(rng = range(Validation))
+
+# # overall stats
+# huma_sum <- humachine %>% group_by(type) %>% 
+#   summarize(med = median(Mean), sd = sd(Mean), min = min(Mean), 
+#             max = max(Mean), mu = mean(Mean)) %>% 
+#   select(type, mu, med, sd, min, max)
+
+segha <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "area") %>% 
+  filter(type == "Segmentation" & Stat == "Mean") %>% pull(value) %>% round(2)
+# segha <- huma_sum %>% filter(type == "seg") %>% pull(mu) %>% round(2)
+valha <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "area") %>% 
+  filter(type == "Validation" & Stat == "Mean") %>% pull(value) %>% round(2)
+# valha <- huma_sum %>% filter(type == "val") %>% pull(mu) %>% round(2)
+
+# seghamed <- huma_sum %>% filter(type == "seg") %>% pull(med) %>% round(2)
+# valhamed <- huma_sum %>% filter(type == "val") %>% pull(med) %>% round(2)
+seghamed <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "area") %>% 
+  filter(type == "Segmentation" & Stat == "Median") %>% pull(value) %>% round(2)
+# segha <- huma_sum %>% filter(type == "seg") %>% pull(mu) %>% round(2)
+valhamed <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "area") %>% 
+  filter(type == "Validation" & Stat == "Median") %>% pull(value) %>% round(2)
+
+# seghasd <- huma_sum %>% filter(type == "seg") %>% pull(sd) %>% round(2)
+seghasd <- humac_sds %>% filter(type == "s" & stat == "sd_ha_mu") %>% 
+  pull(value) %>% round(2)
+# valhasd <- huma_sum %>% filter(type == "val") %>% pull(sd) %>% round(2)
+valhasd <- humac_sds %>% filter(type == "v" & stat == "sd_ha_mu") %>% 
+  pull(value) %>% round(2)
+
+segnsd <- humac_sds %>% filter(type == "s" & stat == "sd_n_mu") %>% 
+  pull(value) %>% round(2)
+valnsd <- humac_sds %>% filter(type == "v" & stat == "sd_n_mu") %>% 
+  pull(value) %>% round(2)
+
+# huma_sumn <- humachine %>% group_by(type) %>% 
+#   summarize(med = median(n), sd = sd(n), min = min(n), max = max(n), 
+#             mu = mean(n)) %>% 
+#   select(type, mu, med, sd, min, max)
+# segn <- huma_sumn %>% filter(type == "seg") %>% pull(mu) %>% round(2)
+# valn <- huma_sumn %>% filter(type == "val") %>% pull(mu) %>% round(2)
+# segnmed <- huma_sumn %>% filter(type == "seg") %>% pull(med) %>% round(2)
+# valnmed <- huma_sumn %>% filter(type == "val") %>% pull(med) %>% round(2)
+# segnsd <- huma_sumn %>% filter(type == "seg") %>% pull(sd) %>% round(2)
+# valnsd <- huma_sumn %>% filter(type == "val") %>% pull(sd) %>% round(2)
+
+segn <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "n") %>% 
+  filter(type == "Segmentation" & Stat == "Mean") %>% pull(value) %>% round(2)
+valn <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "n") %>% 
+  filter(type == "Validation" & Stat == "Mean") %>% pull(value) %>% round(2)
+
+segnmed <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "n") %>% 
+  filter(type == "Segmentation" & Stat == "Median") %>% pull(value) %>% round(2)
+valnmed <- seg_stats_grid %>% 
+  filter(aoi == "All" & measure == "n") %>% 
+  filter(type == "Validation" & Stat == "Median") %>% pull(value) %>% round(2)
+
+```
+The comparisons of digitized versus segmented field boundaries showed that the mean field size across all 1600 validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
+This discrepancy was primarily caused by segmentation results in four AOIs (2, 3, 7, and 15; Figure S5), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
 The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S6).
 
-### Label quality and map accuracy
-```{r, echo = FALSE, out.width="100%", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/figure6.png')
-# library(tidyverse)
-# library(cowplot)
-# set.seed(1)
-# acc_tb <- tibble(aoi = c(1:9, 12, 15), 
-#                  Consensus = runif(11, min = 0.75, max = 0.85),
-#                  Highest = runif(11, min = 0.72, max = 0.8),
-#                  Least = runif(11, min = 0.65, max = 0.75))
-# acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi)
-# ggplot(acc_tb) + 
-#   geom_boxplot(aes(x = type, y = score), fill = "grey") + 
-#   xlab("Labelling strategy") + ylab("Accuracy") + 
-#   theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3))
-```
+<!-- ### Label quality and map accuracy -->
+<!-- ```{r, echo = FALSE, out.width="100%", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.", fig.align='center', message=FALSE} -->
+<!-- knitr::include_graphics('figures/figure6.png') -->
+<!-- # library(tidyverse) -->
+<!-- # library(cowplot) -->
+<!-- # set.seed(1) -->
+<!-- # acc_tb <- tibble(aoi = c(1:9, 12, 15),  -->
+<!-- #                  Consensus = runif(11, min = 0.75, max = 0.85), -->
+<!-- #                  Highest = runif(11, min = 0.72, max = 0.8), -->
+<!-- #                  Least = runif(11, min = 0.65, max = 0.75)) -->
+<!-- # acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi) -->
+<!-- # ggplot(acc_tb) +  -->
+<!-- #   geom_boxplot(aes(x = type, y = score), fill = "grey") +  -->
+<!-- #   xlab("Labelling strategy") + ylab("Accuracy") +  -->
+<!-- #   theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3)) -->
+<!-- ``` -->
 
@@
 
-- Average field size
+Map figure:
+- Panel A: Cropland probability map (0.005 degree resolution)
+  - Panel A insets: Probability close-ups in AOIs 1, 8, 13, 15
+- Panel B: Cropland percentage map (0.005 degree)
+  - Panel B insets: Segment close-ups in AOIs 1, 8, 13, 15
+- Panel C: Median field size
+- Panel D: Field number
+- Median field size
 - Cropland percentage
-- Comparison against other products (croplands.org; Sentinel Landcover)
+
+Stats figure: 
+- Mean field size per AOI
+- Median field size per AOI
+- SD fields per AOI 
+- N field per AOI
 

### 2020-12-24 inst/manuscript/si.Rmd
Updates, through placement of risk-metric correlation in SI.

@@
 ## Image quality
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 data("image_quality")
@@
 
+The association between the average label risk per AOI and several model performance metrics, assessed using Spearman Rank Correlation, is shown in Table \ref{tab:labelrisk_v_metrics}.
+
+```{r labelrisk_v_metrics, echo=FALSE}
+data("label_risk")
+data("iteration_metrics")
+
+risk <- label_risk$stats %>% 
+  filter(usage == "Train") %>% 
+  filter(aoi %in% 1:16) %>% 
+  select(-usage)
+
+mets <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "FPR")
+metrics <- iteration_metrics %>% 
+  filter(Metric %in% mets) %>% 
+  mutate(AOI = as.factor(gsub("labeller", "", AOI))) %>% 
+  filter(AOI %in% 1:16) %>% 
+  # pivot_wider(names_from = Metric, values_from = Score) %>% 
+  group_by(AOI) %>% filter(Iteration == max(Iteration)) %>% 
+  select(-Iteration)
+
+risk_metrics <- left_join(risk, metrics, by = c("aoi" = "AOI"))
+
+cor_fun <- function(df) {
+  cor.test(df$Risk, df$Score, method = "spearman") %>% broom::tidy()
+}
+
+risk_corr <- lapply(unique(risk_metrics$Metric), function(x) {
+  dat <- risk_metrics %>% filter(Metric == x)
+  cor_fun(dat) %>% mutate(Metric = x)
+}) %>% do.call(rbind, .) %>% 
+  select(estimate, Metric) %>% 
+  rename(r = estimate)
+nms <- risk_corr$Metric
+risk_corr <- round(t(risk_corr[, 1]), 3)
+colnames(risk_corr) <- nms
+
+cap <- glue::glue("Spearman's Rank correlation between the average label risk per AOI and a variety of model performance metrics.")
+kable(risk_corr, format = "latex", booktabs = "T", caption = cap) %>% 
+  kable_styling(full_width = TRUE, latex_options = c("hold_position")) #%>% 
+  # column_spec(1, width = "8cm")
+```
+
 ## Map accuracy 
@@
 Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
-knitr::include_graphics('figures/si_validation_stats_fha.png')
+knitr::include_graphics('figures/si_validation_stats_fha2.png')
 ```
@@
 Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
-knitr::include_graphics('figures/si_validation_stats_fn.png')
+knitr::include_graphics('figures/si_validation_stats_fn2.png')
 ```

### 2021-01-04 inst/manuscript/manuscript.Rmd
Updates to manuscript, si, added various ignores. New data-raw script for combining label_risk with accuracy metrics.

@@
 
-A second measure of the impact of training data error can be obtained by comparing the average label risk per AOI against the model performance metrics....
+A second measure of the impact of training data error can be obtained by calculating the Spearman Rank Correlation coefficient between the average label risk per AOI versus various model performance metrics.
 

### 2021-01-04 inst/manuscript/si.Rmd
Updates to manuscript, si, added various ignores. New data-raw script for combining label_risk with accuracy metrics.

@@
 ```{r labelrisk_v_metrics, echo=FALSE}
-data("label_risk")
-data("iteration_metrics")
-
-risk <- label_risk$stats %>% 
-  filter(usage == "Train") %>% 
-  filter(aoi %in% 1:16) %>% 
-  select(-usage)
-
-mets <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "FPR")
-metrics <- iteration_metrics %>% 
-  filter(Metric %in% mets) %>% 
-  mutate(AOI = as.factor(gsub("labeller", "", AOI))) %>% 
-  filter(AOI %in% 1:16) %>% 
-  # pivot_wider(names_from = Metric, values_from = Score) %>% 
-  group_by(AOI) %>% filter(Iteration == max(Iteration)) %>% 
-  select(-Iteration)
-
-risk_metrics <- left_join(risk, metrics, by = c("aoi" = "AOI"))
 
@@
   # column_spec(1, width = "8cm")
+
+# tp <- 10
+# fp <- 2
+# tn <- 20
+# fn <- 3
+# tss <- (tp * tn - fp * fn) / ((tp + fn) * (fp + tn))
+# mcc <- (tp * tn - fp * fn) / sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
+
 ```

### 2021-01-04 inst/manuscript/manuscript.Rmd
Added text on label risk correlations

--- 2020-12-24 inst/manuscript/manuscript.Rmd
@@
 
-A second measure of the impact of training data error can be obtained by comparing the average label risk per AOI against the model performance metrics....
+```{r, echo=FALSE}
+data("label_risk_metrics")
+
+cor_fun <- function(df) {
+  cor.test(df$Risk, df$Score, method = "spearman") %>% broom::tidy()
+}
+
+lr_mets <- unique(label_risk_metrics$Metric)
+risk_corr <- lapply(lr_mets[lr_mets != "TSS"], function(x) {
+  dat <- label_risk_metrics %>% filter(Metric == x)
+  cor_fun(dat) %>% mutate(Metric = x)
+}) %>% do.call(rbind, .) %>% 
+  select(estimate, Metric) %>% 
+  rename(r = estimate)
+nms <- risk_corr$Metric
+risk_corr <- round(t(risk_corr[, 1]), 3)
+colnames(risk_corr) <- nms
+```
+
+A second measure of the impact of training data error can be obtained by calculating the between the average training label risk per AOI versus various model performance metrics (Table S3).
+Accuracy had a strong negative association (Spearman's Rank Correlation =  `r risk_corr[, "Accuracy"]` and AUC had a moderate negative association (r = `r risk_corr[, "AUC"]`), while F1 had a weaker positive association (r = `r risk_corr[, "F1"]`).
+This sign of the latter is counter-intuitive, but relates to the fact that precision was higher in areas of increasing label risk (`r risk_corr[, "Precision"]`), as did false positive error rates (`r risk_corr[, "FPR"]`).
+In contrast, recall had little correlation with label risk (`r risk_corr[, "Recall"]`), thus the higher precision drove F1 higher, even in areas with more false positives.
+**[work on this]**
 

### 2021-01-04 inst/manuscript/si.Rmd
Added text on label risk correlations

@@
 ```{r labelrisk_v_metrics, echo=FALSE}
+data("label_risk_metrics")
 
@@
 
-risk_corr <- lapply(unique(risk_metrics$Metric), function(x) {
-  dat <- risk_metrics %>% filter(Metric == x)
+lr_mets <- unique(label_risk_metrics$Metric)
+risk_corr <- lapply(lr_mets[lr_mets != "TSS"], function(x) {
+  dat <- label_risk_metrics %>% filter(Metric == x)
   cor_fun(dat) %>% mutate(Metric = x)

### 2021-01-28 data-raw/process_validation_sample.R
Added script for processing map_reference_sample, which is written and committed to inst_extdata

@@ -1,0 +1,131 @@
+# Collect data from probability images for accuracy assessment
+# Uses polygons collected from validator
+#
+
+## Map reference site selection
+library(aws.s3)
+library(raster)
+library(sf)
+library(dplyr)
+library(rmapaccuracy)
+
+# get validation samples
+load(system.file("extdata/mapref_tbls.rda", package = "activemapper"))
+# load(system.file("extdata/instance_tbls.rda", package = "activemapper"))
+
+ref_sample <- readr::read_csv(
+  system.file("extdata/map_reference_sample.csv", package = "activemapper")
+)
+
+tiles <- read_sf(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+gcs <- unname(unlist(st_crs(4326))[2])
+
+# create master_grid extent, get list of probability images, download fpool
+# (for checking)
+
+# collect polygons with worker_id
+val_sample <- left_join(mapref_tbls$user_maps, mapref_tbls$assignment_data) %>%
+  select(name, fieldnum, assignment_id, category, worker_id, comment) %>%
+  left_join(., mapref_tbls$categories) %>%
+  filter(name %in% ref_sample$name)  # filter out several not in ref_sample
+
+# val_nms <- unique(val_sample$name)
+# nms_not_in <- val_nms[!val_nms %in% ref_sample$name]
+# mapref_tbls$kml_data %>% filter(name %in% nms_not_in)
+# mapref_tbls$kml_data %>% filter(name %in% ref_sample$name) %>%
+#   group_by(kml_type) %>% count()
+
+# mapref_tbls$kml_data %>% filter(name %in% val_nms) %>%
+#   group_by(kml_type) %>% count()
+# mapref_tbls$kml_data$kml_type
+
+
+# val_sample %>% filter(name == "GH0639733") %>% select(worker_id) %>%
+#   plot()
+# val_sample %>% filter(category == "unsure1")
+# val_sample_1 <- val_sample %>% filter(worker_id == 134)
+
+# separate by validator
+val_sample1 <- val_sample %>% filter(worker_id == 90)
+val_sample2 <- val_sample %>% filter(worker_id == 134)
+overlap_samples <- unique(
+  val_sample1$name[val_sample1$name %in% val_sample2$name]
+)
+
+# remove overlap sites
[…]

### 2021-01-28 inst/manuscript/figures.Rmd
Update ignores, figures, validation process script

@@
 
-The preparation of datasets used here and which install with the `activemapper` package, is shown in `data-raw/data_[from|to]_instaces.R`. These are loaded in here.
+The preparation of datasets used here and which install with the `activemapper` package, is shown in `data-raw/data_[from|to]_instances.R`. These are loaded in here.
 
@@
 
+```{r, eval=FALSE}
+data("label_risk_metrics")
+
+label_risk_metrics %>% 
+  filter(Metric != "TSS") %>%  
+  ggplot() + geom_point(aes(Risk, Score)) + 
+  geom_smooth(aes(Risk, Score), method = "gam") + 
+  facet_wrap(~Metric)
+
+label_risk_metrics %>% select(aoi, Metric, Score) %>% 
+  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
+  ggplot() + geom_point(aes(Precision, Accuracy))
+
+label_risk_metrics %>% select(aoi, Metric, Score) %>% 
+  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
+  ggplot() + geom_point(aes(Precision, F1))
+
+label_risk_metrics %>% select(aoi, Metric, Score) %>% 
+  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
+  ggplot() + geom_point(aes(Recall, F1))
+
+label_risk_metrics %>% select(aoi, Metric, Score) %>% 
+  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
+  ggplot() + geom_point(aes(Precision, Recall))
+
+label_risk_metrics %>% select(aoi, Metric, Score) %>% 
+  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
+  ggplot() + geom_point(aes(Precision, FPR))
+label_risk_metrics %>% select(aoi, Metric, Score) %>% 
+  tidyr::pivot_wider(names_from = "Metric", values_from = "Score") %>% 
+  ggplot() + geom_point(aes(Recall, FPR))
+
+```
+
 ### Figure S7: Random versus active sampling

### 2021-01-28 data-raw/process_validation_sample.R

@@ -114,7 +114,7 @@ cbind(as_tibble(val_sample_1r_sf), tile = val_tiles$tile, val_tile_rcs) %>%
 
 
 # write out
-st_write(ref_labels, here::here("inst/extdata/reference_labels.geojson"))
+st_write(ref_labels, here::here("inst/extdata/map_reference_labels.geojson"))
 
 # # test
 # i <- 2

### 2021-01-28 inst/manuscript/manuscript.Rmd
Manuscript updates

@@
 Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms.
-Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly neural network-based "deep-learning" methods (cite).
+Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly deep-learning methods (cite).
 To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g.
@@
 
-We describe here a cropland mapping system that addresses these requirements on the way to creating an object-based map of individual fields.
+We describe here a cropland mapping system that addresses these requirements on the way to delineating field boundaries.
 The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
 Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016].
+Their daily revisit capacity also is important for being able to develop cloud-free, sub-annual composites in cloudy regions.
 Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g.
@@
 For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or replanted tree crops.
-We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are typically not visible in cocoa crops.
+We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are often not visible within cocoa crops.
 
@@
     - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours. 
+    - Feng et al (2020) importance of temporal information (backs up Debats et al).
+      - temporal information makes labelling harder though
+
+Improvements:    
+- Replace Random Forest with deep learning model. Can still retain the active architecture, but instead use new sites to improve a pre-trained model
+- Map along agroecozone boundaries, rather than random blocks
 

### 2021-01-28 data-raw/refine_reference_sample.R
Created tile_key.rda and commit image_quality.rda

@@ -106,7 +106,6 @@ val_tiles <- tiles %>% slice(ints)
 xys <- val_tiles %>% st_centroid() %>% st_coordinates()
 val_tile_rcs <- rowcol_from_xy(xys[, 1], xys[, 2], res = 0.05)
 
-
 # Combine with labels
 cbind(as_tibble(val_sample_1r_sf), tile = val_tiles$tile, val_tile_rcs) %>%
   select(name, assignment_id, worker_id, category, categ_description, tile,

### 2021-01-28 data-raw/tile_key.R

@@ -1,0 +1,27 @@
+
+# RowCol <> Tile key for AOIs
+library(activemapper)
+library(rmapaccuracy)
+library(sf)
+library(dplyr)
+
+tiles <- read_sf(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+aois <- read_sf(
+  system.file("extdata/aois.geojson", package = "activemapper")
+)
+tiles_aois <- st_join(tiles, aois, largest = TRUE)
+tiles_aois <- tiles_aois %>% select(tile, aoi1) %>% rename(aoi = aoi1)
+
+# extract row_col
+xys <- tiles_aois %>% st_centroid() %>% st_coordinates()
+tile_rcs <- rowcol_from_xy(xys[, 1], xys[, 2], res = 0.05)
+
+# combine into tile, aoi, row, col key
+tile_key <- tibble(
+  as_tibble(tiles_aois), row = tile_rcs[, 1], col = tile_rcs[, 2]
+) %>% select(aoi, tile, col, row, -geometry)
+
+usethis::use_data(tile_key)
+

### 2021-01-29 data-raw/tile_key.R
Fixed tile_key

@@ -20,8 +20,8 @@ tile_rcs <- rowcol_from_xy(xys[, 1], xys[, 2], res = 0.05)
 
 # combine into tile, aoi, row, col key
 tile_key <- tibble(
-  as_tibble(tiles_aois), row = tile_rcs[, 1], col = tile_rcs[, 2]
+  as_tibble(tiles_aois), col = tile_rcs[, 1], row = tile_rcs[, 2]
 ) %>% select(aoi, tile, col, row, -geometry)
 
-usethis::use_data(tile_key)
+usethis::use_data(tile_key, overwrite = TRUE)
 

### 2021-01-30 data-raw/refine_reference_sample.R
Fixed ordering in map reference labels

@@ -22,6 +22,9 @@ tiles <- read_sf(
 )
 gcs <- unname(unlist(st_crs(4326))[2])
 
+# load tile key
+data("tile_key")
+
 # create master_grid extent, get list of probability images, download fpool
 # (for checking)
 
@@ -89,7 +92,8 @@ val_sample_1r <- lapply(unique(val_samples$name), function(x) {
   return(as_tibble(out))
 }) %>% do.call(rbind, .)
 # val_sample_1r <- do.call(rbind, val_sample_1r)
-all(sapply(val_sample_1r, function(x) st_is(x$geom, "POLYGON")))
+all(st_is(st_as_sf(val_sample_1r)$geom, "POLYGON"))
+# all(sapply(st_as_sf(val_sample_1r), function(x) st_is(x$geom, "POLYGON")))
 
 # sum of category
 # val_cats <- val_sample_1r %>% as_tibble() %>%
@@ -98,22 +102,35 @@ all(sapply(val_sample_1r, function(x) st_is(x$geom, "POLYGON")))
 
 # get list of row_col identifiers
 val_sample_1r_sf <- val_sample_1r %>% st_as_sf()
-ints <- sort(unlist(st_intersects(val_sample_1r_sf, tiles)))
-val_tiles <- tiles %>% slice(ints)
-# ggplot2::ggplot(val_tiles) + ggplot2::geom_sf() +
+
+ref_labels <- st_intersection(val_sample_1r_sf, tiles) %>%
+  left_join(., tile_key) %>%
+  select(name, tile, col, row, aoi, assignment_id, worker_id, category,
+         categ_description, tile)
+
+# ggplot2::ggplot(val_tiles) +
+#   ggplot2::geom_sf(ggplot2::aes(fill = as.factor(aoi1))) +
 #   ggplot2::geom_sf(data = val_sample_1r_sf)
 
+
+val_tiles()
 xys <- val_tiles %>% st_centroid() %>% st_coordinates()
+st_centroid(val_tiles)
 val_tile_rcs <- rowcol_from_xy(xys[, 1], xys[, 2], res = 0.05)
 
 # Combine with labels
 cbind(as_tibble(val_sample_1r_sf), tile = val_tiles$tile, val_tile_rcs) %>%
   select(name, assignment_id, worker_id, category, categ_description, tile,
          row, col, geom) %>% st_as_sf() -> ref_labels
[…]

### 2021-01-30 data-raw/extract_probs_for_reference.R
Extracted image probabities and save out to inst/extadata

@@ -1,0 +1,71 @@
+# extract cropland probabilities from prediction images using reference polygons
+# Ran on rstudio instance on EC2, to avoid download into S3
+
+library(sf)
+library(raster)
+library(dplyr)
+library(aws.s3)
+library(activemapper)
+
+# reference polygons
+ref_labels <- read_sf(
+  system.file("extdata/map_reference_labels.geojson", package = "activemapper")
+)
+ref_labels <- ref_labels %>%
+  mutate(image = paste0("image_c", col, "_r", row, ".tif")) %>%
+  select(-geometry)
+
+# pull in tiles and AOIS for filtering
+data("tile_key")
+
+tiles <- read_sf(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+aois <- read_sf(
+  system.file("extdata/aois.geojson", package = "activemapper")
+)
+# tiles_aois <- st_join(tiles, aois, largest = TRUE)
+
+# tiles_aois %>% filter(aoi1 == 8) %>% st_geometry() %>% plot()
+
+
+# ref_labels %>% filter(name == "GH0391094")
+
+# collect image catalog
+bucket_prefixes <- paste0(1:16, "_whole/")
+bucket_prefixes[bucket_prefixes == "3_whole/"] <- "3_whole_retrain/"
+
+img_catalog <- lapply(bucket_prefixes, function(x) { # x <- bucket_prefixes[1]
+  img_root <- glue::glue("/vsis3/activemapper/classified-images/{x}")
+  # aws_string <- glue::glue("aws s3 ls s3://activemapper/classified-images/{x}")
+  # imgs <- system(aws_string, intern = TRUE) %>% as_tibble() %>%
+  #   tidyr::separate(value, sep = c(20, 31),
+  #                   into = c("date", "size", "image")) %>%
+  #   mutate_all(trimws) %>% mutate(prefix = x)
+
+  aoi_id <- gsub("_.*", "", x)
+  # plot(tiles %>% filter(tile %in% tile_ids) %>% st_geometry())
+
+  ref_labels_aoi <- ref_labels %>% filter(aoi %in% aoi_id)
+
+  # plot(ref_labels_aoi$geometry)
+
+  # imgs %>% filter(image == y)
+  tile_ids <- unique(ref_labels_aoi$tile)
+  # which(tile_ids == 486300)
[…]

### 2021-01-30 data-raw/extract_probs_for_reference.R
Update image prob extract script

@@ -42,8 +42,9 @@ img_catalog <- lapply(bucket_prefixes, function(x) { # x <- bucket_prefixes[1]
   #   tidyr::separate(value, sep = c(20, 31),
   #                   into = c("date", "size", "image")) %>%
   #   mutate_all(trimws) %>% mutate(prefix = x)
-
   aoi_id <- gsub("_.*", "", x)
+
+  print(glue::glue("Processing AOI {aoi_id}"))
   # plot(tiles %>% filter(tile %in% tile_ids) %>% st_geometry())
 
   ref_labels_aoi <- ref_labels %>% filter(aoi %in% aoi_id)
@@ -56,7 +57,7 @@ img_catalog <- lapply(bucket_prefixes, function(x) { # x <- bucket_prefixes[1]
   # ref_labels_aoi %>% as_tibble() %>% group_by(tile) %>% count()
   probs <- lapply(1:nrow(ref_labels_aoi), function(y) { # y <- 1
     d <- ref_labels_aoi %>% slice(y)
-    print(d$name)
+    print(glue::glue("..{d$name}"))
     img_nm <- glue::glue("image_c{d$col}_r{d$row}.tif")
     img <- raster(glue::glue("{img_root}{img_nm}"))
     prob <- extract(img, d)[[1]]
@@ -66,6 +67,4 @@ img_catalog <- lapply(bucket_prefixes, function(x) { # x <- bucket_prefixes[1]
   return(probs)
 })
 
-img_catalog_tb <- do.call(rbind, img_catalog) %>% select(prefix, image)
-# img_catalog_tb %>% group_by(image) %>% count() %>%
-#   filter(n > 1)
+save(img_catalog, file = here::here("inst/extdata/image_probs.rda"))

### 2021-02-02 data-raw/intersect_segments_with_reference.R
Fixed extra rows in ref_int_segments by unioning by name

@@ -68,6 +68,15 @@ ref_int_segments <- foreach(x = 1:length(aois), .combine = rbind) %dopar% {
   return(ref_fld_isected)
 }
 
+# union separated segments to produce correct numbers of rows
+# ref_int_segments <- st_read("inst/extdata/ref_int_segments.geojson")
+ref_int_segments <- ref_int_segments %>% group_by(name) %>%
+  summarize(tile = unique(tile), aoi = unique(aoi), category = unique(category),
+            afrac = sum(afrac)) %>% arrange(aoi, tile, name, tile)
+# as_tibble(ref_int_segments) %>% group_by(name) %>% count() %>% filter(n > 1)
+# ref_int_segments %>% filter(name == "GH0005059") %>% plot()
+
 # write out to inst/extdata
-st_write(ref_int_segments, dsn = here("inst/extdata/ref_int_segments.geojson"))
+st_write(ref_int_segments, dsn = here("inst/extdata/ref_int_segments.geojson"),
+         delete_dsn = TRUE)
 

### 2021-02-03 data-raw/cropland_probability_grid.R
Added script for gridding cropland probabilities across the country

@@ -1,0 +1,74 @@
+# Create probability map (grid scale) from RF 3 m probability images
+# Ran on rstudio instance on EC2, to avoid download into S3
+
+library(sf)
+# library(raster)
+library(terra)
+library(dplyr)
+library(aws.s3)
+library(activemapper)
+library(glue)
+library(doMC)
+
+# pull in tiles and AOIs for filtering
+data("tile_key")
+
+bucket_prefixes <- paste0(1:16, "_whole/")
+bucket_prefixes[bucket_prefixes == "3_whole/"] <- "3_whole_retrain/"
+
+tiles <- vect(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+
+r <- rast(ext(tiles), res = 0.05)
+crs(r) <- crs(tiles)
+tilesr <- rasterize(tiles, y = r, field = "tile")
+# tilesr <- fasterize::fasterize(tiles, r, field = "tile")
+tilesr005 <- disaggregate(tilesr, fact = 10)
+tilesr005_sf <- stars::st_as_stars(raster::raster(tilesr005)) %>% st_as_sf()
+# tilesr005_v <- vect(tilesr005)
+# tilesr005_sf <- stars::st_as_stars(tilesr005) %>% st_as_sf() %>%
+#   rename(tile = layer)
+
+lapply(bucket_prefixes, function(x) {
+  # x <- bucket_prefixes[1]
+
+  aoi_prefix <- glue("classified-images/{x}")
+  aoi_id <- gsub("_.*", "", x)
+
+  print(glue::glue("Processing AOI {aoi_id}"))
+
+  # get list of images for AOI and filter using tile_key down to list to read in
+  img_bucket <- get_bucket_df(bucket = "activemapper", prefix = aoi_prefix) %>%
+    select(Key)
+  tile_filter <- tile_key %>% filter(aoi == aoi_id) %>%
+    mutate(image = glue("image_c{col}_r{row}.tif")) %>%
+    mutate(Key = glue("{aoi_prefix}{image}"))
+  # images <- img_bucket %>% filter(Key %in% tile_filter$Key)
+
+  tile_ids <- unique(tile_filter$tile)
+  gridded_probs <- foreach(y = tile_ids, .combine = rbind) %dopar% {
+    # y <- tile_ids[1]
+
+    grids <- tilesr005_sf %>% filter(tile == y) %>%
[…]

### 2021-02-04 data-raw/cropland_probability_grid.R
Updated cropland_probability_grid.R and committed list of aoi probabilities gridded to 0.005 and composited mosaic (as geotiff)

@@ -16,27 +16,39 @@ data("tile_key")
 bucket_prefixes <- paste0(1:16, "_whole/")
 bucket_prefixes[bucket_prefixes == "3_whole/"] <- "3_whole_retrain/"
 
-tiles <- vect(
+# tiles <- vect(
+#   system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+# )
+tiles <- st_read(
   system.file("extdata/ghana_tiles.geojson", package = "activemapper")
 )
 
-r <- rast(ext(tiles), res = 0.05)
-crs(r) <- crs(tiles)
-tilesr <- rasterize(tiles, y = r, field = "tile")
+# r <- rast(ext(tiles), res = 0.05)
+# crs(r) <- crs(tiles)
+# tilesr <- rasterize(tiles, y = r, field = "tile")
+
 # tilesr <- fasterize::fasterize(tiles, r, field = "tile")
-tilesr005 <- disaggregate(tilesr, fact = 10)
-tilesr005_sf <- stars::st_as_stars(raster::raster(tilesr005)) %>% st_as_sf()
+# tilesr005 <- disaggregate(tilesr, fact = 10)
+# tilesr005_sf <- stars::st_as_stars(raster::raster(tilesr005)) %>% st_as_sf()
 # tilesr005_v <- vect(tilesr005)
 # tilesr005_sf <- stars::st_as_stars(tilesr005) %>% st_as_sf() %>%
 #   rename(tile = layer)
+# plot(tilesr)
+# plot(tilesr005_sf$geometry, add = TRUE)
 
-lapply(bucket_prefixes, function(x) {
+logf <- here::here("external/logs/grid_probs.log")
+
+registerDoMC(4)
+aoi_probs <- lapply(bucket_prefixes, function(x) {
   # x <- bucket_prefixes[1]
 
   aoi_prefix <- glue("classified-images/{x}")
   aoi_id <- gsub("_.*", "", x)
 
-  print(glue::glue("Processing AOI {aoi_id}"))
+  cat(glue::glue("Starting {aoi_id} at {Sys.time()}"), file = logf,
+      sep = "\n", append = TRUE)
+
+  # print(glue::glue("Processing AOI {aoi_id}"))
 
   # get list of images for AOI and filter using tile_key down to list to read in
   img_bucket <- get_bucket_df(bucket = "activemapper", prefix = aoi_prefix) %>%
[…]

### 2021-02-04 data-raw/cropland_probability_grid.R
Changed location of cropland probability image

@@ -93,5 +93,5 @@ probs_005 <- do.call(mosaic, aoi_probs)
 probs_005[probs_005 > 100] <- NA
 save(aoi_probs, file = here::here("external/data/results/probs/aoi_probs.rda"))
 writeRaster(probs_005,
-            filename = here::here("external/data/results/probs/probs_005.tif"))
+            filename = here::here("inst/extdata/probs_005.tif"))
 

### 2021-02-04 data-raw/cropland_probability_grid.R
Changed cropland_probability to fraction

@@ -36,21 +36,23 @@ tiles <- st_read(
 # plot(tilesr)
 # plot(tilesr005_sf$geometry, add = TRUE)
 
+# function for estimating proportions of pixels > 0.5
+ctfun <- function(x, na.rm = na.rm) {
+  v <- x[!is.na(x)]
+  length(which(v > 50)) / length(v)
+}
+
+# logfile
 logf <- here::here("external/logs/grid_probs.log")
-
-registerDoMC(4)
-aoi_probs <- lapply(bucket_prefixes, function(x) {
+cropfrac_aoi <- lapply(bucket_prefixes, function(x) {
   # x <- bucket_prefixes[1]
-
   aoi_prefix <- glue("classified-images/{x}")
   aoi_id <- gsub("_.*", "", x)
 
   cat(glue::glue("Starting {aoi_id} at {Sys.time()}"), file = logf,
       sep = "\n", append = TRUE)
 
-  # print(glue::glue("Processing AOI {aoi_id}"))
-
-  # get list of images for AOI and filter using tile_key down to list to read in
+    # get list of images for AOI and filter using tile_key down to list to read in
   img_bucket <- get_bucket_df(bucket = "activemapper", prefix = aoi_prefix) %>%
     select(Key)
   tile_filter <- tile_key %>% filter(aoi == aoi_id) %>%
@@ -59,39 +61,28 @@ aoi_probs <- lapply(bucket_prefixes, function(x) {
   # images <- img_bucket %>% filter(Key %in% tile_filter$Key)
 
   tile_ids <- unique(tile_filter$tile)
-  # gridded_probs <- foreach(y = tile_ids[1:10]) %dopar% {
-  gridded_probs <- foreach(y = tile_ids) %dopar% {
-    # y <- tile_ids[1]
-    cat(glue::glue("...aggregating tile {y}"),
-        file = logf, sep = "\n", append = TRUE)
-
-    # grids <- tilesr005_sf %>% filter(tile == y) %>%
-    #   mutate(id = 1:nrow(.)) %>% select(id) %>% vect(.)
+  cropfracs <- mclapply(tile_ids, function(y) {
+  # cropfracs <- mclapply(tile_ids[1:10], function(y) {
+    # cat(glue::glue("...aggregating tile {y}"),
+    #     file = logf, sep = "\n", append = TRUE)
[…]

### 2021-02-05 inst/manuscript/figures.Rmd
Code for figure 8 (main map) complete

@@
   system.file("extdata", "ghana_grid.csv", package = "activemapper")
-) %>% select(-fwts)
+) %>% dplyr::select(-fwts)
 africa <- st_read(
@@
 africa_cont <- africa %>% st_union %>% st_sf
-ghana <- africa %>% filter(name == "Ghana") %>% select(geometry)
+ghana <- africa %>% filter(name == "Ghana") %>% dplyr::select(geometry)
 
@@
 
+# segment files
 segment_files <- dir(here("external/data/results/segments/merged"),
@@
 
-fdensity <- stars::read_stars(
-  system.file("extdata", "field_density005.tif", package = "activemapper")
+# fdensity <- stars::read_stars(
+#   system.file("extdata", "field_density005.tif", package = "activemapper")
+# )
+
+# entire cropland fraction
+cropfrac <- stars::read_stars(
+  system.file("extdata/probs_005.tif", package = "activemapper")
 )
+
+# individual fractions
+tile_list <- c(486794, 539730, 591243, 616490)
+tile_ss <- tiles %>% filter(tile %in% tile_list)
+
+data("tile_key")
+s3root <- "/vsis3/activemapper"
+prob_tiles <- tile_key %>% filter(tile %in% tile_list) %>% 
+  mutate(impath = glue("{s3root}/classified-images/{aoi}_whole/", 
+                       "image_c{col}_r{row}.tif"))
 
@@
   bucket = "activemapper", prefix = "planet/composite_sr_buf_fix/OS/", max = Inf
-) %>% as_tibble %>% select(Key)
+) %>% as_tibble %>% dplyr::select(Key)
+impath <- osimgs %>% 
+  filter(grepl(paste0(tile_ss$tile, collapse = "|"), Key)) %>% 
+  mutate(pth = glue("{s3root}/{Key}")) %>% pull(pth)
 
@@
 
+# read in field segments
 flds_1 <- read_sf(segment_files$file[1])
@@
 
-tile_ss <- tiles %>% filter(tile %in% c(486794, 539730, 591243, 616490))
 # ggplot(aois %>% filter(aois == 1)) + geom_sf() + 
@@
 
-impath <- osimgs %>% 
-  filter(grepl(paste0(tile_ss$tile, collapse = "|"), Key)) %>% 
-  mutate(pth = paste0("/vsis3/activemapper/", Key)) %>% pull(pth)
-
 # img <- read_stars(impath)
@@
 # imgngb <- st_rgb(img[, , , 4:2], 3, maxColorValue = mxval[4:2])
-
+vpal <- viridis::viridis(20)
+inset_cols <- c("red", "orange", "pink", "brown")
 fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
-isets <- lapply(1:4, function(x) {
+isets <- lapply(1:4, function(x) {  # x <- 1
   imgr <- brick(impath[x])[[2:4]]
+  probr <- stars::read_stars(prob_tiles$impath[x])
   xylims <- st_bbox(imgr)
@@
             size = 0.2) + 
-    geom_sf(data = tile_ss[x, ], color = "red", fill = "transparent") + 
+    geom_sf(data = tile_ss[x, ], color = inset_cols[x], fill = "transparent") + 
     scale_x_continuous(expand = c(0, 0)) + 
@@
     theme_void()
+  iset2 <- ggplot() + 
+    geom_stars(data = probr) +
+    scale_fill_gradientn(name = "", colours = vpal, limits = c(0, 100)) +
+    geom_sf(data = tile_ss[x, ],  color = inset_cols[x], fill = "transparent") +
+    scale_x_continuous(expand = c(0, 0)) + 
+    scale_y_continuous(expand = c(0, 0)) + 
+    coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + 
+    theme_void() + theme(legend.position = "none")
+  cowplot::plot_grid(iset2, iset, ncol = 1, align = "hv")
 })
-isets[[2]]
+# isets[[2]]
 
@@
 # twk <- diff(range(xylims[, 1])) * 0.05
-maprange <-  c(range(xylims[, 1]) + c(-1.25, 1.25), 
+maprange <-  c(range(xylims[, 1]) + c(-1.2, 1.2), 
                range(xylims[, 2]))
 
-inset_cols <- c("red", "orange", "pink", "brown")
+# cropland fractions
 p <- ggplot(africa) + 
   geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = fdensity * 100) +
+  geom_stars(data = cropfrac) +
   geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  geom_sf(data = tile_ss, fill = "red", col = "red") +
+  geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
+  # geom_sf(data = tile_ss, fill = "red", col = "red") +
   scale_fill_viridis_c(name = "% cover", na.value = "transparent") +
@@
   theme_void() + 
-  theme(legend.position = "bottom", 
-        legend.text = element_text(colour = "white"), 
-        legend.title = element_text(colour = "white"))
-
+  theme(legend.position = c(0.65, 0.07), legend.direction = "horizontal", 
+        legend.key.height = unit(0.3, "cm")) 
+        # legend.text = element_text(colour = "white"), 
+        # legend.title = element_text(colour = "white"))
+# with insets added
+wh <- c(0.4, 0.4)
+lx <- -0.1
+rx <- 0.7
 inset_map <- ggdraw() +
   draw_plot(p) +
-  draw_plot(isets[[1]], x = 0.07, y = 0.77, width = 0.2, height = 0.2) + 
-  draw_plot(isets[[2]], x = 0.72, y = 0.6, width = 0.2, height = 0.2) + 
-  draw_plot(isets[[3]], x = 0.04, y = 0.37, width = 0.2, height = 0.2) + 
-  draw_plot(isets[[4]], x = 0.76, y = 0.25, width = 0.2, height = 0.2)
+  draw_plot(isets[[1]], x = lx, y = 0.6, width = wh[1], height = wh[2]) + 
+  draw_plot(isets[[2]], x = rx, y = 0.51, width = wh[1], height = wh[2]) + 
+  draw_plot(isets[[3]], x = lx, y = 0.17, width = wh[1], height = wh[2]) + 
+  draw_plot(isets[[4]], x = rx, y = 0.07, width = wh[1], height = wh[2])
+# inset_map
 
-ggsave(filename = here("presentations/figures/cropcover_with_insets.png"), 
+ggsave(filename = pth("figures/figure8.png"), 
        plot = inset_map, bg = "transparent", width = diff(maprange[1:2]), 

### 2021-02-09 data-raw/error_matrices.R
Added raw scripts for calculating error matrices, Ghana AEZ, mapped areas

@@ -1,0 +1,93 @@
+
+# calculate map accuracy
+# Following Stehman and Foody (2019)
+library(activemapper)
+library(raster)
+library(dplyr)
+library(sf)
+
+# load data
+data("mapped_areas")
+load(system.file("extdata/ref_int_probs.rda", package = "activemapper"))
+ref_int_segments <- read_sf(
+  system.file("extdata/ref_int_segments.geojson", package = "activemapper")
+)
+
+# load aez
+aez <- read_sf(
+  system.file("extdata/ghana_aez.geojson", package = "activemapper")
+) %>% select(ID, Name) %>% rename(aezid = ID, aez = Name)
+
+# Pick up AEZ intersections
+ref_int_segments <- st_intersection(ref_int_segments, aez) %>%
+  select(name, tile, aoi, aezid, aez, category, afrac)
+ref_int_segments %>% as_tibble() %>% group_by(aoi) %>% count()
+
+# set up data for error matrix
+error_matrix_data <- as_tibble(ref_int_segments) %>%
+  # select(name, tile, aoi, aezid, aez, category, afrac) %>%
+  left_join(., ref_int_probs, by = c("name", "tile", "aoi")) %>%
+  select(name, tile, aoi, aezid, aez, category, afrac, mean) %>%
+  rename(seg_pct = afrac, rf_prob = mean) %>%
+  mutate(seg_pct = seg_pct * 100) %>%
+  mutate(ref = case_when(
+    category == "annualcropland" ~ 1,
+    category == "noncropland" ~ 0,
+    category == "unsure1" ~ 1,
+    category == "unsure2" ~ 0
+  )) %>%
+  mutate(seg_class = ifelse(seg_pct < 0.5, 0, 1),
+         prob_class = ifelse(rf_prob < 50, 0, 1)) %>%
+  arrange(aoi)
+
+# Create error matrix based on probability images
+# helper function for error matrix
+error_matrix <- function(pred, ref) {
+  x <- table("Predicted" = pred, "Reference" = ref)
+  rbind(cbind(x, Total = rowSums(x)), Total = c(colSums(x), sum(x)))
+}
+
+# plot(st_geometry(aez))
+# plot(st_geometry(st_centroid(ref_int_segments)), add = TRUE)
+
+# calculate error matrices by AOI
+error_matrices_aoi <- lapply(unique(error_matrix_data$aoi), function(x) {
+  dat <- error_matrix_data %>% filter(aoi == x)
[…]

### 2021-02-09 data-raw/ghana_aez.R

@@ -1,0 +1,13 @@
+## Ghana Agroecozone dataset
+
+# Data accessed from here:
+# https://wheregeospatial.com/agro-ecological-zones-ghana/
+# Stored locally and unzipped
+
+library(sf)
+library(dplyr)
+
+"~/Dropbox/data/agricultural/ghana/aez/Ghana_agroecologicalzones.shp" %>%
+  read_sf() %>%
+  st_transform(., crs = 4326) %>%
+  st_write(., "inst/extdata/ghana_aez.geojson")

### 2021-02-09 data-raw/mapped_areas.R

@@ -1,0 +1,136 @@
+# Calculate cropland area from cropland fraction and segments, for accuracy
+# assessment and area estimation.
+library(sf)
+library(dplyr)
+library(here)
+
+# Datasets
+aois <- st_read(
+  system.file("extdata/aois.geojson", package = "activemapper")
+) # %>% st_transform(crs = st_crs("ESRI:102022")$wkt)
+# sum(aois_proj$area) == sum(as.numeric(st_area(aois_proj)) / 10^6)
+africa <- st_read(
+  system.file("extdata/africa.geojson", package = "activemapper")
+)
+aez <- st_read(
+  system.file("extdata/ghana_aez.geojson", package = "activemapper")
+)
+
+# Cropland fractions from probability images
+cropfrac <- raster(
+  system.file("extdata/cropfrac_005.tif", package = "activemapper")
+)
+
+# Read in segments, drop original AOI3
+segment_files <- dir(here("external/data/results/segments/merged"),
+                     pattern = "boundarymerge", full.names = TRUE)
+iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files)))
+segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid)
+segment_files <- segment_files %>% filter(!grepl("aoi3.*.merge.geo", file))
+
+#------------------------------------------------------------------------------#
+# Cropland fractions - in both AOIs and AEZ
+# First mask out bad pixels on coasts
+
+# mask twice
+cropfrac <- raster(terra::mask(terra::rast(cropfrac), terra::vect(aois)))
+cropfrac <- africa %>%
+  filter(name %in% c("Ghana", "Togo", "Cte d'Ivoire", "Burkina Faso")) %>%
+  mask(cropfrac, .)
+
+# rasterize aois
+aoisr <- terra::vect(aois) %>%
+  terra::rasterize(., y = terra::rast(cropfrac)) %>%
+  terra::mask(., terra::rast(cropfrac)) %>%
+  raster(.)
+# plot(aoisr); plot(cropfrac)
+
+# rasterize aez
+aezr <- terra::vect(aez) %>%
+  terra::rasterize(., y = terra::rast(cropfrac)) %>%
[…]

### 2021-02-11 inst/manuscript/figures.Rmd
Function for outputting accuracy reports, scripts for processing accuracy reports, figure for accuracy tables with AOI zones

@@
 ```{r setup, include=FALSE, cache=FALSE, message = FALSE}
+library(activemapper)
 library(raster)
 library(sf)
+library(stars)
 library(dplyr)
@@
 library(ggplot2)
+library(cowplot)
+library(RColorBrewer)
 library(here)
-library(RColorBrewer)
-library(activemapper)
+library(glue)
+library(aws.s3)
+library(RStoolbox)
+# library(gridExtra)
+library(ggpmisc)
+library(patchwork)
+
 pth <- function(x) file.path(here("inst/manuscript"), x)
@@
 )
+aez <- st_read(
+  system.file("extdata/ghana_aez.geojson", package = "activemapper")
+)
+
 # ggplot(data = tiles) + 
@@
 ### Figure 7: Categorical map accuracy
+Place map accuracy tables and zone report into single figure, for AOI zones.
+
+```{r, eval=FALSE}
+data("accuracy_reports")
+
+# zone maps
+zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
+                  "aez" = list(1, 2:4, 5, 6:7))
+aoi_zones <- aois %>% 
+  mutate(zone = case_when(
+    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
+    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
+    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
+    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(area))
+
+# RF accuracy
+rfreports <-  lapply(accuracy_reports$zones$aoi, function(x) x$rf)
+rfreports[[length(rfreports) + 1]] <- accuracy_reports$ghana$rf
+rf_tabs <- lapply(1:5, function(i) {  # i <- 2
+  m <- rfreports[[i]]
+  d <- data.frame(m, check.names = FALSE)
+  rbind(d, rep("", ncol(d)))
+}) %>% do.call(rbind, .)
+zvec <- rep("", nrow(rf_tabs))
+znames <- paste(c(paste("Zone", 1:4), "Ghana"), "  ")
+zvec[c(3, seq(8, length(zvec), by = 5))] <- znames
+rf_tabs <- cbind(zvec, rf_tabs)
+colnames(rf_tabs) <- c("", colnames(rfreports$North))
+
+rf_tbl <- tibble(x = 1, y = 10, dat = list(rf_tabs))
+
+# segmentation accuracy
+segreports <-  lapply(accuracy_reports$zones$aoi, function(x) x$seg)
+segreports[[length(segreports) + 1]] <- accuracy_reports$ghana$seg
+seg_tabs <- lapply(1:5, function(x) {  # x <- 1
+  m <- segreports[[x]]
+  d <- data.frame(m, check.names = FALSE)
+  rbind(d, rep("", ncol(d)))
+}) %>% do.call(rbind, .)
+colnames(seg_tabs) <- colnames(rfreports$North)
+seg_tbl <- tibble(x = 1, y = 10, dat = list(seg_tabs))
+
+# Assemble plots with maps
+# Table theme
+tt <- ttheme_gtminimal(
+  base_size = 4.5, 
+  padding = unit(c(0.2, 0.2), "char"),
+  core = list(fg_params = list(hjust = 1, x = 1))
+)
+
+m <- ggplot(aoi_zones) + geom_sf(lwd = 0.1) + 
+  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 2) + 
+  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+t1 <- ggplot() + 
+  geom_table(data = rf_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
+  ggtitle("Per-pixel classification") + 
+  theme_void() + 
+  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6))
+t2 <- ggplot() + 
+  geom_table(data = seg_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
+  ggtitle("Segmentation") +
+  theme_void() + 
+  theme(plot.margin = unit(c(0,0,0,0), "cm"), 
+        plot.title = element_text(hjust = 0.5, vjust = 1, size = 6))
+
+# gp <- cowplot::plot_grid(t1, m, t2, nrow = 1, align = "hv", 
+#                          # rel_widths = c(0.2, 0.4, 0.5, 0.4, 0.2),
+#                          rel_widths = c(1, 1, 0.8),
+#                          # rel_heights = c(0.5, 1.5, 0.5),
+#                          scale = c(0.25, 0.95, 0.25),
+#                          label_size = 5, 
+#                          label_x = 0.25, 
+#                          labels = c("Per-pixel classification", "",
+#                                     "Segmentation"))
+gp <- t1 + m + t2
+ggsave(gp, filename = pth("figures/figure7.png"), width = 7.25, height = 2.35, 
+       units = "in", dpi = 300)
+```
 
@@
 ```{r, eval=FALSE}
-library(stars)
-library(ggplot2)
-library(aws.s3)
-library(RStoolbox)
-library(cowplot)
-
 # segment files
@@
 
-# entire cropland fraction
-cropfrac <- stars::read_stars(
-  system.file("extdata/probs_005.tif", package = "activemapper")
+# load in Ghana cropland fractions, mask out tiles over ocean, convert to stars
+# percent
+cropfrac <- raster(
+  system.file("extdata/cropfrac_005.tif", package = "activemapper")
 )
+cropfrac <- africa %>% 
+  filter(name %in% c("Ghana", "Togo", "Cte d'Ivoire", "Burkina Faso")) %>% 
+  mask(cropfrac, .)
+cropfrac <- st_as_stars(cropfrac * 100)
 
@@
 
+# process inset maps
 # img <- read_stars(impath)
@@
   # geom_sf(data = tile_ss, fill = "red", col = "red") +
-  scale_fill_viridis_c(name = "% cover", na.value = "transparent") +
+  scale_fill_viridis_c(name = "%", na.value = "transparent") +
   coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
   theme_void() + 
-  theme(legend.position = c(0.65, 0.07), legend.direction = "horizontal", 
-        legend.key.height = unit(0.3, "cm")) 
-        # legend.text = element_text(colour = "white"), 
-        # legend.title = element_text(colour = "white"))
+  theme(legend.position = c(0.64, 0.08), legend.direction = "horizontal", 
+        legend.key.height = unit(0.3, "cm"), 
+        legend.text = element_text(colour = "black"),
+        legend.title = element_text(size = 10, colour = "black"))
 # with insets added
@@
        height = diff(maprange[3:4]), units = "in", dpi = 300)
-
-  # geom_stars(data = imgngb) + scale_fill_identity()
-  # geom_sf(data = flds_tile)
-
-plot(flds_tile[1])
-
-fboundary <- read_sf(segment_files$file[1])
-
-xylims <- st_coordinates(ghana)
-maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))
-
-pal <- RColorBrewer::brewer.pal(n = 9, "Spectral")[c(1, 2, 3, 4, 9)]
-
-# Panels, organized top to bottom
-# Panel 1: Cropland probabilities, with 3 m probability side maps (tiles)
-p <- ggplot(africa) + 
-  geom_sf(fill = "gray97", color = "grey", size = 0.25) +  
-  geom_stars(data = fdensity * 100) + 
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") + 
-  scale_fill_viridis_c(name = "% cover", na.value = "transparent") + 
-  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
-  theme_void()
-
-# Panel 2: Cropland cover, with polygon side maps over composites (tiles)
-p <- ggplot(africa) + 
-  geom_sf(fill = "gray97", color = "grey", size = 0.25) +  
-  geom_stars(data = fdensity * 100) + 
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") + 
-  geom_sf(data = tile_ss, fill = "red", col = "red") + 
-  scale_fill_viridis_c(name = "% cover", na.value = "transparent") + 
-  coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
-  theme_void()
-p
-
-# Panel 3: Average field size (kriging adjusted?),
-
-# Panel 4: Field number (kriging adjusted)
 

### 2021-02-11 inst/manuscript/manuscript.Rmd
Function for outputting accuracy reports, scripts for processing accuracy reports, figure for accuracy tables with AOI zones

@@
 
+```{r mainmap, echo = FALSE, out.width="90%", fig.cap="The distribution of arable croplands in Ghana.
+The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities.
+The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles.
+Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/figure8.png')
+```
+
 # Discussion

### 2021-02-11 R/accuracy_report.R

@@ -1,0 +1,51 @@
+#' Produce table of accuracy and area measures
+#' @param err_mat The error matrix, with reference sample and map sample in rows
+#' @param acc_mat The output list produced by `accuracy`
+#' @param classes Vector of names for the map/reference classes
+#' @param digits Number of decimal values in outputs
+#' @param z Width of confidence interval for reported standard errors
+#' @details Produces a matrix for reporting that contains the
+#' configuration for reporting map accuracy recommended by Stehman and Foody
+#' (2019), with adjusted areas of each class provided.
+#' @references
+#' Stehman, S.V. & Foody, G.M. (2019) Key issues in rigorous accuracy assessment
+#' of land cover products. Remote Sensing of Environment, 231, 111199.
+#' @return A character matrix.
+#' @export
+#' @examples
+#' # Example 1: Table 4 in Olofsson et al 2014a
+#' err_mat <- cbind(r1 = c(97, 3, 2), r2 = c(0, 279, 1), r3 = c(3, 18, 97))
+#' areas <- c(22353, 1122543, 610228)
+#' acc_mat <- accuracy(err_mat, areas)
+#' accuracy_report(err_mat, acc_mat, classes = paste0("Class", 1:3))
+
+accuracy_report <- function(err_mat, acc_mat, classes, digits = 1, z = 1.96) {
+
+  rnd <- function(tab, d = digits) sprintf(glue("%0.{d}f"), tab)
+  a <- acc_mat$accuracy * 100
+
+  # set up error matrix as percent
+  Pij <- acc_mat$Pij * 100
+  Pij <- round(cbind(Pij, rowSums(Pij)), digits)
+
+  # Areas
+  b <- acc_mat$area
+
+  # U, P, O with SE
+  U <- glue("{rnd(a[, 'U'])} ({rnd(a[, 'U_se'] * z)})")
+  P <- glue("{rnd(a[, 'P'])} ({rnd(a[, 'U_se'] * z)})")
+
+  O <- glue("{rnd(a[1, 'O'])} ({rnd(a[1, 'O_se'] * z)})")
+  A <- glue("{rnd(b[, 1], 0)} ({rnd(b[, 2] * z, 0)})")
+
+  # put together output matrix
+  headers <- c("", classes, "Total", "U", "O", "n", "Area")
+  outmat <- rbind(cbind(Pij, U, c(O, ""), rowSums(err_mat), A),
+                  c(P, rep("", 5)),
[…]

### 2021-02-11 data-raw/accuracy_reports.R

@@ -1,0 +1,94 @@
+# Map accuracy reports (as tables)
+
+library(activemapper)
+library(dplyr)
+library(glue)
+
+data("mapped_areas")
+data("error_matrices")
+
+# Accuracies for AOI and AEZ groups
+# Variables
+grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
+             "aez" = list(1, 2:4, 5, 6:7))
+listkeys <- c("aoi", "aez")
+grpnames <- list("aoi" = c("North", "Central", "SW", "SE"),
+                 "aez" = c("Coast_Sav", "Forest", "Transition", "Guinea_Sav"))
+classes <- c("Non-crop", "Crop")
+
+# Process
+zone_reports <- lapply(1:2, function(x) {  # x <- 2
+
+  listkey <- listkeys[[x]]
+  grp <- grps[[x]]
+  grpname <- grpnames[[x]]
+
+  reports <- lapply(grp, function(y) {  # y <- grp[[2]]
+
+    # sum samples and areas into zones
+    # rf
+    rfl <- lapply(error_matrices[[listkey]][y], function(z) {
+      m <- z$rf[1:2, 1:2]
+      if(!"1" %in% rownames(m)) m[2, ] <- c(0, 0)
+      m <- unname(m)
+      m
+    })
+    rfmat <- Reduce(`+`, rfl)
+    rfarea <- mapped_areas[[glue("rf_areas_{listkey}")]] %>%
+      filter(get(listkey) %in% y) %>% select(nocrop_area, crop_area) %>%
+      summarize_all(sum) %>% unlist()
+
+    # accuracy and report
+    rf_acc <- accuracy(rfmat[1:2, 1:2], areas = rfarea)
+    rf_report_aoi <- accuracy_report(rfmat[1:2, 1:2], rf_acc, classes)
+
+
+    # segs
+    if(listkey == "aoi") {
+      segl <- lapply(error_matrices[[listkey]][y], function(z) {
+        m <- z$seg[1:2, 1:2]
+        if(!"1" %in% rownames(m)) m[2, ] <- c(0, 0)
+        m <- unname(m)
+        m
+      })
+      segmat <- Reduce(`+`, segl)
+      segarea <- mapped_areas[[glue("seg_areas_{listkey}")]] %>%
+        filter(get(listkey) %in% y) %>% select(nocrop_area, crop_area) %>%
[…]

### 2021-02-11 data-raw/mapped_areas.R

@@ -107,7 +107,6 @@ classarea_rf_aez <- as_tibble(class_freq_aez) %>%
 
 #------------------------------------------------------------------------------#
 # Segments (AOIs and Ghana only)
-
 classarea_segments_aoi <- lapply(1:16, function(x) {  # x <- 1
   # i <- segment_files$file[1]
   f <- segment_files$file[x]
@@ -123,14 +122,15 @@ classarea_segments_aoi <- lapply(1:16, function(x) {  # x <- 1
   aoi_area <- aois %>% filter(aois == aoi_id) %>% pull(area)
 
   # output
-  areas <- c("aoi" = aoi_id, "crop_area" = sum(fld_area),
-             "nocrop_area" = aoi_area)
+  areas <- tibble("aoi" = aoi_id, "crop_area" = sum(fld_area),
+                  "nocrop_area" = aoi_area - sum(fld_area))
   return(areas)
-})
+}) %>% do.call(rbind, .)
+
 
 #------------------------------------------------------------------------------#
 # save out
 mapped_areas <- list("rf_areas_aoi" = classarea_rf_aoi,
                      "rf_areas_aez" = classarea_rf_aez,
                      "seg_areas_aoi" = classarea_segments_aoi)
-usethis::use_data(mapped_areas)
+usethis::use_data(mapped_areas, overwrite = TRUE)

### 2021-02-16 inst/manuscript/figures.Rmd
Additional accuracy metrics added to SI.

@@
 
-# zone maps
-zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
-                  "aez" = list(1, 2:4, 5, 6:7))
-aoi_zones <- aois %>% 
-  mutate(zone = case_when(
-    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
-    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
-    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
-    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
-  ) %>%
-  group_by(zone) %>% 
-  summarize(area = sum(area))
-
 # RF accuracy
@@
 
-rf_tbl <- tibble(x = 1, y = 10, dat = list(rf_tabs))
+# rf_tbl <- tibble(x = 0, y = 10, dat = list(rf_tabs))
 
@@
 }) %>% do.call(rbind, .)
-colnames(seg_tabs) <- colnames(rfreports$North)
-seg_tbl <- tibble(x = 1, y = 10, dat = list(seg_tabs))
+seg_tabs <- cbind(zvec, seg_tabs)
+colnames(seg_tabs) <- c("", colnames(rfreports$North))#rep("", ncol(seg_tabs))
+# seg_tbl <- tibble(x = 0, y = 10, dat = list(seg_tabs))
+
+all_tbl <- tibble(x = 0, y = 10, dat = list(rbind(rf_tabs, seg_tabs)))
 
@@
 tt <- ttheme_gtminimal(
-  base_size = 4.5, 
-  padding = unit(c(0.2, 0.2), "char"),
-  core = list(fg_params = list(hjust = 1, x = 1))
+  base_size = 10, 
+  padding = unit(c(0.7, 0.3), "char"),
+  core = list(fg_params = list(hjust = 1, x = 1)), 
+  colhead = list(fg_params = list(hjust = 1, x = 1))
 )
 
-m <- ggplot(aoi_zones) + geom_sf(lwd = 0.1) + 
-  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 2) + 
-  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-t1 <- ggplot() + 
-  geom_table(data = rf_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
-  ggtitle("Per-pixel classification") + 
+t1 <- ggplot(data.frame(x = c(0, 10), y = c(0, 10))) + 
+  geom_point(aes(x = x, y = y), shape = "") +  
+  geom_table(data = all_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
+  # ggtitle("Per-pixel classification") + 
+  annotate(geom = "text", x = -0.2, y = 7, #y = 1, yend = 5, 
+           label = "Per-pixel classification", angle = 90) +
+  annotate(geom = "text", x = -0.2, y = 2.5, label = "Segmentation", 
+           angle = 90, hjust = 1) +
+  annotate(geom = "segment", x = -0.2, xend = 10, y = 9.77, yend = 9.77) + 
+  annotate(geom = "segment", x = -0.2, xend = 10, y = 10.02, yend = 10.02) + 
+  annotate(geom = "segment", x = -0.2, xend = 10, y = 4.7, yend = 4.7, 
+           size = 0.3) + 
   theme_void() +
-  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6))
-t2 <- ggplot() + 
-  geom_table(data = seg_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
-  ggtitle("Segmentation") +
-  theme_void() + 
-  theme(plot.margin = unit(c(0,0,0,0), "cm"), 
-        plot.title = element_text(hjust = 0.5, vjust = 1, size = 6))
-
-# gp <- cowplot::plot_grid(t1, m, t2, nrow = 1, align = "hv", 
-#                          # rel_widths = c(0.2, 0.4, 0.5, 0.4, 0.2),
-#                          rel_widths = c(1, 1, 0.8),
-#                          # rel_heights = c(0.5, 1.5, 0.5),
-#                          scale = c(0.25, 0.95, 0.25),
-#                          label_size = 5, 
-#                          label_x = 0.25, 
-#                          labels = c("Per-pixel classification", "",
-#                                     "Segmentation"))
-gp <- t1 + m + t2
-ggsave(gp, filename = pth("figures/figure7.png"), width = 7.25, height = 2.35, 
+  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6), 
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+# gp <- t1 + m
+ggsave(t1, filename = pth("figures/figure7.png"), width = 7, height = 7.9, 
        units = "in", dpi = 300)
@@
 
-### Figure S9: Segmentation statistics (grid scale)
+### Figure S9: AOI Zones
+
+```{r, eval=FALSE}
+# zone maps
+zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
+                  "aez" = list(1, 2:4, 5, 6:7))
+aoi_zones <- aois %>% 
+  mutate(zone = case_when(
+    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
+    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
+    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
+    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(area))
+
+rmapshaper::ms_simplify(aez) %>% plot()
+aez %>% st_buffer(dist = 0.01) %>% st_union(by_feature = TRUE) %>% 
+  st_buffer(dist = -0.01) %>% 
+  mutate(zone = case_when(
+    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
+    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
+    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
+    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(AREA)) %>% plot()
+
+aez_zones <- rmapshaper::ms_simplify(aez, keep = 1) %>% 
+  mutate(zone = case_when(
+    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
+    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
+    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
+    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(AREA)) 
+# plot(aez_zones["area"])
+# plot(aez["AREA"])
+
+m <- ggplot(aois) + geom_sf(lwd = 0.1) +
+  geom_sf(data = aoi_zones, lwd = 0.5, fill = "transparent") + 
+  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 4) + 
+  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+m2 <- ggplot(aez %>% filter(ID != 8)) + geom_sf(lwd = 0.1) +
+  geom_sf(data = aez_zones, lwd = 0.5, fill = "transparent") + 
+  geom_sf_text(data = st_centroid(aez_zones), aes(label = zone), size = 4) + 
+  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+ggsave(m + m2 + plot_annotation(tag_levels = 'A'), 
+       filename = pth("figures/si_aoi_zones.png"), width = 3.5 * 2, 
+       height = 4.5, units = "in", dpi = 300)
+```
+
+### Figure S10: AEZ accuracies
+
+Grouped by four zones, calculated for per-pixel classifications only.
+
+```{r, eval=FALSE}
+data("accuracy_reports")
+
+# RF accuracy
+rfreports <- lapply(accuracy_reports$zones$aez, function(x) x$rf)
+rf_tabs <- lapply(1:4, function(i) {  # i <- 2
+  m <- rfreports[[i]]
+  d <- data.frame(m, check.names = FALSE)
+  rbind(d, rep("", ncol(d)))
+}) %>% do.call(rbind, .)
+zvec <- rep("", nrow(rf_tabs))
+znames <- paste(paste("Zone", 1:4), "  ")
+zvec[c(3, seq(8, length(zvec), by = 5))] <- znames
+rf_tabs <- cbind(zvec, rf_tabs)
+colnames(rf_tabs) <- c("", colnames(rfreports$Coast_Sav))
+
+all_tbl <- tibble(x = 0, y = 10, dat = list(rf_tabs))
+
+# Assemble plots with maps
+# Table theme
+tt <- ttheme_gtminimal(
+  base_size = 10, 
+  padding = unit(c(0.7, 0.3), "char"),
+  core = list(fg_params = list(hjust = 1, x = 1)), 
+  colhead = list(fg_params = list(hjust = 1, x = 1))
+)
+
+t1 <- ggplot(data.frame(x = c(0, 10), y = c(0, 10))) + 
+  geom_point(aes(x = x, y = y), shape = "") +  
+  geom_table(data = all_tbl, aes(x = x, y = y, label = dat), table.theme = tt) + 
+  annotate(geom = "segment", x = 0, xend = 10, y = 9.5, yend = 9.5) + 
+  annotate(geom = "segment", x = 0, xend = 10, y = 10.01, yend = 10.01) + 
+  theme_void() +
+  theme(plot.title = element_text(hjust = 0.5, vjust = 1, size = 6), 
+        plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+ggsave(t1, filename = pth("figures/si_aez_accuracies.png"), width = 7, 
+       height = 3.5, units = "in", dpi = 300)
+```
+
+### Figure S11: Segmentation statistics (grid scale)
 

### 2021-02-16 data-raw/refine_reference_sample.R

@@ -1,10 +1,11 @@
-# Collect data from probability images for accuracy assessment
-# Uses polygons collected from validator
-#
+# Refine initial reference sample placed using validator
+# Data are a mix of points and polygons, collected by two workers, with a small
+# sample of overlap
+# Refined sample used to extract probabilities and segments from mapped results
+# for accuracy assessment.
+# Overlap samples used to assess inter-rater agreement
 
 ## Map reference site selection
-library(aws.s3)
-library(raster)
 library(sf)
 library(dplyr)
 library(rmapaccuracy)
@@ -91,9 +92,7 @@ val_sample_1r <- lapply(unique(val_samples$name), function(x) {
   }
   return(as_tibble(out))
 }) %>% do.call(rbind, .)
-# val_sample_1r <- do.call(rbind, val_sample_1r)
 all(st_is(st_as_sf(val_sample_1r)$geom, "POLYGON"))
-# all(sapply(st_as_sf(val_sample_1r), function(x) st_is(x$geom, "POLYGON")))
 
 # sum of category
 # val_cats <- val_sample_1r %>% as_tibble() %>%
@@ -108,25 +107,33 @@ ref_labels <- st_intersection(val_sample_1r_sf, tiles) %>%
   select(name, tile, col, row, aoi, assignment_id, worker_id, category,
          categ_description, tile)
 
-# ggplot2::ggplot(val_tiles) +
-#   ggplot2::geom_sf(ggplot2::aes(fill = as.factor(aoi1))) +
-#   ggplot2::geom_sf(data = val_sample_1r_sf)
-
-
-val_tiles()
-xys <- val_tiles %>% st_centroid() %>% st_coordinates()
-st_centroid(val_tiles)
-val_tile_rcs <- rowcol_from_xy(xys[, 1], xys[, 2], res = 0.05)
-
-# Combine with labels
-cbind(as_tibble(val_sample_1r_sf), tile = val_tiles$tile, val_tile_rcs) %>%
-  select(name, assignment_id, worker_id, category, categ_description, tile,
-         row, col, geom) %>% st_as_sf() -> ref_labels
+# # ggplot2::ggplot(val_tiles) +
+# #   ggplot2::geom_sf(ggplot2::aes(fill = as.factor(aoi1))) +
+# #   ggplot2::geom_sf(data = val_sample_1r_sf)
+# xys <- val_tiles %>% st_centroid() %>% st_coordinates()
+# # st_centroid(val_tiles)
[…]

### 2021-02-16 inst/manuscript/si.Rmd
Add changes to SI (missed previous commit)

@@
 ## Map accuracy
-### Categorical
-[To fill]
+
+### Reference label uncertainty
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# reference labels, for total count and uncertainty of class categories
+ref_labels <- st_read(
+  system.file("extdata/map_reference_labels.geojson", 
+              package = "activemapper"), quiet = TRUE
+)
+class_ct <- as_tibble(ref_labels) %>% group_by(category) %>% count()
+ref_ct <- as_tibble(ref_labels) %>% group_by(worker_id) %>% count() %>% pull(n)
+
+crp_ct <- class_ct %>% filter(category %in% c("annualcropland", "unsure1"))
+noncrp_ct <- class_ct %>% filter(category %in% c("noncropland", "unsure2"))
+
+# observer agreement/correlation
+ref_labels_joint <- st_read(
+  system.file("extdata/map_reference_labels_joint.geojson", 
+              package = "activemapper"), quiet = TRUE
+)
+joint_classes <- ref_labels_joint %>% as_tibble() %>% 
+  select(name, category, worker_id) %>% 
+  arrange(name) %>% 
+  mutate(class = case_when(
+    category == "annualcropland" | category == "unsure1" ~ 1, 
+    category == "noncropland" | category == "unsure2" ~ 0 
+  ))
+class_mat <- joint_classes %>% dplyr::select(name, class, worker_id) %>% 
+  tidyr::pivot_wider(names_from = worker_id, values_from = class) %>% 
+  dplyr::select(-name) %>% as.matrix()
+
+usertab <- table("User1" = class_mat[, 1], "User2" = class_mat[, 2])
+
+agreement <- round(sum(diag(usertab)) / sum(usertab) * 100, 1)
+observer_r <- cor.test(class_mat[, 1], class_mat[, 2], method = "spearman")
+
+```
+
+The total validation sample was `r sum(ref_ct)`, with `r ref_ct[2]` samples collected and interpreted by one observer (Su Ye) and `r ref_ct[1]` collected by a second observer (Estes).
+To evaluate the uncertainty inherent in defining the map reference labels, the pair both mapped `r sum(usertab)` sites, showing an overall level of agreement of `r agreement`%, and a Spearman Rank Correlation of `r round(unname(observer_r$estimate), 2)`.
+
+Although this overlap between observers was limited, the map reference classification scheme provided two additional measures of uncertainty, which were classes defined as "unsure but most likely a field" or "unsure but most likely not a field", which the reference labeller would choose when they could not with high confidence state whether the site was either cropland or not cropland.
+Of the `r sum(crp_ct$n)` map reference samples that were identified as cropland, `r round(crp_ct$n[2] / sum(crp_ct$n) * 100, 1)`% fell into the lower confidence category, as did `r round(noncrp_ct$n[2] / sum(noncrp_ct$n) * 100, 1)`% of the `r sum(noncrp_ct$n)` non-cropland samples.
+Across both classes, the reference labellers had lower confidence in `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites labelled.
+
+### Categorical accuracy
+In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{tab:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B) and \ref{fig:aezaccuracy}).
+
+```{r aoizones, echo = FALSE, out.width="100%", fig.cap="Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy.
+", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_aoi_zones.png')
+```
+
+```{r aezaccuracy, echo = FALSE, out.width="100%", fig.cap="Map accuracies and adjusted area estimates within four different groupings of Ghana's agroecozones", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_aez_accuracies.png')
+```
 
@@
 
-```{r numbervalidation, echo = FALSE, out.width="100%", fig.cap="The distributions of the average number of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average number of digitized polygons (shown in red) by the most accurate worker to label each site.
+```{r numbervalidation, echo = FALSE, out.width="90%", fig.cap="The distributions of the average number of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average number of digitized polygons (shown in red) by the most accurate worker to label each site.
 The mean and median of each distribution is shown.

### 2021-02-17 data-raw/error_matrices.R
Fixes to error matrices (removed some samples by accident)

@@ -19,9 +19,9 @@ aez <- read_sf(
 ) %>% select(ID, Name) %>% rename(aezid = ID, aez = Name)
 
 # Pick up AEZ intersections
-ref_int_segments <- st_intersection(ref_int_segments, aez) %>%
+ref_int_segments <- st_join(ref_int_segments, aez) %>%
   select(name, tile, aoi, aezid, aez, category, afrac)
-ref_int_segments %>% as_tibble() %>% group_by(aoi) %>% count()
+# ref_int_segments %>% as_tibble() %>% group_by(aoi) %>% count()
 
 # set up data for error matrix
 error_matrix_data <- as_tibble(ref_int_segments) %>%
@@ -51,6 +51,7 @@ error_matrix <- function(pred, ref) {
 # plot(st_geometry(st_centroid(ref_int_segments)), add = TRUE)
 
 # calculate error matrices by AOI
+
 error_matrices_aoi <- lapply(unique(error_matrix_data$aoi), function(x) {
   dat <- error_matrix_data %>% filter(aoi == x)
   error_matrix_rf <- error_matrix(dat$prob_class, dat$ref)
@@ -90,4 +91,4 @@ error_matrix_ghana <- list("rf" = error_matrix_rf, "seg" = error_matrix_seg)
 # Outputs
 error_matrices <- list("aoi" = error_matrices_aoi, "aez" = error_matrices_aez,
                        "ghana" = error_matrix_ghana)
-usethis::use_data(error_matrices)
+usethis::use_data(error_matrices, overwrite = TRUE)

### 2021-02-22 inst/manuscript/figures.Rmd
Edits for average field size text (SI, ms), and regression kriging adjustments (not to be retained), as well as new tile-wise presentation (figures).

@@
 
-### Figure 7: Categorical map accuracy
+### Table 2: Categorical map accuracy
 Place map accuracy tables and zone report into single figure, for AOI zones.
@@
 # gp <- t1 + m
-ggsave(t1, filename = pth("figures/figure7.png"), width = 7, height = 7.9, 
+ggsave(t1, filename = pth("figures/table2.png"), width = 7, height = 7.9, 
        units = "in", dpi = 300)
@@
 
-### Figure 8: Cropland maps
+### Figure 7: Cropland maps
 ```{r, eval=FALSE}
@@
 
-ggsave(filename = pth("figures/figure8.png"), 
+ggsave(filename = pth("figures/figure7.png"), 
        plot = inset_map, bg = "transparent", width = diff(maprange[1:2]), 
@@
 
+```
+
+### Table 3
+```{r, eval = FALSE}
+fsizes_n_05 <- st_read(
+  system.file("extdata/field_size_n_05.geojson", package = "activemapper")
+) %>% mutate(
+  muarea = ifelse(is.na(muarea), 0, muarea),
+  nflds = ifelse(is.na(nflds), 0, nflds),
+  aoi = as.character(aoi)
+)
+
+# calculate bias factors for adjustment
+data("segment_quality_stats")
+segment_quality_stats %>% filter(measure == "n")
+biases <- segment_quality_stats %>%
+  filter(Stat == "Mean" & scale == "grid") %>% 
+  pivot_wider(names_from = type, values_from = value) %>% 
+  mutate(bias = Validation - Segmentation, 
+         biasfactor = Validation / Segmentation)
+biases <- as_tibble(tiles) %>% group_by(aoi1) %>% 
+  count() %>% ungroup() %>% 
+  rename(aoi = aoi1, ntiles = n) %>% 
+  mutate(aoi = as.character(aoi), ncells = ntiles * 100) %>% 
+  left_join(biases, .) %>% 
+  select(-scale, -Stat)
+# biases %>% filter(measure == "n")
+# tiles %>% slice(1) %>% raster::extract(mgridr, .)
+# st_area(aois) / 10000 / 100
+
+# apply bias adjustments
+# field area first
+fsizes_n_05_adj <- biases %>% filter(measure == "area" & aoi != "All") %>% 
+  select(aoi, biasfactor) %>% 
+  rename(fbias = biasfactor) %>% 
+  left_join(fsizes_n_05, .) %>% 
+  mutate(muarea_adj = muarea * fbias)
+
+# then field count
+fsizes_n_05_adj <- biases %>% filter(measure == "n" & aoi != "All") %>% 
+  select(aoi, biasfactor, ntiles, ncells) %>% 
+  rename(nbias = biasfactor) %>% 
+  left_join(fsizes_n_05_adj, .) %>% 
+  mutate(nflds_adj = nflds * nbias) %>% 
+  select(aoi, tile, muarea, muarea_adj, fbias, nflds, nflds_adj, nbias, ntiles)
+
+# hist(fsizes_n_05$muarea)
+brks <- c(-1, 0.001, 1, 2, 4, 6, 10, 15, 25, 50)
+labs <- c("0", "<1", "1-2", "2-4", "4-6", "6-10", "10-15", "15-25", "25-50")
+labsv <- c(1:(length(labs)))
+fsizes_n_05_adj <- fsizes_n_05_adj %>% 
+  mutate(
+    Area = as.numeric(cut(muarea, breaks = brks, labels = labsv)), 
+    Area_adj = as.numeric(cut(muarea_adj, breaks = brks, labels = labsv))
+  )
+
+nbrks <- c(-1, 0.001, 50, 100, 200, 300, 400, 500, 700, 1000)
+nlabs <- c("0", "<50", "50-100", "100-200", "200-300", "300-400", "400-500",
+           "500-700", "700-1000")
+nlabsv <- c(1:(length(labs)))
+fsizes_n_05_adj <- fsizes_n_05_adj %>% 
+  mutate(
+    N = as.numeric(cut(nflds, breaks = nbrks, labels = nlabsv)), 
+    N_adj = as.numeric(cut(nflds_adj, breaks = nbrks, labels = nlabsv))
+  )
+
+muarea <- st_rasterize(fsizes_n_05_adj["Area"])
+muarea_adj <- st_rasterize(fsizes_n_05_adj["Area_adj"])
+nflds <- st_rasterize(fsizes_n_05_adj["N"])
+nflds_adj <- st_rasterize(fsizes_n_05_adj["N_adj"])
+
+# plot(muarea)
+# plot(nflds)
+# plot(is.na(nflds))
+# 
+# fsizes_n_05 %>% filter(is.na(nflds))
+# fsizes_n_05 %>% filter(is.infinite(nflds))
+# plot(fsizes_n_05 %>% filter(is.na(nflds)))
+thm <- function(x) {
+  theme_void() + 
+    theme(legend.position = c(0.7, 0.07), legend.direction = "horizontal", 
+          legend.key.height = unit(0.2, "cm"), 
+          legend.text = element_text(colour = "black"),
+          legend.title = element_text(size = 9, colour = "black"))
+}
+
+leg <- c(labs, "")
+xylims <- st_bbox(ghana)
+p <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = muarea, aes(x = x, y = y, fill = factor(Area))) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+padj <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = muarea_adj, aes(x = x, y = y, fill = factor(Area_adj))) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+
+leg <- c(nlabs, "")
+p2 <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = nflds, aes(x = x, y = y, fill = factor(N))) +
+  # geom_stars(data = nflds) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+  # scale_fill_viridis_c(name = "N", na.value = "transparent") +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+p2adj <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = nflds_adj, aes(x = x, y = y, fill = factor(N_adj))) +
+  # geom_stars(data = nflds) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+  # scale_fill_viridis_c(name = "N", na.value = "transparent") +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+
+p2adj
+
+op <- p + p2 + plot_annotation(tag_levels = 'A')
+
+ggsave(op, bg = "transparent", width = 6, height = 4, units = "in", dpi = 300, 
+       filename = pth("figures/si_fsize_n_05.png"))
 ```
@@
   geom_sf(data = aez_zones, lwd = 0.5, fill = "transparent") + 
+  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
+               aes(label = ID), 
+               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
+               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
+               size = 3, col = "grey60") + 
   geom_sf_text(data = st_centroid(aez_zones), aes(label = zone), size = 4) + 
   theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
 ggsave(m + m2 + plot_annotation(tag_levels = 'A'), 
@@
 
-### Figure S10: AEZ accuracies
+### Table S4: AEZ accuracies
 
@@
 
-### Figure S11: Segmentation statistics (grid scale)
-
-[Check whether the means need to be weighted or not]
-[Check whether need to redo median]
+### Figures S10 and S11: Segmentation statistics (grid scale)
+
 ```{r segstats, eval=FALSE}
@@
        width = 7, height = 7, units = "in", dpi = 300)
+
+```
+
+### Figure S12: Average fields sizes
+
+Tile scale field size and number, presented without adjustment
+
+```{r, eval=FALSE}
+fsizes_n_05 <- st_read(
+  system.file("extdata/field_size_n_05.geojson", package = "activemapper")
+) %>% mutate(
+  muarea = ifelse(is.na(muarea), 0, muarea),
+  nflds = ifelse(is.na(nflds), 0, nflds),
+  aoi = as.character(aoi)
+)
+
+# # calculate bias factors for adjustment
+# data("segment_quality_stats")
+# segment_quality_stats %>% filter(measure == "n")
+# biases <- segment_quality_stats %>%
+#   filter(Stat == "Mean" & scale == "grid") %>% 
+#   pivot_wider(names_from = type, values_from = value) %>% 
+#   mutate(bias = Validation - Segmentation, 
+#          biasfactor = Validation / Segmentation)
+# biases <- as_tibble(tiles) %>% group_by(aoi1) %>% 
+#   count() %>% ungroup() %>% 
+#   rename(aoi = aoi1, ntiles = n) %>% 
+#   mutate(aoi = as.character(aoi), ncells = ntiles * 100) %>% 
+#   left_join(biases, .) %>% 
+#   select(-scale, -Stat)
+# # biases %>% filter(measure == "n")
+# # tiles %>% slice(1) %>% raster::extract(mgridr, .)
+# # st_area(aois) / 10000 / 100
+# 
+# # apply bias adjustments
+# # field area first
+# fsizes_n_05_adj <- biases %>% filter(measure == "area" & aoi != "All") %>% 
+#   select(aoi, biasfactor) %>% 
+#   rename(fbias = biasfactor) %>% 
+#   left_join(fsizes_n_05, .) %>% 
+#   mutate(muarea_adj = muarea * fbias) 
+# 
+# # then field count
+# fsizes_n_05_adj <- biases %>% filter(measure == "n" & aoi != "All") %>% 
+#   select(aoi, biasfactor, ntiles, ncells) %>% 
+#   rename(nbias = biasfactor) %>% 
+#   left_join(fsizes_n_05_adj, .) %>% 
+#   mutate(nflds_adj = nflds * nbias) %>% 
+#   select(aoi, tile, muarea, muarea_adj, fbias, nflds, nflds_adj, nbias, ntiles)
+# 
+# 
+# # hist(fsizes_n_05$muarea)
+# brks <- c(-1, 0.001, 1, 2, 4, 6, 10, 15, 25, 50)
+# labs <- c("0", "<1", "1-2", "2-4", "4-6", "6-10", "10-15", "15-25", "25-50")
+# labsv <- c(1:(length(labs)))
+# fsizes_n_05_adj <- fsizes_n_05_adj %>% 
+#   mutate(
+#     Area = as.numeric(cut(muarea, breaks = brks, labels = labsv)), 
+#     Area_adj = as.numeric(cut(muarea_adj, breaks = brks, labels = labsv))
+#   )
+# 
+# nbrks <- c(-1, 0.001, 50, 100, 200, 300, 400, 500, 700, 1000)
+# nlabs <- c("0", "<50", "50-100", "100-200", "200-300", "300-400", "400-500",
+#            "500-700", "700-1000")
+# nlabsv <- c(1:(length(labs)))
+# fsizes_n_05_adj <- fsizes_n_05_adj %>% 
+#   mutate(
+#     N = as.numeric(cut(nflds, breaks = nbrks, labels = nlabsv)), 
+#     N_adj = as.numeric(cut(nflds_adj, breaks = nbrks, labels = nlabsv))
+#   )
+# 
+# # stars
+# muarea <- st_rasterize(fsizes_n_05_adj["Area"])
+# muarea_adj <- st_rasterize(fsizes_n_05_adj["Area_adj"])
+# nflds <- st_rasterize(fsizes_n_05_adj["N"])
+# nflds_adj <- st_rasterize(fsizes_n_05_adj["N_adj"])
+
+brks <- c(-1, 0.001, 1, 2, 4, 6, 10, 15, 25, 50)
+labs <- c("0", "<1", "1-2", "2-4", "4-6", "6-10", "10-15", "15-25", "25-50")
+labsv <- c(1:(length(labs)))
+fsizes_n_05 <- fsizes_n_05 %>% 
+  mutate(Area = as.numeric(cut(muarea, breaks = brks, labels = labsv)))
+
+# hist(fsizes_n_05$nflds)
+nbrks <- c(-1, 0.001, 50, 100, 200, 300, 400, 500)
+nlabs <- c("0", "<50", "50-100", "100-200", "200-300", "300-400", "400-500")
+nlabsv <- c(1:(length(nlabs)))
+fsizes_n_05 <- fsizes_n_05 %>% 
+  mutate(N = as.numeric(cut(nflds, breaks = nbrks, labels = nlabsv)))
+
+muarea <- st_rasterize(fsizes_n_05["Area"])
+nflds <- st_rasterize(fsizes_n_05["N"])
+
+thm <- function(x) {
+  theme_void() + 
+    theme(#legend.position = c(0.7, 0.07), 
+          legend.direction = "vertical", 
+          legend.key.height = unit(0.2, "cm"), 
+          legend.text = element_text(colour = "black"),
+          legend.title = element_text(size = 9, colour = "black"))
+}
+
+leg <- c(labs, "")
+xylims <- st_bbox(ghana)
+p <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = muarea, aes(x = x, y = y, fill = factor(Area))) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+# padj <- ggplot(africa) + 
+#   geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+#   geom_stars(data = muarea_adj, aes(x = x, y = y, fill = factor(Area_adj))) +
+#   geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+#   scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+#   coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+
+leg <- c(nlabs, "")
+p2 <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = nflds, aes(x = x, y = y, fill = factor(N))) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_d(name = "N", na.value = "transparent", labels = leg) +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+# p2adj <- ggplot(africa) + 
+#   geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+#   geom_stars(data = nflds_adj, aes(x = x, y = y, fill = factor(N_adj))) +
+#   geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+#   scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
+#   coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
+
+op <- p + p2 + plot_annotation(tag_levels = 'A')
+
+ggsave(op, bg = "transparent", width = 7, height = 4, units = "in", dpi = 300, 
+       filename = pth("figures/si_fsize_n_05.png"))
+```
+
+Both original as derived from segmentations and adjusted using regression kriging.
+
+```{r, eval=FALSE}
+fsizes <- read_stars(
+  system.file("extdata/mean_field_size005.tif", package = "activemapper")
+)
+fsizes_adj <- read_stars(
+  system.file("extdata/mean_field_size005_adj.tif", package = "activemapper")
+)
+
+# Plots
+xylims <- st_bbox(ghana)
+p <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = fsizes) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_c(name = "ha", na.value = "transparent") +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) +
+  theme_void() + 
+  theme(legend.position = c(0.7, 0.07), legend.direction = "horizontal", 
+        legend.key.height = unit(0.2, "cm"), 
+        legend.text = element_text(colour = "black"),
+        legend.title = element_text(size = 9, colour = "black"))
+
+p2 <- ggplot(africa) + 
+  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_stars(data = fsizes_adj[, , , 1]) +
+  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
+  scale_fill_viridis_c(name = "ha", na.value = "transparent") +
+  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) +
+  theme_void() + 
+  theme(legend.position = c(0.7, 0.07), legend.direction = "horizontal", 
+        legend.key.height = unit(0.2, "cm"), 
+        legend.text = element_text(colour = "black"),
+        legend.title = element_text(size = 9, colour = "black"))
+
+op <- p + p2 + plot_annotation(tag_levels = 'A')
+
+ggsave(op, bg = "transparent", width = 6, height = 4, units = "in", dpi = 300, 
+       filename = pth("figures/si_fsizes.png"))
 

### 2021-02-22 inst/manuscript/manuscript.Rmd
Additional accuracy metrics added to SI.
Edits for average field size text (SI, ms), and regression kriging adjustments (not to be retained), as well as new tile-wise presentation (figures).

@@
 ---
-title: "Improving cropland maps through tight integration of human and machine intelligence"
+title: "Mapping the characteristics of small-scale agricultural systems at national extents using small sat data and an active learning approach"
 
@@
 
+- name: Boka Luo
+  affilnum: '1'
+
 - name: Ron Eastman
-  affilnum: '1'
-
-- name: Qi Zhang
-  affilnum: '1'
-
-- name: Boka Luo
   affilnum: '1'
@@
 ### Categorical accuracy
-To assess the overall accuracy of the final maps, we used our map reference sample to evaluate the accuracy of classifications derived from thresholding the posterior probabilities produced by the RandomForest model, as well as the segmented field boundary maps.
-First, we evaluated the inter-rater agreement between reference supervisors in labeling reference polygons.
-The average inter-rater agreement was **X%, with a standard deviation of Y% (see Figure X in SI)**, which provides the upper bound on achievable map accuracy.
-
-Based on this reference sample, we found that the average accuracy of the pixel-wise classifications was **X, with a range of X-X** (Figure 5).
-Overall user's accuracy was slightly **lower/higher** at **X, (range X-X)**, while producer's accuracy averaged **X, (range X-X)**.
-Each accuracy measure was highest in **northern-northeastern 2/3** of the country, and lowest in **southwestern 1/3**, where rainfall and cloud-cover are highest, and where tree crops dominate.
-The segmentation process **increased/decreased** accuracy slightly, with a mean overall, user's and producer's accuracy of **X (range X-X), Y (range X-X), and Z (range X-X) percent**.
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# reference labels, for total count and uncertainty of class categories
+ref_labels <- st_read(
+  system.file("extdata/map_reference_labels.geojson", 
+              package = "activemapper"), quiet = TRUE
+)
+class_ct <- as_tibble(ref_labels) %>% group_by(category) %>% count()
+ref_ct <- as_tibble(ref_labels) %>% group_by(worker_id) %>% count() %>% pull(n)
+
+crp_ct <- class_ct %>% filter(category %in% c("annualcropland", "unsure1"))
+noncrp_ct <- class_ct %>% filter(category %in% c("noncropland", "unsure2"))
+
+# observer agreement/correlation
+ref_labels_joint <- st_read(
+  system.file("extdata/map_reference_labels_joint.geojson", 
+              package = "activemapper"), quiet = TRUE
+)
+joint_classes <- ref_labels_joint %>% as_tibble() %>% 
+  select(name, category, worker_id) %>% 
+  arrange(name) %>% 
+  mutate(class = case_when(
+    category == "annualcropland" | category == "unsure1" ~ 1, 
+    category == "noncropland" | category == "unsure2" ~ 0 
+  ))
+class_mat <- joint_classes %>% dplyr::select(name, class, worker_id) %>% 
+  tidyr::pivot_wider(names_from = worker_id, values_from = class) %>% 
+  dplyr::select(-name) %>% as.matrix()
+
+usertab <- table("User1" = class_mat[, 1], "User2" = class_mat[, 2])
+
+agreement <- round(sum(diag(usertab)) / sum(usertab) * 100, 1)
+observer_r <- cor.test(class_mat[, 1], class_mat[, 2], method = "spearman")
+
+```
+
+To assess the overall accuracy of the final maps, we used our map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of 1) pixel-wise classifications produced by the RandomForest model, and 2) the segmented field boundary maps.
+We first evaluated the uncertainty in the map reference classes using two approaches, the first being an assessment of the overall agreement between map reference labels (at `r sum(usertab)` sites), and the second was the level of confidence in assigned labels (see SI for details).
+The first measure showed that the two individual labellers agreed `r agreement`% of sites, while the second showed that `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites labelled were of lower confidence.
+
+Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table 2).
+This overall level of accuracy varied within four smaller zones (Figure S9A), ranging from 83.3% in Zone 1 (comprised of AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
+Examining the accuracies of the individual classes, the Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1.
+The User's accuracy for cropland was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1.
+Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%, with the lowest accuracies for this class found in Zone 1 (Producer's = 89.3% and User's = 87.7%, respectively).
+
+The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table 2).
+In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map, indicating that the segmentation step helped reduce omission error, but substantially increased false positive errors.
 
@@
 
-The accuracy patterns generally align with the difficulty of the labelling task, that is, **the southwestern regions were hardest for workers to interpret, as shown by Figure 5.**
+\begin{table}
+\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions; top 5 rows) and the segmented map (bottom 5 rows).
+Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country.
+The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.
+}
+\includegraphics[width = 18cm]{figures/table2.png}
+\label{tab:aezaccuracy}
+\end{table}
 
@@
 ## Ghana's croplands
-Two separate maps of cropland were produced for each AOI, one derived from the cropland probability map, and the other from the segmented field boundaries.
-
-Map figure:
-- Panel A: Cropland probability map (0.005 degree resolution)
-  - Panel A insets: Probability close-ups in AOIs 1, 8, 13, 15
-- Panel B: Cropland percentage map (0.005 degree)
-  - Panel B insets: Segment close-ups in AOIs 1, 8, 13, 15
-- Panel C: Median field size
-- Panel D: Field number
-- Median field size
-- Cropland percentage
-
-Stats figure: 
-- Mean field size per AOI
-- Median field size per AOI
-- SD fields per AOI 
-- N field per AOI
+Two separate maps of cropland were produced for each AOI, the per-pixel classified from the cropland probabilities, and the vectorized map of field boundaries (Figure \ref{fig:mainmap}).
+The former map provides the more accurate assessment of the distribution of croplands in Ghana, which are most concentrated in the Southeastern corner (AOI 15), the central-western region (AOI 7, the northeastern and northwestern corners of AOIs 10 and 11, and the south of AOI 8), the northeastern quadrant stretching from AOI 9 through 5 and 6 up to 2 and 3.
+The northern third of AOI 1 also has noticeable densities of cropland.
+Several noteable areas of low cropland density are also apparent, indicating the presence of large protected areas, such as Mole National Park in the the southeastern corner of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
+In contrast, the relative absence of cropland in AOIs 13, 14, and 16 do not reflect the scarcity of agriculture, but rather the predominance of tree-based agriculture in the forested agroecoregions of Ghana, which was not mapped.
 
@@
 Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/figure8.png')
-```
+knitr::include_graphics('figures/figure7.png')
+```
+
+Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of arable croplands in Ghana.
+The cropland extent estimated from the vectorized map is 42,359 km$^2$ (with a margin of error of 4,395 km$^2$), or `r round(42359 / (42359 + 204940) * 100, 1)`% of the mapped area, the estimate based on the per pixel map is 43,233 km$^2$ (margin of error = 4,904 km$^2$), or `r round(43233 / (43233 + 202856) * 100, 1)`% of area.
+
+The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure 7's lower right inset) to more densely packed, less distinctly shaped fields in area such as AOI 5 (upper right inset in Figure 7).
+These characteristics can be summarized and mapped to explore how such characteristics vary geographically For example, average fields sizes were mapped at 0.005 degree resolution (Figure S12A), providing some indication of countrywide patterns.
+Given the over-estimation of fields sizes relative to the validation labels, particularly in several savanna-zone AOIs (2, 3, 6, 7, and 15), the mean sizes likely reflect agglomerations of contiguous smallholder fields, such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).
+A closer approximation of individual field boundaries can be mapped using approaches such as regression kriging to model the relationship between the segmented boundaries and the hand drawn labels collected at validation points (Figure S12B), although such an approach introduces spatial artifacts.
 
 # Discussion
+We present here a method for developing high resolution maps that detail the characteristics of smallholder-dominated croplands, providing information on their geometries and distribution of <5 m resolution, and representing a single agricultural year.
+
+The results
+
 - Label quality
@@
   - Image-interpreted versus ground labels: **However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but may instead be at least partially an indication of how different the perspectives are as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
+
+Compare values
 

### 2021-02-22 inst/manuscript/si.Rmd
Edits for average field size text (SI, ms), and regression kriging adjustments (not to be retained), as well as new tile-wise presentation (figures).

@@
 opts_chunk$set(fig.path = "output/figures/")
+
+knitr::opts_chunk$set(fig.pos = "!ht", out.extra = "")
 
@@
 The total validation sample was `r sum(ref_ct)`, with `r ref_ct[2]` samples collected and interpreted by one observer (Su Ye) and `r ref_ct[1]` collected by a second observer (Estes).
-To evaluate the uncertainty inherent in defining the map reference labels, the pair both mapped `r sum(usertab)` sites, showing an overall level of agreement of `r agreement`%, and a Spearman Rank Correlation of `r round(unname(observer_r$estimate), 2)`.
+To evaluate the uncertainty inherent in defining the map reference labels, the pair mapped `r sum(usertab)` common sites, showing an overall level of agreement of `r agreement`%, and a Spearman Rank Correlation of `r round(unname(observer_r$estimate), 2)`.
 
@@
 ### Categorical accuracy
-In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{tab:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B) and \ref{fig:aezaccuracy}).
+In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).
 
 ```{r aoizones, echo = FALSE, out.width="100%", fig.cap="Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy.
-", fig.align='center', message=FALSE, results='hold'}
+In B, the ID number of Ghana's individual agroecozones is shown in grey: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE, results='hold'}
 knitr::include_graphics('figures/si_aoi_zones.png')
@@
 
-```{r aezaccuracy, echo = FALSE, out.width="100%", fig.cap="Map accuracies and adjusted area estimates within four different groupings of Ghana's agroecozones", fig.align='center', message=FALSE, results='hold'}
-knitr::include_graphics('figures/si_aez_accuracies.png')
-```
+\begin{table}
+\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions).
+Results are provided for four different groupings of Ghana's 8 agroecozones zones (Zone 1 = Coastal savanna; Zone 2 = Wet evergreen, Moist evergreen, and Deciduous forest; Zone 3 = Transitional zone; Zone 4 = Guinea savanna and Sudan savanna) plus the entire country.
+The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.}
+\includegraphics[width = 18cm]{figures/si_aez_accuracies.png}
+\label{tab:aezaccuracy}
+\end{table}
 
@@
 knitr::include_graphics('figures/si_validation_stats_fn2.png')
+```
+
+## Cropland characteristics
+### Average field size
+
+To examine average field sizes, the average area of the vectorized field boundaries intersecting each 0.005 degree cell in the mapping grid was calculated and mapped back onto the grid (Figure \ref{fig:fieldsize}A).
+Given the overestimation bias inherent in the vectorized boundaries relative to the hand-digitized validation labels, we also used a regression kriging approach [@henglRegressionkrigingEquationsCase2007] to these average field sizes for this bias.
+To do this, we used the estimated mean field sizes extracted from the map in Figure S12A to predict the average polygon size of the most accurate mapper's labels for each of the 1600 validation sites, using an ordinary least squares regression to model the relationship, after first log-transforming both variables.
+We then applied ordinary kriging (using an exponential model for the variogram) to the regression residuals, which we added to the surface of the predicted field sizes to create the adjusted surface, and then converted this back to units of hectares (Figure \ref{fig:fieldsize}B).
+
+```{r fieldsizes, echo = FALSE, out.width="100%", fig.cap="Average fields sizes in Ghana at 0.005 degree resolution, as calculated from A) the vectorized field boundaries, and after B) adjustment against hand-drawn validation labels using a regression-kriging approach.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_fsizes.png')
 ```

### 2021-02-22 data-raw/cropcover_adjusted.R

@@ -1,0 +1,206 @@
+# Kriging to get average field size and predicted probabilities.
+# For calculating adjusted field statistics with calculating error metrics.
+
+library(activemapper)
+library(raster)
+library(sf)
+library(dplyr)
+library(ggplot2)
+library(gstat)
+library(sp)
+
+# load in data set
+load(system.file("extdata/top_label_data.rda", package = "activemapper"))
+load(system.file("extdata/field_validation_stats.rda",
+                 package = "activemapper"))
+load(system.file("extdata/segment_validation_data.rda",
+                 package = "activemapper"))
+tiles <- st_read(
+  system.file("extdata/ghana_tiles.geojson", package = "activemapper")
+)
+fsizes <- raster(
+  system.file("extdata/mean_field_size005.tif", package = "activemapper")
+)
+
+# albers ESRI:102022
+aea <- glue::glue("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25 +x_0=0",
+                  " +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
+
+# label data from validation sites and segments
+vlabs <- top_label_data$stats %>% select(aoi, name, x, y, Mean, n) %>%
+  rename(vmu = Mean, vn = n)
+segs <- field_validation_stats %>% select(aoi, name, Mean, n) %>%
+  rename(smu = Mean, sn = n)
+
+# hist(vlabs$vmu)
+# hist(segs$smu)
+val_segs <- left_join(vlabs, segs, by = c("name", "aoi")) %>%
+  arrange(aoi, name) %>% mutate(dist = NA)
+
+# val_segs %>% filter(vmu > 0 & smu > 0) %>%
+#   st_as_sf(coords = c("x", "y")) %>% st_geometry() %>%
+#   plot(pch = 16, cex = 0.5)
+
+val_segs_replaced <- lapply(c("vmu", "smu"), function(x) {  # x <- "vmu"
+  mu0 <- val_segs %>% filter((!!sym(x)) == 0) %>%
+    st_as_sf(coords = c("x", "y"), crs = 4326)
+  munot0 <- val_segs %>% filter((!!sym(x)) > 0) %>%
+    st_as_sf(coords = c("x", "y"), crs = 4326)
+
[…]

### 2021-02-22 data-raw/top_label_data.R

@@ -95,4 +95,8 @@ top_label_data <- list(
 
 # usethis::use_data(top_label_data, overwrite = TRUE)
 save(top_label_data, file = here::here("inst/extdata/top_label_data.rda"))
+f <- here::here(
+  "external/data/aois/labels/validation_labels/top_validation_labels.geojson"
+)
+st_write(top_label_data$maps, dsn = f)
 

### 2021-02-23 inst/manuscript/figures.Rmd
Updated manuscript, SI, figures with text and table, figure for field size and numbers, which is final part of results

@@
 
-```
-
-### Table 3
-```{r, eval = FALSE}
-fsizes_n_05 <- st_read(
-  system.file("extdata/field_size_n_05.geojson", package = "activemapper")
-) %>% mutate(
-  muarea = ifelse(is.na(muarea), 0, muarea),
-  nflds = ifelse(is.na(nflds), 0, nflds),
-  aoi = as.character(aoi)
-)
-
-# calculate bias factors for adjustment
-data("segment_quality_stats")
-segment_quality_stats %>% filter(measure == "n")
-biases <- segment_quality_stats %>%
-  filter(Stat == "Mean" & scale == "grid") %>% 
-  pivot_wider(names_from = type, values_from = value) %>% 
-  mutate(bias = Validation - Segmentation, 
-         biasfactor = Validation / Segmentation)
-biases <- as_tibble(tiles) %>% group_by(aoi1) %>% 
-  count() %>% ungroup() %>% 
-  rename(aoi = aoi1, ntiles = n) %>% 
-  mutate(aoi = as.character(aoi), ncells = ntiles * 100) %>% 
-  left_join(biases, .) %>% 
-  select(-scale, -Stat)
-# biases %>% filter(measure == "n")
-# tiles %>% slice(1) %>% raster::extract(mgridr, .)
-# st_area(aois) / 10000 / 100
-
-# apply bias adjustments
-# field area first
-fsizes_n_05_adj <- biases %>% filter(measure == "area" & aoi != "All") %>% 
-  select(aoi, biasfactor) %>% 
-  rename(fbias = biasfactor) %>% 
-  left_join(fsizes_n_05, .) %>% 
-  mutate(muarea_adj = muarea * fbias)
-
-# then field count
-fsizes_n_05_adj <- biases %>% filter(measure == "n" & aoi != "All") %>% 
-  select(aoi, biasfactor, ntiles, ncells) %>% 
-  rename(nbias = biasfactor) %>% 
-  left_join(fsizes_n_05_adj, .) %>% 
-  mutate(nflds_adj = nflds * nbias) %>% 
-  select(aoi, tile, muarea, muarea_adj, fbias, nflds, nflds_adj, nbias, ntiles)
-
-# hist(fsizes_n_05$muarea)
-brks <- c(-1, 0.001, 1, 2, 4, 6, 10, 15, 25, 50)
-labs <- c("0", "<1", "1-2", "2-4", "4-6", "6-10", "10-15", "15-25", "25-50")
-labsv <- c(1:(length(labs)))
-fsizes_n_05_adj <- fsizes_n_05_adj %>% 
-  mutate(
-    Area = as.numeric(cut(muarea, breaks = brks, labels = labsv)), 
-    Area_adj = as.numeric(cut(muarea_adj, breaks = brks, labels = labsv))
-  )
-
-nbrks <- c(-1, 0.001, 50, 100, 200, 300, 400, 500, 700, 1000)
-nlabs <- c("0", "<50", "50-100", "100-200", "200-300", "300-400", "400-500",
-           "500-700", "700-1000")
-nlabsv <- c(1:(length(labs)))
-fsizes_n_05_adj <- fsizes_n_05_adj %>% 
-  mutate(
-    N = as.numeric(cut(nflds, breaks = nbrks, labels = nlabsv)), 
-    N_adj = as.numeric(cut(nflds_adj, breaks = nbrks, labels = nlabsv))
-  )
-
-muarea <- st_rasterize(fsizes_n_05_adj["Area"])
-muarea_adj <- st_rasterize(fsizes_n_05_adj["Area_adj"])
-nflds <- st_rasterize(fsizes_n_05_adj["N"])
-nflds_adj <- st_rasterize(fsizes_n_05_adj["N_adj"])
-
-# plot(muarea)
-# plot(nflds)
-# plot(is.na(nflds))
-# 
-# fsizes_n_05 %>% filter(is.na(nflds))
-# fsizes_n_05 %>% filter(is.infinite(nflds))
-# plot(fsizes_n_05 %>% filter(is.na(nflds)))
-thm <- function(x) {
-  theme_void() + 
-    theme(legend.position = c(0.7, 0.07), legend.direction = "horizontal", 
-          legend.key.height = unit(0.2, "cm"), 
-          legend.text = element_text(colour = "black"),
-          legend.title = element_text(size = 9, colour = "black"))
-}
-
-leg <- c(labs, "")
-xylims <- st_bbox(ghana)
-p <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = muarea, aes(x = x, y = y, fill = factor(Area))) +
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
-  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
-padj <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = muarea_adj, aes(x = x, y = y, fill = factor(Area_adj))) +
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
-  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
-
-leg <- c(nlabs, "")
-p2 <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = nflds, aes(x = x, y = y, fill = factor(N))) +
-  # geom_stars(data = nflds) +
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
-  # scale_fill_viridis_c(name = "N", na.value = "transparent") +
-  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
-p2adj <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = nflds_adj, aes(x = x, y = y, fill = factor(N_adj))) +
-  # geom_stars(data = nflds) +
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  scale_fill_viridis_d(name = "ha", na.value = "transparent", labels = leg) +
-  # scale_fill_viridis_c(name = "N", na.value = "transparent") +
-  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) + thm()
-
-p2adj
-
-op <- p + p2 + plot_annotation(tag_levels = 'A')
-
-ggsave(op, bg = "transparent", width = 6, height = 4, units = "in", dpi = 300, 
-       filename = pth("figures/si_fsize_n_05.png"))
 ```

### 2021-02-23 inst/manuscript/manuscript.Rmd
Updated manuscript, SI, figures with text and table, figure for field size and numbers, which is final part of results

@@
 ```{r setup, include=FALSE, cache=FALSE, message = FALSE}
-
 library(knitr)
@@
 library(activemapper)
+library(kableExtra)
 
@@
 
-The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure 7's lower right inset) to more densely packed, less distinctly shaped fields in area such as AOI 5 (upper right inset in Figure 7).
-These characteristics can be summarized and mapped to explore how such characteristics vary geographically For example, average fields sizes were mapped at 0.005 degree resolution (Figure S12A), providing some indication of countrywide patterns.
-Given the over-estimation of fields sizes relative to the validation labels, particularly in several savanna-zone AOIs (2, 3, 6, 7, and 15), the mean sizes likely reflect agglomerations of contiguous smallholder fields, such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).
-A closer approximation of individual field boundaries can be mapped using approaches such as regression kriging to model the relationship between the segmented boundaries and the hand drawn labels collected at validation points (Figure S12B), although such an approach introduces spatial artifacts.
+The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in area such as AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
+To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S12).
+These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally aligning in the areas of higher field densities, although the biases inherent in both measures (Figures S10-11) complicate the interpretation of those variations.
+However, we can use the estimated to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana as a whole (Table \ref{tab:sizentab}).
+The bias-adjusted estimates show that the typical field size in Ghana is 1.73 ha, ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern half of the country (Table \ref{tab:sizentab}).
+The total number of fields was estimated to be 1,662,281 overall, or 205 fields per tile on average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.
+
+```{r sizentab, echo=FALSE, warning=FALSE, message=FALSE}
+fsizes_n_05 <- st_read(
+  system.file("extdata/field_size_n_05.geojson", package = "activemapper"),
+  quiet = TRUE
+) %>% mutate(
+  muarea = ifelse(is.na(muarea), 0, muarea),
+  nflds = ifelse(is.na(nflds), 0, nflds),
+  aoi = as.character(aoi)
+)
+
+# calculate bias factors for adjustment
+data("segment_quality_stats")
+# segment_quality_stats %>% filter(measure == "n")
+biases <- segment_quality_stats %>%
+  filter(Stat == "Mean" & scale == "grid") %>% 
+  pivot_wider(names_from = type, values_from = value) %>% 
+  mutate(biasfactor = Validation / Segmentation)
+biases <- tile_key %>% group_by(aoi) %>% 
+  count() %>% ungroup() %>% rename(ntiles = n) %>% 
+  mutate(aoi = as.character(aoi)) %>% 
+  left_join(biases, .) %>% 
+  select(-scale, -Stat)
+
+fsizes_05_adj <- biases %>% filter(measure == "area" & aoi != "All") %>% 
+  select(aoi, biasfactor) %>% 
+  rename(fbias = biasfactor) %>% 
+  left_join(as_tibble(fsizes_n_05), .) %>% 
+  mutate(muarea_adj = muarea * fbias) %>% select(-geometry) %>% 
+  group_by(aoi) %>% 
+  summarize(fsize = mean(muarea), fsizeadj = mean(muarea_adj))
+
+# then field count
+n_05_adj <- biases %>% filter(measure == "n" & aoi != "All") %>% 
+  select(aoi, biasfactor, ntiles) %>% 
+  rename(nbias = biasfactor) %>% 
+  left_join(as_tibble(fsizes_n_05), .) %>% 
+  mutate(nfldsadj = nflds * nbias) %>%
+  group_by(aoi) %>% 
+  summarize(N = sum(nflds), Nadj = sum(nfldsadj))
+
+aoi_sizes_n <- left_join(fsizes_05_adj, n_05_adj) %>% 
+  left_join(., biases %>% filter(measure == "area") %>% select(aoi, ntiles))
+
+ghana_sizes_n <- aoi_sizes_n %>% 
+  summarize(fsize = weighted.mean(fsize, ntiles), 
+            fsizeadj = weighted.mean(fsizeadj, ntiles), 
+            N = sum(N), Nadj = sum(Nadj), 
+            ntiles = round(sum(ntiles)))
+
+size_n_tab <- bind_rows(aoi_sizes_n, ghana_sizes_n) %>% 
+  mutate(aoi = as.numeric(ifelse(is.na(aoi), 17, aoi))) %>% 
+  arrange(aoi) %>% mutate(aoi = as.character(aoi)) %>% 
+  mutate(aoi = ifelse(aoi == "17", "Ghana", aoi), 
+         N_tile = round(N / ntiles),
+         Nadj_tile = round(Nadj / ntiles)) %>% 
+  select(aoi, ntiles, fsize, fsizeadj, N, N_tile, Nadj, Nadj_tile) %>% 
+  mutate(
+    ntiles = scales::number(
+      round(ntiles), accuracy = 1, big.mark = ",", trim = FALSE
+    ), 
+    fsize = round(fsize, 2), fsizeadj = round(fsizeadj, 2), 
+    N = scales::number(N, big.mark = ",", trim = FALSE), 
+    N_tile = scales::number(N_tile, big.mark = ",", accuracy = 1, trim = FALSE), 
+    Nadj = scales::number(round(Nadj), big.mark = ",", trim = FALSE),
+    Nadj_tile = scales::number(
+      round(Nadj_tile), accuracy = 1, big.mark = ",", trim = FALSE
+    )
+  ) %>% 
+  rename(AOI = aoi, `N tiles` = ntiles, Size = fsize, `Size (adj)` = fsizeadj, 
+         `N / tile` = N_tile, `N (adj)` = Nadj, 
+         `N (adj) / tile` = Nadj_tile)
+# size_n_tab %>% select(`N (adj) / tile`)
+
+cap <- glue::glue("The average size and total number of crop fields for each ", 
+                  "AOI and for Ghana overall. The original and bias-adjusted values", 
+                  " for each measure are provided, as well as the total number", 
+                  " of 0.05$^\\circ$ degree tiles in each AOI.")
+knitr::kable(size_n_tab, format = "latex", booktabs = "T", caption = cap, 
+             align = rep("r", 8)) %>% 
+  kableExtra::kable_styling(latex_options = c("hold_position")) %>%
+  kableExtra::column_spec(1:3, width = "0.45in") %>%
+  kableExtra::column_spec(4:8, width = "0.85in")
+```
 
@@
 The results 
+Bias in mean sizes likely reflect agglomerations of contiguous smallholder fields, and the difficulty of segmenting indistinct adjacent boundaries.
+Such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).
 

### 2021-02-23 inst/manuscript/si.Rmd
Updated manuscript, SI, figures with text and table, figure for field size and numbers, which is final part of results

@@
 
-```{r setup, include=FALSE, cache=FALSE, message = FALSE}
+```{r setup, include=FALSE, cache=FALSE, message=FALSE}
 library(knitr)
@@
 ## Cropland characteristics
-### Average field size
+### Field area and number
 
-To examine average field sizes, the average area of the vectorized field boundaries intersecting each 0.005 degree cell in the mapping grid was calculated and mapped back onto the grid (Figure \ref{fig:fieldsize}A).
-Given the overestimation bias inherent in the vectorized boundaries relative to the hand-digitized validation labels, we also used a regression kriging approach [@henglRegressionkrigingEquationsCase2007] to these average field sizes for this bias.
-To do this, we used the estimated mean field sizes extracted from the map in Figure S12A to predict the average polygon size of the most accurate mapper's labels for each of the 1600 validation sites, using an ordinary least squares regression to model the relationship, after first log-transforming both variables.
-We then applied ordinary kriging (using an exponential model for the variogram) to the regression residuals, which we added to the surface of the predicted field sizes to create the adjusted surface, and then converted this back to units of hectares (Figure \ref{fig:fieldsize}B).
+To examine average field sizes, and the total number, the mean segment size per 0.05 degree tiles was calculated and mapped, as well as the total number of fields per tile (Figure \ref{fig:fieldsizes}).
 
-```{r fieldsizes, echo = FALSE, out.width="100%", fig.cap="Average fields sizes in Ghana at 0.005 degree resolution, as calculated from A) the vectorized field boundaries, and after B) adjustment against hand-drawn validation labels using a regression-kriging approach.", fig.align='center', message=FALSE, results='hold'}
-knitr::include_graphics('figures/si_fsizes.png')
+```{r fieldsizes, echo = FALSE, out.width="100%", fig.cap="The A) average sizes of fields and B) total number of fields in each 0.05 degree tile, as calculated from the vectorized field boundaries.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_fsize_n_05.png')
 ```

### 2021-02-23 inst/manuscript/manuscript.Rmd
Merge in Su's edits to manuscript and SI

@@
 
-<!-- values. The first is based on a simplified formula of the Haze Optimized Transformation (HOT) index [@Zhangimagetransformcharacterize2002; @ZhuObjectbasedcloudcloud2012]:  -->
+<!-- values. The first is based on a simplified formula of the Haze Optimized Transformation (HOT) index [√; @ZhuObjectbasedcloudcloud2012]:  -->
 
@@
 \mathrm{W2_t} =\begin{cases}
-    1, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_1}$, $\mathrm{NIR_2}$, ..., $\mathrm{NIR_i}$\}}.\\
-    \frac{1}{\mathrm{NIR_t}^4}, & \text{otherwise}.
+    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_1}$, $\mathrm{NIR_2}$, ..., $\mathrm{NIR_i}$\}}.\\
+    1, & \text{otherwise}.
   \end{cases}
@@
 
-Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
-Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels, while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow.
+Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @ShiCirrusCloudsThatAffect2020].
 
@@
 \begin{equation}
-\mathrm{score_i}=\beta0\mathrm{I}+\beta1\mathrm{O}+\beta2\mathrm{F}+\beta3\mathrm{E}+\beta4\mathrm{C}
+\mathrm{score_i}=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}
 \end{equation}
@@
 
-In the second step, a Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands [**I assume**], and weighted edge image is computed using weights of 1, 1, and 2 to each band for the dry season only [**or growing season?**].
-A watershed algorithm (**cite**) is then run on the weighted edge image, with a high level of segmentation specified.
-In this case, we specified **2400** segments per tile.
+In the second step, a Sobel filter is then applied to the green, red, near-infrared mean-shifted bands and probability map , and an combined edge image is computed using the sum of these four edge images for the dry season only.
+A compact watershed algorithm [@NeubertCompactWatershedPreemptiveSlic2014] is then run on the weighted edge image, with a high level of segmentation specified.
+In this case, we specified **6400** segments per tile.
 
 In the third step, a region adjacency graph is constructed for each tile, in which each node represents all pixels within each polygon created in the previous step.
-The edge between two adjacent regions (polygons) is calculated as the difference between the means of normalized colors of all bands.
-Hierarchical merging is then applied, in which the most similar pairs of adjacent nodes are merged until there are no edges remaining below a predetermined threshold of **X**.
+The edge between two adjacent regions (polygons) is calculated as the norm of the difference between the means of normalized colors of all bands.
+Hierarchical merging is then applied, in which the most similar pairs of adjacent nodes are merged until there are no edges remaining below a predetermined threshold of 0.05.
 
@@
 
-In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Douglas-Peucker algorithm (cite).
+In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@VisvalingamLineGeneralisation1993].
 Neighboring polygon that touch along tile boundaries are merged.
@@
 Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map.
-[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy (**cite**), because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
+[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@YeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
 On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).
@@
 # Discussion
+
 We present here a method for developing high resolution maps that detail the characteristics of smallholder-dominated croplands, providing information on their geometries and distribution of <5 m resolution, and representing a single agricultural year.

### 2021-02-23 inst/manuscript/si.Rmd
Merge in Su's edits to manuscript and SI

@@
 
-After creating the consensus label, the degree of confidence in the resulting label value is measured by Bayes Risk:
+After creating the consensus label, the degree of confidence in the resulting label value is measured by Bayesian Risk:
 
@@
 
-Using equation \ref{eq:pixelrisk}, the corresponding risk associated with this particular pixel's label is thus `r rj * (1 - 1) + (1 - rj) * 1`.
+Using equation \ref{eq:pixelrisk}, the corresponding risk associated with this particular pixel's label is thus `r rj * (1 - 1) + (1 - rj) * 1` (i.e., 1 - `r rj <- round((wtA * 1 + wtB * 0) / (wtA + wtB), 3); rj`).
 
@@
 This entailed the creation of a map reference sample, which first entailed designing a sample, and then designing how the sample response would be collected.
+
+#### Worker's assignment evaluation
+We assigned a score for each workers' assign; the score history determines the weights of each workers for creating a consensus label for each 0.005 degree resolution grid.
+The score for a particular assignment *i* is defined as a weighted sum of five accuracy metrics:
+$$ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}$$
+Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.
+*I* is "inside the box" accuracy which is defined as a proportion of the area correctly mapped within a 0.005 degree resolution grid (*I_c*) over total "inside the box" area for this grid (*I_t*).
+$$ I=\frac{I_c}{I_t}$$
+*O* is "outside the box" accuracy which refers to a proportion of the field area correctly mapped outside the grid over total "outside the box" region (*O_t*, the region within the bounding box of the workers' polygons but not within 0.005 degree resolution grid).
+$$ O=\frac{O_c}{O_t}$$
+*F* is the fragmentation accuracy which is defined as a proportion of matched polygon number (*N_m*, the number of the workers' polygons that has at least 50% of its region overlapped with a reference polygon) over total workers' polygon number (*N_t*).
+$$ F=\frac{N_m}{N_t}$$
+*E* is the average edge accuracy for all pairs of matched workers and refernece polygons; the edge accuracy for a single pair is defined as the length of 'correctly mapped edges' (*L_c*, the partial boundary of a workers' polygon that are within a three-planet-pixel buffer region of the matched reference polygon boundary) over the total boundary length of its matched reference polygon (*L_t*).
+$$ E=\frac{L_c}{L_t}$$
+*C* is the categorical accuracy, i.e., a proportion of the area that has been correctly labeled with field category within intersected regions between worker's and reference polygons (*T_c*) over the total intersecting area (*T_t*).
+$$ C=\frac{T_c}{T_t}$$
 

### 2021-02-24 inst/manuscript/manuscript.Rmd
First section of Discussion with outline

@@
 We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
-These factors make Ghana emblematic of the substantial challenges facing cropland classification over the broader region.
+These factors make Ghana emblematic of the substantial challenges facing cropland mapping over the broader region.
 
@@
 
-Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @ShiCirrusCloudsThatAffect2020].
+Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
 
@@
 In the second step, a Sobel filter is then applied to the green, red, near-infrared mean-shifted bands and probability map , and an combined edge image is computed using the sum of these four edge images for the dry season only.
-A compact watershed algorithm [@NeubertCompactWatershedPreemptiveSlic2014] is then run on the weighted edge image, with a high level of segmentation specified.
+A compact watershed algorithm [@neubertCompactWatershedPreemptive2014] is then run on the weighted edge image, with a high level of segmentation specified.
 In this case, we specified **6400** segments per tile.
@@
 
-In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@VisvalingamLineGeneralisation1993].
+In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@visvalingamLineGeneralisationRepeated1993].
 Neighboring polygon that touch along tile boundaries are merged.
@@
 Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map.
-[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@YeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
+[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@yeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
 On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).
@@
 
-Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table 2).
+Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table \href{tab:mapaccuracy}).
 This overall level of accuracy varied within four smaller zones (Figure S9A), ranging from 83.3% in Zone 1 (comprised of AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
@@
 
-The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table 2).
+The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table \href{tab:mapaccuracy}).
 In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map, indicating that the segmentation step helped reduce omission error, but substantially increased false positive errors.
@@
 \includegraphics[width = 18cm]{figures/table2.png}
-\label{tab:aezaccuracy}
+\label{tab:mapaccuracy}
 \end{table}
@@
 # Discussion
-
-We present here a method for developing high resolution maps that detail the characteristics of smallholder-dominated croplands, providing information on their geometries and distribution of <5 m resolution, and representing a single agricultural year.
-
-The results 
-Bias in mean sizes likely reflect agglomerations of contiguous smallholder fields, and the difficulty of segmenting indistinct adjacent boundaries.
+```{r, echo=FALSE}
+# from lesiv et al (2019) SI field sizes for Ghana
+lesiv_fsizes <- tibble(n = c(6, 7, 25, 71, 154), 
+                       sz = c(100, mean(c(100, 16)), mean(c(2.56, 16)), 
+                              mean(c(0.64, 2.56)), 0.32))
+lesiv_est <- sum(with(lesiv_fsizes, (n / sum(n))  * sz))
+
+aoi_nofor <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
+  summarize(weighted.mean(fsize, ntiles)) %>% pull() %>% round(., 2)
+aoi_nofor_adj <- aoi_sizes_n %>% filter(aoi %in% c(1:9, 12, 15)) %>% 
+  summarize(weighted.mean(fsizeadj, ntiles)) %>% pull() %>% round(., 2)
+```
+
+These results demonstrate our system's ability to create high resolution maps of the distribution and characteristics of cropland across a diverse and hard-to-map agricultural systems, at country-scales and at sub-annual to annual time steps.
+Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
+These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our maps of cropland in 2018, which cover 16.1-23.2% of the mapped area (Table \href{tab:mapaccuracy}), and yield a sample-corrected estimate `r round(c(42359) / (42359 + 204940) * 100, 1)` to `r round(c(43233) / (43233 + 202856) * 100, 1)`%, which is lower than both prior estimates.
+
+In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \href{fig:mainmap}, S12).
+Prior efforts to map such agricultural characteristics have often focused on estimating farm rather than field size using census datasets [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004].
+Efforts to map fields in small-scale agricultural systems have either been field-based [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small areas [e.g.
+@forkuorIntegrationOpticalSynthetic2014].
+The most notable efforts to estimate field sizes were two global initiatives that used crowdsourced volunteers to categorize five broad field size classes based on their interpretations of imagery sampled from virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
+The most recent of these [@lesivEstimatingGlobalDistribution2019] provided country-specific results for Ghana (n = 263) that return an average field size of `r round(lesiv_est, 2)` ha (calculated as the weighted mean of the class counts and the middle values of the four smallest size classes and the lower bound of the largest size class).
+This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \href{tab:sizentab}), but is closer to the mean segment size (`r aoi_nofor` ha) across AOIs 1-9, 12, and 15, which is where most of their sample appears to have been collected.
+However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha are much smaller.
+
+## Performance of the system
+Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
+The overall map accuracies are good (86.7-88%, Table \href{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence).
+However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map, indicating that both commission and omission errors were above 30% for this class.
+The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%).
+These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017]
+
+Although the raw maps are biased, data from the labelling platform provides information that can be used to correct for these discrepancies.
+The detail, large extent, and temporal precision of the map were enabled by the system's ability to process small-sat data, which currently provide the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with the frequency necessary to create clean imagery in areas of frequent cloud cover across multiple crop growing periods, which is important for distinguishing cropland from non-cropland [@Debatsgeneralizedcomputervision2016].
+This map represents an capability represents an advance over Although free and providing higher radiometric quality and depth, a moderate-resolution sensor such as Sentinel-2 may be too infrequent to successfully penetrate clouds in an area such as southern Ghana (meaning that dry season imagery is likely to be all that is available), while lacking the spatial resolution necessary for labelling and classifying smaller fields.
+
+A possible reason for the discrepancy between earlier maps and ours is that those earlier maps include more areas of active or abandoned cropland in their estimates.
+
+using a mapping system that is designed to work with high spatio-temporal resolution small-sat data.
+
+- Map at large scale using PlanetScope, one of the first (any citations). 
+  - Done using small sat data for a single agricultural year, for two primary seasons
+  - Follows recommended best practice for training ML algorithms
+
+- Takes step towards answering important questions about the characteristics of fields and how many there are. 
+  - Provide insights into the characteristics of croplands
+  - Discussion about the characteristics of croplands: estimates of field sizes
+  - What is says about Ghana's croplands.
+
+## Performance of the system
+- Pixel-wise system, reasonable overall accuracy, but low performance on the agricultural class
+- Low performance primarily due to several factors:
+  - Model simplicity and scale of compute
+  - Challenge of labelling task - uncertainty in labels
+    - Should also be noted substantial uncertainty in identifying fields (map reference uncertainty)
+  - Difficulty of getting clean imagery in highly cloudy areas
+- Design of system helped to mitigate these factors:
+  - Consensus labels minimize error substantially
+  - Active learning selects cites more efficiently, some evidence it improves more efficiently and overall relative to random selection.
+Less marked gain here, but doing this at much larger scale than Debats
+  - Multi-season imagery helps improve it    
+    - Feng et al (2020) importance of temporal information (backs up Debats et al).
+    - temporal information makes labelling harder though
+  - Image quality:
+    - Developed our own mean-weighting not median composite (artifacts) (maybe methods)
+    - impact on accuracy--residual cloud issues
+    - Important to develop the method because best-on-top method of Planet's basemaps leaves lots of residual cloud in key seasons, and require large volumes of imagery in highly cloudy areas such as southerwestern Ghana.
+
+- Segmentation accuracy issues:
+  - Over-merging
+  - Bias in mean sizes likely reflect agglomerations of contiguous smallholder fields, and the difficulty of segmenting indistinct adjacent boundaries.
 Such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).
+  - Importance of having high quality probabilities underlying for filtering segments
+
+## Looking forward 
+- Sets up capability for repeated annual mapping of croplands to understand critical issues of how these systems are changing (sizes and shapes of croplands, where they are expanding) (Jayne papers, recent agricultural expansion paper)
+- Key components will be useful in other applications and regions
+- Provides means for improving crop type and crop yield estimates
+- Improvements to help realize these capabilities: promise of CNNs (cite Waldner paper, others), for segmentation to realize this.
+This systems is well designed to adapt to it, and capabilities for assessing various aspects of label quality, e.g.
+edge accuracy and fragmentation accuracy, can help improve quality of labels designed to emphasize edges
+  - Can still retain the active architecture, but instead use new sites to improve a pre-trained model
+- Map along agroecozone boundaries, rather than random blocks  
+- Another interesting question to answer: whether guided approach to site selection would result in more informative sample for CNNs also.
 
@@
 - Left unquantified is how much training on base maps matters
-- Image quality:
-    - why mean-weighting not median composite (artifacts) (maybe methods)
-    - impact on accuracy
+
 - Assessment of segmentation quality
@@
     - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.
-    - Feng et al (2020) importance of temporal information (backs up Debats et al).
-      - temporal information makes labelling harder though
-
-Improvements:    
-- Replace Random Forest with deep learning model. Can still retain the active architecture, but instead use new sites to improve a pre-trained model
-- Map along agroecozone boundaries, rather than random blocks
 

### 2021-02-24 inst/manuscript/si.Rmd
First section of Discussion with outline

--- 2021-02-23 inst/manuscript/si.Rmd
@@
 This entailed the creation of a map reference sample, which first entailed designing a sample, and then designing how the sample response would be collected.
+
+#### Worker's assignment evaluation
+We assigned a score for each workers' assign; the score history determines the weights of each workers for creating a consensus label for each 0.005 degree resolution grid.
+The score for a particular assignment *i* is defined as a weighted sum of five accuracy metrics:
+
+$$ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}$$
+
+Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.
+
+*I* is "inside the box" accuracy which is defined as a proportion of the area correctly mapped within a 0.005 degree resolution grid (*I$_c$*) over total "inside the box" area for this grid (*I$_t$*).
+
+$$ I=\frac{I_c}{I_t}$$
+
+*O* is "outside the box" accuracy which refers to a proportion of the field area correctly mapped outside the grid over total "outside the box" region (*O$_t$*, the region within the bounding box of the workers' polygons but not within 0.005 degree resolution grid).
+
+$$ O=\frac{O_c}{O_t}$$
+
+*F* is the fragmentation accuracy which is defined as a proportion of matched polygon number (*N$_m$*, the number of the workers' polygons that has at least 50% of its region overlapped with a reference polygon) over total workers' polygon number (*N$_t$*).
+
+$$ F=\frac{N_m}{N_t}$$
+
+*E* is the average edge accuracy for all pairs of matched workers and refernece polygons; the edge accuracy for a single pair is defined as the length of 'correctly mapped edges' (*L$_c$*, the partial boundary of a workers' polygon that are within a three-planet-pixel buffer region of the matched reference polygon boundary) over the total boundary length of its matched reference polygon (*L$_t$*).
+
+$$ E=\frac{L_c}{L_t}$$
+
+*C* is the categorical accuracy, i.e., a proportion of the area that has been correctly labeled with field category within intersected regions between worker's and reference polygons (*T$_c$*) over the total intersecting area (*T$_t$*).
+
+$$ C=\frac{T_c}{T_t}$$
 
@@
 
-The average Bayes Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}).
+The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}).
 The three initial clusters include the second mapping of Cluster 2.
@@
 
-```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayes Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
+```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
 The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}

### 2021-02-24 inst/manuscript/manuscript.Rmd
Discussion drafted through error sources

@@
 
-## Performance of the system
+## Sources of error
 Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
@@
 
-Although the raw maps are biased, data from the labelling platform provides information that can be used to correct for these discrepancies.
-The detail, large extent, and temporal precision of the map were enabled by the system's ability to process small-sat data, which currently provide the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with the frequency necessary to create clean imagery in areas of frequent cloud cover across multiple crop growing periods, which is important for distinguishing cropland from non-cropland [@Debatsgeneralizedcomputervision2016].
-This map represents an capability represents an advance over Although free and providing higher radiometric quality and depth, a moderate-resolution sensor such as Sentinel-2 may be too infrequent to successfully penetrate clouds in an area such as southern Ghana (meaning that dry season imagery is likely to be all that is available), while lacking the spatial resolution necessary for labelling and classifying smaller fields.
-
-A possible reason for the discrepancy between earlier maps and ours is that those earlier maps include more areas of active or abandoned cropland in their estimates.
-
-using a mapping system that is designed to work with high spatio-temporal resolution small-sat data.
-
-- Map at large scale using PlanetScope, one of the first (any citations). 
-  - Done using small sat data for a single agricultural year, for two primary seasons
-  - Follows recommended best practice for training ML algorithms
-
-- Takes step towards answering important questions about the characteristics of fields and how many there are. 
-  - Provide insights into the characteristics of croplands
-  - Discussion about the characteristics of croplands: estimates of field sizes
-  - What is says about Ghana's croplands.
-
-## Performance of the system
-- Pixel-wise system, reasonable overall accuracy, but low performance on the agricultural class
-- Low performance primarily due to several factors:
-  - Model simplicity and scale of compute
-  - Challenge of labelling task - uncertainty in labels
-    - Should also be noted substantial uncertainty in identifying fields (map reference uncertainty)
-  - Difficulty of getting clean imagery in highly cloudy areas
+These relatively low accuracies for the cropland class can be attributed to several important factors.
+The first of these was the overall mapping extent combined with the resolution of the analysis.
+Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended [@MaxwellImplementationmachinelearningclassification2018] tree number of depth  in our RandomForest implementation.
+Previous work shows that RandomForest achieves much better performance on small-scale croplands when trained on a large number of randomly generated features [@Debatsgeneralizedcomputervision2016], which is a commonality with deep learning.
+However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the typical 10-14 hours required from training through prediction on 800 CPUs, for a single iteration.
+This left the model less able to discern cropland from surrounding savanna vegetation with bare patches or sparse grass cover, which was common in most AOIs.
+A larger feature set with more variable textures filters could have helped distinguish such areas from crop fields, helping to particularly reduce commission error.
+
+Another major limiting factor was the inherent difficulty of the labelling task.
+The system is designed to minimize the error inherent in labelling, but determining croplands from non-croplands over these system is difficult and requires educated guessing by each labeller.
+Interpretation is particularly hard over areas where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3.
+Such changes in scene circumstances require labellers to adopt different heuristics, for example by delineating the field boundary on the growing season rather than the dry season image.
+Smaller field sizes also complicated the labelling task, as smaller fields become increasingly indistinct in the ~4 m PlanetScope composites.
+These challenges caused substantial differences in labels, as evident in the Bayesian Risk metrics (Figure S5) and labeller accuracies that averaged 71%, as measured against our reference set.
+The design of our system allowed us to measure how much label quality impacts model performance.
+Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] has reported that RandomForest is particularly robust to training data error, we found its impact to be substantial (Figure \href{fig:trainingimpact}), suggesting that the greatest single investment towards improved accuracy lies in improving training data quality.
+
+Image quality was another issue, although primarily in the forest zone (Zone 4) AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \href{fig:imqual}).
+This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions.
+There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance--prior work shows that imagery from multiple seasons is important for improving classifier performance [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+Alternatively, composite quality could be improved by integrating imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would confound the ability to delineate field boundaries.
+Alternative compositing approaches might also be ineffective, as the "best-on-top" approach\footnote{https://assets.planet.com/products/basemap/planet-basemaps-product-specifications.pdf} used to develop PlanetScope basemaps tends to result in substantial residual clouds in this region.
+
+The final major error sources relates to the segmentation algorithm, which had two particular problems.
+The first of these relates to the substantial commission errors in the vector maps, which is attributable to errors in the RandomForest predictions.
+Uncertainty in that model resulted in substantial numbers of pixels just above the 0.5 classification threshold, meaning a number of segments over highly uncertain areas were retained as fields.
+This problem would resolve with a more accurate classifier, or potentially by using locally varying thresholds for distinguishing croplands from non-cropland [e.g.
+@waldnerDeepLearningEdge2020].
+The second problem was over-merging by the segmentation algorithm, which caused over-estimation of field sizes and often irregular shapes, particularly in high density croplands, such as in AOIs 2 and 8 (Figure \href{fig:mainmap}), where closely adjacent small-scale fields have boundaries that are often indistinct in the PlanetScope imagery.
+In this case, minimizing or preventing merging could help, resulting in an over-segmented solution that would likely underestimate field sizes.
+However, from the standpoint of other map applications, such as filtering pixels for crop type mapping or yield estimation, over-segmented solutions could help by isolating distinct within field units, such as those caused by soil type differences.
+
+## Key system features
 - Design of system helped to mitigate these factors:
@@
   - Image quality:
-    - Developed our own mean-weighting not median composite (artifacts) (maybe methods)
-    - impact on accuracy--residual cloud issues
-    - Important to develop the method because best-on-top method of Planet's basemaps leaves lots of residual cloud in key seasons, and require large volumes of imagery in highly cloudy areas such as southerwestern Ghana.
-
-- Segmentation accuracy issues:
-  - Over-merging
-  - Bias in mean sizes likely reflect agglomerations of contiguous smallholder fields, and the difficulty of segmenting indistinct adjacent boundaries.
-Such as in AOI 2,and the southern portion of AOI 8 (Figure S12A).
-  - Importance of having high quality probabilities underlying for filtering segments
+
+Although the raw maps are biased, data from the labelling platform provides information that can be used to correct for these discrepancies.
+The detail, large extent, and temporal precision of the map were enabled by the system's ability to process small-sat data, which currently provide the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with the frequency necessary to create clean imagery in areas of frequent cloud cover across multiple crop growing periods, which is important for distinguishing cropland from non-cropland [@Debatsgeneralizedcomputervision2016].
+This map represents an capability represents an advance over Although free and providing higher radiometric quality and depth, a moderate-resolution sensor such as Sentinel-2 may be too infrequent to successfully penetrate clouds in an area such as southern Ghana (meaning that dry season imagery is likely to be all that is available), while lacking the spatial resolution necessary for labelling and classifying smaller fields.
+
+A possible reason for the discrepancy between earlier maps and ours is that those earlier maps include more areas of active or abandoned cropland in their estimates.
+
+using a mapping system that is designed to work with high spatio-temporal resolution small-sat data.
 
 ## Looking forward 
+- Provide new insight 
+- Given errors, should be considered generation 1 maps
 - Sets up capability for repeated annual mapping of croplands to understand critical issues of how these systems are changing (sizes and shapes of croplands, where they are expanding) (Jayne papers, recent agricultural expansion paper)
@@
 - Provides means for improving crop type and crop yield estimates
-- Improvements to help realize these capabilities: promise of CNNs (cite Waldner paper, others), for segmentation to realize this.
+- Improvements to help realize these capabilities: 
+-     [This suggests that replacing the Random Forest more advanced deep learning model].
+promise of CNNs (cite Waldner paper, others), for segmentation to realize this.
 This systems is well designed to adapt to it, and capabilities for assessing various aspects of label quality, e.g.
@@
 - Another interesting question to answer: whether guided approach to site selection would result in more informative sample for CNNs also.
+- Higher number of high quality reference labels can achieve this.
+
+# Suggested uses
+- Fallow versus active
 

### 2021-02-25 inst/manuscript/manuscript.Rmd
Discussion updated through penultimate section

@@
 
-Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table \href{tab:mapaccuracy}).
+Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table \ref{tab:mapaccuracy}).
 This overall level of accuracy varied within four smaller zones (Figure S9A), ranging from 83.3% in Zone 1 (comprised of AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
@@
 
-The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table \href{tab:mapaccuracy}).
+The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table \ref{tab:mapaccuracy}).
 In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map, indicating that the segmentation step helped reduce omission error, but substantially increased false positive errors.
@@
 cap <- glue::glue("The average size and total number of crop fields for each ", 
-                  "AOI and for Ghana overall. The original and bias-adjusted values", 
-                  " for each measure are provided, as well as the total number", 
-                  " of 0.05$^\\circ$ degree tiles in each AOI.")
+                  "AOI and for Ghana overall. The original and bias-adjusted", 
+                  " values for each measure are provided, as well as the total", 
+                  " number of 0.05$^\\circ$ degree tiles in each AOI.")
 knitr::kable(size_n_tab, format = "latex", booktabs = "T", caption = cap, 
@@
 Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
-These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our maps of cropland in 2018, which cover 16.1-23.2% of the mapped area (Table \href{tab:mapaccuracy}), and yield a sample-corrected estimate `r round(c(42359) / (42359 + 204940) * 100, 1)` to `r round(c(43233) / (43233 + 202856) * 100, 1)`%, which is lower than both prior estimates.
-
-In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \href{fig:mainmap}, S12).
+These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our 2018 maps, in which cropland covers 16.1-23.2% of the mapped area (Table \ref{tab:mapaccuracy}), and yield a sample-corrected estimate `r round(c(42359) / (42359 + 204940) * 100, 1)` to `r round(c(43233) / (43233 + 202856) * 100, 1)`%.
+This result could suggest that Ghana's arable cropland occupies is less than these previous estimates, or that these previous maps used a more expansive definition of cropland than we used.
+
+In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S12).
 Prior efforts to map such agricultural characteristics have often focused on estimating farm rather than field size using census datasets [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004].
@@
 The most recent of these [@lesivEstimatingGlobalDistribution2019] provided country-specific results for Ghana (n = 263) that return an average field size of `r round(lesiv_est, 2)` ha (calculated as the weighted mean of the class counts and the middle values of the four smallest size classes and the lower bound of the largest size class).
-This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \href{tab:sizentab}), but is closer to the mean segment size (`r aoi_nofor` ha) across AOIs 1-9, 12, and 15, which is where most of their sample appears to have been collected.
+This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean segment size (`r aoi_nofor` ha) across AOIs 1-9, 12, and 15, which is where most of their sample appears to have been collected.
 However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha are much smaller.
@@
 Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
-The overall map accuracies are good (86.7-88%, Table \href{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence).
+The overall map accuracies are good (86.7-88%, Table \ref{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence).
 However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map, indicating that both commission and omission errors were above 30% for this class.
@@
 Another major limiting factor was the inherent difficulty of the labelling task.
-The system is designed to minimize the error inherent in labelling, but determining croplands from non-croplands over these system is difficult and requires educated guessing by each labeller.
+Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands over these system is difficult and requires the evaluation of multiple image sources and considerable judgement by each labeller.
 Interpretation is particularly hard over areas where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3.
 Such changes in scene circumstances require labellers to adopt different heuristics, for example by delineating the field boundary on the growing season rather than the dry season image.
-Smaller field sizes also complicated the labelling task, as smaller fields become increasingly indistinct in the ~4 m PlanetScope composites.
-These challenges caused substantial differences in labels, as evident in the Bayesian Risk metrics (Figure S5) and labeller accuracies that averaged 71%, as measured against our reference set.
+Smaller field sizes also complicated the labelling task, becoming increasingly indistinct in the ~4 m PlanetScope composites.
+The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by labeller accuracies that averaged 71% (Figure S4B).
 The design of our system allowed us to measure how much label quality impacts model performance.
-Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] has reported that RandomForest is particularly robust to training data error, we found its impact to be substantial (Figure \href{fig:trainingimpact}), suggesting that the greatest single investment towards improved accuracy lies in improving training data quality.
-
-Image quality was another issue, although primarily in the forest zone (Zone 4) AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \href{fig:imqual}).
+Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] has reported that RandomForest is particularly robust to training data error, we found its impact to be substantial (Figure \ref{fig:trainingimpact}), suggesting that the greatest single investment towards improved accuracy lies in improving training data quality.
+
+Image quality was another issue, although primarily in the forest zone (Zone 4) AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}).
 This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions.
 There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance--prior work shows that imagery from multiple seasons is important for improving classifier performance [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
-Alternatively, composite quality could be improved by integrating imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would confound the ability to delineate field boundaries.
-Alternative compositing approaches might also be ineffective, as the "best-on-top" approach\footnote{https://assets.planet.com/products/basemap/planet-basemaps-product-specifications.pdf} used to develop PlanetScope basemaps tends to result in substantial residual clouds in this region.
-
-The final major error sources relates to the segmentation algorithm, which had two particular problems.
-The first of these relates to the substantial commission errors in the vector maps, which is attributable to errors in the RandomForest predictions.
+Composite quality could be improved by integrating imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would confound the ability to delineate field boundaries.
+
+The final major source of error arose from the segmentation process.
+The vectorized maps had high commission errors that are attributable to the RandomForest predictions.
 Uncertainty in that model resulted in substantial numbers of pixels just above the 0.5 classification threshold, meaning a number of segments over highly uncertain areas were retained as fields.
-This problem would resolve with a more accurate classifier, or potentially by using locally varying thresholds for distinguishing croplands from non-cropland [e.g.
+A more accurate classifier would reduce such errors.
+Another possible solution could be to apply a locally varying classification threshold to distinguish croplands from non-cropland [e.g.
 @waldnerDeepLearningEdge2020].
-The second problem was over-merging by the segmentation algorithm, which caused over-estimation of field sizes and often irregular shapes, particularly in high density croplands, such as in AOIs 2 and 8 (Figure \href{fig:mainmap}), where closely adjacent small-scale fields have boundaries that are often indistinct in the PlanetScope imagery.
-In this case, minimizing or preventing merging could help, resulting in an over-segmented solution that would likely underestimate field sizes.
-However, from the standpoint of other map applications, such as filtering pixels for crop type mapping or yield estimation, over-segmented solutions could help by isolating distinct within field units, such as those caused by soil type differences.
+Over-merging was another source of error in the segmentation algorithm.
+This led in some areas to over-estimation of field sizes and often irregular shapes, particularly in high density croplands (e.g.
+in AOIs 2 and 8; Figure \ref{fig:mainmap}) where adjacent fields have boundaries that are indistinct in the PlanetScope imagery.
+Minimizing or preventing merging would help in such cases, although it would result in  over-segmentation and thus would likely underestimate field sizes.
+However, over-segmentation can help to identify important within-field distinctions (e.g.
+different soil types), which can be useful for other applications, such as filtering pixels for crop type mapping or yield estimation [e.g.
+@EstesComparingmechanisticempirical2013; @EstesProjectedclimateimpacts2013]
 
 ## Key system features
-- Design of system helped to mitigate these factors:
-  - Consensus labels minimize error substantially
-  - Active learning selects cites more efficiently, some evidence it improves more efficiently and overall relative to random selection.
-Less marked gain here, but doing this at much larger scale than Debats
-  - Multi-season imagery helps improve it    
-    - Feng et al (2020) importance of temporal information (backs up Debats et al).
-    - temporal information makes labelling harder though
-  - Image quality:
-
-Although the raw maps are biased, data from the labelling platform provides information that can be used to correct for these discrepancies.
-The detail, large extent, and temporal precision of the map were enabled by the system's ability to process small-sat data, which currently provide the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with the frequency necessary to create clean imagery in areas of frequent cloud cover across multiple crop growing periods, which is important for distinguishing cropland from non-cropland [@Debatsgeneralizedcomputervision2016].
-This map represents an capability represents an advance over Although free and providing higher radiometric quality and depth, a moderate-resolution sensor such as Sentinel-2 may be too infrequent to successfully penetrate clouds in an area such as southern Ghana (meaning that dry season imagery is likely to be all that is available), while lacking the spatial resolution necessary for labelling and classifying smaller fields.
-
-A possible reason for the discrepancy between earlier maps and ours is that those earlier maps include more areas of active or abandoned cropland in their estimates.
-
-using a mapping system that is designed to work with high spatio-temporal resolution small-sat data.
+Despite the sources of errors mentioned above, the design of the system proved effective in mitigating error, resulting in a higher overall accuracy than would have otherwise been possible.
+The label accuracy assessment and consensus labelling methods appeared to be the most effective tools for error mitigation.
+Label accuracy assessment allowed us to quantify the impact of trainnig data error on model performance, in keeping with recommended best practices [@ElmesAccountingtrainingdata2020].
+Label quality proved to have a large impact on model performance (Figure \ref{fig:trainingimpact}), and consensus labeling was effective in reducing individual digitization errors and between-labeller differences in interpretation, improving label, and thus map, quality.
+Worker-specific label accuracy measures also helped to improve the overall quality of the consensus labels, by enabling higher weight to be placed on labels that were more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels.
+The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we could in turn use to bias-correct map-based size class estimates (Table \ref{tab:sizentab}).
+The ability to select the most accurate label will also help when different training strategies are needed, such as tasking a model to classify both field edge and interior classes [e.g.
+@waldnerDeepLearningEdge2020; @aungFarmlandParcelDelineation2020].
+In such cases, the vectorized labels are needed to preserve information about the field edges, which are removed during consensus label generation.
+
+```{r, message = FALSE, warning = FALSE, echo=FALSE}
+prdelta <- iteration_metrics %>% 
+  filter(Metric %in% c("FPR", "TPR")) %>% 
+  group_by(Iteration, Metric) %>% 
+  summarize(Score = mean(Score)) %>% 
+  filter(Iteration %in% c(0, 3)) %>% 
+  tidyr::pivot_wider(names_from = Iteration, values_from = Score) %>% 
+  mutate(rate = round((`0` - `3`) / `0` * 100, 1))
+```
+The active learning approach helped to improve overall model performance relative to a randomized approach to training site selection.
+Although we only compared active to random site selection within three AOIs, the average gains across all three AOIs and three performance metrics demonstrated the effectiveness of the active approach, most notably by the gains in the F1 score (Figure S7), a combined metric of commission and omission error.
+The gains in accuracy and AUC were smaller.
+For accuracy, this was due to the larger drop in the active learning metric after iteration 1 in AOI 15, which proved one of the hardest AOIs to map.
+The plateau in AUC gains at 0.5% relative to random training reflects the fact that each iteration tended to reduce both the false and true positive rates simultaneously.
+Although the false positive rate overall declined (an average of `r prdelta$rate[1]`% between Iterations 0 and 3), the reductions in true positive rate (averaging `r prdelta$rate[2]`%) offset some of the improved predictive power.
+
+The gain due to active learning were substantially less than those reported by Debats et al (2017), who found that active site selection improved classifier performance by `r round((0.45-0.35) / 0.35 * 100)` over randomized site selection after one iteration, and after 45 iterations by `r round((0.69-0.64) / 0.64 * 100)`%.
+These larger gains likely reflect the fact that this evaluation began from the smallest possible training sample--one image chip--and tested the effect of adding a single additional chip each iteration.
+In contrast, we started with a large initial random sample (500) and added 100 samples per iteration, thus our models received significant training at the outset, making the marginal gain due to each new sample smaller.
+This suggests that we could have benefitted by reducing the size of the initial random training sample while increasing the size of actively selected samples, not only in terms of larger accuracy gains, but also in terms of reducing--by as much as 62%, per Debats et al (2017)--the overall size of the training sample needed to achieve a given level of performance.
+
+The detail, large extent, and temporal precision of the maps we created by virtue of our system's intentional design around PlanetScope imagery, which currently provides the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with daily frequency.
+Daily frequency is important for creating multiple seasonal composites during a single agricultural year in cloudy regions.
+Although unquantified, our compositing technique may be more effective at providing complete image coverage during a season than the "best-on-top" approach\footnote{https://assets.planet.com/products/basemap/planet-basemaps-product-specifications.pdf} used by Planet to develop their analytic basemap, which anecdotally appears to contains substantial residual cloud cover during the growing season in southern Ghana.
+Although free and providing higher radiometric quality and depth, and of sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
+@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], Sentinel-2's 5-day revist may be too infrequent to provide enough clear pixels to create a cloud-free growing season composite in the southern half of Ghana.
+Sentinel-1 is not affected by the same problem, but labelling fields in a more coarsely resolved radar images is difficult.
+
+## Lingering questions
+Several potential issues not answered by our assessment merit further exploration.
+These include the correspondence between image- and ground-collected labels.
+We believe that such comparisons will in most circumstances will reveal unresolvable discrepancies.
+The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels can produce substantial differences.
+These discrepancies could be further exacerbated by differences in an understanding of what constitutes a field, which might vary on the ground depending on the farmer or the data collecter's collectors judgement.
+On top of this, field-collected boundaries are typically highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is.
+These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth".
+Nevertheless, a comparison against ground data can help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.
+
+The temporal discrepancies mentioned above [and discussed in @ElmesAccountingtrainingdata2020] are another reason why we chose not to  label on basemap imagery (in addition to restrictive usage terms), which is typically several years old [@LesivCharacterizingspatialtemporal2018].
+However, we did not assess whether the higher label accuracy one might achieve by digitizing on <1-2 m basemap would offset model errors due temporal mismatches between labels and imagery.
+
+Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpacts}) was influenced by the validation labels we used, which were generated using the consensus method.
+This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels.
+However, the visual assessment of probability images generated from each labelling strategy show that these two strategies produce nearly identical maps, and that the model trained with low quality labels generated a markedly less certain map (Figure S6).
 
@@
   - Can still retain the active architecture, but instead use new sites to improve a pre-trained model
+  - Deep learning may be more resistant to label errors (cites)
 - Map along agroecozone boundaries, rather than random blocks  
@@
 
-# Suggested uses
-- Fallow versus active
-
-- Label quality
-  - Accuracy asssessment at training reference sites maybe over-inflated. Each worker saw training reference site multiple times. Need larger number of reference samples
-  - Training reference to training ratio was artificially inflated. Workers not always notified when training labels complete, and thus only served training reference sites. 
-  - Image-interpreted versus ground labels: **However, these discrepancies are not necessarily an indication of how far workers were from the "truth", but may instead be at least partially an indication of how different the perspectives are as to what constitutes a field when looking at an image versus seeing it from the ground (which in many cases was based on the field team's judgement, rather than farmer guidance).**
-
-Compare values
-
-- The impact of training data error on map quality:
-  - Label quality makes a large difference. 
-  - Although these comparisons were assessed against consensus-generated validation labels, a visual assessment of probability images generated from each labelling strategy shows markedly greater uncertainty between consensus and most accurate individual labels maps relative to those generated with the lowest quality labels are clearly less certain (Figure S6).
-
-- Left unquantified is how much training on base maps matters
-
-- Assessment of segmentation quality
 - Next steps: 
-    - Annually updated cropland maps
     - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]

### 2021-02-26 inst/manuscript/manuscript.Rmd
Full discussion drafted

@@
 
+- name: Zhenhua Meng
+  affilnum: '1'
+
+- name: Qi Zhang
+  affilnum: '1'
+
 - name: Dennis McRitchie
@@
 
+- name: Justus Muhando
+  affilnum: '5'
+
 - name: SpatialCollective
   affilnum: '5'
-
-- name: Meridia
-  affilnum: '6'
 
@@
 
+```{r, echo=FALSE}
+data("accuracy_reports")
+
+# Adjusted areas
+rfa <- accuracy_reports$ghana$rf[, "Area"]
+rfa <- rfrep[nchar(rfa) > 0]
+rfas <- as.numeric(gsub("\\(.*", "", rfa[nchar(rfa) > 0]))
+rfma <- as.numeric(stringr::str_match(rfa[1], "\\(\\s*(.*?)\\s*\\)")[, 2])
+cipct <- function(x, moe) {
+  apct <- x[2] / sum(x)
+  moepct <- moe / sum(x)
+  round(c(apct, apct - moepct, apct + moepct) * 100, 1)
+}
+sega <- accuracy_reports$ghana$seg[, "Area"]
+sega <- segrep[nchar(sega) > 0]
+segas <- as.numeric(gsub("\\(.*", "", sega[nchar(sega) > 0]))
+segma <- as.numeric(stringr::str_match(sega[1], "\\(\\s*(.*?)\\s*\\)")[, 2])
+
+rfcipct <- cipct(rfas, rfma)
+segcipct <- cipct(segas, segma)
+
+fmt <- function(x) format(x, big.mark = ",")
+```
+
 Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of arable croplands in Ghana.
-The cropland extent estimated from the vectorized map is 42,359 km$^2$ (with a margin of error of 4,395 km$^2$), or `r round(42359 / (42359 + 204940) * 100, 1)`% of the mapped area, the estimate based on the per pixel map is 43,233 km$^2$ (margin of error = 4,904 km$^2$), or `r round(43233 / (43233 + 202856) * 100, 1)`% of area.
+The cropland extent estimated from the vectorized map is `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area.
+The estimate based on the per pixel map is `r fmt(rfas[2])` km$^2$ (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.
 
@@
   summarize(weighted.mean(fsizeadj, ntiles)) %>% pull() %>% round(., 2)
+
 ```
@@
 Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
-These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our 2018 maps, in which cropland covers 16.1-23.2% of the mapped area (Table \ref{tab:mapaccuracy}), and yield a sample-corrected estimate `r round(c(42359) / (42359 + 204940) * 100, 1)` to `r round(c(43233) / (43233 + 202856) * 100, 1)`%.
-This result could suggest that Ghana's arable cropland occupies is less than these previous estimates, or that these previous maps used a more expansive definition of cropland than we used.
+These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our 2018 maps, in which cropland covers 16.1-23.2% of the mapped area (Table \ref{tab:mapaccuracy}), with sample-corrected estimates of `r segcipct[1]`-`r rfcipct[1]`%.
+Our findings suggests that Ghana's arable cropland is slightly less than previous estimates, although the difference may also be attributable to the earlier maps using a more expansive definition of cropland than ours (active to recently active fields).
 
@@
 The overall map accuracies are good (86.7-88%, Table \ref{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence).
-However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map, indicating that both commission and omission errors were above 30% for this class.
+However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map across the country, indicating that both commission and omission errors were above 30% for this class.
 The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%).
-These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017]
-
-These relatively low accuracies for the cropland class can be attributed to several important factors.
+These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
+
+The patterns of cropland-class accuracies varied by zone, which largely align with agroecozones and thus somewhat reflective of the inherent challenge each poses for mapping.
+Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savanna agroecoregions (Figure S9), and lowest in zones 3 and 4, which are either forest or coastal savannas.
+User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample.
+Confining the samples to specific agroecozones provides further insight into error patterns (Table S4).
+The coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably due to the numerous areas of high density cropland and low woody cover in surrounding uncultivated areas, which promotes commission error.
+The northern savannas had the best balance between omission and commission error, and had the highest overall User's accuracy.
+The transition zone had a very low Producer's accuracy (21%), which likely reflects the fact that this region, which also has high cropland density, was divided between several AOIs for mapping.
+
+The relatively low accuracies for the cropland class can be attributed to several important factors.
 The first of these was the overall mapping extent combined with the resolution of the analysis.
-Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended [@MaxwellImplementationmachinelearningclassification2018] tree number of depth  in our RandomForest implementation.
+Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forest implementation.
 Previous work shows that RandomForest achieves much better performance on small-scale croplands when trained on a large number of randomly generated features [@Debatsgeneralizedcomputervision2016], which is a commonality with deep learning.
@@
 The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we could in turn use to bias-correct map-based size class estimates (Table \ref{tab:sizentab}).
-The ability to select the most accurate label will also help when different training strategies are needed, such as tasking a model to classify both field edge and interior classes [e.g.
-@waldnerDeepLearningEdge2020; @aungFarmlandParcelDelineation2020].
-In such cases, the vectorized labels are needed to preserve information about the field edges, which are removed during consensus label generation.
+The ability to select the most accurate label will also help when different training strategies are needed, such as when training a model to perform instance segmentation [@carvalhoInstanceSegmentationLarge2021], in which case the field identities have to be preserved.
 
@@
 
-Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpacts}) was influenced by the validation labels we used, which were generated using the consensus method.
+Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpact}) was influenced by the validation labels we used, which were generated using the consensus method.
 This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels.
@@
 
-## Looking forward 
-- Provide new insight 
-- Given errors, should be considered generation 1 maps
-- Sets up capability for repeated annual mapping of croplands to understand critical issues of how these systems are changing (sizes and shapes of croplands, where they are expanding) (Jayne papers, recent agricultural expansion paper)
-- Key components will be useful in other applications and regions
-- Provides means for improving crop type and crop yield estimates
-- Improvements to help realize these capabilities: 
--     [This suggests that replacing the Random Forest more advanced deep learning model].
-promise of CNNs (cite Waldner paper, others), for segmentation to realize this.
-This systems is well designed to adapt to it, and capabilities for assessing various aspects of label quality, e.g.
-edge accuracy and fragmentation accuracy, can help improve quality of labels designed to emphasize edges
-  - Can still retain the active architecture, but instead use new sites to improve a pre-trained model
-  - Deep learning may be more resistant to label errors (cites)
-- Map along agroecozone boundaries, rather than random blocks  
-- Another interesting question to answer: whether guided approach to site selection would result in more informative sample for CNNs also.
-- Higher number of high quality reference labels can achieve this.
-
-- Next steps: 
-    - Improvement with deep learning approaches, even with simpler labelling methods [@WangWeaklySupervisedDeep2020]
-- Relevant literature: 
-    - [@Tongforgottenlanduse2020] discuss Sentinel-based mapping in Sahel, to detect fallows (reference to our intermediate probability values).
-Semi-automated approach to generating reference data from imagery (based on phenology)
-    - [@BeyMappingsmallholderlargescale2020] use CollectEarth to train an algorithm. Multi-stage. No reporting of training or reference data accuracy. 
-    - [@LuoDevelopingclearskycloud2008] relevant compositing method to cite, which sounds similar to ours.
+## Next steps
+The maps presented here represent a version 1 product that is freely available to use **(links)**, along with its underlying code.
+These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020].
+In their current form, the maps may be useful for a variety of research applications.
+For example, analyzing the  distributions of values in the probability maps may provide additional insight into the relative extents of active versus fallow croplands [@Tongforgottenlanduse2020].
+However, use of these data, particularly for decision-making processes (e.g.
+cropped area estimates), should be careful to account for the reported errors [@OlofssonGoodpracticesestimating2014; @StehmanKeyissuesrigorous2019].
+
+To facilitate the next step, generating more accurate version 2 maps, several improvements will be made.
+The first is to replace Random Forest with a more advanced convolutional neural network (CNN), which can generate and learn from a large number of features representing a variety of spatial scales [@maDeepLearningRemote2019].
+Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, provide an effective means for delineating field boundaries [@waldnerDeepLearningEdge2020].
+Our system can readily incorporate such a model.
+The active learning framework is already analagous to transfer learning, in which a trained deep learning model is applied to  a different region and updated with local training samples [@panSurveyTransferLearning2010; @russwurmMetalearningFewshotLand2020].
+The existing tooling can thus be adjusted to guide training site selection in sub-regions of Ghana for a CNN trained from scratch from a larger initial random draw.
+Instead of the current AOIs, a more effective strategy may be to update the model within agroecozones (Figure S9B) The labelling methods already provide for a case in which edge and interior classes need to be labelled, as the most accurate individual training samples can be identified and used, rather than the consensus strategy.
+The segmentation algorithm will also be updated to work with the resulting score maps.
+
+## Conclusion
+This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands over national to regional scales, using a flexible and framework that can be readily updated to improve their accuracy as technologies improve.
+Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.
 
 # Acknowledgements
+The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund.
+Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393).
+We thank Joe Okyere and his colleagues at Meridia for assistance in data collection.
 

### 2021-02-27 inst/manuscript/manuscript.Rmd
Revised p1 of intro

@@
 ---
-title: "Mapping the characteristics of small-scale agricultural systems at national extents using small sat data and an active learning approach"
+title: "High resolution, annual maps of the characteristics of small-scale croplands at national extents"
 
@@
 # Introduction
-How much cropland is on the planet, and where is it located?
-How much more will be needed to meet humanity's rapidly growing food demands?
-The answer to the first question is unclear, as existing estimates tend to vary widely and cropland maps show substantial spatial disagreements [e.g.
-@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], which makes it difficult to answer the second.
-The first question can't be clearly answered because cropland estimates rely heavily on remote sensing-derived landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for an accurate baseline is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015].
-*[Fill in more reason why cropland maps are important]*
+Amidst all the challenges posed by global change, a particular concern is how agriculture will adapt to meet humanity's growing food demands, and the impacts that a transforming and expanding agriculture will have on societies, economies, and environment [@searchingerCreatingSustainableFood2019].
+Significant efforts are being made to address the various aspects of this challenge, including work on  diagnosing and closing yield gaps [e.g.
+@LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts.
+The ability to answer the questions these efforts seek to address depends on data that describes the location and nature of cropland, and how that is changing over time.
+Unfortunately, the data that do exist are in many places inaccurate.
+Existing estimates of how much cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
+@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013].
+These discrepancies are due to the fact that cropland data in many regions are derived from remotey sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]
 

### 2021-03-02 inst/manuscript/manuscript.Rmd
MS edited through methods

@@
 ---
-title: "High resolution, annual maps of the characteristics of small-scale croplands at national extents"
+title: "High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"
 
@@
 
-- name: \*Provisional author list\*
-  affilnum: '1'
-
-- name: Lyndon D. Estes
+- name: Lyndon Estes
   affilnum: '1'
@@
 - name: Dennis McRitchie
+  affilnum: ''
+
+- name: Stephanie Debats
+  affilnum: ''
+
+- name: Justus Muhando
   affilnum: '2'
@@
 - name: Kelly Caylor
-  affilnum: '3'
-
-- name: Stephanie Debats
-  affilnum: '4'
-
-- name: Justus Muhando
-  affilnum: '5'
-
-- name: SpatialCollective
-  affilnum: '5'
+  affilnum: '3-5'
 
@@
 - affilnum: 2
-  affil: Dennis and Sons, Tucson, AZ, USA
+  affil: SpatialCollective
 
 - affilnum: 3
-  affil: UCSB
+  affil: Department of Geography, University of California Santa Barbara, Santa Barbara, CA, USA
 
 - affilnum: 4
-  affil: Uber
+  affil: Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA, USA
 
 - affilnum: 5
-  affil: SpatialCollective
-
-- affilnum: 6
-  affil: Meridia
+  affil: Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA, USA
 
@@
   - references.bib
-  - knitcitations.bib
 csl: ecology.csl
@@
 @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts.
-The ability to answer the questions these efforts seek to address depends on data that describes the location and nature of cropland, and how that is changing over time.
-Unfortunately, the data that do exist are in many places inaccurate.
+Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time.
+Unfortunately, the data that do exist are in many places inaccurate or missing.
 Existing estimates of how much cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
 @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013].
-These discrepancies are due to the fact that cropland data in many regions are derived from remotey sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]
+Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018].
+Beyond cropland distributions, few data are available on cropland characteristics such as field size, which provide important information on agricultural systems  [@lesivEstimatingGlobalDistribution2019].
+These inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]
 
@@
 
-Cropland mapping over a region such as Africa is difficult for several reasons.
-The primary reason relates to the characteristics of the region's smallholder-dominated croplands, where field size averages 1-2 ha [@Debatsgeneralizedcomputervision2016; @FritzMappingglobalcropland2015].
-This size is small relative to the 30-250 m resolution of the sensors typically used for most landcover mapping efforts, which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979], in this case the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
+Cropland mapping over Africa is difficult for several reasons.
+The primary reason relates to the characteristics of the contintent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019].
+This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g.
+@ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979].
+In the latter case, the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
 On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons.
@@
 Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
-This spatial and temporal resolution addresses the sensor-scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+This high spatial *and* temporal resolution addresses the sensor-scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
 On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale.
-These capabilities are aleady being used to create a new generation of higher resolution (20-30 m) cropland and landcover maps for Africa and other regions [@ChenGloballandcover2015; @ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
+These capabilities are aleady being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
 
 Despite these advances, the highest resolution (<5 m) image sources are still not used to map cropland over very large extents, presumably because they are commercial and relatively high cost to acquire, in addition to the greater computational demands.
-As such, map accuracy can still be a challenge, particular for user's accuracy, which ranges between 46 and 76% for for the cropland class [e.g.
+As such, map accuracy can still be a challenge, particular for User's accuracy, which ranges between 46 and 76% for the cropland class [e.g.
 @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
@@
 Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms.
-Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large training datasets [@MaxwellImplementationmachinelearningclassification2018], particularly deep-learning methods (cite).
+Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large, high-quality training datasets [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @MaxwellImplementationmachinelearningclassification2018].
 To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g.
@@
 The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland.
-Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial image sources that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018].
-This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which are usually from a different source (e.g.
-Landsat, Sentinel; @XiongNominal30mCropland2017) and often represent a different year.
+Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial images that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018].
+This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which is usually from a different source (e.g.
+Landsat, Sentinel; @XiongNominal30mCropland2017).
 If a land change occurs in the interval between the two image sets (e.g.
 a new field was created), this can introduce error into the training data that is then passed on to the classifier.
-This source of error may be elevated in smallholder-dominated systems, where swidden practices are common (cite), or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
-<!-- - Why cropland mapping is hard -->
-<!--     - 3 problems -->
-<!--         - The target itself and its spectral variability and the spectral variability of its background -->
-<!--             - Spectral and spatial resolution of sensors -->
-<!--             - Poses a challenge for classification algorithms, which want both high resolution to be smaller than characteristic field size, but temporal contrast also -->
-
-To improve the accuracy of cropland maps over smallholder-dominated systems therefore requires an approach that meets three requirements.
-First, it should be based on high spatial and temporal resolution imagery, to be able to capture the fine grain and temporal variability of these fields.
+This source of error may be elevated in smallholder-dominated systems in the tropics, where swidden practices are common [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
+
+Improving the accuracy of cropland maps over smallholder-dominated systems thus requires an approach that meets three requirements.
+First, it should be based on high spatial and temporal resolution imagery, to be able to capture the fine grain and temporal variability of smallholders' fields.
 Second, an algorithm with suitable skill for classifying these images must be selected, and combined with the computational resources needed to process large imagery volumes.
@@
 
-<!-- - Training data dependence issues -->
-<!--     - Also interact with image resolution issues -->
-<!--     - Increasingly rely on high amounts of training data -->
-<!--     - Human interpretation error (training data quality) -->
-<!--       - Temporal mismatch with classification feedstock (Lesiv paper) -->
-<!--       - Errors in interpreting these sources of imagery -->
-<!-- - Review of who does what to solve the problem (probably integrated with problem statement) -->
-<!--   - croplands.org -->
-
-We describe here a cropland mapping system that addresses these requirements on the way to delineating field boundaries.
-The first requirement is enabled by the recent availability of cubesat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
-Although these data are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country-continent scale image mosaics to be created for multiple periods during the crop growing calendars, and this representation of intra-annual variability is more important for cropland classification than spectral depth [@Debatsgeneralizedcomputervision2016].
-Their daily revisit capacity also is important for being able to develop cloud-free, sub-annual composites in cloudy regions.
-Additionally, despite this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g.
+We describe here a cropland mapping system that follows these requirements on the way to delineating field boundaries.
+The first requirement is enabled by the recent availability of CubeSat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
+Although these data, currently collected by Planet's CubeSat fleet, are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country to continental scale image mosaics to be created for multiple periods during the crop growing calendars, and capturing the intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016].
+This daily revisit capacity is also important for developing seasonal composites in cloudy regions, where satellites with longer return intervals may fail.
+Lastly, although this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g.
 see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009].
 This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3).
-The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016] and was re-engineered to run on high performance, cloud-based computing clusters.
-This classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols [@Estesplatformcrowdsourcingcreation2016] and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error.
-The training and machine learning components are combined within an "active learning" framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017].
-Our framework automates and formalizes this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g.
-@XiongNominal30mCropland2017; others add].
-Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@DebatsIntegratingactivelearning2017].
-Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, **resulting in a further sharpening of accuracy** and resulting in a final vector-based field boundary map.
-
-We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
+
+The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020].
+The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020].
+The training and machine learning components are combined within an "active learning" [@cohnImprovingGeneralizationActive1994] framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017].
+Our framework automates this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g.
+@XiongNominal30mCropland2017].
+Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@cohnImprovingGeneralizationActive1994; @DebatsIntegratingactivelearning2017].
+Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a final vector-based field boundary map that provides important information on field geometry.
+
+We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of climates and agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
 These factors make Ghana emblematic of the substantial challenges facing cropland mapping over the broader region.
-
-<!-- - Our approach that tries to solve these three problems -->
-<!-- - Within a framework that tightly integrates tightly training data collection -->
-<!-- - And handles large-scale computation -->
-<!-- - Segmented boundaries -->
 
@@
 ## System overview 
-The mapping system is centered around four components: 1) an image acquisition and compositing component, 2) a training data collection platform, 3) a machine learning component, and 4) a segmentation component.
-The first component is applied as a once-off step to generate an image catalog covering the mapping geography, while the second component initiates an interactive image training and classification (component 3) pipeline that results in a pixel-wise posterior probability map of cropland probabilities on its final iteration (Fig.
-1).
-The final segmentation step (component 4) is then initiated and applied to both the pre-processed primary images and the posterior probability maps, resulting in vectorized field boundaries.
+The mapping system design has four primary components focused on 1) image acquisition and compositing, 2) training and validation data collection, 3) image classification, and 4) segmentation.
+The first component is applied as a once-off step to generate an image catalog covering the mapping geography.
+The second component provides tools for labelling imagery, and controls an interactive model training and classification (component 3) pipeline that produces a map of cropland probabilities for each image pixel (Figure \ref{fig:systemoverview}).
+The final segmentation step (component 4) is then initiated and applied to both the input image catalog and the posterior probability maps, resulting in vectorized field boundaries.
 Each system component comprises an individual software module designed to be implemented on cloud computing architecture, and available on a GitHub repository (see data and software availability section for details).
 
+```{r systemoverview, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
+The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
+knitr::include_graphics('figures/figure1.png')
+```
+
 ## System components
-### Imagery
-**[Note: This will require updating to reflect changes to compositing methods]**
-The primary image source used in our system is the PlanetScope Analytic surface reflectance product [@PlanetTeamPlanetApplicationProgram2017], although it can be readily adapted to apply to other images (e.g.
-Sentinel and Landsat).
+### Image compositing
+The image processing components of our system was designed to work with PlanetScope Analytic surface reflectance data [@PlanetTeamPlanetApplicationProgram2017].
 PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency.
@@
 
-```{r aois, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
-The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
-knitr::include_graphics('figures/figure1.png')
-```
-
-PlanetScope imagery is accessed via the Planet API, and an initial order is placed wherein the imagery representing all available dates in the two compositing periods and within a bounding box covering the mapping geography is collected and transferred directly to the system's cloud storage (Amazon Web Services S3).
-
-We then transform all images into an "analysis ready data" (ARD) product [@DwyerAnalysisReadyData2018] that is prepared against a tiling grid that is situated within a larger 1 degree resolution grid that covers the entire continent.
-Each 1 degree cell defines the minimum mapping area of interest (Fig.
-2A) is divided into 400 0.05 degree cells (Fig.
-2B), which provides the scale for creating image composite tiles.
-The ARD process simply clips all available imagery to the extent of each 0.05 cell it intersects, regardless of cloud cover.
+PlanetScope imagery is accessed via the Planet API, and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods.
+The imagery is collected and transferred directly to the system's cloud storage (Amazon Web Services [AWS] S3).
+
+Individual images are then transformed into "analysis ready data" (ARD) [@DwyerAnalysisReadyData2018], by cropping each downloaded scene into a 0.05 degree tile, regardless of cloud cover.
+Tiles are organized within a larger 1 degree resolution grid that covers the entire continent, which defines the minimum mapping area of interest (AOI; Figure \ref{fig:aois}).
 
@@
 
-```{r, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
+```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
 knitr::include_graphics('figures/figure2.png')
 ```
-The temporal compositing process is applied to the ARD imagery for the time period of interest, which in the case of this study is one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year.
+The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year.
 Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in same year helps to minimize differences caused by land change.
-For each pixel in each image in each ARD temporal stack for a given season, we calculate two weights, with the first being:
+For each pixel in each image in each ARD temporal stack for a given season, two weights are calculated:
 
@@
 \end{equation}
-
-And the second:
 
@@
 
-Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
-
-Once these two weights were calculated, the final composited pixel value for each of the four PlanetScope bands was calculated as:
+Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
+Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
+
+Once these two weights are calculated, the final composited pixel value for each of the four PlanetScope bands is:
 
@@
 Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
-The composited tiles were then added to the S3 store (Fig.
-1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using RasterFoundry (cite).
-This web-based rendering of the image is used in the training data platform (see **section X**).
+The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$.
+The web-rendered imagery is presented within the training data platform (next section).
 
 ### Labelling platform
-Both training and reference data were collected by a custom crowdsourcing platform, which we refer to here for simplicity's sake as *labeller*.
-The original version of *labeller* was designed to work with Amazon Web Services' Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
-The basic structure of the system remains the same, but we have re-engineered it into a standalone platform that 1) allows us to enroll and pay workers (the human supervisors) directly to conduct digitizing work, and 2) drives and assesses the overall machine learning process.
-The system runs on a Linux Virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing mapping and assessing assignments and converting them to rasterized labels for the machine learning process.
-Each *labeller* instance is focused on a specific mapping geography (e.g.
-one to several 1 degree cells in Fig.
-2A)
-
-Here we provide an overview of *labeller*'s architecture.
+Training and reference data are collected by a custom labelling platform, which was originally designed for AWS's Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
+The basic structure of the system remains the same, but we converted it into a standalone platform that allows us to enroll and pay people directly for their labelling, and is designed to control and supervise the machine learning process.
+The platform runs on a Linux virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing, assessing, and converting digitization work into rasterized labels for training a machine learning algorithm.
+Each instance of the platform focuses on a specific AOI (Figure \ref{fig:aois}A)
+
+The following sections provide an overview of the labelling platform's architecture.
 
@@
 ##### Selecting training and reference sites
-The labelling process begins with a python script that randomly selects a subset (e.g.
-100) of the 0.005 degree grid (Fig.
-2C) cells that intersect the *labeller* instance's specific area of interest (e.g.
-shaded grey cell in Fig.
-2A).
-These site names are stored in a Postgres table with the PostGIS extension that represents the mapping queue.
-Another python process converts each site into a mapping *task* that has a specified number of *assignments* (maps drawn by an individual worker) that must be completed before the task is complete.
+The labelling process begins with the random selection of a subset (e.g.
+500) of cells from a 0.005 degree grid, with the selection itself potentially split into a training and validation sample, according to predetermined proportions.
+The grid, which is nested within the tiling and larger 1 degree grids (Figure \ref{fig:aois}C) defines the spatial unit for a labelling job.
+The selected cells are placed into a queue within the platform's database, and then converted into a mapping *task* that has a specified number of *assignments* (boundaries drawn by an individual labeller) that must be completed before the task is complete.
 
 ##### Mapping assignments
-Workers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Fig.
-3).
-There the worker is presented with a white target box representing the 0.005 degree cell that was randomly selected to label, a set of digitizing tools, and different image backdrops, which include false and true color "slippy map" renderings (generated by Raster Foundry) of both the growing season and off-growing season PlanetScope images covering that cell, as well as basemap imagery provided by Bing and ESRI.
-The worker uses the polygon drawing tool to digitize the boundaries in crop fields visible within the PlanetScope overlays that intersect the grid target, following a set of pre-defined digitizing rules.
-To aid with interpretation, the worker toggles between the various seasonal renderings and the satellite base map to help form a judgement about what constitutes a field within the PlanetScope imagery.
-The worker assigns each completed polygon a label that describes the field class (e.g.
-annual cropland, tree crop, fallow), and after completing all visible fields saves the full set of polygons to the server.
-The worker is then automatically taken to their next assignment at a different randomly selected site.
-
-```{r, echo = FALSE, out.width="100%", fig.cap="An overview of mapper's interface", fig.align='center'}
+Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps.
+Following a set of pre-defined digitizing rules, the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays.
+To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
+The labeller assigns each polygon a field class category (e.g.
+annual cropland, tree crop, fallow), and upon completing all fields submits the assignment to the database.
+In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete.
+The labeller is then directed to the next available assignment from a different labelling task.
+
+```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of mapper's interface", fig.align='center'}
 knitr::include_graphics('figures/figure3.png')
@@
 ##### Processing completed assignments
-Upon submission, the workers' polygons, if there are any--the worker will submit the assignment without mapping if they see no fields in the image--are cleaned to fix any topological irregularities that arose during digitization (see supporting information (SI)) and stored in a PostGIS table.
-The assignment that was just completed represents one of two possible types of tasks: 1) an accuracy task; 2) a labeling task for training the machine learning algorithm.
-If it was the former type, an `R` script is called that executes a series of map comparisons between the worker's polygons and the reference polygons, assigning a score based on the following formula:
-
-\begin{equation}
+All submitted polygons are cleaned to fix topological irregularities that arose during digitization (see supporting information [SI]) and stored in a PostGIS table.
+Each completed assignment represents one of two types of tasks: 1) accuracy assessment, or 2) model training or validation.
+For the former type, an accuracy assessment routine is invoked that executes a series of comparisons between the labeller's results and a training reference dataset, resulting in a assignment score:
+
+\begin{equation} \label{eq:qaqc}
 \mathrm{score_i}=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}
@@
 
-Where *i* represents the particular assignment, and $\beta_{0-5}$ represent varying weights that sum to 1.
-*I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the worker's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the worker delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the worker's thematic labels (see SI for individual formulae).
-
-Over time, workers are assessed multiple times across a range of accuracy tasks that are selected to represent the variability of the agricultural system being mapped.
-Each worker's score history is averaged to provide an overall assessment of accuracy, and this information is used for creating labels, the second task.
-
-If the worker's completed assignment was a labelling task, their maps remain stored in the database until the remaining assignments needed to complete the task are submitted by other workers.
-Once complete, a `R`-based routine is invoked, which combines the maps from all workers who undertook the task into a single consensus label using a Bayesian merging approach:
+Where *i* indicates the particular assignment, and $\beta_{0-5}$ represent varying weights that sum to 1.
+*I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the labeller's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the labeller delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the labeller's thematic labels (see SI for individual formulae).
+Equation \ref{eq:qaqc} is an extension of the approach described by Estes et al.
+(2016).
+
+Over time, labellers are assessed multiple times across a range of accuracy tasks, which are selected to represent the variability of the agricultural system being mapped.
+Each labeller's score history is averaged to provide an overall accuracy measure, and this information is used for creating labels, the second task.
+
+If the labeller's completed assignment was a training/validation task, their maps remain stored in the database until the task's outstanding assignments are completed by other labellers.
+Once complete, another routine is invoked, which combines the task's completed assignments into a single consensus label using a Bayesian merging approach:
 
@@
 
-Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and $W_i$ is an individual worker.
-P($\theta$|D) is therefore the probability that the actual cover type is what the workers who labelled it says it is, while P(W$_i$|D) is the worker's average score over the accuracy assessment assignments they have mapped to date, and P(W$\theta$|D, *W$_i$*) is the worker's label for that pixel.
-This approach therefore uses the overall accuracy of each worker to weight their labels as they are combined with other workers' labels for a given pixel (see SI for further details).
-As a further measure of confidence in the final consensus label, its average Bayes Risk is calculated (see SI).
+Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the label assigned to that pixel by a labeller, and $W_i$ is an individual labeller.
+P($\theta$|D) is therefore the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$|D) is an individual labeller's average score over all the accuracy assessment assignments they have completed, and P(W$\theta$|D, *W$_i$*) is the labeller's label for that pixel.
+This approach therefore uses the overall accuracy of each labeller to weight their labels when combined with those made by other labellers' for the same pixel (see SI for further details).
+As a further measure of confidence in the final consensus label, its average Bayesian Risk can be calculated (see SI).
 This measure ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels (n = 40000) in the label, and 1 indicating complete disagreement.
-
-Once all training sites are completed, they are randomly partitioned into training and validation sets according to predetermined proportions.
 
@@
 ### Classification pipeline
-Upon completion of the labels, *labeller* automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens of instances, depending on the size of the mapping geography.
+Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens of instances, depending on the size of the AOI.
 
 #### Feature extraction
-The first step is the extraction of additional features from the growing season/non-growing season PlanetScope imagery.
-Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetatation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
-We followed this logic in this study, but opted for a smaller feature space to accomodate the storage and memory requirements of mapping geography that were several orders of magnitude larger.
-Here, we used a set of 24 features, comprising the eight original composite bands (four from each season), an 11X11 mean of each band, as well as the standard deviation of each band within a 5X5 window (Table 1).
+The first step is the extraction of additional features from each seasonal image composite.
+Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
+We followed this logic in this study, but were constrained to use a smaller feature space because the storage and memory requirements for our mapping geographies in this case were several orders of magnitude larger.
+For this implemention, we thus extract a set of 16 features, which are the 11X11 mean and 5X5 standard deviation of each band for both seasons (initial tests revealed these window sizes most effective), resulting in 24 overall features with the bands (Table 1).
 
@@
 
-Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`, `GeoPySpark`, `rasterio`, and `RasterFrames`.
-These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features (Table 1), and convert these into `RasterFrames`, which are an Apache Spark compatible data object that is designed specifically for raster data.
-Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs (COGs), which allows fast windowed reads of image subsets.
+Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`$\footnote{https://github.com/locationtech/geotrellis}$, `rasterio`$\footnote{https://rasterio.readthedocs.io/en/latest/}$, and `RasterFrames`$\footnote{https://www.cogeo.org/}$.
+These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into `RasterFrames`, a data structure designed for handling raster data within Apache Spark.
+Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs$\footnote{https://www.cogeo.org/}$ (COGs), which allows rapid reading of image subsets.
 
 #### Classification
-Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of RandomForest [@BreimanRandomForests2001].
-\textcolor{red}{The model is trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.}
-
-After fitting, the model is applied to the features of the model validation set, and a set of performance metrics are calculated, including ordinary binary accuracy, the F1 score (the geometric mean of user's and producer's accuracy), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
+Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
+For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.
+
+After fitting, the model is applied to the features of the model validation set, and a set of performance metrics are calculated, including binary accuracy, the F1 score (the geometric mean of user's and producer's accuracy), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
 
 ### The active learning loop
-After fitting and model evaluation a second prediction process is undertaken that initiates the active learning process.
-The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sets, but applied to a subset of randomly drawn pixels from each cell in order to reduce computational demand.
-The fitted model is then applied to these features in order to calculate the posterior probability of cropland presence, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:
+After fitting and model evaluation a second prediction is undertaken to enable active learning.
+The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sample, but applied to a subset of randomly drawn pixels from each cell in order to reduce computational demand.
+The fitted model is applied to predict the cropland probability for these selected pixels, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:
 
@@
 ```
-Where Q is the uncertainty for cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
-Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus images having the lowest values of $Q_I$ represent sites posing the most difficulty for the classifier.
-
-After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to *mapper*, where workers then digitize new labels.
-These new consensus labels are then added to *learner*'s initial training pool upon completion, and a new machine learning cluster is launched, resulting in a reassessment of uncertainty for unlabelled sites, a selection from the most uncertain sites, and a labelling of those sites by the workers.
-This loop repeats until the map accuracy metrics, as assessed against the validation set, show no further gains.
+Where Q is the uncertainty for grid cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
+Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus images with the lowest values of $Q$ represent sites posing the most difficulty for the classifier.
+
+After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to the labelling platform, which are then digitized by the labellers.
+The resulting consensus labels from the actively selected sample are added to the initial randomly selected sampe, and a new cluster is launched.
+The model is retrained, assesses uncertainty across the remaining unlabelled sites, and selects the next most uncertain sites for labelling.
+This loop repeats until model performance gains against the validation set show diminishing returns.
 
@@
 After the final iteration, the segmentation algorithm is invoked, which entails several steps.
-In the first step, the meanshift algorithm [@YizongChengMeanshiftmode1995a] is applied to the original bands of the growing season/off-growing season PlanetScope image pair.
-
-In the second step, a Sobel filter is then applied to the green, red, near-infrared mean-shifted bands and probability map , and an combined edge image is computed using the sum of these four edge images for the dry season only.
+In the first step, the meanshift algorithm [@YizongChengMeanshiftmode1995a] is applied to the original bands of the dry season composite.
+A Sobel filter is then tapplied to the green, red, and near-infrared mean-shifted bands and the probability map, and an combined edge image is computed using the sum of these four edge images for the dry season only.
 A compact watershed algorithm [@neubertCompactWatershedPreemptive2014] is then run on the weighted edge image, with a high level of segmentation specified.
@@
 
-In the third step, a region adjacency graph is constructed for each tile, in which each node represents all pixels within each polygon created in the previous step.
+Third, a region adjacency graph is constructed for each image tile, in which each node represents all pixels within each polygon created in the previous step.
 The edge between two adjacent regions (polygons) is calculated as the norm of the difference between the means of normalized colors of all bands.
@@
 
-In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (typically 0.5, but can vary locally) are retained as field polygons.
+In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined  threshold (here 0.5, but could vary locally) are retained as field polygons.
 
 In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@visvalingamLineGeneralisationRepeated1993].
-Neighboring polygon that touch along tile boundaries are merged.
+Neighboring polygons that overlap along tile boundaries are then merged.
 
@@
 Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map.
-[**Note: the following maybe better for discussion**] We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@yeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and human interpreter are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
-On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of a "field" may vary between farmers (**cite**) and their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries (**see SI**).
-
-## Application and evaluation of the system
-To create a map of Ghana, we divided the country into 16 distinct mapping zones, or areas of interest (AOIs), which were developed by grouping together each one degree cell (Figure 2A) fully contained with the tiles (Figure 2B) belonging to any adjacent one degree cells that overlapped with neighboring countries.
-The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), and each was used to run a separate active learning process and to create a final segmented map.
-
-Prior to initiating active learning in each AOI, we grouped them into three zones: a northern zone comprising the 6 northernmost AOIs, a central to southeastern zone comprising the 3 middle and 3 southeastern AOIs, and a southwestern zone made up of the AOIs falling in the forest zone (see Figure SX in SI).
-Within each zone group, we randomly selected 500 points for initial labelling by the mapping team, which were used to train the first iteration of each RandomForest model run for the zone group.
-At the outset of mapping in each AOI, the mapping team labelled 100 randomly selected points within each AOI that were set aside for model validation, and subsequent iterations targeted new training sites for collection in each AOI based on classification uncertainty.
-
-To evaluate the results, we undertook a number of analyses.
+We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@yeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and labellers are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
+On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of what constitutes a "field" could vary between farmers, while their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries.
+
+## Applying the system to map Ghana
+To create a map of Ghana, we divided the country into 16 distinct AOIs, which were developed by grouping together each one degree cell fully contained within Ghana with the tiles belonging to any adjacent degree cell that overlapped neighboring countries (Figure \ref{fig:aois}A).
+The exception was AOI 16, which consisted of the four degree cells along Ghana's southern coast.
+The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), A separate active learning and segmentation process was run for each of these AOIs.
+
+To collect the initial randomized samples for model training, we grouped the AOIs into three blocks: a northern block comprising the 6 northernmost AOIs, a central to southeastern blocks comprising the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern block made up of the forest zone AOIs (10, 11, 13, 14, 16).
+Within each block, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort.
+In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.
+
+After collecting the initial training and validation samples, we trained a starter model for each block and applied it to each of the block's AOI.
+For each iteration, 100 samples were actively selected within each AOI, and added to the training pool.
+
+To evaluate the performance of the system, we undertook several different analyses.
 
 ### Image quality
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 data("image_quality")
-tile_quality <- image_quality %>% filter(kml_type == "F") %>% 
+tile_quality <- image_quality$img %>% filter(kml_type == "F") %>% 
   group_by(name, season, worker_id) %>% 
@@
 
-We evaluated the overall quality of the resulting seasonal image composites across the region using a random selection of `r tilen`, which we graded according to a four category quality score that evaluated the degree of residual 1) cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image in which resolution was degraded below the 3-4 m PlanetScope resolution (e.g.
+We evaluated the overall quality of the resulting seasonal image composites by assessing a random selection of `r tilen` tiles.
+We graded both seasonal composites for each tile using a four category quality score, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its  resolution was degraded below the typical 3-4 m PlanetScope resolution (e.g.
 because of between-date image mis-registrations).
@@
 
-### Map assessment
-#### Validation performance
-To assess the gain in RandomForest performance due to active learning, we assessed the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
-To assess whether active learning improved the performance relative to a purely random approach to selecting new training sites, we ran an additional test at a subset of AOIs (1, 8, and 15).
+### Model gains per iteration
+To assess the gain in model performance due to active learning, we measured the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
+
+To evaluate whether active learning improved model performance relative to a purely random approach to selecting new training sites, we ran additional tests within a subset of AOIs (1, 8, and 15).
 We first randomly selected and labelled 300-400 sites in each AOI.
-After initial training of AOIs 1, 8, and 15 with their initial pools of 500 training sites (Cluster 1 for AOI 1 and Cluster 2 for AOIs 8 and 15), we then retrained the model for each AOI by progressively adding 100 of the new randomly selected labels, evaluating model performance against the validation set after each retraining.
-We repeated this process so that the number of retrainings and training samples matched those from the active learning process for each AOI.
-We then compared the difference in accuracy, AUC, and F1 between these randomly trained models and those trained with active learning [@DebatsIntegratingactivelearning2017].
-
-#### Evaluating the impact of training data error
-To quantify the potential impact of training data error on our map results, we compared maps created by training the RandomForest algorithm with three sets of labels: 1) those from the lowest scoring worker to map each training site, 2) those from the most accurate worker, and 3) those trained as usual with the consensus labels.
-We then quantified the prediction performance against the validation sites, and compared the differences in performance metrics for each variant.
+We then progressively added 100 of the randomly selected samples to the relevant training pool and retrained the model,  repeating the process so that the number of iterations and samples matched those from the active learning process.
+We then compared the difference in accuracy, AUC, and F1 between the randomly trained models and those trained with active learning [@DebatsIntegratingactivelearning2017].
+
+### Accounting for label error
+To quantify the potential impact of label error on our results, we evaluated the performance differences between models trained with three different sets of labels: 1) those from the lowest scoring labeller to map each training site, 2) those from the highest scoring labeller, and 3) the consensus labels.
 This assessment follows recommended Tier 1 (i.e.
-best practice) standards to account for training data errors in machine learning-based Earth Observation applications [@ElmesAccountingtrainingdata2020].
+best practice) standards to account for training data errors [@ElmesAccountingtrainingdata2020].
 
@@
 
-As a final test of training data quality on map accuracy, we retrained the model for AOI 8 for 2018 using new labels that were generated against an updated training reference library (n = `r kml_types`), and compared the performance of this model against the previous version of the model.
-
-#### Accuracy assessment
-The accuracy assessment procedure described above was not strictly independent because it was re-used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020].
-To independently assess the accuracy of final map products, we followed guidelines provided by [@StehmanKeyissuesrigorous2019] to create a separate map reference sample.
-To develop this sample, we first used the segmented field boundaries to define a cropland and non-cropland stratum for Ghana, and then randomly selected a sample of the centroids of grid cells intersecting each stratum.
-Next, a primary supervisor examined the image composites in each randomly selected cell, and placed a square polygon of ~0.1 ha on the primary class intersecting the centerpoint of the cell.
-The square was shifted so that it was contained entirety with the intersected class: cropland or non-cropland.
-Where the supervisor was uncertain about the class label, one of two additional classes were selected: likely cropland or likely non-cropland.
-After placing these polygons, the supervisor removed their classes from the polygons and sent them to three separate experts, who labelled the polygons according to their interpretation of the underlying imagery.
-The four sets of labels were then combined and the overall agreement between them was calculated.
-This agreement provided the upper bound on knowable map accuracy.
-See the SI for further details on the creation and analysis of the map reference sample.
-These points were then intersected with both the RandomForest probability images and the segmented field boundaries, and confusion matrixes between the map reference sample labels and the extracted map classes was constructed to assess the categorical accuracy of each map product.
+### Accuracy assessment
+The model performance assessments described above (2.3.2-3) were not fully independent because they used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020].
+To independently assess the accuracy of our final map products, we followed recommended guidelines [@StehmanKeyissuesrigorous2019] to create a separate map reference sample.
+To develop this sample, we first defined cropland and non-cropland strata for Ghana by intersecting the centroids of the 0.005 degree grid (Figure \ref{fig:aois}C) with the segmented field boundaries, and then randomly selected samples from the points representing each stratum.
+The selected sample was loaded into labelling platform, were it was presented as a ~0.1 ha square polygon (representing the minimum mapping unit).
+An expert supervisor who was not involved in labelling evaluated the PlanetScope composites at each site to determine its class membership.
+If the 0.1 ha target was not completely contained within a single class, the supervisor drew a square polygon of ~0.1 ha and shifted it the minimum distance necessary to make it contained entirety within the class that the initial target had the greatest overlap.
+Four classes were used for the map reference sample: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
+The latter two classes were used to provide insight into the degree of uncertainty in the map reference sample.
+For efficiency, two separate supervisors evaluated separate portions of the reference sample, but both assessed a small subset.
+We calculated their level of agreement on this subset to provide an additional assessment of uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019].
+The SI contains further details on the creation and analysis of the map reference sample.
+
+The map reference polygons were then intersected with both the probability images and the segmented field boundaries, and confusion matrixes between the map reference labels and the extracted map classes were constructed to assess the categorical accuracy of each map product.
+We calculated the overall accuracy for each map, as well as the class-wise User's and Producer's accuracy, as well as the 95% confidence intervals for each accuracy measure [@OlofssonGoodpracticesestimating2014; @OlofssonMakingbetteruse2013; @StehmanKeyissuesrigorous2019].
 
@@
 ## Assessing the characteristics of Ghanaian cropland
-Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the characteristics of field sizes through the country.
-We also used the uncertainty information encoded in the cropland probability images produced by the Random Forest classifier to provide further insight into the characteristics of cropland active-fallow cycles.
+Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size of fields and number of fields in the different AOIs.
+We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the biases between labellers' polygons and segmented boundaries at validation labels to calculate adjusted estimates of mean field sizes and the total number of fields.
 
@@
 rfa <- accuracy_reports$ghana$rf[, "Area"]
-rfa <- rfrep[nchar(rfa) > 0]
+rfa <- rfa[nchar(rfa) > 0]
 rfas <- as.numeric(gsub("\\(.*", "", rfa[nchar(rfa) > 0]))
@@
 sega <- accuracy_reports$ghana$seg[, "Area"]
-sega <- segrep[nchar(sega) > 0]
+sega <- sega[nchar(sega) > 0]
 segas <- as.numeric(gsub("\\(.*", "", sega[nchar(sega) > 0]))
@@
 Efforts to map fields in small-scale agricultural systems have either been field-based [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small areas [e.g.
-@forkuorIntegrationOpticalSynthetic2014].
+@forkuorIntegrationOpticalSynthetic2014] (**include Persello**).
 The most notable efforts to estimate field sizes were two global initiatives that used crowdsourced volunteers to categorize five broad field size classes based on their interpretations of imagery sampled from virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
@@
 These include the correspondence between image- and ground-collected labels.
-We believe that such comparisons will in most circumstances will reveal unresolvable discrepancies.
+We believe that such comparisons will in most circumstances reveal unresolvable discrepancies.
 The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels can produce substantial differences.
@@
 The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund.
+Support for computing was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative.
 Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393).

### 2021-03-02 inst/manuscript/manuscript.Rmd
Edited Results and SI

@@
 
+fmt <- function(x) format(x, big.mark = ",")
 #opts_knit$set(root.dir=normalizePath('../'))
@@
 Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps.
-Following a set of pre-defined digitizing rules, the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays.
+Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays.
 To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
@@
 
-To collect the initial randomized samples for model training, we grouped the AOIs into three blocks: a northern block comprising the 6 northernmost AOIs, a central to southeastern blocks comprising the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern block made up of the forest zone AOIs (10, 11, 13, 14, 16).
-Within each block, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort.
+To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16).
+Within each cluster, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort.
 In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.
 
-After collecting the initial training and validation samples, we trained a starter model for each block and applied it to each of the block's AOI.
+After collecting the initial training and validation samples, we trained a starter model for each cluster and applied it to each of the block's AOI.
 For each iteration, 100 samples were actively selected within each AOI, and added to the training pool.
@@
 To independently assess the accuracy of our final map products, we followed recommended guidelines [@StehmanKeyissuesrigorous2019] to create a separate map reference sample.
-To develop this sample, we first defined cropland and non-cropland strata for Ghana by intersecting the centroids of the 0.005 degree grid (Figure \ref{fig:aois}C) with the segmented field boundaries, and then randomly selected samples from the points representing each stratum.
-The selected sample was loaded into labelling platform, were it was presented as a ~0.1 ha square polygon (representing the minimum mapping unit).
-An expert supervisor who was not involved in labelling evaluated the PlanetScope composites at each site to determine its class membership.
-If the 0.1 ha target was not completely contained within a single class, the supervisor drew a square polygon of ~0.1 ha and shifted it the minimum distance necessary to make it contained entirety within the class that the initial target had the greatest overlap.
+We used a stratified design, randomly assigning square polygons of ~0.1 ha extent into cropland and non-cropland strata, developed the map of segmented field boundaries.
 Four classes were used for the map reference sample: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
 The latter two classes were used to provide insight into the degree of uncertainty in the map reference sample.
-For efficiency, two separate supervisors evaluated separate portions of the reference sample, but both assessed a small subset.
+For efficiency, two separate supervisors evaluated separate portions of the reference sample, but both jointly assessed a small subset of the sample.
 We calculated their level of agreement on this subset to provide an additional assessment of uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019].
-The SI contains further details on the creation and analysis of the map reference sample.
+The SI contains further details on the design and collection of the map reference sample.
 
@@
 ## Assessing the characteristics of Ghanaian cropland
-Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size of fields and number of fields in the different AOIs.
-We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the biases between labellers' polygons and segmented boundaries at validation labels to calculate adjusted estimates of mean field sizes and the total number of fields.
+Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
+We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the differences between labellers' polygons and segmented boundaries at validation sites to calculate bias-adjusted estimates of mean field sizes and the total number of fields.
 
@@
 km2 <- round(as.numeric(units::set_units(sum(sf::st_area(tiles)), "km2")))
-km2 <- format(km2, big.mark = ",", scientific = FALSE)
-```
-We developed maps for all of Ghana (`r km2` km$^2$).
+km2 <- fmt(km2)
+```
+We developed maps of Ghana's arable croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles.
 Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
-For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or replanted tree crops.
-We made this decision because the resolution of PlanetScope imagery is insufficient to distinguish tree crops from the surrounding forest, and clear field boundaries are often not visible within cocoa crops.
+For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops.
+We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g.
+cocoa) are often not visible.
 
 ## Image catalog and quality
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 data("image_quality")
@@
 names(tqct) <- tqual_ct %>% pull(season)
-```
-
-We generated a complete image catalog for the `r nrow(tiles)` tiles intersecting Ghana, spanning `r km2` km$^2$.
-This entailed processing all PlanetScope imagery intersecting the total extent of these tiles for the periods May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
-The longer period was necessary for the growing season because of the high frequency of cloud cover, which severely limits the number of clear scenes over any given area.
-For cloudier regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
-The compositing processes was applied to render the resulting seasonally contrasting image pairs for each tile.
-We also composited imagery for AOI 8 (Figure \ref{fig:aois}A) for the two seasons of the 2019 agricultural year.
-All images were converted into Raster Foundry slippy maps for presentation in *labeller*.
-
-An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["gs"]`, with `r tqct["gs"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["os"]` (`r tqct["os"]` percent $\geq$ 0.85).
+
+nt <- fmt(nrow(tiles))
+```
+
+To develop the map, we first generated an image catalog for the `r nt` tiles covering Ghana.
+This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
+The longer period was necessary for the growing season because of the frequent cloud cover, which severely limits the number of clear scenes for any tile.
+For the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
+
+An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85).
 Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period.
@@
 
-```{r imqual, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D) [This a hypothetical placeholder figure].", fig.align='center'}
+```{r imqual, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D).", fig.align='center'}
 knitr::include_graphics('figures/figure4.png')
@@
 ### Training data collection
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
@@
   filter(usage == "validate") %>% distinct(name) %>% count()
-```
-
-After initial training, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (Figure S2A).
-AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement, thus were trained with `r aoi10n` -  `r aoi14n` samples.
-AOI 15 was run for 4 iterations and had a training sample size of `r aoi15n`, while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
-In this second run, `r aoi3init`  initial training sites randomly selected within the AOI were used (Figure S2A), followed by 2 subsequent active learning subsequent iterations, resulting in a training sample of `r aoi3n`.
-The total number of unique training and validation sites across the country were `r initial_total + active_total` and `r validate_total`, respectively.
+
+ttrain <- fmt(initial_total + active_total)
+tval <- fmt(validate_total$n)
+```
+
+After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (see Figure S3A for sample distributions).
+AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement.
+The models for these two AOIs were thus trained with `r aoi10n` -  `r aoi14n` samples.
+AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
+In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
+The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.
 
@@
   mutate(trate = tref / tr) %>% filter(tr > 10)
-```
-
-The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core team of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each.
-As each training site was mapped by 4 separate labellers, `r tsites$tot[1]` sets of vector labels were made.
-Each labeller digitized an average of `r tsites$mu[1]` (see Figure S3A for more details on labelling effort).
-
-This accuracy of the labellers was assessed `r tsites$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
-The average of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S3B for detailed score distributions).
+
+tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
+# nvecs <- tsites$tot[1])
+# muvecs <- fmt(tsites$mu[1])
+
+```
+
+The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each.
+As each training site was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made.
+Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S4A for more details on labelling effort).
+Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
+The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S4B for detailed score distributions).
 
@@
 The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
-Risk was highest in the northern cluster (AOIs 1-6; Figure S5), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figure S6), and lowest in the southwestern cluster (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
-Label risk in the central/eastern cluster (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
-Greater labeller experience also appeared to reduce risk, as the mean risk for the second mapping of Cluster 2's 500 initial training sites was `r cl2brisk`, compared to `r cl2arisk` for the first mapping.
-
-**[To further evaluate label quality, we assessed the degree of similarity between worker polygons with the field boundaries collected by our ground validation team.
-Overall, we found that on average a worker's field boundary interpreted from a PlanetScope image had only an *X%* agreement (as determined by area of intersection over union) with the corresponding field-mapped boundary, and the average field size mapped by a worker was on average *X% smaller/larger* than those mapped on the ground.]**
+Risk was highest in the northen AOIs (AOIs 1-6; Figures S5-6), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S5-6), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
+Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
+Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster (see SI); the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.
 
 ### Validation performance
-```{r, echo=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 data("iteration_metrics")
@@
 
-The gain in model performance was assessed for each AOI against several accuracy metrics.
+Model performance was calculated for each iteration within each AOI.
 The average accuracy, AUC, and F1 at iteration 0 were `r imu0$Accuracy$score`, `r imu0$AUC$score`, and `r imu0$F1$score`, respectively, increasing to `r imu3$Accuracy$score`, `r imu3$AUC$score`, and `r imu3$F1$score` by iteration 3 (Figure \ref{fig:alperformance}).
 These differences represent respective gains of `r imu3$Accuracy$diff`, `r imu3$AUC$diff`, and `r imu3$F1$diff` percent for the three metrics.
-The largest gains in each metric were registered on iteration 1, averaging `r idmu$I1$Accuracy$score`, `r idmu$I1$AUC$score`, and `r idmu$I1$F1$score` percent for accuracy, AUC, and F1.
-The lowest gains were received on Iteration 3 for accuracy (`r idmu$I3$Accuracy$score`%) and F1 (`r idmu$I3$F1$score`%), while AUC gains were least on iteration 2 (`r idmu$I2$AUC$score`%).
-The range in scores varied substantially across AOIs and metrics.
+The largest gains for each metric occurred on iteration 1, averaging `r idmu$I1$Accuracy$score`, `r idmu$I1$AUC$score`, and `r idmu$I1$F1$score` percent for accuracy, AUC, and F1, while the lowest gains were realized on iteration 3, with accuracy, F1, and AUC respectively increasing by just `r idmu$I3$Accuracy$score`%, `r idmu$I3$F1$score`%, and `r idmu$I2$AUC$score`%.
+The scores achieved on the final iteration varied substantially across AOIs and metrics.
 Accuracy ranged between `r imin[imin$Metric == "Accuracy", ]$Score` (AOI `r imin[imin$Metric == "Accuracy", ]$AOI`) and `r imax[imax$Metric == "Accuracy", ]$Score` (AOI `r imax[imax$Metric == "Accuracy", ]$AOI`), while AUC varied from `r imin[imin$Metric == "AUC", ]$Score` (AOI `r imin[imin$Metric == "AUC", ]$AOI`) and `r imax[imax$Metric == "AUC", ]$Score` (AOI `r imax[imax$Metric == "AUC", ]$AOI`), and F1 from `r imin[imin$Metric == "F1", ]$Score` (AOI `r imin[imin$Metric == "F1", ]$AOI`) and `r imax[imax$Metric == "F1", ]$Score` (AOI `r imax[imax$Metric == "F1", ]$AOI`).
 
-```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting for the Random Forest model after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
+```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/figure5.png')
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # random versus active data
@@
 
-The training tests at AOIs 1, 8, and 15 showed that active learning produced higher performance than randomized model training after three iterations, with accuracy, AUC, and F1 being respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher after three iterations (Figure S5).
+The comparison of active versus randomized training sample collection (in AOIs 1, 8, and 15) showed that the former approach outperformed the latter.
+After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S7).
 However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2.
-The lower accuracy under active leaving was caused by results at AOI 15, where accuracy gain was `r abs(u(i1acc["15"]))` percent lower under active training (although accuracy slightly between iterations; see Figure \ref{fig:alperformance}).
-In comparison, active learning accuracies after iteration 1 were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
+The negative results for accuracy was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:alperformance}).
+In comparison, iteration 1 active learning accuracies 1 were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
+Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.
 
@@
 
-The potential impact of training data errors on map quality was assessed in four AOIs (1, 2, 8, and 15).
+The potential impact of label errors on map quality was assessed in four AOIs (1, 2, 8, and 15).
 The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
-Models trained with the most accurate worker's labels had scores that were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than models trained with the most accurate worker's lablers (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate worker's labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
-
-```{r trainingimpact, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from Random Forest models trained with consensus labels, and labels made by the most and least accurate worker to label each site.
+Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
+
+```{r trainingimpact, echo = FALSE, out.width="100%", fig.cap="Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from models trained with consensus labels, and labels made by the most and least accurate labellers to map each site.
 Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.", fig.align='center', message=FALSE}
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 data("label_risk_metrics")
@@
 
-A second measure of the impact of training data error can be obtained by calculating the between the average training label risk per AOI versus various model performance metrics (Table S3).
-Accuracy had a strong negative association (Spearman's Rank Correlation =  `r risk_corr[, "Accuracy"]` and AUC had a moderate negative association (r = `r risk_corr[, "AUC"]`), while F1 had a weaker positive association (r = `r risk_corr[, "F1"]`).
-This sign of the latter is counter-intuitive, but relates to the fact that precision was higher in areas of increasing label risk (`r risk_corr[, "Precision"]`), as did false positive error rates (`r risk_corr[, "FPR"]`).
-In contrast, recall had little correlation with label risk (`r risk_corr[, "Recall"]`), thus the higher precision drove F1 higher, even in areas with more false positives.
-**[work on this]**
+A second measure of the impact of label error is found within the correlations between the mean label risk per AOI and the model performance metrics (Table S3).
+Accuracy and AUC had strong (Spearman's Rank Correlation =  `r risk_corr[, "Accuracy"]` to moderate (r = `r risk_corr[, "AUC"]`) negative correlations with label risk, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
+The positive sign of the latter relationship is counter-intuitive, but is explained by risk's association with precision, one of two inputs to F1, which was moderately positive (`r risk_corr[, "Precision"]`), whereas risk had a negligible correlation with recall (`r risk_corr[, "Recall"]`), F1's other component.
+The correlation between risk and the false positive rate (`r risk_corr[, "FPR"]`), another important performance metric, shows that labelling uncertainty may increase model commission error.
 
@@
 
-To assess the overall accuracy of the final maps, we used our map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of 1) pixel-wise classifications produced by the RandomForest model, and 2) the segmented field boundary maps.
-We first evaluated the uncertainty in the map reference classes using two approaches, the first being an assessment of the overall agreement between map reference labels (at `r sum(usertab)` sites), and the second was the level of confidence in assigned labels (see SI for details).
-The first measure showed that the two individual labellers agreed `r agreement`% of sites, while the second showed that `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites labelled were of lower confidence.
-
-Based on this reference sample, we found that the overall accuracy of the pixel-wise classifications was 88% (Table \ref{tab:mapaccuracy}).
-This overall level of accuracy varied within four smaller zones (Figure S9A), ranging from 83.3% in Zone 1 (comprised of AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
-Examining the accuracies of the individual classes, the Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1.
-The User's accuracy for cropland was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1.
-Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%, with the lowest accuracies for this class found in Zone 1 (Producer's = 89.3% and User's = 87.7%, respectively).
+We used a map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of the per-pixel classifications (resulting from thresholding the Random Forests probability), as well as the segmented field boundary maps.
+We first evaluated the uncertainty in the map reference classes by assessing 1) the overall agreement between map reference labels collected by two separate supervisors at `r sum(usertab)` sites, and 2) the confidence of the labels assigned by the supervisors (see SI for details).
+The first measure showed that the two individual supervisors' labels agreed at `r agreement`% of common sites, while the second showed that `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites were labelled with the two classes that indicated a level of uncertainty.
+
+We found that the overall accuracy of the pixel-wise classifications was 88% against this map reference sample (Table \ref{tab:mapaccuracy}).
+Confining the map reference sample to four distinct zones (Figure S9A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
+The Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1, while the User's accuracy for was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1.
+Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%.
+The lowest accuracies for the non-cropland class was in Zone 1 (Producer's = 89.3%; User's = 87.7%).
 
 The overall accuracies obtained from the segmented maps were generally 1-2 percentage points lower than those of the per-pixel maps, while User's accuracies tended to be 8-10 percentage points less (Table \ref{tab:mapaccuracy}).
-In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map, indicating that the segmentation step helped reduce omission error, but substantially increased false positive errors.
+In contrast, Producer's accuracies were 15-20 points higher than in the per-pixel map.
+The segmentation step therefore helped to reduce omission error while substantially increasing commission error.
 
@@
 Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country.
-The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.
+The error matrix (with reference values in columns) provides the areal percentage for each cell, and the Producer's (P), User's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.
 }
@@
 ```
-The comparisons of digitized versus segmented field boundaries showed that the mean field size across all 1600 validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
-This discrepancy was primarily caused by segmentation results in four AOIs (2, 3, 7, and 15; Figure S5), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
-The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S6).
+The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
+This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S10), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
+The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S11).
 
@@
 ## Ghana's croplands
-Two separate maps of cropland were produced for each AOI, the per-pixel classified from the cropland probabilities, and the vectorized map of field boundaries (Figure \ref{fig:mainmap}).
-The former map provides the more accurate assessment of the distribution of croplands in Ghana, which are most concentrated in the Southeastern corner (AOI 15), the central-western region (AOI 7, the northeastern and northwestern corners of AOIs 10 and 11, and the south of AOI 8), the northeastern quadrant stretching from AOI 9 through 5 and 6 up to 2 and 3.
+Two separate maps of cropland were produced for each AOI, a per-pixel map derived from the cropland probabilities, and the vectorized map of field boundaries (Figure \ref{fig:mainmap}).
+The former provides the more picture of cropland distributions in Ghana, which are most concentrated in the Southeastern corner (AOI 15), the central-western region (AOI 7, the northeastern and northwestern corners of AOIs 10 and 11, and the south of AOI 8), and the northeastern quadrant stretching from AOI 9 through AOIs 5 and 6 and up to AOIs 2 and 3.
 The northern third of AOI 1 also has noticeable densities of cropland.
 Several noteable areas of low cropland density are also apparent, indicating the presence of large protected areas, such as Mole National Park in the the southeastern corner of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
-In contrast, the relative absence of cropland in AOIs 13, 14, and 16 do not reflect the scarcity of agriculture, but rather the predominance of tree-based agriculture in the forested agroecoregions of Ghana, which was not mapped.
+In contrast, the relative absence of cropland in AOIs 13, 14, and 16 does not reflect the scarcity of agriculture, but rather the predominance of tree crops, which we did not map, in these areas.
 
@@
 The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities.
-The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles.
+The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original image resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles.
 Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
@@
 
-fmt <- function(x) format(x, big.mark = ",")
 ```
@@
 
-The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in area such as AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
+The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
 To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S12).
-These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally aligning in the areas of higher field densities, although the biases inherent in both measures (Figures S10-11) complicate the interpretation of those variations.
-However, we can use the estimated to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana as a whole (Table \ref{tab:sizentab}).
-The bias-adjusted estimates show that the typical field size in Ghana is 1.73 ha, ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern half of the country (Table \ref{tab:sizentab}).
+These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally occurring in areas of higher field density, although the biases inherent in both measures (Figures S10-11) complicate the interpretation of those variations.
+However, we can use the estimated biases to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana overall (Table \ref{tab:sizentab}).
+These adjusted estimates show that the typical field size in Ghana is 1.73 ha, ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern half of the country (Table \ref{tab:sizentab}).
 The total number of fields was estimated to be 1,662,281 overall, or 205 fields per tile on average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.

### 2021-03-02 inst/manuscript/si.Rmd
MS edited through methods
Edited Results and SI

@@
 ---
-title: "Supporting Information: Improving cropland maps through tight integration of human and machine intelligence"
+title: "Supporting Information: High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"
 
@@
 ## Mapping platform
-
 ### Digitizing tools
@@
 Upon submission, the PostGIS [`ST_MakeValid`](https://www.postgis.net/docs/ST_MakeValid.html) function is applied to each polygon's geometry to clean remaining topological errors upon insertion into the database.
+
+### Mapping rules
+Labellers were asked to follow a particular set of rules when digitizing field boundaries.
+The following rules are reproduced (with some modification to wording) from a [website](https://github.com/agroimpacts/mappingrules) that we developed to host the rules.
+
+#### Rules for interpreting and mapping imagery
+
+- **Rule 1**: Map the fields that are visible in the Planet imagery, not the fields that you see in the basemap.
+
+- **Rule 2**: Please map the fields as they appear in the off-season Planet images. If it is too hard to make these out clearly, the next choice is to map them in growing season imagery.
+
+- **Rule 3**: Use all available images (Planet scenes and base map) to help you interpret what is and is not a field.
+Toggling back and forth between the different image sources can help you identify and classify the field boundaries.
+
+- **Rule 4**: Map fields that intersect the mapping target, making sure to finish the entire boundary.
+
+#### Rules for classifying different types of agriculture
+
+- **Rule 5**: Choose the annual croplands category if the fields looks like they contain crops that are planted and harvested during a single year (or occasionally after a few seasons).
+
+- **Rule 6**: Fields that look like they are under tree crops, such as orchards, woodlots, or other planted forest types are agriculture, but we will not map these in the current campaign.
+Therefore, do not draw polygons around these types of fields.
+
+- **Rule 7**: Choose the fallow category if the fields looks like it was an annual cropland that is somewhat overgrown, and hasn't been planted in a few seasons.
+[Note: this category was eventually not used, as such fields were not mapped]
+
+- **Rule 8**: Choose the field, but unsure of category if you can't tell whether the field is annual cropland, a tree crop, or fallow.
+Try hard to avoid this category.
+[Note: this category was also not used]
+
+- **Rule 9**: Choose the unsure if agricultural field or not category if you think the field might be agriculture (whether annual cropland, tree crops, or fallow), but it could also be non-agricultural land.
+Try hard to avoid this category also.
+[Note: this category was also not used]
+
+- **Rule 10**: Don't map any polygons if there is no agricultural to map, or if the fields look like they are tree crops (see updated Rule 6).
+
+### Assessing label accuracy
+For each accuracy assessment assignment, the labeller's maps are scored against  training reference polygons digitized by one of the map supervisors (e.g.
+Estes or Ye).
+A proportion of training reference sites were of the non-cropland class and thus had no corresponding polygons.
+
+As described in the main text, the score for a particular assignment *i* is calculated as the weighted sum of five accuracy metrics:
+
+$$ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}$$
+
+Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.
+
+*I* is "inside the box" accuracy which is defined as a proportion of the area correctly mapped within a 0.005 degree resolution grid (*I$_c$*) over total "inside the box" area for this grid (*I$_t$*).
+
+$$ I=\frac{I_c}{I_t}$$
+
+*O* is "outside the box" accuracy which refers to a proportion of the field area correctly mapped outside the grid over total "outside the box" region (*O$_t$*, the region within the bounding box of the workers' polygons but not within 0.005 degree resolution grid).
+
+$$ O=\frac{O_c}{O_t}$$
+
+*F* is the fragmentation accuracy which is defined as a proportion of matched polygon number (*N$_m$*, the number of the workers' polygons that has at least 50% of its region overlapped with a reference polygon) over total workers' polygon number (*N$_t$*).
+
+$$ F=\frac{N_m}{N_t}$$
+
+*E* is the average edge accuracy for all pairs of matched workers and reference polygons; the edge accuracy for a single pair is defined as the length of 'correctly mapped edges' (*L$_c$*, the partial boundary of a workers' polygon that are within a three-pixel buffer region of the matched reference polygon boundary) over the total boundary length of its matched reference polygon (*L$_t$*).
+
+$$ E=\frac{L_c}{L_t}$$
+
+*C* is the categorical accuracy, i.e., a proportion of the area that has been correctly labeled with field category within intersected regions between worker's and reference polygons (*T$_c$*) over the total intersecting area (*T$_t$*).
+
+$$ C=\frac{T_c}{T_t}$$
 
@@
 Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the worker's label of that field, and *W$_i$* is an individual worker.
-Looking in greater details at this equation, the first half of the righthand side of the equation, P(W$_i$|D), is the "prior" for worker *i* for the current site based on their history of scores from prior accuracy assessment assignments.
+Looking in greater details at this equation, the first half of the righthand side of the equation, P(W$_i$|D), is the "prior" for worker *i* for the current site based on their history of scores from accuracy assessment assignments.
 The second term, P($\theta$|D, W$_i$), is the probability that the actual class of the pixel in the current assignment is the class that worker *i*'s says that it is, which is either 0 or 1.
@@
 
+Accuracy assessment and consensus label generation is performed using an `R` package `rmapaccuracy` that is internal to the labelling platform.
+The package relies on spatial data classes and operations provided by the `sf` [@pebesmaSimpleFeaturesStandardized2018] and `raster` [@hijmansRasterGeographicData2020] packages.
+
 #### Example
-
 To provide an example of this approach in practice, we'll imagine two workers A and B, each with the following histories:
@@
 
-#### Worker's assignment evaluation
-We assigned a score for each workers' assign; the score history determines the weights of each workers for creating a consensus label for each 0.005 degree resolution grid.
-The score for a particular assignment *i* is defined as a weighted sum of five accuracy metrics:
-
-$$ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}$$
-
-Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.
-
-*I* is "inside the box" accuracy which is defined as a proportion of the area correctly mapped within a 0.005 degree resolution grid (*I$_c$*) over total "inside the box" area for this grid (*I$_t$*).
-
-$$ I=\frac{I_c}{I_t}$$
-
-*O* is "outside the box" accuracy which refers to a proportion of the field area correctly mapped outside the grid over total "outside the box" region (*O$_t$*, the region within the bounding box of the workers' polygons but not within 0.005 degree resolution grid).
-
-$$ O=\frac{O_c}{O_t}$$
-
-*F* is the fragmentation accuracy which is defined as a proportion of matched polygon number (*N$_m$*, the number of the workers' polygons that has at least 50% of its region overlapped with a reference polygon) over total workers' polygon number (*N$_t$*).
-
-$$ F=\frac{N_m}{N_t}$$
-
-*E* is the average edge accuracy for all pairs of matched workers and refernece polygons; the edge accuracy for a single pair is defined as the length of 'correctly mapped edges' (*L$_c$*, the partial boundary of a workers' polygon that are within a three-planet-pixel buffer region of the matched reference polygon boundary) over the total boundary length of its matched reference polygon (*L$_t$*).
-
-$$ E=\frac{L_c}{L_t}$$
-
-*C* is the categorical accuracy, i.e., a proportion of the area that has been correctly labeled with field category within intersected regions between worker's and reference polygons (*T$_c$*) over the total intersecting area (*T$_t$*).
-
-$$ C=\frac{T_c}{T_t}$$
-
 #### Map reference sample design
-A reference sample should ideally have some randomization involved in the sample selection protocol, and the probabilities for a sample being included should be provided.
-There are multiple sample designs that can be chosen, but we employed a stratified sampling design, using the segmented field boundaries to define the strata for field/non-field.
+We employed a stratified design for collecting the map reference sample, using the segmented field boundaries to define the strata for cropland/non-cropland.
 To create the sample, we first extracted the centroids for each of the sample grid cells in Ghana.
 We then intersected the centroid points with the field segments, assigned a class of 1 (cropland) where points intersected a field, and 0 where they didn't (non-cropland).
-We then removed from this set of points all those that corresponded to model training, validation, or training reference sites.
-We then extracted a random sample from both the cropland and non-cropland points.
+We removed from this set of points all those that corresponded to model training, validation, or training reference sites, and extracted a random sample from both the cropland and non-cropland points.
 To determine the sample size of each, we specified a desired confidence interval using the following formula [@StehmanKeyissuesrigorous2019]:
@@
 #### Response design
-
-##### Stage 1
-We collected the map reference sample on a separate instance of `labeller` set up for the purpose (`validator.crowdmapper.org`).
-For the sampling unit, we selected a rectangular polygon of ~0.1 ha (0.0002866$^\circ$ resolution), centered on the centroid of each grid cell selected for map reference sample.
+We collected the map reference sample on a separate instance of the labelling platform set up for the purpose.
+For the sampling unit, we selected a rectangular polygon of ~0.1 ha (0.0002866$^\circ$ resolution).
+This polygon was centered on the centroid of each grid cell selected for map reference sample, and was presented as the target grid in the labelling platform.
 Four classes were established for the validation: **cropland**, **non-cropland**, **uncertain but likely cropland**, and **uncertain but likely non-cropland**.
@@
 This uncertainty also captures information about the inherent difficulty of the mapping task.
-Samples were collected by visually interpreting the overlays of PlanetScope composites presented in `labeller`, following the same interpretation protocols used by the labelling team, with the exception that the polygons placed were square and of 0.0002866$^\circ$ resolution.
+Samples were collected by visually interpreting the overlays of PlanetScope composites, following the same interpretation protocols used by the labelling team, with the exception that the polygons placed were square and of 0.0002866$^\circ$ resolution.
 
-The map reference sample was first placed and defined in an initial stage.
-In this stage, the class of the initial square polygon was determined, and then the polygon was adjusted if it overlapped more than one class.
-In sites where the initial map reference grid fell on one of the two uncertain classes, a second polygon was placed in the nearest location where a definite cropland or non-cropland interpretation could be made, with the choice determined by whether the uncertain class was more likely cropland or not cropland.
-The uncertain class was only used for providing insight into fallow cycles, while the second, more certain polygon was used in a second stage of the map reference sample, which was given to multiple independent interpreters (see below).
+Two supervisors (Ye and Estes) who were not involved in labelling developed the map reference sample evaluated the PlanetScope composites at each site to determine its class membership.
+The following set of rules were followed in collecting the sample:
 
-The following set of rules were followed in this first, sample placing phase (undertaken by Lyndon Estes):
-
-1. When determining the class corresponding with the initial location of the target grid, IF:
-  - More than half of the target falls within what appears to be a clear arable crop field, then assign it to **cropland**. 
-  - More than half falls in what is clearly not a field, then assign it to **non-cropland**. 
+1. When determining the class corresponding with the initial location of the target grid, if:
+  - More than half of the target falls within what appears to be a clear arable crop field, then classify it as **cropland**; 
+  - More than half falls in what is clearly not a field, then classify it as **non-cropland**; 
   - More than half falls in a location where it is harder to tell whether it is cropland or non-cropland, determine whether it is more likely a crop field or not a crop field, and then assign either **uncertain but likely cropland**, or **uncertain but likely non-cropland**.
 
-2. After determining the class, IF:
-  a. The target polygon is contained entirely within a single clear class AND
-      - The class is either **cropland** or **non-cropland** THEN:
-        - Digitize a point within the center of the target box, assign appropriate class, and complete the assignment
-      - The class belongs to one of the two **uncertain** classes THEN:
+2. After determining the class, if
+  - The target polygon is contained entirely within a single clear class, then simply digitize a point within the center of the target box, assign the appropriate class label, and complete the assignment;
+  - Digitize a square polygon exactly aligned with the initial target, choose the correct class label, and then move the new polygon to the nearest location where it can be contained entirely by the assigned class.
 
-        - Digitize a square aligned exactly with the initial target polygon, and choose the class **cropland** or **non-cropland** (following the most likely interpretation) and move the new polygon to the nearest location where it can be contained entirely within that clear **cropland** or **non-cropland** class;
-
-        - Digitize a point in the middle of the target polygon, and assign it the appropriate label: **cropland** or **non-cropland**.
-
-  b. The target polygon partially overlaps a second class AND:
-      - That class is either **cropland** or **non-cropland** THEN:
-
-        - Digitize a square aligned exactly with the target polygon, assign the appropriate class, and move the box to the nearest location (shortest possible distance) where it can be contained entirely within that same class.
-
-      - That class is **uncertain** THEN:
-
-        - Digitize a square aligned exactly with the target polygon, assign the appropriate **uncertain** class, then move the box to the nearest location (shortest possible distance) where it can be contained entirely within that same **uncertain** class.
-        - Digitize a second square aligned exactly with the initial target polygon, and choose the class **cropland** or **non-cropland** (following the most likely interpretation) and move the new polygon to the nearest location where it can be contained entirely within that clear **cropland** or **non-cropland** class.
-
-We then 
-  - Definition of spatial assessment
-  - Definition of classes
-  - Reference source
-  - Specific information collected from each source
-  - Rules for reference class labels
-  - How agreement between map and reference classes is defined
-
-##### Stage 2
+After collecting the sample, the geometries were further refined by converting the points (sites were the target didn't have to be shifted) to polygons with the same  0.0002866$^\circ$ resolution, and then the full set of map reference polygons was used to extract the classified values from both the cropland probability and vectorized field boundary maps.
 
@@
 
-The distribution of data collection effort and scores from training reference sites by labelling team (Figure \ref{fig:trainqual}).
-These results include those from repeat labelling of the initial training in Cluster 2.
+The distribution of data collection effort and scores from training reference sites by labelling team are shown in Figure \ref{fig:trainqual}.
+These results include those from re-labelling the initial training in Cluster 2 (AOIs 7-9, 12, 15).
 This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
-We replaced the image overlays in the instance for Cluster 2 \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
+We replaced the image overlays in the instance for central-southeast \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
 The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.
@@
 
-The total validation sample was `r sum(ref_ct)`, with `r ref_ct[2]` samples collected and interpreted by one observer (Su Ye) and `r ref_ct[1]` collected by a second observer (Estes).
+The size of the collected validation sample was `r sum(ref_ct)`, with `r ref_ct[2]` samples collected and interpreted by one observer (Su Ye) and `r ref_ct[1]` collected by a second observer (Estes).
 To evaluate the uncertainty inherent in defining the map reference labels, the pair mapped `r sum(usertab)` common sites, showing an overall level of agreement of `r agreement`%, and a Spearman Rank Correlation of `r round(unname(observer_r$estimate), 2)`.

### 2021-03-03 inst/manuscript/manuscript.Rmd
Edited start of discussion. Started committing figures

@@
 
-These results demonstrate our system's ability to create high resolution maps of the distribution and characteristics of cropland across a diverse and hard-to-map agricultural systems, at country-scales and at sub-annual to annual time steps.
-Our maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
-These earlier maps estimated that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana's areas in the 2015, in contrast to our 2018 maps, in which cropland covers 16.1-23.2% of the mapped area (Table \ref{tab:mapaccuracy}), with sample-corrected estimates of `r segcipct[1]`-`r rfcipct[1]`%.
-Our findings suggests that Ghana's arable cropland is slightly less than previous estimates, although the difference may also be attributable to the earlier maps using a more expansive definition of cropland than ours (active to recently active fields).
+These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales.
+The resulting maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
+This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover.
+Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference may be attributable to the earlier maps using a more expansive definition of cropland than we did (active to recently active fields).
 

### 2021-03-04 inst/manuscript/figures.Rmd
Edits to Discussion nearly complete

@@
           size = 0.05) +   
+  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
+               aes(label = ID), 
+               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
+               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
+               size = 3, col = "grey60") + 
+  geom_sf(data = aez, fill = "transparent", color = "gray60", size = 0.25) +
   scale_color_manual(values = pal) +
@@
   geom_sf(data = train_val_sites$active, aes(color = Type), size = 0.05) +
+  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
+               aes(label = ID), 
+               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
+               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
+               size = 3, col = "grey60") + 
+  geom_sf(data = aez, fill = "transparent", color = "gray60", size = 0.25) +
   scale_color_manual(values = pal) +

### 2021-03-04 inst/manuscript/manuscript.Rmd
Edits to Discussion nearly complete

@@
 - name: Su Ye
-  affilnum: '1'
+  affilnum: '1,2'
 
@@
 - name: Justus Muhando
-  affilnum: '2'
+  affilnum: '3'
 
 - name: Ryan Avery
-  affilnum: '3'
+  affilnum: '4'
 
 - name: Kelly Caylor
-  affilnum: '3-5'
+  affilnum: '4-6'
 
@@
 - affilnum: 1
-  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA 01610 USA
+  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA, USA
 
 - affilnum: 2
-  affil: SpatialCollective
+  affil: Department of Natural Resources and the Environment, University of Connecticut, Storrs, CT, USA
+
+- affilnum: 2
+  affil: SpatialCollective, Nairobi, Kenya
 
@@
 To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
-The labeller assigns each polygon a field class category (e.g.
-annual cropland, tree crop, fallow), and upon completing all fields submits the assignment to the database.
+The labeller assigns each polygon a class category (e.g.
+annual cropland), and upon completing all fields submits the assignment to the database.
 In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete.
@@
 #### Classification
+
 Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
@@
 The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), A separate active learning and segmentation process was run for each of these AOIs.
-
 To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16).
@@
 
-To evaluate the performance of the system, we undertook several different analyses.
+During the collection of training and validation samples, labellers were tasked to only digitize active or recently active crop fields, avoiding tree crops, and fallow or potentially abandoned fields (see SI for the digitizing rules).
+
+To evaluate the performance of the system, we performed several analyses described in sections 2.3.1-4.
 
@@
 
-After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI (see Figure S3A for sample distributions).
+After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI.
 AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement.
@@
 In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
+Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often falling along ecotones, such as the boundaries between agroecozones (see Figure S3A).
 The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.
@@
 This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover.
-Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference may be attributable to the earlier maps using a more expansive definition of cropland than we did (active to recently active fields).
-
-In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S12).
-Prior efforts to map such agricultural characteristics have often focused on estimating farm rather than field size using census datasets [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004].
-Efforts to map fields in small-scale agricultural systems have either been field-based [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small areas [e.g.
-@forkuorIntegrationOpticalSynthetic2014] (**include Persello**).
-The most notable efforts to estimate field sizes were two global initiatives that used crowdsourced volunteers to categorize five broad field size classes based on their interpretations of imagery sampled from virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
-The most recent of these [@lesivEstimatingGlobalDistribution2019] provided country-specific results for Ghana (n = 263) that return an average field size of `r round(lesiv_est, 2)` ha (calculated as the weighted mean of the class counts and the middle values of the four smallest size classes and the lower bound of the largest size class).
-This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean segment size (`r aoi_nofor` ha) across AOIs 1-9, 12, and 15, which is where most of their sample appears to have been collected.
-However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha are much smaller.
-
-## Sources of error
+Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].
+
+In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S11-12).
+Previous work to estimate such agricultural characteristics have often focused on farm, rather than field, size using census data [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004].
+Efforts to map field boundaries in smallholder-dominated agricultural systems have either used *in situ* data collection [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small [e.g.
+@forkuorIntegrationOpticalSynthetic2014; @perselloDelineationAgriculturalFields2019] or discontiguous [@Estesplatformcrowdsourcingcreation2016] areas.
+The most extensive studies to date used crowdsourced volunteers to classify fields into broad size classes, based on their interpretations of imagery sampled from high resolution virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
+Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha$\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class.
+Data sourced from Table S3 in Lesiv et al.
+2019.}$.
+This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean (`r aoi_nofor` ha) within AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample appears to have been collected.
+However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha were much smaller.
+
+## Map accuracy and key sources of error
 Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
-The overall map accuracies are good (86.7-88%, Table \ref{tab:mapaccuracy}), which is at the boundary of what could be considered achievable map accuracy [@ElmesAccountingtrainingdata2020], given the level of confidence in the map reference sample (only 85% of reference labels were assigned with confidence).
-However, accuracies for the cropland class were between 62 (Producer's) and 67 (User's) percent in the per-pixel map across the country, indicating that both commission and omission errors were above 30% for this class.
-The field boundary has lower omission error (Producer's accuracy = 79%), but that comes at the expense of increased false positives (User's accuracy = 58.2%).
+The overall map accuracies are decent (86.7-88%, Table \ref{tab:mapaccuracy}), and are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given the inherent uncertainty in the map reference sample, our best estimate of the "truth", in which we have roughly 85% confidence.
+However, accuracies for the cropland class were much lower, falling between 62 (Producer's) to 67 (User's) percent country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}), meaning the model produced substantial commission and omission errors for this class.
+The segmented boundary maps had fewer omission errors (Producer's accuracy = 79%), but higher false positive errors (User's accuracy = 58.2%).
 These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
 
-The patterns of cropland-class accuracies varied by zone, which largely align with agroecozones and thus somewhat reflective of the inherent challenge each poses for mapping.
-Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savanna agroecoregions (Figure S9), and lowest in zones 3 and 4, which are either forest or coastal savannas.
+The patterns of cropland-class accuracies varied by zone.
+These zones largely align, albeit with some discrepancies, with the country's agroecoregions, thus the accuracy patterns may be partially attributed to some regions being harder to map than others.
+Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S9), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas.
 User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample.
-Confining the samples to specific agroecozones provides further insight into error patterns (Table S4).
-The coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably due to the numerous areas of high density cropland and low woody cover in surrounding uncultivated areas, which promotes commission error.
-The northern savannas had the best balance between omission and commission error, and had the highest overall User's accuracy.
-The transition zone had a very low Producer's accuracy (21%), which likely reflects the fact that this region, which also has high cropland density, was divided between several AOIs for mapping.
-
-The relatively low accuracies for the cropland class can be attributed to several important factors.
-The first of these was the overall mapping extent combined with the resolution of the analysis.
-Given the goal of country-scale mapping at high resolution, we found that the computational cost required us to use a relatively small and simple set of features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forest implementation.
-Previous work shows that RandomForest achieves much better performance on small-scale croplands when trained on a large number of randomly generated features [@Debatsgeneralizedcomputervision2016], which is a commonality with deep learning.
-However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the typical 10-14 hours required from training through prediction on 800 CPUs, for a single iteration.
-This left the model less able to discern cropland from surrounding savanna vegetation with bare patches or sparse grass cover, which was common in most AOIs.
-A larger feature set with more variable textures filters could have helped distinguish such areas from crop fields, helping to particularly reduce commission error.
-
-Another major limiting factor was the inherent difficulty of the labelling task.
-Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands over these system is difficult and requires the evaluation of multiple image sources and considerable judgement by each labeller.
-Interpretation is particularly hard over areas where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3.
-Such changes in scene circumstances require labellers to adopt different heuristics, for example by delineating the field boundary on the growing season rather than the dry season image.
-Smaller field sizes also complicated the labelling task, becoming increasingly indistinct in the ~4 m PlanetScope composites.
-The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by labeller accuracies that averaged 71% (Figure S4B).
-The design of our system allowed us to measure how much label quality impacts model performance.
-Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] has reported that RandomForest is particularly robust to training data error, we found its impact to be substantial (Figure \ref{fig:trainingimpact}), suggesting that the greatest single investment towards improved accuracy lies in improving training data quality.
-
-Image quality was another issue, although primarily in the forest zone (Zone 4) AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}).
+Aligning the reference samples more precisely with agroecozone boundaries (Figure S9B) provides further insight into error patterns (Table S4).
+Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error.
+Maps in the two northern savanna agroecoregions had the best balance between omission and commission error, and had the highest overall User's accuracy.
+The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S3), within which it typically covered a smaller share of area relative to the other agroecoregions.
+This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S3B).
+
+Beyond the errors linked to regional differences, several other important factors contributed to reducing the accuracies in the cropland class.
+The first of these stems from the overall mapping extent and the high resolution of the data.
+Given the goal of developing a country-scale map at high resolution, the attendant data volume required us to use a relatively small set of image features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forests implementation, in order to limit computational costs.
+Previous work found that Random Forests achieves much better performance on small-scale croplands when trained on a much larger number of features [@Debatsgeneralizedcomputervision2016].
+However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the ~4-8 hours of runtime on 800 CPUs required for a single active learning iteration, followed by ~10-14 hours for prediction.
+This reduced the skill of the model, particularly when it came to differentiating cropland from adjacent bare patches or natural vegetation with sparse herbaceous cover, which were common in many AOIs.
+
+The inherent difficulty of the labelling task was another major limiting factor.
+Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands in these agricultural systems can be a difficult task.
+Labellers have to evaluate multiple image sources and to rely heavily on judgement, which inevitably leads to errors.
+Interpretation is particularly hard where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3.
+Smaller field sizes also complicate labelling, as these become increasingly indistinct in the ~4 m PlanetScope composites.
+The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by the average score achieved by each labeller against our training reference dataset  (71%; Figure S4B).
+Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label errors, we found that they have substantial impact (Figure \ref{fig:trainingimpact}), which suggest that simply improving label quality may be one of the single most important investments towards improved model accuracy.
+
+Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}).
 This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions.
-There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance--prior work shows that imagery from multiple seasons is important for improving classifier performance [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
-Composite quality could be improved by integrating imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would confound the ability to delineate field boundaries.
+There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance, given the importance of multi-temporal imagery for cropland classification [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+Composite quality could be improved by using imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would blur field boundaries within the imagery.
 
 The final major source of error arose from the segmentation process.
-The vectorized maps had high commission errors that are attributable to the RandomForest predictions.
-Uncertainty in that model resulted in substantial numbers of pixels just above the 0.5 classification threshold, meaning a number of segments over highly uncertain areas were retained as fields.
-A more accurate classifier would reduce such errors.
-Another possible solution could be to apply a locally varying classification threshold to distinguish croplands from non-cropland [e.g.
+The vectorized maps had high commission errors caused by uncertainties in the Random Forests predictions.
+Model uncertainty led to many pixels over non-crop areas with probabilities straddling the 0.5 classification threshold.
+Segments that intersected such areas were retained as fields when the average probability of intersecting pixels exceeded 0.5.
+A more accurate classifier would reduce such errors, or the application of a locally varying classification threshold [e.g.
 @waldnerDeepLearningEdge2020].
-Over-merging was another source of error in the segmentation algorithm.
-This led in some areas to over-estimation of field sizes and often irregular shapes, particularly in high density croplands (e.g.
-in AOIs 2 and 8; Figure \ref{fig:mainmap}) where adjacent fields have boundaries that are indistinct in the PlanetScope imagery.
-Minimizing or preventing merging would help in such cases, although it would result in  over-segmentation and thus would likely underestimate field sizes.
-However, over-segmentation can help to identify important within-field distinctions (e.g.
-different soil types), which can be useful for other applications, such as filtering pixels for crop type mapping or yield estimation [e.g.
-@EstesComparingmechanisticempirical2013; @EstesProjectedclimateimpacts2013]
-
-## Key system features
-Despite the sources of errors mentioned above, the design of the system proved effective in mitigating error, resulting in a higher overall accuracy than would have otherwise been possible.
-The label accuracy assessment and consensus labelling methods appeared to be the most effective tools for error mitigation.
-Label accuracy assessment allowed us to quantify the impact of trainnig data error on model performance, in keeping with recommended best practices [@ElmesAccountingtrainingdata2020].
-Label quality proved to have a large impact on model performance (Figure \ref{fig:trainingimpact}), and consensus labeling was effective in reducing individual digitization errors and between-labeller differences in interpretation, improving label, and thus map, quality.
-Worker-specific label accuracy measures also helped to improve the overall quality of the consensus labels, by enabling higher weight to be placed on labels that were more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels.
-The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we could in turn use to bias-correct map-based size class estimates (Table \ref{tab:sizentab}).
-The ability to select the most accurate label will also help when different training strategies are needed, such as when training a model to perform instance segmentation [@carvalhoInstanceSegmentationLarge2021], in which case the field identities have to be preserved.
+Over-merging was another source of error in the segmentation algorithm, which in some areas led to overestimated field sizes and irregular shapes, particularly in high density croplands (e.g.
+in AOIs 2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent fields are often indistinct in the PlanetScope imagery.
+Minimizing or preventing merging would help in such cases, although could would result in the opposite problem, over-segmentation, and thereby underestimate field size.
+
+## Error mitigation features
+Despite the error sources mentioned above, several features of the system proved effective in mitigating error, leading to a higher overall accuracy than would have otherwise been possible.
+Label accuracy assessment and consensus labelling appeared to be the most effective error mitigation tools.
+Label accuracy measures allowed us to quantify the substantial impact of label error on model performance (Figure \ref{fig:trainingimpact}), while consensus labels substantially reduced individual labelling errors, resulting in maps that were more accurate than they would have been had we used individually generated labels.
+Labeller-specific accuracy measures also helped to improve the quality of the consensus labels, by placing higher weight on labels more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels.
+The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).
 
@@
 ```
-The active learning approach helped to improve overall model performance relative to a randomized approach to training site selection.
-Although we only compared active to random site selection within three AOIs, the average gains across all three AOIs and three performance metrics demonstrated the effectiveness of the active approach, most notably by the gains in the F1 score (Figure S7), a combined metric of commission and omission error.
+The active learning approach helped to improve overall model performance relative to randomized training site selection, confirming previous findings reported by @DebatsIntegratingactivelearning2017.
+Although the performance gains relative to randomized model training that we observed were substantially less than in the previous study, which reported `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher after 45 iterations [@DebatsIntegratingactivelearning2017], those comparisons were made from a low initial base, beginning with a single training site with one more site added each iteration.
+In contrast, our large initial random sample (500) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance.
+Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness.
+Most notable were the larger gains in the F1 score (Figure S7), a balanced performance metric.
 The gains in accuracy and AUC were smaller.
-For accuracy, this was due to the larger drop in the active learning metric after iteration 1 in AOI 15, which proved one of the hardest AOIs to map.
-The plateau in AUC gains at 0.5% relative to random training reflects the fact that each iteration tended to reduce both the false and true positive rates simultaneously.
-Although the false positive rate overall declined (an average of `r prdelta$rate[1]`% between Iterations 0 and 3), the reductions in true positive rate (averaging `r prdelta$rate[2]`%) offset some of the improved predictive power.
-
-The gain due to active learning were substantially less than those reported by Debats et al (2017), who found that active site selection improved classifier performance by `r round((0.45-0.35) / 0.35 * 100)` over randomized site selection after one iteration, and after 45 iterations by `r round((0.69-0.64) / 0.64 * 100)`%.
-These larger gains likely reflect the fact that this evaluation began from the smallest possible training sample--one image chip--and tested the effect of adding a single additional chip each iteration.
-In contrast, we started with a large initial random sample (500) and added 100 samples per iteration, thus our models received significant training at the outset, making the marginal gain due to each new sample smaller.
-This suggests that we could have benefitted by reducing the size of the initial random training sample while increasing the size of actively selected samples, not only in terms of larger accuracy gains, but also in terms of reducing--by as much as 62%, per Debats et al (2017)--the overall size of the training sample needed to achieve a given level of performance.
-
-The detail, large extent, and temporal precision of the maps we created by virtue of our system's intentional design around PlanetScope imagery, which currently provides the only wall-to-wall, sub-5 meter imaging capability [@McCabefutureEarthobservation2017] with daily frequency.
-Daily frequency is important for creating multiple seasonal composites during a single agricultural year in cloudy regions.
-Although unquantified, our compositing technique may be more effective at providing complete image coverage during a season than the "best-on-top" approach\footnote{https://assets.planet.com/products/basemap/planet-basemaps-product-specifications.pdf} used by Planet to develop their analytic basemap, which anecdotally appears to contains substantial residual cloud cover during the growing season in southern Ghana.
-Although free and providing higher radiometric quality and depth, and of sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
-@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], Sentinel-2's 5-day revist may be too infrequent to provide enough clear pixels to create a cloud-free growing season composite in the southern half of Ghana.
-Sentinel-1 is not affected by the same problem, but labelling fields in a more coarsely resolved radar images is difficult.
+For accuracy, the reduced advantage was primarily due to active learning being outperformed by randomized training after the first iteration in AOI 15, which proved one of the hardest AOIs to both map and label.
+Active learning likely resulted in the selection of sites that were harder to label than randomly selected ones, leading to more label error, and thus initially lower model accuracy.
+However, this deficit was overcome by the 5th iteration.
+The plateau in AUC gains at 0.5% better than randomized training reflects the findings that active learning reduced both the false and true positive rates, the two inputs to AUC.
+Although the false positive rate declined by `r prdelta$rate[1]`% between Iterations 0 and 3, the `r prdelta$rate[2]`% reduction in the true positive rate offset some of these gains, and the effect of this tradeoff may have been more pronounced in the threshold independent AUC.
+
+The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
+Daily revisits are necessary for creating seasonal composites within a single agricultural year over cloudy areas.
+The compositing technique we developed to work with PlanetScope was able to provide complete image coverage over the entire country for both seasons.
+Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
+@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana.
+Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is difficult.
 
@@
 Several potential issues not answered by our assessment merit further exploration.
-These include the correspondence between image- and ground-collected labels.
-We believe that such comparisons will in most circumstances reveal unresolvable discrepancies.
-The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels can produce substantial differences.
+One of these is the degree of correspondence between image- and ground-collected labels.
+However, such comparisons will likely reveal unresolvable discrepancies in most circumstances.
+The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels may show substantial differences.
 These discrepancies could be further exacerbated by differences in an understanding of what constitutes a field, which might vary on the ground depending on the farmer or the data collecter's collectors judgement.
@@
 ## Next steps
-The maps presented here represent a version 1 product that is freely available to use **(links)**, along with its underlying code.
+The maps presented here represent a version 1 product that is freely available to use, along with its underlying code.
 These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020].
@@
 To facilitate the next step, generating more accurate version 2 maps, several improvements will be made.
-The first is to replace Random Forest with a more advanced convolutional neural network (CNN), which can generate and learn from a large number of features representing a variety of spatial scales [@maDeepLearningRemote2019].
-Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, provide an effective means for delineating field boundaries [@waldnerDeepLearningEdge2020].
+The first is to replace Random Forests with a more advanced convolutional neural network (CNN), which can generate and learn from a large number of features representing a variety of spatial scales [@maDeepLearningRemote2019].
+Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, are effective in delineating field boundaries [@waldnerDeepLearningEdge2020].
 Our system can readily incorporate such a model.
@@
 The existing tooling can thus be adjusted to guide training site selection in sub-regions of Ghana for a CNN trained from scratch from a larger initial random draw.
-Instead of the current AOIs, a more effective strategy may be to update the model within agroecozones (Figure S9B) The labelling methods already provide for a case in which edge and interior classes need to be labelled, as the most accurate individual training samples can be identified and used, rather than the consensus strategy.
+Instead of the current AOIs, a more effective strategy may be to update the model within agroecozones (Figure S9B) The labelling methods already provide for a case in which edge and interior classes need to be labelled, as the most accurate individual training samples can be identified and used, rather than binary consensus labels.
 The segmentation algorithm will also be updated to work with the resulting score maps.
@@
 ## Conclusion
-This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands over national to regional scales, using a flexible and framework that can be readily updated to improve their accuracy as technologies improve.
+This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve.
 Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.
-
 # Acknowledgements
 The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund.
-Support for computing was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative.
 Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393).
-We thank Joe Okyere and his colleagues at Meridia for assistance in data collection.
+Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative.
+Azavea provided significant support in developing the machine learning pipeline.
+We thank Meridia for assistance in initial data about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
 

### 2021-03-04 inst/manuscript/si.Rmd
Edits to Discussion nearly complete

@@
 
-#### Rules for interpreting and mapping imagery
+#### Interpreting and mapping imagery
 
-- **Rule 1**: Map the fields that are visible in the Planet imagery, not the fields that you see in the basemap.
+- Map the fields that are visible in the PlanetScope images, not the fields that you see in the basemap.
 
-- **Rule 2**: Please map the fields as they appear in the off-season Planet images. If it is too hard to make these out clearly, the next choice is to map them in growing season imagery.
+- As the first choice, map the field outlines as they appear in the dry season PlanetScope images.
+If a field is not clearly visible in the dry season image, but is visible in the growing season image, then the next choice is to map it in the growing season image.
 
-- **Rule 3**: Use all available images (Planet scenes and base map) to help you interpret what is and is not a field.
-Toggling back and forth between the different image sources can help you identify and classify the field boundaries.
+- Use all available images (PlanetScope scenes and base map) to help you interpret what is and is not a field.
+Toggling back and forth between the different image sources can help identify and classify the field boundaries.
 
-- **Rule 4**: Map fields that intersect the mapping target, making sure to finish the entire boundary.
+- Map fields that intersect the mapping target, making sure to finish the entire boundary.
 
-#### Rules for classifying different types of agriculture
+#### Choosing what to map
 
-- **Rule 5**: Choose the annual croplands category if the fields looks like they contain crops that are planted and harvested during a single year (or occasionally after a few seasons).
+- Draw polygons around fields that look like they contain crops that are planted and harvested during a single year (or occasionally slightly longer, in the case of crops such as sugarcane).
+These fields may look recently ploughed or harvested, or contain actively growing crops.
 
-- **Rule 6**: Fields that look like they are under tree crops, such as orchards, woodlots, or other planted forest types are agriculture, but we will not map these in the current campaign.
-Therefore, do not draw polygons around these types of fields.
+- Do not draw polygons around fields that are under tree crops, such as orchards, woodlots, or other planted forest types.
 
-- **Rule 7**: Choose the fallow category if the fields looks like it was an annual cropland that is somewhat overgrown, and hasn't been planted in a few seasons.
-[Note: this category was eventually not used, as such fields were not mapped]
+- Do not draw polygons around fields that look like they are annual croplands, but are overgrown and haven't been planted for a few years. These are possibly fallows or abandoned fields.
 
-- **Rule 8**: Choose the field, but unsure of category if you can't tell whether the field is annual cropland, a tree crop, or fallow.
-Try hard to avoid this category.
-[Note: this category was also not used]
+- Do not draw polygons if you cannot tell whether a piece of land is an annual crop field or if it is perhaps another type of land cover (e.g. bare land, or a young orchard).
 
-- **Rule 9**: Choose the unsure if agricultural field or not category if you think the field might be agriculture (whether annual cropland, tree crops, or fallow), but it could also be non-agricultural land.
-Try hard to avoid this category also.
-[Note: this category was also not used]
-
-- **Rule 10**: Don't map any polygons if there is no agricultural to map, or if the fields look like they are tree crops (see updated Rule 6).
+- Do not draw any polygons if there are no annual crop fields to map.
 
@@
 
-```{r refsample, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class. "}
+```{r refsample, echo = FALSE, fig.align='center', out.width="80%", message=FALSE, warning=FALSE, fig.cap="Distribution of the selected map reference sample for the cropland and non-cropland class.
+Classes represent the values extracted from the map strata, rather than those assigned during the classification of the sample."}
 knitr::include_graphics('figures/si_map_reference_sample.png')
@@
 Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
-```{r trainval, echo = FALSE, out.width="80%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected within AOI 3 to retrain a second active learning run for this region, as well as the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.", fig.align='center', message=FALSE, results='hold'}
+```{r trainval, echo = FALSE, out.width="80%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected within AOI 3 to retrain a second active learning run for this region, as well as the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.
+Grey background lines indicate the boundaries of Ghana's agroecozones, with the grey numbers indicating the agroecoregion ID: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE, results='hold'}
 knitr::include_graphics('figures/si_training_validation_pts.png')

### 2021-03-04 inst/manuscript/manuscript.Rmd
Edited through discussion, abstract still missing

@@
 - name: Justus Muhando
+  affilnum: '3'
+
+- name: Angeline Hokar
+  affilnum: '3'
+
+- name: Charles Juma 
+  affilnum: '3'
+
+- name: Jackson Makuru
+  affilnum: '3'
+
+- name: Fredrick Matasa
+  affilnum: '3'
+
+- name: Ben Mbatia
+  affilnum: '3'
+
+- name: Judith Mbaye
+  affilnum: '3'
+
+- name: Isaac Muasa
+  affilnum: '3'
+
+- name: Julius Mucha
+  affilnum: '3'
+
+- name: Adelide Mugami
+  affilnum: '3'
+
+- name: Purent Oduor
+  affilnum: '3'
+
+- name: Jeff Ochieng
+  affilnum: '3'
+
+- name: Francis Wambua
+  affilnum: '3'
+
+- name: Thuo Wanjiku
+  affilnum: '3'
+
+- name: Joe Wanyoike 
   affilnum: '3'
@@
 footerdate: yes
-abstract: The abstract can go either here or below
+abstract:
+
 keywords: rmarkdown, reproducible science
@@
 The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020].
-The training and machine learning components are combined within an "active learning" [@cohnImprovingGeneralizationActive1994] framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017].
-Our framework automates this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g.
-@XiongNominal30mCropland2017].
-Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@cohnImprovingGeneralizationActive1994; @DebatsIntegratingactivelearning2017].
-Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a final vector-based field boundary map that provides important information on field geometry.
+The training and machine learning components are combined within an active learning framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011].
+Our framework automates this interactive approach to label selection, which is effective in boosting the performance of classification models while reducing the overall number of training samples required to achieve a given level of performance [e.g.
+@hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017].
+Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a vectorized field boundary map that provides important information on field geometry.
 
@@
 ```
-The active learning approach helped to improve overall model performance relative to randomized training site selection, confirming previous findings reported by @DebatsIntegratingactivelearning2017.
-Although the performance gains relative to randomized model training that we observed were substantially less than in the previous study, which reported `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher after 45 iterations [@DebatsIntegratingactivelearning2017], those comparisons were made from a low initial base, beginning with a single training site with one more site added each iteration.
-In contrast, our large initial random sample (500) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance.
+The active learning approach helped to improve overall model performance relative to randomized training site selection, in line with findings from two recent efforts [@DebatsIntegratingactivelearning2017; @hamrouniLocalGlobalTransfer2021].
+Although the performance gains relative to randomized model training that we observed were smaller (e.g.
+@DebatsIntegratingactivelearning2017 `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher on the final iterations), those comparisons were made from lower initial bases, with initial training samples that were less than 1/10 the size, in terms of pixels, of our initial training sample.
+Our large initial randomly selected sample (500 grid cells) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance.
 Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness.
-Most notable were the larger gains in the F1 score (Figure S7), a balanced performance metric.
-The gains in accuracy and AUC were smaller.
+Most notable were the larger improvements seen in the F1 score (Figure S7), a balanced performance metric.
+Gains in accuracy and AUC were smaller.
 For accuracy, the reduced advantage was primarily due to active learning being outperformed by randomized training after the first iteration in AOI 15, which proved one of the hardest AOIs to both map and label.
@@
 The plateau in AUC gains at 0.5% better than randomized training reflects the findings that active learning reduced both the false and true positive rates, the two inputs to AUC.
-Although the false positive rate declined by `r prdelta$rate[1]`% between Iterations 0 and 3, the `r prdelta$rate[2]`% reduction in the true positive rate offset some of these gains, and the effect of this tradeoff may have been more pronounced in the threshold independent AUC.
+Although the decline of the false positive rate (`r prdelta$rate[1]`% between Iterations 0 and 3) was nearly three times larger than that of the true positive rate (`r prdelta$rate[2]`%), AUC should be quite sensitive to the reduction in the latter, as it assesses how the tradeoff between the two rates varies across a full range of possible classification thresholds [@Pontiustotaloperatingcharacteristic2014].
 
 The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
-Daily revisits are necessary for creating seasonal composites within a single agricultural year over cloudy areas.
-The compositing technique we developed to work with PlanetScope was able to provide complete image coverage over the entire country for both seasons.
+Daily revisits are necessary for creating seasonal composites within a single  year over cloudy areas.
+The compositing technique we developed allowed us to develop a complete image catalog for the country representing the two seasons for 2018 agricultural year.
 Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
 @Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana.
-Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is difficult.
+Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is challenging.
 
@@
 One of these is the degree of correspondence between image- and ground-collected labels.
-However, such comparisons will likely reveal unresolvable discrepancies in most circumstances.
-The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels may show substantial differences.
-These discrepancies could be further exacerbated by differences in an understanding of what constitutes a field, which might vary on the ground depending on the farmer or the data collecter's collectors judgement.
-On top of this, field-collected boundaries are typically highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is.
+However, such comparisons will likely reveal unresolvable differences between the two perspectives.
+The highly dynamic nature of many agricultural systems means that relatively narrow differences between the dates of ground- and image-based digitizing campaigns can lead to substantial disagreement between the resulting field boundaries, simply because the fields themselves may have shifted during the interval [@ElmesAccountingtrainingdata2020].
+These discrepancies could be further exacerbated by differences in the definition of what constitutes a field, which might vary on the ground depending on who is being asked, or who is doing the collecting.
+On top of this, ground-based collection efforts are usually highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is, or may be unrepresentative.
 These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth".
-Nevertheless, a comparison against ground data can help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.
+Nevertheless, a comparison against ground data would help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.
 
 The temporal discrepancies mentioned above [and discussed in @ElmesAccountingtrainingdata2020] are another reason why we chose not to label on basemap imagery (in addition to restrictive usage terms), which is typically several years old [@LesivCharacterizingspatialtemporal2018].
-However, we did not assess whether the higher label accuracy one might achieve by digitizing on <1-2 m basemap would offset model errors due temporal mismatches between labels and imagery.
+However, we did not assess whether the higher label accuracy one might achieve by digitizing on a <1-2 m resolution basemap would offset model errors caused by temporal mismatches.
 
@@
 This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels.
-However, the visual assessment of probability images generated from each labelling strategy show that these two strategies produce nearly identical maps, and that the model trained with low quality labels generated a markedly less certain map (Figure S6).
+However, the visual assessment of their resulting probability maps confirm the differences in scores: consensus and most accurate individual labels produce nearly identical maps with relatively high certainty, while low quality labels led to a markedly less certain map (Figure S8).
 
@@
 Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.
+
 # Acknowledgements
@@
 Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative.
-Azavea provided significant support in developing the machine learning pipeline.
-We thank Meridia for assistance in initial data about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
+Azavea provided significant contributions in engineering the machine learning pipeline.
+We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
 

### 2021-03-05 inst/manuscript/manuscript.Rmd
Near-final draft of MS and SI

@@
 Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
-This high spatial *and* temporal resolution addresses the sensor-scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+This high spatial *and* temporal resolution addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
 On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale.
@@
 We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of climates and agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
-These factors make Ghana emblematic of the substantial challenges facing cropland mapping over the broader region.
 
@@
 ## Assessing the characteristics of Ghanaian cropland
-Using the final mapped results, we calculated the estimated area of arable cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
+Using the final mapped results, we calculated the estimated area of cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
 We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the differences between labellers' polygons and segmented boundaries at validation sites to calculate bias-adjusted estimates of mean field sizes and the total number of fields.
@@
 ```
-We developed maps of Ghana's arable croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles.
+We developed maps of Ghana's croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles.
 Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
@@
 
-### Validation performance
+### Performance gains during active learning
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
@@
 The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
-Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`'; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
+Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
 
@@
 
-```{r mainmap, echo = FALSE, out.width="90%", fig.cap="The distribution of arable croplands in Ghana.
+```{r mainmap, echo = FALSE, out.width="90%", fig.cap="The distribution of croplands in Ghana.
 The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities.
@@
 
-Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of arable croplands in Ghana.
+Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of croplands in Ghana.
 The cropland extent estimated from the vectorized map is `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area.
@@
 These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales.
-The resulting maps provide an updated and more granular view of the distribution and extent of arable croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
+The resulting maps provide an updated and more granular view of the distribution and extent of croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
 This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover.
-Our results thus suggest that Ghana's arable cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].
+Our results thus suggest that Ghana's cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].
 
@@
 Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}).
-This impacted labellers' abilities to discern arable fields, and doubtless affected model predictions.
+This impacted labellers' abilities to discern fields, and doubtless affected model predictions.
 There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance, given the importance of multi-temporal imagery for cropland classification [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
@@
 We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
+We thank Manushi Trivedi, Sitian Xiong, and Tammy Woodard are thanked for their contributions to the underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and Primoz Kovacic for support and management of the labelling efforts.
 

### 2021-03-05 inst/manuscript/manuscript.Rmd
Abstract added

@@
 footerdate: yes
-abstract:
+abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands.
+Such data are often unavailable for regions dominated by smallholder agricultural systems, which are particularly challenging for remote sensing.
+To overcome these challenges, we designed a system to minimize several sources of error that arise when mapping smallholder croplands.
+To overcome errors caused by mismatches between image resolution and cropland scales, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year.
+To reduce errors that occur when training classifiers, we built a labelling platform that rigorously assesses label accuracy, and creates more accurate consensus labels that train a Random Forests model.
+The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which combines with a segmentation process to delineate individual field boundaries.
+We applied this system to map Ghana’s croplands for the year 2018.
+We divided Ghana into 16 mapping regions (12,160-23,535 km$^2$), training separate models for each using a total of 6,299 labels, plus 1,600 for validation.
+Using an independent map reference sample (n=1,207), we found that overall accuracies of the resulting cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%.
+Croplands covered 16.1-23.2% of the mapped area, comprising 1,131,146 total fields with an average size of 3.92 ha.
+Unbiased estimates based on the map reference sample yield indicate a cropland percentage of 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on whether the cropland probability or field boundary map was assessed.
+Using the labellers’ digitized field boundaries to estimate biases in field boundary statistics, we calculated an adjusted mean field size of 1.73 ha and total field count of 1,662,281.
+Although the cropland class contained substantial errors, the system was effective in mitigating error and quantifying resulting performance gains.
+By minimizing training errors, consensus labelling improved the model’s F1 scores by up to 25%, while 3 iterations of active learning increased the F1 score by 9.1%, on average, which was 2.3% higher than training models with randomly selected labels.
+Map accuracy may be further improved by replacing Random Forests with a convolutional neural network.
+These results demonstrate a readily adapted, transferrable framework for developing high resolution, annual, nation-scale maps that provide important details about smallholder-dominated croplands.
 
@@
 # Introduction
-Amidst all the challenges posed by global change, a particular concern is how agriculture will adapt to meet humanity's growing food demands, and the impacts that a transforming and expanding agriculture will have on societies, economies, and environment [@searchingerCreatingSustainableFood2019].
+Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019].
 Significant efforts are being made to address the various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g.
@@
 Unfortunately, the data that do exist are in many places inaccurate or missing.
-Existing estimates of how much cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
+Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
 @FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013].
@@
 Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
-For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 80, which initial testing showed to provide a reasonable balance between computational time and performance.
+For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 60, which initial testing showed to provide a reasonable balance between computational time and performance.
 

### 2021-03-07 inst/manuscript/manuscript.Rmd
Edits to ms and si, plus commit of many si figures

@@
 
-- name: Lyndon Estes
+- name: L.D. Estes
   affilnum: '1'
@@
 
-- name: Su Ye
+- name: S. Ye
   affilnum: '1,2'
 
-- name: Lei Song
+- name: L. Song
   affilnum: '1'
 
-- name: Boka Luo
+- name: B. Luo
+  affilnum: '1,3'
+
+- name: J. Ronald Eastman
+  affilnum: '1,3'
+
+- name: Z. Meng
   affilnum: '1'
 
-- name: Ron Eastman
+- name: Q. Zhang
   affilnum: '1'
 
-- name: Zhenhua Meng
-  affilnum: '1'
-
-- name: Qi Zhang
-  affilnum: '1'
-
-- name: Dennis McRitchie
+- name: D. McRitchie
   affilnum: ''
 
-- name: Stephanie Debats
+- name: S. Debats
   affilnum: ''
 
-- name: Justus Muhando
-  affilnum: '3'
-
-- name: Angeline Hokar
-  affilnum: '3'
-
-- name: Charles Juma 
-  affilnum: '3'
-
-- name: Jackson Makuru
-  affilnum: '3'
-
-- name: Fredrick Matasa
-  affilnum: '3'
-
-- name: Ben Mbatia
-  affilnum: '3'
-
-- name: Judith Mbaye
-  affilnum: '3'
-
-- name: Isaac Muasa
-  affilnum: '3'
-
-- name: Julius Mucha
-  affilnum: '3'
-
-- name: Adelide Mugami
-  affilnum: '3'
-
-- name: Purent Oduor
-  affilnum: '3'
-
-- name: Jeff Ochieng
-  affilnum: '3'
-
-- name: Francis Wambua
-  affilnum: '3'
-
-- name: Thuo Wanjiku
-  affilnum: '3'
-
-- name: Joe Wanyoike 
-  affilnum: '3'
-
-- name: Ryan Avery
+- name: J. Muhando
   affilnum: '4'
 
-- name: Kelly Caylor
-  affilnum: '4-6'
+- name: A. Hokar
+  affilnum: '4'
+
+- name: C. Juma 
+  affilnum: '4'
+
+- name: J. Makuru
+  affilnum: '4'
+
+- name: F. Matasa
+  affilnum: '4'
+
+- name: B. Mbatia
+  affilnum: '4'
+
+- name: J. Mbaye
+  affilnum: '4'
+
+- name: I. Muasa
+  affilnum: '4'
+
+- name: J. Mucha
+  affilnum: '4'
+
+- name: A. Mugami
+  affilnum: '4'
+
+- name: P. Oduor
+  affilnum: '4'
+
+- name: J. Ochieng
+  affilnum: '4'
+
+- name: F. Wambua
+  affilnum: '4'
+
+- name: T. Wanjiku
+  affilnum: '4'
+
+- name: J. Wanyoike 
+  affilnum: '4'
+
+- name: R. Avery
+  affilnum: '5'
+
+- name: K. Caylor
+  affilnum: '5,6,7'
 
@@
 - affilnum: 1
-  affil: Graduate School of Geography, Clark University, 950 Main Street, Worcester, MA, USA
+  affil: Graduate School of Geography, Clark University,  Worcester, MA, USA
 
@@
 
-- affilnum: 2
+- affilnum: 3
+  affil: Clark Labs, Clark University, Worcester, MA, USA
+
+- affilnum: 4
   affil: SpatialCollective, Nairobi, Kenya
 
-- affilnum: 3
+- affilnum: 5
   affil: Department of Geography, University of California Santa Barbara, Santa Barbara, CA, USA
 
-- affilnum: 4
+- affilnum: 6
   affil: Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA, USA
 
-- affilnum: 5
+- affilnum: 7
   affil: Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA, USA
@@
 To overcome these challenges, we designed a system to minimize several sources of error that arise when mapping smallholder croplands.
-To overcome errors caused by mismatches between image resolution and cropland scales, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year.
+To overcome errors caused by mismatches between image resolution and cropland scales, as well as persistent cloud cover, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year.
 To reduce errors that occur when training classifiers, we built a labelling platform that rigorously assesses label accuracy, and creates more accurate consensus labels that train a Random Forests model.
-The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which combines with a segmentation process to delineate individual field boundaries.
+The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which is used in a segmentation process to delineate individual field boundaries.
 We applied this system to map Ghana’s croplands for the year 2018.
@@
 Croplands covered 16.1-23.2% of the mapped area, comprising 1,131,146 total fields with an average size of 3.92 ha.
-Unbiased estimates based on the map reference sample yield indicate a cropland percentage of 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on whether the cropland probability or field boundary map was assessed.
+Estimates based on the map reference sample indicate the cropland percentage is 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on the map used to estimate the standard error.
 Using the labellers’ digitized field boundaries to estimate biases in field boundary statistics, we calculated an adjusted mean field size of 1.73 ha and total field count of 1,662,281.
@@
 By minimizing training errors, consensus labelling improved the model’s F1 scores by up to 25%, while 3 iterations of active learning increased the F1 score by 9.1%, on average, which was 2.3% higher than training models with randomly selected labels.
-Map accuracy may be further improved by replacing Random Forests with a convolutional neural network.
+Map accuracy can be improved by replacing Random Forests with a convolutional neural network.
 These results demonstrate a readily adapted, transferrable framework for developing high resolution, annual, nation-scale maps that provide important details about smallholder-dominated croplands.
 
-keywords: rmarkdown, reproducible science
+keywords: active learning, CubeSat, smallholder, cropland, segmentation, Ghana, consensus labels, machine learning
+
 bibliography: 
@@
 
-\doublespace
+\singlespace
 
@@
 Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time.
-Unfortunately, the data that do exist are in many places inaccurate or missing.
+Unfortunately, the data that do exist are in many places inaccurate.
 Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
@@
 Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018].
-Beyond cropland distributions, few data are available on cropland characteristics such as field size, which provide important information on agricultural systems  [@lesivEstimatingGlobalDistribution2019].
-These inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021]
-
-<!-- - The problem in general  -->
-<!--     - Africa in particular -->
-<!--     - Types of analyses that depend on cropland maps -->
-
+Beyond cropland distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016;@levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
+
+These informational inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021].
 Cropland mapping over Africa is difficult for several reasons.
-The primary reason relates to the characteristics of the contintent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019].
+The primary reason relates to the characteristics of the continent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019].
 This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g.
@@
 
-These problems arising from cropland characteristics are increasingly being overcome due to technological advances.
+These problems arising from cropland characteristics are increasingly being addressed due to technological advances.
 Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
@@
 
-We describe here a cropland mapping system that follows these requirements on the way to delineating field boundaries.
+We describe here a cropland mapping system that follows these requirements, with an emphasis on delineating field boundaries.
 The first requirement is enabled by the recent availability of CubeSat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
@@
 
-The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020].
+The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters with a simplified feature set, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020].
 The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020].
@@
 
-We demonstrate this approach to map cropland in Ghana, a country where agriculture is predominantly small-scale and comprises a broad mix of climates and agricultural systems, including large areas where shifting agricultural practices predominate [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
+We demonstrate this approach to map cropland in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
 
@@
 ### Image compositing
-The image processing components of our system was designed to work with PlanetScope Analytic surface reflectance data [@PlanetTeamPlanetApplicationProgram2017].
+The image processing components of our system were designed to work with PlanetScope Analytic surface reflectance data [@planetteamPlanetApplicationProgram2018].
 PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency.
@@
 
-PlanetScope imagery is accessed via the Planet API, and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods.
-The imagery is collected and transferred directly to the system's cloud storage (Amazon Web Services [AWS] S3).
-
-Individual images are then transformed into "analysis ready data" (ARD) [@DwyerAnalysisReadyData2018], by cropping each downloaded scene into a 0.05 degree tile, regardless of cloud cover.
+PlanetScope imagery is accessed via the Planet API [@planetteamPlanetApplicationProgram2018], and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods.
+The imagery is collected and transferred directly to a cloud storage platform (Amazon Web Services [AWS] S3).
+
+Individual images are then transformed into analysis ready data (ARD) [@DwyerAnalysisReadyData2018], by subsetting each downloaded image into 0.05 degree tiles, regardless of cloud cover.
 Tiles are organized within a larger 1 degree resolution grid that covers the entire continent, which defines the minimum mapping area of interest (AOI; Figure \ref{fig:aois}).
@@
 The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year.
-Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in same year helps to minimize differences caused by land change.
+Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in the same year helps to minimize differences caused by land change.
 For each pixel in each image in each ARD temporal stack for a given season, two weights are calculated:
@@
 \mathrm{W2_t} =\begin{cases}
-    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_1}$, $\mathrm{NIR_2}$, ..., $\mathrm{NIR_i}$\}}.\\
+    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
     1, & \text{otherwise}.
@@
 Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
-The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$.
+The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy_Map}" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$.
 The web-rendered imagery is presented within the training data platform (next section).
@@
 ##### Mapping assignments
-Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps.
+Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps.
 Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays.
@@
 
-```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of mapper's interface", fig.align='center'}
+```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
 knitr::include_graphics('figures/figure3.png')
@@
 
-Where *i* indicates the particular assignment, and $\beta_{0-5}$ represent varying weights that sum to 1.
+Where *i* indicates the particular assignment, and $\beta_{0-4}$ represent varying weights that sum to 1.
 *I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the labeller's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the labeller delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the labeller's thematic labels (see SI for individual formulae).
@@
 ### Classification pipeline
-Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce cluster consisting of tens of instances, depending on the size of the AOI.
+Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf} cluster consisting of tens of instances, depending on the size of the AOI.
 
@@
 We followed this logic in this study, but were constrained to use a smaller feature space because the storage and memory requirements for our mapping geographies in this case were several orders of magnitude larger.
-For this implemention, we thus extract a set of 16 features, which are the 11X11 mean and 5X5 standard deviation of each band for both seasons (initial tests revealed these window sizes most effective), resulting in 24 overall features with the bands (Table 1).
+For this implemention, we thus extract a set of 16 features, which are the mean and standard deviations calculated within an 11X11 and 5X5 moving window, respectively (initial tests revealed these two window sizes to be most effective), resulting in 24 overall features, including the original bands (Table 1).
 
@@
 
-Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`$\footnote{https://github.com/locationtech/geotrellis}$, `rasterio`$\footnote{https://rasterio.readthedocs.io/en/latest/}$, and `RasterFrames`$\footnote{https://www.cogeo.org/}$.
+Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://www.cogeo.org/}.
 These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into `RasterFrames`, a data structure designed for handling raster data within Apache Spark.
-Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs$\footnote{https://www.cogeo.org/}$ (COGs), which allows rapid reading of image subsets.
+Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs), which allows rapid reading of image subsets.
 
@@
 
-After fitting, the model is applied to the features of the model validation set, and a set of performance metrics are calculated, including binary accuracy, the F1 score (the geometric mean of user's and producer's accuracy), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
+After fitting, the model is applied to the features of the model validation set, and a set of performance metrics is calculated, including binary accuracy, the F1 score (the geometric mean of precision and recall), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
 
@@
 After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to the labelling platform, which are then digitized by the labellers.
-The resulting consensus labels from the actively selected sample are added to the initial randomly selected sampe, and a new cluster is launched.
+The resulting consensus labels from the actively selected sample are added to the initial randomly selected sample, and a new cluster is launched.
 The model is retrained, assesses uncertainty across the remaining unlabelled sites, and selects the next most uncertain sites for labelling.
@@
 In the first step, the meanshift algorithm [@YizongChengMeanshiftmode1995a] is applied to the original bands of the dry season composite.
-A Sobel filter is then tapplied to the green, red, and near-infrared mean-shifted bands and the probability map, and an combined edge image is computed using the sum of these four edge images for the dry season only.
+A Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands and the probability map, and a combined edge image is computed using the sum of these four edge images for the dry season only.
 A compact watershed algorithm [@neubertCompactWatershedPreemptive2014] is then run on the weighted edge image, with a high level of segmentation specified.
@@
 We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@yeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and labellers are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
-On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult and potentially impossible to evaluate performance against the boundaries that farmers use to delineate their fields, as the conception of what constitutes a "field" could vary between farmers, while their field edges may be invisible or span multiple units that appear to a segmentation to have distinct boundaries.
+On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult to evaluate performance against real-world boundaries as the spectral distinction of field boundaries will vary across different crop types and land use arrangements.
 
 ## Applying the system to map Ghana
-To create a map of Ghana, we divided the country into 16 distinct AOIs, which were developed by grouping together each one degree cell fully contained within Ghana with the tiles belonging to any adjacent degree cell that overlapped neighboring countries (Figure \ref{fig:aois}A).
-The exception was AOI 16, which consisted of the four degree cells along Ghana's southern coast.
+We applied the system to map Ghana's croplands, excluding areas primarily cultivated with tree crops.
+Ghana has several distinct agricultural regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
+For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops.
+We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g.
+cocoa) are often not visible.
+
+To create the cropland maps, we divided the country into 16 distinct AOIs, which were developed by grouping together each one degree cell fully contained within Ghana with the tiles belonging to any adjacent degree cell that overlapped neighboring countries (Figure \ref{fig:aois}A).
+The exception was AOI 16, which consisted of the four degree cells intersecting Ghana's southern coast.
 The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), A separate active learning and segmentation process was run for each of these AOIs.
+
 To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16).
@@
 We evaluated the overall quality of the resulting seasonal image composites by assessing a random selection of `r tilen` tiles.
-We graded both seasonal composites for each tile using a four category quality score, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its  resolution was degraded below the typical 3-4 m PlanetScope resolution (e.g.
+We graded both seasonal composites for each tile using a four category quality score, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g.
 because of between-date image mis-registrations).
@@
 ### Accounting for label error
-To quantify the potential impact of label error on our results, we evaluated the performance differences between models trained with three different sets of labels: 1) those from the lowest scoring labeller to map each training site, 2) those from the highest scoring labeller, and 3) the consensus labels.
+To quantify the potential impact of label error on classification results, we evaluated the performance differences between models trained with three different sets of labels: 1) those from the lowest scoring labeller to map each training site, 2) those from the highest scoring labeller, and 3) the consensus labels.
 This assessment follows recommended Tier 1 (i.e.
@@
 To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against those of the  workers' digitized polygons at map validation sites.
-We selected the field polygons from the most accurate labeller to digitize each of the 100 validation sites in each AOI, and calculated the site-wise average area and number of polygons.
+We chose this approach because of existing uncertainties in polygon-based accuracy assessment methods [@yeReviewAccuracyAssessment2018], and because the map's ability to represent field sizes was of greatest interest.
+To undertake this comparison, we selected the field polygons from the most accurate labeller to digitize each of the 100 validation sites in each AOI, and calculated the site-wise average area and number of polygons.
 We then calculated the same statistics from the segmented boundaries that intersected each validation grid.
@@
 ```
-We developed maps of Ghana's croplands, covering an area of `r km2` km$^2$ that included portions of neighboring countries overlapped by images tiles.
-Ghana has several agro-ecologically distinct regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
-For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops.
-We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g.
-cocoa) are often not visible.
+We developed maps of Ghana's cultivated croplands within an area of `r km2` km$^2$, which included portions of neighboring countries overlapped by images tiles.
 
@@
 
-To develop the map, we first generated an image catalog for the `r nt` tiles covering Ghana.
+To develop the maps, we first generated an image catalog for the `r nt` tiles covering Ghana.
 This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
@@
 In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
-Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often falling along ecotones, such as the boundaries between agroecozones (see Figure S3A).
+Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agroecozones (see Figure S3A).
 The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.
@@
 The negative results for accuracy was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:alperformance}).
-In comparison, iteration 1 active learning accuracies 1 were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
+In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
 Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.
@@
 \begin{table}
-\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions; top 5 rows) and the segmented map (bottom 5 rows).
+\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on Random Forests predictions; top 5 rows) and the segmented map (bottom 5 rows).
 Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country.
@@
 Two separate maps of cropland were produced for each AOI, a per-pixel map derived from the cropland probabilities, and the vectorized map of field boundaries (Figure \ref{fig:mainmap}).
-The former provides the more picture of cropland distributions in Ghana, which are most concentrated in the Southeastern corner (AOI 15), the central-western region (AOI 7, the northeastern and northwestern corners of AOIs 10 and 11, and the south of AOI 8), and the northeastern quadrant stretching from AOI 9 through AOIs 5 and 6 and up to AOIs 2 and 3.
+The former provides the more accurate picture of cropland distributions in Ghana, which are most concentrated in the Southeastern corner (AOI 15), the central-western region (AOI 7, the northeastern and northwestern corners of AOIs 10 and 11, and the south of AOI 8), and the northeastern quadrant stretching from AOI 9 through AOIs 5 and 6 and up to AOIs 2 and 3.
 The northern third of AOI 1 also has noticeable densities of cropland.
-Several noteable areas of low cropland density are also apparent, indicating the presence of large protected areas, such as Mole National Park in the the southeastern corner of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
-In contrast, the relative absence of cropland in AOIs 13, 14, and 16 does not reflect the scarcity of agriculture, but rather the predominance of tree crops, which we did not map, in these areas.
+Several notable areas of low cropland density are also apparent, indicating the presence of large protected areas, such as Mole National Park in the southeastern corner of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
+In contrast, the relative absence of cropland in AOIs 13, 14, and 16 does not reflect the scarcity of agriculture in these areas, but rather the predominance of tree crops, which we did not map.
 
@@
 Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
-The overall map accuracies are decent (86.7-88%, Table \ref{tab:mapaccuracy}), and are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given the inherent uncertainty in the map reference sample, our best estimate of the "truth", in which we have roughly 85% confidence.
+The overall map accuracies (86.7-88%, Table \ref{tab:mapaccuracy}) are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given the inherent uncertainty in the map reference sample, our best estimate of the "truth", in which we have roughly 85% confidence.
 However, accuracies for the cropland class were much lower, falling between 62 (Producer's) to 67 (User's) percent country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}), meaning the model produced substantial commission and omission errors for this class.
@@
 The patterns of cropland-class accuracies varied by zone.
-These zones largely align, albeit with some discrepancies, with the country's agroecoregions, thus the accuracy patterns may be partially attributed to some regions being harder to map than others.
+These zones largely align, albeit with some discrepancies, with the country's agroecozones, thus the accuracy patterns may be partially attributed to some regions being harder to map than others.
 Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S9), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas.
@@
 Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error.
-Maps in the two northern savanna agroecoregions had the best balance between omission and commission error, and had the highest overall User's accuracy.
-The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S3), within which it typically covered a smaller share of area relative to the other agroecoregions.
+Maps in the two northern savanna agroecozones had the best balance between omission and commission error, and had the highest overall User's accuracy.
+The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S3), within which it typically covered a smaller share of area relative to the other agroecozones.
 This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S3B).
@@
 @waldnerDeepLearningEdge2020].
-Over-merging was another source of error in the segmentation algorithm, which in some areas led to overestimated field sizes and irregular shapes, particularly in high density croplands (e.g.
+Over-merging was another source of error in the segmentation algorithm, which in some areas led to overestimated field sizes and unrealistic shapes, particularly in high density croplands (e.g.
 in AOIs 2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent fields are often indistinct in the PlanetScope imagery.
@@
 The first is to replace Random Forests with a more advanced convolutional neural network (CNN), which can generate and learn from a large number of features representing a variety of spatial scales [@maDeepLearningRemote2019].
-Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, are effective in delineating field boundaries [@waldnerDeepLearningEdge2020].
+Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, is effective in delineating field boundaries [@waldnerDeepLearningEdge2020].
 Our system can readily incorporate such a model.
-The active learning framework is already analagous to transfer learning, in which a trained deep learning model is applied to a different region and updated with local training samples [@panSurveyTransferLearning2010; @russwurmMetalearningFewshotLand2020].
-The existing tooling can thus be adjusted to guide training site selection in sub-regions of Ghana for a CNN trained from scratch from a larger initial random draw.
-Instead of the current AOIs, a more effective strategy may be to update the model within agroecozones (Figure S9B) The labelling methods already provide for a case in which edge and interior classes need to be labelled, as the most accurate individual training samples can be identified and used, rather than binary consensus labels.
-The segmentation algorithm will also be updated to work with the resulting score maps.
+The labelling platform already provides the methods needed to develop and assess the quality of labels that include field edges and interior classes, while active learning has proven to be effective for training deep learning models [@caoHyperspectralImageClassification2020; @liuActiveDeepLearning2017].
+Our current framework can be adjusted so that it starts by training a CNN from scratch with a large initial random sample, and then uses a transfer learning approach [@panSurveyTransferLearning2010] to update the model with the most informative samples from different AOIs or agroecozones.
 
 ## Conclusion
-This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve.
+This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve.
+Maps that include information on field boundaries can help improve remote estimation of crop planted area and yield, and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and livelihoods.
 Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.
@@
 We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
-We thank Manushi Trivedi, Sitian Xiong, and Tammy Woodard are thanked for their contributions to the underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and Primoz Kovacic for support and management of the labelling efforts.
+We thank Manushi Trivedi, Sitian Xiong, and Tammy Woodard for their contributions to the underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and Primoz Kovacic for support and management of the labelling efforts.
 

### 2021-03-07 inst/manuscript/si.Rmd
Edits to ms and si, plus commit of many si figures

@@
 geometry: margin=1in
-linenumbers: false
+linenumbers: yes
 spacing: doublespacing
@@
 
-Where $\beta_{0-5}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-5}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.
+Where $\beta_{0-4}$ are the weights assigned to each accuracy metric that sum to 1. For the current production version, $\beta_{0-4}$ were assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.
 
@@
 Looking in greater details at this equation, the first half of the righthand side of the equation, P(W$_i$|D), is the "prior" for worker *i* for the current site based on their history of scores from accuracy assessment assignments.
-The second term, P($\theta$|D, W$_i$), is the probability that the actual class of the pixel in the current assignment is the class that worker *i*'s says that it is, which is either 0 or 1.
+The second term, P($\theta$|D, W$_i$), is the probability that the actual class of the pixel in the current assignment is the class that worker *i* says that it is, which is either 0 or 1.
 There are four possible values for this second term:
@@
 
-Equations \ref{eq:pacc1} and \ref{eq:pacc2} are producer's accuracies, thus the maximum worker likelihood is equivalent to the worker's average producer's accuracy.
+Equations \ref{eq:pacc1} and \ref{eq:pacc2} are Producer's accuracies, thus the maximum worker likelihood is equivalent to the worker's average Producer's accuracy.
 
@@
 
-The other component of equation \ref{eq:prior}, P(W$_i$), is the worker's average score over *m* accuracy assessment assignment:  
+The other component of equation \ref{eq:prior}, P(W$_i$), is the worker's average score over *m* accuracy assessment assignments:  
 \begin{equation} 
@@
 
-Thus equation \ref{eq:prior} uses two measures of worker accuracy, 1) their overall average accuracy score multipled by 2) their average producer's accuracy to create a *weight* for their individual maps for the given site.
+Thus equation \ref{eq:prior} uses two measures of worker accuracy, 1) their overall average accuracy score multipled by 2) their average Producer's accuracy to create a *weight* for their individual maps for the given site.
 Equation \ref{eq:main} becomes:
@@
 
-\clearpage
 $$
@@
 
-```{r labelriskhist, echo = FALSE, out.width="45%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
+```{r labelriskhist, echo = FALSE, out.width="40%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
 The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}

### 2021-03-08 inst/manuscript/manuscript.Rmd
Small edits to ms and si, separated raw data script for label_summary.R, various uncommitted files added

@@
 - name: D. McRitchie
-  affilnum: ''
+  affilnum: '4'
 
 - name: S. Debats
-  affilnum: ''
+  affilnum: '4'
 
 - name: J. Muhando
-  affilnum: '4'
+  affilnum: '5'
 
 - name: A. Hokar
-  affilnum: '4'
+  affilnum: '5'
 
 - name: C. Juma 
-  affilnum: '4'
+  affilnum: '5'
 
 - name: J. Makuru
-  affilnum: '4'
+  affilnum: '5'
 
 - name: F. Matasa
-  affilnum: '4'
+  affilnum: '5'
 
 - name: B. Mbatia
-  affilnum: '4'
+  affilnum: '5'
 
 - name: J. Mbaye
-  affilnum: '4'
+  affilnum: '5'
 
 - name: I. Muasa
-  affilnum: '4'
+  affilnum: '5'
 
 - name: J. Mucha
-  affilnum: '4'
+  affilnum: '5'
 
 - name: A. Mugami
-  affilnum: '4'
+  affilnum: '5'
 
 - name: P. Oduor
-  affilnum: '4'
+  affilnum: '5'
 
 - name: J. Ochieng
-  affilnum: '4'
+  affilnum: '5'
 
 - name: F. Wambua
-  affilnum: '4'
+  affilnum: '5'
 
 - name: T. Wanjiku
-  affilnum: '4'
+  affilnum: '5'
 
 - name: J. Wanyoike 
-  affilnum: '4'
+  affilnum: '5'
 
 - name: R. Avery
-  affilnum: '5'
+  affilnum: '6'
 
 - name: K. Caylor
-  affilnum: '5,6,7'
+  affilnum: '6,7,8'
 
@@
 - affilnum: 4
+  affil: Independent contributor
+
+- affilnum: 5
   affil: SpatialCollective, Nairobi, Kenya
 
-- affilnum: 5
+- affilnum: 6
   affil: Department of Geography, University of California Santa Barbara, Santa Barbara, CA, USA
 
-- affilnum: 6
+- affilnum: 7
   affil: Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA, USA
 
-- affilnum: 7
+- affilnum: 8
   affil: Bren School of Environmental Science and Management, University of California Santa Barbara, Santa Barbara, CA, USA
@@
 geometry: margin=1in
-linenumbers: yes
+linenumbers: no
 spacing: doublespacing
-footerdate: yes
+footerdate: no
+note: This paper is a non-peer reviewed preprint submitted to EarthArXiv. It will be periodically updated as it is revised, including information on its peer review status. 
 abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands.
@@
 
-keywords: active learning, CubeSat, smallholder, cropland, segmentation, Ghana, consensus labels, machine learning
+keywords: active learning, machine learning, CubeSat, PlanetScope, smallholder, cropland, segmentation, field size, consensus labels, Ghana, Africa
 
@@
 Our system can readily incorporate such a model.
-The labelling platform already provides the methods needed to develop and assess the quality of labels that include field edges and interior classes, while active learning has proven to be effective for training deep learning models [@caoHyperspectralImageClassification2020; @liuActiveDeepLearning2017].
+The labelling platform already provides the methods needed to develop and assess the quality of labels that include field edges and interior classes, while active learning has proven to be effective for optimizing training datasets for deep learning models [@caoHyperspectralImageClassification2020; @liuActiveDeepLearning2017].
 Our current framework can be adjusted so that it starts by training a CNN from scratch with a large initial random sample, and then uses a transfer learning approach [@panSurveyTransferLearning2010] to update the model with the most informative samples from different AOIs or agroecozones.

### 2021-03-08 inst/manuscript/si.Rmd
Small edits to ms and si, separated raw data script for label_summary.R, various uncommitted files added

@@
 ### Categorical accuracy
-In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).
+In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones\footnote{sourced from https://wheregeospatial.com/agro-ecological-zones-ghana/} (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).
 

### 2021-03-08 data-raw/data_from_instances.R

@@ -259,69 +259,6 @@ train_val_sites <- list("initial" = initial_train_pts,
 usethis::use_data(train_val_sites, overwrite = TRUE)
 
 #------------------------------------------------------------------------------#
-# Results from all instances: Average N labels and label accuracy
-load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
-
-# instance_tbls$hit_data %>%
-#   filter(aoi == "labeller3" & delete_time > "2019-12-05") %>%
-#   left_join(., instance_tbls$assignment_data %>% filter(aoi == "labeller3"),
-#             by = c("aoi", "hit_id")) %>%
-#   filter(!is.na(score))
-# instance_tbls$assignment_data %>%
-#   filter(!(aoi == "labeller3" & completion_time < "2019-12-05" & !is.na(score)))
-
-# select scores from Q assignments, dropping Q sites in labeller3 done during
-# run 0. Filter by completion_time earlier than 5/12/2019, which was a few days
-# before run 1 started. Anything after that is a Q assignment done during run 1.
-# Also including Q sites done during labeller, labellertrainsouth,
-# labellertrainsouth, labellertrainsouthnl, and labellertrainsw. labellernl not
-# included because not used for map-making
-
-# aoinms <- paste0("labeller", 1:16)  # original
-aoinms <- c(
-  "labeller",
-  paste0("labeller", c("trainsouth", "trainsouthnl", "trainsw", 1:16))
-)
-quality_scores <- instance_tbls$assignment_data %>%
-  select(-one_of("comment", "status")) %>%
-  filter(!is.na(score)) %>% rename(id = worker_id) %>%
-  filter(aoi %in% aoinms) %>%
-  filter(!(aoi == "labeller3" & completion_time < "2019-12-05")) %>%
-  left_join(., instance_tbls$users, by = c("aoi", "id")) %>%
-  left_join(., instance_tbls$hit_data %>% select(aoi, hit_id, name)) %>%
-  select(-contains("time")) %>% select(aoi, name, !!names(.))
-
-# summary statistics for label quality (from Q sites)
-quality_sum <- quality_scores %>% group_by(uuid) %>%
-  summarize(N = n(), Mean = mean(score), StDev = sd(score)) %>%
-  arrange(-Mean)
-
-# overall number of labels
-nlabels <- instance_tbls$assignment_data %>%
-  filter(aoi %in% aoinms) %>%
[…]

### 2021-03-08 data-raw/label_summary.R

@@ -1,0 +1,63 @@
+library(dplyr)
+
+# Results from all instances: Average N labels and label accuracy
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+# instance_tbls$hit_data %>%
+#   filter(aoi == "labeller3" & delete_time > "2019-12-05") %>%
+#   left_join(., instance_tbls$assignment_data %>% filter(aoi == "labeller3"),
+#             by = c("aoi", "hit_id")) %>%
+#   filter(!is.na(score))
+# instance_tbls$assignment_data %>%
+#   filter(!(aoi == "labeller3" & completion_time < "2019-12-05" & !is.na(score)))
+
+# select scores from Q assignments, dropping Q sites in labeller3 done during
+# run 0. Filter by completion_time earlier than 5/12/2019, which was a few days
+# before run 1 started. Anything after that is a Q assignment done during run 1.
+# Also including Q sites done during labeller, labellertrainsouth,
+# labellertrainsouth, labellertrainsouthnl, and labellertrainsw. labellernl not
+# included because not used for map-making
+
+# aoinms <- paste0("labeller", 1:16)  # original
+aoinms <- c(
+  "labeller",
+  paste0("labeller", c("trainsouth", "trainsouthnl", "trainsw", 1:16))
+)
+quality_scores <- instance_tbls$assignment_data %>%
+  select(-one_of("comment", "status")) %>%
+  filter(!is.na(score)) %>% rename(id = worker_id) %>%
+  filter(aoi %in% aoinms) %>%
+  filter(!(aoi == "labeller3" & completion_time < "2019-12-05")) %>%
+  left_join(., instance_tbls$users, by = c("aoi", "id")) %>%
+  left_join(., instance_tbls$hit_data %>% select(aoi, hit_id, name)) %>%
+  select(-contains("time")) %>% select(aoi, name, !!names(.))
+
+# summary statistics for label quality (from Q sites)
+quality_sum <- quality_scores %>% group_by(uuid) %>%
+  summarize(N = n(), Mean = mean(score), StDev = sd(score)) %>%
+  arrange(-Mean)
+
+# overall number of labels
+nlabels <- instance_tbls$assignment_data %>%
+  filter(aoi %in% aoinms) %>%
+  left_join(., instance_tbls$hit_data, by = c("aoi", "hit_id")) %>%
[…]

### 2021-03-09 inst/manuscript/figures.Rmd
Edited ms and si: first upload to EarthArXiv. Edited scripts for data_for_instances.R, added new script to data-raw (aois_tiles.R) to show creation of degree and tiling grids

@@
 
-### Figure S3: Initial training site distribution
+### Figure S3: PlanetScope image density
+The density of imagery available for download from the image catalog.
+
+```{r imagedensity, eval=FALSE}
+img_density <- st_read(
+  system.file("extdata/img_catalog_count_201712_201902.geojson", 
+              package = "activemapper")
+)
+
+tile_count <- img_density %>% mutate(season = case_when(
+  year == 2018 & month %in% 5:9 ~ "GS",
+  year == 2018 & month %in% 12 ~ "OS", 
+  year == 2019 & month %in% 1:2 ~ "OS", 
+  TRUE ~ NA_character_)
+) %>% filter(season %in% c("GS", "OS")) %>% 
+  as_tibble() %>%
+  group_by(tile, season) %>%
+  summarize(Count = sum(count))
+
+gs_count <- tiles %>% dplyr::select(tile) %>% 
+  left_join(., tile_count %>% filter(season == "GS"))
+os_count <- tiles %>% dplyr::select(tile) %>% 
+  left_join(., tile_count %>% filter(season == "OS")) # %>%
+  # fasterize::fasterize(., field = "count")
+
+pal <- colorRampPalette(rev(RColorBrewer::brewer.pal(9, "RdBu")))
+zlim <- range(c(gs_count$Count, os_count$Count))
+p <- ggplot() + 
+  geom_sf(data = gs_count, aes(fill = Count), col = "transparent") + 
+  geom_sf(data = aois, fill = "transparent", size = 0.2) + 
+  scale_fill_gradientn(colours = pal(20), limits = zlim) +
+  ggtitle("May - Sept 2018") + 
+  theme_void() + 
+  theme(legend.position = "none", 
+        title = element_text(size = 9))
+
+p2 <- ggplot() + 
+  geom_sf(data = os_count, aes(fill = Count), col = "transparent") + 
+  scale_fill_gradientn(colours = pal(20), limits = zlim) +
+  geom_sf(data = aois, fill = "transparent", size = 0.2) + 
+  ggtitle("Dec 2018 - Feb 2019") +
+  theme_void() + 
+  theme(title = element_text(size = 9),
+        legend.text = element_text(size = 9))
+
+op <- p + p2
+ggsave(op, bg = "transparent",
+       filename = pth("figures/si_planet_coverage.png"), 
+       width = 5, height = 3, dpi = 300, units = "in")
+
+```
+
+### Figure S4: Initial training site distribution
 Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
@@
 
-### Figure S4: Label count and quality
+### Figure S5: Label count and quality
 ```{r labelquality, eval=FALSE}
@@
 
-### Figures S5 and S6: Consensus label (Bayesian) risk
+### Figures S6 and S7: Consensus label (Bayesian) risk
 
@@
 
-### Figure S7: Random versus active sampling
+### Figure S8: Random versus active sampling
 ```{r, eval=FALSE}
@@
 
-### Figure S8: Training data error impact on probability images
+### Figure S9: Training data error impact on probability images
 ```{r, eval = FALSE}
@@
 
-### Figure S9: AOI Zones
+### Figure S10: AOI Zones
 
@@
 
-### Figures S10 and S11: Segmentation statistics (grid scale)
+### Figures S11 and S12: Segmentation statistics (grid scale)
 
@@
 
-### Figure S12: Average fields sizes
+### Figure S13: Average fields sizes
 
@@
 ```
-
-Both original as derived from segmentations and adjusted using regression kriging.
-
-```{r, eval=FALSE}
-fsizes <- read_stars(
-  system.file("extdata/mean_field_size005.tif", package = "activemapper")
-)
-fsizes_adj <- read_stars(
-  system.file("extdata/mean_field_size005_adj.tif", package = "activemapper")
-)
-
-# Plots
-xylims <- st_bbox(ghana)
-p <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = fsizes) +
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  scale_fill_viridis_c(name = "ha", na.value = "transparent") +
-  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) +
-  theme_void() + 
-  theme(legend.position = c(0.7, 0.07), legend.direction = "horizontal", 
-        legend.key.height = unit(0.2, "cm"), 
-        legend.text = element_text(colour = "black"),
-        legend.title = element_text(size = 9, colour = "black"))
-
-p2 <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
-  geom_stars(data = fsizes_adj[, , , 1]) +
-  geom_sf(data = aois, size = 0.1, fill = "transparent", col = "grey") +
-  scale_fill_viridis_c(name = "ha", na.value = "transparent") +
-  coord_sf(xlim = xylims[c(1, 3)], ylim = xylims[c(2, 4)]) +
-  theme_void() + 
-  theme(legend.position = c(0.7, 0.07), legend.direction = "horizontal", 
-        legend.key.height = unit(0.2, "cm"), 
-        legend.text = element_text(colour = "black"),
-        legend.title = element_text(size = 9, colour = "black"))
-
-op <- p + p2 + plot_annotation(tag_levels = 'A')
-
-ggsave(op, bg = "transparent", width = 6, height = 4, units = "in", dpi = 300, 
-       filename = pth("figures/si_fsizes.png"))
-
-```
-
-<!-- ### Figure S10: Segmentation statistics, tile scale -->
-<!-- ```{r, eval=FALSE} -->
-<!-- data("tile_aoi_validation_stats") -->
-
-<!-- humachine <- tile_aoi_validation_stats$tile %>%  -->
-<!--   mutate(type = ifelse(grepl("segments", type), "Segmentation", type), -->
-<!--          type = ifelse(grepl("validation", type), "Validation", type)) %>%  -->
-<!--   select(type, aoi, tile, area_mu, area_med, n_mu) %>%  -->
-<!--   rename(Mean = area_mu, Median = area_med) -->
-
-<!-- humachine_sum <- humachine %>% group_by(aoi, type) %>%  -->
-<!--   summarize(Median = median(Mean), Mean = mean(Mean)) %>%  -->
-<!--   ungroup() %>%  -->
-<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") -->
-
-<!-- humachine_sum_n <- humachine %>% group_by(aoi, type) %>%  -->
-<!--   summarize(Median = median(n_mu), Mean = mean(n_mu)) %>%  -->
-<!--   ungroup() %>%  -->
-<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") -->
-
-<!-- # Field size -->
-<!-- brks <- seq(0, 20, 5) -->
-<!-- p <- ggplot(humachine) +  -->
-<!--   geom_density(aes(x = Mean, y = ..count.., fill = type), col = "transparent",  -->
-<!--                linetype = "blank") +   -->
-<!--   geom_vline(data = humachine_sum,  -->
-<!--              aes(xintercept = value, color = type, linetype = Stat),  -->
-<!--              size = 0.5) + -->
-<!--   scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 20)) + -->
-<!--   scale_y_continuous(expand = c(0, 0)) + -->
-<!--   scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) +  -->
-<!--   scale_fill_manual(name = "", values = fcols) +   -->
-<!--   # scale_linetype_manual(name = "", values = c("dashed", "solid")) +   -->
-<!--   scale_linetype("") +   -->
-<!--   facet_wrap(vars(aoi)) +  -->
-<!--   xlab("Field size (ha)") + ylab("Count") +  -->
-<!--   theme_fun() -->
-<!-- ggsave(pth("figures/si_tile_validation_stats_fha.png"),  -->
-<!--        width = 7, height = 7, units = "in", dpi = 300) -->
-
-<!-- # field number -->
-<!-- brks <- seq(0, 20, 5) -->
-<!-- p2 <- ggplot(humachine) +  -->
-<!--   geom_density(aes(x = n_mu, y = ..count.., fill = type), col = "transparent",  -->
-<!--                linetype = "blank") +   -->
-<!--   geom_vline(data = humachine_sum_n,  -->
-<!--              aes(xintercept = value, color = type, linetype = Stat),  -->
-<!--              size = 0.5) + -->
-<!--   scale_x_continuous(expand = c(0.03, 0.03), breaks = brks, limits = c(0, 20)) + -->
-<!--   scale_y_continuous(expand = c(0, 0)) + #, limits = c(0, 40)) + -->
-<!--   scale_color_manual(name = "", values = c("blue", "red"), guide = FALSE) +  -->
-<!--   scale_fill_manual(name = "", values = fcols) +   -->
-<!--   # scale_linetype_manual(name = "", values = c("dashed", "solid")) +   -->
-<!--   scale_linetype("") +   -->
-<!--   facet_wrap(vars(aoi)) +  -->
-<!--   xlab("Number of fields") + ylab("Count") +  -->
-<!--   theme_fun() -->
-<!-- ggsave(pth("figures/si_tile_validation_stats_fn.png"),  -->
-<!--        width = 7, height = 7, units = "in", dpi = 300) -->
-<!-- ``` -->
-
-<!-- ### Figure S11: Segmentation statistics: grid, tile, and AOI scales -->
-
-<!-- Read back in segments to get distributions -->
-<!-- ```{r, eval = FALSE} -->
-<!-- data("tile_aoi_validation_stats") -->
-
-<!-- # # Read in segments, drop old AOI3 -->
-<!-- # segment_files <- dir(here("external/data/results/segments/merged"), -->
-<!-- #                      pattern = "boundarymerge", full.names = TRUE) -->
-<!-- # iid <- as.numeric(gsub("[[:alpha:]]|_|\\.", "", basename(segment_files))) -->
-<!-- # segment_files <- tibble(file = segment_files, order = iid) %>% arrange(iid) -->
-<!-- # segment_files <- segment_files %>% filter(!grepl("aoi3_*.*merge.geojson", file)) -->
-<!-- #  -->
-<!-- # # area and quiet func -->
-<!-- # ha <- function(x) as.numeric(units::set_units(st_area(x), "ha")) -->
-<!-- # quiet <- function(x) suppressWarnings(suppressMessages(x)) -->
-
-<!-- humachine_sum <- tile_aoi_validation_stats$aoi %>%  -->
-<!--   select(type, aoi, area_mu, area_med) %>%  -->
-<!--   rename(Mean = area_mu, Median = area_med) %>%  -->
-<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") %>%  -->
-<!--   tidyr::pivot_wider(names_from = type, values_from = value) -->
-
-<!-- aoi_wide <- tile_aoi_validation_stats$aoi %>% group_by(type) %>%  -->
-<!--   summarize(Mean = weighted.mean(area_mu, w = n),  -->
-<!--             Median = weighted.mean(area_med, w = n)) %>%  -->
-<!--   tidyr::pivot_longer(cols = Median:Mean, names_to = "Stat") %>%  -->
-<!--   tidyr::pivot_wider(names_from = type, values_from = value) %>%  -->
-<!--   mutate(aoi = "All") %>% select(aoi, !!names(.)) -->
-
-<!-- p <- ggplot(humachine_sum) + -->
-<!--   geom_point(aes(x = segments, y = validation)) +  -->
-<!--   geom_point(data = aoi_wide, aes(x = segments, y = validation), color = "red") +  -->
-<!--   scale_x_continuous(limits = c(0, 8)) +  -->
-<!--   scale_y_continuous(limits = c(0, 8)) +  -->
-<!--   ggrepel::geom_text_repel( -->
-<!--     aes(label = aoi, x = segments, y = validation), col = "grey50",  -->
-<!--     segment.color = "grey60", point.padding = 0.1, min.segment.length = 0.2, -->
-<!--     segment.size = 0.2, size = 3    -->
-<!--   ) +  -->
-<!--   xlab("Segmentation field size (ha)") + ylab("Validation field size (ha)") +  -->
-<!--   geom_abline(slope = 1) + facet_wrap(vars(Stat)) +  -->
-<!--   theme_linedraw() +  -->
-<!--   theme(strip.background = element_rect(fill = "white"),  -->
-<!--         strip.text = element_text(colour = "black")) -->
-
-<!-- ``` -->

### 2021-03-09 inst/manuscript/manuscript.Rmd
Edited ms and si: first upload to EarthArXiv. Edited scripts for data_for_instances.R, added new script to data-raw (aois_tiles.R) to show creation of degree and tiling grids

@@
 
-- name: L.D. Estes
+- name: Lyndon D. Estes
   affilnum: '1'
@@
 
-- name: S. Ye
+- name: Su Ye
   affilnum: '1,2'
 
-- name: L. Song
+- name: Lei Song
   affilnum: '1'
 
-- name: B. Luo
+- name: Boka Luo
   affilnum: '1,3'
@@
 
-- name: Z. Meng
+- name: Zhenhua Meng
   affilnum: '1'
 
-- name: Q. Zhang
+- name: Qi Zhang
   affilnum: '1'
 
-- name: D. McRitchie
+- name: Dennis McRitchie
   affilnum: '4'
 
-- name: S. Debats
+- name: Stephanie R. Debats
   affilnum: '4'
 
-- name: J. Muhando
-  affilnum: '5'
-
-- name: A. Hokar
-  affilnum: '5'
-
-- name: C. Juma 
-  affilnum: '5'
-
-- name: J. Makuru
-  affilnum: '5'
-
-- name: F. Matasa
-  affilnum: '5'
-
-- name: B. Mbatia
-  affilnum: '5'
-
-- name: J. Mbaye
-  affilnum: '5'
-
-- name: I. Muasa
-  affilnum: '5'
-
-- name: J. Mucha
-  affilnum: '5'
-
-- name: A. Mugami
-  affilnum: '5'
-
-- name: P. Oduor
-  affilnum: '5'
-
-- name: J. Ochieng
-  affilnum: '5'
-
-- name: F. Wambua
-  affilnum: '5'
-
-- name: T. Wanjiku
-  affilnum: '5'
-
-- name: J. Wanyoike 
-  affilnum: '5'
-
-- name: R. Avery
+- name: Justus Muhando
+  affilnum: '5'
+
+- name: Angeline H. Amukoa
+  affilnum: '5'
+
+- name: Brian W. Kaloo
+  affilnum: '5'
+
+- name: Jackson Makuru
+  affilnum: '5'
+
+- name: Ben K. Mbatia
+  affilnum: '5'
+
+- name: Isaac M. Muasa
+  affilnum: '5'
+
+- name: Julius Mucha
+  affilnum: '5'
+
+- name: Adelide M. Mugami
+  affilnum: '5'
+
+- name: Judith M. Mugami
+  affilnum: '5'
+
+- name: Francis W. Muinde
+  affilnum: '5'
+
+- name: Fredrick M. Mwawaza
+  affilnum: '5'
+
+- name: Jeff Ochieng
+  affilnum: '5'
+
+- name: Charles J. Oduol
+  affilnum: '5'
+
+- name: Purent Oduor
+  affilnum: '5'
+
+- name: Thuo Wanjiku
+  affilnum: '5'
+
+- name: Joseph G. Wanyoike 
+  affilnum: '5'
+
+- name: Ryan B. Avery
   affilnum: '6'
 
-- name: K. Caylor
+- name: Kelly K. Caylor
   affilnum: '6,7,8'
@@
 Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
-The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview} 1), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy_Map}" rendering is created for each composite using Raster Foundry$\footnote{https://github.com/raster-foundry/raster-foundry}$.
+The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview}), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering is created for each composite using Raster Foundry [@azaveaRasterFoundry2020].
 The web-rendered imagery is presented within the training data platform (next section).
@@
 The basic structure of the system remains the same, but we converted it into a standalone platform that allows us to enroll and pay people directly for their labelling, and is designed to control and supervise the machine learning process.
-The platform runs on a Linux virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (RasterFoundry), and a set of utilities for managing, assessing, and converting digitization work into rasterized labels for training a machine learning algorithm.
+The platform runs on a Linux virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (Raster Foundry), and a set of utilities for managing, assessing, and converting digitization work into rasterized labels for training a machine learning algorithm.
 Each instance of the platform focuses on a specific AOI (Figure \ref{fig:aois}A)
@@
 
-```{r labeller, echo = FALSE, out.width="100%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
+```{r labeller, echo = FALSE, out.width="95%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
 knitr::include_graphics('figures/figure3.png')
@@
 
-\begin{center}Table 1. List of features.\end{center}
-
+\begin{center}Table 1. List of image features.\end{center}
 Feature             Window Size  N Features
@@
 ------------------ ------------- ------------
-
-Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://www.cogeo.org/}.
-These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into `RasterFrames`, a data structure designed for handling raster data within Apache Spark.
-Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs), which allows rapid reading of image subsets.
+Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://rasterframes.io/}.
+These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into Apache Spark DataFrames.
+Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs).
+
+\clearpage
 
 #### Classification
-
 Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
@@
 This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
-The longer period was necessary for the growing season because of the frequent cloud cover, which severely limits the number of clear scenes for any tile.
+The longer period was necessary for the growing season because of the frequent cloud cover, which substantially limits the number of clear scenes for any tile (Figure S3).
 For the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
@@
 An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85).
-Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period.
+Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S3).
 
@@
 AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
-In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S3A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
-Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agroecozones (see Figure S3A).
+In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
+Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agroecozones (see Figure S4A).
 The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.
@@
 As each training site was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made.
-Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S4A for more details on labelling effort).
-Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S3A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
-The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S4B for detailed score distributions).
+Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S5A for more details on labelling effort).
+Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S4A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
+The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S5B for detailed score distributions).
 
@@
 The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
-Risk was highest in the northen AOIs (AOIs 1-6; Figures S5-6), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S5-6), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
+Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
 Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
@@
 
-```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall Accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
+```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/figure5.png')
@@
 The comparison of active versus randomized training sample collection (in AOIs 1, 8, and 15) showed that the former approach outperformed the latter.
-After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S7).
+After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S8).
 However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2.
@@
 We found that the overall accuracy of the pixel-wise classifications was 88% against this map reference sample (Table \ref{tab:mapaccuracy}).
-Confining the map reference sample to four distinct zones (Figure S9A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
+Confining the map reference sample to four distinct zones (Figure S10A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
 The Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1, while the User's accuracy for was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1.
@@
 The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
-This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S10), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
-The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S11).
+This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S11), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
+The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S12).
 
@@
 The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
-To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S12).
-These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally occurring in areas of higher field density, although the biases inherent in both measures (Figures S10-11) complicate the interpretation of those variations.
+To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S13).
+These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally occurring in areas of higher field density, although the biases inherent in both measures (Figures S11-12) complicate the interpretation of those variations.
 However, we can use the estimated biases to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana overall (Table \ref{tab:sizentab}).
@@
 The most extensive studies to date used crowdsourced volunteers to classify fields into broad size classes, based on their interpretations of imagery sampled from high resolution virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
-Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha$\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class.
+Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class.
 Data sourced from Table S3 in Lesiv et al.
-2019.}$.
+2019.}.
 This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean (`r aoi_nofor` ha) within AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample appears to have been collected.
@@
 These zones largely align, albeit with some discrepancies, with the country's agroecozones, thus the accuracy patterns may be partially attributed to some regions being harder to map than others.
-Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S9), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas.
+Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S10), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas.
 User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample.
-Aligning the reference samples more precisely with agroecozone boundaries (Figure S9B) provides further insight into error patterns (Table S4).
+Aligning the reference samples more precisely with agroecozone boundaries (Figure S10B) provides further insight into error patterns (Table S4).
 Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error.
 Maps in the two northern savanna agroecozones had the best balance between omission and commission error, and had the highest overall User's accuracy.
-The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S3), within which it typically covered a smaller share of area relative to the other agroecozones.
-This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S3B).
+The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S4), within which it typically covered a smaller share of area relative to the other agroecozones.
+This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S4B).
 
@@
 Smaller field sizes also complicate labelling, as these become increasingly indistinct in the ~4 m PlanetScope composites.
-The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S5), and by the average score achieved by each labeller against our training reference dataset  (71%; Figure S4B).
+The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S6), and by the average score achieved by each labeller against our training reference dataset (71%; Figure S5B).
 Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label errors, we found that they have substantial impact (Figure \ref{fig:trainingimpact}), which suggest that simply improving label quality may be one of the single most important investments towards improved model accuracy.
@@
 Labeller-specific accuracy measures also helped to improve the quality of the consensus labels, by placing higher weight on labels more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels.
-The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S4B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).
+The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S5B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).
 
@@
 Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness.
-Most notable were the larger improvements seen in the F1 score (Figure S7), a balanced performance metric.
+Most notable were the larger improvements seen in the F1 score (Figure S8), a balanced performance metric.
 Gains in accuracy and AUC were smaller.
@@
 The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
-Daily revisits are necessary for creating seasonal composites within a single  year over cloudy areas.
+Daily revisits are important for creating seasonal composites within a single  year over cloudy areas.
 The compositing technique we developed allowed us to develop a complete image catalog for the country representing the two seasons for 2018 agricultural year.
 Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
-@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana.
+@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval may be too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana.
 Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is challenging.
@@
 ## Lingering questions
-Several potential issues not answered by our assessment merit further exploration.
+Several potential issues not addressed in our assessment merit further exploration.
 One of these is the degree of correspondence between image- and ground-collected labels.
-However, such comparisons will likely reveal unresolvable differences between the two perspectives.
+However, such comparisons may reveal unresolvable differences between the two perspectives.
 The highly dynamic nature of many agricultural systems means that relatively narrow differences between the dates of ground- and image-based digitizing campaigns can lead to substantial disagreement between the resulting field boundaries, simply because the fields themselves may have shifted during the interval [@ElmesAccountingtrainingdata2020].
 These discrepancies could be further exacerbated by differences in the definition of what constitutes a field, which might vary on the ground depending on who is being asked, or who is doing the collecting.
-On top of this, ground-based collection efforts are usually highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is, or may be unrepresentative.
 These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth".
@@
 This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels.
-However, the visual assessment of their resulting probability maps confirm the differences in scores: consensus and most accurate individual labels produce nearly identical maps with relatively high certainty, while low quality labels led to a markedly less certain map (Figure S8).
+However, the visual assessment of their resulting probability maps confirm the differences in scores: consensus and most accurate individual labels produce nearly identical maps with relatively high certainty, while low quality labels led to a markedly less certain map (Figure S9).
 
@@
 # Acknowledgements
-The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund.
+The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PLACE.
 Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393).
@@
 
-<!-- ```{r Table1, results='asis', echo=FALSE, cache=FALSE} -->
-<!-- kable(head(iris), caption = "A glimpse of the famous *Iris* dataset.") -->
-<!-- ``` -->
-
-\newpage
-
-<!-- ```{r Table2, results='asis', echo=FALSE, cache=FALSE} -->
-<!-- kable(mtcars[10:16, ], caption = "Now a subset of mtcars dataset.") -->
-<!-- ``` -->
-
-\clearpage
-
-<!-- \listoffigures -->
-
-\newpage
-
-<!-- ```{r Fig1, echo=FALSE, fig.cap="Just my first figure with a very fantastic caption.", cache=FALSE} -->
-<!-- x <- rnorm(100) -->
-<!-- y <- jitter(x, 1000) -->
-<!-- plot(x, y) -->
-<!-- ``` -->
-
-\newpage
-
-\blandscape
-
-<!-- ```{r Fig2, echo=FALSE, fig.cap="Second figure in landscape format.", cache=FALSE} -->
-<!-- a <- sort(rnorm(100)) -->
-<!-- b <- c(rep("Group Small", 35), rep("Group Big", 65)) -->
-<!-- boxplot(a ~ b) -->
-<!-- ``` -->
-
-\elandscape
-
-\clearpage
-
 ```{r sessioninfo, echo = FALSE, eval = FALSE}

### 2021-03-09 inst/manuscript/si.Rmd
Edited ms and si: first upload to EarthArXiv. Edited scripts for data_for_instances.R, added new script to data-raw (aois_tiles.R) to show creation of degree and tiling grids

--- 2021-03-07 inst/manuscript/si.Rmd
@@
 ---
-title: "Supporting Information: High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"
+title: "High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"
+
+subtitle: "Supporting Information"
 
@@
     keep_tex: yes
-    number_sections: yes
-    template: manuscript.latex
+    number_sections: no
+    extra_dependencies: ["float", "titling", "setspace"]
     includes:
@@
 knitr::opts_chunk$set(fig.pos = "!ht", out.extra = "")
+# knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
 
@@
 The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites.
-The imagery was evaluated by examining their Raster Foundry overlays within a `labeller` instance set up for the purpose.
+The imagery was evaluated by examining their Raster Foundry [@azaveaRasterFoundry2020] overlays within a `labeller` instance set up for the purpose.
 
@@
 ### Mapping rules
-Labellers were asked to follow a particular set of rules when digitizing field boundaries.
-The following rules are reproduced (with some modification to wording) from a [website](https://github.com/agroimpacts/mappingrules) that we developed to host the rules.
+Labellers followed set of rules described in the next two sections when interpreting and digitizing field boundaries.
 
@@
 After collecting the sample, the geometries were further refined by converting the points (sites were the target didn't have to be shifted) to polygons with the same  0.0002866$^\circ$ resolution, and then the full set of map reference polygons was used to extract the classified values from both the cropland probability and vectorized field boundary maps.
+Accuracies were assessed for the entire country, and with several zones consisting of different groupings of AOIs or agroecozones\footnote{sourced from https://wheregeospatial.com/agro-ecological-zones-ghana/}
 
-\clearpage
 # Results
-## Image quality
+## Image catalog and quality
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
               names_prefix = "w") %>%
-  select(tile, season, w92, w99) %>% 
+  dplyr::select(tile, season, w92, w99) %>% 
   mutate(d = w92 - w99, 
          absd = abs(d))
+```
+
+The total number of PlanetScope images available for download from the Planet API [@planetteamPlanetApplicationProgram2018] per tile for each of the two major seasons is shown in Figure \ref{fig:imagedensity}.
+
+```{r imagedensity, echo = FALSE, out.width="90%", fig.cap="The density of images available through the Planet API for the five month growing season in 2018 and the subsequent three month dry season (excluding the month of November), shown in relation to the AOI boundaries.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_planet_coverage.png')
 ```
@@
 ### Labelling
-The distributions of the initial 500 sites used to train the RandomForest model for each AOI are shown in Figure \ref{fig:trainval}.
+The locations of training, training reference, and validation points are shown in Figure \ref{fig:trainval}.
 In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season.
 Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
-```{r trainval, echo = FALSE, out.width="80%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected within AOI 3 to retrain a second active learning run for this region, as well as the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.
-Grey background lines indicate the boundaries of Ghana's agroecozones, with the grey numbers indicating the agroecoregion ID: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE, results='hold'}
+```{r trainval, echo = FALSE, out.width="70%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.
+Grey background lines and show the boundaries and ID number of Ghana's agroecozones: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_training_validation_pts.png')
@@
 This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
-We replaced the image overlays in the instance for central-southeast \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
+We replaced the image overlays in the instance for central-southeast (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
 The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.
 
-```{r trainqual, echo = FALSE, out.width="60%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
-Labellers' identities are anonymized.", fig.align='center', message=FALSE, results='hold'}
+```{r trainqual, echo = FALSE, out.width="50%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
+Labellers' identities are anonymized.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_training_qual_n.png')
@@
 
-```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE, results='hold'}
+```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_label_risk_map.png')
@@
 ```{r labelriskhist, echo = FALSE, out.width="40%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
-The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE, results='hold'}
+The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_label_risk_hists.png')
@@
 ```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites.
-Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE, results='hold'}
+Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_random_vs_active.png')
@@
 ### The impact of training data error
-Probability images resulting from Random Forest models trained with labels generated by 1) consensus and the 2) most and 3) least accurate worker to digitize each training site are illustrated in Figure \ref{fig:labelstrategy}.
+Probability images resulting from Random Forests models trained with labels generated under three different labeling strategies are illustrated in Figure \ref{fig:labelstrategy}.
+These included consensus labels, and those individual labels that were likely to be the most and least accurate for each training site.
+Label accuracy was based on the mean score of each labeller against the training reference sites (Figure \ref{fig:trainqual}), as assessed when labelling a given AOI.
 These images were created for a single tile in AOI 1.
 
-```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by RandomForest models trained with A) consensus labels, B) the most accurate worker's labels, and C) the least accurate worker's labels.", fig.align='center', message=FALSE}
+```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by Random Forests models trained with A) consensus labels, B) the most accurate individual labels, and C) the least accurate individual labels.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_label_strategy_probs.png')
@@
   # column_spec(1, width = "8cm")
-
-# tp <- 10
-# fp <- 2
-# tn <- 20
-# fn <- 3
-# tss <- (tp * tn - fp * fn) / ((tp + fn) * (fp + tn))
-# mcc <- (tp * tn - fp * fn) / sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
-
 ```
@@
 
-```{r aoizones, echo = FALSE, out.width="100%", fig.cap="Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy.
+```{r aoizones, echo = FALSE, out.width="90%", fig.cap="Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy.
 In B, the ID number of Ghana's individual agroecozones is shown in grey: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE, results='hold'}
@@
 The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.}
-\includegraphics[width = 18cm]{figures/si_aez_accuracies.png}
+\begin{center}
+\includegraphics[width = 16cm]{figures/si_aez_accuracies.png}
+\end{center}
 \label{tab:aezaccuracy}
@@
 The same statistics for average number of segments versus average number of labelled polygons across validation sites in each AOI are shown in Figure \ref{fig:numbervalidation}.
-```{r areavalidation, echo = FALSE, out.width="100%", fig.cap="The distributions of the average areas (in hectares) of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average areas of field boundaries digitized (shown in red) by the most accurate worker to label each site.
-Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
+```{r areavalidation, echo = FALSE, out.width="90%", fig.cap="The distributions of the average areas (in hectares) of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average areas of field boundaries digitized (shown in red) by the most accurate worker to label each site.
+Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_validation_stats_fha2.png')
@@
 The mean and median of each distribution is shown.
-Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE, results='hold'}
+Vertical lines indicate the mean and median of each distribution.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_validation_stats_fn2.png')
@@
 
-```{r fieldsizes, echo = FALSE, out.width="100%", fig.cap="The A) average sizes of fields and B) total number of fields in each 0.05 degree tile, as calculated from the vectorized field boundaries.", fig.align='center', message=FALSE, results='hold'}
+```{r fieldsizes, echo = FALSE, out.width="90%", fig.cap="The A) average sizes of fields and B) total number of fields in each 0.05 degree tile, as calculated from the vectorized field boundaries.", fig.align='center', message=FALSE}
 knitr::include_graphics('figures/si_fsize_n_05.png')

### 2021-03-09 data-raw/data_for_instances.R

@@ -1,4 +1,5 @@
 # Data set up
+## Follows on from aoi_tiles.R
 
 #------------------------------------------------------------------------------#
 # Prepare main grid for Ghana
@@ -9,16 +10,19 @@ library(sf)
 library(dplyr)
 
 # get data from labeller (get from S3)
-data_path <- file.path("/Users", unname(Sys.info()["user"]),
-                       "Dropbox/projects/activelearning/mapper/spatial/data")
-master_grid <- brick(file.path(data_path, "processed/master_grid.tif"))
-v <- values(master_grid[[1]])
+mgrid_path <- file.path("/vsis3/activemapper/grid/master_grid.tif")
+mgrid <- raster(mgrid_path)
+# data_path <- file.path("/Users", unname(Sys.info()["user"]),
+#                        "Dropbox/projects/activelearning/mapper/spatial/data")
+# mgrid <- brick(file.path(data_path, "processed/master_grid.tif"))
+# v <- values(mgrid[[1]])
 # length(v[!is.na(v)])
 # length(unique(v[!is.na(v)]))
 
 # crop master_grid to ghana_grid
-tiles <- st_read("inst/extdata/ghana_tiles.geojson")
-ghana_grid <- crop(master_grid[[1]], extent(tiles))
+tiles <- st_read(here::here("external/data/ghana_tiles.geojson"))
+# tiles <- tiles %>% rename(aoi = aoi1)
+ghana_grid <- crop(mgrid[[1]], extent(tiles))
 # length(ghana_grid[(!is.na(ghana_grid))])
 # length(ghana_grid[!is.na(ghana_grid)])
 # length(unique(ghana_grid[!is.na(ghana_grid)]))
@@ -26,16 +30,14 @@ ghana_grid <- crop(master_grid[[1]], extent(tiles))
 
 writeRaster(ghana_grid, file = "inst/extdata/ghana_grid.tif",
             overwrite = TRUE, datatype = "INT4U")  # careful for datatype
-r <- raster("inst/extdata/ghana_grid.tif")
+# r <- raster("inst/extdata/ghana_grid.tif")
 # length(r[!is.na(r)])
 # length(unique(r[!is.na(r)]))
 
 #------------------------------------------------------------------------------#
 # Create AOIs, tiles, degrees
 # slim down Africa data
[…]

### 2021-03-10 data-raw/consensus_risk.R
Committing data files and scripts

@@ -1,4 +1,6 @@
-# Calculate Bayesian risk metrics for labels
+# Calculate Bayesian Risk for each labels
+# Reprocessing to include measure on mean risk as well as proportion risky
+
 library(rmapaccuracy)
 library(dplyr)
 library(data.table)

### 2021-03-10 data-raw/iteration_metrics.R

@@ -1,0 +1,45 @@
+# Iteration performance metrics -
+# Analyze and reshape for figure making, including calculating percentage
+# change per iteration
+
+library(dplyr)
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+# function to calculate change over time
+delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)
+
+# process individual AOIs
+metrics <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "TPR",
+             "FPR")
+aoinms <- paste0("labeller", c(1:16, "1r", "8r", "15r", "8v2", "82019"))
+iteration_metrics <- instance_tbls$iteration_metrics %>%
+  filter(aoi %in% aoinms) %>%
+  filter(iteration > 0) %>%
+  filter(!(aoi == "labeller3" & run == 0)) %>% # remove run 0 labeller3
+  filter(!(aoi %in% paste0("labeller", c("1r", "8r")) & run == 1)) %>% # no run1
+  mutate(iteration = iteration - 1) %>% # rebase iteration counter on 0
+  mutate(F1 = 2 * (precision * recall) / (precision + recall)) %>%
+  rename(AOI = aoi, Iteration = iteration, Accuracy = accuracy, AUC = auc,
+         TSS = tss, Precision = precision, Recall = recall, TPR = tpr,
+         FPR = fpr) %>%
+  select(AOI, Iteration, !!metrics) %>%
+  group_by(AOI) %>%  # calculate deltas
+  mutate_at(metrics, list("change" = delta_func)) %>%
+  ungroup() %>%
+  tidyr::pivot_longer(cols = Accuracy:FPR_change, names_to = "Metric",
+                      values_to = "Score")
+
+# process across AOI means and combine with AOIs
+aois <- paste0("labeller", 1:16)  # calculate summary across just AOIs 1-16
+iteration_metrics %>%
+  filter(AOI %in% aois) %>%
+  filter(Iteration < 4) %>%
+  group_by(Iteration, Metric) %>%
+  summarize(Score = mean(Score)) %>%
+  mutate(AOI = "All") %>%
+  ungroup %>%
+  select(AOI, !!names(.)) %>%
+  bind_rows(iteration_metrics, .) -> iteration_metrics
+
+usethis::use_data(iteration_metrics, overwrite = TRUE)
+

### 2021-03-11 inst/manuscript/manuscript.Rmd
Updated acknowledgements in MS, and added TOC and data availability statement to SI.

@@
 ## Next steps
-The maps presented here represent a version 1 product that is freely available to use, along with its underlying code.
+The maps presented here represent a version 1 product that is freely available to use, along with its underlying code (see SI for details).
 These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020].
@@
 The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PLACE.
-Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393).
+Additional support was provided by NASA (80NSSC18K0158), the National Science Foundation (SES-1801251; SES-1832393), and Princeton University.
 Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative.
 Azavea provided significant contributions in engineering the machine learning pipeline.
-We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
+We thank Meridia for providing information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
 We thank Manushi Trivedi, Sitian Xiong, and Tammy Woodard for their contributions to the underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and Primoz Kovacic for support and management of the labelling efforts.

### 2021-03-11 inst/manuscript/si.Rmd
Updated acknowledgements in MS, and added TOC and data availability statement to SI.

@@
     keep_tex: yes
-    number_sections: no
+    number_sections: yes
+    toc_depth: 4
+    toc: yes
     extra_dependencies: ["float", "titling", "setspace"]
@@
 
-# Methods
+# Supplemental Methods
 ## Assessing quality of composites
@@
 
-# Results
+# Supplemental Results
 ## Image catalog and quality
@@
 
+# Code and Data Availability
+The repositories containing the data, code, and manuscript source files used to produce this paper are listed below:
+
+- [`activemapper`](https://github.com/agroimpacts/activemapper): Contains the source code and derived data for this paper.
+The repository\footnote{github.com/agroimpacts/activemapper} will be opened following publication in a peer-reviewed journal.
+- [`croplandmaps`](https://github.com/agroimpacts/croplandmaps_V2): Source code for web map displaying image composites and cropland probability maps.
+The repository\footnote{github.com/agroimpacts/croplandmaps\_V2} is open, and the field boundary datasets are available for download.
+- [`imager`](https:github.com/agroimpacts/imager): Source code for the image processing software.
+The repository\footnote{github.com/agroimpacts/imager} is currently private, but will be opened after completing a security audit.
+- [`labeller`](https:github.com/agroimpacts/labeller): Source code for the labelling platform.
+The repository\footnote{github.com/agroimpacts/labeller} is currently private, but will be opened after completing a security audit.
+- [`learner`](https:github.com/agroimpacts/learner): Source code for the active learning component. The repository\footnote{github.com/agroimpacts/learner} is open.
+- [`segmenter`](https:github.com/agroimpacts/segmenter): Source code for the segmentation algorithm.
+The repository \footnote{github.com/agroimpacts/segmenter} is currently private, but will be opened after completing a security audit.
+
 \clearpage

### 2021-03-12 inst/manuscript/manuscript.tex
Updating raw scripts and package vignette. Added tex files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Original default rstudio/pandoc latex file
%% upated by @jhollist 09/15/2014
%% inspired by @cboetting https://github.com/cboettig/template and
%% @rmflight blog posts:
%% http://rmflight.github.io/posts/2014/07/analyses_as_packages.html
%% http://rmflight.github.io/posts/2014/07/vignetteAnalysis.html).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{float}  % added 9/3/2021 LDE
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes borrowed from @cboettig, added by @jhollist
% A modified page layout
\textwidth 6.75in
\oddsidemargin -0.15in
\evensidemargin -0.15in
\textheight 9in
\topmargin -0.5in
\usepackage{lineno} % add
  \linenumbers % turns line numbering on
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Packages and layout changes by @jhollist 09/15/2014
\usepackage{ragged2e}
\usepackage[font=normalsize]{caption}
  \usepackage[doublespacing]{setspace}
\usepackage{parskip}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
  \rfoot{\today}
\lfoot{\thepage}
%%Changed default abstract width and added lines
\renewenvironment{abstract}{
  \hfill\begin{minipage}{1\textwidth}
  \rule{\textwidth}{1pt}\vspace{5pt}
  \normalsize
  \begin{justify}
  \bfseries\abstractname\vspace{5pt}
  \end{justify}}
  {\par\noindent\rule{\textwidth}{1pt}\end{minipage}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Added this on 28/2/2021 to deal with CSLReferences issue:
%% >> LaTeX Error: Environment CSLReferences undefined.
%% Occurs after pandoc upgraded, per https://github.com/mpark/wg21/issues/54

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales}
\author{
Lyndon D. Estes
Su Ye
Lei Song
Boka Luo
J. Ronald Eastman
Zhenhua Meng
Qi Zhang
Dennis McRitchie
Stephanie R. Debats
Justus Muhando
Angeline H. Amukoa
Brian W. Kaloo
Jackson Makuru
Ben K. Mbatia
Isaac M. Muasa
Julius Mucha
Adelide M. Mugami
Judith M. Mugami
Francis W. Muinde
Fredrick M. Mwawaza
Jeff Ochieng
Charles J. Oduol
Purent Oduor
Thuo Wanjiku
Joseph G. Wanyoike
Ryan B. Avery
Kelly K. Caylor
}
\date{}
% Allowing for landscape pages
\usepackage{lscape}
\usepackage{graphicx}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

% Left justification of the text: see https://www.sharelatex.com/learn/Text_alignment
% \usepackage[document]{ragged2e} % already in the latex template
\newcommand{\bleft}{\begin{flushleft}}
\newcommand{\eleft}{\end{flushleft}}
\usepackage{flafter}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

%%Fix tightlist error: https://stackoverflow.com/questions/40438037/tightlist-error-using-pandoc-with-markdown
%%Added 2018-03-26
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
%%%

\begin{document}
%%Edited by @jhollist 09/15/2014
%%Adds title from YAML
\begin{singlespace}
\begin{center}
\huge High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales
\end{center}
%%Adds subtitle from YAML
%%Adds Author, correspond email asterisk, and affilnum from YAML
\begin{center}
\large
%% removed space here
Lyndon D. Estes\textsuperscript{*} \textsuperscript{1}, 
%% removed space here
Su Ye \textsuperscript{1,2}, 
%% removed space here
Lei Song \textsuperscript{1}, 
%% removed space here
Boka Luo \textsuperscript{1,3}, 
%% removed space here
J. Ronald Eastman \textsuperscript{1,3}, 
%% removed space here
Zhenhua Meng \textsuperscript{1}, 
%% removed space here
Qi Zhang \textsuperscript{1}, 
%% removed space here
Dennis McRitchie \textsuperscript{4}, 
%% removed space here
Stephanie R. Debats \textsuperscript{4}, 
%% removed space here
Justus Muhando \textsuperscript{5}, 
%% removed space here
Angeline H. Amukoa \textsuperscript{5}, 
%% removed space here
Brian W. Kaloo \textsuperscript{5}, 
%% removed space here
Jackson Makuru \textsuperscript{5}, 
%% removed space here
Ben K. Mbatia \textsuperscript{5}, 
%% removed space here
Isaac M. Muasa \textsuperscript{5}, 
%% removed space here
Julius Mucha \textsuperscript{5}, 
%% removed space here
Adelide M. Mugami \textsuperscript{5}, 
%% removed space here
Judith M. Mugami \textsuperscript{5}, 
%% removed space here
Francis W. Muinde \textsuperscript{5}, 
%% removed space here
Fredrick M. Mwawaza \textsuperscript{5}, 
%% removed space here
Jeff Ochieng \textsuperscript{5}, 
%% removed space here
Charles J. Oduol \textsuperscript{5}, 
%% removed space here
Purent Oduor \textsuperscript{5}, 
%% removed space here
Thuo Wanjiku \textsuperscript{5}, 
%% removed space here
Joseph G. Wanyoike \textsuperscript{5}, 
%% removed space here
Ryan B. Avery \textsuperscript{6}, 
%% removed space here
Kelly K. Caylor \textsuperscript{6,7,8}, 
\end{center}
%%Adds affiliations from YAML
\begin{justify}
% \footnotesize \emph{
% \\*
\footnotesize\textsuperscript{1}Graduate School of Geography, Clark
University, Worcester, MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{2}Department of Natural Resources and the
Environment, University of Connecticut, Storrs, CT, USA\\\\*
% \\*
\footnotesize\textsuperscript{3}Clark Labs, Clark University, Worcester,
MA, USA\\\\*
% \\*
\footnotesize\textsuperscript{4}Independent contributor\\\\*
% \\*
\footnotesize\textsuperscript{5}SpatialCollective, Nairobi, Kenya\\\\*
% \\*
\footnotesize\textsuperscript{6}Department of Geography, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{7}Earth Research Institute, University of
California Santa Barbara, Santa Barbara, CA, USA\\\\*
% \\*
\footnotesize\textsuperscript{8}Bren School of Environmental Science and
Management, University of California Santa Barbara, Santa Barbara, CA,
USA\\\\*
% }
%%Adds corresponding author email(s) from YAML
\newcounter{num}
\setcounter{num}{1}
\\[0.1cm]
\footnotesize \emph{
\ifnum\value{num}=1%
\textsuperscript{*} corresponding author:
\fi
\href{mailto:lestes@clarku.edu}{\nolinkurl{lestes@clarku.edu}}
\stepcounter{num}
}

%\begin{center}
This paper is a non-peer reviewed preprint submitted to EarthArXiv. It
will be periodically updated as it is revised, including information on
its peer review status.
%\end{center}

\end{justify}
%%Adds date from YAML
\normalsize

\begin{abstract}
Understanding agricultural change requires reliable, frequently updated
maps that describe the characteristics of croplands. Such data are often
unavailable for regions dominated by smallholder agricultural systems,
which are particularly challenging for remote sensing. To overcome these
challenges, we designed a system to minimize several sources of error
that arise when mapping smallholder croplands. To overcome errors caused
by mismatches between image resolution and cropland scales, as well as
persistent cloud cover, the system converts daily, 3.7 m PlanetScope
imagery into two seasonal composites within a single agricultural year.
To reduce errors that occur when training classifiers, we built a
labelling platform that rigorously assesses label accuracy, and creates
more accurate consensus labels that train a Random Forests model. The
labelling platform and model interact within an active learning process
that boosts the accuracy of the resulting cropland probability map,
which is used in a segmentation process to delineate individual field
boundaries. We applied this system to map Ghana's croplands for the year
2018. We divided Ghana into 16 mapping regions (12,160-23,535 km\(^2\)),
training separate models for each using a total of 6,299 labels, plus
1,600 for validation. Using an independent map reference sample
(n=1,207), we found that overall accuracies of the resulting cropland
probability and field boundary maps were 88\% and 86.7\%, respectively,
with User's accuracies for the cropland class of 61.2\% and 78.9\%, and
Producer's accuracies of 67.3\% and 58.2\%. Croplands covered
16.1-23.2\% of the mapped area, comprising 1,131,146 total fields with
an average size of 3.92 ha. Estimates based on the map reference sample
indicate the cropland percentage is 17.1\% (15.4-18.9\%) or 17.6\%
(15.6-19.6\%), depending on the map used to estimate the standard error.
Using the labellers' digitized field boundaries to estimate biases in
field boundary statistics, we calculated an adjusted mean field size of
1.73 ha and total field count of 1,662,281. Although the cropland class
contained substantial errors, the system was effective in mitigating
error and quantifying resulting performance gains. By minimizing
training errors, consensus labelling improved the model's F1 scores by
up to 25\%, while 3 iterations of active learning increased the F1 score
by 9.1\%, on average, which was 2.3\% higher than training models with
randomly selected labels. Map accuracy can be improved by replacing
Random Forests with a convolutional neural network. These results
demonstrate a readily adapted, transferrable framework for developing
high resolution, annual, nation-scale maps that provide important
details about smallholder-dominated croplands.
\end{abstract}
\end{singlespace}

\bleft

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Amidst all the challenges posed by global change, a particular concern
is how agricultural systems will adapt to meet humanity's growing food
demands, and the impacts that transforming and expanding food systems
will have on societies, economies, and the environment (Searchinger et
al. 2019). Significant efforts are being made to address the various
aspects of this challenge, including work on diagnosing and closing
yield gaps (Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al.
2012), expanding and commercializing production (Morris and Byerlee
2009), and to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017,
Davis et al. 2020) and mitigate (Estes et al. 2016b) agriculture's
ecological impacts. Answering many of the questions these efforts seek
to address depends on reliable data that describes the location and
characteristics of cropland (Fritz et al. 2015), and how these are
changing over time. Unfortunately, the data that do exist are in many
places inaccurate. Existing estimates of how much global cropland there
is tend to vary widely, and they often disagree about where cropland is
located (e.g. Fritz et al. 2011, 2013). Such errors can propagate in
subsequent analyses that rely on cropland data as inputs, resulting in
potentially misleading answers (Estes et al. 2018). Beyond cropland
distributions, few data are available on key cropland characteristics
such as field size, an important variable needed to estimate yield and
other key food security variables (Carletto et al. 2015), and as an
indicator of farm size (Levin 2006, Samberg et al. 2016), a critical
component of rural livelihoods given increasing population densities and
longstanding debates about the relationship between farm size and
productivity (Feder 1985, Carletto et al. 2013, Desiere and Jolliffe
2018).

These informational inadequacies are due to the fact that cropland data
in much of the world are derived from remotely sensed landcover maps,
which can be notoriously high in error, particularly over regions such
as Africa (Fritz et al. 2010, Estes et al. 2018), where agricultural
changes will be largest and the need for accurate baseline data is thus
greatest (Searchinger et al. 2015, Estes et al. 2016b, Bullock et al.
2021). Cropland mapping over Africa is difficult for several reasons.
The primary reason relates to the characteristics of the continent's
smallholder-dominated croplands, where half of all fields are smaller
than 1 ha (Lesiv et al. 2019). This size is small relative to the 30-250
m resolution of the sensors typically used in many landcover mapping
efforts (e.g. Chen et al. 2015, Sulla-Menashe et al. 2019), which
results in errors due to mixed pixels and aspects of the modifiable area
unit problem (Openshaw and Taylor 1979). In the latter case, the pixel's
shape may be poorly matched to that of cropland, and is too coarse to
aggregate to approximate that shape at the characteristic scales of crop
fields (Dark and Bram 2007, Estes et al. 2018). On top of the matter of
scale is 1) high intra-class variability of the cropland class,
compounded by the fact that these particular croplands can be heavily
intergraded with surrounding vegetation (Debats et al. 2016, Estes et
al. 2016a), and 2) the substantial temporal variability within
croplands, both within and between seasons. These latter two aspects
pose challenges for the classification algorithms that are applied to
the imagery.

These problems arising from cropland characteristics are increasingly
being addressed due to technological advances. Recent advances in
satellite technology have increased the coverage of high (\textless5 m)
or near-high (10 m) resolution imagery with weekly to near-daily return
intervals (Drusch et al. 2012, McCabe et al. 2017). This high spatial
\emph{and} temporal resolution addresses the sensor-field scale
mismatch, and more effectively captures the intra-seasonal dynamics of
cropland, which helps classifiers distinguish cropland from surrounding
cover types (Debats et al. 2016, Defourny et al. 2019). On top of this,
advances in cloud computing (Gorelick et al. 2017), the opening of image
archives (Wulder et al. 2016), and next generation machine learning
approaches (Zhu et al. 2017, Maxwell et al. 2018) are placing large
volumes of these moderate to near-high resolution imagery together with
the computational and algorithmic resources necessary to classify them
at scale. These capabilities are aleady being used to create a new
generation of higher resolution (10-30 m) cropland and landcover maps
for Africa and other regions (Xiong et al. 2017, Lesiv et al. 2017, ESA
n.d.).

Despite these advances, the highest resolution (\textless5 m) image
sources are still not used to map cropland over very large extents,
presumably because they are commercial and relatively high cost to
acquire, in addition to the greater computational demands. As such, map
accuracy can still be a challenge, particular for User's accuracy, which
ranges between 46 and 76\% for the cropland class (Xiong et al. 2017,
e.g. Lesiv et al. 2017).

Accuracy may also suffer due to error-inducing factors that are becoming
somewhat more pronounced as a consequence of these technology advances,
particularly with respect to algorithms. Advances in machine learning
are helping to greatly improve classification skill, but these
algorithms generally require large, high-quality training datasets
(Maxwell et al. 2018, Ma et al. 2019, Elmes et al. 2020). To satisfy
this need for more training (and reference) samples, map-makers
increasingly rely on visual interpretation of high resolution satellite
or aerial imagery to collect training (or validation) samples (Chen et
al. 2015, e.g. Xiong et al. 2017, Stehman and Foody 2019). Several
web-based platforms have been developed to facilitate such efforts,
which provide convenient and highly scalable tools for training data
collection (Fritz et al. 2012, Estes et al. 2016a, e.g. Bey et al.
2016). Visually interpreted training labels present two particular
problems. The first is that such labels have inevitable interpretation
errors that can vary substantially according to the skill of the
interpreter (Estes et al. 2016a, Waldner et al. 2019). These errors are
typically not accounted for in reported accuracy metrics, despite the
fact that they can introduce substantial error into the resulting maps
and subsequent analyses (Estes et al. 2018, Elmes et al. 2020). The
second problem is that visual interpretation depends on high resolution
imagery (\textless5 m), as lower resolutions make it difficult for a
human analyst to discern cropland. Typically the only practical source
for such imagery are ``virtual globe'' basemaps provided by Microsoft
and Google, which are composed of mosaics of various high resolution
satellite and aerial images that typically span 3-5 years of time within
a single country (Lesiv et al. 2018). This within-mosaic temporal
variation can set up a temporal mismatch between the imagery being
interpreted and the imagery being classified, which is usually from a
different source (e.g.~Landsat, Sentinel; Xiong et al. (2017)). If a
land change occurs in the interval between the two image sets (e.g.~a
new field was created), this can introduce error into the training data
that is then passed on to the classifier. This source of error may be
elevated in smallholder-dominated systems in the tropics, where swidden
practices are common (Van Vliet et al. 2013), or in rapidly developing
agricultural frontiers (Zeng et al. 2018).

Improving the accuracy of cropland maps over smallholder-dominated
systems thus requires an approach that meets three requirements. First,
it should be based on high spatial and temporal resolution imagery, to
be able to capture the fine grain and temporal variability of
smallholders' fields. Second, an algorithm with suitable skill for
classifying these images must be selected, and combined with the
computational resources needed to process large imagery volumes. Third,
a method for collecting large volumes of high quality training and
validation data based on image interpretation is essential. This method
should quantify and minimize the errors associated with image
interpretation. It should also ensure that labels are collected either
from the same imagery that is being classified, or from contemporanous
imagery, in order to reduce errors introduced by land change processes.

We describe here a cropland mapping system that follows these
requirements, with an emphasis on delineating field boundaries. The
first requirement is enabled by the recent availability of CubeSat data
that provides 3-4 m resolution imagery over large areas at near daily
intervals (McCabe et al. 2017). Although these data, currently collected
by Planet's CubeSat fleet, are of lower spectral depth and quality than
Landsat, Sentinel, or Worldview imagery, they enable country to
continental scale image mosaics to be created for multiple periods
during the crop growing calendars, and capturing the intra-annual
variability can be more important for classifying cropland than spectral
depth (Debats et al. 2016). This daily revisit capacity is also
important for developing seasonal composites in cloudy regions, where
satellites with longer return intervals may fail. Lastly, although this
imagery being up to \textasciitilde16 times coarser than much of Bing or
Google imagery, it is sufficiently resolved for humans to discern most
fields (Fourie 2009, e.g.~see Estes et al. 2018). This allows labels to
be generated on the same imagery processed by the classifier, thereby
addressing one of the two needs related to training data (requirement
3).

The second requirement is addressed by a computer vision/machine
learning classifer that is effective for classifying smallholder
croplands (Debats et al. 2016), re-engineered to run on high
performance, cloud-based computing clusters with a simplified feature
set, and following recommended practices for controlling for and
measuring the errors that occur when training machine learning models
for remote sensing applications (Elmes et al. 2020). The classifier is
tightly coupled to a front-end platform for collecting label data, which
includes rigorous accuracy assessment protocols and a novel approach for
merging multiple maps into consensus labels, thereby minimizing image
interpretation error (Estes et al. 2016a, Elmes et al. 2020). The
training and machine learning components are combined within an active
learning framework, wherein the machine learning process assesses
classification uncertainty in unlabelled areas after a training step,
and selects sites from areas of highest uncertainty for additional
labelling (Cohn et al. 1994, Tuia et al. 2011). Our framework automates
this interactive approach to label selection, which is effective in
boosting the performance of classification models while reducing the
overall number of training samples required to achieve a given level of
performance (Debats et al. 2017, e.g. Hamrouni et al. 2021). Finally, an
unsupervised segmentation step is applied to the imagery and merged with
the pixel-wise classifications from the machine learning process,
resulting in a vectorized field boundary map that provides important
information on field geometry.

We demonstrate this approach to map cropland in Ghana, a country where
smallholder farming predominates, and which has a broad mix of climates
and agricultural systems, including large areas where shifting
agriculture is practiced (Samberg et al. 2016, Kansanga et al. 2019).

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{system-overview}{%
\subsection{System overview}\label{system-overview}}

The mapping system design has four primary components focused on 1)
image acquisition and compositing, 2) training and validation data
collection, 3) image classification, and 4) segmentation. The first
component is applied as a once-off step to generate an image catalog
covering the mapping geography. The second component provides tools for
labelling imagery, and controls an interactive model training and
classification (component 3) pipeline that produces a map of cropland
probabilities for each image pixel (Figure \ref{fig:systemoverview}).
The final segmentation step (component 4) is then initiated and applied
to both the input image catalog and the posterior probability maps,
resulting in vectorized field boundaries. Each system component
comprises an individual software module designed to be implemented on
cloud computing architecture, and available on a GitHub repository (see
data and software availability section for details).

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{figures/figure1}

}

\caption{An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.}\label{fig:systemoverview}
\end{figure}

\hypertarget{system-components}{%
\subsection{System components}\label{system-components}}

\hypertarget{image-compositing}{%
\subsubsection{Image compositing}\label{image-compositing}}

The image processing components of our system were designed to work with
PlanetScope Analytic surface reflectance data (PlanetTeam 2018).
PlanetScope provides three visual (red, green, blue) and a near-infrared
band at 3-4 m resolution at nominal daily frequency. Although these
images are already pre-preprocessed and corrected for atmospheric
effects, there are residual errors from inter-sensor differences and the
radiometric normalization process (Houborg and McCabe 2018), variation
in the orientation of scene footprints, as well as a high frequency of
cloud cover over the study region (Wilson and Jetz 2016). To correct for
these factors, we developed a procedure for creating temporal composites
representing the primary growing and non-growing seasons within a single
year.

PlanetScope imagery is accessed via the Planet API (PlanetTeam 2018),
and an initial order is placed for all imagery falling within the
mapping geography and the date ranges for the two compositing periods.
The imagery is collected and transferred directly to a cloud storage
platform (Amazon Web Services {[}AWS{]} S3).

Individual images are then transformed into analysis ready data (ARD)
(Dwyer et al. 2018), by subsetting each downloaded image into 0.05
degree tiles, regardless of cloud cover. Tiles are organized within a
larger 1 degree resolution grid that covers the entire continent, which
defines the minimum mapping area of interest (AOI; Figure
\ref{fig:aois}).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/figure2}

}

\caption{The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.}\label{fig:aois}
\end{figure}

The temporal compositing process is applied to the tiled daily images
for the time period of interest, which in this case was one of two
multi-month seasons, the primary growing and dry seasons for a single
agricultural year. Imagery from two seasons helps to improve the
performance of cropland classifiers (Debats et al. 2016), while having
the seasons in the same year helps to minimize differences caused by
land change. For each pixel in each image in each ARD temporal stack for
a given season, two weights are calculated:

\begin{equation} \label{eq:cloud}
\mathrm{W1_t} = \frac{1}{\mathrm{blue_t}^2}
\end{equation}

\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}

Where \emph{t} is a particular date in near-daily time series of
PlanetScope images, which begins at date 1 for the given compositing
period and ends on date \emph{i}, \emph{blue} is the blue band, and
\emph{NIR} the near infrared band. Equation \ref{eq:cloud} assigns lower
weights to hazy and clouded pixels as the blue band is sensitive to haze
and cloud pixels (Zhang et al. 2002), while Equation \ref{eq:shadow}
assigns low weights to pixels in cloud shadow considering the
significant darkening effect of the cloud shadows in the Near Infrared
band (Zhu and Woodcock 2012, Qiu et al. 2020)

Once these two weights are calculated, the final composited pixel value
for each of the four PlanetScope bands is:

\begin{equation}
\mathrm{\bar{B} = \frac{\sum_{t=1}^{T}B_t * W1_t * W2_t}{\sum_{t=1}^{T}W1_t * W2_t}}
\end{equation}

Which is a weighted mean for each pixel for each band \emph{B} for the
particular compositing period. The composited tiles were then added to
the S3 store (Figure \ref{fig:systemoverview}), where they are stored as
cloud-optimized geotiffs, and a ``slippy
map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}''
rendering is created for each composite using Raster Foundry (Azavea
2020). The web-rendered imagery is presented within the training data
platform (next section).

\hypertarget{labelling-platform}{%
\subsubsection{Labelling platform}\label{labelling-platform}}

Training and reference data are collected by a custom labelling
platform, which was originally designed for AWS's Mechanical Turk job
marketplace (Estes et al. 2016a). The basic structure of the system
remains the same, but we converted it into a standalone platform that
allows us to enroll and pay people directly for their labelling, and is
designed to control and supervise the machine learning process. The
platform runs on a Linux virtual machine hosted on an AWS EC2 instance
and is comprised of a database (PostGIS/Postgres), a mapping interface
(OpenLayers 3), an image server (Raster Foundry), and a set of utilities
for managing, assessing, and converting digitization work into
rasterized labels for training a machine learning algorithm. Each
instance of the platform focuses on a specific AOI (Figure
\ref{fig:aois}A)

The following sections provide an overview of the labelling platform's
architecture.

\hypertarget{mapping-workflow}{%
\paragraph{Mapping workflow}\label{mapping-workflow}}

\hypertarget{selecting-training-and-reference-sites}{%
\subparagraph{Selecting training and reference
sites}\label{selecting-training-and-reference-sites}}

The labelling process begins with the random selection of a subset
(e.g.~500) of cells from a 0.005 degree grid, with the selection itself
potentially split into a training and validation sample, according to
predetermined proportions. The grid, which is nested within the tiling
and larger 1 degree grids (Figure \ref{fig:aois}C) defines the spatial
unit for a labelling job. The selected cells are placed into a queue
within the platform's database, and then converted into a mapping
\emph{task} that has a specified number of \emph{assignments}
(boundaries drawn by an individual labeller) that must be completed
before the task is complete.

\hypertarget{mapping-assignments}{%
\subparagraph{Mapping assignments}\label{mapping-assignments}}

Labellers registered in the system log in to the mapping platform (built
with Flask) and navigate to the OpenLayers-based field mapping interface
(Figure \ref{fig:labeller}), where they are presented with a white
target box representing the randomly selected grid cell, a set of
digitizing tools, and different image backdrops, including true and
false color renderings of the growing season and off-growing season
PlanetScope composites, and several virtual globe basemaps. Following a
set of pre-defined digitizing rules (see SI), the labeller uses the
polygon drawing tool to digitize the boundaries of all crop fields
intersecting the target grid cell that are visible within the
PlanetScope overlays. To aid with interpretation, the labeller can
toggle between the PlanetScope renderings and the basemaps to help form
a judgement about what constitutes a field. The labeller assigns each
polygon a class category (e.g.~annual cropland), and upon completing all
fields submits the assignment to the database. In cases where the target
grid cell does not contain any fields, the labeller simply submits the
assignment to mark it complete. The labeller is then directed to the
next available assignment from a different labelling task.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{figures/figure3}

}

\caption{An overview of the labelling platform's interface}\label{fig:labeller}
\end{figure}

\hypertarget{processing-completed-assignments}{%
\subparagraph{Processing completed
assignments}\label{processing-completed-assignments}}

All submitted polygons are cleaned to fix topological irregularities
that arose during digitization (see supporting information {[}SI{]}) and
stored in a PostGIS table. Each completed assignment represents one of
two types of tasks: 1) accuracy assessment, or 2) model training or
validation. For the former type, an accuracy assessment routine is
invoked that executes a series of comparisons between the labeller's
results and a training reference dataset, resulting in a assignment
score:

\begin{equation} \label{eq:qaqc}
\mathrm{score_i}=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}
\end{equation}

Where \emph{i} indicates the particular assignment, and \(\beta_{0-4}\)
represent varying weights that sum to 1. \emph{I} refers to ``inside the
box'' accuracy, \emph{O} is the accuracy of those portions of the
labeller's polygons extending beyond the target grid boundaries,
\emph{F} is fragmentation accuracy, a measure of how many individual
polygons the labeller delineated relative to the reference, \emph{E}
measures how closely each polygon's boundary matched its corresponding
reference polygon boundary, and \emph{C} assesses the accuracy of the
labeller's thematic labels (see SI for individual formulae). Equation
\ref{eq:qaqc} is an extension of the approach described by Estes et
al.~(2016).

Over time, labellers are assessed multiple times across a range of
accuracy tasks, which are selected to represent the variability of the
agricultural system being mapped. Each labeller's score history is
averaged to provide an overall accuracy measure, and this information is
used for creating labels, the second task.

If the labeller's completed assignment was a training/validation task,
their maps remain stored in the database until the task's outstanding
assignments are completed by other labellers. Once complete, another
routine is invoked, which combines the task's completed assignments into
a single consensus label using a Bayesian merging approach:

\begin{equation}
P(\theta|\mathrm{D})=\sum_{i=1}^{n}\mathrm{P}(\mathrm{W_i}|\mathrm{D})\mathrm{P}(\mathrm{\theta}|\mathrm{D}, \mathrm{W_i})
\end{equation}

Where \(\theta\) represents the true cover type of a pixel (field or not
field), \emph{D} is the label assigned to that pixel by a labeller, and
\(W_i\) is an individual labeller. P(\(\theta\)\textbar D) is therefore
the probability that the actual cover type is what the labellers who
mapped it says it is, while P(W\(_i\)\textbar D) is an individual
labeller's average score over all the accuracy assessment assignments
they have completed, and P(W\(\theta\)\textbar D, \emph{W\(_i\)}) is the
labeller's label for that pixel. This approach therefore uses the
overall accuracy of each labeller to weight their labels when combined
with those made by other labellers' for the same pixel (see SI for
further details). As a further measure of confidence in the final
consensus label, its average Bayesian Risk can be calculated (see SI).
This measure ranges between 0 and 1, with 0 indicating full agreement
between labellers for all pixels (n = 40000) in the label, and 1
indicating complete disagreement.

\hypertarget{classification-pipeline}{%
\subsubsection{Classification pipeline}\label{classification-pipeline}}

Upon completion of a batch of labels, the platform automatically
launches an ephemeral Elastic Map
Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf}
cluster consisting of tens of instances, depending on the size of the
AOI.

\hypertarget{feature-extraction}{%
\paragraph{Feature extraction}\label{feature-extraction}}

The first step is the extraction of additional features from each
seasonal image composite. Previous work showed that a large number of
simple features that summarize the statistical properties of reflectance
and vegetation indices in local neighborhoods are highly effective for
classifying smallholder croplands (Debats et al. 2016). We followed this
logic in this study, but were constrained to use a smaller feature space
because the storage and memory requirements for our mapping geographies
in this case were several orders of magnitude larger. For this
implemention, we thus extract a set of 16 features, which are the mean
and standard deviations calculated within an 11X11 and 5X5 moving
window, respectively (initial tests revealed these two window sizes to
be most effective), resulting in 24 overall features, including the
original bands (Table 1).

\begin{center}Table 1. List of image features.\end{center}

\begin{longtable}[]{@{}lll@{}}
\toprule
Feature & Window Size & N Features \\
\midrule
\endhead
RGB-NIR & 1X1 & 8 \\
Mean & 11X11 & 8 \\
Standard deviation & 5X5 & 8 \\
\bottomrule
\end{longtable}

Feature extraction and the conversion of image features is handled by a
combination of
\texttt{GeoTrellis}\footnote{https://github.com/locationtech/geotrellis},
\texttt{rasterio}\footnote{https://rasterio.readthedocs.io/en/latest/},
and \texttt{RasterFrames}\footnote{https://rasterframes.io/}. These
collectively extract subsets of imagery from the PlanetScope temporal
composites, derive the features, and convert these into Apache Spark
DataFrames. Features are extracted on the fly for each cell in the
training and validation sets, a functionality enabled by storing the
image composites as Cloud-optimized
Geotiffs\footnote{https://www.cogeo.org/} (COGs).

\hypertarget{classification}{%
\paragraph{Classification}\label{classification}}

Once the features from the training sites are extracted into
\texttt{RasterFrames}, these are combined with their corresponding
labels and passed to the machine learning classifier, a
\texttt{SparkMLlib} implementation of Random Forests (Breiman 2001). For
this study, the model was trained with a balanced sample and a tree
depth of 15 and total tree number of 60, which initial testing showed to
provide a reasonable balance between computational time and performance.

After fitting, the model is applied to the features of the model
validation set, and a set of performance metrics is calculated,
including binary accuracy, the F1 score (the geometric mean of precision
and recall), and the area under the curve of the Receiver Operating
Characteristic (Pontius and Si 2014).

\hypertarget{the-active-learning-loop}{%
\subsubsection{The active learning
loop}\label{the-active-learning-loop}}

After fitting and model evaluation a second prediction is undertaken to
enable active learning. The feature extraction process is repeated for
the rest of the mapping geography that falls outside of the training and
validation sample, but applied to a subset of randomly drawn pixels from
each cell in order to reduce computational demand. The fitted model is
applied to predict the cropland probability for these selected pixels,
and an uncertainty criterion (Debats et al. 2017) is calculated for each
grid cell:

\begin{equation}
\mathrm{Q_I = \sum_{I(x, y) \epsilon I} (p(x, y) - 0.5)^2}
\end{equation}

Where Q is the uncertainty for grid cell I, calculated from the
predicted probability \emph{p} of the randomly selected subset of pixels
(x, y) drawn from it. Pixels with predicted probabilities closer to 0.5
are least certain as to their classification, thus images with the
lowest values of \(Q\) represent sites posing the most difficulty for
the classifier.

After scoring with the uncertainty criterion, the top \emph{N} most
uncertain grid cells are selected and sent back to the labelling
platform, which are then digitized by the labellers. The resulting
consensus labels from the actively selected sample are added to the
initial randomly selected sample, and a new cluster is launched. The
model is retrained, assesses uncertainty across the remaining unlabelled
sites, and selects the next most uncertain sites for labelling. This
loop repeats until model performance gains against the validation set
show diminishing returns.

\hypertarget{segmentation}{%
\subsubsection{Segmentation}\label{segmentation}}

After the final iteration, the segmentation algorithm is invoked, which
entails several steps. In the first step, the meanshift algorithm
(Yizong Cheng 1995) is applied to the original bands of the dry season
composite. A Sobel filter is then applied to the green, red, and
near-infrared mean-shifted bands and the probability map, and a combined
edge image is computed using the sum of these four edge images for the
dry season only. A compact watershed algorithm (Neubert and Protzel
2014) is then run on the weighted edge image, with a high level of
segmentation specified. In this case, we specified \textbf{6400}
segments per tile.

Third, a region adjacency graph is constructed for each image tile, in
which each node represents all pixels within each polygon created in the
previous step. The edge between two adjacent regions (polygons) is
calculated as the norm of the difference between the means of normalized
colors of all bands. Hierarchical merging is then applied, in which the
most similar pairs of adjacent nodes are merged until there are no edges
remaining below a predetermined threshold of 0.05.

In the fourth step, the merged polygons are overlaid with the posterior
probabilities resulting from the final active learning loop, and
polygons in which the average posterior probability is greater than a
predetermined threshold (here 0.5, but could vary locally) are retained
as field polygons.

In the final step, the retained polygons are refined by removing holes
and smoothing their boundaries using the Visvalingam algorithm
(Visvalingam and Whyatt 1993). Neighboring polygons that overlap along
tile boundaries are then merged.

To assess the accuracy of the final segmented boundaries, we used a
two-step approach. First, we assessed the overall thematic accuracy of
the resulting classification against our map reference data. Second, to
assess the quality of the segmentation, we compared the mean area and
relative frequencies of the segmented polygons within different size
classes against the same metrics derived from the digitized fields of
the most accurate worker to create the given map. We selected this
relatively simple procedure, as opposed to more complex measures of
object accuracy (Ye et al. 2018), because, on the one hand, both the
automated segmentation algorithm and labellers are cueing in on the same
features--abrupt, physically detectable breaks within the imagery. On
the other hand, no matter how well the intepreted/segmented boundaries
align with the boundaries of fields in the imagery, it is logistically
difficult to evaluate performance against real-world boundaries as the
spectral distinction of field boundaries will vary across different crop
types and land use arrangements.

\hypertarget{applying-the-system-to-map-ghana}{%
\subsection{Applying the system to map
Ghana}\label{applying-the-system-to-map-ghana}}

We applied the system to map Ghana's croplands, excluding areas
primarily cultivated with tree crops. Ghana has several distinct
agricultural regions, ranging from the primarily grain and vegetable
crop producing regions in the northern savannas to tree crop-dominated
system in the forested southwest, where cocoa and oil palm are among the
dominant crops. For these latter regions, we did not attempt to classify
tree crops, and instead mapped clearings that potentially contain field
crops or newly felled or recently replanted tree crops. We made this
decision because PlanetScope's resolution is not high enough for
labellers to distinguish many tree crops from surrounding forest, and
the boundaries of many tree crops (e.g.~cocoa) are often not visible.

To create the cropland maps, we divided the country into 16 distinct
AOIs, which were developed by grouping together each one degree cell
fully contained within Ghana with the tiles belonging to any adjacent
degree cell that overlapped neighboring countries (Figure
\ref{fig:aois}A). The exception was AOI 16, which consisted of the four
degree cells intersecting Ghana's southern coast. The resulting AOIs
ranged from 12,160 to 23,535 km\(^2\) in extent (average = 15,457
km\(^2\)), A separate active learning and segmentation process was run
for each of these AOIs.

To collect the initial randomized samples for model training, we grouped
the AOIs into three clusters: a northern cluster comprising the 6
northernmost AOIs (Cluster 1), a central to southeastern cluster
(Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern
AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the
forest zone AOIs (10, 11, 13, 14, 16). Within each cluster, we randomly
selected and labelled 500 grid cells, which provided relatively large
initial training samples for these agro-ecologically similar regions,
while helping to minimize the overall amount of labelling effort. In
addition to these samples, we randomly selected and labelled 100 grid
cells within each AOI to provide a validation sample.

After collecting the initial training and validation samples, we trained
a starter model for each cluster and applied it to each of the block's
AOI. For each iteration, 100 samples were actively selected within each
AOI, and added to the training pool.

During the collection of training and validation samples, labellers were
tasked to only digitize active or recently active crop fields, avoiding
tree crops, and fallow or potentially abandoned fields (see SI for the
digitizing rules).

To evaluate the performance of the system, we performed several analyses
described in sections 2.3.1-4.

\hypertarget{image-quality}{%
\subsubsection{Image quality}\label{image-quality}}

We evaluated the overall quality of the resulting seasonal image
composites by assessing a random selection of 50 tiles. We graded both
seasonal composites for each tile using a four category quality score,
which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3)
the number of visible scene boundary artifacts, and 4) the proportion of
the image that had its resolution degraded below the typical 3-4 m
PlanetScope resolution (e.g.~because of between-date image
mis-registrations). Each category was qualitatively ranked from 0-3,
with 0 being the lowest quality, and 3 the highest (see SI for complete
protocol), making the highest possible score 12. We rescaled scores to
fall between 0 and 1.

\hypertarget{model-gains-per-iteration}{%
\subsubsection{Model gains per
iteration}\label{model-gains-per-iteration}}

To assess the gain in model performance due to active learning, we
measured the change in accuracy, F1, and AUC (see 2.2.3.2) between each
iteration and between the first and last iterations for each AOI.

To evaluate whether active learning improved model performance relative
to a purely random approach to selecting new training sites, we ran
additional tests within a subset of AOIs (1, 8, and 15). We first
randomly selected and labelled 300-400 sites in each AOI. We then
progressively added 100 of the randomly selected samples to the relevant
training pool and retrained the model, repeating the process so that the
number of iterations and samples matched those from the active learning
process. We then compared the difference in accuracy, AUC, and F1
between the randomly trained models and those trained with active
learning (Debats et al. 2017).

\hypertarget{accounting-for-label-error}{%
\subsubsection{Accounting for label
error}\label{accounting-for-label-error}}

To quantify the potential impact of label error on classification
results, we evaluated the performance differences between models trained
with three different sets of labels: 1) those from the lowest scoring
labeller to map each training site, 2) those from the highest scoring
labeller, and 3) the consensus labels. This assessment follows
recommended Tier 1 (i.e.~best practice) standards to account for
training data errors (Elmes et al. 2020).

\hypertarget{accuracy-assessment}{%
\subsubsection{Accuracy assessment}\label{accuracy-assessment}}

The model performance assessments described above (2.3.2-3) were not
fully independent because they used the same validation sites over
multiple iterations (Elmes et al. 2020). To independently assess the
accuracy of our final map products, we followed recommended guidelines
(Stehman and Foody 2019) to create a separate map reference sample. We
used a stratified design, randomly assigning square polygons of
\textasciitilde0.1 ha extent into cropland and non-cropland strata,
developed the map of segmented field boundaries. Four classes were used
for the map reference sample: cropland; non-cropland; unsure but likely
cropland; unsure but likely non-cropland. The latter two classes were
used to provide insight into the degree of uncertainty in the map
reference sample. For efficiency, two separate supervisors evaluated
separate portions of the reference sample, but both jointly assessed a
small subset of the sample. We calculated their level of agreement on
this subset to provide an additional assessment of uncertainty in the
map reference sample (Stehman and Foody 2019). The SI contains further
details on the design and collection of the map reference sample.

The map reference polygons were then intersected with both the
probability images and the segmented field boundaries, and confusion
matrixes between the map reference labels and the extracted map classes
were constructed to assess the categorical accuracy of each map product.
We calculated the overall accuracy for each map, as well as the
class-wise User's and Producer's accuracy, as well as the 95\%
confidence intervals for each accuracy measure (Olofsson et al. 2013,
2014, Stehman and Foody 2019).

To assess the accuracy of the segmented field boundaries, we compared
the size class distributions of the segmented field boundaries against
those of the workers' digitized polygons at map validation sites. We
chose this approach because of existing uncertainties in polygon-based
accuracy assessment methods (Ye et al. 2018), and because the map's
ability to represent field sizes was of greatest interest. To undertake
this comparison, we selected the field polygons from the most accurate
labeller to digitize each of the 100 validation sites in each AOI, and
calculated the site-wise average area and number of polygons. We then
calculated the same statistics from the segmented boundaries that
intersected each validation grid. We compared the distributions and
proximity of two measures of central tendency (mean and median)
calculated from the two datasets for each AOI, and across all AOIs.

\hypertarget{assessing-the-characteristics-of-ghanaian-cropland}{%
\subsection{Assessing the characteristics of Ghanaian
cropland}\label{assessing-the-characteristics-of-ghanaian-cropland}}

Using the final mapped results, we calculated the estimated area of
cropland in Ghana, as well as the average size and total number of
fields in the different AOIs. We used the map reference sample to
calculate adjusted area estimates and confidence intervals for each map
class, and used the differences between labellers' polygons and
segmented boundaries at validation sites to calculate bias-adjusted
estimates of mean field sizes and the total number of fields.

\hypertarget{results}{%
\section{Results}\label{results}}

We developed maps of Ghana's cultivated croplands within an area of
247,299 km\(^2\), which included portions of neighboring countries
overlapped by images tiles.

\hypertarget{image-catalog-and-quality}{%
\subsection{Image catalog and quality}\label{image-catalog-and-quality}}

To develop the maps, we first generated an image catalog for the 8,116
tiles covering Ghana. This entailed processing all PlanetScope imagery
intersecting these tiles between May-September, 2018 (the growing
season) and December, 2018 to February, 2019 (the subsequent dry
season). The longer period was necessary for the growing season because
of the frequent cloud cover, which substantially limits the number of
clear scenes for any tile (Figure S3). For the cloudiest regions (AOIs
10, 11, 13, 14, 16) we started the dry season window in November.

An assessment with two observers (see SI for observer details) found
that average quality per growing season composite tile was 0.88, with 70
percent having scores \(\geq\) 0.85, while the average quality of dry
season composites was 0.92 (74 percent \(\geq\) 0.85). Composite quality
in both seasons was highest in the northern half of the country and
lowest in the southwest (Figure \ref{fig:imqual}A), where the
substantially greater cloud cover resulted in a much lower density of
available PlanetScope imagery for each time period (Figure S3).

\begin{figure}

{\centering \includegraphics[width=0.7\linewidth]{figures/figure4}

}

\caption{The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D).}\label{fig:imqual}
\end{figure}

\hypertarget{active-learning}{%
\subsection{Active learning}\label{active-learning}}

\hypertarget{training-data-collection}{%
\subsubsection{Training data
collection}\label{training-data-collection}}

After training of models with the initial randomly selected label sets,
the active learning process was run for 3 iterations for 12 of the 16
AOIs, resulting in 800 labels per AOI. AOIs 10 and 14 stopped after one
and two iterations, respectively, as they started with high initial
validation accuracies (\textgreater83\%) and showed little subsequent
improvement. The models for these two AOIs were thus trained with 600 -
700 samples. AOI 15 was run for 4 iterations (900 samples), while AOI 3
underwent a second active learning cycle because the model produced
during the first cycle was inaccurate (see SI). In this second run, 300
initial training sites randomly selected within the AOI were used
(Figure S4A), followed by 2 subsequent active learning iterations,
resulting in a training sample of 500. Labels collected during the
active learning iterations showed distinct patterns in several AOIs,
which often fell along ecotones, such as the boundaries between
agroecozones (see Figure S4A). The total number of unique training and
validation sites across the country were 6,299 and 1,600, respectively.

The distribution of training and validation sample collection effort was
divided across 20 labellers, with a core group of 13 who mapped more
than 1,000 sites each. As each training site was mapped by 4 separate
labellers, 34,014 sets of vector labels were made. Each labeller
digitized an average of 2,001 (see Figure S5A for more details on
labelling effort). Labeller accuracy was scored 9,389 times against 98
unique training reference sites (Figure S4A), with each labeller
assessed an average of 552 times at a rate of 1 training reference site
for every 3.62 training site mapped. The mean of each labeller's average
accuracy score was 0.71 (range 0.6 to 0.85; see Figure S5B for detailed
score distributions).

After each site was mapped by four labellers, consensus labels were
generated. The Bayesian Risk (see SI) of each training and validation
label was calculated as an additional measure of label quality. The
average risk was 0.122 for training labels and 0.127 for validation
labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7),
falling between 0.157 for training and 0.173 for validation labels
(Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13,
14, 16; training risk = 0.079; validation risk = 0.065). Label risk in
the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower
(training = 0.127; validation = 0.136) than in the north. Labeller
experience also appeared to reduce risk, which we observed during a
relabelling of the 500 initial random site in this cluster (see SI); the
mean risk of the updated labels was 0.055, compared to 0.172 for
original labels.

\hypertarget{performance-gains-during-active-learning}{%
\subsubsection{Performance gains during active
learning}\label{performance-gains-during-active-learning}}

Model performance was calculated for each iteration within each AOI. The
average accuracy, AUC, and F1 at iteration 0 were 0.786, 0.809, and
0.464, respectively, increasing to 0.825, 0.818, and 0.507 by iteration
3 (Figure \ref{fig:alperformance}). These differences represent
respective gains of 4.9, 1.1, and 9.1 percent for the three metrics. The
largest gains for each metric occurred on iteration 1, averaging 2.9, 1,
and 3.8 percent for accuracy, AUC, and F1, while the lowest gains were
realized on iteration 3, with accuracy, F1, and AUC respectively
increasing by just 1.2\%, 0.9\%, and 0.3\%. The scores achieved on the
final iteration varied substantially across AOIs and metrics. Accuracy
ranged between 0.725 (AOI 15) and 0.948 (AOI 16), while AUC varied from
0.725 (AOI 4) and 0.93 (AOI 11), and F1 from 0.252 (AOI 13) and 0.636
(AOI 8).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/figure5}

}

\caption{Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).}\label{fig:alperformance}
\end{figure}

The comparison of active versus randomized training sample collection
(in AOIs 1, 8, and 15) showed that the former approach outperformed the
latter. After three iterations, the accuracy, AUC, and F1 scores
resulting from active learning were respectively 0.8, 0.6, and 2.3
percent higher than the scores from a randomly trained model (Figure
S8). However, there was more variability in earlier iterations, with
average score differences of -1.7 (accuracy), 0.6 (AUC), and 0.8 percent
(F1) after iteration 1, and -0.3 (accuracy), 0.4 (AUC), and 1.8 (F1)
percent after iteration 2. The negative results for accuracy was caused
by results at AOI 15, where active learning accuracy was 8.37 percent
lower than random training after iteration 1 (see Figure
\ref{fig:alperformance}). In comparison, iteration 1 active learning
accuracies were 2.88 and 0.45 percent higher than random training for
AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15
exceeded randomized training after 4 iterations.

\hypertarget{the-impact-of-training-data-error}{%
\subsubsection{The impact of training data
error}\label{the-impact-of-training-data-error}}

The potential impact of label errors on map quality was assessed in four
AOIs (1, 2, 8, and 15). The results of these tests showed that the
average accuracy, AUC, and F1 scores for models trained with the
consensus labels were respectively 0.772, 0.8, and 0.555 (Figure
\ref{fig:trainingimpact}). Performance metrics from consensus-trained
models were just 0.5 - 1.2 percent higher than those models trained with
the most accurate individuals' labels (accuracy = 0.762; AUC = 0.796; F1
= 0.55), but were 11.6 - 27.4 higher than models trained with the least
accurate individual labels (accuracy = 0.606; AUC = 0.716; F1 = 0.44).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{figures/figure6}

}

\caption{Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 score resulting from models trained with consensus labels, and labels made by the most and least accurate labellers to map each site. Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.}\label{fig:trainingimpact}
\end{figure}

A second measure of the impact of label error is found within the
correlations between the mean label risk per AOI and the model
performance metrics (Table S3). Accuracy and AUC had strong (Spearman's
Rank Correlation = -0.824 to moderate (r = -0.568) negative correlations
with label risk, while F1 had a weaker but moderate positive association
(r = 0.456). The positive sign of the latter relationship is
counter-intuitive, but is explained by risk's association with
precision, one of two inputs to F1, which was moderately positive
(0.629), whereas risk had a negligible correlation with recall (0.206),
F1's other component. The correlation between risk and the false
positive rate (0.688), another important performance metric, shows that
labelling uncertainty may increase model commission error.

\hypertarget{map-accuracy}{%
\subsection{Map accuracy}\label{map-accuracy}}

\hypertarget{categorical-accuracy}{%
\subsubsection{Categorical accuracy}\label{categorical-accuracy}}

We used a map reference sample of 1207 sites (487 cropland; 720
non-cropland) to evaluate the accuracy of the per-pixel classifications
(resulting from thresholding the Random Forests probability), as well as
the segmented field boundary maps. We first evaluated the uncertainty in
the map reference classes by assessing 1) the overall agreement between
map reference labels collected by two separate supervisors at 23 sites,
and 2) the confidence of the labels assigned by the supervisors (see SI
for details). The first measure showed that the two individual
supervisors' labels agreed at 87\% of common sites, while the second
showed that 15.7 of sites were labelled with the two classes that
indicated a level of uncertainty.

We found that the overall accuracy of the pixel-wise classifications was
88\% against this map reference sample (Table \ref{tab:mapaccuracy}).
Confining the map reference sample to four distinct zones (Figure S10A)
shows that overall accuracy ranged from 83.3\% in Zone 1 (AOIs 1-3) to
93.6\% in Zone 3 (AOIs 10, 11, 13, 15, and 16). The Producer's accuracy
of the cropland class was 61.7\% across Ghana, ranging from 45.6\% in
Zone 3 to 67.9\% in Zone 1, while the User's accuracy for was 67.3\%
overall, ranging from 59.8\% in Zone 4 to 71.2\% in Zone 1. Both
measures of accuracy were substantially higher for the non-cropland
class across all zones, typically exceeding 90\%. The lowest accuracies
for the non-cropland class was in Zone 1 (Producer's = 89.3\%; User's =
87.7\%).

The overall accuracies obtained from the segmented maps were generally
1-2 percentage points lower than those of the per-pixel maps, while
User's accuracies tended to be 8-10 percentage points less (Table
\ref{tab:mapaccuracy}). In contrast, Producer's accuracies were 15-20
points higher than in the per-pixel map. The segmentation step therefore
helped to reduce omission error while substantially increasing
commission error.

\begin{table}
\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on Random Forests predictions; top 5 rows) and the segmented map (bottom 5 rows). Results are provided for 4 zones (Zone 1 = AOIs 1-3; Zone 2 = AOIs 4-9; Zone 3 = AOIs 10, 11, 13, 14, 16; Zone 4 = AOIs 12, 15) plus the entire country. The error matrix (with reference values in columns) provides the areal percentage for each cell, and the Producer's (P), User's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error. }
\includegraphics[width = 18cm]{figures/table2.png}
\label{tab:mapaccuracy}
\end{table}

\hypertarget{segmentation-accuracy}{%
\subsubsection{Segmentation accuracy}\label{segmentation-accuracy}}

The comparisons of digitized versus segmented field boundaries showed
that the mean field size across all validation sites averaged 4.97 ha
(Median = 3.75; StDev = 6.04), which was 1.41 times larger than the 2.06
ha (Median = 1.35; StDev = 3.26) mean area of labeller-digitized
polygons. This discrepancy was primarily caused by results in four AOIs
(2, 3, 7, and 15; Figure S11), where segments averaged between 7.76 and
10.76 ha, compared to 2.18 - 2.77 ha for the corresponding
hand-digitized polygons. The number of segmented fields per validation
site averaged 3.08 (median = 2.66; StDev = 2.9) compared to 4.4 (median
= 3.38; StDev = 4.52) for digitized polygons (Figure S12).

\hypertarget{ghanas-croplands}{%
\subsection{Ghana's croplands}\label{ghanas-croplands}}

Two separate maps of cropland were produced for each AOI, a per-pixel
map derived from the cropland probabilities, and the vectorized map of
field boundaries (Figure \ref{fig:mainmap}). The former provides the
more accurate picture of cropland distributions in Ghana, which are most
concentrated in the Southeastern corner (AOI 15), the central-western
region (AOI 7, the northeastern and northwestern corners of AOIs 10 and
11, and the south of AOI 8), and the northeastern quadrant stretching
from AOI 9 through AOIs 5 and 6 and up to AOIs 2 and 3. The northern
third of AOI 1 also has noticeable densities of cropland. Several
notable areas of low cropland density are also apparent, indicating the
presence of large protected areas, such as Mole National Park in the
southeastern corner of AOI 1 and Digya National Park in the northwestern
corner of AOI 12. In contrast, the relative absence of cropland in AOIs
13, 14, and 16 does not reflect the scarcity of agriculture in these
areas, but rather the predominance of tree crops, which we did not map.

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{figures/figure7}

}

\caption{The distribution of croplands in Ghana. The main map shows the percentage of croplands in each 0.005 degree grid cell, derived from the predicted cropland probabilities. The insets  on the margins illustrate predicted probabilities (top map in each couplet) at original image resolution (0.000025 degrees) and segmented field boundaries overlaid on the dry season PlanetScope composite, for four separate tiles. Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.}\label{fig:mainmap}
\end{figure}

Both the per-pixel and vectorized maps, when combined with the map
reference sample, enable separate estimates of the total extent of
croplands in Ghana. The cropland extent estimated from the vectorized
map is 42,359 km\(^2\) (with a margin of error of 4,395 km\(^2\)), or
17.1 (15.4-18.9\%) of the mapped area. The estimate based on the per
pixel map is 43,233 km\(^2\) (margin of error = 4,904 km\(^2\)), or 17.6
(15.6-19.6\%) of area.

The vectorized map provides additional information on how the
characteristics of croplands can vary geographically, ranging from
narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s
lower right inset) to more densely packed, less distinctly shaped fields
in AOI 5 (upper right inset in Figure \ref{fig:mainmap}). To explore how
field characteristics varied geographically, we mapped the average field
size and total number of fields within each 0.05 degree tile (Figure
S13). These patterns generally correspond to those seen in the cropland
density map (Figure \ref{fig:mainmap}), with larger sizes and field
counts generally occurring in areas of higher field density, although
the biases inherent in both measures (Figures S11-12) complicate the
interpretation of those variations. However, we can use the estimated
biases to develop adjusted estimates of field sizes and counts for each
AOI, and for Ghana overall (Table \ref{tab:sizentab}). These adjusted
estimates show that the typical field size in Ghana is 1.73 ha, ranging
from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone
AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern
half of the country (Table \ref{tab:sizentab}). The total number of
fields was estimated to be 1,662,281 overall, or 205 fields per tile on
average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.

\begin{table}[!h]

\caption{\label{tab:sizentab}The average size and total number of crop fields for each AOI and for Ghana overall. The original and bias-adjusted values for each measure are provided, as well as the total number of 0.05$^\circ$ degree tiles in each AOI.}
\centering
\begin{tabular}[t]{>{\raggedleft\arraybackslash}p{0.45in}>{\raggedleft\arraybackslash}p{0.45in}>{\raggedleft\arraybackslash}p{0.45in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}>{\raggedleft\arraybackslash}p{0.85in}}
\toprule
AOI & N tiles & Size & Size (adj) & N & N / tile & N (adj) & N (adj) / tile\\
\midrule
1 & 777 & 3.71 & 1.26 & 97,822 & 126 & 127,580 & 164\\
2 & 597 & 7.66 & 1.96 & 87,666 & 147 & 120,651 & 202\\
3 & 501 & 8.24 & 2.18 & 108,819 & 217 & 104,422 & 208\\
4 & 465 & 2.44 & 2.82 & 26,276 & 57 & 50,163 & 108\\
5 & 400 & 4.24 & 2.09 & 43,290 & 108 & 53,756 & 134\\
\addlinespace
6 & 429 & 5.10 & 2.15 & 81,363 & 190 & 145,347 & 339\\
7 & 471 & 5.64 & 1.49 & 93,282 & 198 & 123,005 & 261\\
8 & 400 & 4.89 & 1.98 & 55,500 & 139 & 78,868 & 197\\
9 & 479 & 4.10 & 1.82 & 72,081 & 150 & 89,840 & 188\\
10 & 630 & 2.24 & 1.04 & 119,019 & 189 & 170,907 & 271\\
\addlinespace
11 & 400 & 3.65 & 1.52 & 52,510 & 131 & 94,709 & 237\\
12 & 471 & 3.44 & 1.77 & 44,667 & 95 & 52,947 & 112\\
13 & 627 & 0.84 & 0.96 & 67,996 & 108 & 125,368 & 200\\
14 & 400 & 1.09 & 2.72 & 56,006 & 140 & 101,767 & 254\\
15 & 548 & 4.95 & 1.54 & 75,752 & 138 & 105,681 & 193\\
\addlinespace
16 & 521 & 0.95 & 1.41 & 49,097 & 94 & 117,268 & 225\\
Ghana & 8,116 & 3.92 & 1.73 & 1,131,146 & 139 & 1,662,281 & 205\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

These results demonstrate a capability for mapping the characteristics
of smallholder-dominated cropping systems at high spatial resolution,
annual time steps, and national scales. The resulting maps provide an
updated and more granular view of the distribution and extent of
croplands in Ghana, complementing existing national to regional land
cover maps derived from moderate resolution sensors (Hackman et al.
2017, Xiong et al. 2017, ESA n.d.). This prior work found that cropland
covered 19.4 (Xiong et al. 2017) to 32\% (Hackman et al. 2017) of Ghana
in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2\% (Table
\ref{tab:mapaccuracy}), and our map reference sample-based estimates
finds 17.1-17.6\% cover. Our results thus suggest that Ghana's cropland
is less than previously estimated, but the difference is perhaps
attributable to our use of a cropland definition that excluded longer
fallows and abandoned fields, which in some regions can account for over
half of the total area that could be counted as cropland (Tong et al.
2020).

In addition to the more detailed update of cropland extent, our maps
also provide new information on the size and number of fields in Ghana
(Figures \ref{fig:mainmap}, S11-12). Previous work to estimate such
agricultural characteristics have often focused on farm, rather than
field, size using census data (Von Braun 2004, Samberg et al. 2016,
Jayne et al. 2016, Lowder et al. 2016). Efforts to map field boundaries
in smallholder-dominated agricultural systems have either used \emph{in
situ} data collection (Carletto et al. 2013, 2015) or remote sensing
studies over relatively small (e.g. Forkuor et al. 2014, Persello et al.
2019) or discontiguous (Estes et al. 2016a) areas. The most extensive
studies to date used crowdsourced volunteers to classify fields into
broad size classes, based on their interpretations of imagery sampled
from high resolution virtual globes (Fritz et al. 2015, Lesiv et al.
2019). Those efforts included country-specific results for Ghana (n =
263), which can be converted into an average field size estimate of 5.33
ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class. Data sourced from Table S3 in Lesiv et al. 2019.}.
This estimate exceeds our Ghana-wide average segment size (3.92 ha;
Table \ref{tab:sizentab}), but is closer to the mean (4.97 ha) within
AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample
appears to have been collected. However, our bias-corrected estimates of
1.73 (Ghana-wide) and 1.87 (AOIs 1-9, 12, and 15) ha were much smaller.

\hypertarget{map-accuracy-and-key-sources-of-error}{%
\subsection{Map accuracy and key sources of
error}\label{map-accuracy-and-key-sources-of-error}}

Although the maps generated by our system provide valuable new
information, they nevertheless contain substantial errors. The overall
map accuracies (86.7-88\%, Table \ref{tab:mapaccuracy}) are near the
boundary of what might be considered \emph{achievable} map accuracy
(Elmes et al. 2020), given the inherent uncertainty in the map reference
sample, our best estimate of the ``truth,'' in which we have roughly
85\% confidence. However, accuracies for the cropland class were much
lower, falling between 62 (Producer's) to 67 (User's) percent
country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}),
meaning the model produced substantial commission and omission errors
for this class. The segmented boundary maps had fewer omission errors
(Producer's accuracy = 79\%), but higher false positive errors (User's
accuracy = 58.2\%). These accuracies are near the middle to upper ranges
of those reported for the cropland class in other large-area mapping
studies (Hackman et al. 2017, Xiong et al. 2017, Lesiv et al. 2017).

The patterns of cropland-class accuracies varied by zone. These zones
largely align, albeit with some discrepancies, with the country's
agroecozones, thus the accuracy patterns may be partially attributed to
some regions being harder to map than others. Producer's accuracies for
both maps were highest in the two northern zones (1 and 2), which are
primarily savannas (Figure S10), and lowest in zones 3 and 4, which are
comprised of forest or coastal savannas. User's accuracies followed a
similar pattern, with the exception of Zone 3, which had the highest
User's accuracy, albeit from a very small sample. Aligning the reference
samples more precisely with agroecozone boundaries (Figure S10B)
provides further insight into error patterns (Table S4). Coastal
savannas in the southeast had the highest Producer's accuracy but lowest
User's accuracy for the per-pixel map, presumably because this region's
numerous areas of high density cropland, combined with low woody cover
in surrounding uncultivated areas, helped to promote commission error.
Maps in the two northern savanna agroecozones had the best balance
between omission and commission error, and had the highest overall
User's accuracy. The transitional (between forest and savanna)
agroecozone had a very low Producer's accuracy (21\%), which likely
reflects the fact that it was divided between several AOIs for mapping
(Figure S4), within which it typically covered a smaller share of area
relative to the other agroecozones. This likely caused insufficient
representation of this AEZ in training samples, particularly in AOIs 10
and 11 (Figure S4B).

Beyond the errors linked to regional differences, several other
important factors contributed to reducing the accuracies in the cropland
class. The first of these stems from the overall mapping extent and the
high resolution of the data. Given the goal of developing a
country-scale map at high resolution, the attendant data volume required
us to use a relatively small set of image features and less than the
recommended tree number and depth (Maxwell et al. 2018) in our Random
Forests implementation, in order to limit computational costs. Previous
work found that Random Forests achieves much better performance on
small-scale croplands when trained on a much larger number of features
(Debats et al. 2016). However, applying such a large feature set within
the extent of our AOIs was intractable, as the computing time would have
been several-fold larger than the \textasciitilde4-8 hours of runtime on
800 CPUs required for a single active learning iteration, followed by
\textasciitilde10-14 hours for prediction. This reduced the skill of the
model, particularly when it came to differentiating cropland from
adjacent bare patches or natural vegetation with sparse herbaceous
cover, which were common in many AOIs.

The inherent difficulty of the labelling task was another major limiting
factor. Our system was designed to minimize the error inherent in
labelling, but determining croplands from non-croplands in these
agricultural systems can be a difficult task. Labellers have to evaluate
multiple image sources and to rely heavily on judgement, which
inevitably leads to errors. Interpretation is particularly hard where
the background savanna vegetation and croplands have similar reflectance
during the dry season, which is a particular problem in AOIs 2 and 3.
Smaller field sizes also complicate labelling, as these become
increasingly indistinct in the \textasciitilde4 m PlanetScope
composites. The difficulty of labelling is reflected in the magnitude of
the Bayesian Risk metrics (Figure S6), and by the average score achieved
by each labeller against our training reference dataset (71\%; Figure
S5B). Although prior work (Rodriguez-Galiano et al. 2012, Mellor et al.
2015) found that Random Forests are robust to label errors, we found
that they have substantial impact (Figure \ref{fig:trainingimpact}),
which suggest that simply improving label quality may be one of the
single most important investments towards improved model accuracy.

Image quality was another issue, although primarily in the forested
AOIs, where frequent cloud cover reduced the number of available images
in all seasons, resulting in composites with more brightness artifacts
and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities
to discern fields, and doubtless affected model predictions. There is
little to be done to mitigate these errors, short of confining imagery
to the less cloudy dry season, which may further undermine model
performance, given the importance of multi-temporal imagery for cropland
classification (Debats et al. 2016, Defourny et al. 2019). Composite
quality could be improved by using imagery from the same seasons over
multiple years, but this would undermine the goal of developing annual
maps, while the dynamism of the croplands would blur field boundaries
within the imagery.

The final major source of error arose from the segmentation process. The
vectorized maps had high commission errors caused by uncertainties in
the Random Forests predictions. Model uncertainty led to many pixels
over non-crop areas with probabilities straddling the 0.5 classification
threshold. Segments that intersected such areas were retained as fields
when the average probability of intersecting pixels exceeded 0.5. A more
accurate classifier would reduce such errors, or the application of a
locally varying classification threshold (e.g. Waldner and Diakogiannis
2020). Over-merging was another source of error in the segmentation
algorithm, which in some areas led to overestimated field sizes and
unrealistic shapes, particularly in high density croplands (e.g.~in AOIs
2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent
fields are often indistinct in the PlanetScope imagery. Minimizing or
preventing merging would help in such cases, although could would result
in the opposite problem, over-segmentation, and thereby underestimate
field size.

\hypertarget{error-mitigation-features}{%
\subsection{Error mitigation features}\label{error-mitigation-features}}

Despite the error sources mentioned above, several features of the
system proved effective in mitigating error, leading to a higher overall
accuracy than would have otherwise been possible. Label accuracy
assessment and consensus labelling appeared to be the most effective
error mitigation tools. Label accuracy measures allowed us to quantify
the substantial impact of label error on model performance (Figure
\ref{fig:trainingimpact}), while consensus labels substantially reduced
individual labelling errors, resulting in maps that were more accurate
than they would have been had we used individually generated labels.
Labeller-specific accuracy measures also helped to improve the quality
of the consensus labels, by placing higher weight on labels more likely
to be accurate during the merging process, rather than giving equal
weight to potentially less accurate labels. The ability to select the
most accurate individual labels for a site also allowed us to develop
independent estimates of field size to which measures of confidence can
be attached (Figure S5B), which we were in turn able to correct
estimates of field sizes and numbers (Table \ref{tab:sizentab}).

The active learning approach helped to improve overall model performance
relative to randomized training site selection, in line with findings
from two recent efforts (Debats et al. 2017, Hamrouni et al. 2021).
Although the performance gains relative to randomized model training
that we observed were smaller (e.g. Debats et al. (2017) 29\% higher
model performance after one iteration, and 8\% higher on the final
iterations), those comparisons were made from lower initial bases, with
initial training samples that were less than 1/10 the size, in terms of
pixels, of our initial training sample. Our large initial randomly
selected sample (500 grid cells) meant that our models were
substantially trained before they were exposed to actively selected
labels, thereby diluting their impact on performance. Nevertheless, the
higher average performance of the active approach across three
performance metrics demonstrated its effectiveness. Most notable were
the larger improvements seen in the F1 score (Figure S8), a balanced
performance metric. Gains in accuracy and AUC were smaller. For
accuracy, the reduced advantage was primarily due to active learning
being outperformed by randomized training after the first iteration in
AOI 15, which proved one of the hardest AOIs to both map and label.
Active learning likely resulted in the selection of sites that were
harder to label than randomly selected ones, leading to more label
error, and thus initially lower model accuracy. However, this deficit
was overcome by the 5th iteration. The plateau in AUC gains at 0.5\%
better than randomized training reflects the findings that active
learning reduced both the false and true positive rates, the two inputs
to AUC. Although the decline of the false positive rate (30.7\% between
Iterations 0 and 3) was nearly three times larger than that of the true
positive rate (10.9\%), AUC should be quite sensitive to the reduction
in the latter, as it assesses how the tradeoff between the two rates
varies across a full range of possible classification thresholds
(Pontius and Si 2014).

The detail, temporal precision, and large extent of these maps was
enabled by the system's ability to process PlanetScope data, which is
currently the only source of sub-5 meter imagery with daily coverage
(McCabe et al. 2017). Daily revisits are important for creating seasonal
composites within a single year over cloudy areas. The compositing
technique we developed allowed us to develop a complete image catalog
for the country representing the two seasons for 2018 agricultural year.
Although Sentinel-2 is free, has better radiometric quality, and has
sufficiently high resolution (10 m) to accurately classify small-scale
agricultural systems (e.g. Defourny et al. 2019, Kerner et al. 2020),
its 5-day interval may be too infrequent to generate sufficiently
cloud-free composites during the growing season over southern Ghana.
Sentinel-1 is not affected by the same problem, but labelling fields in
more coarsely resolved radar images is challenging.

\hypertarget{lingering-questions}{%
\subsection{Lingering questions}\label{lingering-questions}}

Several potential issues not addressed in our assessment merit further
exploration. One of these is the degree of correspondence between image-
and ground-collected labels. However, such comparisons may reveal
unresolvable differences between the two perspectives. The highly
dynamic nature of many agricultural systems means that relatively narrow
differences between the dates of ground- and image-based digitizing
campaigns can lead to substantial disagreement between the resulting
field boundaries, simply because the fields themselves may have shifted
during the interval (Elmes et al. 2020). These discrepancies could be
further exacerbated by differences in the definition of what constitutes
a field, which might vary on the ground depending on who is being asked,
or who is doing the collecting. These factors suggest that ground versus
image label differences would not necessarily indicate how far
image-based labellers were from the ``truth.'' Nevertheless, a
comparison against ground data would help to assess how accurately
image-collected labels capture the typical size of fields, and thus
merits further investigation for this reason.

The temporal discrepancies mentioned above (and discussed in Elmes et
al. 2020) are another reason why we chose not to label on basemap
imagery (in addition to restrictive usage terms), which is typically
several years old (Lesiv et al. 2018). However, we did not assess
whether the higher label accuracy one might achieve by digitizing on a
\textless1-2 m resolution basemap would offset model errors caused by
temporal mismatches.

Another potential issue is the degree to which our assessment of the
impact of label error on model performance (Figure
\ref{fig:trainingimpact}) was influenced by the validation labels we
used, which were generated using the consensus method. This could have
confounded the assessment, particularly when comparing models trained
with the most accurate individual label and those trained with consensus
labels. However, the visual assessment of their resulting probability
maps confirm the differences in scores: consensus and most accurate
individual labels produce nearly identical maps with relatively high
certainty, while low quality labels led to a markedly less certain map
(Figure S9).

\hypertarget{next-steps}{%
\subsection{Next steps}\label{next-steps}}

The maps presented here represent a version 1 product that is freely
available to use, along with its underlying code (see SI for details).
These data were developed according to the recommended best practices
for training and assessing error in machine learning models (Elmes et
al. 2020). In their current form, the maps may be useful for a variety
of research applications. For example, analyzing the distributions of
values in the probability maps may provide additional insight into the
relative extents of active versus fallow croplands (Tong et al. 2020).
However, use of these data, particularly for decision-making processes
(e.g.~cropped area estimates), should be careful to account for the
reported errors (Olofsson et al. 2014, Stehman and Foody 2019).

To facilitate the next step, generating more accurate version 2 maps,
several improvements will be made. The first is to replace Random
Forests with a more advanced convolutional neural network (CNN), which
can generate and learn from a large number of features representing a
variety of spatial scales (Ma et al. 2019). Recent work suggests that a
common architecture such as U-Net, when trained to distinguish field
edges from interiors and combined with a post-hoc segmentation routine,
is effective in delineating field boundaries (Waldner and Diakogiannis
2020). Our system can readily incorporate such a model. The labelling
platform already provides the methods needed to develop and assess the
quality of labels that include field edges and interior classes, while
active learning has proven to be effective for optimizing training
datasets for deep learning models (Liu et al. 2017, Cao et al. 2020).
Our current framework can be adjusted so that it starts by training a
CNN from scratch with a large initial random sample, and then uses a
transfer learning approach (Pan and Yang 2010) to update the model with
the most informative samples from different AOIs or agroecozones.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

This work demonstrates a proof of concept for developing high
resolution, annual maps of smallholder-dominated croplands at national
to regional scales, using a framework that can be readily updated to
improve map accuracy as technologies improve. Maps that include
information on field boundaries can help improve remote estimation of
crop planted area and yield, and provide deeper insights into important
socioeconomic aspects of agricultural systems, such as the relationships
between agricultural productivity and livelihoods. Such maps will be
important for developing an understanding of the rapid agricultural
change that is currently unfolding throughout much of the continent.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

The primary support for this work was provided by Omidyar Network's
Property Rights Initiative, now PLACE. Additional support was provided
by NASA (80NSSC18K0158), the National Science Foundation (SES-1801251;
SES-1832393), and Princeton University. Computing support was provided
by the AWS Cloud Credits for Research program and the Amazon
Sustainability Data Initiative. Azavea provided significant
contributions in engineering the machine learning pipeline. We thank
Meridia for providing information about local cropping systems and the
characteristics of fields, and Radiant Earth Foundation for advice and
guidance regarding machine learning best practices. We thank Manushi
Trivedi, Sitian Xiong, and Tammy Woodard for their contributions to the
underlying datasets and methods, and Michelle Gathigi, Omar Shehe, and
Primoz Kovacic for support and management of the labelling efforts.

\hypertarget{references}{%
\section{References}\label{references}}

\singlespace

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-azaveaRasterFoundry2020}{}%
Azavea. 2020. Raster {Foundry}.
https://github.com/raster-foundry/raster-foundry.

\leavevmode\hypertarget{ref-BeyCollectEarthLand2016}{}%
Bey, A., A. Sánchez-Paus Díaz, D. Maniatis, G. Marchi, D. Mollicone, S.
Ricci, J.-F. Bastin, R. Moore, S. Federici, M. Rezende, C. Patriarca, R.
Turia, G. Gamoga, H. Abe, E. Kaidong, and G. Miceli. 2016. Collect
{Earth}: {Land Use} and {Land Cover Assessment} through {Augmented
Visual Interpretation}. Remote Sensing 8:807.

\leavevmode\hypertarget{ref-BreimanRandomForests2001}{}%
Breiman, L. 2001. Random {Forests}. Machine Learning 45:5--32.

\leavevmode\hypertarget{ref-bullockThreeDecadesLand2021}{}%
Bullock, E. L., S. P. Healey, Z. Yang, P. Oduor, N. Gorelick, S. Omondi,
E. Ouko, and W. B. Cohen. 2021. Three {Decades} of {Land Cover Change}
in {East Africa}. Land 10:150.

\leavevmode\hypertarget{ref-caoHyperspectralImageClassification2020}{}%
Cao, X., J. Yao, Z. Xu, and D. Meng. 2020. Hyperspectral {Image
Classification With Convolutional Neural Network} and {Active Learning}.
IEEE Transactions on Geoscience and Remote Sensing 58:4604--4616.

\leavevmode\hypertarget{ref-CarlettoGuesstimatesGPStimatesLand2015}{}%
Carletto, C., S. Gourlay, and P. Winters. 2015. From {Guesstimates} to
{GPStimates}: {Land Area Measurement} and {Implications} for
{Agricultural Analysis}. Journal of African Economies 24:593--628.

\leavevmode\hypertarget{ref-CarlettoFactartifactimpact2013}{}%
Carletto, C., S. Savastano, and A. Zezza. 2013. Fact or artifact: {The}
impact of measurement errors on the farm size{}productivity
relationship. Journal of Development Economics 103:254--261.

\leavevmode\hypertarget{ref-ChenGloballandcover2015}{}%
Chen, J., J. Chen, A. Liao, X. Cao, L. Chen, X. Chen, C. He, G. Han, S.
Peng, M. Lu, W. Zhang, X. Tong, and J. Mills. 2015. Global land cover
mapping at 30 m resolution: {A POK}-based operational approach. ISPRS
Journal of Photogrammetry and Remote Sensing 103:7--27.

\leavevmode\hypertarget{ref-cohnImprovingGeneralizationActive1994}{}%
Cohn, D., L. Atlas, and R. Ladner. 1994. Improving generalization with
active learning. Machine Learning 15:201--221.

\leavevmode\hypertarget{ref-Darkmodifiablearealunit2007}{}%
Dark, S. J., and D. Bram. 2007. The modifiable areal unit problem
({MAUP}) in physical geography. Progress in Physical Geography
31:471--479.

\leavevmode\hypertarget{ref-davisTropicalForestLoss2020}{}%
Davis, K. F., H. I. Koo, J. Dell'Angelo, P. D'Odorico, L. Estes, L. J.
Kehoe, M. Kharratzadeh, T. Kuemmerle, D. Machava, A. de J. R. Pais, N.
Ribeiro, M. C. Rulli, and M. Tatlhego. 2020. Tropical forest loss
enhanced by large-scale land acquisitions. Nature Geoscience:1--7.

\leavevmode\hypertarget{ref-DebatsIntegratingactivelearning2017}{}%
Debats, S. R., L. D. Estes, D. R. Thompson, and K. K. Caylor. 2017.
Integrating active learning and crowdsourcing into large-scale
supervised landcover mapping algorithms. {PeerJ Preprints}.

\leavevmode\hypertarget{ref-Debatsgeneralizedcomputervision2016}{}%
Debats, S. R., D. Luo, L. D. Estes, T. J. Fuchs, and K. K. Caylor. 2016.
A generalized computer vision approach to mapping crop fields in
heterogeneous agricultural landscapes. Remote Sensing of Environment
179:210--221.

\leavevmode\hypertarget{ref-Defournyrealtimeagriculturemonitoring2019}{}%
Defourny, P., S. Bontemps, N. Bellemans, C. Cara, G. Dedieu, E.
Guzzonato, O. Hagolle, J. Inglada, L. Nicola, T. Rabaute, M. Savinaud,
C. Udroiu, S. Valero, A. Bégué, J.-F. Dejoux, A. El Harti, J. Ezzahar,
N. Kussul, K. Labbassi, V. Lebourgeois, Z. Miao, T. Newby, A. Nyamugama,
N. Salh, A. Shelestov, V. Simonneaux, P. S. Traore, S. S. Traore, and B.
Koetz. 2019. Near real-time agriculture monitoring at national scale at
parcel resolution: {Performance} assessment of the {Sen2}-{Agri}
automated system in various cropping systems around the world. Remote
Sensing of Environment 221:551--568.

\leavevmode\hypertarget{ref-desiereLandProductivityPlot2018}{}%
Desiere, S., and D. Jolliffe. 2018. Land productivity and plot size:
{Is} measurement error driving the inverse relationship? Journal of
Development Economics 130:84--98.

\leavevmode\hypertarget{ref-DruschSentinel2ESAOptical2012}{}%
Drusch, M., U. Del Bello, S. Carlier, O. Colin, V. Fernandez, F. Gascon,
B. Hoersch, C. Isola, P. Laberinti, P. Martimort, A. Meygret, F. Spoto,
O. Sy, F. Marchese, and P. Bargellini. 2012. Sentinel-2: {ESA}'s
{Optical High}-{Resolution Mission} for {GMES Operational Services}.
Remote Sensing of Environment 120:25--36.

\leavevmode\hypertarget{ref-DwyerAnalysisReadyData2018}{}%
Dwyer, J. L., D. P. Roy, B. Sauer, C. B. Jenkerson, H. K. Zhang, and L.
Lymburner. 2018. Analysis {Ready Data}: {Enabling Analysis} of the
{Landsat Archive}. Remote Sensing 10:1363.

\leavevmode\hypertarget{ref-ElmesAccountingtrainingdata2020}{}%
Elmes, A., H. Alemohammad, R. Avery, K. Caylor, J. R. Eastman, L.
Fishgold, M. A. Friedl, M. Jain, D. Kohli, J. C. Laso Bayas, D. Lunga,
J. L. McCarty, R. G. Pontius, A. B. Reinmann, J. Rogan, L. Song, H.
Stoynova, S. Ye, Z.-F. Yi, and L. Estes. 2020. Accounting for training
data error in machine learning applied to {Earth Observations}. Remote
Sensing 12:1034.

\leavevmode\hypertarget{ref-ESAESACCILAND}{}%
ESA. (n.d.). {ESA CCI LAND COVER} {} {S2} prototype {Land Cover} 20m map
of {Africa} 2016. http://2016africalandcover20m.esrin.esa.int/.

\leavevmode\hypertarget{ref-Esteslargeareaspatiallycontinuous2018}{}%
Estes, L., P. Chen, S. Debats, T. Evans, S. Ferreira, T. Kuemmerle, G.
Ragazzo, J. Sheffield, A. Wolf, E. Wood, and K. Caylor. 2018. A
large-area, spatially continuous assessment of land cover map error and
its impact on downstream analyses. Global Change Biology 24:322--337.

\leavevmode\hypertarget{ref-Estesplatformcrowdsourcingcreation2016}{}%
Estes, L. D., D. McRitchie, J. Choi, S. Debats, T. Evans, W. Guthe, D.
Luo, G. Ragazzo, R. Zempleni, and K. K. Caylor. 2016a. A platform for
crowdsourcing the creation of representative, accurate landcover maps.
Environmental Modelling \& Software 80:41--53.

\leavevmode\hypertarget{ref-EstesReconcilingagriculturecarbon2016}{}%
Estes, L. D., T. Searchinger, M. Spiegel, D. Tian, S. Sichinga, M.
Mwale, L. Kehoe, T. Kuemmerle, A. Berven, N. Chaney, J. Sheffield, E. F.
Wood, and K. K. Caylor. 2016b. Reconciling agriculture, carbon and
biodiversity in a savannah transformation frontier. Phil. Trans. R. Soc.
B 371:20150316.

\leavevmode\hypertarget{ref-federRelationFarmSize1985}{}%
Feder, G. 1985. The relation between farm size and farm productivity:
{The} role of family labor, supervision and credit constraints. Journal
of Development Economics 18:297--313.

\leavevmode\hypertarget{ref-forkuorIntegrationOpticalSynthetic2014}{}%
Forkuor, G., C. Conrad, M. Thiel, T. Ullmann, and E. Zoungrana. 2014.
Integration of {Optical} and {Synthetic Aperture Radar Imagery} for
{Improving Crop Mapping} in {Northwestern Benin}, {West Africa}. Remote
Sensing 6:6472--6499.

\leavevmode\hypertarget{ref-FourieBetterCropEstimates2009}{}%
Fourie, A. 2009. Better {Crop Estimates} in {South Africa}. ArcUser
Online.

\leavevmode\hypertarget{ref-FritzGeoWikionlineplatform2012}{}%
Fritz, S., I. McCallum, C. Schill, C. Perger, L. See, D. Schepaschenko,
M. van der Velde, F. Kraxner, and M. Obersteiner. 2012. Geo-{Wiki}: {An}
online platform for improving global land cover. Environmental Modelling
\& Software 31:110--123.

\leavevmode\hypertarget{ref-FritzHighlightingcontinueduncertainty2011}{}%
Fritz, S., L. See, I. McCallum, C. Schill, M. Obersteiner, M. van der
Velde, H. Boettcher, P. Havlík, and F. Achard. 2011. Highlighting
continued uncertainty in global land cover maps for the user community.
Environmental Research Letters 6:044005.

\leavevmode\hypertarget{ref-FritzMappingglobalcropland2015}{}%
Fritz, S., L. See, I. McCallum, L. You, A. Bun, E. Moltchanova, M.
Duerauer, F. Albrecht, C. Schill, C. Perger, P. Havlik, A. Mosnier, P.
Thornton, U. Wood-Sichra, M. Herrero, I. Becker-Reshef, C. Justice, M.
Hansen, P. Gong, S. Abdel Aziz, A. Cipriani, R. Cumani, G. Cecchi, G.
Conchedda, S. Ferreira, A. Gomez, M. Haffani, F. Kayitakire, J.
Malanding, R. Mueller, T. Newby, A. Nonguierma, A. Olusegun, S. Ortner,
D. R. Rajak, J. Rocha, D. Schepaschenko, M. Schepaschenko, A. Terekhov,
A. Tiangwa, C. Vancutsem, E. Vintrou, W. Wenbin, M. van der Velde, A.
Dunwoody, F. Kraxner, and M. Obersteiner. 2015. Mapping global cropland
and field size. Global Change Biology 21:1980--1992.

\leavevmode\hypertarget{ref-FritzComparisonglobalregional2010}{}%
Fritz, S., L. See, and F. Rembold. 2010. Comparison of global and
regional land cover maps with statistical information for the
agricultural domain in {Africa}. International Journal of Remote Sensing
31:2237--2256.

\leavevmode\hypertarget{ref-Fritzneedimprovedmaps2013}{}%
Fritz, S., L. See, L. You, C. Justice, I. Becker-Reshef, L. Bydekerke,
R. Cumani, P. Defourny, K. Erb, J. Foley, S. Gilliams, P. Gong, M.
Hansen, T. Hertel, M. Herold, M. Herrero, F. Kayitakire, J. Latham, O.
Leo, I. McCallum, M. Obersteiner, N. Ramankutty, J. Rocha, H. Tang, P.
Thornton, C. Vancutsem, M. van der Velde, S. Wood, and C. Woodcock.
2013. The need for improved maps of global cropland. Eos, Transactions
American Geophysical Union 94:31--32.

\leavevmode\hypertarget{ref-GorelickGoogleEarthEngine2017}{}%
Gorelick, N., M. Hancher, M. Dixon, S. Ilyushchenko, D. Thau, and R.
Moore. 2017. Google {Earth Engine}: {Planetary}-scale geospatial
analysis for everyone. Remote Sensing of Environment 202:18--27.

\leavevmode\hypertarget{ref-hackmanNewLandcoverMaps2017}{}%
Hackman, K. O., P. Gong, and J. Wang. 2017. New land-cover maps of
{Ghana} for 2015 using {Landsat} 8 and three popular classifiers for
biodiversity assessment. International Journal of Remote Sensing
38:4008--4021.

\leavevmode\hypertarget{ref-hamrouniLocalGlobalTransfer2021}{}%
Hamrouni, Y., E. Paillassa, V. Chéret, C. Monteil, and D. Sheeren. 2021.
From local to global: {A} transfer learning-based approach for mapping
poplar plantations at national scale using {Sentinel}-2. ISPRS Journal
of Photogrammetry and Remote Sensing 171:76--100.

\leavevmode\hypertarget{ref-HouborgDailyRetrievalNDVI2018}{}%
Houborg, R., and M. McCabe. 2018. Daily {Retrieval} of {NDVI} and {LAI}
at 3 m {Resolution} via the {Fusion} of {CubeSat}, {Landsat}, and {MODIS
Data}. Remote Sensing 10:890.

\leavevmode\hypertarget{ref-jayneAfricaChangingFarm2016}{}%
Jayne, T. s., J. Chamberlin, L. Traub, N. Sitko, M. Muyanga, F. K.
Yeboah, W. Anseeuw, A. Chapoto, A. Wineman, C. Nkonde, and R. Kachule.
2016. Africa's changing farm size distribution patterns: The rise of
medium-scale farms. Agricultural Economics 47:197--214.

\leavevmode\hypertarget{ref-KansangaTraditionalagriculturetransition2019}{}%
Kansanga, M., P. Andersen, D. Kpienbaareh, S. Mason-Renton, K. Atuoye,
Y. Sano, R. Antabe, and I. Luginaah. 2019. Traditional agriculture in
transition: Examining the impacts of agricultural modernization on
smallholder farming in {Ghana} under the new {Green Revolution}.
International Journal of Sustainable Development \& World Ecology
26:11--24.

\leavevmode\hypertarget{ref-kehoeNatureRiskFuture2017}{}%
Kehoe, L., A. Romero-Muñoz, L. Estes, H. Kreft, E. Polaina, and T.
Kuemmerle. 2017. Nature at risk under future agricultural expansion and
intensification. Nature Ecology and Evolution 1:1129--1135.

\leavevmode\hypertarget{ref-kernerRapidResponseCrop2020}{}%
Kerner, H., G. Tseng, I. Becker-Reshef, C. Nakalembe, B. Barker, B.
Munshell, M. Paliyam, and M. Hosseini. 2020. Rapid {Response Crop Maps}
in {Data Sparse Regions}. arXiv:2006.16866 {[}cs, eess{]}.

\leavevmode\hypertarget{ref-LesivEvaluationESACCI2017}{}%
Lesiv, M., S. Fritz, I. McCallum, N. Tsendbazar, M. Herold, J.-F. Pekel,
M. Buchhorn, B. Smets, and R. Van De Kerchove. 2017, November.
Evaluation of {ESA CCI} prototype land cover map at 20m. Monograph,
http://pure.iiasa.ac.at/id/eprint/14979/.

\leavevmode\hypertarget{ref-lesivEstimatingGlobalDistribution2019}{}%
Lesiv, M., J. C. Laso Bayas, L. See, M. Duerauer, D. Dahlia, N. Durando,
R. Hazarika, P. Kumar Sahariah, M. Vakolyuk, V. Blyshchyk, A. Bilous, A.
Perez-Hoyos, S. Gengler, R. Prestele, S. Bilous, I. ul H. Akhtar, K.
Singha, S. B. Choudhury, T. Chetri, Ž. Malek, K. Bungnamei, A. Saikia,
D. Sahariah, W. Narzary, O. Danylo, T. Sturn, M. Karner, I. McCallum, D.
Schepaschenko, E. Moltchanova, D. Fraisl, I. Moorthy, and S. Fritz.
2019. Estimating the global distribution of field size using
crowdsourcing. Global Change Biology 25:174--186.

\leavevmode\hypertarget{ref-LesivCharacterizingspatialtemporal2018}{}%
Lesiv, M., L. See, J. Laso Bayas, T. Sturn, D. Schepaschenko, M. Karner,
I. Moorthy, I. McCallum, and S. Fritz. 2018. Characterizing the spatial
and temporal availability of very high resolution satellite imagery in
{Google Earth} and {Microsoft Bing} maps as a source of reference data.
Land 7:118.

\leavevmode\hypertarget{ref-levinFarmSizeLandscape2006}{}%
Levin, G. 2006. Farm size and landscape composition in relation to
landscape changes in {Denmark}. Geografisk Tidsskrift-Danish Journal of
Geography 106:45--59.

\leavevmode\hypertarget{ref-LickerMindgaphow2010}{}%
Licker, R., M. Johnston, J. A. Foley, C. Barford, C. J. Kucharik, C.
Monfreda, and N. Ramankutty. 2010. Mind the gap: How do climate and
agricultural management explain the {`yield gap'} of croplands around
the world? Global Ecology and Biogeography 19:769--782.

\leavevmode\hypertarget{ref-liuActiveDeepLearning2017}{}%
Liu, P., H. Zhang, and K. B. Eom. 2017. Active {Deep Learning} for
{Classification} of {Hyperspectral Images}. IEEE Journal of Selected
Topics in Applied Earth Observations and Remote Sensing 10:712--724.

\leavevmode\hypertarget{ref-lobellCropYieldGaps2009}{}%
Lobell, D. B., K. G. Cassman, and C. B. Field. 2009. Crop {Yield Gaps}:
{Their Importance}, {Magnitudes}, and {Causes}. Annual Review of
Environment and Resources 34:179--204.

\leavevmode\hypertarget{ref-LowderNumberSizeDistribution2016}{}%
Lowder, S. K., J. Skoet, and T. Raney. 2016. The {Number}, {Size}, and
{Distribution} of {Farms}, {Smallholder Farms}, and {Family Farms
Worldwide}. World Development 87:16--29.

\leavevmode\hypertarget{ref-maDeepLearningRemote2019}{}%
Ma, L., Y. Liu, X. Zhang, Y. Ye, G. Yin, and B. A. Johnson. 2019. Deep
learning in remote sensing applications: {A} meta-analysis and review.
ISPRS Journal of Photogrammetry and Remote Sensing 152:166--177.

\leavevmode\hypertarget{ref-MaxwellImplementationmachinelearningclassification2018}{}%
Maxwell, A. E., T. A. Warner, and F. Fang. 2018. Implementation of
machine-learning classification in remote sensing: An applied review.
International Journal of Remote Sensing 39:2784--2817.

\leavevmode\hypertarget{ref-McCabefutureEarthobservation2017}{}%
McCabe, M. F., M. Rodell, D. E. Alsdorf, D. G. Miralles, R. Uijlenhoet,
W. Wagner, A. Lucieer, R. Houborg, N. E. C. Verhoest, T. E. Franz, J.
Shi, H. Gao, and E. F. Wood. 2017. The future of {Earth} observation in
hydrology. Hydrology and Earth System Sciences 21:3879--3914.

\leavevmode\hypertarget{ref-mellorExploringIssuesTraining2015}{}%
Mellor, A., S. Boukir, A. Haywood, and S. Jones. 2015. Exploring issues
of training data imbalance and mislabelling on random forest performance
for large area land cover classification using the ensemble margin.
ISPRS Journal of Photogrammetry and Remote Sensing 105:155--168.

\leavevmode\hypertarget{ref-morrisAwakeningAfricaSleeping2009}{}%
Morris, M., and D. Byerlee. 2009. Awakening {Africa}'s {Sleeping Giant}.
{World Bank and FAO}, {Washington, DC}.

\leavevmode\hypertarget{ref-muellerClosingYieldGaps2012}{}%
Mueller, N. D., J. S. Gerber, M. Johnston, D. K. Ray, N. Ramankutty, and
J. A. Foley. 2012. Closing yield gaps through nutrient and water
management. Nature 490:254--257.

\leavevmode\hypertarget{ref-neubertCompactWatershedPreemptive2014}{}%
Neubert, P., and P. Protzel. 2014. Compact {Watershed} and {Preemptive
SLIC}: {On Improving Trade}-offs of {Superpixel Segmentation
Algorithms}. Pages 996--1001 2014 22nd {International Conference} on
{Pattern Recognition}. {IEEE}, {Stockholm, Sweden}.

\leavevmode\hypertarget{ref-OlofssonGoodpracticesestimating2014}{}%
Olofsson, P., G. M. Foody, M. Herold, S. V. Stehman, C. E. Woodcock, and
M. A. Wulder. 2014. Good practices for estimating area and assessing
accuracy of land change. Remote Sensing of Environment 148:42--57.

\leavevmode\hypertarget{ref-OlofssonMakingbetteruse2013}{}%
Olofsson, P., G. M. Foody, S. V. Stehman, and C. E. Woodcock. 2013.
Making better use of accuracy data in land change studies: {Estimating}
accuracy and area and quantifying uncertainty using stratified
estimation. Remote Sensing of Environment 129:122--131.

\leavevmode\hypertarget{ref-Openshawmillioncorrelationcoefficients1979}{}%
Openshaw, S., and P. J. Taylor. 1979. A million or so correlation
coefficients: Three experiments on the modifiable areal unit problem.
Statistical applications in the spatial sciences 21:127--144.

\leavevmode\hypertarget{ref-panSurveyTransferLearning2010}{}%
Pan, S. J., and Q. Yang. 2010. A {Survey} on {Transfer Learning}. IEEE
Transactions on Knowledge and Data Engineering 22:1345--1359.

\leavevmode\hypertarget{ref-perselloDelineationAgriculturalFields2019}{}%
Persello, C., V. A. Tolpekin, J. R. Bergado, and R. A. de By. 2019.
Delineation of agricultural fields in smallholder farms from satellite
images using fully convolutional networks and combinatorial grouping.
Remote Sensing of Environment 231:111253.

\leavevmode\hypertarget{ref-planetteamPlanetApplicationProgram2018}{}%
PlanetTeam. 2018. Planet application program interface: In space for
life on {Earth}. {https://api.planet.com}, {San Francisco, CA}.

\leavevmode\hypertarget{ref-Pontiustotaloperatingcharacteristic2014}{}%
Pontius, R. G., and K. Si. 2014. The total operating characteristic to
measure diagnostic ability for multiple thresholds. International
Journal of Geographical Information Science 28:570--583.

\leavevmode\hypertarget{ref-qiuCirrusCloudsThat2020}{}%
Qiu, S., Z. Zhu, and C. E. Woodcock. 2020. Cirrus clouds that adversely
affect {Landsat} 8 images: {What} are they and how to detect them?
Remote Sensing of Environment 246:111884.

\leavevmode\hypertarget{ref-rodriguez-galianoAssessmentEffectivenessRandom2012}{}%
Rodriguez-Galiano, V. F., B. Ghimire, J. Rogan, M. Chica-Olmo, and J. P.
Rigol-Sanchez. 2012. An assessment of the effectiveness of a random
forest classifier for land-cover classification. ISPRS Journal of
Photogrammetry and Remote Sensing 67:93--104.

\leavevmode\hypertarget{ref-RulliFoodappropriationlarge2014}{}%
Rulli, M. C., and P. D'Odorico. 2014. Food appropriation through large
scale land acquisitions. Environmental Research Letters 9:064030.

\leavevmode\hypertarget{ref-SambergSubnationaldistributionaverage2016}{}%
Samberg, L. H., J. S. Gerber, N. Ramankutty, M. Herrero, and P. C. West.
2016. Subnational distribution of average farm size and smallholder
contributions to global food production. Environmental Research Letters
11:124010.

\leavevmode\hypertarget{ref-SearchingerHighcarbonbiodiversity2015}{}%
Searchinger, T. D., L. Estes, P. K. Thornton, T. Beringer, A.
Notenbaert, D. Rubenstein, R. Heimlich, R. Licker, and M. Herrero. 2015.
High carbon and biodiversity costs from converting {Africa}'s wet
savannahs to cropland. Nature Climate Change 5:481--486.

\leavevmode\hypertarget{ref-searchingerCreatingSustainableFood2019}{}%
Searchinger, T., R. Waite, C. Hanson, J. Ranganathan, P. Dumas, E.
Matthews, and C. Klirs. 2019. Creating a sustainable food future: {A}
menu of solutions to feed nearly 10 billion people by 2050. {Final}
report. {WRI}.

\leavevmode\hypertarget{ref-StehmanKeyissuesrigorous2019}{}%
Stehman, S. V., and G. M. Foody. 2019. Key issues in rigorous accuracy
assessment of land cover products. Remote Sensing of Environment
231:111199.

\leavevmode\hypertarget{ref-sulla-menasheHierarchicalMappingAnnual2019}{}%
Sulla-Menashe, D., J. M. Gray, S. P. Abercrombie, and M. A. Friedl.
2019. Hierarchical mapping of annual global land cover 2001 to present:
{The MODIS Collection} 6 {Land Cover} product. Remote Sensing of
Environment 222:183--194.

\leavevmode\hypertarget{ref-Tongforgottenlanduse2020}{}%
Tong, X., M. Brandt, P. Hiernaux, S. Herrmann, L. V. Rasmussen, K.
Rasmussen, F. Tian, T. Tagesson, W. Zhang, and R. Fensholt. 2020. The
forgotten land use class: {Mapping} of fallow fields across the {Sahel}
using {Sentinel}-2. Remote Sensing of Environment 239:111598.

\leavevmode\hypertarget{ref-tuiaSurveyActiveLearning2011}{}%
Tuia, D., M. Volpi, L. Copa, M. Kanevski, and J. Munoz-Mari. 2011. A
{Survey} of {Active Learning Algorithms} for {Supervised Remote Sensing
Image Classification}. IEEE Journal of Selected Topics in Signal
Processing 5:606--617.

\leavevmode\hypertarget{ref-vanvlietThereContinuingRationale2013}{}%
Van Vliet, N., O. Mertz, T. Birch-Thomsen, and B. Schmook. 2013. Is
{There} a {Continuing Rationale} for {Swidden Cultivation} in the 21st
{Century}? Human Ecology 41:1--5.

\leavevmode\hypertarget{ref-visvalingamLineGeneralisationRepeated1993}{}%
Visvalingam, M., and J. D. Whyatt. 1993. Line generalisation by repeated
elimination of points. The Cartographic Journal 30:46--51.

\leavevmode\hypertarget{ref-VonBraunSmallscalefarmersliberalized2004}{}%
Von Braun, J. 2004. Small-scale farmers in a liberalized trade
environment. Page 21.

\leavevmode\hypertarget{ref-waldnerDeepLearningEdge2020}{}%
Waldner, F., and F. I. Diakogiannis. 2020. Deep learning on edge:
{Extracting} field boundaries from satellite images with a convolutional
neural network. Remote Sensing of Environment 245:111741.

\leavevmode\hypertarget{ref-WaldnerConflationexpertcrowd2019}{}%
Waldner, F., A. Schucknecht, M. Lesiv, J. Gallego, L. See, A.
Pérez-Hoyos, R. d'Andrimont, T. de Maet, J. C. L. Bayas, S. Fritz, O.
Leo, H. Kerdiles, M. Díez, K. Van Tricht, S. Gilliams, A. Shelestov, M.
Lavreniuk, M. Simões, R. Ferraz, B. Bellón, A. Bégué, G. Hazeu, V.
Stonacek, J. Kolomaznik, J. Misurec, S. R. Verón, D. de Abelleyra, D.
Plotnikov, L. Mingyong, M. Singha, P. Patil, M. Zhang, and P. Defourny.
2019. Conflation of expert and crowd reference data to validate global
binary thematic maps. Remote Sensing of Environment 221:235--246.

\leavevmode\hypertarget{ref-WilsonRemotelySensedHighResolution2016}{}%
Wilson, A. M., and W. Jetz. 2016. Remotely {Sensed High}-{Resolution
Global Cloud Dynamics} for {Predicting Ecosystem} and {Biodiversity
Distributions}. PLOS Biology 14:e1002415.

\leavevmode\hypertarget{ref-WulderglobalLandsatarchive2016}{}%
Wulder, M. A., J. C. White, T. R. Loveland, C. E. Woodcock, A. S.
Belward, W. B. Cohen, E. A. Fosnight, J. Shaw, J. G. Masek, and D. P.
Roy. 2016. The global {Landsat} archive: {Status}, consolidation, and
direction. Remote Sensing of Environment 185:271--283.

\leavevmode\hypertarget{ref-XiongNominal30mCropland2017}{}%
Xiong, J., P. S. Thenkabail, J. C. Tilton, M. K. Gumma, P. Teluguntla,
A. Oliphant, R. G. Congalton, K. Yadav, and N. Gorelick. 2017. Nominal
30-m {Cropland Extent Map} of {Continental Africa} by {Integrating
Pixel}-{Based} and {Object}-{Based Algorithms Using Sentinel}-2 and
{Landsat}-8 {Data} on {Google Earth Engine}. Remote Sensing 9:1065.

\leavevmode\hypertarget{ref-yeReviewAccuracyAssessment2018}{}%
Ye, S., R. G. Pontius, and R. Rakshit. 2018. A review of accuracy
assessment for object-based image analysis: {From} per-pixel to
per-polygon approaches. ISPRS Journal of Photogrammetry and Remote
Sensing 141:137--147.

\leavevmode\hypertarget{ref-YizongChengMeanshiftmode1995a}{}%
Yizong Cheng. 1995. Mean shift, mode seeking, and clustering. IEEE
Transactions on Pattern Analysis and Machine Intelligence 17:790--799.

\leavevmode\hypertarget{ref-ZengHighlandcroplandexpansion2018}{}%
Zeng, Z., L. Estes, A. D. Ziegler, A. Chen, T. Searchinger, F. Hua, K.
Guan, A. Jintrawet, and E. F. Wood. 2018. Highland cropland expansion
and forest loss in {Southeast Asia} in the twenty-first century. Nature
Geoscience 11:556--562.

\leavevmode\hypertarget{ref-Zhangimagetransformcharacterize2002}{}%
Zhang, Y., B. Guindon, and J. Cihlar. 2002. An image transform to
characterize and compensate for spatial variations in thin cloud
contamination of {Landsat} images. Remote Sensing of Environment
82:173--187.

\leavevmode\hypertarget{ref-ZhuDeeplearningremote2017}{}%
Zhu, X. X., D. Tuia, L. Mou, G.-S. Xia, L. Zhang, F. Xu, and F.
Fraundorfer. 2017. Deep learning in remote sensing: A review. IEEE
Geoscience and Remote Sensing Magazine 5:8--36.

\leavevmode\hypertarget{ref-ZhuObjectbasedcloudcloud2012}{}%
Zhu, Z., and C. E. Woodcock. 2012. Object-based cloud and cloud shadow
detection in {Landsat} imagery. Remote Sensing of Environment
118:83--94.

\end{CSLReferences}

\eleft

\clearpage

\newpage

\end{document}

### 2021-03-12 inst/manuscript/si.tex
Updating raw scripts and package vignette. Added tex files

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
  a4paper]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales},
  pdfkeywords={rmarkdown, reproducible science},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Allowing for landscape pages
\usepackage{lscape}
\usepackage{graphicx}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}

% Left justification of the text: see https://www.sharelatex.com/learn/Text_alignment
% \usepackage[document]{ragged2e} % already in the latex template
\newcommand{\bleft}{\begin{flushleft}}
\newcommand{\eleft}{\end{flushleft}}
\usepackage{float}
\usepackage{titling}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{High resolution, annual maps of the characteristics of
smallholder-dominated croplands at national scales}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Supporting Information}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{4}
\tableofcontents
}
\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%

\singlespace

\bleft

\hypertarget{supplemental-methods}{%
\section{Supplemental Methods}\label{supplemental-methods}}

\hypertarget{assessing-quality-of-composites}{%
\subsection{Assessing quality of
composites}\label{assessing-quality-of-composites}}

The rubric presented in Table \ref{tab:imqcriteria} was used to assess
the quality of the seasonal image composites. The imagery was evaluated
by examining their Raster Foundry (Azavea 2020) overlays within a
\texttt{labeller} instance set up for the purpose.

\begin{table}[!h]

\caption{\label{tab:imqcriteria}Four dimensions used to assess the assess the quality
of the temporally composited image tiles, including the
criteria used to award points for scoring each dimension.}
\centering
\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{8cm}>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\toprule
Quality dimension & 3 pts & 2 pts & 1 pts & 0 pts\\
\midrule
Percent affected by residual cloud & <1\% & 1-5\% & 5-10\% & >10\%\\
Percent affected by cloud shadow & <1\% & 1-5\% & 5-10\% & >10\%\\
Number of visible scene boundaries & None & 1 & 2-3 & >3\\
Percent blurred & None & <20\% & 20-50\% & >50\%\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{mapping-platform}{%
\subsection{Mapping platform}\label{mapping-platform}}

\hypertarget{digitizing-tools}{%
\subsubsection{Digitizing tools}\label{digitizing-tools}}

To minimize the risk of topological errors, \emph{mapper}'s polygon
digitizing tools prevent drawing that results in self-intersections and
overlaps between adjacent polygons. Upon submission, the PostGIS
\href{https://www.postgis.net/docs/ST_MakeValid.html}{\texttt{ST\_MakeValid}}
function is applied to each polygon's geometry to clean remaining
topological errors upon insertion into the database.

\hypertarget{mapping-rules}{%
\subsubsection{Mapping rules}\label{mapping-rules}}

Labellers followed set of rules described in the next two sections when
interpreting and digitizing field boundaries.

\hypertarget{interpreting-and-mapping-imagery}{%
\paragraph{Interpreting and mapping
imagery}\label{interpreting-and-mapping-imagery}}

\begin{itemize}
\item
  Map the fields that are visible in the PlanetScope images, not the
  fields that you see in the basemap.
\item
  As the first choice, map the field outlines as they appear in the dry
  season PlanetScope images. If a field is not clearly visible in the
  dry season image, but is visible in the growing season image, then the
  next choice is to map it in the growing season image.
\item
  Use all available images (PlanetScope scenes and base map) to help you
  interpret what is and is not a field. Toggling back and forth between
  the different image sources can help identify and classify the field
  boundaries.
\item
  Map fields that intersect the mapping target, making sure to finish
  the entire boundary.
\end{itemize}

\hypertarget{choosing-what-to-map}{%
\paragraph{Choosing what to map}\label{choosing-what-to-map}}

\begin{itemize}
\item
  Draw polygons around fields that look like they contain crops that are
  planted and harvested during a single year (or occasionally slightly
  longer, in the case of crops such as sugarcane). These fields may look
  recently ploughed or harvested, or contain actively growing crops.
\item
  Do not draw polygons around fields that are under tree crops, such as
  orchards, woodlots, or other planted forest types.
\item
  Do not draw polygons around fields that look like they are annual
  croplands, but are overgrown and haven't been planted for a few years.
  These are possibly fallows or abandoned fields.
\item
  Do not draw polygons if you cannot tell whether a piece of land is an
  annual crop field or if it is perhaps another type of land cover
  (e.g.~bare land, or a young orchard).
\item
  Do not draw any polygons if there are no annual crop fields to map.
\end{itemize}

\hypertarget{assessing-label-accuracy}{%
\subsubsection{Assessing label
accuracy}\label{assessing-label-accuracy}}

For each accuracy assessment assignment, the labeller's maps are scored
against training reference polygons digitized by one of the map
supervisors (e.g.~Estes or Ye). A proportion of training reference sites
were of the non-cropland class and thus had no corresponding polygons.

As described in the main text, the score for a particular assignment
\emph{i} is calculated as the weighted sum of five accuracy metrics:

\[ score_i=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}\]

Where \(\beta_{0-4}\) are the weights assigned to each accuracy metric
that sum to 1. For the current production version, \(\beta_{0-4}\) were
assigned as 0.4, 0.2, 0.2, 0.1, and 0.1.

\emph{I} is ``inside the box'' accuracy which is defined as a proportion
of the area correctly mapped within a 0.005 degree resolution grid
(\emph{I\(_c\)}) over total ``inside the box'' area for this grid
(\emph{I\(_t\)}).

\[ I=\frac{I_c}{I_t}\]

\emph{O} is ``outside the box'' accuracy which refers to a proportion of
the field area correctly mapped outside the grid over total ``outside
the box'' region (\emph{O\(_t\)}, the region within the bounding box of
the workers' polygons but not within 0.005 degree resolution grid).

\[ O=\frac{O_c}{O_t}\]

\emph{F} is the fragmentation accuracy which is defined as a proportion
of matched polygon number (\emph{N\(_m\)}, the number of the workers'
polygons that has at least 50\% of its region overlapped with a
reference polygon) over total workers' polygon number (\emph{N\(_t\)}).

\[ F=\frac{N_m}{N_t}\]

\emph{E} is the average edge accuracy for all pairs of matched workers
and reference polygons; the edge accuracy for a single pair is defined
as the length of `correctly mapped edges' (\emph{L\(_c\)}, the partial
boundary of a workers' polygon that are within a three-pixel buffer
region of the matched reference polygon boundary) over the total
boundary length of its matched reference polygon (\emph{L\(_t\)}).

\[ E=\frac{L_c}{L_t}\]

\emph{C} is the categorical accuracy, i.e., a proportion of the area
that has been correctly labeled with field category within intersected
regions between worker's and reference polygons (\emph{T\(_c\)}) over
the total intersecting area (\emph{T\(_t\)}).

\[ C=\frac{T_c}{T_t}\]

\hypertarget{consensus-labelling}{%
\subsubsection{Consensus labelling}\label{consensus-labelling}}

As described in the main text, the formula used for creating a consensus
label is:

\begin{equation} \label{eq:main}
\mathrm{P(\theta|D)=\sum_{i=1}^{n}P(W_i|D)P(\theta|D, W_i)}
\end{equation}

Where \(\theta\) represents the true cover type of a pixel (field or not
field), \emph{D} is the worker's label of that field, and \emph{W\(_i\)}
is an individual worker. Looking in greater details at this equation,
the first half of the righthand side of the equation,
P(W\(_i\)\textbar D), is the ``prior'' for worker \emph{i} for the
current site based on their history of scores from accuracy assessment
assignments. The second term, P(\(\theta\)\textbar D, W\(_i\)), is the
probability that the actual class of the pixel in the current assignment
is the class that worker \emph{i} says that it is, which is either 0 or
1. There are four possible values for this second term:

\begin{equation} \label{eq:tp}
P(\theta = field|D_i = field) = 1
\end{equation} \begin{equation} \label{eq:fp}
P(\theta = no field|D_i = field) = 0
\end{equation} \begin{equation} \label{eq:tn}
P(\theta = no field|D_i = no field) = 1
\end{equation} \begin{equation} \label{eq:fn}
P(\theta = field|D_i = no field) = 0
\end{equation}

Where equations \ref{eq:tp} and \ref{eq:tn} represent true positives and
negatives, respectively, and equation \ref{eq:fp} is a false positive,
and equation \ref{eq:fn} is a false negative.

Coming back to the first term, the calculation of prior probability can
be re-expressed as:

\begin{equation} \label{eq:prior}
\mathrm{P(W_i|D) \approx P(D|W_i)P(W_i)}
\end{equation}

Where:

\begin{equation} 
\mathrm{P(D|W_i) \propto exp\left(-\frac{1}{2}BIC_i\right)}
\end{equation}

With BIC being the Bayesian information criterion:

\begin{equation}
\mathrm{BIC = ln(n)k - 2ln(\hat{L})}
\end{equation}

In which \emph{n} is the sample size, \emph{k} is the number of
parameters to estimate, and \(\hat{L}\) is the maximum likeihood
function. In this case, we are only interested in one parameter (the
label that maximizes the likelihood function), thus the BIC becomes:

\begin{equation}
\mathrm{BIC \approx -2ln(\hat{L}) = 2ln(p(D|\hat{\theta}, W))}
\end{equation}

After rearranging, we have:

\begin{equation}
\mathrm{P(D|W_i) \propto p(D|\hat{\theta}, W_i))}
\end{equation}

Which is the worker maximum likelihood, which can be computed as:

\begin{equation} \label{eq:pacc1}
\mathrm{P(\theta = field|\hat{\theta}, M_I) = P(D = field|\theta = field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tp_j}{tp_j + fn_j} \right)}
\end{equation}

\begin{equation} \label{eq:pacc2}
\mathrm{P(\theta = no field|\hat{\theta}, M_I) = P(D = no field|\theta = no field, M_I) = \frac{1}{m}\left(\sum_{j}^{m}\frac{tn_j}{tn_j + fp_j} \right)}
\end{equation}

Equations \ref{eq:pacc1} and \ref{eq:pacc2} are Producer's accuracies,
thus the maximum worker likelihood is equivalent to the worker's average
Producer's accuracy.

The other component of equation \ref{eq:prior}, P(W\(_i\)), is the
worker's average score over \emph{m} accuracy assessment assignments:\\
\begin{equation} 
\mathrm{P(W_i) \propto \frac{1}{m}\sum_{j=1}^{m}score_j}
\end{equation}

Thus equation \ref{eq:prior} uses two measures of worker accuracy, 1)
their overall average accuracy score multipled by 2) their average
Producer's accuracy to create a \emph{weight} for their individual maps
for the given site. Equation \ref{eq:main} becomes:

\begin{equation} 
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(\theta|D, W_i)}{\sum_{i=1}^{n}weight_i}}
\end{equation}

With \(\mathrm{P(\theta|D, W_i)}\) being either 0 or 1. In labelling, if
the consensus result for a pixel is: \(\mathrm{P(\theta = field|D)}\)
\textgreater{} 0.5, then we assign that pixel to the field category,
otherwise to the no field category.

After creating the consensus label, the degree of confidence in the
resulting label value is measured by Bayesian Risk:

\begin{equation} \label{eq:pixelrisk}
\mathrm{r=C(1 - L) + (1 - C)L}
\end{equation}

Where \emph{C} is the consensus probability that a given pixel is a
field (\(\mathrm{P(\theta = field|D)}\)), and \emph{L} is the consensus
label (i.e.~non-field if \emph{C} \textless{} 0.5, field if \emph{C}
\textgreater{} 0.5) for that pixel. The risk values across the entire
sample site can be processed in two ways to provide useful information
about the confidence in the consensus label for that site. The first is
a simple average of all risk values in the site, where the slope of the
risk varies depending on whether \emph{L} is a field or not a field
(Figure \ref{fig:riskcurve}). The closer to 0 the lower the risk that
the \emph{L} is mislabelled, while values approaching 1 indicate
increasing risk of mislabelling. A second approach is to calculate the
proportion of pixels having risk values that exceed a certain threshold.

\begin{figure}[!ht]

{\centering \includegraphics[width=0.7\linewidth,]{figures/si_label_risk}

}

\caption{Bayesian risk values (Y-axis) for consensus values (X axis) ranging from 0 to 1 (0 indicates no consensus that a pixel falls into the field class, 1 means complete consensus) for field and non-field consensus labels.}\label{fig:riskcurve}
\end{figure}

Accuracy assessment and consensus label generation is performed using an
\texttt{R} package \texttt{rmapaccuracy} that is internal to the
labelling platform. The package relies on spatial data classes and
operations provided by the \texttt{sf} (Pebesma 2018) and
\texttt{raster} (Hijmans 2020) packages.

\hypertarget{example}{%
\paragraph{Example}\label{example}}

To provide an example of this approach in practice, we'll imagine two
workers A and B, each with the following histories:

\begin{longtable}[]{@{}rccr@{}}
\toprule
Worker & Prod. Acc. (Field) & Prod. Acc (no field) & Score \\
\midrule
\endhead
A & 0.80 & 0.81 & 0.75 \\
B & 0.62 & 0.61 & 0.60 \\
\bottomrule
\end{longtable}

In this scenario, worker A thinks that the given pixel falls within a
field, and worker B thinks it is not a field. First, we calculate the
weights for each worker:

\[
Weight_A = score_A * PA_A(field) = P(W_A)P(D=field|W_A) = 0.8 * 0.75 = 0.6
\]

\[
Weight_B = score_B * PA_B(field) = P(W_B)P(D=no field|W_B) = 0.61 * 0.6 = 0.366
\]

And then we plug these weights into the full equation:

\[
\mathrm{P(\theta|D)=\frac{\sum_{i=1}^{n}weight_iP(L = field|D, W_i)}{\sum_{i=1}^{n}weight_i}} = \frac{0.6 * 1 + 0.366 * 0}{0.6 + 0.366} = 0.621
\]

Since 0.621 \textgreater{} 0.5, we label the particular pixel a field.

Using equation \ref{eq:pixelrisk}, the corresponding risk associated
with this particular pixel's label is thus 0.379 (i.e., 1 - 0.621).

\hypertarget{accuracy-assessment}{%
\subsubsection{Accuracy assessment}\label{accuracy-assessment}}

We designed and implemented a map accuracy assessment protocol following
procedures summarized by Stehman and Foody (2019). This entailed the
creation of a map reference sample, which first entailed designing a
sample, and then designing how the sample response would be collected.

\hypertarget{map-reference-sample-design}{%
\paragraph{Map reference sample
design}\label{map-reference-sample-design}}

We employed a stratified design for collecting the map reference sample,
using the segmented field boundaries to define the strata for
cropland/non-cropland. To create the sample, we first extracted the
centroids for each of the sample grid cells in Ghana. We then
intersected the centroid points with the field segments, assigned a
class of 1 (cropland) where points intersected a field, and 0 where they
didn't (non-cropland). We removed from this set of points all those that
corresponded to model training, validation, or training reference sites,
and extracted a random sample from both the cropland and non-cropland
points. To determine the sample size of each, we specified a desired
confidence interval using the following formula (Stehman and Foody
2019):

\[n = \frac{z^2p(1-p)}{d^2}\]

Where p is the estimated probability (or mapped class accuracy), and d
is the size of the margin of error (1/2 the confidence interval). We
selected a d value of 0.03 and assumed that the user's accuracy of the
field class would be 0.75 and that of the non-cropland class would be
0.8, returning sample sizes of 800 and 683, respectively. The
distribution of the resulting map reference sample is shown in Figure
\ref{fig:refsample}.

\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\linewidth,]{figures/si_map_reference_sample}

}

\caption{Distribution of the selected map reference sample for the cropland and non-cropland class. Classes represent the values extracted from the map strata, rather than those assigned during the classification of the sample.}\label{fig:refsample}
\end{figure}

\hypertarget{response-design}{%
\paragraph{Response design}\label{response-design}}

We collected the map reference sample on a separate instance of the
labelling platform set up for the purpose. For the sampling unit, we
selected a rectangular polygon of \textasciitilde0.1 ha
(0.0002866\(^\circ\) resolution). This polygon was centered on the
centroid of each grid cell selected for map reference sample, and was
presented as the target grid in the labelling platform. Four classes
were established for the validation: \textbf{cropland},
\textbf{non-cropland}, \textbf{uncertain but likely cropland}, and
\textbf{uncertain but likely non-cropland}. The latter two classes were
designed to capture information related to swidden dynamics, following
the rationale that uncertainty and time since last cropping are likely
to be positively correlated. This uncertainty also captures information
about the inherent difficulty of the mapping task. Samples were
collected by visually interpreting the overlays of PlanetScope
composites, following the same interpretation protocols used by the
labelling team, with the exception that the polygons placed were square
and of 0.0002866\(^\circ\) resolution.

Two supervisors (Ye and Estes) who were not involved in labelling
developed the map reference sample evaluated the PlanetScope composites
at each site to determine its class membership. The following set of
rules were followed in collecting the sample:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  When determining the class corresponding with the initial location of
  the target grid, if:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  More than half of the target falls within what appears to be a clear
  arable crop field, then classify it as \textbf{cropland};
\item
  More than half falls in what is clearly not a field, then classify it
  as \textbf{non-cropland};
\item
  More than half falls in a location where it is harder to tell whether
  it is cropland or non-cropland, determine whether it is more likely a
  crop field or not a crop field, and then assign either
  \textbf{uncertain but likely cropland}, or \textbf{uncertain but
  likely non-cropland}.
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  After determining the class, if
\end{enumerate}

\begin{itemize}
\tightlist
\item
  The target polygon is contained entirely within a single clear class,
  then simply digitize a point within the center of the target box,
  assign the appropriate class label, and complete the assignment;
\item
  Digitize a square polygon exactly aligned with the initial target,
  choose the correct class label, and then move the new polygon to the
  nearest location where it can be contained entirely by the assigned
  class.
\end{itemize}

After collecting the sample, the geometries were further refined by
converting the points (sites were the target didn't have to be shifted)
to polygons with the same 0.0002866\(^\circ\) resolution, and then the
full set of map reference polygons was used to extract the classified
values from both the cropland probability and vectorized field boundary
maps. Accuracies were assessed for the entire country, and with several
zones consisting of different groupings of AOIs or
agroecozones\footnote{sourced from https://wheregeospatial.com/agro-ecological-zones-ghana/}

\hypertarget{supplemental-results}{%
\section{Supplemental Results}\label{supplemental-results}}

\hypertarget{image-catalog-and-quality}{%
\subsection{Image catalog and quality}\label{image-catalog-and-quality}}

The total number of PlanetScope images available for download from the
Planet API (PlanetTeam 2018) per tile for each of the two major seasons
is shown in Figure \ref{fig:imagedensity}.

\begin{figure}[!ht]

{\centering \includegraphics[width=0.9\linewidth,]{figures/si_planet_coverage}

}

\caption{The density of images available through the Planet API for the five month growing season in 2018 and the subsequent three month dry season (excluding the month of November), shown in relation to the AOI boundaries.}\label{fig:imagedensity}
\end{figure}

The image quality assessment was conducted by two separate observers (L.
Song and Q. Zhang). Each observer assessed the composites for each
season at all 50 of the randomly selected tiles (see main text)
following the defined criteria (Table \ref{tab:imqcriteria}). To
calculate the score for a given tile for each season, we summed the
ranking across the four categories for each observer, rescaled the
values, and then calculated the mean tile score across both observers by
season. The mean difference between the two observers was -0.057 (sd =
0.081) and the mean absolute difference was 0.063 (sd = 0.076), thus one
observer scored tiles about 6 percent lower than the other observer, on
average.

\hypertarget{active-learning-process}{%
\subsection{Active learning process}\label{active-learning-process}}

\hypertarget{labelling}{%
\subsubsection{Labelling}\label{labelling}}

The locations of training, training reference, and validation points are
shown in Figure \ref{fig:trainval}. In AOI 3, the initial active
learning cycle resulted in low accuracy because the northern part of the
AOI shows low contrast between fields and the surrounding vegetation in
the dry season. Training the model with the initial 500 samples resulted
in large commission errors in this part of the AOI, thus we ran a second
active learning cycle that began with an inital random draw of 300
training sites confined to this AOI (blue points in Figure
\ref{fig:trainval}A)

\begin{figure}[!ht]

{\centering \includegraphics[width=0.7\linewidth,]{figures/si_training_validation_pts}

}

\caption{The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration. Grey background lines and show the boundaries and ID number of Ghana's agroecozones: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.}\label{fig:trainval}
\end{figure}

The distribution of data collection effort and scores from training
reference sites by labelling team are shown in Figure
\ref{fig:trainqual}. These results include those from re-labelling the
initial training in Cluster 2 (AOIs 7-9, 12, 15). This was done because
we discovered a small spatial offset in the original image composites,
which we corrected by reprocessing the images. We replaced the image
overlays in the instance for central-southeast (Figure
\ref{fig:trainval}A), and the labelling team mapped these sites a second
time during the production run in late 2019. The reprocessed labels were
used to initially train the model for AOIs 9, 12, and 15.

\begin{figure}[!ht]

{\centering \includegraphics[width=0.5\linewidth,]{figures/si_training_qual_n}

}

\caption{The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller. Labellers' identities are anonymized.}\label{fig:trainqual}
\end{figure}

The average Bayesian Risk of each training and validation site is shown
in Figure \ref{fig:labelrisk}, as well as the distribution of risk
values per AOI and the initial training clusters (Figure
\ref{fig:labelriskhist}). The three initial clusters include the second
mapping of Cluster 2.

\begin{figure}[!ht]

{\centering \includegraphics[width=0.8\linewidth,]{figures/si_label_risk_map}

}

\caption{The average Bayes Risk of each training and validation site in Ghana.}\label{fig:labelrisk}
\end{figure}

\begin{figure}[!ht]

{\centering \includegraphics[width=0.4\linewidth,]{figures/si_label_risk_hists}

}

\caption{The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines. The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.}\label{fig:labelriskhist}
\end{figure}

\hypertarget{validation}{%
\subsubsection{Validation}\label{validation}}

The differences in accuracy, AUC, and F1 between the active learning
process and the random retraining at each iteration for AOIs 1, 8, and
15 (Figure \ref{fig:randomvactive}). Small differences due to random
variations in the RandomForest models are the reason for the non-zero
differences at iteration 0, when both models were trained with the same
level set.\\

\begin{figure}[!ht]

{\centering \includegraphics[width=0.9\linewidth,]{figures/si_random_vs_active}

}

\caption{The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites. Postive percentages indicate superior performance by active learning, negative percentages the inverse.}\label{fig:randomvactive}
\end{figure}

\hypertarget{the-impact-of-training-data-error}{%
\subsubsection{The impact of training data
error}\label{the-impact-of-training-data-error}}

Probability images resulting from Random Forests models trained with
labels generated under three different labeling strategies are
illustrated in Figure \ref{fig:labelstrategy}. These included consensus
labels, and those individual labels that were likely to be the most and
least accurate for each training site. Label accuracy was based on the
mean score of each labeller against the training reference sites (Figure
\ref{fig:trainqual}), as assessed when labelling a given AOI. These
images were created for a single tile in AOI 1.

\begin{figure}[!ht]

{\centering \includegraphics[width=1\linewidth,]{figures/si_label_strategy_probs}

}

\caption{Cropland probability images produced by Random Forests models trained with A) consensus labels, B) the most accurate individual labels, and C) the least accurate individual labels.}\label{fig:labelstrategy}
\end{figure}

The association between the average label risk per AOI and several model
performance metrics, assessed using Spearman Rank Correlation, is shown
in Table \ref{tab:labelrisk_v_metrics}.

\begin{table}[!h]

\caption{\label{tab:labelrisk_v_metrics}Spearman's Rank correlation between the average label risk per AOI and a variety of model performance metrics.}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
  & Accuracy & AUC & F1 & Precision & Recall & FPR\\
\midrule
r & -0.824 & -0.568 & 0.456 & 0.629 & 0.206 & 0.688\\
\bottomrule
\end{tabu}
\end{table}

\hypertarget{map-accuracy}{%
\subsection{Map accuracy}\label{map-accuracy}}

\hypertarget{reference-label-uncertainty}{%
\subsubsection{Reference label
uncertainty}\label{reference-label-uncertainty}}

The size of the collected validation sample was 1207, with 1036 samples
collected and interpreted by one observer (Su Ye) and 171 collected by a
second observer (Estes). To evaluate the uncertainty inherent in
defining the map reference labels, the pair mapped 23 common sites,
showing an overall level of agreement of 87\%, and a Spearman Rank
Correlation of 0.76.

Although this overlap between observers was limited, the map reference
classification scheme provided two additional measures of uncertainty,
which were classes defined as ``unsure but most likely a field'' or
``unsure but most likely not a field,'' which the reference labeller
would choose when they could not with high confidence state whether the
site was either cropland or not cropland. Of the 487 map reference
samples that were identified as cropland, 22\% fell into the lower
confidence category, as did 11.5\% of the 720 non-cropland samples.
Across both classes, the reference labellers had lower confidence in
15.7 of sites labelled.

\hypertarget{categorical-accuracy}{%
\subsubsection{Categorical accuracy}\label{categorical-accuracy}}

In addition to the map accuracies and area estimates calculated per AOI
zone (reported in main text; see Figure \ref{fig:aoizones}A), the
accuracies were also assessed within several different groupings of
agroecozones (Figures \ref{fig:aoizones}B and Table
\ref{tab:aezaccuracy}).

\begin{figure}[!ht]

{\centering \includegraphics[width=0.9\linewidth,]{figures/si_aoi_zones}

}

\caption{Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy. In B, the ID number of Ghana's individual agroecozones is shown in grey: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.}\label{fig:aoizones}
\end{figure}

\begin{table}
\caption{Map accuracies and adjusted area estimates for the ~3 m pixel-wise classifications (based on RandomForest predictions). Results are provided for four different groupings of Ghana's 8 agroecozones zones (Zone 1 = Coastal savanna; Zone 2 = Wet evergreen, Moist evergreen, and Deciduous forest; Zone 3 = Transitional zone; Zone 4 = Guinea savanna and Sudan savanna) plus the entire country. The error matrix (with reference values in columns) provides the areal percentage for each cell, and the producer's (P), user's (U), and overall (O) map accuracies and their margins of error (in parenthesis) are provided, as well as the sample-adjusted area estimates (in km$^{2}$) and margins of error.}
\begin{center}
\includegraphics[width = 16cm]{figures/si_aez_accuracies.png}
\end{center}
\label{tab:aezaccuracy}
\end{table}

\hypertarget{field-area-and-number}{%
\subsubsection{Field area and number}\label{field-area-and-number}}

To mean, median, and distributions of the average area of segmented
field boundaries over the 100 validation sites in each AOI are compared
to the areas of the polygons digitized by the most accurate labeller
over the same sites in Figure \ref{fig:areavalidation}. The same
statistics for average number of segments versus average number of
labelled polygons across validation sites in each AOI are shown in
Figure \ref{fig:numbervalidation}.\\

\begin{figure}[!ht]

{\centering \includegraphics[width=0.9\linewidth,]{figures/si_validation_stats_fha2}

}

\caption{The distributions of the average areas (in hectares) of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average areas of field boundaries digitized (shown in red) by the most accurate worker to label each site. Vertical lines indicate the mean and median of each distribution.}\label{fig:areavalidation}
\end{figure}

\begin{figure}[!ht]

{\centering \includegraphics[width=0.9\linewidth,]{figures/si_validation_stats_fn2}

}

\caption{The distributions of the average number of segmented field boundaries (shown in blue) at the 100 validation sites per AOI, compared with the average number of digitized polygons (shown in red) by the most accurate worker to label each site. The mean and median of each distribution is shown. Vertical lines indicate the mean and median of each distribution.}\label{fig:numbervalidation}
\end{figure}

\hypertarget{cropland-characteristics}{%
\subsection{Cropland characteristics}\label{cropland-characteristics}}

\hypertarget{field-area-and-number-1}{%
\subsubsection{Field area and number}\label{field-area-and-number-1}}

To examine average field sizes, and the total number, the mean segment
size per 0.05 degree tiles was calculated and mapped, as well as the
total number of fields per tile (Figure \ref{fig:fieldsizes}).

\begin{figure}[!ht]

{\centering \includegraphics[width=0.9\linewidth,]{figures/si_fsize_n_05}

}

\caption{The A) average sizes of fields and B) total number of fields in each 0.05 degree tile, as calculated from the vectorized field boundaries.}\label{fig:fieldsizes}
\end{figure}

\hypertarget{code-and-data-availability}{%
\section{Code and Data Availability}\label{code-and-data-availability}}

The repositories containing the data, code, and manuscript source files
used to produce this paper are listed below:

\begin{itemize}
\tightlist
\item
  \href{https://github.com/agroimpacts/activemapper}{\texttt{activemapper}}:
  Contains the source code and derived data for this paper. The
  repository\footnote{github.com/agroimpacts/activemapper} will be
  opened following publication in a peer-reviewed journal.
\item
  \href{https://github.com/agroimpacts/croplandmaps_V2}{\texttt{croplandmaps}}:
  Source code for web map displaying image composites and cropland
  probability maps. The
  repository\footnote{github.com/agroimpacts/croplandmaps\_V2} is open,
  and the field boundary datasets are available for download.
\item
  \href{https:github.com/agroimpacts/imager}{\texttt{imager}}: Source
  code for the image processing software. The
  repository\footnote{github.com/agroimpacts/imager} is currently
  private, but will be opened after completing a security audit.
\item
  \href{https:github.com/agroimpacts/labeller}{\texttt{labeller}}:
  Source code for the labelling platform. The
  repository\footnote{github.com/agroimpacts/labeller} is currently
  private, but will be opened after completing a security audit.
\item
  \href{https:github.com/agroimpacts/learner}{\texttt{learner}}: Source
  code for the active learning component. The
  repository\footnote{github.com/agroimpacts/learner} is open.
\item
  \href{https:github.com/agroimpacts/segmenter}{\texttt{segmenter}}:
  Source code for the segmentation algorithm. The repository
  \footnote{github.com/agroimpacts/segmenter} is currently private, but
  will be opened after completing a security audit.
\end{itemize}

\clearpage

\hypertarget{references}{%
\section{References}\label{references}}

\singlespace

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-azaveaRasterFoundry2020}{}%
Azavea. 2020. Raster {Foundry}.
https://github.com/raster-foundry/raster-foundry.

\leavevmode\hypertarget{ref-hijmansRasterGeographicData2020}{}%
Hijmans, R. J. 2020. Raster: {Geographic} data analysis and modeling.
Manual.

\leavevmode\hypertarget{ref-pebesmaSimpleFeaturesStandardized2018}{}%
Pebesma, E. 2018. Simple features for r: {Standardized} support for
spatial vector data. The R Journal 10:439--446.

\leavevmode\hypertarget{ref-planetteamPlanetApplicationProgram2018}{}%
PlanetTeam. 2018. Planet application program interface: In space for
life on {Earth}. {https://api.planet.com}, {San Francisco, CA}.

\leavevmode\hypertarget{ref-StehmanKeyissuesrigorous2019}{}%
Stehman, S. V., and G. M. Foody. 2019. Key issues in rigorous accuracy
assessment of land cover products. Remote Sensing of Environment
231:111199.

\end{CSLReferences}

\eleft

\clearpage

\end{document}

### 2021-03-12 data-raw/data_from_instances.R

@@ -4,8 +4,6 @@
 # Heavier data go into inst/extdata
 # Intermediate (untracked) inputs go into external/..
 
-# usethis::use_data_raw("data_from_instances")
-
 # Packages
 library(aws.s3)
 library(dplyr)
@@ -181,7 +179,7 @@ save(instance_dbases, file = "external/data/aois/dbases/instance_dbases.rda")
 #------------------------------------------------------------------------------#
 # reshape instance databases into combined tables across instances
 
-load("external/data/aois/dbases/instance_dbases.rda")
+# load("external/data/aois/dbases/instance_dbases.rda")
 
 tbl_names <- names(instance_dbases$labeller)
 instance_tbls <- lapply(tbl_names, function(x) {  # x <- tbl_names[11]
@@ -206,236 +204,6 @@ instance_tbls$users <- instance_tbls$users %>% distinct(aoi, id, email) %>%
 
 # save(instance_tbls, file = "external/data/aois/dbases/instance_tables.rda")
 save(instance_tbls, file = here::here("inst/extdata/instance_tbls.rda"))
-
-#------------------------------------------------------------------------------#
-# Training and validation sites for AOIs 1-16
-
-# inputs
-load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
-mgrid <- data.table::fread(
-  system.file("extdata/ghana_grid.csv", package = "activemapper")
-)
-train_val <- readr::read_csv(here::here("inst/extdata/train_val_sites.csv"))
-
-# Initial random training draw
-initial_train_pts <- train_val %>% filter(grepl("static", file)) %>%
-  filter(!grepl("spatial|sub", file)) %>% distinct() %>%
-  rename(Cluster = file) %>%
-  mutate(Cluster = case_when(
-    grepl("cluster1", Cluster) ~ "1",
-    grepl("cluster2", Cluster) ~ "2",
-    grepl("cluster3", Cluster) ~ "3",
-  )) %>% select(-id) %>%
-  st_as_sf(coords = c("x", "y"), crs = 4326)
-
-# Active learning points
-activelearn_pts <- instance_tbls$incoming_names %>%
-  filter(aoi %in% paste0("labeller", 1:16)) %>% # distinct(aoi)
-  filter(!(aoi == "labeller3" & run == 0)) %>%
[…]

### 2021-03-12 data-raw/train_val_sites.R

@@ -1,0 +1,56 @@
+#------------------------------------------------------------------------------#
+# Training and validation sites for AOIs 1-16
+
+#
+# Collect training and validation sites for AOIs 1-16
+library(dplyr)
+library(sf)
+
+# inputs
+load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+mgrid <- data.table::fread(
+  system.file("extdata/ghana_grid.csv", package = "activemapper")
+)
+train_val <- readr::read_csv(here::here("inst/extdata/train_val_sites.csv"))
+
+# Initial random training draw
+initial_train_pts <- train_val %>% filter(grepl("static", file)) %>%
+  filter(!grepl("spatial|sub", file)) %>% distinct() %>%
+  rename(Cluster = file) %>%
+  mutate(Cluster = case_when(
+    grepl("cluster1", Cluster) ~ "1",
+    grepl("cluster2", Cluster) ~ "2",
+    grepl("cluster3", Cluster) ~ "3",
+  )) %>% select(-id) %>%
+  st_as_sf(coords = c("x", "y"), crs = 4326)
+
+# Active learning points
+activelearn_pts <- instance_tbls$incoming_names %>%
+  filter(aoi %in% paste0("labeller", 1:16)) %>% # distinct(aoi)
+  filter(!(aoi == "labeller3" & run == 0)) %>%
+  filter(processed == TRUE) %>%
+  select(-one_of("processed", "label")) %>%
+  left_join(., mgrid, by = "name") %>%
+  st_as_sf(coords = c("x", "y"), crs = 4326) %>%
+  select(aoi, name, run, iteration, usage) %>%
+  mutate(Type = case_when(
+    usage == "validate" ~ "Validation",
+    usage == "train" ~ paste("Iteration", iteration)
+  ))
+
+# pull out retraining sites for AOI3 and combine with initial train sites
+initial_train_pts <- activelearn_pts %>%
+  filter(aoi == "labeller3" & iteration == 0 & usage =="train") %>%
+  rename(Cluster = aoi) %>% mutate(Cluster = "AOI 3 retrain") %>%
+  select(Cluster, name, iteration) %>%
+  rbind(initial_train_pts, .)  # bind original to to AOI3 new points
+
+# Drop AOI3 retraining points from run 1
+activelearn_pts <- activelearn_pts %>%
+  filter(!(aoi == "labeller3" & iteration == 0 & usage == "train"))
+
+# list and put in data/
[…]

### 2021-03-19 data-raw/data_from_instances.R
Committing scripts and datasets

@@ -4,85 +4,6 @@
 # Heavier data go into inst/extdata
 # Intermediate (untracked) inputs go into external/..
 
-# Packages
-library(aws.s3)
-library(dplyr)
-
-#------------------------------------------------------------------------------#
-# Train/validation/training reference sites: collected from csvs on S3, as these
-# included initial 500 sites used to train each model that were not on the
-# instance databases
-
-# read in incoming and outgoing names files
-# get names of csv files holding sample sites
-# also new Q sites from labeller 8
-str <- glue::glue("aws s3 ls s3://activemapper/planet/")
-planet_bucket <- system(str, intern = TRUE)
-fnames <- sapply(strsplit(planet_bucket, split = " "), function(x) x[length(x)])
-sample_files <- as_tibble(fnames) %>%
-  filter(grepl("incoming_names|outgoing_names|q_sites", value)) %>%
-  filter(!grepl("congo|BF|GH|tanzania|empty|test", value)) %>% pull()
-
-# read in in loop
-calval_pts <- lapply(sample_files, function(x) {  # x <- sample_files[1]
-  print(x)
-  sites <- s3read_using(readr::read_csv, bucket = "activemapper",
-                        object = glue::glue("planet/{x}"))
-  sites %>% mutate(file = gsub("\\.csv", "", x)) %>% select(file, !!names(.))
-})
-
-# new Q sites from labeller8 (after turning on labeller8). These were sites used
-# after initial maps were completed, and created from best labels made during
-# that process
-params <- yaml::yaml.load_file(here::here("common/config.yaml"))
-dinfo <- params$mapper
-host <- "labeller8.crowdmapper.org"
-con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
-                      dbname = "Africa", user = dinfo$db_username,
-                      password = dinfo$db_password)
-newqsites <- tbl(con, "kml_data_static") %>% collect() %>%
-  mutate(file = "qsites_new", name) %>% select(file, name)
-
-# combine with calval_pts
-calval_pts <- append(calval_pts, list(newqsites))
-
-# combine into single tibble
-sapply(calval_pts, function(x) {  # x <- calval_pts[[62]]
-  if(nrow(x) > 0) {
-    # print(unique(x$file))
-    if(substr(unique(x$file), 1, 1) == "q") {
-      out <- x %>% mutate(usage = "qsite", iteration = 0)
[…]

### 2021-03-19 data-raw/train_val_sites.R

@@ -1,11 +1,87 @@
 #------------------------------------------------------------------------------#
 # Training and validation sites for AOIs 1-16
 
-#
 # Collect training and validation sites for AOIs 1-16
+library(aws.s3)
 library(dplyr)
 library(sf)
 
+#------------------------------------------------------------------------------#
+# Train/validation/training reference sites: collected from csvs on S3, as these
+# included initial 500 sites used to train each model that were not on the
+# instance databases
+
+# read in incoming and outgoing names files
+# get names of csv files holding sample sites
+# also new Q sites from labeller 8
+str <- glue::glue("aws s3 ls s3://activemapper/planet/")
+planet_bucket <- system(str, intern = TRUE)
+fnames <- sapply(strsplit(planet_bucket, split = " "), function(x) x[length(x)])
+sample_files <- as_tibble(fnames) %>%
+  filter(grepl("incoming_names|outgoing_names|q_sites", value)) %>%
+  filter(!grepl("congo|BF|GH|tanzania|empty|test", value)) %>% pull()
+
+# read in in loop
+calval_pts <- lapply(sample_files, function(x) {  # x <- sample_files[1]
+  print(x)
+  sites <- s3read_using(readr::read_csv, bucket = "activemapper",
+                        object = glue::glue("planet/{x}"))
+  sites %>% mutate(file = gsub("\\.csv", "", x)) %>% select(file, !!names(.))
+})
+
+# new Q sites from labeller8 (after turning on labeller8). These were sites used
+# after initial maps were completed, and created from best labels made during
+# that process
+params <- yaml::yaml.load_file(here::here("common/config.yaml"))
+dinfo <- params$mapper
+host <- "labeller8.crowdmapper.org"
+con <- DBI::dbConnect(RPostgreSQL::PostgreSQL(), host = host,
+                      dbname = "Africa", user = dinfo$db_username,
+                      password = dinfo$db_password)
+newqsites <- tbl(con, "kml_data_static") %>% collect() %>%
+  mutate(file = "qsites_new", name) %>% select(file, name)
+
+# combine with calval_pts
+calval_pts <- append(calval_pts, list(newqsites))
+
+# combine into single tibble
+sapply(calval_pts, function(x) {  # x <- calval_pts[[62]]
+  if(nrow(x) > 0) {
+    # print(unique(x$file))
[…]

### 2021-06-23 inst/manuscript/manuscript.Rmd
Updating raw scripts and package vignette. Added tex files
Restore from botched checkout

@@
 footerdate: no
-note: This paper is a non-peer reviewed preprint submitted to EarthArXiv. It will be periodically updated as it is revised, including information on its peer review status. 
+note: This pre-print not yet undergone peer review.
+It has been submitted to the \emph{ISPRS Journal of Photogrammetry and Remote Sensing}.
+This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
 abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands.
@@
 
-\singlespace
+<!-- \singlespace -->
 
@@
 
-\clearpage
+<!-- \clearpage -->
 

### 2021-06-28 inst/manuscript/manuscript.Rmd
Intro revised, shortened

@@
 note: This pre-print not yet undergone peer review.
-It has been submitted to the \emph{ISPRS Journal of Photogrammetry and Remote Sensing}.
+It has been submitted to the \emph{Frontiers in Artificial Intelligence}.
 This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
@@
 csl: ecology.csl
-
 ---
@@
 <!-- \vspace{2mm}\hrule -->
+
 <!-- \vspace{-20pt} -->
+
 <!-- <!-- # Abstract -->
@@
 # Introduction
+
 Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019].
@@
 @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts.
-Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time.
-Unfortunately, the data that do exist are in many places inaccurate.
+Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time, ideally on an annual basis, given the rapid pace of change [@bullockThreeDecadesLand2021; @GibbsTropicalforestswere2010].
+Unfortunately, the data that do exist are in many places inaccurate and infrequently updated.
 Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
@@
 Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018].
-Beyond cropland distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016;@levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
+Beyond cropland distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016; @levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
 
@@
 This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g.
-@ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979].
+@ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979; @boschettiAnalysisConflictOmission2004]].
 In the latter case, the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
@@
 
-These problems arising from cropland characteristics are increasingly being addressed due to technological advances.
-Recent advances in satellite technology have increased the coverage of high (<5 m) or near-high (10 m) resolution imagery with weekly to near-daily return intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
-This high spatial *and* temporal resolution addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
-On top of this, advances in cloud computing [@GorelickGoogleEarthEngine2017], the opening of image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches [@MaxwellImplementationmachinelearningclassification2018; @ZhuDeeplearningremote2017] are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale.
-These capabilities are aleady being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
-
-Despite these advances, the highest resolution (<5 m) image sources are still not used to map cropland over very large extents, presumably because they are commercial and relatively high cost to acquire, in addition to the greater computational demands.
-As such, map accuracy can still be a challenge, particular for User's accuracy, which ranges between 46 and 76% for the cropland class [e.g.
-@LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
-
-Accuracy may also suffer due to error-inducing factors that are becoming somewhat more pronounced as a consequence of these technology advances, particularly with respect to algorithms.
-Advances in machine learning are helping to greatly improve classification skill, but these algorithms generally require large, high-quality training datasets [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @MaxwellImplementationmachinelearningclassification2018].
-To satisfy this need for more training (and reference) samples, map-makers increasingly rely on visual interpretation of high resolution satellite or aerial imagery to collect training (or validation) samples [e.g.
-@XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019].
+Recent technological advances are helping to overcome these challenges.
+Chief among these are the growing numbers of satellites that collect high (\<5 m) to near-high (10 m) resolution imagery at sub-weekly intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
+This high spatial *and* temporal resolution imagery addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+On top of this, advances in cloud computing and the opening of satellite image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale [@GorelickGoogleEarthEngine2017].
+These capabilities are already being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017; [@zhangGLCFCS30Global2021a]].
+However, the potential of the highest resolution (\<5 m) imagery to map cropland over very large extents (e.g.
+country scales) has yet to be realized, presumably because these data are commercial and relatively expensive, and require significant computational resource to process.
+
+Beyond the imagery and computational gains, machine learning algorithms are rapidly advancing, providing large gains in classification performance [@maDeepLearningRemote2019; @MaxwellImplementationmachinelearningclassification2018].
+However, the ability to take advantage of these gains is often limited by newer models' need for large training datasets, which are typically unavailable, hard to collect, or contain numerous errors in developing regions [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @burkeUsingSatelliteImagery2021].
+To build sufficient samples for training cropland mapping models, as well as the reference data needed to objectively assess their performance (we refer collectively to both sample types as "labels", distinguishing between each type where needed), map-makers rely heavily on visual interpretation of high resolution satellite or aerial imagery [e.g.
+@XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019], as it is impractical and expensive to collect these data in the field over large areas, particularly on an ongoing basis.
 Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g.
 @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012].
-Visually interpreted training labels present two particular problems.
-The first is that such labels have inevitable interpretation errors that can vary substantially according to the skill of the interpreter [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019].
-These errors are typically not accounted for in reported accuracy metrics, despite the fact that they can introduce substantial error into the resulting maps and subsequent analyses [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018].
-The second problem is that visual interpretation depends on high resolution imagery (<5 m), as lower resolutions make it difficult for a human analyst to discern cropland.
-Typically the only practical source for such imagery are "virtual globe" basemaps provided by Microsoft and Google, which are composed of mosaics of various high resolution satellite and aerial images that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018].
+Visually interpreted labels present two particular problems.
+The first is that they inevitably contain errors of interpretation that can vary substantially according to the skill of the labeller, particularly over complex croplands with small field sizes [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019].
+The second problem is that visual interpretation depends on high resolution imagery (\<5 m), as field boundaries become increasingly difficult to discern as image resolution decreases.
+Typically the only practical source of high resolution imagery is "virtual globe" basemaps (e.g.
+Bing or Google Maps), which are composed of mosaics of various high resolution satellite and aerial images that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018].
 This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which is usually from a different source (e.g.
@@
 If a land change occurs in the interval between the two image sets (e.g.
-a new field was created), this can introduce error into the training data that is then passed on to the classifier.
-This source of error may be elevated in smallholder-dominated systems in the tropics, where swidden practices are common [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
-
-Improving the accuracy of cropland maps over smallholder-dominated systems thus requires an approach that meets three requirements.
-First, it should be based on high spatial and temporal resolution imagery, to be able to capture the fine grain and temporal variability of smallholders' fields.
-Second, an algorithm with suitable skill for classifying these images must be selected, and combined with the computational resources needed to process large imagery volumes.
-Third, a method for collecting large volumes of high quality training and validation data based on image interpretation is essential.
-This method should quantify and minimize the errors associated with image interpretation.
-It should also ensure that labels are collected either from the same imagery that is being classified, or from contemporanous imagery, in order to reduce errors introduced by land change processes.
-
-We describe here a cropland mapping system that follows these requirements, with an emphasis on delineating field boundaries.
-The first requirement is enabled by the recent availability of CubeSat data that provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017].
-Although these data, currently collected by Planet's CubeSat fleet, are of lower spectral depth and quality than Landsat, Sentinel, or Worldview imagery, they enable country to continental scale image mosaics to be created for multiple periods during the crop growing calendars, and capturing the intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016].
-This daily revisit capacity is also important for developing seasonal composites in cloudy regions, where satellites with longer return intervals may fail.
-Lastly, although this imagery being up to ~16 times coarser than much of Bing or Google imagery, it is sufficiently resolved for humans to discern most fields [e.g.
+a new field was created), the label, even if accurately drawn, introduces error into the classifier.
+This source of error may be elevated in tropical smallholder systems, where swidden agricultural is common [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
+Despite the high potential for label error, they are typically not accounted for during model training and map accuracy assessment, resulting not only in the potential for maps to be misused or misinterpreted in subsequent analyses,  but in missed opportunities to improve model performance [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018; @StehmanKeyissuesrigorous2019].
+
+Taking into consideration the advances and remaining limitations described above, the ability to map smallholder-dominated croplands can be further improved by 1) more fully exploiting the profusion of high frequency, high resolution imagery provided by CubeSats [@McCabefutureEarthobservation2017], and 2) by implementing methods that improve the ability to collect and minimize errors in image-interpreted labels.
+We developed a mapping approach that focuses on these two sources of improvement.
+Our approach uses PlanetScope imagery collected by Planet's fleet of Dove satellite, which provides 3-4 m resolution imagery over large areas at near daily intervals [@McCabefutureEarthobservation2017; @planetteamPlanetApplicationProgram2018], at relatively low to no cost for academic research\footnote{www.planet.com/markets/education-and-research/} and non-commercial, sustainability-oriented applications\footnote{assets.planet.com/docs/Planet\_ParticipantLicenseAgreement\_NICFI.pdf}.
+Although these data are of lower spectral depth and, in some cases, quality, than Landsat, Sentinel, or Worldview imagery, their daily revisit enables country- to continent-scale image mosaics to be created for multiple periods during a single agricultural year, even over the cloudiest forest regions where it is hard to successfully construct cloud-free composites from optical imagery with return intervals (even by a few days).
+This ability to capture intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016].
+Beyond the frequency, PlanetScope's 3.7 m resolution--although substantially coarser the 0.5-1 m available in most areas covered by virtual globes--is sufficiently resolved for humans to discern small fields under many conditions [e.g.
 see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009].
-This allows labels to be generated on the same imagery processed by the classifier, thereby addressing one of the two needs related to training data (requirement 3).
-
-The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters with a simplified feature set, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020].
-The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020].
-The training and machine learning components are combined within an active learning framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011].
-Our framework automates this interactive approach to label selection, which is effective in boosting the performance of classification models while reducing the overall number of training samples required to achieve a given level of performance [e.g.
+This allows labels to be made using the same imagery that is classified, which helps to minimize label error.
+To further reduce label noise, we developed a platform that includes rigorous label accuracy assessment protocols and a novel approach for creating consensus labels, which helps reduce mistakes made by individual labellers [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020].
+We couple the labelling platform with a machine learning model inside an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] framework, in which model training is done iteratively and interactively, wherein the model's prediction uncertainty over unlabelled areas is assessed after each training step, and a sample of the most uncertain sites is selected for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011].
+This approach helps boost the performance of the classifier while reducing the overall number of labels required to achieve a given level of performance [e.g.
 @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017].
-Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a vectorized field boundary map that provides important information on field geometry.
-
-We demonstrate this approach to map cropland in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
-
-# Methods
+Finally, an unsupervised segmentation step is applied to convert pixel-wise cropland predictions into vectorized maps of individual field boundaries.
+
+Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
+The map we develop represents a single agricultural year (2018-2019), as opposed to a multi-year composite, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size.
+In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, our findings provide insights into how maps can be improved by adhering to best practices for model training [@ElmesAccountingtrainingdata2020].
+
+# Materials and Methods
+
 ## System overview
+
 The mapping system design has four primary components focused on 1) image acquisition and compositing, 2) training and validation data collection, 3) image classification, and 4) segmentation.
@@
 ## System components
+
 ### Image compositing
+
 The image processing components of our system were designed to work with PlanetScope Analytic surface reflectance data [@planetteamPlanetApplicationProgram2018].
@@
 ```
+
 The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year.
@@
 
+```{=tex}
 \begin{equation} \label{eq:cloud}
@@
 \end{equation}
-
+```
 <!-- Equation solution from here: https://stackoverflow.com/questions/4027363/two-statements-next-to-curly-brace-in-an-equation -->
 
+```{=tex}
 \begin{equation} \label{eq:shadow}  
@@
 \end{equation}
-
+```
 Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
@@
 
+```{=tex}
 \begin{equation}
@@
 \end{equation}
-
+```
 Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
@@
 ### Labelling platform
+
 Training and reference data are collected by a custom labelling platform, which was originally designed for AWS's Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
@@
 #### Mapping workflow
+
 ##### Selecting training and reference sites
+
 The labelling process begins with the random selection of a subset (e.g.
@@
 ##### Mapping assignments
+
 Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps.
@@
 ##### Processing completed assignments
+
 All submitted polygons are cleaned to fix topological irregularities that arose during digitization (see supporting information [SI]) and stored in a PostGIS table.
@@
 
+```{=tex}
 \begin{equation} \label{eq:qaqc}
@@
 \end{equation}
-
+```
 Where *i* indicates the particular assignment, and $\beta_{0-4}$ represent varying weights that sum to 1.
@@
 \end{equation}
-
+```
 Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the label assigned to that pixel by a labeller, and $W_i$ is an individual labeller.
-P($\theta$|D) is therefore the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$|D) is an individual labeller's average score over all the accuracy assessment assignments they have completed, and P(W$\theta$|D, *W$_i$*) is the labeller's label for that pixel.
+P($\theta$\|D) is therefore the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is an individual labeller's average score over all the accuracy assessment assignments they have completed, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
 This approach therefore uses the overall accuracy of each labeller to weight their labels when combined with those made by other labellers' for the same pixel (see SI for further details).
@@
 <!-- - Training and validation data -->
+
 <!-- - Accuracy assessment -->
@@
 ### Classification pipeline
+
 Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf} cluster consisting of tens of instances, depending on the size of the AOI.
@@
 #### Feature extraction
+
 The first step is the extraction of additional features from each seasonal image composite.
@@
 \begin{center}Table 1. List of image features.\end{center}
-Feature             Window Size  N Features
-------------------- ------------ ------------
-RGB-NIR                      1X1            8 
-Mean                       11X11            8
-Standard deviation           5X5            8
------------------- ------------- ------------
+
+| Feature            | Window Size | N Features |
+|:-------------------|:------------|:-----------|
+| RGB-NIR            | 1X1         | 8          |
+| Mean               | 11X11       | 8          |
+| Standard deviation | 5X5         | 8          |
+
 Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://rasterframes.io/}.
@@
 ### The active learning loop
+
 After fitting and model evaluation a second prediction is undertaken to enable active learning.
@@
 \end{equation}
-
+```
 ```{r, echo = FALSE, eval = FALSE}
@@
 ```
+
 Where Q is the uncertainty for grid cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
@@
 ### Segmentation
+
 After the final iteration, the segmentation algorithm is invoked, which entails several steps.
@@
 ## Applying the system to map Ghana
+
 We applied the system to map Ghana's croplands, excluding areas primarily cultivated with tree crops.
@@
 ### Image quality
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ### Model gains per iteration
+
 To assess the gain in model performance due to active learning, we measured the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
@@
 ### Accounting for label error
+
 To quantify the potential impact of label error on classification results, we evaluated the performance differences between models trained with three different sets of labels: 1) those from the lowest scoring labeller to map each training site, 2) those from the highest scoring labeller, and 3) the consensus labels.
@@
 ### Accuracy assessment
+
 The model performance assessments described above (2.3.2-3) were not fully independent because they used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020].
 To independently assess the accuracy of our final map products, we followed recommended guidelines [@StehmanKeyissuesrigorous2019] to create a separate map reference sample.
-We used a stratified design, randomly assigning square polygons of ~0.1 ha extent into cropland and non-cropland strata, developed the map of segmented field boundaries.
+We used a stratified design, randomly assigning square polygons of \~0.1 ha extent into cropland and non-cropland strata, developed the map of segmented field boundaries.
 Four classes were used for the map reference sample: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
@@
 ## Assessing the characteristics of Ghanaian cropland
+
 Using the final mapped results, we calculated the estimated area of cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
@@
 # Results
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
+
 We developed maps of Ghana's cultivated croplands within an area of `r km2` km$^2$, which included portions of neighboring countries overlapped by images tiles.
@@
 ## Image catalog and quality
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ## Active learning
+
 ### Training data collection
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI.
-AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (>83%) and showed little subsequent improvement.
+AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (\>83%) and showed little subsequent improvement.
 The models for these two AOIs were thus trained with `r aoi10n` - `r aoi14n` samples.
@@
 ### Performance gains during active learning
+
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
@@
 ### The impact of training data error
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ## Map accuracy
+
 ### Categorical accuracy
+
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 <!-- ```{r, echo = FALSE, out.width="80%", fig.cap="The overall, user's, and producer's accuracies (indicated by fill color) for each of the 16 mapping zones (zones numbers are shown on maps) [This a hypothetical placeholder figure].", fig.align='center'} -->
+
 <!-- knitr::include_graphics('figures/si_mapping_blocks.png') -->
+
 <!-- ``` -->
 
+```{=tex}
 \begin{table}
@@
 \end{table}
-
+```
 ### Segmentation accuracy
+
 ```{r, echo=FALSE}
@@
 ```
+
 The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
@@
 <!-- ### Label quality and map accuracy -->
+
 <!-- ```{r, echo = FALSE, out.width="100%", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.", fig.align='center', message=FALSE} -->
+
 <!-- knitr::include_graphics('figures/figure6.png') -->
+
 <!-- # library(tidyverse) -->
+
 <!-- # library(cowplot) -->
+
 <!-- # set.seed(1) -->
+
 <!-- # acc_tb <- tibble(aoi = c(1:9, 12, 15),  -->
+
 <!-- #                  Consensus = runif(11, min = 0.75, max = 0.85), -->
+
 <!-- #                  Highest = runif(11, min = 0.72, max = 0.8), -->
+
 <!-- #                  Least = runif(11, min = 0.65, max = 0.75)) -->
+
 <!-- # acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi) -->
+
 <!-- # ggplot(acc_tb) +  -->
+
 <!-- #   geom_boxplot(aes(x = type, y = score), fill = "grey") +  -->
+
 <!-- #   xlab("Labelling strategy") + ylab("Accuracy") +  -->
+
 <!-- #   theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3)) -->
+
 <!-- ``` -->
@@
 ## Ghana's croplands
+
 Two separate maps of cropland were produced for each AOI, a per-pixel map derived from the cropland probabilities, and the vectorized map of field boundaries (Figure \ref{fig:mainmap}).
@@
 # Discussion
+
 ```{r, echo=FALSE}
@@
 ## Map accuracy and key sources of error
+
 Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
@@
 Previous work found that Random Forests achieves much better performance on small-scale croplands when trained on a much larger number of features [@Debatsgeneralizedcomputervision2016].
-However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the ~4-8 hours of runtime on 800 CPUs required for a single active learning iteration, followed by ~10-14 hours for prediction.
+However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the \~4-8 hours of runtime on 800 CPUs required for a single active learning iteration, followed by \~10-14 hours for prediction.
 This reduced the skill of the model, particularly when it came to differentiating cropland from adjacent bare patches or natural vegetation with sparse herbaceous cover, which were common in many AOIs.
@@
 Interpretation is particularly hard where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3.
-Smaller field sizes also complicate labelling, as these become increasingly indistinct in the ~4 m PlanetScope composites.
+Smaller field sizes also complicate labelling, as these become increasingly indistinct in the \~4 m PlanetScope composites.
 The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S6), and by the average score achieved by each labeller against our training reference dataset (71%; Figure S5B).
@@
 ## Error mitigation features
+
 Despite the error sources mentioned above, several features of the system proved effective in mitigating error, leading to a higher overall accuracy than would have otherwise been possible.
@@
 ```
+
 The active learning approach helped to improve overall model performance relative to randomized training site selection, in line with findings from two recent efforts [@DebatsIntegratingactivelearning2017; @hamrouniLocalGlobalTransfer2021].
@@
 ## Lingering questions
+
 Several potential issues not addressed in our assessment merit further exploration.
@@
 The temporal discrepancies mentioned above [and discussed in @ElmesAccountingtrainingdata2020] are another reason why we chose not to label on basemap imagery (in addition to restrictive usage terms), which is typically several years old [@LesivCharacterizingspatialtemporal2018].
-However, we did not assess whether the higher label accuracy one might achieve by digitizing on a <1-2 m resolution basemap would offset model errors caused by temporal mismatches.
+However, we did not assess whether the higher label accuracy one might achieve by digitizing on a \<1-2 m resolution basemap would offset model errors caused by temporal mismatches.
 
@@
 ## Next steps
+
 The maps presented here represent a version 1 product that is freely available to use, along with its underlying code (see SI for details).
@@
 ## Conclusion
+
 This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve.
@@
 # Acknowledgements
+
 The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PLACE.
@@
 # References
+
 \singlespace
@@
 <!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
+
 <!-- write.bibtex(file="knitcitations.bib") -->
+
 <!-- ``` -->
 
-<div id = "refs"></div>
+::: {#refs}
+:::
 

### 2021-06-28 inst/manuscript/manuscript.tex
Methods first section revised, including fig 1

@@
 %\begin{center}
-This paper is a non-peer reviewed preprint submitted to EarthArXiv. It
-will be periodically updated as it is revised, including information on
-its peer review status.
+This pre-print not yet undergone peer review. It has been submitted to
+the \emph{Frontiers in Artificial Intelligence}. This version will be
+updated as it is revised, and the final published version will be
+accessible through its DOI link.
 %\end{center}
@@
 characteristics of cropland (Fritz et al. 2015), and how these are
-changing over time. Unfortunately, the data that do exist are in many
-places inaccurate. Existing estimates of how much global cropland there
-is tend to vary widely, and they often disagree about where cropland is
-located (e.g. Fritz et al. 2011, 2013). Such errors can propagate in
-subsequent analyses that rely on cropland data as inputs, resulting in
-potentially misleading answers (Estes et al. 2018). Beyond cropland
-distributions, few data are available on key cropland characteristics
-such as field size, an important variable needed to estimate yield and
-other key food security variables (Carletto et al. 2015), and as an
-indicator of farm size (Levin 2006, Samberg et al. 2016), a critical
-component of rural livelihoods given increasing population densities and
-longstanding debates about the relationship between farm size and
-productivity (Feder 1985, Carletto et al. 2013, Desiere and Jolliffe
-2018).
+changing over time, ideally on an annual basis, given the rapid pace of
+change (Gibbs et al. 2010, Bullock et al. 2021). Unfortunately, the data
+that do exist are in many places inaccurate and infrequently updated.
+Existing estimates of how much global cropland there is tend to vary
+widely, and they often disagree about where cropland is located (e.g.
+Fritz et al. 2011, 2013). Such errors can propagate in subsequent
+analyses that rely on cropland data as inputs, resulting in potentially
+misleading answers (Estes et al. 2018). Beyond cropland distributions,
+few data are available on key cropland characteristics such as field
+size, an important variable needed to estimate yield and other key food
+security variables (Carletto et al. 2015), and as an indicator of farm
+size (Levin 2006, Samberg et al. 2016), a critical component of rural
+livelihoods given increasing population densities and longstanding
+debates about the relationship between farm size and productivity (Feder
+1985, Carletto et al. 2013, Desiere and Jolliffe 2018).
 
@@
 results in errors due to mixed pixels and aspects of the modifiable area
-unit problem (Openshaw and Taylor 1979). In the latter case, the pixel's
-shape may be poorly matched to that of cropland, and is too coarse to
-aggregate to approximate that shape at the characteristic scales of crop
-fields (Dark and Bram 2007, Estes et al. 2018). On top of the matter of
-scale is 1) high intra-class variability of the cropland class,
-compounded by the fact that these particular croplands can be heavily
-intergraded with surrounding vegetation (Debats et al. 2016, Estes et
-al. 2016a), and 2) the substantial temporal variability within
-croplands, both within and between seasons. These latter two aspects
-pose challenges for the classification algorithms that are applied to
-the imagery.
-
-These problems arising from cropland characteristics are increasingly
-being addressed due to technological advances. Recent advances in
-satellite technology have increased the coverage of high (\textless5 m)
-or near-high (10 m) resolution imagery with weekly to near-daily return
+unit problem (Openshaw and Taylor 1979, Boschetti et al. 2004){]}. In
+the latter case, the pixel's shape may be poorly matched to that of
+cropland, and is too coarse to aggregate to approximate that shape at
+the characteristic scales of crop fields (Dark and Bram 2007, Estes et
+al. 2018). On top of the matter of scale is 1) high intra-class
+variability of the cropland class, compounded by the fact that these
+particular croplands can be heavily intergraded with surrounding
+vegetation (Debats et al. 2016, Estes et al. 2016a), and 2) the
+substantial temporal variability within croplands, both within and
+between seasons. These latter two aspects pose challenges for the
+classification algorithms that are applied to the imagery.
+
+Recent technological advances are helping to overcome these challenges.
+Chief among these are the growing numbers of satellites that collect
+high (\textless5 m) to near-high (10 m) resolution imagery at sub-weekly
 intervals (Drusch et al. 2012, McCabe et al. 2017). This high spatial
-\emph{and} temporal resolution addresses the sensor-field scale
+\emph{and} temporal resolution imagery addresses the sensor-field scale
 mismatch, and more effectively captures the intra-seasonal dynamics of
@@
 cover types (Debats et al. 2016, Defourny et al. 2019). On top of this,
-advances in cloud computing (Gorelick et al. 2017), the opening of image
-archives (Wulder et al. 2016), and next generation machine learning
-approaches (Zhu et al. 2017, Maxwell et al. 2018) are placing large
-volumes of these moderate to near-high resolution imagery together with
-the computational and algorithmic resources necessary to classify them
-at scale. These capabilities are aleady being used to create a new
-generation of higher resolution (10-30 m) cropland and landcover maps
-for Africa and other regions (Xiong et al. 2017, Lesiv et al. 2017, ESA
-n.d.).
-
-Despite these advances, the highest resolution (\textless5 m) image
-sources are still not used to map cropland over very large extents,
-presumably because they are commercial and relatively high cost to
-acquire, in addition to the greater computational demands. As such, map
-accuracy can still be a challenge, particular for User's accuracy, which
-ranges between 46 and 76\% for the cropland class (Xiong et al. 2017,
-e.g. Lesiv et al. 2017).
-
-Accuracy may also suffer due to error-inducing factors that are becoming
-somewhat more pronounced as a consequence of these technology advances,
-particularly with respect to algorithms. Advances in machine learning
-are helping to greatly improve classification skill, but these
-algorithms generally require large, high-quality training datasets
-(Maxwell et al. 2018, Ma et al. 2019, Elmes et al. 2020). To satisfy
-this need for more training (and reference) samples, map-makers
-increasingly rely on visual interpretation of high resolution satellite
-or aerial imagery to collect training (or validation) samples (Chen et
-al. 2015, e.g. Xiong et al. 2017, Stehman and Foody 2019). Several
-web-based platforms have been developed to facilitate such efforts,
-which provide convenient and highly scalable tools for training data
-collection (Fritz et al. 2012, Estes et al. 2016a, e.g. Bey et al.
-2016). Visually interpreted training labels present two particular
-problems. The first is that such labels have inevitable interpretation
-errors that can vary substantially according to the skill of the
-interpreter (Estes et al. 2016a, Waldner et al. 2019). These errors are
-typically not accounted for in reported accuracy metrics, despite the
-fact that they can introduce substantial error into the resulting maps
-and subsequent analyses (Estes et al. 2018, Elmes et al. 2020). The
-second problem is that visual interpretation depends on high resolution
-imagery (\textless5 m), as lower resolutions make it difficult for a
-human analyst to discern cropland. Typically the only practical source
-for such imagery are ``virtual globe'' basemaps provided by Microsoft
-and Google, which are composed of mosaics of various high resolution
+advances in cloud computing and the opening of satellite image archives
+(Wulder et al. 2016), and next generation machine learning approaches
+are placing large volumes of these moderate to near-high resolution
+imagery together with the computational and algorithmic resources
+necessary to classify them at scale (Gorelick et al. 2017). These
+capabilities are already being used to create a new generation of higher
+resolution (10-30 m) cropland and landcover maps for Africa and other
+regions {[}ESA (n.d.); Lesiv et al. (2017); Xiong et al. (2017); (Zhang
+et al. 2021){]}. However, the potential of the highest resolution
+(\textless5 m) imagery to map cropland over very large extents
+(e.g.~country scales) has yet to be realized, presumably because these
+data are commercial and relatively expensive, and require significant
+computational resource to process.
+
+Beyond the imagery and computational gains, machine learning algorithms
+are rapidly advancing, providing large gains in classification
+performance (Maxwell et al. 2018, Ma et al. 2019). However, the ability
+to take advantage of these gains is often limited by newer models' need
+for large training datasets, which are typically unavailable, hard to
+collect, or contain numerous errors in developing regions (Ma et al.
+2019, Elmes et al. 2020, Burke et al. 2021). To build sufficient samples
+for training cropland mapping models, as well as the reference data
+needed to objectively assess their performance (we refer collectively to
+both sample types as ``labels,'' distinguishing between each type where
+needed), map-makers rely heavily on visual interpretation of high
+resolution satellite or aerial imagery (Chen et al. 2015, e.g. Xiong et
+al. 2017, Stehman and Foody 2019), as it is impractical and expensive to
+collect these data in the field over large areas, particularly on an
+ongoing basis. Several web-based platforms have been developed to
+facilitate such efforts, which provide convenient and highly scalable
+tools for training data collection (Fritz et al. 2012, Estes et al.
+2016a, e.g. Bey et al. 2016). Visually interpreted labels present two
+particular problems. The first is that they inevitably contain errors of
+interpretation that can vary substantially according to the skill of the
+labeller, particularly over complex croplands with small field sizes
+(Estes et al. 2016a, Waldner et al. 2019). The second problem is that
+visual interpretation depends on high resolution imagery (\textless5 m),
+as field boundaries become increasingly difficult to discern as image
+resolution decreases. Typically the only practical source of high
+resolution imagery is ``virtual globe'' basemaps (e.g.~Bing or Google
+Maps), which are composed of mosaics of various high resolution
 satellite and aerial images that typically span 3-5 years of time within
@@
 land change occurs in the interval between the two image sets (e.g.~a
-new field was created), this can introduce error into the training data
-that is then passed on to the classifier. This source of error may be
-elevated in smallholder-dominated systems in the tropics, where swidden
-practices are common (Van Vliet et al. 2013), or in rapidly developing
-agricultural frontiers (Zeng et al. 2018).
-
-Improving the accuracy of cropland maps over smallholder-dominated
-systems thus requires an approach that meets three requirements. First,
-it should be based on high spatial and temporal resolution imagery, to
-be able to capture the fine grain and temporal variability of
-smallholders' fields. Second, an algorithm with suitable skill for
-classifying these images must be selected, and combined with the
-computational resources needed to process large imagery volumes. Third,
-a method for collecting large volumes of high quality training and
-validation data based on image interpretation is essential. This method
-should quantify and minimize the errors associated with image
-interpretation. It should also ensure that labels are collected either
-from the same imagery that is being classified, or from contemporanous
-imagery, in order to reduce errors introduced by land change processes.
-
-We describe here a cropland mapping system that follows these
-requirements, with an emphasis on delineating field boundaries. The
-first requirement is enabled by the recent availability of CubeSat data
-that provides 3-4 m resolution imagery over large areas at near daily
-intervals (McCabe et al. 2017). Although these data, currently collected
-by Planet's CubeSat fleet, are of lower spectral depth and quality than
-Landsat, Sentinel, or Worldview imagery, they enable country to
-continental scale image mosaics to be created for multiple periods
-during the crop growing calendars, and capturing the intra-annual
-variability can be more important for classifying cropland than spectral
-depth (Debats et al. 2016). This daily revisit capacity is also
-important for developing seasonal composites in cloudy regions, where
-satellites with longer return intervals may fail. Lastly, although this
-imagery being up to \textasciitilde16 times coarser than much of Bing or
-Google imagery, it is sufficiently resolved for humans to discern most
-fields (Fourie 2009, e.g.~see Estes et al. 2018). This allows labels to
-be generated on the same imagery processed by the classifier, thereby
-addressing one of the two needs related to training data (requirement
-3).
-
-The second requirement is addressed by a computer vision/machine
-learning classifer that is effective for classifying smallholder
-croplands (Debats et al. 2016), re-engineered to run on high
-performance, cloud-based computing clusters with a simplified feature
-set, and following recommended practices for controlling for and
-measuring the errors that occur when training machine learning models
-for remote sensing applications (Elmes et al. 2020). The classifier is
-tightly coupled to a front-end platform for collecting label data, which
-includes rigorous accuracy assessment protocols and a novel approach for
-merging multiple maps into consensus labels, thereby minimizing image
-interpretation error (Estes et al. 2016a, Elmes et al. 2020). The
-training and machine learning components are combined within an active
-learning framework, wherein the machine learning process assesses
-classification uncertainty in unlabelled areas after a training step,
-and selects sites from areas of highest uncertainty for additional
-labelling (Cohn et al. 1994, Tuia et al. 2011). Our framework automates
-this interactive approach to label selection, which is effective in
-boosting the performance of classification models while reducing the
-overall number of training samples required to achieve a given level of
-performance (Debats et al. 2017, e.g. Hamrouni et al. 2021). Finally, an
-unsupervised segmentation step is applied to the imagery and merged with
-the pixel-wise classifications from the machine learning process,
-resulting in a vectorized field boundary map that provides important
-information on field geometry.
-
-We demonstrate this approach to map cropland in Ghana, a country where
-smallholder farming predominates, and which has a broad mix of climates
-and agricultural systems, including large areas where shifting
-agriculture is practiced (Samberg et al. 2016, Kansanga et al. 2019).
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{system-overview}{%
-\subsection{System overview}\label{system-overview}}
-
-The mapping system design has four primary components focused on 1)
-image acquisition and compositing, 2) training and validation data
-collection, 3) image classification, and 4) segmentation. The first
-component is applied as a once-off step to generate an image catalog
-covering the mapping geography. The second component provides tools for
-labelling imagery, and controls an interactive model training and
-classification (component 3) pipeline that produces a map of cropland
-probabilities for each image pixel (Figure \ref{fig:systemoverview}).
-The final segmentation step (component 4) is then initiated and applied
-to both the input image catalog and the posterior probability maps,
-resulting in vectorized field boundaries. Each system component
-comprises an individual software module designed to be implemented on
-cloud computing architecture, and available on a GitHub repository (see
-data and software availability section for details).
+new field was created), the label, even if accurately drawn, introduces
+error into the classifier. This source of error may be elevated in
+tropical smallholder systems, where swidden agricultural is common (Van
+Vliet et al. 2013), or in rapidly developing agricultural frontiers
+(Zeng et al. 2018). Despite the high potential for label error, they are
+typically not accounted for during model training and map accuracy
+assessment, resulting not only in the potential for maps to be misused
+or misinterpreted in subsequent analyses, but in missed opportunities to
+improve model performance (Estes et al. 2018, Stehman and Foody 2019,
+Elmes et al. 2020).
+
+Taking into consideration the advances and remaining limitations
+described above, the ability to map smallholder-dominated croplands can
+be further improved by 1) more fully exploiting the profusion of high
+frequency, high resolution imagery provided by CubeSats (McCabe et al.
+2017), and 2) by implementing methods that improve the ability to
+collect and minimize errors in image-interpreted labels. We developed a
+mapping approach that focuses on these two sources of improvement. Our
+approach uses PlanetScope imagery collected by Planet's fleet of Dove
+satellite, which provides 3-4 m resolution imagery over large areas at
+near daily intervals (McCabe et al. 2017, PlanetTeam 2018), at
+relatively low to no cost for academic
+research\footnote{www.planet.com/markets/education-and-research/} and
+non-commercial, sustainability-oriented
+applications\footnote{assets.planet.com/docs/Planet\_ParticipantLicenseAgreement\_NICFI.pdf}.
+Although these data are of lower spectral depth and, in some cases,
+quality, than Landsat, Sentinel, or Worldview imagery, their daily
+revisit enables country- to continent-scale image mosaics to be created
+for multiple periods during a single agricultural year, even over the
+cloudiest forest regions where it is hard to successfully construct
+cloud-free composites from optical imagery with return intervals (even
+by a few days). This ability to capture intra-annual variability can be
+more important for classifying cropland than spectral depth (Debats et
+al. 2016). Beyond the frequency, PlanetScope's 3.7 m
+resolution--although substantially coarser the 0.5-1 m available in most
+areas covered by virtual globes--is sufficiently resolved for humans to
+discern small fields under many conditions (Fourie 2009, e.g.~see Estes
+et al. 2018). This allows labels to be made using the same imagery that
+is classified, which helps to minimize label error. To further reduce
+label noise, we developed a platform that includes rigorous label
+accuracy assessment protocols and a novel approach for creating
+consensus labels, which helps reduce mistakes made by individual
+labellers (Estes et al. 2016a, Elmes et al. 2020). We couple the
+labelling platform with a machine learning model inside an active
+learning (Cohn et al. 1994, Tuia et al. 2011) framework, in which model
+training is done iteratively and interactively, wherein the model's
+prediction uncertainty over unlabelled areas is assessed after each
+training step, and a sample of the most uncertain sites is selected for
+additional labelling (Cohn et al. 1994, Tuia et al. 2011). This approach
+helps boost the performance of the classifier while reducing the overall
+number of labels required to achieve a given level of performance
+(Debats et al. 2017, e.g. Hamrouni et al. 2021). Finally, an
+unsupervised segmentation step is applied to convert pixel-wise cropland
+predictions into vectorized maps of individual field boundaries.
+
+Here we use this approach to create a high resolution, country-scale map
+of crop field boundaries in Ghana, a country where smallholder farming
+predominates, and which has a broad mix of climates and agricultural
+systems, including large areas where shifting agriculture is practiced
+(Samberg et al. 2016, Kansanga et al. 2019). The map we develop
+represents a single agricultural year (2018-2019), as opposed to a
+multi-year composite, thereby demonstrating a capacity for annual, high
+resolution maps that can be used to monitor rapidly evolving small-scale
+agricultural systems, including key characteristics such as field size.
+In addition to presenting one of the most spatially extensive
+agricultural applications of CubeSats to date, we provide a new
+technique for converting daily imagery into seasonal composites, and
+show how best practices for model training and label collection can be
+applied to improve map accuracy (Elmes et al. 2020).
+
+\hypertarget{materials-and-methods}{%
+\section{Materials and Methods}\label{materials-and-methods}}
+
+The cropland mapping approach we developed is comprised of four primary
+components (Figure \ref{fig:systemoverview}) that were developed using
+open source software, and is designed to run in a cloud computing
+environment. The first component collects daily PlanetScope imagery for
+a period of several months over a particular mapping geography, and
+converts these into cloud-free seasonal composites. The second component
+is a custom-built platform that a provides a set tools for labelling the
+PlanetScope composites, and built-in procedures for assessing and
+minimizing label error. This platform interacts with the third
+component, a machine learning process running in Apache Spark, which
+includes a routine for 1) selecting new training sites based on their
+prediction uncertainty, and 2) returning these to the labelling platform
+for labeling. This interaction, which represents active learning (Cohn
+et al. 1994, Tuia et al. 2011), repeats until a model performance
+threshold is reached. The fourth and final component is a segmentation
+process that is applied to the seasonal image composites to segment the
+entire image, and then merged with the pixel-wise probability maps
+generated by active learning to identify the segments corresponding to
+croplands, resulting in a final vectorized field boundaries. Each system
+component is available on an open repository (see data and software
+availability section for details).
 
@@
 
-\hypertarget{system-components}{%
-\subsection{System components}\label{system-components}}
-
 \hypertarget{image-compositing}{%
-\subsubsection{Image compositing}\label{image-compositing}}
+\subsection{Image compositing}\label{image-compositing}}
 
@@
 labeller's average score over all the accuracy assessment assignments
-they have completed, and P(W\(\theta\)\textbar D, \emph{W\(_i\)}) is the
+they have completed, and P(W\(\theta\)\textbar D, \emph{W}\(_i\)) is the
 labeller's label for that pixel. This approach therefore uses the
@@
 
+\leavevmode\hypertarget{ref-boschettiAnalysisConflictOmission2004}{}%
+Boschetti, L., S. P. Flasse, and P. A. Brivio. 2004. Analysis of the
+conflict between omission and commission in low spatial resolution
+dichotomic thematic products: {The Pareto Boundary}. Remote Sensing of
+Environment 91:280--292.
+
 \leavevmode\hypertarget{ref-BreimanRandomForests2001}{}%
@@
 in {East Africa}. Land 10:150.
+
+\leavevmode\hypertarget{ref-burkeUsingSatelliteImagery2021}{}%
+Burke, M., A. Driscoll, D. B. Lobell, and S. Ermon. 2021. Using
+satellite imagery to understand and promote sustainable development.
+Science 371.
 
@@
 
+\leavevmode\hypertarget{ref-GibbsTropicalforestswere2010}{}%
+Gibbs, H. K., A. S. Ruesch, F. Achard, M. K. Clayton, P. Holmgren, N.
+Ramankutty, and J. A. Foley. 2010. Tropical forests were the primary
+sources of new agricultural land in the 1980s and 1990s. Proceedings of
+the National Academy of Sciences 107:16732--16737.
+
 \leavevmode\hypertarget{ref-GorelickGoogleEarthEngine2017}{}%
@@
 
+\leavevmode\hypertarget{ref-zhangGLCFCS30Global2021a}{}%
+Zhang, X., L. Liu, X. Chen, Y. Gao, S. Xie, and J. Mi. 2021.
+{GLC}\_{FCS30}: Global land-cover product with fine classification
+system at 30 m using time-series {Landsat} imagery. Earth System Science
+Data 13:2753--2776.
+
 \leavevmode\hypertarget{ref-Zhangimagetransformcharacterize2002}{}%
@@
 
-\leavevmode\hypertarget{ref-ZhuDeeplearningremote2017}{}%
-Zhu, X. X., D. Tuia, L. Mou, G.-S. Xia, L. Zhang, F. Xu, and F.
-Fraundorfer. 2017. Deep learning in remote sensing: A review. IEEE
-Geoscience and Remote Sensing Magazine 5:8--36.
-
 \leavevmode\hypertarget{ref-ZhuObjectbasedcloudcloud2012}{}%

### 2021-06-28 inst/manuscript/figures.Rmd
Revised through image compositing methods, removed figure 2 to SI, reorganized figures.Rmd to accommodate.

@@
 load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
+
+# Ghana data
+ghana_aois <- tiles %>% pull(aoi1) %>% unique
+ghana_degrees <- tiles %>% pull(aoi2) %>% unique
 ```
@@
 
-### Figure 2: Grid system and AOIs
+### Figure 2: Labeller interface
+Taken as a screenshot.
+
+### Figure S1: Grid reference system
 
 ```{r, eval = FALSE}
-ghana_aois <- tiles %>% pull(aoi1) %>% unique
-ghana_degrees <- tiles %>% pull(aoi2) %>% unique
-
 ind <- 8  # index of selected AOI for plotting (from ghana_aois)
@@
                          labels = LETTERS[1:3], label_x = 0.05)
-ggsave(pth("figures/figure2.png"), width = 7, height = 3,
+ggsave(pth("figures/reference_system.png"), width = 7, height = 3,
        units = "in", dpi = 300)
@@
 
-### Figure 3: Labeller interface
-Taken as a screenshot.
-
-### Figure S1: Bayesian Risk
+### Figure S2: Bayesian Risk
 
@@
 
-### Figure S2: Map reference sample
+### Figure S3: Map reference sample
 ```{r, eval = FALSE}
@@
 ## Results 
-### Figure 4: Composite quality
+### Figure 3: Composite quality
 ```{r, eval=FALSE}
@@
 
-### Figure S3: PlanetScope image density
+### Figure S4: PlanetScope image density
 The density of imagery available for download from the image catalog.
@@
 
-### Figure S4: Initial training site distribution
+### Figure S5: Initial training site distribution
 Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
@@
 
-### Figure S5: Label count and quality
+### Figure S6: Label count and quality
 ```{r labelquality, eval=FALSE}
@@
 
-### Figures S6 and S7: Consensus label (Bayesian) risk
+### Figures S7 and S8: Consensus label (Bayesian) risk
 
@@
 
-### Figure S8: Random versus active sampling
+### Figure S9: Random versus active sampling
 ```{r, eval=FALSE}
@@
 
-### Figure S9: Training data error impact on probability images
+### Figure S10: Training data error impact on probability images
 ```{r, eval = FALSE}
@@
 
-### Figure S10: AOI Zones
+### Figure S11: AOI Zones
 
@@
 
-### Figures S11 and S12: Segmentation statistics (grid scale)
+### Figures S12 and S13: Segmentation statistics (grid scale)
 
@@
 
-### Figure S13: Average fields sizes
+### Figure S14: Average fields sizes
 

### 2021-06-28 inst/manuscript/manuscript.Rmd
Revised through image compositing methods, removed figure 2 to SI, reorganized figures.Rmd to accommodate.

@@
 The map we develop represents a single agricultural year (2018-2019), as opposed to a multi-year composite, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size.
-In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, our findings provide insights into how maps can be improved by adhering to best practices for model training [@ElmesAccountingtrainingdata2020].
+In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, we provide a new technique for converting daily imagery into seasonal composites, and show how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].
 
@@
 
-## System overview
-
-The mapping system design has four primary components focused on 1) image acquisition and compositing, 2) training and validation data collection, 3) image classification, and 4) segmentation.
-The first component is applied as a once-off step to generate an image catalog covering the mapping geography.
-The second component provides tools for labelling imagery, and controls an interactive model training and classification (component 3) pipeline that produces a map of cropland probabilities for each image pixel (Figure \ref{fig:systemoverview}).
-The final segmentation step (component 4) is then initiated and applied to both the input image catalog and the posterior probability maps, resulting in vectorized field boundaries.
-Each system component comprises an individual software module designed to be implemented on cloud computing architecture, and available on a GitHub repository (see data and software availability section for details).
-
-```{r systemoverview, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
+The cropland mapping approach we developed is comprised of four primary components (Figure \ref{fig:systemoverview}) that were developed using open source software, and is designed to run in a cloud computing environment.
+The first component collects daily PlanetScope imagery collected over into cloud-free seasonal composites.
+The second component is a custom-built platform that a provides a set tools for labelling the PlanetScope composites, and built-in procedures for assessing and minimizing label error.
+This platform interacts with the third component, a machine learning process running in Apache Spark, which includes a routine for 1) selecting new training sites based on their prediction uncertainty, and 2) returning these to the labelling platform for labeling.
+This interaction, which represents active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011], repeats until a model performance threshold is reached.
+The fourth and final component is a segmentation process that is applied to the seasonal image composites to segment the entire image, and then filter using the pixel-wise probability maps resulting from active learning, resulting in a final set of vectorized field boundaries.
+Each system component is available on an open repository (see data and software availability section for details).
+
+```{r systemoverview, echo = FALSE, out.width="50%", fig.cap="An overview of the primary mapping components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
 The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
@@
 
-## System components
-
-### Image compositing
-
-The image processing components of our system were designed to work with PlanetScope Analytic surface reflectance data [@planetteamPlanetApplicationProgram2018].
-PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency.
-Although these images are already pre-preprocessed and corrected for atmospheric effects, there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016].
-To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.
-
-PlanetScope imagery is accessed via the Planet API [@planetteamPlanetApplicationProgram2018], and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods.
-The imagery is collected and transferred directly to a cloud storage platform (Amazon Web Services [AWS] S3).
-
-Individual images are then transformed into analysis ready data (ARD) [@DwyerAnalysisReadyData2018], by subsetting each downloaded image into 0.05 degree tiles, regardless of cloud cover.
-Tiles are organized within a larger 1 degree resolution grid that covers the entire continent, which defines the minimum mapping area of interest (AOI; Figure \ref{fig:aois}).
-
-<!-- and then applies the usable data mask (UDM) supplied with the imagery to remove low quality and cloudy pixels.  -->
-
-```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
-knitr::include_graphics('figures/figure2.png')
-```
-
-The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year.
-Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in the same year helps to minimize differences caused by land change.
-For each pixel in each image in each ARD temporal stack for a given season, two weights are calculated:
+## Image compositing
+
+The image processing component was designed for PlanetScope Analytic surface reflectance imagery [@planetteamPlanetApplicationProgram2018], which provides three visual (red, green, blue) and near-infrared bands at 3.7 m resolution at nominal daily frequency.
+The images are provided as ortho-rectified and converted to surface reflectance, although there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016] that are not fully captured by the provided cloud masks.
+To minimize the effect of these residual errors, we developed a procedure for creating temporal composites of the primary growing and non-growing seasons within a single 12-month period.
+For Ghana, we defined the primary growing season as May through September, followed by the off (or dry) season from November or December through February.
+We chose these two seasons because prior work shows that the contrast improves cropland classifications [@Debatsgeneralizedcomputervision2016], while having the seasons captured in sequence in the same year helps minimize differences caused by land change.
+The wide time intervals we used to define each season were necessary for collecting a sufficient number of images to make high quality composites, as Ghana's cloud cover renders many scenes unusable (and thus unavailable in Planet's catalog), thus the frequency of available imagery can be substantially less than daily during several months.
+
+We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (beginning in May and ending in February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3).
+We then converted each scene into analysis ready  data [@DwyerAnalysisReadyData2018], by cropping each downloaded to the dimensions of a 0.05 degree grid (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for our composited image tiles.
+We chose this dimension for tiling because it  is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, helping to minimize edge artifacts in the composites.
+
+To create a composite tile for a given season, we calculated two weights for each pixel in each image within the ARD stack for that season:
 
@@
 ```
-Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
+Where *t* is a particular date in the time series of images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
 Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
 
-Once these two weights are calculated, the final composited pixel value for each of the four PlanetScope bands is:
+After assigning these two weights, we calculated the final composited pixel value:
 
@@
 ```
-Which is a weighted mean for each pixel for each band *B* for the particular compositing period.
-The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview}), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering is created for each composite using Raster Foundry [@azaveaRasterFoundry2020].
-The web-rendered imagery is presented within the training data platform (next section).
-
-### Labelling platform
+Which is the weighted mean for each pixel for each band *B* for the particular compositing period.
+The composited tiles were then saved as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering was created for each composite using Raster Foundry [@azaveaRasterFoundry2020], for display within the labelling platform (next section).
+
+## Labelling platform
 

### 2021-06-28 inst/manuscript/si.Rmd
Revised through image compositing methods, removed figure 2 to SI, reorganized figures.Rmd to accommodate.

@@
 # Supplemental Methods
-## Assessing quality of composites
+## Map reference system
+The map reference system used in our mapping approach had three different levels.
+At the coarsest level, Ghana was divided into 16 different mapping zones, or areas of interest (AOIS; Figure \ref{fig:aois}A), that were used to create AOI-specific mapping models.
+Each AOIs was comprised of 400 to 777 adjacent tiles.
+These tiles were used to create seasonal image composites, and were defined within a 0.05 degree grid (Figure \ref{fig:aois}B)), with each tile numbered to correspond to a larger 1 degree grid cell that it sits within (dotted lines in Figure \ref{fig:aois}B).
+Training and reference labels were created within a 0.005 degree grid that nested within each tile (Figure \ref{fig:aois}C).
+Therefore, each tile has 100 grid cells, and there are 400 tiles per 1X1 degree.
+The smallest AOI consists of a single 1X1 degree, which fall in the center of the country (AOIs 5, 8, 11, and 14).
+AOIs falling along Ghana's boundaries were created by tiles from 1X1 degree cells that straddled Ghana's border with those from the closest degree cells that were fully contained within Ghana (e.g.
+AOI 1), with the exception of AOI 16, which was comprised of tiles in three partial 1X1 degree cells along Ghana's coast.
+
+```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in this mapping approach, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
+knitr::include_graphics('figures/reference_system.png')
+```
+
+## Composites Assessing quality of composites
 The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites.

### 2021-07-01 inst/manuscript/manuscript.Rmd
Methods reorganization complete

@@
 
-Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
-The map we develop represents a single agricultural year (2018-2019), as opposed to a multi-year composite, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size.
+Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, ranging from primarily grain and vegetable crop production in the northern savannas to tree crop-dominated systems in the forested southwest, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
+The map represents a single agricultural year (2018-2019), as opposed to a multi-year epoch, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size.
 In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, we provide a new technique for converting daily imagery into seasonal composites, and show how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].
@@
 # Materials and Methods
-
-The cropland mapping approach we developed is comprised of four primary components (Figure \ref{fig:systemoverview}) that were developed using open source software, and is designed to run in a cloud computing environment.
-The first component collects daily PlanetScope imagery collected over into cloud-free seasonal composites.
-The second component is a custom-built platform that a provides a set tools for labelling the PlanetScope composites, and built-in procedures for assessing and minimizing label error.
-This platform interacts with the third component, a machine learning process running in Apache Spark, which includes a routine for 1) selecting new training sites based on their prediction uncertainty, and 2) returning these to the labelling platform for labeling.
-This interaction, which represents active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011], repeats until a model performance threshold is reached.
-The fourth and final component is a segmentation process that is applied to the seasonal image composites to segment the entire image, and then filter using the pixel-wise probability maps resulting from active learning, resulting in a final set of vectorized field boundaries.
-Each system component is available on an open repository (see data and software availability section for details).
+The mapping approach we developed is comprised of four open source components (Figure \ref{fig:systemoverview}) that are designed to run in a cloud computing environment.
+The first component collects daily PlanetScope imagery and converts them into cloud-free seasonal composites.
+The second is a custom-built platform that provides tools for labelling the composites, along with procedures to assess and minimize label error.
+This platform interacts with the third component, a machine learning process, in an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] loop, to produce a map of predicted cropland probabilities for each image pixel.
+The fourth and final component is an algorithm that segments the image composites, then filters the resulting polygons using the pixel-wise cropland predictions produced by the active learning classifier, resulting in a final set of vectorized field boundaries.
+
+<!-- Each system component is available on an open repository (see data and software availability section for details). -->
 
@@
 
+We describe each component in further detail in the following section, and how we applied them to map Ghana's annual cropland boundaries, excluding tree crops.
+
 ## Image compositing
-
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("image_quality")
+data("tile_key")
+tile_quality <- image_quality$img %>% filter(kml_type == "F") %>% 
+  group_by(name, season, worker_id) %>% 
+  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() 
+tilen <- tile_quality %>% distinct(tile) %>% count() %>% pull(n)
+```
 The image processing component was designed for PlanetScope Analytic surface reflectance imagery [@planetteamPlanetApplicationProgram2018], which provides three visual (red, green, blue) and near-infrared bands at 3.7 m resolution at nominal daily frequency.
@@
 We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (beginning in May and ending in February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3).
-We then converted each scene into analysis ready  data [@DwyerAnalysisReadyData2018], by cropping each downloaded to the dimensions of a 0.05 degree grid (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for our composited image tiles.
-We chose this dimension for tiling because it  is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, helping to minimize edge artifacts in the composites.
-
-To create a composite tile for a given season, we calculated two weights for each pixel in each image within the ARD stack for that season:
+We then converted each scene into analysis ready  data [@DwyerAnalysisReadyData2018], by cropping each to the boundaries of a 0.05$^\circ$ grid that it intersected (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for making composite image tiles.
+We chose this cell size for tiling because it is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, helping to minimize edge artifacts in the composites.
+
+To create a seasonal composite tile, we calculated two weights for the time series of each pixel within the ARD stack for a given season:
 
@@
 ```
-Where *t* is a particular date in the time series of images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
+Where *t* is a particular date in the pixel time series, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
 Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
@@
 ```
-Which is the weighted mean for each pixel for each band *B* for the particular compositing period.
-The composited tiles were then saved as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering was created for each composite using Raster Foundry [@azaveaRasterFoundry2020], for display within the labelling platform (next section).
-
-## Labelling platform
-
-Training and reference data are collected by a custom labelling platform, which was originally designed for AWS's Mechanical Turk job marketplace [@Estesplatformcrowdsourcingcreation2016].
-The basic structure of the system remains the same, but we converted it into a standalone platform that allows us to enroll and pay people directly for their labelling, and is designed to control and supervise the machine learning process.
-The platform runs on a Linux virtual machine hosted on an AWS EC2 instance and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (Raster Foundry), and a set of utilities for managing, assessing, and converting digitization work into rasterized labels for training a machine learning algorithm.
-Each instance of the platform focuses on a specific AOI (Figure \ref{fig:aois}A)
-
-The following sections provide an overview of the labelling platform's architecture.
-
-#### Mapping workflow
-
-##### Selecting training and reference sites
-
-The labelling process begins with the random selection of a subset (e.g.
-500) of cells from a 0.005 degree grid, with the selection itself potentially split into a training and validation sample, according to predetermined proportions.
-The grid, which is nested within the tiling and larger 1 degree grids (Figure \ref{fig:aois}C) defines the spatial unit for a labelling job.
-The selected cells are placed into a queue within the platform's database, and then converted into a mapping *task* that has a specified number of *assignments* (boundaries drawn by an individual labeller) that must be completed before the task is complete.
-
-##### Mapping assignments
-
-Labellers registered in the system log in to the mapping platform (built with Flask) and navigate to the OpenLayers-based field mapping interface (Figure \ref{fig:labeller}), where they are presented with a white target box representing the randomly selected grid cell, a set of digitizing tools, and different image backdrops, including true and false color renderings of the growing season and off-growing season PlanetScope composites, and several virtual globe basemaps.
-Following a set of pre-defined digitizing rules (see SI), the labeller uses the polygon drawing tool to digitize the boundaries of all crop fields intersecting the target grid cell that are visible within the PlanetScope overlays.
-To aid with interpretation, the labeller can toggle between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
-The labeller assigns each polygon a class category (e.g.
-annual cropland), and upon completing all fields submits the assignment to the database.
-In cases where the target grid cell does not contain any fields, the labeller simply submits the assignment to mark it complete.
-The labeller is then directed to the next available assignment from a different labelling task.
-
-```{r labeller, echo = FALSE, out.width="95%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
-knitr::include_graphics('figures/figure3.png')
-```
-
-##### Processing completed assignments
-
-All submitted polygons are cleaned to fix topological irregularities that arose during digitization (see supporting information [SI]) and stored in a PostGIS table.
-Each completed assignment represents one of two types of tasks: 1) accuracy assessment, or 2) model training or validation.
-For the former type, an accuracy assessment routine is invoked that executes a series of comparisons between the labeller's results and a training reference dataset, resulting in a assignment score:
-
-```{=tex}
-\begin{equation} \label{eq:qaqc}
-\mathrm{score_i}=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}
-\end{equation}
-```
-Where *i* indicates the particular assignment, and $\beta_{0-4}$ represent varying weights that sum to 1.
-*I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the labeller's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the labeller delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the labeller's thematic labels (see SI for individual formulae).
-Equation \ref{eq:qaqc} is an extension of the approach described by Estes et al.
-(2016).
-
-Over time, labellers are assessed multiple times across a range of accuracy tasks, which are selected to represent the variability of the agricultural system being mapped.
-Each labeller's score history is averaged to provide an overall accuracy measure, and this information is used for creating labels, the second task.
-
-If the labeller's completed assignment was a training/validation task, their maps remain stored in the database until the task's outstanding assignments are completed by other labellers.
-Once complete, another routine is invoked, which combines the task's completed assignments into a single consensus label using a Bayesian merging approach:
-
-```{=tex}
-\begin{equation}
-P(\theta|\mathrm{D})=\sum_{i=1}^{n}\mathrm{P}(\mathrm{W_i}|\mathrm{D})\mathrm{P}(\mathrm{\theta}|\mathrm{D}, \mathrm{W_i})
-\end{equation}
-```
-Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the label assigned to that pixel by a labeller, and $W_i$ is an individual labeller.
-P($\theta$\|D) is therefore the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is an individual labeller's average score over all the accuracy assessment assignments they have completed, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
-This approach therefore uses the overall accuracy of each labeller to weight their labels when combined with those made by other labellers' for the same pixel (see SI for further details).
-As a further measure of confidence in the final consensus label, its average Bayesian Risk can be calculated (see SI).
-This measure ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels (n = 40000) in the label, and 1 indicating complete disagreement.
-
-<!-- - Training and validation data -->
-
-<!-- - Accuracy assessment -->
-
-### Classification pipeline
-
-Upon completion of a batch of labels, the platform automatically launches an ephemeral Elastic Map Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf} cluster consisting of tens of instances, depending on the size of the AOI.
-
-#### Feature extraction
-
-The first step is the extraction of additional features from each seasonal image composite.
-Previous work showed that a large number of simple features that summarize the statistical properties of reflectance and vegetation indices in local neighborhoods are highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
-We followed this logic in this study, but were constrained to use a smaller feature space because the storage and memory requirements for our mapping geographies in this case were several orders of magnitude larger.
-For this implemention, we thus extract a set of 16 features, which are the mean and standard deviations calculated within an 11X11 and 5X5 moving window, respectively (initial tests revealed these two window sizes to be most effective), resulting in 24 overall features, including the original bands (Table 1).
-
-\begin{center}Table 1. List of image features.\end{center}
-
-| Feature            | Window Size | N Features |
-|:-------------------|:------------|:-----------|
-| RGB-NIR            | 1X1         | 8          |
-| Mean               | 11X11       | 8          |
-| Standard deviation | 5X5         | 8          |
-
-Feature extraction and the conversion of image features is handled by a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://rasterframes.io/}.
-These collectively extract subsets of imagery from the PlanetScope temporal composites, derive the features, and convert these into Apache Spark DataFrames.
-Features are extracted on the fly for each cell in the training and validation sets, a functionality enabled by storing the image composites as Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/} (COGs).
-
-<!-- \clearpage -->
-
-#### Classification
-
-Once the features from the training sites are extracted into `RasterFrames`, these are combined with their corresponding labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
-For this study, the model was trained with a balanced sample and a tree depth of 15 and total tree number of 60, which initial testing showed to provide a reasonable balance between computational time and performance.
-
-After fitting, the model is applied to the features of the model validation set, and a set of performance metrics is calculated, including binary accuracy, the F1 score (the geometric mean of precision and recall), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
-
-### The active learning loop
-
-After fitting and model evaluation a second prediction is undertaken to enable active learning.
-The feature extraction process is repeated for the rest of the mapping geography that falls outside of the training and validation sample, but applied to a subset of randomly drawn pixels from each cell in order to reduce computational demand.
-The fitted model is applied to predict the cropland probability for these selected pixels, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] is calculated for each grid cell:
+Which is the weighted mean for each pixel for each band *B* for the given season.
+
+Each composited seasonal tile was saved as a cloud-optimized geotiff, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering was created for each composite using Raster Foundry [@azaveaRasterFoundry2020], for display within the labelling platform (next section).
+
+We generated a catalog of `r nrow(tile_key) * 2` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana.
+To assess the quality of the resulting composites, we graded the tiles drawn from `r tilen` randomly selected tile grid cells using a four category score that evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g.
+because of between-date image mis-registrations).
+Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), making the highest possible score 12.
+We rescaled scores to fall between 0 and 1.
+
+## Cropland classification using active learning 
+The first step in creating a country-wide field boundary map of Ghana was to create a pixel-wise classification of cropland probabilities throughout the country.
+Given the high resolution of the imagery, we divided Ghana into 16 distinct mapping regions, or Areas of Interest (AOIs), to minimize the computational burden.
+We constructed the AOIs by grouping together tile grids into blocks representing the larger 1$^\circ$ cells used to assign tile grid identifiers (Figure S1A).
+We grouped tile grids from 1$^\circ$ degree cells that overlapped Ghana's boundaries together with tile grids from the adjacent 1$^\circ$ cell contained entirely within Ghana, with the exception was AOI 16, comprised of tile grids from the 1$^\circ$ cells intersecting Ghana's southern coast.
+The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$).
+
+We used the active learning process to develop a separate cropland classification model for each of these AOIs, based on an approach described by Debats et al (2017).
+We initiated the process by training a starter model using labels from a set of randomly selected training sites drawn from a 0.005$^\circ$ grid that was nested within the tiling grid.
+This finer grid, which we refer to as the "primary grid" for simplicity, provided the target area for creating labels (section 2.2.1), as well as the unit for distributing computing jobs (section 2.2.2).
+We then assessed the performance of the starter model against a separate set of validation labels developed for each AOI.
+The trained model was then applied to predict cropland probabilities for pixels in unlabelled primary grid cells in each AOI, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] was calculated:
 
@@
 
-Where Q is the uncertainty for grid cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
-Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus images with the lowest values of $Q$ represent sites posing the most difficulty for the classifier.
-
-After scoring with the uncertainty criterion, the top *N* most uncertain grid cells are selected and sent back to the labelling platform, which are then digitized by the labellers.
-The resulting consensus labels from the actively selected sample are added to the initial randomly selected sample, and a new cluster is launched.
-The model is retrained, assesses uncertainty across the remaining unlabelled sites, and selects the next most uncertain sites for labelling.
-This loop repeats until model performance gains against the validation set show diminishing returns.
-
-### Segmentation
-
-After the final iteration, the segmentation algorithm is invoked, which entails several steps.
-In the first step, the meanshift algorithm [@YizongChengMeanshiftmode1995a] is applied to the original bands of the dry season composite.
-A Sobel filter is then applied to the green, red, and near-infrared mean-shifted bands and the probability map, and a combined edge image is computed using the sum of these four edge images for the dry season only.
-A compact watershed algorithm [@neubertCompactWatershedPreemptive2014] is then run on the weighted edge image, with a high level of segmentation specified.
-In this case, we specified **6400** segments per tile.
-
-Third, a region adjacency graph is constructed for each image tile, in which each node represents all pixels within each polygon created in the previous step.
-The edge between two adjacent regions (polygons) is calculated as the norm of the difference between the means of normalized colors of all bands.
-Hierarchical merging is then applied, in which the most similar pairs of adjacent nodes are merged until there are no edges remaining below a predetermined threshold of 0.05.
-
-In the fourth step, the merged polygons are overlaid with the posterior probabilities resulting from the final active learning loop, and polygons in which the average posterior probability is greater than a predetermined threshold (here 0.5, but could vary locally) are retained as field polygons.
-
-In the final step, the retained polygons are refined by removing holes and smoothing their boundaries using the Visvalingam algorithm [@visvalingamLineGeneralisationRepeated1993].
-Neighboring polygons that overlap along tile boundaries are then merged.
-
-To assess the accuracy of the final segmented boundaries, we used a two-step approach.
-First, we assessed the overall thematic accuracy of the resulting classification against our map reference data.
-Second, to assess the quality of the segmentation, we compared the mean area and relative frequencies of the segmented polygons within different size classes against the same metrics derived from the digitized fields of the most accurate worker to create the given map.
-We selected this relatively simple procedure, as opposed to more complex measures of object accuracy [@yeReviewAccuracyAssessment2018], because, on the one hand, both the automated segmentation algorithm and labellers are cueing in on the same features--abrupt, physically detectable breaks within the imagery.
-On the other hand, no matter how well the intepreted/segmented boundaries align with the boundaries of fields in the imagery, it is logistically difficult to evaluate performance against real-world boundaries as the spectral distinction of field boundaries will vary across different crop types and land use arrangements.
-
-## Applying the system to map Ghana
-
-We applied the system to map Ghana's croplands, excluding areas primarily cultivated with tree crops.
-Ghana has several distinct agricultural regions, ranging from the primarily grain and vegetable crop producing regions in the northern savannas to tree crop-dominated system in the forested southwest, where cocoa and oil palm are among the dominant crops.
-For these latter regions, we did not attempt to classify tree crops, and instead mapped clearings that potentially contain field crops or newly felled or recently replanted tree crops.
-We made this decision because PlanetScope's resolution is not high enough for labellers to distinguish many tree crops from surrounding forest, and the boundaries of many tree crops (e.g.
-cocoa) are often not visible.
-
-To create the cropland maps, we divided the country into 16 distinct AOIs, which were developed by grouping together each one degree cell fully contained within Ghana with the tiles belonging to any adjacent degree cell that overlapped neighboring countries (Figure \ref{fig:aois}A).
-The exception was AOI 16, which consisted of the four degree cells intersecting Ghana's southern coast.
-The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$), A separate active learning and segmentation process was run for each of these AOIs.
-
-To collect the initial randomized samples for model training, we grouped the AOIs into three clusters: a northern cluster comprising the 6 northernmost AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16).
-Within each cluster, we randomly selected and labelled 500 grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort.
-In addition to these samples, we randomly selected and labelled 100 grid cells within each AOI to provide a validation sample.
-
-After collecting the initial training and validation samples, we trained a starter model for each cluster and applied it to each of the block's AOI.
-For each iteration, 100 samples were actively selected within each AOI, and added to the training pool.
-
-During the collection of training and validation samples, labellers were tasked to only digitize active or recently active crop fields, avoiding tree crops, and fallow or potentially abandoned fields (see SI for the digitizing rules).
-
-To evaluate the performance of the system, we performed several analyses described in sections 2.3.1-4.
-
-### Image quality
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("image_quality")
-tile_quality <- image_quality$img %>% filter(kml_type == "F") %>% 
-  group_by(name, season, worker_id) %>% 
-  summarize(tile = min(tile), Score = sum(score) / 12) %>% ungroup() 
-tilen <- tile_quality %>% distinct(tile) %>% count() %>% pull(n)
-```
-
-We evaluated the overall quality of the resulting seasonal image composites by assessing a random selection of `r tilen` tiles.
-We graded both seasonal composites for each tile using a four category quality score, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g.
-because of between-date image mis-registrations).
-Each category was qualitatively ranked from 0-3, with 0 being the lowest quality, and 3 the highest (see SI for complete protocol), making the highest possible score 12.
-We rescaled scores to fall between 0 and 1.
-
-### Model gains per iteration
-
-To assess the gain in model performance due to active learning, we measured the change in accuracy, F1, and AUC (see 2.2.3.2) between each iteration and between the first and last iterations for each AOI.
-
-To evaluate whether active learning improved model performance relative to a purely random approach to selecting new training sites, we ran additional tests within a subset of AOIs (1, 8, and 15).
-We first randomly selected and labelled 300-400 sites in each AOI.
-We then progressively added 100 of the randomly selected samples to the relevant training pool and retrained the model, repeating the process so that the number of iterations and samples matched those from the active learning process.
-We then compared the difference in accuracy, AUC, and F1 between the randomly trained models and those trained with active learning [@DebatsIntegratingactivelearning2017].
-
-### Accounting for label error
+Where Q is the uncertainty for each unlabelled primary grid cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
+Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus the lowest values of $Q$ represent locations posing the most difficulty for the classifier.
+
+We ranked the unlabelled primary grid cells in from least to most certain, and randomly selected a subset cells from the top 30% of the ranking (to minimize the risk of spatial autocorrelation) and sent these back to the labelling platform.
+After these new sites were labelled, they were added to the starter pool of labels, the model was retrained with the larger training set, its performance and prediction uncertainty was reassessed, and a new sample of the most uncertain primary grid cells was sent for labelling.
+This loop repeated until model performance gains saturated or reached a threshold, and then a final map of predicted cropland probabilities was made for each AOI.
+
+In the next two sections, we describe the labelling and machine learning components of the active learning process in more detail
+
+### Labelling
+To collect the initial randomized samples for model training, we grouped the AOIs (Figure S1A) into three clusters based on approximate agro-ecological similarity: the 6 northernmost savanna-zone AOIs (Cluster 1), a central to southeastern cluster (Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the forest zone AOIs (10, 11, 13, 14, 16).
+Within each cluster, we randomly selected and labelled 500 primary grid cells, which provided relatively large initial training samples for these agro-ecologically similar regions, while helping to minimize the overall amount of labelling effort.
+To create validation samples, we randomly selected and labelled 100 primary grid cells per AOI, and a further 100 cells were labelled in each AOI during each active learning iteration.
+
+In addition to training and validation labels, we also collected training reference labels and map reference labels [@ElmesAccountingtrainingdata2020].
+The former were a selection of primary grid cells selected to represent the range of cropland types and densities in Ghana, which were labelled by expert analysts (the lead researchers on this project).
+We were used these to assess the performance of the individual labellers collecting training and validation labels.
+Map reference labels were collected and used to assess the accuracy of the final map (see Section X).
+
+We collected labels using a custom-built platform that we adapted from an earlier prototype used for crowdsourced labelling [@Estesplatformcrowdsourcingcreation2016].
+We enhanced this platform by making several major additions, including an independent backend that allowed us to recruit and manage our own labelling teams, improved procedures for assessing and improving label accuracy, and processes for automating the machine learning component.
+The platform runs on a cloud-hosted Linux virtual server (AWS EC2) and is comprised of a database (PostGIS/Postgres), a mapping interface with a full set of labelling tools (OpenLayers 3), an image server (Raster Foundry), and a set of utilities for managing, assessing, and converting digitized field boundaries into rasterized labels.
+We created a separate labelling instance to manage the active learning process for each AOI.
+To create training and validation labels, labellers (the co-authors of this paper) would log into  the website (built with Flask) for a particular AOI and navigate to the mapping interface (Figure \ref{fig:labeller}), where they would be presented with a white target box representing a primary grid cell to label, a set of digitizing tools, and several different sources of imagery.
+These included true and false color renderings of the growing season and dry season PlanetScope composites, and several virtual globe basemaps.
+They then used the polygon drawing tool to digitize the boundaries of all crop fields visible within the PlanetScope overlays that intersect the target grid cell.
+For this project, labellers were instructed to digitize active or recently active crop fields, avoiding tree crops, and fallow or potentially abandoned fields (see SI for digitizing rules) To aid with interpretation, labellers toggled between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
+The labeller assigned each digitized polygon a class category (e.g.
+annual cropland), and upon completing all fields saved the fields to the database, and were then presented with the next target to label.
+If the target grid cell did not contain any fields, the labeller simply pressed save to go to the next cell.
+
+```{r labeller, echo = FALSE, out.width="95%", fig.cap="An overview of the labelling platform's interface", fig.align='center'}
+knitr::include_graphics('figures/figure2.png')
+```
+
+The flow of labelling targets presented to each worker was determined by the platform's built-in scheduler.
+Each primary grid cell selected for labeling was placed into a queue within the platform's database, and converted into a labelling *task* with a specified number of *assignments* (the boundaries drawn by an individual labeller) that had to be completed in order to finish the task.
+There were two types of tasks, accuracy assessment or model training/validation, with the assignments for each indistinguishable to labellers.
+Upon completing an accuracy assessment assignment, the platform invoked an label scoring algorithm that compared the labeller's digitized boundaries against a set of training reference polygons, resulting in an assignment quality score:
+
+```{=tex}
+\begin{equation} \label{eq:qaqc}
+\mathrm{score_i}=\beta_0\mathrm{I}+\beta_1\mathrm{O}+\beta_2\mathrm{F}+\beta_3\mathrm{E}+\beta_4\mathrm{C}
+\end{equation}
+```
+Where *i* indicates the particular assignment, and $\beta_{0-4}$ represent varying weights that sum to 1.
+*I* refers to "inside the box" accuracy, *O* is the accuracy of those portions of the labeller's polygons extending beyond the target grid boundaries, *F* is fragmentation accuracy, a measure of how many individual polygons the labeller delineated relative to the reference, *E* measures how closely each polygon's boundary matched its corresponding reference polygon boundary, and *C* assesses the accuracy of the labeller's thematic labels (see SI for individual formulae).
+Equation \ref{eq:qaqc} is an extension of the approach described by Estes et al.
+(2016).
+
+We configured the platform's scheduler to present workers with accuracy assessment assignmetns at a rate of 1 for every 5 assignments mapped.
+This generated a history of accuracy assessment scores that we used to assess label quality and minimize label error.
+
+For training and validation, where there was no reference data to assess label accuracy, we set each task to have four assignments, i.e.
+each was completed by four separate labellers.
+When all four assignments were complete, a consensus labelling routine was invoked to combine the four sets of labels into a single consensus label using a Bayesian merging approach:
+
+```{=tex}
+\begin{equation}
+P(\theta|\mathrm{D})=\sum_{i=1}^{n}\mathrm{P}(\mathrm{W_i}|\mathrm{D})\mathrm{P}(\mathrm{\theta}|\mathrm{D}, \mathrm{W_i})
+\end{equation}
+```
+Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the label assigned to that pixel by a labeller, and $W_i$ is an individual labeller.
+P($\theta$\|D) is the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is the average score of the accuracy assessment assignments an individual labeller completed within the AOI, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
+This approach therefore used the overall accuracy of each labeller to weight their label for a given pixel when combining it with those made by other labellers' for the same pixel (see SI for further details).
+Each pixel in the target grid cell was merged using this approach (n = 40000), which helps to minimize individual labellers' errors.
+We estimated a confidence for each consensus label by  calculating its Bayesian Risk (see SI), which ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels in the label, and 1 indicating complete disagreement.
+
+### Machine learning
+Upon completing each batch of labels, the platform automatically launched a machine learning cluster (Elastic Map Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf}) comprised of several hundred to a thousand CPUs, depending on the size of the AOI.
+
+The first step in the process was to derive a set of features from the image composites.
+Previous work showed that a large number of simple features summarizing image reflectance and vegetation indices within local neighborhoods were highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
+We followed that logic in this study, but used a smaller feature space because the storage and memory required for our mapping geographies were several orders of magnitude larger than in the previous study.
+We therefore extracted a set of 8 features from each seasonal composite, which were the mean calculated within an 11X11 moving applied to each band, and the standard deviation within a 5X5 moving window (initial tests revealed these two window sizes to be most effective).
+This provided an overall set of 24 features, including the unmodified bands of both composites (Table 1).
+
+\begin{center}Table 1. List of image features.\end{center}
+
+| Feature            | Window Size | N Features |
+|:-------------------|:------------|:-----------|
+| RGB-NIR            | 1X1         | 8          |
+| Mean               | 11X11       | 8          |
+| Standard deviation | 5X5         | 8          |
+
+We used a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://rasterframes.io/} to derive the features on the fly (which was enabled by converting the composites to Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/}) and convert them into Apache Spark DataFrames.
+The extracted features were combined with their corresponding training and validation labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
+We trained the model with a balanced sample and a tree depth of 15 and total tree number of 60, which initial testing showed to provide a reasonable balance between computational time and performance.
+
+### Classifier performance
+To assess performance of the Random Forests classifier, we used the validation sample to calculate binary accuracy, the F1 score (the geometric mean of precision and recall), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
+We calculated these measures each time the model was retrained for a given AOI, in order to assess the change in classifier performance with each active learning iteration.
+
+To evaluate whether active learning improved model performance relative to randomized label selection, we ran an additional test within three AOIs (1, 8, and 15), in which we retrained the model with 100 randomly selected labels for each iteration.
+We then compared the differences in accuracy, AUC, and F1 between the actively and randomly trained models  [@DebatsIntegratingactivelearning2017].
 
@@
 best practice) standards to account for training data errors [@ElmesAccountingtrainingdata2020].
+
+## Segmentation
+Upon completion of the active learning process, we deployed a five-step algorithm to create a segmented map of field boundaries.
+In the first step, we identified edge features within the imagery.
+To do this, we applied the meanshift algorithm [@YizongChengMeanshiftmode1995a] to each dry-season composite tile, and then passed a Sobel filter over the mean-shifted green, red, and near-infrared bands, and the corresponding map of predicted cropland probabilities.
+We then summed the four resulting edge images to produce a combined edge image.
+
+In the second step, we used a compact watershed algorithm [@neubertCompactWatershedPreemptive2014] to segment the weighted edge image, specifying a high number of segments (6,400) segments per tile, so that the mean segment size (<0.5 ha) was finer than the expected mean field size (1-2 ha).
+
+In the third step, we hierarchically merged the resulting polygons.
+We first constructed a region adjacency graph for each tile, with each node representing all image pixels within each polygon.
+The edge between two adjacent regions (polygons) was calculated as the difference between the means of the normalized colors of all bands.
+We then merged the most similar pairs of adjacent nodes until there were no edges remaining below the predetermined threshold of 0.05.
+
+In the fourth step, we overlaid the merged polygons with the cropland probability images, and polygons in which the mean probability was greater than 0.5 were retained as crop fields.
+
+In the fifth and final step, we refined the crop field polygons, by removing holes and smoothing boundaries using the Visvalingam algorithm [@visvalingamLineGeneralisationRepeated1993].
+We then merged neighboring polygons that overlapped along tile boundaries.
+
+The resulting map represents dry season crop field boundaries, as we did not segment growing season images.
+We made this choice because labels were primarily drawn on dry season composites, when boundaries were typically more visible.
 
@@
 
-### Accuracy assessment
-
-The model performance assessments described above (2.3.2-3) were not fully independent because they used the same validation sites over multiple iterations [@ElmesAccountingtrainingdata2020].
-To independently assess the accuracy of our final map products, we followed recommended guidelines [@StehmanKeyissuesrigorous2019] to create a separate map reference sample.
-We used a stratified design, randomly assigning square polygons of \~0.1 ha extent into cropland and non-cropland strata, developed the map of segmented field boundaries.
-Four classes were used for the map reference sample: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
-The latter two classes were used to provide insight into the degree of uncertainty in the map reference sample.
-For efficiency, two separate supervisors evaluated separate portions of the reference sample, but both jointly assessed a small subset of the sample.
-We calculated their level of agreement on this subset to provide an additional assessment of uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019].
-The SI contains further details on the design and collection of the map reference sample.
-
-The map reference polygons were then intersected with both the probability images and the segmented field boundaries, and confusion matrixes between the map reference labels and the extracted map classes were constructed to assess the categorical accuracy of each map product.
-We calculated the overall accuracy for each map, as well as the class-wise User's and Producer's accuracy, as well as the 95% confidence intervals for each accuracy measure [@OlofssonGoodpracticesestimating2014; @OlofssonMakingbetteruse2013; @StehmanKeyissuesrigorous2019].
-
-To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against those of the workers' digitized polygons at map validation sites.
+## Map assessment
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# reference labels, for total count and uncertainty of class categories
+ref_labels <- st_read(
+  system.file("extdata/map_reference_labels.geojson", 
+              package = "activemapper"), quiet = TRUE
+)
+class_ct <- as_tibble(ref_labels) %>% group_by(category) %>% count()
+ref_ct <- as_tibble(ref_labels) %>% group_by(worker_id) %>% count() %>% pull(n)
+
+crp_ct <- class_ct %>% filter(category %in% c("annualcropland", "unsure1"))
+noncrp_ct <- class_ct %>% filter(category %in% c("noncropland", "unsure2"))
+
+# observer agreement/correlation
+ref_labels_joint <- st_read(
+  system.file("extdata/map_reference_labels_joint.geojson", 
+              package = "activemapper"), quiet = TRUE
+)
+joint_classes <- ref_labels_joint %>% as_tibble() %>% 
+  select(name, category, worker_id) %>% 
+  arrange(name) %>% 
+  mutate(class = case_when(
+    category == "annualcropland" | category == "unsure1" ~ 1, 
+    category == "noncropland" | category == "unsure2" ~ 0 
+  ))
+class_mat <- joint_classes %>% dplyr::select(name, class, worker_id) %>% 
+  tidyr::pivot_wider(names_from = worker_id, values_from = class) %>% 
+  dplyr::select(-name) %>% as.matrix()
+
+usertab <- table("User1" = class_mat[, 1], "User2" = class_mat[, 2])
+
+agreement <- round(sum(diag(usertab)) / sum(usertab) * 100, 1)
+observer_r <- cor.test(class_mat[, 1], class_mat[, 2], method = "spearman")
+
+```
+
+We followed recommended guidelines [@StehmanKeyissuesrigorous2019] to conduct an independent assessment of the categorical accuracy of the final maps, using a set of `r sum(ref_ct)` (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) point-based, map reference labels, which were placed using a stratified random sample design, and collected through the labelling platform by two expert supervisors (see SI for full details on sample design and collection).
+For efficiency, the supervisors labelled separate portions of the sample, but a small subset was labelled by both.
+We used the label agreement on this subset to estimate uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019].
+The sample was labelled with four classes: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
+The last two classes provided a further measure of uncertainty in the map reference sample.
+We used the sample to calculate the overall accuracy for each map, the class-wise User's and Producer's accuracy, and the 95% confidence intervals for each accuracy measure [@OlofssonGoodpracticesestimating2014; @OlofssonMakingbetteruse2013; @StehmanKeyissuesrigorous2019].
+
+To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against the field boundaries digitized by by the labellers' over validation sites (primary grid cells).
 We chose this approach because of existing uncertainties in polygon-based accuracy assessment methods [@yeReviewAccuracyAssessment2018], and because the map's ability to represent field sizes was of greatest interest.
-To undertake this comparison, we selected the field polygons from the most accurate labeller to digitize each of the 100 validation sites in each AOI, and calculated the site-wise average area and number of polygons.
-We then calculated the same statistics from the segmented boundaries that intersected each validation grid.
-We compared the distributions and proximity of two measures of central tendency (mean and median) calculated from the two datasets for each AOI, and across all AOIs.
-
-<!-- Related to this, we evaluated two different versions of the active learning uncertainty criterion to evaluate the impact on performance gain.  -->
-
-<!-- Finally, we evaluated production performance in terms of speed of labelling and accuracy of results by comparing the generation time (label creation, training, and then classification) and accuracy of maps trained by 1) our core team of 10-12 experienced workers, and 2) a team of 100 novice workers from three different countries.
--->
-
-## Assessing the characteristics of Ghanaian cropland
-
-Using the final mapped results, we calculated the estimated area of cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
+To undertake this comparison, we selected the polygons from the most accurate labeller to digitize the 100 validation grids in each AOI, and calculated the average area and number of polygons in each cell.
+We then calculated the same statistics from the segmented boundaries that intersected each validation grid, and compared the two sets of statistics.
+
+We used to the final maps to evaluate the characteristics of Ghana's croplands.
+We calculated the estimated area of cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
 We used the map reference sample to calculate adjusted area estimates and confidence intervals for each map class, and used the differences between labellers' polygons and segmented boundaries at validation sites to calculate bias-adjusted estimates of mean field sizes and the total number of fields.
@@
 In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
-Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agroecozones (see Figure S4A).
+Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agro-ecozones (see Figure S4A).
 The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.
@@
 ### Categorical accuracy
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-# reference labels, for total count and uncertainty of class categories
-ref_labels <- st_read(
-  system.file("extdata/map_reference_labels.geojson", 
-              package = "activemapper"), quiet = TRUE
-)
-class_ct <- as_tibble(ref_labels) %>% group_by(category) %>% count()
-ref_ct <- as_tibble(ref_labels) %>% group_by(worker_id) %>% count() %>% pull(n)
-
-crp_ct <- class_ct %>% filter(category %in% c("annualcropland", "unsure1"))
-noncrp_ct <- class_ct %>% filter(category %in% c("noncropland", "unsure2"))
-
-# observer agreement/correlation
-ref_labels_joint <- st_read(
-  system.file("extdata/map_reference_labels_joint.geojson", 
-              package = "activemapper"), quiet = TRUE
-)
-joint_classes <- ref_labels_joint %>% as_tibble() %>% 
-  select(name, category, worker_id) %>% 
-  arrange(name) %>% 
-  mutate(class = case_when(
-    category == "annualcropland" | category == "unsure1" ~ 1, 
-    category == "noncropland" | category == "unsure2" ~ 0 
-  ))
-class_mat <- joint_classes %>% dplyr::select(name, class, worker_id) %>% 
-  tidyr::pivot_wider(names_from = worker_id, values_from = class) %>% 
-  dplyr::select(-name) %>% as.matrix()
-
-usertab <- table("User1" = class_mat[, 1], "User2" = class_mat[, 2])
-
-agreement <- round(sum(diag(usertab)) / sum(usertab) * 100, 1)
-observer_r <- cor.test(class_mat[, 1], class_mat[, 2], method = "spearman")
-
-```
-
 We used a map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of the per-pixel classifications (resulting from thresholding the Random Forests probability), as well as the segmented field boundary maps.

### 2021-07-01 inst/manuscript/si.Rmd
Methods reorganization complete

@@
 
-## Composites Assessing quality of composites
+## Assessing quality of composites
 The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites.
@@
 
-Two supervisors (Ye and Estes) who were not involved in labelling developed the map reference sample evaluated the PlanetScope composites at each site to determine its class membership.
+Two supervisors (Ye and Estes), who were not involved in collecting training and validation labels, collected the map reference sample by evaluating the PlanetScope composites at each site to determine its class membership.
 The following set of rules were followed in collecting the sample:

### 2021-07-02 inst/manuscript/figures.Rmd
Manuscript and SI revised through Results (map accuracy section).

@@
        height = 4, units = "in", dpi = 300, bg = "transparent")
+```
+
+### Figure S4: AOI Zones
+
+```{r, eval=FALSE}
+# zone maps
+zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
+                  "aez" = list(1, 2:4, 5, 6:7))
+aoi_zones <- aois %>% 
+  mutate(zone = case_when(
+    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
+    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
+    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
+    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(area))
+
+rmapshaper::ms_simplify(aez) %>% plot()
+aez %>% st_buffer(dist = 0.01) %>% st_union(by_feature = TRUE) %>% 
+  st_buffer(dist = -0.01) %>% 
+  mutate(zone = case_when(
+    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
+    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
+    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
+    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(AREA)) %>% plot()
+
+aez_zones <- rmapshaper::ms_simplify(aez, keep = 1) %>% 
+  mutate(zone = case_when(
+    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
+    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
+    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
+    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
+  ) %>%
+  group_by(zone) %>% 
+  summarize(area = sum(AREA)) 
+# plot(aez_zones["area"])
+# plot(aez["AREA"])
+
+m <- ggplot(aois) + geom_sf(lwd = 0.1) +
+  geom_sf(data = aoi_zones, lwd = 0.5, fill = "transparent") + 
+  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 4) + 
+  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+
+m2 <- ggplot(aez %>% filter(ID != 8)) + geom_sf(lwd = 0.1) +
+  geom_sf(data = aez_zones, lwd = 0.5, fill = "transparent") + 
+  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
+               aes(label = ID), 
+               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
+               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
+               size = 3, col = "grey60") + 
+  geom_sf_text(data = st_centroid(aez_zones), aes(label = zone), size = 4) + 
+  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
+ggsave(m + m2 + plot_annotation(tag_levels = 'A'), 
+       filename = pth("figures/si_aoi_zones.png"), width = 3.5 * 2, 
+       height = 4.5, units = "in", dpi = 300)
 ```
@@
 
-### Figure S4: PlanetScope image density
+### Figure S5: PlanetScope image density
 The density of imagery available for download from the image catalog.
@@
 
-### Figure S5: Initial training site distribution
+### Figure S6: Initial training site distribution
 Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
@@
 
-### Figure S6: Label count and quality
+### Figure S7: Label count and quality
 ```{r labelquality, eval=FALSE}
@@
 
-### Figures S7 and S8: Consensus label (Bayesian) risk
+### Figures S8 and S9: Consensus label (Bayesian) risk
 
@@
 
-### Figure S9: Random versus active sampling
+### Figure S10: Random versus active sampling
 ```{r, eval=FALSE}
@@
 
-### Figure S10: Training data error impact on probability images
+### Figure S11: Training data error impact on probability images
 ```{r, eval = FALSE}
@@
 
-```
-
-### Figure S11: AOI Zones
-
-```{r, eval=FALSE}
-# zone maps
-zone_grps <- list("aoi" = list(1:3, 4:9, c(10:11, 13, 14, 16), c(12, 15)),
-                  "aez" = list(1, 2:4, 5, 6:7))
-aoi_zones <- aois %>% 
-  mutate(zone = case_when(
-    aois %in% zone_grps$aoi[[1]] ~ "Zone 1", 
-    aois %in% zone_grps$aoi[[2]] ~ "Zone 2", 
-    aois %in% zone_grps$aoi[[3]] ~ "Zone 3", 
-    aois %in% zone_grps$aoi[[4]] ~ "Zone 4")
-  ) %>%
-  group_by(zone) %>% 
-  summarize(area = sum(area))
-
-rmapshaper::ms_simplify(aez) %>% plot()
-aez %>% st_buffer(dist = 0.01) %>% st_union(by_feature = TRUE) %>% 
-  st_buffer(dist = -0.01) %>% 
-  mutate(zone = case_when(
-    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
-    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
-    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
-    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
-  ) %>%
-  group_by(zone) %>% 
-  summarize(area = sum(AREA)) %>% plot()
-
-aez_zones <- rmapshaper::ms_simplify(aez, keep = 1) %>% 
-  mutate(zone = case_when(
-    ID %in% zone_grps$aez[[1]] ~ "Zone 1", 
-    ID %in% zone_grps$aez[[2]] ~ "Zone 2", 
-    ID %in% zone_grps$aez[[3]] ~ "Zone 3", 
-    ID %in% zone_grps$aez[[4]] ~ "Zone 4")
-  ) %>%
-  group_by(zone) %>% 
-  summarize(area = sum(AREA)) 
-# plot(aez_zones["area"])
-# plot(aez["AREA"])
-
-m <- ggplot(aois) + geom_sf(lwd = 0.1) +
-  geom_sf(data = aoi_zones, lwd = 0.5, fill = "transparent") + 
-  geom_sf_text(data = st_centroid(aoi_zones), aes(label = zone), size = 4) + 
-  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
-
-m2 <- ggplot(aez %>% filter(ID != 8)) + geom_sf(lwd = 0.1) +
-  geom_sf(data = aez_zones, lwd = 0.5, fill = "transparent") + 
-  geom_sf_text(data = st_centroid(aez) %>% filter(ID != 8) %>% arrange(ID),
-               aes(label = ID), 
-               nudge_x = c(0.75, 0, 0, -1.25, -1, -1.25, 0),
-               nudge_y = c(0.25, 0, 0, 0.7, 0.25, 1, 0),
-               size = 3, col = "grey60") + 
-  geom_sf_text(data = st_centroid(aez_zones), aes(label = zone), size = 4) + 
-  theme_void() + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))
-ggsave(m + m2 + plot_annotation(tag_levels = 'A'), 
-       filename = pth("figures/si_aoi_zones.png"), width = 3.5 * 2, 
-       height = 4.5, units = "in", dpi = 300)
 ```

### 2021-07-02 inst/manuscript/manuscript.Rmd
Manuscript and SI revised through Results (map accuracy section).

@@
 We generated a catalog of `r nrow(tile_key) * 2` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana.
-To assess the quality of the resulting composites, we graded the tiles drawn from `r tilen` randomly selected tile grid cells using a four category score that evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g.
+To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using a four category scoring system, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g.
 because of between-date image mis-registrations).
@@
 
-## Cropland classification using active learning 
+## Mapping cropland probabilities with active learning 
 The first step in creating a country-wide field boundary map of Ghana was to create a pixel-wise classification of cropland probabilities throughout the country.
@@
 
-In the next two sections, we describe the labelling and machine learning components of the active learning process in more detail
+In the next two sections, we describe the labelling and machine learning components of the active learning process in more detail.
 
@@
 In addition to training and validation labels, we also collected training reference labels and map reference labels [@ElmesAccountingtrainingdata2020].
-The former were a selection of primary grid cells selected to represent the range of cropland types and densities in Ghana, which were labelled by expert analysts (the lead researchers on this project).
-We were used these to assess the performance of the individual labellers collecting training and validation labels.
-Map reference labels were collected and used to assess the accuracy of the final map (see Section X).
+The former were a set of 98 primary grid cells selected to represent the range of cropland types and densities in Ghana, which were labelled by expert analysts (the lead researchers on this project).
+We used these to assess the performance of the individual labellers collecting training and validation labels.
+Map reference labels were collected and used to assess the accuracy of the final map (see Section 2.4).
 
@@
 Where $\theta$ represents the true cover type of a pixel (field or not field), *D* is the label assigned to that pixel by a labeller, and $W_i$ is an individual labeller.
-P($\theta$\|D) is the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is the average score of the accuracy assessment assignments an individual labeller completed within the AOI, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
+P($\theta$\|D) is the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is the average score (ranging between 0 and 1) of the accuracy assessment assignments an individual labeller completed within the AOI, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
 This approach therefore used the overall accuracy of each labeller to weight their label for a given pixel when combining it with those made by other labellers' for the same pixel (see SI for further details).
 Each pixel in the target grid cell was merged using this approach (n = 40000), which helps to minimize individual labellers' errors.
-We estimated a confidence for each consensus label by  calculating its Bayesian Risk (see SI), which ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels in the label, and 1 indicating complete disagreement.
-
-### Machine learning
+We estimated a confidence measure for each consensus label by  calculating its Bayesian Risk (see SI), which ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels in the label, and 1 indicating complete disagreement.
+
+### Cropland classification model
 Upon completing each batch of labels, the platform automatically launched a machine learning cluster (Elastic Map Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf}) comprised of several hundred to a thousand CPUs, depending on the size of the AOI.
@@
 
-### Classifier performance
-To assess performance of the Random Forests classifier, we used the validation sample to calculate binary accuracy, the F1 score (the geometric mean of precision and recall), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014].
+### Model performance
+To assess performance of the Random Forests classifier, we used the validation sample to calculate binary accuracy, the F1 score (the geometric mean of precision and recall), and the area under the curve of the Receiver Operating Characteristic [@Pontiustotaloperatingcharacteristic2014], as well as the false positive rate.
 We calculated these measures each time the model was retrained for a given AOI, in order to assess the change in classifier performance with each active learning iteration.
@@
 
-To quantify the potential impact of label error on classification results, we evaluated the performance differences between models trained with three different sets of labels: 1) those from the lowest scoring labeller to map each training site, 2) those from the highest scoring labeller, and 3) the consensus labels.
-This assessment follows recommended Tier 1 (i.e.
-best practice) standards to account for training data errors [@ElmesAccountingtrainingdata2020].
+To quantify the potential impact of label error on classification results, we conducted two further analyses.
+We evaluated the performance differences between models trained with three different sets of labels: 1) those from the lowest scoring labeller to map each training site, 2) those from the highest scoring labeller, and 3) the consensus labels.
+We also calculated the correlations between the mean Bayesian Risk of labels in each AOI and the corresponding model performance metrics (Table S3).
 
@@
 
-In the second step, we used a compact watershed algorithm [@neubertCompactWatershedPreemptive2014] to segment the weighted edge image, specifying a high number of segments (6,400) segments per tile, so that the mean segment size (<0.5 ha) was finer than the expected mean field size (1-2 ha).
+In the second step, we used a compact watershed algorithm [@neubertCompactWatershedPreemptive2014] to segment the edge image, specifying a high number of segments (6,400) segments per tile, so that the mean segment size (<0.5 ha) was finer than the expected mean field size (>1 ha).
 
@@
 
-We followed recommended guidelines [@StehmanKeyissuesrigorous2019] to conduct an independent assessment of the categorical accuracy of the final maps, using a set of `r sum(ref_ct)` (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) point-based, map reference labels, which were placed using a stratified random sample design, and collected through the labelling platform by two expert supervisors (see SI for full details on sample design and collection).
-For efficiency, the supervisors labelled separate portions of the sample, but a small subset was labelled by both.
-We used the label agreement on this subset to estimate uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019].
-The sample was labelled with four classes: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
-The last two classes provided a further measure of uncertainty in the map reference sample.
+We followed recommended guidelines [@StehmanKeyissuesrigorous2019] to conduct an independent assessment of the categorical accuracy of the final maps, using a set of `r sum(ref_ct)` (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) point-based, map reference labels, which were placed across Ghana using a stratified random sample design, and collected through the labelling platform by two expert supervisors (see SI for full details on sample design and collection).
+For efficiency, the supervisors labelled separate portions of the sample, but overlapped on a small subset (n = `r sum(usertab)`).
+We calculated the label agreement (`r agreement`%) on this subset to estimate uncertainty in the map reference sample [@StehmanKeyissuesrigorous2019].
+In addition to this, the sample was labelled with four classes: cropland; non-cropland; unsure but likely cropland; unsure but likely non-cropland.
+The last two classes, which constituted `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)`% of the sample, provided a further measure of uncertainty in the map reference sample
+
 We used the sample to calculate the overall accuracy for each map, the class-wise User's and Producer's accuracy, and the 95% confidence intervals for each accuracy measure [@OlofssonGoodpracticesestimating2014; @OlofssonMakingbetteruse2013; @StehmanKeyissuesrigorous2019].
+We calculated these measures across the entire country, as well as several different zones, to evaluate regional difference in accuracy.
+We defined two sets of zonations (Figure S4), each containing four zones, the first created by grouping 1) the three northern AOIs (1-3), 2) the six central AOIs (4-9), 3) the four southwestern AOIs (10, 11, 13, 14, 16), and 4) the two southeastern zones (13, 15).
+This grouping differs from the three clusters used to collect initial model training samples, as we designed these to divide the country more finely, and to isolate the less forested southeastern third of Ghana from the more forest northwest.
+The second zonation was developed by grouping the country's eight agro-ecological zones into four broader clusters (Figure S4B).
+We confined use of this zonation to better understand patterns in the per-pixel classifications.
 
@@
 
-We used to the final maps to evaluate the characteristics of Ghana's croplands.
+We used the final maps to evaluate the characteristics of Ghana's croplands.
 We calculated the estimated area of cropland in Ghana, as well as the average size and total number of fields in the different AOIs.
@@
 
-We developed maps of Ghana's cultivated croplands within an area of `r km2` km$^2$, which included portions of neighboring countries overlapped by images tiles.
-
-## Image catalog and quality
+Our results produced two separate maps of Ghana's annual croplands, over a total area of `r km2` km$^2$ that included portions of the neighboring countries overlapped by image tiles.
+
+## Image quality
 
@@
 ```
-
-To develop the maps, we first generated an image catalog for the `r nt` tiles covering Ghana.
-This entailed processing all PlanetScope imagery intersecting these tiles between May-September, 2018 (the growing season) and December, 2018 to February, 2019 (the subsequent dry season).
-The longer period was necessary for the growing season because of the frequent cloud cover, which substantially limits the number of clear scenes for any tile (Figure S3).
-For the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November.
-
-An assessment with two observers (see SI for observer details) found that average quality per growing season composite tile was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85, while the average quality of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85).
-Composite quality in both seasons was highest in the northern half of the country and lowest in the southwest (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S3).
+The assessment of image composites found that their quality in both seasons was highest in the northern half of the country and lowest in the southwest, (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S3).
+The average quality score of growing season composites was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85 (out of 1; Figure \ref{fig:imqual}B), while the mean score of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85).
 
@@
 
-## Active learning
-
-### Training data collection
-
+## Cropland probabilities
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-After training of models with the initial randomly selected label sets, the active learning process was run for 3 iterations for 12 of the 16 AOIs, resulting in `r typicaln` labels per AOI.
-AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (\>83%) and showed little subsequent improvement.
-The models for these two AOIs were thus trained with `r aoi10n` - `r aoi14n` samples.
-AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
-In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
-Labels collected during the active learning iterations showed distinct patterns in several AOIs, which often fell along ecotones, such as the boundaries between agro-ecozones (see Figure S4A).
-The total number of unique training and validation sites across the country were `r ttrain` and `r tval`, respectively.
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("label_summary")  
-toofew <- label_summary$nlabels %>% filter(n < 10) %>% pull(ID)
-tsites <- label_summary$nlabels %>% filter(!ID %in% toofew) %>% 
-  group_by(kml_type) %>% 
-  summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
-            stdev = round(sd(n)), lower = min(n), upper = max(n))
-
-# training rate
-train_rate <- label_summary$nlabels %>% 
-  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
-  rename_all(funs(c("ID", "tr", "tref"))) %>% 
-  mutate(trate = tref / tr) %>% filter(tr > 10)
-
-tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
-# nvecs <- tsites$tot[1])
-# muvecs <- fmt(tsites$mu[1])
-
-```
-
-The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each.
-As each training site was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made.
-Each labeller digitized an average of `r tsitesc$mu[1]` (see Figure S5A for more details on labelling effort).
-Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites (Figure S4A), with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
-The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figure S5B for detailed score distributions).
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("label_risk")
-murisk <- label_risk$stats %>% group_by(usage) %>% 
-  summarize(Risk = round(mean(Risk), 3)) %>% 
-  tidyr::pivot_wider(names_from = usage, values_from = Risk)
-mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
-  filter(Risk == min(Risk))
-maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
-  filter(Risk == max(Risk))
-minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
-  filter(Risk == min(Risk))
-maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
-  filter(Risk == max(Risk))
-
-# label clusters
-cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
-  summarize(Risk = round(mean(Risk), 3))
-cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
-  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
-cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
-  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
-
-# training clusters
-cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
-  round(3)
-cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
-  round(3)
-
-```
-
-After each site was mapped by four labellers, consensus labels were generated.
-The Bayesian Risk (see SI) of each training and validation label was calculated as an additional measure of label quality.
-The average risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
-Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
-Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
-Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster (see SI); the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.
+To make the initial maps of cropland probabilities, the active learning process ran for 3 iterations in 12 of 16 AOIs, varying from as little as 1 to as many as 4 iterations across the other 4 AOIs, with the number of iterations associated with the performance of the starter models against the validation labels (i.e.
+AOIs with higher starting performance stopped after fewer iterations, see SI).
+A total of `r ttrain` training and `r tval` validation labels were collected by 20 labellers to develop and assess model performance.
+Each AOI's model was trained by the 500 randomly selected labels, plus an additional `r aoi10n` - `r aoi15n` (typically `r typicaln`) labels within each the AOI that were selected by active learning.
+Actively selected labels showed distinctive patterns in several AOIs, such as concentrating along ecotones, including the boundaries between agro-ecozones (see Figure S4A).
 
@@
 
-Model performance was calculated for each iteration within each AOI.
+The performance of the Random Forest classifier typically improved with each active learning iteration.
 The average accuracy, AUC, and F1 at iteration 0 were `r imu0$Accuracy$score`, `r imu0$AUC$score`, and `r imu0$F1$score`, respectively, increasing to `r imu3$Accuracy$score`, `r imu3$AUC$score`, and `r imu3$F1$score` by iteration 3 (Figure \ref{fig:alperformance}).
@@
 
-The comparison of active versus randomized training sample collection (in AOIs 1, 8, and 15) showed that the former approach outperformed the latter.
-After three iterations, the accuracy, AUC, and F1 scores resulting from active learning were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than the scores from a randomly trained model (Figure S8).
-However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2.
-The negative results for accuracy was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:alperformance}).
-In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
-Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.
-
-### The impact of training data error
+The experiment conducted in three AOIs (in AOIs 1, 8, and 15) showed that training models with active learning improved performance compared to random approaches to training label selection.
+After three iterations, the accuracy, AUC, and F1 scores for the actively trained models were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than those for randomly trained models (Figure S8).
+However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2 (see SI for more details).
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("label_summary")  
+# toofew <- label_summary$nlabels %>% filter(n < 10) %>% pull(ID)
+# tsites <- label_summary$nlabels %>% filter(!ID %in% toofew) %>% 
+#   group_by(kml_type) %>% 
+#   summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
+#             stdev = round(sd(n)), lower = min(n), upper = max(n))
+# 
+# # training rate
+# train_rate <- label_summary$nlabels %>% 
+#   tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
+#   rename_all(funs(c("ID", "tr", "tref"))) %>% 
+#   mutate(trate = tref / tr) %>% filter(tr > 10)
+# 
+# tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
+# # nvecs <- tsites$tot[1])
+# # muvecs <- fmt(tsites$mu[1])
+
+data("label_risk")
+murisk <- label_risk$stats %>% #group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3)) #%>% 
+  # tidyr::pivot_wider(names_from = usage, values_from = Risk)
+
+# label clusters
+cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% #group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3))
+cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
+  # group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3))
+cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
+  # group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3))
+```
+
+### The impact of label error on model performance
+We used the two measures of label quality calculated by the platform, the average accuracy score of each labeller and Bayesian Risk (or simply "label risk"), to assess the potential impacts of label error and uncertainty on model performance.
+The average of each labeller's AOI-specific accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figures S4 and S5 for details on label scores and number of assignments per labeller).
+The average Bayesian Risk was `r murisk$Risk`, with highest label risk (`r cl1$Risk`) in the northern AOIs (AOIs 1-6; Figures S6-7), lowest (`r cl1$Risk`) in the southwestern AOIs (AOIs 10, 11, 13, 14, 16), and intermediate (`r cl2$Risk`) in the central-southeastern AOIs (AOIs 7-9, 12, 15).
 
@@
 
-The potential impact of label errors on map quality was assessed in four AOIs (1, 2, 8, and 15).
-The results of these tests showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
+Treating average labeller accuracy as a proxy for label error, we used these scores to select the labels used in the experiments to assess how label error impacted model performance.
+The results of these tests, which were conducted in AOIs 1, 2, 8, and 15, showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
 Performance metrics from consensus-trained models were just `r metdiffs$AUC$chdelt` - `r metdiffs$Accuracy$chdelt` percent higher than those models trained with the most accurate individuals' labels (accuracy = `r metdiffs$Accuracy$high`; AUC = `r metdiffs$AUC$high`; F1 = `r metdiffs$F1$high`), but were `r metdiffs$AUC$cldelt` - `r metdiffs$Accuracy$cldelt` higher than models trained with the least accurate individual labels (accuracy = `r metdiffs$Accuracy$low`; AUC = `r metdiffs$AUC$low`; F1 = `r metdiffs$F1$low`).
@@
 
-A second measure of the impact of label error is found within the correlations between the mean label risk per AOI and the model performance metrics (Table S3).
-Accuracy and AUC had strong (Spearman's Rank Correlation = `r risk_corr[, "Accuracy"]` to moderate (r = `r risk_corr[, "AUC"]`) negative correlations with label risk, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
+Correlations between the mean label risk per AOI and model performance metrics showed strong (Spearman's Rank Correlation = `r risk_corr[, "Accuracy"]`) to moderate (r = `r risk_corr[, "AUC"]`) negative correlations between label risk and accuracy and AUC, respectively, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
 The positive sign of the latter relationship is counter-intuitive, but is explained by risk's association with precision, one of two inputs to F1, which was moderately positive (`r risk_corr[, "Precision"]`), whereas risk had a negligible correlation with recall (`r risk_corr[, "Recall"]`), F1's other component.
@@
 ## Map accuracy
-
 ### Categorical accuracy
-We used a map reference sample of `r sum(ref_ct)` sites (`r sum(crp_ct$n)` cropland; `r sum(noncrp_ct$n)` non-cropland) to evaluate the accuracy of the per-pixel classifications (resulting from thresholding the Random Forests probability), as well as the segmented field boundary maps.
-We first evaluated the uncertainty in the map reference classes by assessing 1) the overall agreement between map reference labels collected by two separate supervisors at `r sum(usertab)` sites, and 2) the confidence of the labels assigned by the supervisors (see SI for details).
-The first measure showed that the two individual supervisors' labels agreed at `r agreement`% of common sites, while the second showed that `r round(sum(crp_ct$n[2], noncrp_ct$n[2]) / sum(ref_ct) * 100, 1)` of sites were labelled with the two classes that indicated a level of uncertainty.
-
+We used the map reference sample to evaluate the accuracy of the cropland probability map (after converting to a classification using a threshold probability of 0.5) and the map of segmented field boundary maps.
 We found that the overall accuracy of the pixel-wise classifications was 88% against this map reference sample (Table \ref{tab:mapaccuracy}).
@@
 The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S12).
-
-<!-- ### Label quality and map accuracy -->
-
-<!-- ```{r, echo = FALSE, out.width="100%", fig.cap="Difference in overall, user's, and producer's accuracy between maps made with consensus labels, the least accurate worker's, and most accurate worker's labels.", fig.align='center', message=FALSE} -->
-
-<!-- knitr::include_graphics('figures/figure6.png') -->
-
-<!-- # library(tidyverse) -->
-
-<!-- # library(cowplot) -->
-
-<!-- # set.seed(1) -->
-
-<!-- # acc_tb <- tibble(aoi = c(1:9, 12, 15),  -->
-
-<!-- #                  Consensus = runif(11, min = 0.75, max = 0.85), -->
-
-<!-- #                  Highest = runif(11, min = 0.72, max = 0.8), -->
-
-<!-- #                  Least = runif(11, min = 0.65, max = 0.75)) -->
-
-<!-- # acc_tb <- gather(acc_tb, key = "type", value = "score", -aoi) -->
-
-<!-- # ggplot(acc_tb) +  -->
-
-<!-- #   geom_boxplot(aes(x = type, y = score), fill = "grey") +  -->
-
-<!-- #   xlab("Labelling strategy") + ylab("Accuracy") +  -->
-
-<!-- #   theme(panel.grid.major.y = element_line(colour = "grey80", linetype = 3)) -->
-
-<!-- ``` -->
 

### 2021-07-02 inst/manuscript/si.Rmd
Manuscript and SI revised through Results (map accuracy section).

@@
 
+fmt <- function(x) format(x, big.mark = ",")
 ```
@@
 
-## Assessing quality of composites
+## Image compositing 
+### Variation in lengths of temporal window used
+The typical window for compositing dry season imagery was December, 2018 to February, 2019, but in the cloudiest regions (AOIs 10, 11, 13, 14, 16) we started the dry season window in November, to ensure a sufficient density of images for compositing.
+
+### Assessing composite quality
 The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites.
@@
 
-# Supplemental Results
-## Image catalog and quality
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("image_quality")
-wqual_diff <- image_quality$img %>% filter(kml_type == "F") %>%
-  group_by(name, season, worker_id) %>%
-  summarize(tile = min(tile), score = sum(score) / 12) %>% 
-  ungroup() %>% 
-  pivot_wider(names_from = worker_id, values_from = score,
-              names_prefix = "w") %>%
-  dplyr::select(tile, season, w92, w99) %>% 
-  mutate(d = w92 - w99, 
-         absd = abs(d))
-```
-
-The total number of PlanetScope images available for download from the Planet API [@planetteamPlanetApplicationProgram2018] per tile for each of the two major seasons is shown in Figure \ref{fig:imagedensity}.
-
-```{r imagedensity, echo = FALSE, out.width="90%", fig.cap="The density of images available through the Planet API for the five month growing season in 2018 and the subsequent three month dry season (excluding the month of November), shown in relation to the AOI boundaries.", fig.align='center', message=FALSE, results='hold'}
-knitr::include_graphics('figures/si_planet_coverage.png')
-```
-
-The image quality assessment was conducted by two separate observers (L.
-Song and Q.
-Zhang).
-Each observer assessed the composites for each season at all 50 of the randomly selected tiles (see main text) following the defined criteria (Table \ref{tab:imqcriteria}).
-To calculate the score for a given tile for each season, we summed the ranking across the four categories for each observer, rescaled the values, and then calculated the mean tile score across both observers by season.
-The mean difference between the two observers was `r d <- round(mean(wqual_diff$d), 3); d` (sd = `r absd <- round(sd(wqual_diff$d), 3); absd`) and the mean absolute difference was `r round(mean(wqual_diff$absd), 3)` (sd = `r round(sd(wqual_diff$absd), 3)`), thus one observer scored tiles about `r abs(round(d * 100))` percent lower than the other observer, on average.
-
-## Active learning process
-### Labelling
-The locations of training, training reference, and validation points are shown in Figure \ref{fig:trainval}.
-In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season.
-Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
-```{r trainval, echo = FALSE, out.width="70%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.
-Grey background lines and show the boundaries and ID number of Ghana's agroecozones: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_training_validation_pts.png')
-```
-
-The distribution of data collection effort and scores from training reference sites by labelling team are shown in Figure \ref{fig:trainqual}.
-These results include those from re-labelling the initial training in Cluster 2 (AOIs 7-9, 12, 15).
-This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
-We replaced the image overlays in the instance for central-southeast (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
-The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.
-
-```{r trainqual, echo = FALSE, out.width="50%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
-Labellers' identities are anonymized.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_training_qual_n.png')
-```
-
-The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}).
-The three initial clusters include the second mapping of Cluster 2.
-
-```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_label_risk_map.png')
-```
-
-```{r labelriskhist, echo = FALSE, out.width="40%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
-The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_label_risk_hists.png')
-```
-
-### Validation
-The differences in accuracy, AUC, and F1 between the active learning process and the random retraining at each iteration for AOIs 1, 8, and 15 (Figure \ref{fig:randomvactive}).
-Small differences due to random variations in the RandomForest models are the reason for the non-zero differences at iteration 0, when both models were trained with the same level set.
-```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites.
-Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_random_vs_active.png')
-```
-
-### The impact of training data error
-Probability images resulting from Random Forests models trained with labels generated under three different labeling strategies are illustrated in Figure \ref{fig:labelstrategy}.
-These included consensus labels, and those individual labels that were likely to be the most and least accurate for each training site.
-Label accuracy was based on the mean score of each labeller against the training reference sites (Figure \ref{fig:trainqual}), as assessed when labelling a given AOI.
-These images were created for a single tile in AOI 1.
-
-```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by Random Forests models trained with A) consensus labels, B) the most accurate individual labels, and C) the least accurate individual labels.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_label_strategy_probs.png')
-```
-
-The association between the average label risk per AOI and several model performance metrics, assessed using Spearman Rank Correlation, is shown in Table \ref{tab:labelrisk_v_metrics}.
-
-```{r labelrisk_v_metrics, echo=FALSE}
-data("label_risk_metrics")
-
-cor_fun <- function(df) {
-  cor.test(df$Risk, df$Score, method = "spearman") %>% broom::tidy()
-}
-
-lr_mets <- unique(label_risk_metrics$Metric)
-risk_corr <- lapply(lr_mets[lr_mets != "TSS"], function(x) {
-  dat <- label_risk_metrics %>% filter(Metric == x)
-  cor_fun(dat) %>% mutate(Metric = x)
-}) %>% do.call(rbind, .) %>% 
-  select(estimate, Metric) %>% 
-  rename(r = estimate)
-nms <- risk_corr$Metric
-risk_corr <- round(t(risk_corr[, 1]), 3)
-colnames(risk_corr) <- nms
-
-cap <- glue::glue("Spearman's Rank correlation between the average label risk per AOI and a variety of model performance metrics.")
-kable(risk_corr, format = "latex", booktabs = "T", caption = cap) %>% 
-  kable_styling(full_width = TRUE, latex_options = c("hold_position")) #%>% 
-  # column_spec(1, width = "8cm")
-```
-
-## Map accuracy
-
-### Reference label uncertainty
+#### Map reference label uncertainty
 
@@
 
-### Categorical accuracy
-In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).
+#### Accuracy assessment zones
+The zones used to assess regional variation in map accuracy are illustrated in Figure \ref{fig:aoizones}.
 
@@
 ```
+
+# Supplemental Results
+## Image catalog and quality
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("image_quality")
+wqual_diff <- image_quality$img %>% filter(kml_type == "F") %>%
+  group_by(name, season, worker_id) %>%
+  summarize(tile = min(tile), score = sum(score) / 12) %>% 
+  ungroup() %>% 
+  pivot_wider(names_from = worker_id, values_from = score,
+              names_prefix = "w") %>%
+  dplyr::select(tile, season, w92, w99) %>% 
+  mutate(d = w92 - w99, 
+         absd = abs(d))
+```
+
+The total number of PlanetScope images available for download from the Planet API [@planetteamPlanetApplicationProgram2018] per tile for each of the two major seasons is shown in Figure \ref{fig:imagedensity}.
+
+```{r imagedensity, echo = FALSE, out.width="90%", fig.cap="The density of images available through the Planet API for the five month growing season in 2018 and the subsequent three month dry season (excluding the month of November), shown in relation to the AOI boundaries.", fig.align='center', message=FALSE, results='hold'}
+knitr::include_graphics('figures/si_planet_coverage.png')
+```
+
+The image quality assessment was conducted by two separate observers (L.
+Song and Q.
+Zhang).
+Each observer assessed the composites for each season at all 50 of the randomly selected tiles (see main text) following the defined criteria (Table \ref{tab:imqcriteria}).
+To calculate the score for a given tile for each season, we summed the ranking across the four categories for each observer, rescaled the values, and then calculated the mean tile score across both observers by season.
+The mean difference between the two observers was `r d <- round(mean(wqual_diff$d), 3); d` (sd = `r absd <- round(sd(wqual_diff$d), 3); absd`) and the mean absolute difference was `r round(mean(wqual_diff$absd), 3)` (sd = `r round(sd(wqual_diff$absd), 3)`), thus one observer scored tiles about `r abs(round(d * 100))` percent lower than the other observer, on average.
+
+## Mapping cropland probabilities with active learning
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+initial <- train_val_sites$initial %>% as_tibble %>% group_by(Cluster) %>%
+  count()
+active <- train_val_sites$active %>% as_tibble %>% filter(usage == "train") %>% 
+  group_by(aoi) %>% count()
+typicaln <- raster::modal(initial$n) + raster::modal(active$n)
+aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
+  (active %>% filter(aoi == "labeller10") %>% pull(n))
+aoi14n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
+  (active %>% filter(aoi == "labeller14") %>% pull(n))
+aoi15n <- (initial %>% filter(Cluster %in% 2) %>% pull(n)) + 
+  (active %>% filter(aoi == "labeller15") %>% pull(n))
+aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
+aoi3n <- aoi3init + (active %>% filter(aoi == "labeller3") %>% pull(n))
+```
+
+The number of active learning iterations per AOI was three, for all but four AOIs.
+AOIs 10 and 14 stopped after one and two iterations, respectively, as they started with high initial validation accuracies (\>83%) and showed little subsequent improvement.
+The models for these two AOIs were thus trained with `r aoi10n` - `r aoi14n` labels.
+AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
+In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
+
+### Labelling
+The locations of training, training reference, and validation points are shown in Figure \ref{fig:trainval}.
+In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season.
+Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
+```{r trainval, echo = FALSE, out.width="70%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.
+Grey background lines and show the boundaries and ID number of Ghana's agroecozones: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_training_validation_pts.png')
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("label_summary")  
+toofew <- label_summary$nlabels %>% filter(n < 10) %>% pull(ID)
+tsites <- label_summary$nlabels %>% filter(!ID %in% toofew) %>% 
+  group_by(kml_type) %>% 
+  summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
+            stdev = round(sd(n)), lower = min(n), upper = max(n))
+
+# training rate
+train_rate <- label_summary$nlabels %>% 
+  tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
+  rename_all(funs(c("ID", "tr", "tref"))) %>% 
+  mutate(trate = tref / tr) %>% filter(tr > 10)
+
+tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
+# nvecs <- tsites$tot[1])
+# muvecs <- fmt(tsites$mu[1])
+
+```
+
+The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each (Figure \ref{fig:trainqual}A).
+As each training and reference tasks was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made.
+Each labeller digitized an average of `r tsitesc$mu[1]`.
+Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites, with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
+The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)` (Figure \ref{fig:trainqual}B).
+
+These results include those from re-labelling the initial training in Cluster 2 (AOIs 7-9, 12, 15).
+This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
+We replaced the image overlays in the instance for central-southeast (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
+The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.
+
+```{r trainqual, echo = FALSE, out.width="50%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
+Labellers' identities are anonymized.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_training_qual_n.png')
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("label_risk")
+murisk <- label_risk$stats %>% group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3)) %>% 
+  tidyr::pivot_wider(names_from = usage, values_from = Risk)
+mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
+  filter(Risk == min(Risk))
+maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
+  filter(Risk == max(Risk))
+minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
+  filter(Risk == min(Risk))
+maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
+  filter(Risk == max(Risk))
+
+# label clusters
+cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3))
+cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
+  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
+cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
+  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
+
+# training clusters
+cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
+  round(3)
+cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
+  round(3)
+```
+
+The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}).
+The three initial clusters include the second mapping of Cluster 2.
+
+The average Bayesian risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
+Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
+Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
+Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster; the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.
+
+```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_risk_map.png')
+```
+
+```{r labelriskhist, echo = FALSE, out.width="40%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
+The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_risk_hists.png')
+```
+
+### Model performance
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# random versus active data
+data("iteration_metrics")
+
+# process data: remove labeller from AOI and select AOIs 1-16
+V1 <- c("Accuracy", "AUC", "F1")
+aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
+imetrics <- iteration_metrics %>% 
+  filter(Metric %in% V1 & AOI %in% aois) %>% 
+  mutate(AOI = gsub("labeller", "", AOI)) %>% 
+  mutate(type = case_when(
+    AOI %in% c(1, 8, 15) ~ "active", 
+    AOI %in% c("1r", "8r", "15r") ~ "random"
+  )) %>% 
+  mutate(AOI = gsub("r", "", AOI)) %>% 
+  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
+  mutate(delta = (active - random) / random * 100) 
+imetrics_all <- imetrics %>% 
+  filter(Iteration < 4) %>% 
+  group_by(Iteration, Metric) %>% 
+  summarize(active = mean(active), random = mean(random), 
+            delta = mean(delta)) %>%
+  ungroup() %>% mutate(AOI = "All") %>% 
+  select(AOI, Iteration, Metric, active, random, delta)
+
+ixl <- lapply(1:3, function(x) {
+  ix <- imetrics_all %>% filter(Iteration == x) %>% pull(delta)
+  ix <- round(ix, 1)
+  names(ix) <- V1
+  ix
+})
+names(ixl) <- as.roman(1:3)
+
+i1acc <- imetrics %>% 
+  filter(Iteration == 1 & Metric == "Accuracy") %>% 
+  pull(delta) %>% round(2)
+names(i1acc) <- imetrics %>% distinct(AOI) %>% pull()
+
+u <- function(x) unname(x)
+```
+
+The differences in accuracy, AUC, and F1 between the active learning process and the random retraining at each iteration for AOIs 1, 8, and 15 (Figure \ref{fig:randomvactive}).
+Small differences due to random variations in the RandomForest models are the reason for the non-zero differences at iteration 0, when both models were trained with the same level set.
+
+The lower accuracy of actively versus randomly trained models in earlier iterations was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:randomvactive}).
+In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
+Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.
+
+```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites.
+Postive percentages indicate superior performance by active learning, negative percentages the inverse.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_random_vs_active.png')
+```
+
+### The impact of training data error
+Probability images resulting from Random Forests models trained with labels generated under three different labeling strategies are illustrated in Figure \ref{fig:labelstrategy}.
+These included consensus labels, and those individual labels that were likely to be the most and least accurate for each training site.
+Label accuracy was based on the mean score of each labeller against the training reference sites (Figure \ref{fig:trainqual}), as assessed when labelling a given AOI.
+These images were created for a single tile in AOI 1.
+
+```{r labelstrategy, echo = FALSE, out.width="100%", fig.cap="Cropland probability images produced by Random Forests models trained with A) consensus labels, B) the most accurate individual labels, and C) the least accurate individual labels.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_strategy_probs.png')
+```
+
+The association between the average label risk per AOI and several model performance metrics, assessed using Spearman Rank Correlation, is shown in Table \ref{tab:labelrisk_v_metrics}.
+
+```{r labelrisk_v_metrics, echo=FALSE}
+data("label_risk_metrics")
+
+cor_fun <- function(df) {
+  cor.test(df$Risk, df$Score, method = "spearman") %>% broom::tidy()
+}
+
+lr_mets <- unique(label_risk_metrics$Metric)
+risk_corr <- lapply(lr_mets[lr_mets != "TSS"], function(x) {
+  dat <- label_risk_metrics %>% filter(Metric == x)
+  cor_fun(dat) %>% mutate(Metric = x)
+}) %>% do.call(rbind, .) %>% 
+  select(estimate, Metric) %>% 
+  rename(r = estimate)
+nms <- risk_corr$Metric
+risk_corr <- round(t(risk_corr[, 1]), 3)
+colnames(risk_corr) <- nms
+
+cap <- glue::glue("Spearman's Rank correlation between the average label risk per AOI and a variety of model performance metrics.")
+kable(risk_corr, format = "latex", booktabs = "T", caption = cap) %>% 
+  kable_styling(full_width = TRUE, latex_options = c("hold_position")) #%>% 
+  # column_spec(1, width = "8cm")
+```
+
+## Map accuracy 
+### Categorical accuracy
+In addition to the map accuracies and area estimates calculated per AOI zone (reported in main text; see Figure \ref{fig:aoizones}A), the accuracies were also assessed within several different groupings of agroecozones (Figures \ref{fig:aoizones}B and Table \ref{tab:aezaccuracy}).
 

### 2021-07-20 inst/manuscript/figures.Rmd
Fully revised manuscript (except abstract), SI, and figures.

@@
                         scale = c(1, 1), rel_heights = c(1, 0.7))
-ggsave(pth("figures/figure4.png"), 
+ggsave(pth("figures/figure3.png"), 
        width = 4, height = 6.5, units = "in", dpi = 300)
@@
 
-### Figure 5: Active learning gain
+### Figure 4: Active learning gain
 ```{r, eval=FALSE}
@@
 
-ggsave(pth("figures/figure5.png"), 
+ggsave(pth("figures/figure4.png"), 
        width = 7, height = 4, units = "in", dpi = 300)
@@
 
-### Figure 6: The impact of training data error
+### Figure 5: The impact of training data error
 ```{r, eval=FALSE}
@@
         strip.text = element_text(colour = "black"))
-ggsave(pth("figures/figure6.png"), height = 4, width = 6, dpi = 300)
+ggsave(pth("figures/figure5.png"), height = 4, width = 6, dpi = 300)
 ```
@@
 
-### Figure 7: Cropland maps
+### Figure 6: Cropland maps
 ```{r, eval=FALSE}
@@
 
-ggsave(filename = pth("figures/figure7.png"), 
+ggsave(filename = pth("figures/figure6.png"), 
        plot = inset_map, bg = "transparent", width = diff(maprange[1:2]), 
@@
 
-### Figure S7: Label count and quality
-```{r labelquality, eval=FALSE}
+### Figure S7: Assignment count
+```{r assignmentcount, eval=FALSE}
 data("label_summary")
@@
 
+# N sites by type
+yl <- c(0, ceiling(max(label_summary$nlabels$n) * 1.01))
+p <- label_summary$nlabels %>% filter(!ID %in% toofewq$ID) %>% 
+  group_by(ID, kml_type) %>% 
+  ggplot() + 
+  geom_bar(aes(x = reorder(ID, -n), y = n, fill = kml_type), 
+           position = "dodge", stat = "identity") +
+  xlab("ID") + ylab("N") + 
+  labs(fill = "Label Type") + 
+  scale_fill_manual(labels = c("Training/validation", "Accuracy assessment"),
+                    values = c("grey40", "grey70")) + 
+  scale_y_continuous(breaks = seq(yl[1], yl[2], by = 200), 
+                     expand = c(0, 0.01), limits = yl) + 
+  theme_linedraw() + 
+  # theme(legend.position = c(0.7, 0.85), axis.text = element_text(size = 11),
+  theme(axis.text = element_text(size = 11), 
+        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
+
+ggsave(pth("figures/si_training_n.png"), width = 5.5, 
+       height = 4, units = "in", dpi = 300, bg = "transparent")
+```
+
+### Figure S8: Random versus active sampling
+```{r, eval=FALSE}
+data("iteration_metrics")
+
+# Plots
+V1 <- c("Accuracy", "AUC", "F1")
+
+# prepare data
+# remove labeller from AOI and select AOIs 1-16
+aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
+imetrics <- iteration_metrics %>% 
+  filter(Metric %in% V1 & AOI %in% aois) %>% 
+  mutate(AOI = gsub("labeller", "", AOI)) %>% 
+  mutate(type = case_when(
+    AOI %in% c(1, 8, 15) ~ "active", 
+    AOI %in% c("1r", "8r", "15r") ~ "random"
+  )) %>% 
+  mutate(AOI = gsub("r", "", AOI)) %>% 
+  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
+  mutate(delta = (active - random) / random * 100)
+
+imetrics_all <- imetrics %>% filter(Iteration < 4) %>% 
+  group_by(Iteration, Metric) %>% 
+  summarize(active = mean(active), random = mean(random), 
+            delta = mean(delta)) %>%
+  ungroup() %>% mutate(AOI = "All") %>% 
+  select(AOI, Iteration, Metric, active, random, delta)
+imetrics <- bind_rows(imetrics, imetrics_all)
+
+# select out final iteration for each AOI
+mx_metrics <- imetrics %>% group_by(AOI) %>% 
+  filter(Iteration == max(Iteration)) %>% filter(Metric %in% V1)
+
+# plot
+cols <- c(rep("grey50", length(unique(imetrics$AOI)) -1), "black")
+lsize <- c(rep(0.3, length(unique(imetrics$AOI)) -1), 1)
+p <- ggplot(imetrics) + 
+  geom_hline(yintercept = 0, linetype = "dotted", col = "grey40") +
+  geom_line(aes(Iteration, delta, color = AOI, size = AOI)) +
+  geom_point(data = mx_metrics, aes(Iteration, delta, color = AOI), 
+             size = 1.2) +
+  ggrepel::geom_text_repel(
+    data = mx_metrics %>% filter(AOI != "All"), 
+    aes(label = AOI, x = Iteration, y = delta), col = "grey50", 
+    segment.color = "grey60", point.padding = 0.2, min.segment.length = 0.2,
+    segment.size = 0.2, size = 3, nudge_x = 0 
+  ) +
+  scale_color_manual(values = cols, guide = FALSE) +
+  scale_size_manual(values = lsize, guide = FALSE) + 
+  scale_x_continuous(expand = c(0.01, 0.01)) + 
+  scale_y_continuous(expand = c(0.01, 0.01), breaks = seq(-8, 4, 1)) + 
+  ylab("% difference") +                    
+  facet_wrap(vars(Metric), ncol = 3) +
+  theme_linedraw() + 
+  theme(panel.grid.minor.x = element_blank(), 
+        strip.background = element_rect(fill = "white"), 
+        strip.text = element_text(colour = "black"))
+
+ggsave(pth("figures/si_random_vs_active.png"), 
+       width = 6, height = 4, units = "in", dpi = 300)
+
+```
+
+### Figure S9: Label quality
+```{r labelquality, eval=FALSE}
+# Plot data from Figure S7
+
 # Q site accuracy
-p1 <- ggplot(label_quality2) + 
+p <- ggplot(label_quality2) + 
   geom_boxplot(aes(x = reorder(ID, -score), y = score, fill = ID), 
@@
 
-# N sites by type
-yl <- c(0, ceiling(max(label_summary$nlabels$n) * 1.01))
-p2 <- label_summary$nlabels %>% filter(!ID %in% toofewq$ID) %>% 
-  group_by(ID, kml_type) %>% 
-  ggplot() + 
-  geom_bar(aes(x = reorder(ID, -n), y = n, fill = kml_type), 
-           position = "dodge", stat = "identity") +
-  xlab("ID") + ylab("N") + 
-  labs(fill = "Map Type") + 
-  scale_fill_manual(values = c("grey40", "grey70")) + 
-  scale_y_continuous(breaks = seq(yl[1], yl[2], by = 200), 
-                     expand = c(0, 0.01), limits = yl) + 
-  theme_linedraw() + 
-  theme(legend.position = c(0.7, 0.85), axis.text = element_text(size = 11), 
-        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
-
-cp_theme <- theme(axis.text = element_text(size = 11),  
-                  legend.text = element_text(size = 11),
-                  legend.title = element_text(size = 11),
-                  axis.title = element_text(size = 12))
-gp <- cowplot::plot_grid(p2 + cp_theme, p1 + cp_theme, nrow = 2, 
-                         labels = LETTERS[1:2], label_x = 0.0) 
-                         #rel_widths = c(0.65, 0.45))
-ggsave(pth("figures/si_training_qual_n.png"), width = 4, 
-       height = 8, units = "in", dpi = 300, bg = "transparent")
-```
-
-### Figures S8 and S9: Consensus label (Bayesian) risk
+ggsave(pth("figures/si_label_quality.png"), width = 4, 
+       height = 4, units = "in", dpi = 300, bg = "transparent")
+```
+
+### Figures S10 and S11: Consensus label (Bayesian) risk
 
@@
 
-### Figure S10: Random versus active sampling
-```{r, eval=FALSE}
-data("iteration_metrics")
-
-# Plots
-V1 <- c("Accuracy", "AUC", "F1")
-
-# prepare data
-# remove labeller from AOI and select AOIs 1-16
-aois <- c(paste0("labeller", c(1, 8, 15, "1r", "8r", "15r")))
-imetrics <- iteration_metrics %>% 
-  filter(Metric %in% V1 & AOI %in% aois) %>% 
-  mutate(AOI = gsub("labeller", "", AOI)) %>% 
-  mutate(type = case_when(
-    AOI %in% c(1, 8, 15) ~ "active", 
-    AOI %in% c("1r", "8r", "15r") ~ "random"
-  )) %>% 
-  mutate(AOI = gsub("r", "", AOI)) %>% 
-  tidyr::pivot_wider(names_from = "type", values_from = "Score") %>% 
-  mutate(delta = (active - random) / random * 100)
-
-imetrics_all <- imetrics %>% filter(Iteration < 4) %>% 
-  group_by(Iteration, Metric) %>% 
-  summarize(active = mean(active), random = mean(random), 
-            delta = mean(delta)) %>%
-  ungroup() %>% mutate(AOI = "All") %>% 
-  select(AOI, Iteration, Metric, active, random, delta)
-imetrics <- bind_rows(imetrics, imetrics_all)
-
-# select out final iteration for each AOI
-mx_metrics <- imetrics %>% group_by(AOI) %>% 
-  filter(Iteration == max(Iteration)) %>% filter(Metric %in% V1)
-
-# plot
-cols <- c(rep("grey50", length(unique(imetrics$AOI)) -1), "black")
-lsize <- c(rep(0.3, length(unique(imetrics$AOI)) -1), 1)
-p <- ggplot(imetrics) + 
-  geom_hline(yintercept = 0, linetype = "dotted", col = "grey40") +
-  geom_line(aes(Iteration, delta, color = AOI, size = AOI)) +
-  geom_point(data = mx_metrics, aes(Iteration, delta, color = AOI), 
-             size = 1.2) +
-  ggrepel::geom_text_repel(
-    data = mx_metrics %>% filter(AOI != "All"), 
-    aes(label = AOI, x = Iteration, y = delta), col = "grey50", 
-    segment.color = "grey60", point.padding = 0.2, min.segment.length = 0.2,
-    segment.size = 0.2, size = 3, nudge_x = 0 
-  ) +
-  scale_color_manual(values = cols, guide = FALSE) +
-  scale_size_manual(values = lsize, guide = FALSE) + 
-  scale_x_continuous(expand = c(0.01, 0.01)) + 
-  scale_y_continuous(expand = c(0.01, 0.01), breaks = seq(-8, 4, 1)) + 
-  ylab("% difference") +                    
-  facet_wrap(vars(Metric), ncol = 3) +
-  theme_linedraw() + 
-  theme(panel.grid.minor.x = element_blank(), 
-        strip.background = element_rect(fill = "white"), 
-        strip.text = element_text(colour = "black"))
-
-ggsave(pth("figures/si_random_vs_active.png"), 
-       width = 6, height = 4, units = "in", dpi = 300)
-
-```
-
-### Figure S11: Training data error impact on probability images
+### Figure S12: Training data error impact on probability images
 ```{r, eval = FALSE}
@@
 
-### Figures S12 and S13: Segmentation statistics (grid scale)
+### Figures S13 and S14: Segmentation statistics (grid scale)
 
@@
 
-### Figure S14: Average fields sizes
+### Figure S15: Average fields sizes
 

### 2021-07-20 inst/manuscript/manuscript.Rmd
Fully revised manuscript (except abstract), SI, and figures.

@@
 Amidst all the challenges posed by global change, a particular concern is how agricultural systems will adapt to meet humanity's growing food demands, and the impacts that transforming and expanding food systems will have on societies, economies, and the environment [@searchingerCreatingSustainableFood2019].
-Significant efforts are being made to address the various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g.
+A number of efforts are underway to address various aspects of this challenge, including work on diagnosing and closing yield gaps [e.g.
 @LickerMindgaphow2010; @lobellCropYieldGaps2009; @muellerClosingYieldGaps2012], expanding and commercializing production [@morrisAwakeningAfricaSleeping2009], and to understand [@davisTropicalForestLoss2020; @kehoeNatureRiskFuture2017; @RulliFoodappropriationlarge2014] and mitigate [@EstesReconcilingagriculturecarbon2016] agriculture's ecological impacts.
-Answering many of the questions these efforts seek to address depends on reliable data that describes the location and characteristics of cropland [@FritzMappingglobalcropland2015], and how these are changing over time, ideally on an annual basis, given the rapid pace of change [@bullockThreeDecadesLand2021; @GibbsTropicalforestswere2010].
-Unfortunately, the data that do exist are in many places inaccurate and infrequently updated.
-Existing estimates of how much global cropland there is tend to vary widely, and they often disagree about where cropland is located [e.g.
-@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013].
-Such errors can propagate in subsequent analyses that rely on cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018].
-Beyond cropland distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016; @levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
-
-These informational inadequacies are due to the fact that cropland data in much of the world are derived from remotely sensed landcover maps, which can be notoriously high in error, particularly over regions such as Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021].
-Cropland mapping over Africa is difficult for several reasons.
-The primary reason relates to the characteristics of the continent's smallholder-dominated croplands, where half of all fields are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019].
+The success of these efforts depends heavily on data that accurately describes the location and characteristics of croplands [@FritzMappingglobalcropland2015], and, given the rapid pace of agricultural change [@bullockThreeDecadesLand2021; @GibbsTropicalforestswere2010; @ZengHighlandcroplandexpansion2018], how these are changing from one year to the next.
+Unfortunately, for many regions, existing cropland datasets are inaccurate, and are usually created as once-off or infrequently update products.
+As such, estimates of global cropland area tend to vary widely, often disagree about where croplands are located [e.g.
+@FritzHighlightingcontinueduncertainty2011; @Fritzneedimprovedmaps2013], and become rapidly outdated.
+Errors in these maps can propagate in subsequent analyses that use cropland data as inputs, resulting in potentially misleading answers [@Esteslargeareaspatiallycontinuous2018].
+Beyond distributions, few data are available on key cropland characteristics such as field size, an important variable needed to estimate yield and other key food security variables [@CarlettoGuesstimatesGPStimatesLand2015], and as an indicator of farm size [@SambergSubnationaldistributionaverage2016; @levinFarmSizeLandscape2006], a critical component of rural livelihoods given increasing population densities and longstanding debates about the relationship between farm size and productivity [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
+
+The deficit of information is due to the fact that in many regions the only source of cropland data are remotely sensed land cover maps, which are prone to error.
+This is particularly true in Africa [@Esteslargeareaspatiallycontinuous2018; @FritzComparisonglobalregional2010], where agricultural changes will be largest and the need for accurate baseline data is thus greatest [@EstesReconcilingagriculturecarbon2016; @SearchingerHighcarbonbiodiversity2015; @bullockThreeDecadesLand2021], and where the characteristics of croplands exacerbate the error inherent in remote sensing analyses.
+Half of all fields in Africa's smallholder-dominated agricultural systems are smaller than 1 ha [@lesivEstimatingGlobalDistribution2019].
 This size is small relative to the 30-250 m resolution of the sensors typically used in many landcover mapping efforts [e.g.
-@ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979; @boschettiAnalysisConflictOmission2004]].
-In the latter case, the pixel's shape may be poorly matched to that of cropland, and is too coarse to aggregate to approximate that shape at the characteristic scales of crop fields [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
-On top of the matter of scale is 1) high intra-class variability of the cropland class, compounded by the fact that these particular croplands can be heavily intergraded with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and 2) the substantial temporal variability within croplands, both within and between seasons.
-These latter two aspects pose challenges for the classification algorithms that are applied to the imagery.
+@ChenGloballandcover2015; @sulla-menasheHierarchicalMappingAnnual2019], which results in errors due to mixed pixels and aspects of the modifiable area unit problem [@Openshawmillioncorrelationcoefficients1979; @boschettiAnalysisConflictOmission2004], wherein the pixel's shape does not match that of crop fields, and is too coarse to aggregate into an approximation of that shape [@Darkmodifiablearealunit2007; @Esteslargeareaspatiallycontinuous2018].
+On top of the matter of scale is the high variability within and between fields, their tendency to intergrade with surrounding vegetation [@Debatsgeneralizedcomputervision2016; @Estesplatformcrowdsourcingcreation2016], and the high temporal variability within croplands.
+These last three aspects pose challenges for the classification algorithms that are applied to the imagery.
 
@@
 Chief among these are the growing numbers of satellites that collect high (\<5 m) to near-high (10 m) resolution imagery at sub-weekly intervals [@DruschSentinel2ESAOptical2012; @McCabefutureEarthobservation2017].
-This high spatial *and* temporal resolution imagery addresses the sensor-field scale mismatch, and more effectively captures the intra-seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
-On top of this, advances in cloud computing and the opening of satellite image archives [@WulderglobalLandsatarchive2016], and next generation machine learning approaches are placing large volumes of these moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale [@GorelickGoogleEarthEngine2017].
-These capabilities are already being used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017; [@zhangGLCFCS30Global2021a]].
+The spatial resolution of these imagery addresses the scale mismatch between sensor and field, and their high frequency captures the seasonal dynamics of cropland, which helps classifiers distinguish cropland from surrounding cover types [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+On top of this, the opening of satellite image archives [@WulderglobalLandsatarchive2016] and advances in cloud computing are placing large volumes of moderate to near-high resolution imagery together with the computational and algorithmic resources necessary to classify them at scale [@GorelickGoogleEarthEngine2017].
+These capabilities have already been used to create a new generation of higher resolution (10-30 m) cropland and landcover maps for Africa and other regions [@ESAESACCILAND; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017; [@zhangGLCFCS30Global2021a]].
 However, the potential of the highest resolution (\<5 m) imagery to map cropland over very large extents (e.g.
@@
 Beyond the imagery and computational gains, machine learning algorithms are rapidly advancing, providing large gains in classification performance [@maDeepLearningRemote2019; @MaxwellImplementationmachinelearningclassification2018].
-However, the ability to take advantage of these gains is often limited by newer models' need for large training datasets, which are typically unavailable, hard to collect, or contain numerous errors in developing regions [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @burkeUsingSatelliteImagery2021].
-To build sufficient samples for training cropland mapping models, as well as the reference data needed to objectively assess their performance (we refer collectively to both sample types as "labels", distinguishing between each type where needed), map-makers rely heavily on visual interpretation of high resolution satellite or aerial imagery [e.g.
+However, the ability to take advantage of these gains is often limited by newer models' need for large training datasets, which are typically unavailable, hard to collect, or contain numerous errors [@ElmesAccountingtrainingdata2020; @maDeepLearningRemote2019; @burkeUsingSatelliteImagery2021].
+To build sufficient training samples, as well as the reference data needed to objectively assess their performance (we refer collectively to both types as "labels", distinguishing between each as needed), map-makers rely heavily on visual interpretation of high resolution satellite or aerial imagery [e.g.
 @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019], as it is impractical and expensive to collect these data in the field over large areas, particularly on an ongoing basis.
-Several web-based platforms have been developed to facilitate such efforts, which provide convenient and highly scalable tools for training data collection [e.g.
+As such, a number of web-based platforms have been developed to collect such labels [e.g.
 @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012].
-Visually interpreted labels present two particular problems.
-The first is that they inevitably contain errors of interpretation that can vary substantially according to the skill of the labeller, particularly over complex croplands with small field sizes [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019].
-The second problem is that visual interpretation depends on high resolution imagery (\<5 m), as field boundaries become increasingly difficult to discern as image resolution decreases.
-Typically the only practical source of high resolution imagery is "virtual globe" basemaps (e.g.
-Bing or Google Maps), which are composed of mosaics of various high resolution satellite and aerial images that typically span 3-5 years of time within a single country [@LesivCharacterizingspatialtemporal2018].
-This within-mosaic temporal variation can set up a temporal mismatch between the imagery being interpreted and the imagery being classified, which is usually from a different source (e.g.
+Image-drawn labels present two particular problems.
+The first is that they inevitably contain errors of interpretation, which can vary substantially according to the skill of the labeller, particularly over complex croplands with small field sizes [@Estesplatformcrowdsourcingcreation2016; @WaldnerConflationexpertcrowd2019].
+The second problem is that visual interpretation depends on high resolution imagery (\<5 m), as fields are increasingly difficult to discern as image resolution decreases.
+Typically the only available source of high resolution imagery is "virtual globe" basemaps (e.g.
+Bing or Google Maps), which present mosaics of high resolution satellite and aerial images collected over a span of several years [@LesivCharacterizingspatialtemporal2018].
+This within-mosaic temporal variation can create a temporal mismatch between the labels and the imagery being classified, which is usually from a different source (e.g.
 Landsat, Sentinel; @XiongNominal30mCropland2017).
@@
 a new field was created), the label, even if accurately drawn, introduces error into the classifier.
-This source of error may be elevated in tropical smallholder systems, where swidden agricultural is common [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
-Despite the high potential for label error, they are typically not accounted for during model training and map accuracy assessment, resulting not only in the potential for maps to be misused or misinterpreted in subsequent analyses,  but in missed opportunities to improve model performance [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018; @StehmanKeyissuesrigorous2019].
+This source of error may be elevated in croplands where swidden agriculture is practiced [@vanvlietThereContinuingRationale2013], or in rapidly developing agricultural frontiers [@ZengHighlandcroplandexpansion2018].
+Despite the high potential for it, label error is often not considered during model training and map accuracy assessment, resulting not only in the potential for maps to be misused or misinterpreted, but in missed opportunities to improve model performance [@ElmesAccountingtrainingdata2020; @Esteslargeareaspatiallycontinuous2018; @StehmanKeyissuesrigorous2019].
 
@@
 This ability to capture intra-annual variability can be more important for classifying cropland than spectral depth [@Debatsgeneralizedcomputervision2016].
-Beyond the frequency, PlanetScope's 3.7 m resolution--although substantially coarser the 0.5-1 m available in most areas covered by virtual globes--is sufficiently resolved for humans to discern small fields under many conditions [e.g.
+Beyond the frequency, PlanetScope's 3.7 m resolution--although substantially coarser than the 0.5-1 m imagery available in most areas covered by virtual globes--is sufficiently resolved for humans to discern small fields under many conditions [e.g.
 see @Esteslargeareaspatiallycontinuous2018; @FourieBetterCropEstimates2009].
@@
 To further reduce label noise, we developed a platform that includes rigorous label accuracy assessment protocols and a novel approach for creating consensus labels, which helps reduce mistakes made by individual labellers [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020].
-We couple the labelling platform with a machine learning model inside an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] framework, in which model training is done iteratively and interactively, wherein the model's prediction uncertainty over unlabelled areas is assessed after each training step, and a sample of the most uncertain sites is selected for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011].
+We couple the labelling platform with a machine learning model inside an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] framework, in which the model is trained interactively, using the model's prediction uncertainty over unlabelled areas to select new sites for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011].
 This approach helps boost the performance of the classifier while reducing the overall number of labels required to achieve a given level of performance [e.g.
 @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017].
-Finally, an unsupervised segmentation step is applied to convert pixel-wise cropland predictions into vectorized maps of individual field boundaries.
-
-Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, ranging from primarily grain and vegetable crop production in the northern savannas to tree crop-dominated systems in the forested southwest, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
+An unsupervised segmentation step is then applied to convert pixel-wise cropland predictions into vectorized maps of individual field boundaries.
+
+Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates across a broad mix of climate and agricultural systems, ranging from primarily grain and vegetable crop production in the northern savannas to tree crop-dominated systems in the forested southwest, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016].
 The map represents a single agricultural year (2018-2019), as opposed to a multi-year epoch, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size.
-In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, we provide a new technique for converting daily imagery into seasonal composites, and show how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].
+In addition to providing valuable new data and insight into Ghana's agriculture, our study demonstrates one of the most spatially extensive agricultural applications of CubeSats to date, provides a new technique for converting daily imagery into seasonal composites, and shows how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].
 
@@
 The second is a custom-built platform that provides tools for labelling the composites, along with procedures to assess and minimize label error.
-This platform interacts with the third component, a machine learning process, in an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] loop, to produce a map of predicted cropland probabilities for each image pixel.
+This platform interacts with the third component, a machine learning process, within an active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011] loop, to produce a map of predicted cropland probabilities for each image pixel.
 The fourth and final component is an algorithm that segments the image composites, then filters the resulting polygons using the pixel-wise cropland predictions produced by the active learning classifier, resulting in a final set of vectorized field boundaries.
@@
 The image processing component was designed for PlanetScope Analytic surface reflectance imagery [@planetteamPlanetApplicationProgram2018], which provides three visual (red, green, blue) and near-infrared bands at 3.7 m resolution at nominal daily frequency.
-The images are provided as ortho-rectified and converted to surface reflectance, although there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016] that are not fully captured by the provided cloud masks.
+The images are provided as ortho-rectified and converted to surface reflectance, although there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016; @royGlobalAnalysisTemporal2021] that are not fully captured by the provided cloud masks.
 To minimize the effect of these residual errors, we developed a procedure for creating temporal composites of the primary growing and non-growing seasons within a single 12-month period.
 For Ghana, we defined the primary growing season as May through September, followed by the off (or dry) season from November or December through February.
-We chose these two seasons because prior work shows that the contrast improves cropland classifications [@Debatsgeneralizedcomputervision2016], while having the seasons captured in sequence in the same year helps minimize differences caused by land change.
-The wide time intervals we used to define each season were necessary for collecting a sufficient number of images to make high quality composites, as Ghana's cloud cover renders many scenes unusable (and thus unavailable in Planet's catalog), thus the frequency of available imagery can be substantially less than daily during several months.
-
-We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (beginning in May and ending in February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3).
-We then converted each scene into analysis ready  data [@DwyerAnalysisReadyData2018], by cropping each to the boundaries of a 0.05$^\circ$ grid that it intersected (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for making composite image tiles.
-We chose this cell size for tiling because it is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, helping to minimize edge artifacts in the composites.
-
-To create a seasonal composite tile, we calculated two weights for the time series of each pixel within the ARD stack for a given season:
+We chose these two seasons because prior work shows that the contrast between them improves cropland classifications [@Debatsgeneralizedcomputervision2016], Furthermore, capturing the seasons in this sequence during the same year helps minimize differences caused by land change.
+The wide time intervals we used to define each season were necessary for collecting a sufficient number of images to make high quality composites, as Ghana's cloud cover renders many scenes unusable and therefore unavailable in Planet's catalog, thus the effective return interval can be substantially longer than 24 hours during the cloudiest months [@royGlobalAnalysisTemporal2021].
+
+We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (defined here as March, 2018-February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3).
+We then converted each scene into analysis ready data [@DwyerAnalysisReadyData2018] by cropping each to the boundaries of a 0.05$^\circ$ grid that it intersected (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for making composited image tiles.
+We chose this cell size for tiling because it is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, thereby helping to minimize edge artifacts in the composites.
+
+To create a seasonal composite, we calculated two weights for the time series of each pixel within the ARD stack for a given season:
 
@@
 Where *t* is a particular date in the pixel time series, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band.
-Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
+Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to these atmospheric features [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]
 
@@
 
-We generated a catalog of `r nrow(tile_key) * 2` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana.
-To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using a four category scoring system, which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image that had its resolution degraded below the typical 3-4 m PlanetScope resolution (e.g.
+We generated a catalog of `r as.integer(nrow(tile_key) * 2)` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana.
+To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using a four category that evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image with resolution degraded below the 3.7 m PlanetScope resolution (e.g.
 because of between-date image mis-registrations).
@@
 The first step in creating a country-wide field boundary map of Ghana was to create a pixel-wise classification of cropland probabilities throughout the country.
-Given the high resolution of the imagery, we divided Ghana into 16 distinct mapping regions, or Areas of Interest (AOIs), to minimize the computational burden.
-We constructed the AOIs by grouping together tile grids into blocks representing the larger 1$^\circ$ cells used to assign tile grid identifiers (Figure S1A).
-We grouped tile grids from 1$^\circ$ degree cells that overlapped Ghana's boundaries together with tile grids from the adjacent 1$^\circ$ cell contained entirely within Ghana, with the exception was AOI 16, comprised of tile grids from the 1$^\circ$ cells intersecting Ghana's southern coast.
-The resulting AOIs ranged from 12,160 to 23,535 km$^2$ in extent (average = 15,457 km$^2$).
+Given the high resolution of the imagery and the need to minimize the computational burden, we divided Ghana into 16 distinct mapping regions, or Areas of Interest (AOIs).
+We constructed the AOIs by grouping together tile grids into blocks representing the larger 1$^\circ$ cells used to assign tile identifiers (Figure S1A).
+We grouped tile cells from 1$^\circ$ degree cells that overlapped Ghana's boundaries together with those from the nearest 1$^\circ$ cell contained entirely within Ghana (with the exception of AOI 16, which was comprised of tile grids from the 1$^\circ$ cells along Ghana's southern coast.
+The average extent of the resulting AOIs was 15,457 km$^2$ (range 12,160-23,535 km$^2$).
 
@@
 This finer grid, which we refer to as the "primary grid" for simplicity, provided the target area for creating labels (section 2.2.1), as well as the unit for distributing computing jobs (section 2.2.2).
-We then assessed the performance of the starter model against a separate set of validation labels developed for each AOI.
-The trained model was then applied to predict cropland probabilities for pixels in unlabelled primary grid cells in each AOI, and an uncertainty criterion [@DebatsIntegratingactivelearning2017] was calculated:
+We then assessed the performance of the starter model against a separate set of validation labels developed for each AOI, applied the model to predict cropland probabilities for pixels in unlabelled primary grid cells in each AOI, and calculated an uncertainty criterion [@DebatsIntegratingactivelearning2017]:
 
@@
 
-Where Q is the uncertainty for each unlabelled primary grid cell I, calculated from the predicted probability *p* of the randomly selected subset of pixels (x, y) drawn from it.
-Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus the lowest values of $Q$ represent locations posing the most difficulty for the classifier.
-
-We ranked the unlabelled primary grid cells in from least to most certain, and randomly selected a subset cells from the top 30% of the ranking (to minimize the risk of spatial autocorrelation) and sent these back to the labelling platform.
-After these new sites were labelled, they were added to the starter pool of labels, the model was retrained with the larger training set, its performance and prediction uncertainty was reassessed, and a new sample of the most uncertain primary grid cells was sent for labelling.
-This loop repeated until model performance gains saturated or reached a threshold, and then a final map of predicted cropland probabilities was made for each AOI.
+Where Q is the uncertainty for each unlabelled primary grid cell I, calculated from the predicted probability *p* of a randomly selected subset of pixels (x, y) drawn from it.
+Pixels with predicted probabilities closer to 0.5 are least certain as to their classification, thus the lowest values of $Q$ represent primary grid cells posing the most difficulty for the classifier.
+
+We ranked the unlabelled primary grid cells from least to most certain, randomly selected a subset of cells from the top 30% of the ranking (to minimize the risk of spatial autocorrelation), and sent these back to the labelling platform.
+After these new sites were labelled, they were added to the starter pool of labels, the model was retrained with the larger training set, its performance and prediction uncertainty was reassessed, and a new sample of the most uncertain primary grid cells was again sent for labelling.
+This loop repeated until model performance gains saturated or reached a pre-defined threshold, after which a final map of cropland probabilities was made for the AOI.
 
@@
 
-We collected labels using a custom-built platform that we adapted from an earlier prototype used for crowdsourced labelling [@Estesplatformcrowdsourcingcreation2016].
+We collected all labels using a custom-built platform that we adapted from an earlier prototype we developed for crowdsourced labelling [@Estesplatformcrowdsourcingcreation2016].
 We enhanced this platform by making several major additions, including an independent backend that allowed us to recruit and manage our own labelling teams, improved procedures for assessing and improving label accuracy, and processes for automating the machine learning component.
-The platform runs on a cloud-hosted Linux virtual server (AWS EC2) and is comprised of a database (PostGIS/Postgres), a mapping interface with a full set of labelling tools (OpenLayers 3), an image server (Raster Foundry), and a set of utilities for managing, assessing, and converting digitized field boundaries into rasterized labels.
-We created a separate labelling instance to manage the active learning process for each AOI.
-To create training and validation labels, labellers (the co-authors of this paper) would log into  the website (built with Flask) for a particular AOI and navigate to the mapping interface (Figure \ref{fig:labeller}), where they would be presented with a white target box representing a primary grid cell to label, a set of digitizing tools, and several different sources of imagery.
+The platform runs on a cloud-hosted Linux virtual server (AWS EC2) and is comprised of a database (PostGIS/Postgres), a mapping interface (OpenLayers 3), an image server (Raster Foundry), and a set of utilities for managing, assessing, and converting digitized field boundaries into rasterized labels.
+
+We created a separate labelling instance for each AOI.
+To create training and validation labels, labellers (the co-authors of this paper) logged into the website (built with Flask) for a particular AOI and navigated to the mapping interface (Figure \ref{fig:labeller}), where they were presented with a white target box representing a primary grid cell to label, a set of digitizing tools, and several different sources of imagery.
 These included true and false color renderings of the growing season and dry season PlanetScope composites, and several virtual globe basemaps.
@@
 The labeller assigned each digitized polygon a class category (e.g.
-annual cropland), and upon completing all fields saved the fields to the database, and were then presented with the next target to label.
-If the target grid cell did not contain any fields, the labeller simply pressed save to go to the next cell.
+annual cropland), saved all completed fields to the database, and were then presented with the next target to label.
+If the target grid cell did not contain any fields, labellers simply pressed save to go to the next cell.
 
@@
 There were two types of tasks, accuracy assessment or model training/validation, with the assignments for each indistinguishable to labellers.
-Upon completing an accuracy assessment assignment, the platform invoked an label scoring algorithm that compared the labeller's digitized boundaries against a set of training reference polygons, resulting in an assignment quality score:
+Upon completing an accuracy assessment assignment, the platform invoked a scoring algorithm that compared the labeller's digitized boundaries against a set of training reference polygons, resulting in a label quality score:
 
@@
 
-We configured the platform's scheduler to present workers with accuracy assessment assignmetns at a rate of 1 for every 5 assignments mapped.
+We configured the platform's scheduler to present workers with accuracy assessment assignments at a rate of 1 for every 5 assignments mapped.
 This generated a history of accuracy assessment scores that we used to assess label quality and minimize label error.
@@
 each was completed by four separate labellers.
-When all four assignments were complete, a consensus labelling routine was invoked to combine the four sets of labels into a single consensus label using a Bayesian merging approach:
+When all four assignments were complete, a Bayesian merging routine was invoked to combine the four sets of labels into a single consensus label:
 
@@
 P($\theta$\|D) is the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is the average score (ranging between 0 and 1) of the accuracy assessment assignments an individual labeller completed within the AOI, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
-This approach therefore used the overall accuracy of each labeller to weight their label for a given pixel when combining it with those made by other labellers' for the same pixel (see SI for further details).
+This approach therefore used the average assignment quality score for to weight each labeller's label for a given pixel (see SI for further details).
 Each pixel in the target grid cell was merged using this approach (n = 40000), which helps to minimize individual labellers' errors.
-We estimated a confidence measure for each consensus label by  calculating its Bayesian Risk (see SI), which ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels in the label, and 1 indicating complete disagreement.
+We estimated a confidence measure for each consensus label by  calculating its Bayesian Risk (see SI), which ranges between 0 and 1, with 0 indicating full agreement between labellers for all pixels, and 1 indicating complete disagreement.
 
@@
 Previous work showed that a large number of simple features summarizing image reflectance and vegetation indices within local neighborhoods were highly effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016].
-We followed that logic in this study, but used a smaller feature space because the storage and memory required for our mapping geographies were several orders of magnitude larger than in the previous study.
-We therefore extracted a set of 8 features from each seasonal composite, which were the mean calculated within an 11X11 moving applied to each band, and the standard deviation within a 5X5 moving window (initial tests revealed these two window sizes to be most effective).
+We followed that logic in this study, but used a smaller feature set because the storage and memory required for our mapping geographies were several orders of magnitude larger.
+For each seasonal composite, we calculated the mean and standard deviation of each band within an 11X11 and 5X5 moving window, respectively (initial tests revealed these two window sizes to be most effective).
 This provided an overall set of 24 features, including the unmodified bands of both composites (Table 1).
@@
 The second zonation was developed by grouping the country's eight agro-ecological zones into four broader clusters (Figure S4B).
-We confined use of this zonation to better understand patterns in the per-pixel classifications.
-
-To assess the accuracy of the segmented field boundaries, we compared the size class distributions of the segmented field boundaries against the field boundaries digitized by by the labellers' over validation sites (primary grid cells).
+We applied this zonation only to the per-pixel classification, to better understand patterns of error in the model.
+
+To assess how effectively the segmentations captured field characteristics, we compared the size class distributions of the segmented field boundaries against the field boundaries digitized by the labellers' over the 100 validation sites in each AOI.
 We chose this approach because of existing uncertainties in polygon-based accuracy assessment methods [@yeReviewAccuracyAssessment2018], and because the map's ability to represent field sizes was of greatest interest.
@@
 ```
-The assessment of image composites found that their quality in both seasons was highest in the northern half of the country and lowest in the southwest, (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S3).
+The assessment of image composites found that their quality in both seasons was highest in the northern half of the country and lowest in the southwest, (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S5).
 The average quality score of growing season composites was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85 (out of 1; Figure \ref{fig:imqual}B), while the mean score of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85).
@@
 ```{r imqual, echo = FALSE, out.width="70%", fig.cap="The location and quality scores of 100 randomly selected tiles for the growing (A) and off-growing season (B), and the corresponding distributions of the quality scores for each season, respectively (C and D).", fig.align='center'}
-knitr::include_graphics('figures/figure4.png')
+knitr::include_graphics('figures/figure3.png')
 ```
@@
 initial_total <- train_val_sites$initial %>% as_tibble() %>% distinct(name) %>% 
-  count()
+  count() %>% pull()
 active_total <- sum(active$n)
@@
 
-To make the initial maps of cropland probabilities, the active learning process ran for 3 iterations in 12 of 16 AOIs, varying from as little as 1 to as many as 4 iterations across the other 4 AOIs, with the number of iterations associated with the performance of the starter models against the validation labels (i.e.
+To make the initial maps of cropland probabilities, the active learning process ran for 3 iterations in 12 of 16 AOIs, varying from as little as 1 to as many as 4 iterations across the other 4 AOIs, with the number of iterations varying according to the performance of the starter models (i.e.
 AOIs with higher starting performance stopped after fewer iterations, see SI).
-A total of `r ttrain` training and `r tval` validation labels were collected by 20 labellers to develop and assess model performance.
-Each AOI's model was trained by the 500 randomly selected labels, plus an additional `r aoi10n` - `r aoi15n` (typically `r typicaln`) labels within each the AOI that were selected by active learning.
-Actively selected labels showed distinctive patterns in several AOIs, such as concentrating along ecotones, including the boundaries between agro-ecozones (see Figure S4A).
+Each AOI's model was trained by 300-500 randomly selected labels (Figure S6A), plus an additional `r aoi10n` - `r aoi15n` (typically `r typicaln`) within each the AOI that were selected by active learning.
+Actively selected labels showed distinctive patterns in several AOIs (Figure S6B), such as concentrating along ecotones or the boundaries of agro-ecological zones.
+A total of `r ttrain` training and `r tval` validation labels were collected by 20 labellers to develop and assess model performance (Figure S7).
 
@@
 ```{r alperformance, echo = FALSE, out.width="100%", fig.cap="Scores for overall accuracy, area under the curve of the Receiver Operating Characteristic, and the F1 scores for the Random Forests model results after each iteration of the active learning loop for each AOI (gray lines), as well as the mean score per iteration across all AOIs (black lines).", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/figure5.png')
+knitr::include_graphics('figures/figure4.png')
 ```
@@
 
-The experiment conducted in three AOIs (in AOIs 1, 8, and 15) showed that training models with active learning improved performance compared to random approaches to training label selection.
+The experiment conducted in three AOIs (in AOIs 1, 8, and 15) showed that training models with active learning improved performance compared to randomized approaches to label selection.
 After three iterations, the accuracy, AUC, and F1 scores for the actively trained models were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than those for randomly trained models (Figure S8).
@@
 
-### The impact of label error on model performance
-We used the two measures of label quality calculated by the platform, the average accuracy score of each labeller and Bayesian Risk (or simply "label risk"), to assess the potential impacts of label error and uncertainty on model performance.
+### The impact of label error and uncertainty on model performance
+We used the two measures of label quality calculated by the platform, the average quality score of each labeller and Bayesian Risk (or simply "label risk"), to assess the potential impacts of label error on model performance.
 The average of each labeller's AOI-specific accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figures S4 and S5 for details on label scores and number of assignments per labeller).
@@
 
-Treating average labeller accuracy as a proxy for label error, we used these scores to select the labels used in the experiments to assess how label error impacted model performance.
+Treating each labeller's average label quality scores (Figure S9) as a proxy for error, we used these scores to develop training sets to test the impact of label error on model performance.
 The results of these tests, which were conducted in AOIs 1, 2, 8, and 15, showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
@@
 Comparisons were made for AOIs 1, 2, 8, and 15, denoted by grey symbols, while the mean scores across these AOIs are shown for each metric.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/figure6.png')
+knitr::include_graphics('figures/figure5.png')
 ```
@@
 
-Correlations between the mean label risk per AOI and model performance metrics showed strong (Spearman's Rank Correlation = `r risk_corr[, "Accuracy"]`) to moderate (r = `r risk_corr[, "AUC"]`) negative correlations between label risk and accuracy and AUC, respectively, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
-The positive sign of the latter relationship is counter-intuitive, but is explained by risk's association with precision, one of two inputs to F1, which was moderately positive (`r risk_corr[, "Precision"]`), whereas risk had a negligible correlation with recall (`r risk_corr[, "Recall"]`), F1's other component.
-The correlation between risk and the false positive rate (`r risk_corr[, "FPR"]`), another important performance metric, shows that labelling uncertainty may increase model commission error.
+Correlations (Table S3) between the mean label risk per AOI (Figures S10-11) and model performance metrics showed strong (Spearman's Rank Correlation = `r risk_corr[, "Accuracy"]`) to moderate (r = `r risk_corr[, "AUC"]`) negative correlations between label risk and accuracy and AUC, respectively, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
+The positive sign of the latter relationship is counter-intuitive, but is explained by risk's association with precision, one of two inputs to F1, which was moderately positive (r = `r risk_corr[, "Precision"]`), whereas risk had a negligible correlation with recall (r = `r risk_corr[, "Recall"]`), F1's other component.
+The correlation between risk and the false positive rate (r = `r risk_corr[, "FPR"]`), another important performance metric, shows that labelling uncertainty may increase model commission error.
 
@@
 ### Categorical accuracy
-We used the map reference sample to evaluate the accuracy of the cropland probability map (after converting to a classification using a threshold probability of 0.5) and the map of segmented field boundary maps.
+We used the map reference sample to evaluate the accuracy of the cropland probability map (after classifying it using a threshold probability of 0.5) and the map of segmented field boundary maps.
 We found that the overall accuracy of the pixel-wise classifications was 88% against this map reference sample (Table \ref{tab:mapaccuracy}).
-Confining the map reference sample to four distinct zones (Figure S10A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
-The Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1, while the User's accuracy for was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1.
+Confining the map reference sample to four distinct zones (Figure S4A) shows that overall accuracy ranged from 83.3% in Zone 1 (AOIs 1-3) to 93.6% in Zone 3 (AOIs 10, 11, 13, 15, and 16).
+The Producer's accuracy of the cropland class was 61.7% across Ghana, ranging from 45.6% in Zone 3 to 67.9% in Zone 1, while the User's accuracy was 67.3% overall, ranging from 59.8% in Zone 4 to 71.2% in Zone 1.
 Both measures of accuracy were substantially higher for the non-cropland class across all zones, typically exceeding 90%.
@@
 ```
-### Segmentation accuracy
+### Segmentation quality
 
@@
 The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
-This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S11), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
-The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S12).
+This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S13), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
+The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S14).
 
 ## Ghana's croplands
-
 Two separate maps of cropland were produced for each AOI, a per-pixel map derived from the cropland probabilities, and the vectorized map of field boundaries (Figure \ref{fig:mainmap}).
@@
 The northern third of AOI 1 also has noticeable densities of cropland.
-Several notable areas of low cropland density are also apparent, indicating the presence of large protected areas, such as Mole National Park in the southeastern corner of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
-In contrast, the relative absence of cropland in AOIs 13, 14, and 16 does not reflect the scarcity of agriculture in these areas, but rather the predominance of tree crops, which we did not map.
+Several prominent areas of low cropland density indicate the presence of large protected areas, such as Mole National Park in the southeastern corner of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
+The relative absence of cropland in AOIs 13, 14, and 16 does not reflect the scarcity of agriculture in these areas, but rather the predominance of tree crops, which we did not map.
 
@@
 Each tile's position is shown on the main map, and is color-coded to the boundary lines around its corresponding inset.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/figure7.png')
+knitr::include_graphics('figures/figure6.png')
 ```
@@
 
-Both the per-pixel and vectorized maps, when combined with the map reference sample, enable separate estimates of the total extent of croplands in Ghana.
-The cropland extent estimated from the vectorized map is `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area.
-The estimate based on the per pixel map is `r fmt(rfas[2])` km$^2$ (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.
-
-The vectorized map provides additional information on how the characteristics of croplands can vary geographically, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
-To explore how field characteristics varied geographically, we mapped the average field size and total number of fields within each 0.05 degree tile (Figure S13).
-These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts generally occurring in areas of higher field density, although the biases inherent in both measures (Figures S11-12) complicate the interpretation of those variations.
-However, we can use the estimated biases to develop adjusted estimates of field sizes and counts for each AOI, and for Ghana overall (Table \ref{tab:sizentab}).
+Using the map reference sample and each map, we made two separate estimates of the total cropland area in Ghana in 2018.
+The cropland extent estimated from the field boundary map was `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area.
+The estimate from the per pixel map was `r fmt(rfas[2])` km$^2$% (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.
+
+The field boundary map provides additional information on how the characteristics of croplands vary across Ghana, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
+To explore how field characteristics varied geographically, we mapped the average size and total number of fields within each 0.05 degree tile grid (Figure S15).
+These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts occurring where field densities are higher, although the biases (relative to the validation labels) in both measures (Figures S13-14) complicate interpretations of those variations.
+To minimize this complication, we used the calculated biases to develop adjusted estimates of field size and count (Table \ref{tab:sizentab}).
 These adjusted estimates show that the typical field size in Ghana is 1.73 ha, ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern half of the country (Table \ref{tab:sizentab}).
-The total number of fields was estimated to be 1,662,281 overall, or 205 fields per tile on average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.
+The estimated total number of fields is 1,662,281, or 205 fields per tile grid cells, varying from 108 fields/tile cell in AOI 4 to 399 in AOI 6.
 
@@
 
-These results demonstrate a capability for mapping the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales.
-The resulting maps provide an updated and more granular view of the distribution and extent of croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution sensors [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
-This prior work found that cropland covered `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2% (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimates finds `r segcipct[1]`-`r rfcipct[1]`% cover.
-Our results thus suggest that Ghana's cropland is less than previously estimated, but the difference is perhaps attributable to our use of a cropland definition that excluded longer fallows and abandoned fields, which in some regions can account for over half of the total area that could be counted as cropland [@Tongforgottenlanduse2020].
-
-In addition to the more detailed update of cropland extent, our maps also provide new information on the size and number of fields in Ghana (Figures \ref{fig:mainmap}, S11-12).
-Previous work to estimate such agricultural characteristics have often focused on farm, rather than field, size using census data [@jayneAfricaChangingFarm2016; @LowderNumberSizeDistribution2016; @SambergSubnationaldistributionaverage2016; @VonBraunSmallscalefarmersliberalized2004].
-Efforts to map field boundaries in smallholder-dominated agricultural systems have either used *in situ* data collection [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small [e.g.
+These results demonstrate a capability to map the characteristics of smallholder-dominated cropping systems at high spatial resolution, annual time steps, and national scales.
+The resulting maps provide an updated and more granular view of the distribution and extent of croplands in Ghana, complementing existing national to regional land cover maps derived from moderate resolution imagery [@ESAESACCILAND; @hackmanNewLandcoverMaps2017; @XiongNominal30mCropland2017].
+Those prior studies estimated that cropland covers `r round(4.62 / 23.86 * 100, 1)` [@XiongNominal30mCropland2017] to 32% [@hackmanNewLandcoverMaps2017] of Ghana in 2015.
+In contrast, our 2018 maps provide a raw estimate of 16.1-23.2% cover (Table \ref{tab:mapaccuracy}), and our map reference sample-based estimate was `r segcipct[1]`-`r rfcipct[1]`%.
+Our results thus suggest that Ghana's croplands are less extensive than those previous estimates.
+However, this difference may arise from our use of a cropland definition that excludes longer fallows and abandoned fields, which in some regions may comprise over half of total cropland area [@Tongforgottenlanduse2020].
+
+In addition to this updated information on Ghana's cropland extent and distribution, our results provide new insights into field size and number at a national scale (Figures \ref{fig:mainmap}, S11-12).
+Previous efforts to map smallholder field boundaries have either used *in situ* data collection [@CarlettoFactartifactimpact2013; @CarlettoGuesstimatesGPStimatesLand2015] or remote sensing studies over relatively small [e.g.
 @forkuorIntegrationOpticalSynthetic2014; @perselloDelineationAgriculturalFields2019] or discontiguous [@Estesplatformcrowdsourcingcreation2016] areas.
-The most extensive studies to date used crowdsourced volunteers to classify fields into broad size classes, based on their interpretations of imagery sampled from high resolution virtual globes [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
-Those efforts included country-specific results for Ghana (n = 263), which can be converted into an average field size estimate of `r round(lesiv_est, 2)` ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class.
+The most extensive studies to date enlisted crowdsourced volunteers to classify fields visible within high resolution imagery sampled from virtual globes into broad size categories [@lesivEstimatingGlobalDistribution2019; @FritzMappingglobalcropland2015].
+Those efforts included country-specific results for Ghana (n = 263), which yield an average field size estimate of `r round(lesiv_est, 2)` ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided for the largest size class.
 Data sourced from Table S3 in Lesiv et al.
@@
 This estimate exceeds our Ghana-wide average segment size (3.92 ha; Table \ref{tab:sizentab}), but is closer to the mean (`r aoi_nofor` ha) within AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample appears to have been collected.
-However, our bias-corrected estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha were much smaller.
+However, our bias-adjusted estimates of 1.73 (Ghana-wide) and `r aoi_nofor_adj` (AOIs 1-9, 12, and 15) ha were much smaller.
 
 ## Map accuracy and key sources of error
-
-Although the maps generated by our system provide valuable new information, they nevertheless contain substantial errors.
-The overall map accuracies (86.7-88%, Table \ref{tab:mapaccuracy}) are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given the inherent uncertainty in the map reference sample, our best estimate of the "truth", in which we have roughly 85% confidence.
-However, accuracies for the cropland class were much lower, falling between 62 (Producer's) to 67 (User's) percent country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}), meaning the model produced substantial commission and omission errors for this class.
-The segmented boundary maps had fewer omission errors (Producer's accuracy = 79%), but higher false positive errors (User's accuracy = 58.2%).
+Although these maps provide valuable new information, they nevertheless contain substantial errors.
+The overall accuracies (86.7-88%, Table \ref{tab:mapaccuracy}) are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given that we only have ~85% confidence in our map reference sample, which is our best estimate of the "truth".
+However, accuracies for the cropland class were much lower, falling between 62 (producer's) to 67 (user's) percent country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}), meaning the model produced substantial commission and omission errors for this class.
+The segmented boundary maps had fewer omission errors (producer's accuracy = 79%), but higher false positives (user's accuracy = 58.2%).
 These accuracies are near the middle to upper ranges of those reported for the cropland class in other large-area mapping studies [@hackmanNewLandcoverMaps2017; @LesivEvaluationESACCI2017; @XiongNominal30mCropland2017].
 
-The patterns of cropland-class accuracies varied by zone.
-These zones largely align, albeit with some discrepancies, with the country's agroecozones, thus the accuracy patterns may be partially attributed to some regions being harder to map than others.
-Producer's accuracies for both maps were highest in the two northern zones (1 and 2), which are primarily savannas (Figure S10), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas.
-User's accuracies followed a similar pattern, with the exception of Zone 3, which had the highest User's accuracy, albeit from a very small sample.
-Aligning the reference samples more precisely with agroecozone boundaries (Figure S10B) provides further insight into error patterns (Table S4).
-Coastal savannas in the southeast had the highest Producer's accuracy but lowest User's accuracy for the per-pixel map, presumably because this region's numerous areas of high density cropland, combined with low woody cover in surrounding uncultivated areas, helped to promote commission error.
-Maps in the two northern savanna agroecozones had the best balance between omission and commission error, and had the highest overall User's accuracy.
-The transitional (between forest and savanna) agroecozone had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S4), within which it typically covered a smaller share of area relative to the other agroecozones.
-This likely caused insufficient representation of this AEZ in training samples, particularly in AOIs 10 and 11 (Figure S4B).
-
-Beyond the errors linked to regional differences, several other important factors contributed to reducing the accuracies in the cropland class.
-The first of these stems from the overall mapping extent and the high resolution of the data.
-Given the goal of developing a country-scale map at high resolution, the attendant data volume required us to use a relatively small set of image features and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forests implementation, in order to limit computational costs.
-Previous work found that Random Forests achieves much better performance on small-scale croplands when trained on a much larger number of features [@Debatsgeneralizedcomputervision2016].
-However, applying such a large feature set within the extent of our AOIs was intractable, as the computing time would have been several-fold larger than the \~4-8 hours of runtime on 800 CPUs required for a single active learning iteration, followed by \~10-14 hours for prediction.
-This reduced the skill of the model, particularly when it came to differentiating cropland from adjacent bare patches or natural vegetation with sparse herbaceous cover, which were common in many AOIs.
+The patterns of accuracies within the cropland class varied by zone.
+These zones largely align, albeit with some discrepancies, with the country's agro-ecological zones (AEZs), thus the accuracy patterns may be in part because some regions are simply more difficult to map.
+Producer's accuracy for both maps was highest in the two northern zones (1 and 2), which are primarily savannas (Figure S4), and lowest in zones 3 and 4, which are comprised of forest or coastal savannas.
+User's accuracy followed a similar pattern, with the exception of Zone 3, which had the highest user's accuracy, albeit from a very small sample.
+Aligning the reference samples more precisely with agroecozone boundaries (Figure S4B) provides further insight into error patterns within the per-pixel map's cropland class (Table S4).
+Coastal savannas in the southeast had the highest producer's and lowest user's accuracy, perhaps because this region has high density cropland inter-mixed with uncultivated areas that have low woody cover, which could help promote commission error.
+Maps in the northern savannas had the best balance between omission and commission error, and had the highest overall user's accuracy.
+The transitional zone between forest and savanna had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S4), and thus was under-represented in the training samples, particularly in AOIs 10 and 11 (Figure S6B).
+
+Beyond the errors linked to regional differences, several other important factors contributed to map error.
+The first of these related to the large mapping extent and image resolution.
+Given the goal of developing a high resolution, country-scale map, the large data volume constrained us to use a relatively small feature set and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forests models, in order to limit computational costs.
+Previous work found that Random Forests achieves much better performance on small-scale croplands when trained on a much larger number of features [@Debatsgeneralizedcomputervision2016;@lebourgeoisCombinedRandomForest2017].
+However, applying such a large feature set within the extent of our AOIs was not possible, as the computing time and costs would have been several times larger\footnote{Each active learning iteration ran for $\sim$4-8 hours on 800 CPUs, followed by a final $\sim$10-14 hours for prediction}.
+This reduced the skill of the model, particularly when it came to differentiating cropland from bare or sparsely vegetated patches, which were common in many AOIs.
 
 The inherent difficulty of the labelling task was another major limiting factor.
-Our system was designed to minimize the error inherent in labelling, but determining croplands from non-croplands in these agricultural systems can be a difficult task.
-Labellers have to evaluate multiple image sources and to rely heavily on judgement, which inevitably leads to errors.
-Interpretation is particularly hard where the background savanna vegetation and croplands have similar reflectance during the dry season, which is a particular problem in AOIs 2 and 3.
-Smaller field sizes also complicate labelling, as these become increasingly indistinct in the \~4 m PlanetScope composites.
-The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figure S6), and by the average score achieved by each labeller against our training reference dataset (71%; Figure S5B).
-Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label errors, we found that they have substantial impact (Figure \ref{fig:trainingimpact}), which suggest that simply improving label quality may be one of the single most important investments towards improved model accuracy.
-
-Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover reduced the number of available images in all seasons, resulting in composites with more brightness artifacts and blur (Figure \ref{fig:imqual}).
+Our platform was designed to minimize label errors, but determining croplands from non-croplands in these agricultural systems can be difficult.
+Labellers had to evaluate multiple image sources and to rely heavily on their judgment, which inevitably led to errors.
+Interpretation was particularly hard where croplands and surrounding landscapes had similar dry season reflectances, which was a particular problem in the northernmost savannas.
+Smaller field sizes also complicated labelling, as these become increasingly indistinct in the \~4 m PlanetScope composites.
+The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figures S10-11), and by the average assignment quality scores of each labeller (71%; Figure S9).
+Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label error, we found that it has substantial impact (Figure \ref{fig:trainingimpact}), which suggests that improving label quality is one of the most important factors in increasing model accuracy.
+Newer models, such as convolutional neural networks, may be less sensitive to label error, provided the error is random and the map reference samples are of high quality [@burkeUsingSatelliteImagery2021].
+However, over many smallholder systems training label errors will likely be biased in a particular direction (e.g.
+towards omission when fields are not easily distinguished from the background), and our results show that reference labels can have substantial uncertainty.
+
+Image quality was another issue, although primarily in the forested AOIs, where frequent cloud cover and the corresponding lower number of available images resulted in lower quality composites (Figure \ref{fig:imqual}), with more brightness artifacts and blur.
 This impacted labellers' abilities to discern fields, and doubtless affected model predictions.
-There is little to be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which may further undermine model performance, given the importance of multi-temporal imagery for cropland classification [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019].
+Little can be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which could reduce model performance by removing the temporal contrast [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019], or by adding radar data to the predictor set, which would reduce map resolution.
 Composite quality could be improved by using imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would blur field boundaries within the imagery.
@@
 The vectorized maps had high commission errors caused by uncertainties in the Random Forests predictions.
-Model uncertainty led to many pixels over non-crop areas with probabilities straddling the 0.5 classification threshold.
-Segments that intersected such areas were retained as fields when the average probability of intersecting pixels exceeded 0.5.
-A more accurate classifier would reduce such errors, or the application of a locally varying classification threshold [e.g.
+Model uncertainty meant that many pixels in non-cropland areas had probabilities with values near 0.5.
+Segments in these areas were retained if the average probability of intersected pixels exceeded the 0.5 classification threshold.
+A more accurate classifier would reduce such errors, as would a locally varying classification threshold [e.g.
 @waldnerDeepLearningEdge2020].
-Over-merging was another source of error in the segmentation algorithm, which in some areas led to overestimated field sizes and unrealistic shapes, particularly in high density croplands (e.g.
-in AOIs 2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent fields are often indistinct in the PlanetScope imagery.
-Minimizing or preventing merging would help in such cases, although could would result in the opposite problem, over-segmentation, and thereby underestimate field size.
+Over-merging was another source of error in the segmentation algorithm, which led to overestimated field sizes and unrealistic shapes in some areas, particularly in high density croplands (e.g.
+in AOIs 2 and 8; Figure \ref{fig:mainmap}) where boundaries between adjacent fields were indistinct in the imagery.
+Preventing merging could help in such cases, but potentially lead to over-segmentation, thereby underestimating field sizes.
 
 ## Error mitigation features
-
-Despite the error sources mentioned above, several features of the system proved effective in mitigating error, leading to a higher overall accuracy than would have otherwise been possible.
-Label accuracy assessment and consensus labelling appeared to be the most effective error mitigation tools.
-Label accuracy measures allowed us to quantify the substantial impact of label error on model performance (Figure \ref{fig:trainingimpact}), while consensus labels substantially reduced individual labelling errors, resulting in maps that were more accurate than they would have been had we used individually generated labels.
-Labeller-specific accuracy measures also helped to improve the quality of the consensus labels, by placing higher weight on labels more likely to be accurate during the merging process, rather than giving equal weight to potentially less accurate labels.
-The ability to select the most accurate individual labels for a site also allowed us to develop independent estimates of field size to which measures of confidence can be attached (Figure S5B), which we were in turn able to correct estimates of field sizes and numbers (Table \ref{tab:sizentab}).
+Despite these numerous source of errors, our approach was effective in mitigating several of these error sources.
+Label quality assessment and consensus labelling were the most effective error mitigation tools.
+Label quality scores allowed us to quantify the impact of label error on model performance (Figure \ref{fig:trainingimpact}), while consensus labels produced maps that were more accurate than they would have been if we had relied on individually generated labels.
+The quality scores also helped to improve the overall accuracy of consensus labels, by placing higher weight on the work of more accurate labellers.
+In addition to these benefits, label quality scores (Figure S9) also allowed us to select the labels most likely to accurately capture field sizes and numbers, which we used to estimated and correct the biases in these two measures derived from the segmented field boundaries.
 
@@
 
-The active learning approach helped to improve overall model performance relative to randomized training site selection, in line with findings from two recent efforts [@DebatsIntegratingactivelearning2017; @hamrouniLocalGlobalTransfer2021].
-Although the performance gains relative to randomized model training that we observed were smaller (e.g.
-@DebatsIntegratingactivelearning2017 `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher on the final iterations), those comparisons were made from lower initial bases, with initial training samples that were less than 1/10 the size, in terms of pixels, of our initial training sample.
-Our large initial randomly selected sample (500 grid cells) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance.
-Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness.
-Most notable were the larger improvements seen in the F1 score (Figure S8), a balanced performance metric.
-Gains in accuracy and AUC were smaller.
-For accuracy, the reduced advantage was primarily due to active learning being outperformed by randomized training after the first iteration in AOI 15, which proved one of the hardest AOIs to both map and label.
-Active learning likely resulted in the selection of sites that were harder to label than randomly selected ones, leading to more label error, and thus initially lower model accuracy.
-However, this deficit was overcome by the 5th iteration.
-The plateau in AUC gains at 0.5% better than randomized training reflects the findings that active learning reduced both the false and true positive rates, the two inputs to AUC.
-Although the decline of the false positive rate (`r prdelta$rate[1]`% between Iterations 0 and 3) was nearly three times larger than that of the true positive rate (`r prdelta$rate[2]`%), AUC should be quite sensitive to the reduction in the latter, as it assesses how the tradeoff between the two rates varies across a full range of possible classification thresholds [@Pontiustotaloperatingcharacteristic2014].
-
-The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
+Active learning improved overall model performance relative to randomized training site selection, in line with findings from two recent efforts [@DebatsIntegratingactivelearning2017; @hamrouniLocalGlobalTransfer2021].
+Although the relative performance gains that we observed were smaller (e.g.
+@DebatsIntegratingactivelearning2017 `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher on the final iterations), those comparisons were made by starting with a training sample that was <1/10 the size of ours.
+Our large starter sample meant that the models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance.
+Nevertheless, we found higher performance from active learning, most notably in the F1 score (Figure S8), a balanced performance metric, which further demonstrates its effectiveness.
+
+The detail, temporal precision, and large extent of our maps was enabled by the ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
 Daily revisits are important for creating seasonal composites within a single year over cloudy areas.
 The compositing technique we developed allowed us to develop a complete image catalog for the country representing the two seasons for 2018 agricultural year.
-Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
-@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval may be too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana.
-Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is challenging.
+Although Sentinel-2 is free, has better radiometric quality, and has sufficient spatial resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
+@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval may be too infrequent to generate cloud-free composites during the growing season over southern Ghana.
+Sentinel-1 is not affected by the same problem, but labeling fields in radar images is more challenging.
 
 ## Lingering questions
-
 Several potential issues not addressed in our assessment merit further exploration.
-One of these is the degree of correspondence between image- and ground-collected labels.
-However, such comparisons may reveal unresolvable differences between the two perspectives.
-The highly dynamic nature of many agricultural systems means that relatively narrow differences between the dates of ground- and image-based digitizing campaigns can lead to substantial disagreement between the resulting field boundaries, simply because the fields themselves may have shifted during the interval [@ElmesAccountingtrainingdata2020].
-These discrepancies could be further exacerbated by differences in the definition of what constitutes a field, which might vary on the ground depending on who is being asked, or who is doing the collecting.
-These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth".
-Nevertheless, a comparison against ground data would help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.
+One of these was the degree of correspondence between image- and ground-collected labels.
+However, such comparisons may reveal unresolvable discrepancies between the two perspectives.
+The highly dynamic nature of these agricultural systems means that relatively narrow differences between the dates of ground- and image-based label collection can lead to substantial disagreement, simply because the fields themselves may have shifted during the interval [@ElmesAccountingtrainingdata2020].
+These discrepancies can be exacerbated by the definition used to determine what constitutes a field, which might vary on the ground depending on who is being asked, or who is doing the collecting.
+These factors suggest that difference between ground- and image-collected labels would not necessarily indicate how far image labellers were from the "truth".
+Nevertheless, a comparison against ground data would help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation.
 
@@
 
-Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpact}) was influenced by the validation labels we used, which were generated using the consensus method.
-This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels.
-However, the visual assessment of their resulting probability maps confirm the differences in scores: consensus and most accurate individual labels produce nearly identical maps with relatively high certainty, while low quality labels led to a markedly less certain map (Figure S9).
-
-## Next steps
-
-The maps presented here represent a version 1 product that is freely available to use, along with its underlying code (see SI for details).
-These data were developed according to the recommended best practices for training and assessing error in machine learning models [@ElmesAccountingtrainingdata2020].
-In their current form, the maps may be useful for a variety of research applications.
+Another potential issue is the degree to which our assessment of label error on model performance (Figure \ref{fig:trainingimpact}) was influenced by the validation dataset we used, which was based on consensus labels.
+This could have confounded the analysis, particularly when comparing the consensus label-trained models with those trained with the most accurate individual labels.
+However, a visual assessment of the resulting probability maps confirms that models trained with the consensus and most accurate individual labels were more precise than the model trained with lower quality labels (Figure S12).
+
+## Broader applications
+This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using an approach that follows recommended best practices for training and assessing machine learning models [@ElmesAccountingtrainingdata2020].
+This approach can be readily updated to integrate improvements, such as newer machine learning models.
+Beyond providing valuable insights into field characteristics, field boundary maps can help improve remote estimation of crop areas and yield [e.g.
+@EstesProjectedclimateimpacts2013], and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and farm size [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
+Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of Africa.
+
+## Data availability and usage
+The maps presented here are a version 1 product that is freely available to use, along with its underlying code (see SI).
+In their current form, they may be useful for a variety of research applications.
 For example, analyzing the distributions of values in the probability maps may provide additional insight into the relative extents of active versus fallow croplands [@Tongforgottenlanduse2020].
@@
 
-To facilitate the next step, generating more accurate version 2 maps, several improvements will be made.
-The first is to replace Random Forests with a more advanced convolutional neural network (CNN), which can generate and learn from a large number of features representing a variety of spatial scales [@maDeepLearningRemote2019].
-Recent work suggests that a common architecture such as U-Net, when trained to distinguish field edges from interiors and combined with a post-hoc segmentation routine, is effective in delineating field boundaries [@waldnerDeepLearningEdge2020].
-Our system can readily incorporate such a model.
-The labelling platform already provides the methods needed to develop and assess the quality of labels that include field edges and interior classes, while active learning has proven to be effective for optimizing training datasets for deep learning models [@caoHyperspectralImageClassification2020; @liuActiveDeepLearning2017].
-Our current framework can be adjusted so that it starts by training a CNN from scratch with a large initial random sample, and then uses a transfer learning approach [@panSurveyTransferLearning2010] to update the model with the most informative samples from different AOIs or agroecozones.
-
-## Conclusion
-
-This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve.
-Maps that include information on field boundaries can help improve remote estimation of crop planted area and yield, and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and livelihoods.
-Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.
-
 # Acknowledgements
-
 The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PLACE.

### 2021-07-20 inst/manuscript/si.Rmd
Fully revised manuscript (except abstract), SI, and figures.

@@
 
-2. After determining the class, if
+2. After determining the class, if:
   - The target polygon is contained entirely within a single clear class, then simply digitize a point within the center of the target box, assign the appropriate class label, and complete the assignment;
@@
 The models for these two AOIs were thus trained with `r aoi10n` - `r aoi14n` labels.
-AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see SI).
-In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used (Figure S4A), followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
+AOI 15 was run for 4 iterations (`r aoi15n` samples), while AOI 3 underwent a second active learning cycle because the model produced during the first cycle was inaccurate (see next section).
+In this second run, `r aoi3init` initial training sites randomly selected within the AOI were used, followed by 2 subsequent active learning iterations, resulting in a training sample of `r aoi3n`.
 
@@
 In AOI 3, the initial active learning cycle resulted in low accuracy because the northern part of the AOI shows low contrast between fields and the surrounding vegetation in the dry season.
-Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A)
+Training the model with the initial 500 samples resulted in large commission errors in this part of the AOI, thus we ran a second active learning cycle that began with an inital random draw of 300 training sites confined to this AOI (blue points in Figure \ref{fig:trainval}A).
 ```{r trainval, echo = FALSE, out.width="70%", fig.cap="The distribution of A) initial randomly selected training sites, including 300 points selected to initialize a second run for retrain AOI 3, and the locations of training reference sites, and B) validation points and training sites selected during each active learning iteration.
@@
 
-The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who mapped more than 1,000 sites each (Figure \ref{fig:trainqual}A).
-As each training and reference tasks was mapped by 4 separate labellers, `r tsitesc$tot[1]` sets of vector labels were made.
-Each labeller digitized an average of `r tsitesc$mu[1]`.
-Labeller accuracy was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites, with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
-The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)` (Figure \ref{fig:trainqual}B).
+The distribution of training and validation sample collection effort was divided across `r nrow(label_summary$score_stats)` labellers, with a core group of `r label_summary$nlabels %>% filter(kml_type == "Training" & n > 1000) %>% nrow()` who completed more than 1,000 assignment each (Figure \ref{fig:assignmentcount}).
+As each training/validation task was undertaken by 4 separate labellers, `r tsitesc$tot[1]` sets of labels were made.
+Each labeller digitized an average of `r tsitesc$mu[1]` training/validation assignments.
+
+```{r assignmentcount, echo = FALSE, out.width="80%", fig.cap="The A) number of training/validation and accuracy assessment assignments completed by each labeller, and B) the distributions of quality scores at training reference sites for each labeller (means indicated by X in boxplots).
+Labellers' identities are anonymized.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_training_n.png')
+```
 
@@
 This was done because we discovered a small spatial offset in the original image composites, which we corrected by reprocessing the images.
-We replaced the image overlays in the instance for central-southeast (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
+We replaced the image overlays in the instances for this Cluster (Figure \ref{fig:trainval}A), and the labelling team mapped these sites a second time during the production run in late 2019.
 The reprocessed labels were used to initially train the model for AOIs 9, 12, and 15.
-
-```{r trainqual, echo = FALSE, out.width="50%", fig.cap="The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
-Labellers' identities are anonymized.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_training_qual_n.png')
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("label_risk")
-murisk <- label_risk$stats %>% group_by(usage) %>% 
-  summarize(Risk = round(mean(Risk), 3)) %>% 
-  tidyr::pivot_wider(names_from = usage, values_from = Risk)
-mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
-  filter(Risk == min(Risk))
-maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
-  filter(Risk == max(Risk))
-minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
-  filter(Risk == min(Risk))
-maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
-  filter(Risk == max(Risk))
-
-# label clusters
-cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
-  summarize(Risk = round(mean(Risk), 3))
-cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
-  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
-cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
-  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
-
-# training clusters
-cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
-  round(3)
-cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
-  round(3)
-```
-
-The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, as well as the distribution of risk values per AOI and the initial training clusters (Figure \ref{fig:labelriskhist}).
-The three initial clusters include the second mapping of Cluster 2.
-
-The average Bayesian risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
-Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels (Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
-Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
-Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster; the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.
-
-```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_label_risk_map.png')
-```
-
-```{r labelriskhist, echo = FALSE, out.width="40%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
-The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE}
-knitr::include_graphics('figures/si_label_risk_hists.png')
-```
 
@@
 
-The lower accuracy of actively versus randomly trained models in earlier iterations was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:randomvactive}).
-In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
-Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.
-
 ```{r randomvactive, echo = FALSE, out.width="90%", fig.cap="The percent difference in performance metrics per iteration for AOIs 1, 8, and 15 (grey lines with numbers indicating AOI; the black line indicates average difference across the three AOIs) when comparing models trained using active learning versus those trained using randomly selected sites.
@@
 
+The lower accuracy of actively versus randomly trained models in earlier iterations was caused by results at AOI 15, where active learning accuracy was `r abs(u(i1acc["15"]))` percent lower than random training after iteration 1 (see Figure \ref{fig:randomvactive}).
+In comparison, iteration 1 active learning accuracies were `r u(i1acc["1"])` and `r u(i1acc["8"])` percent higher than random training for AOIs 1 and 8, respectively.
+Accuracy under active learning for AOI 15 exceeded randomized training after 4 iterations.
+
 ### The impact of training data error
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+data("label_risk")
+murisk <- label_risk$stats %>% group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3)) %>% 
+  tidyr::pivot_wider(names_from = usage, values_from = Risk)
+mintrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
+  filter(Risk == min(Risk))
+maxtrainrisk <- label_risk$stats %>% filter(usage == "Train") %>% 
+  filter(Risk == max(Risk))
+minvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
+  filter(Risk == min(Risk))
+maxvalrisk <- label_risk$stats %>% filter(usage == "Validate") %>% 
+  filter(Risk == max(Risk))
+
+# label clusters
+cl1 <- label_risk$stats %>% filter(aoi %in% 1:6) %>% group_by(usage) %>% 
+  summarize(Risk = round(mean(Risk), 3))
+cl2 <- label_risk$stats %>% filter(aoi %in% c(7:9, 12, 15)) %>% 
+  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
+cl3 <- label_risk$stats %>% filter(aoi %in% c(10:11, 13:14, 16)) %>% 
+  group_by(usage) %>% summarize(Risk = round(mean(Risk), 3))
+
+# training clusters
+cl2arisk <- label_risk$stats %>% filter(aoi %in% c("Cl2a")) %>% pull(Risk) %>% 
+  round(3)
+cl2brisk <- label_risk$stats %>% filter(aoi %in% c("Cl2b")) %>% pull(Risk) %>% 
+  round(3)
+```
+
+Two measures of label quality, average quality score per labeller and the Bayesian Risk of labels, were calculated to provide proxy measures of label error.
+Labeller quality was scored `r tsitesc$tot[2]` times against `r length(unique(label_summary$scores$name))` unique training reference sites, with each labeller assessed an average of `r tsites$mu[2]` times at a rate of 1 training reference site for every `r round(1 / (tsites$mu[2] / tsites$mu[1]), 2)` training site mapped.
+The mean of each labeller's average accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; Figure \ref{fig:labelqual}).
+
+```{r labelqual, echo = FALSE, out.width="70%", fig.cap="Boxplots showing the distributions of quality scores from the accuracy assessment assignments undertaken by each labeller (means indicated by X in boxplots).
+Labellers' identities are anonymized.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_quality.png')
+```
+
+The average Bayesian Risk of each training and validation site is shown in Figure \ref{fig:labelrisk}, and the distribution of risk values per AOI and the initial training clusters in Figure \ref{fig:labelriskhist}.
+The three initial clusters include the second mapping of Cluster 2 (Figure \ref{fig:trainval}A).
+
+```{r labelrisk, echo = FALSE, out.width="80%", fig.cap="The average Bayes Risk of each training and validation site in Ghana.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_risk_map.png')
+```
+
+```{r labelriskhist, echo = FALSE, out.width="40%", fig.cap="The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
+The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.", fig.align='center', message=FALSE}
+knitr::include_graphics('figures/si_label_risk_hists.png')
+```
+
+The average Bayesian risk was `r murisk$Train` for training labels and `r murisk$Validate` for validation labels.
+Risk was highest in the northern AOIs (AOIs 1-6; Figures \ref{fig:labelrisk}-\ref{fig:labelriskhist}), falling between `r cl1$Risk[1]` for training and `r cl1$Risk[2]` for validation labels, and lowest in the southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = `r cl3$Risk[1]`; validation risk = `r cl3$Risk[2]`).
+Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower (training = `r cl2$Risk[1]`; validation = `r cl2$Risk[2]`) than in the north.
+Labeller experience also appeared to reduce risk, which we observed during a relabelling of the 500 initial random site in this cluster; the mean risk of the updated labels was `r cl2brisk`, compared to `r cl2arisk` for original labels.
+
 Probability images resulting from Random Forests models trained with labels generated under three different labeling strategies are illustrated in Figure \ref{fig:labelstrategy}.
 These included consensus labels, and those individual labels that were likely to be the most and least accurate for each training site.
-Label accuracy was based on the mean score of each labeller against the training reference sites (Figure \ref{fig:trainqual}), as assessed when labelling a given AOI.
+Label accuracy was based on the mean score of each labeller against the training reference sites (Figure \ref{fig:labelqual}), as assessed when labelling a given AOI.
 These images were created for a single tile in AOI 1.
@@
 - [`learner`](https:github.com/agroimpacts/learner): Source code for the active learning component. The repository\footnote{github.com/agroimpacts/learner} is open.
-- [`segmenter`](https:github.com/agroimpacts/segmenter): Source code for the segmentation algorithm.
-The repository \footnote{github.com/agroimpacts/segmenter} is currently private, but will be opened after completing a security audit.
+- [`segmenter`](https:github.com/agroimpacts/segmenter): Source code for the segmentation algorithm. The repository \footnote{github.com/agroimpacts/segmenter} is open.
 

### 2021-07-20 inst/manuscript/manuscript.Rmd
Abstract redrafted

@@
 This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
-abstract: Understanding agricultural change requires reliable, frequently updated maps that describe the characteristics of croplands.
-Such data are often unavailable for regions dominated by smallholder agricultural systems, which are particularly challenging for remote sensing.
-To overcome these challenges, we designed a system to minimize several sources of error that arise when mapping smallholder croplands.
-To overcome errors caused by mismatches between image resolution and cropland scales, as well as persistent cloud cover, the system converts daily, 3.7 m PlanetScope imagery into two seasonal composites within a single agricultural year.
-To reduce errors that occur when training classifiers, we built a labelling platform that rigorously assesses label accuracy, and creates more accurate consensus labels that train a Random Forests model.
-The labelling platform and model interact within an active learning process that boosts the accuracy of the resulting cropland probability map, which is used in a segmentation process to delineate individual field boundaries.
-We applied this system to map Ghana’s croplands for the year 2018.
-We divided Ghana into 16 mapping regions (12,160-23,535 km$^2$), training separate models for each using a total of 6,299 labels, plus 1,600 for validation.
-Using an independent map reference sample (n=1,207), we found that overall accuracies of the resulting cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%.
-Croplands covered 16.1-23.2% of the mapped area, comprising 1,131,146 total fields with an average size of 3.92 ha.
-Estimates based on the map reference sample indicate the cropland percentage is 17.1% (15.4-18.9%) or 17.6% (15.6-19.6%), depending on the map used to estimate the standard error.
-Using the labellers’ digitized field boundaries to estimate biases in field boundary statistics, we calculated an adjusted mean field size of 1.73 ha and total field count of 1,662,281.
-Although the cropland class contained substantial errors, the system was effective in mitigating error and quantifying resulting performance gains.
-By minimizing training errors, consensus labelling improved the model’s F1 scores by up to 25%, while 3 iterations of active learning increased the F1 score by 9.1%, on average, which was 2.3% higher than training models with randomly selected labels.
-Map accuracy can be improved by replacing Random Forests with a convolutional neural network.
-These results demonstrate a readily adapted, transferrable framework for developing high resolution, annual, nation-scale maps that provide important details about smallholder-dominated croplands.
+abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is key to understanding critical questions related to food security and other pressing socioeconomic and environmental concerns.
+However, accurately mapping these systems is difficult because of two key problems: 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers.
+We developed a cropland mapping approach designed to address these two problems, which we used to map Ghana’s croplands (excluding those under tree crops) for the year 2018.
+To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering the 2018 agricultural calendar.
+To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a machine learning classifier (Random Forests) using an active learning approach, in which we used the uncertainty in model predictions to select more informative training samples.
+Minimizing label errors improved model F1 scores by up to 25%, while active learning increased the F1 score by an average of 9.1% between first and last training iterations, which was 2.3% higher than models trained with randomly selected labels.
+We used the resulting <4 m map of cropland probabilities within a final segmentation step to delineate crop field boundaries.
+Using an independent map reference sample (n=1,207), we found that overall accuracies of the cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%.
+Relying on the map reference sample to generate an unbiased area estimate from the field boundary map, we calculate that cropland covers 17.1 (15.4-18.9%).
+Using the labellers’ digitized field boundaries to estimate biases in the size and number of segments in the field boundary map, we estimate that there are 1,662,281 crop fields in Ghana with an average size of 1.73 ha.
+Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively minimize the errors inherent in remote sensing of smallholder-dominated agriculture.
 
@@
 @EstesProjectedclimateimpacts2013], and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and farm size [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
-Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of Africa.
+Such maps will be important for understanding the rapid agricultural change that is currently occurring in Africa.
 

### 2021-07-21 inst/manuscript/manuscript.Rmd
Tweaked abstract, incremented package

@@
 Using the labellers’ digitized field boundaries to estimate biases in the size and number of segments in the field boundary map, we estimate that there are 1,662,281 crop fields in Ghana with an average size of 1.73 ha.
-Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively minimize the errors inherent in remote sensing of smallholder-dominated agriculture.
+Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively mitigate the errors inherent in remote sensing of smallholder-dominated agriculture.
 

### 2021-07-21 inst/manuscript/manuscript.Rmd
A few more edits.

@@
 note: This pre-print not yet undergone peer review.
-It has been submitted to the \emph{Frontiers in Artificial Intelligence}.
+It has been submitted to \emph{Frontiers in Artificial Intelligence}.
 This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
 abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is key to understanding critical questions related to food security and other pressing socioeconomic and environmental concerns.
-However, accurately mapping these systems is difficult because of two key problems: 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers.
+However, accurately mapping these systems is difficult because of two key problems$:$ 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers.
 We developed a cropland mapping approach designed to address these two problems, which we used to map Ghana’s croplands (excluding those under tree crops) for the year 2018.
@@
 @XiongNominal30mCropland2017; @ChenGloballandcover2015; @StehmanKeyissuesrigorous2019], as it is impractical and expensive to collect these data in the field over large areas, particularly on an ongoing basis.
-As such, a number of web-based platforms have been developed to collect such labels [e.g.
+Consequently, a number of web-based platforms have been developed to collect such labels [e.g.
 @BeyCollectEarthLand2016; @Estesplatformcrowdsourcingcreation2016; @FritzGeoWikionlineplatform2012].
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("image_quality")
-data("tile_key")
+data(image_quality)
+data(tile_key)
 tile_quality <- image_quality$img %>% filter(kml_type == "F") %>% 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("image_quality")
+data(image_quality)
 tqual_mu <- image_quality$tile %>%
@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
-data("iteration_metrics")
+data(iteration_metrics)
 delta_func <- function(x, rnd = 4) round((x - lag(x)) / lag(x), rnd)
@@
 # random versus active data
-data("iteration_metrics")
+data(iteration_metrics)
 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("label_summary")  
-# toofew <- label_summary$nlabels %>% filter(n < 10) %>% pull(ID)
-# tsites <- label_summary$nlabels %>% filter(!ID %in% toofew) %>% 
-#   group_by(kml_type) %>% 
-#   summarize(tot = sum(n), mu = round(mean(n)), med = round(median(n)), 
-#             stdev = round(sd(n)), lower = min(n), upper = max(n))
-# 
-# # training rate
-# train_rate <- label_summary$nlabels %>% 
-#   tidyr::pivot_wider(names_from = kml_type, values_from = n) %>% 
-#   rename_all(funs(c("ID", "tr", "tref"))) %>% 
-#   mutate(trate = tref / tr) %>% filter(tr > 10)
-# 
-# tsitesc <- tsites %>% mutate(across(where(is.numeric), fmt))
-# # nvecs <- tsites$tot[1])
-# # muvecs <- fmt(tsites$mu[1])
-
-data("label_risk")
+data(label_risk)
 murisk <- label_risk$stats %>% #group_by(usage) %>% 
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("consensus_high_low")
+data(consensus_high_low)
 V1 <- c("Accuracy", "AUC", "F1")
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE}
-data("label_risk_metrics")
+data(label_risk_metrics)
 
@@
 # validation statistics
-data("segment_quality_stats")
+data(segment_quality_stats)
 
@@
 ```{r, echo=FALSE}
-data("accuracy_reports")
+data(accuracy_reports)
 
@@
 Using the map reference sample and each map, we made two separate estimates of the total cropland area in Ghana in 2018.
-The cropland extent estimated from the field boundary map was `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]` (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area.
-The estimate from the per pixel map was `r fmt(rfas[2])` km$^2$% (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]` (`r paste(rfcipct[2:3], collapse = "-")`%) of area.
+The cropland extent estimated from the field boundary map was `r fmt(segas[2])` km$^2$ (with a margin of error of `r fmt(segma[1])` km$^2$), or `r segcipct[1]`% (`r paste(segcipct[2:3], collapse = "-")`%) of the mapped area.
+The estimate from the per pixel map was `r fmt(rfas[2])` km$^2$ (margin of error = `r fmt(rfma[1])` km$^2$), or `r rfcipct[1]`% (`r paste(rfcipct[2:3], collapse = "-")`%) of area.
 
@@
 # calculate bias factors for adjustment
-data("segment_quality_stats")
+data(segment_quality_stats)
 # segment_quality_stats %>% filter(measure == "n")

### 2021-07-21 inst/manuscript/manuscript.Rmd
Frontiers version

@@
 footerdate: no
-note: This pre-print not yet undergone peer review.
-It has been submitted to \emph{Frontiers in Artificial Intelligence}.
-This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
-abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is key to understanding critical questions related to food security and other pressing socioeconomic and environmental concerns.
-However, accurately mapping these systems is difficult because of two key problems$:$ 1) the spatial and temporal mismatch between most publicly satellite sensors and smallholder fields; 2) the difficulty of collecting high quality labels for training and assessing machine learning classifiers.
-We developed a cropland mapping approach designed to address these two problems, which we used to map Ghana’s croplands (excluding those under tree crops) for the year 2018.
-To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering the 2018 agricultural calendar.
-To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a machine learning classifier (Random Forests) using an active learning approach, in which we used the uncertainty in model predictions to select more informative training samples.
-Minimizing label errors improved model F1 scores by up to 25%, while active learning increased the F1 score by an average of 9.1% between first and last training iterations, which was 2.3% higher than models trained with randomly selected labels.
-We used the resulting <4 m map of cropland probabilities within a final segmentation step to delineate crop field boundaries.
-Using an independent map reference sample (n=1,207), we found that overall accuracies of the cropland probability and field boundary maps were 88% and 86.7%, respectively, with User’s accuracies for the cropland class of 61.2% and 78.9%, and Producer’s accuracies of 67.3% and 58.2%.
-Relying on the map reference sample to generate an unbiased area estimate from the field boundary map, we calculate that cropland covers 17.1 (15.4-18.9%).
-Using the labellers’ digitized field boundaries to estimate biases in the size and number of segments in the field boundary map, we estimate that there are 1,662,281 crop fields in Ghana with an average size of 1.73 ha.
+abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is crucial for understanding food security and a range of other socioeconomic and environmental concerns.
+However, accurately mapping these systems is difficult because of 1) the spatial and temporal mismatch between satellite sensors and smallholder fields, and 2) the lack of high-quality labels needed to train and assess machine learning classifiers.
+We developed an approach designed to address these two problems, which we used to map Ghana’s annual croplands for the year 2018.
+To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering a single agricultural year.
+To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a Random Forests classifier with active learning, which identifies the most informative training sample based on prediction uncertainty.
+Minimizing label errors improved model F1 scores by up to 25%.
+Active learning increased F1 scores by an average of 9.1% between first and last training iterations, and 2.3% more than models trained with randomly selected labels.
+We used the resulting 3.7 m map of cropland probabilities within a segmentation algorithm to delineate crop field boundaries.
+Based on an independent map reference sample (n=1,207), the cropland probability and field boundary maps have respective overall accuracies of 88% and 86.7%, user’s accuracies for the cropland class of 61.2% and 78.9%, and producer’s accuracies of 67.3% and 58.2%.
+Using the map reference sample to calculate an unbiased area estimate from the field boundary map, we found that cropland covers 17.1% (15.4-18.9%) of Ghana.
+Using the most accurately digitized labels to calculate and correct for biases in the segmented field boundaries map, we further estimated the average size (1.73 ha) and total number (1,662,281) of crop fields in Ghana.
 Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively mitigate the errors inherent in remote sensing of smallholder-dominated agriculture.

### 2021-07-23 inst/manuscript/manuscript.Rmd
Added dataset documentation, manuscript statement (EarthArxiv), several datasets

@@
 footerdate: no
+note: This pre-print not yet undergone peer review.
+It has been submitted to \emph{Frontiers in Artificial Intelligence}.
+This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
 abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is crucial for understanding food security and a range of other socioeconomic and environmental concerns.
@@
 
-keywords: active learning, machine learning, CubeSat, PlanetScope, smallholder, cropland, segmentation, field size, consensus labels, Ghana, Africa
+keywords: active learning, machine learning, CubeSat, PlanetScope, smallholder, cropland, segmentation, field size, label error, Ghana, Africa
 

### 2021-07-23 inst/manuscript/manuscript.tex
Added dataset documentation, manuscript statement (EarthArxiv), several datasets

@@
 This pre-print not yet undergone peer review. It has been submitted to
-the \emph{Frontiers in Artificial Intelligence}. This version will be
+\emph{Frontiers in Artificial Intelligence}. This version will be
 updated as it is revised, and the final published version will be
@@
 \begin{abstract}
-Understanding agricultural change requires reliable, frequently updated
-maps that describe the characteristics of croplands. Such data are often
-unavailable for regions dominated by smallholder agricultural systems,
-which are particularly challenging for remote sensing. To overcome these
-challenges, we designed a system to minimize several sources of error
-that arise when mapping smallholder croplands. To overcome errors caused
-by mismatches between image resolution and cropland scales, as well as
-persistent cloud cover, the system converts daily, 3.7 m PlanetScope
-imagery into two seasonal composites within a single agricultural year.
-To reduce errors that occur when training classifiers, we built a
-labelling platform that rigorously assesses label accuracy, and creates
-more accurate consensus labels that train a Random Forests model. The
-labelling platform and model interact within an active learning process
-that boosts the accuracy of the resulting cropland probability map,
-which is used in a segmentation process to delineate individual field
-boundaries. We applied this system to map Ghana's croplands for the year
-2018. We divided Ghana into 16 mapping regions (12,160-23,535 km\(^2\)),
-training separate models for each using a total of 6,299 labels, plus
-1,600 for validation. Using an independent map reference sample
-(n=1,207), we found that overall accuracies of the resulting cropland
-probability and field boundary maps were 88\% and 86.7\%, respectively,
-with User's accuracies for the cropland class of 61.2\% and 78.9\%, and
-Producer's accuracies of 67.3\% and 58.2\%. Croplands covered
-16.1-23.2\% of the mapped area, comprising 1,131,146 total fields with
-an average size of 3.92 ha. Estimates based on the map reference sample
-indicate the cropland percentage is 17.1\% (15.4-18.9\%) or 17.6\%
-(15.6-19.6\%), depending on the map used to estimate the standard error.
-Using the labellers' digitized field boundaries to estimate biases in
-field boundary statistics, we calculated an adjusted mean field size of
-1.73 ha and total field count of 1,662,281. Although the cropland class
-contained substantial errors, the system was effective in mitigating
-error and quantifying resulting performance gains. By minimizing
-training errors, consensus labelling improved the model's F1 scores by
-up to 25\%, while 3 iterations of active learning increased the F1 score
-by 9.1\%, on average, which was 2.3\% higher than training models with
-randomly selected labels. Map accuracy can be improved by replacing
-Random Forests with a convolutional neural network. These results
-demonstrate a readily adapted, transferrable framework for developing
-high resolution, annual, nation-scale maps that provide important
-details about smallholder-dominated croplands.
+Mapping the changing characteristics of Africa's smallholder-dominated
+agricultural systems, including the sizes and numbers of fields, is
+crucial for understanding food security and a range of other
+socioeconomic and environmental concerns. However, accurately mapping
+these systems is difficult because of 1) the spatial and temporal
+mismatch between satellite sensors and smallholder fields, and 2) the
+lack of high-quality labels needed to train and assess machine learning
+classifiers. We developed an approach designed to address these two
+problems, which we used to map Ghana's annual croplands for the year
+2018. To overcome the first problem, we converted daily, high resolution
+CubeSat (PlanetScope) imagery into two cloud-free seasonal composites
+covering a single agricultural year. To address the second problem, we
+created a labelling platform that rigorously assesses and minimizes
+label error, and used it to iteratively train a Random Forests
+classifier with active learning, which identifies the most informative
+training sample based on prediction uncertainty. Minimizing label errors
+improved model F1 scores by up to 25\%. Active learning increased F1
+scores by an average of 9.1\% between first and last training
+iterations, and 2.3\% more than models trained with randomly selected
+labels. We used the resulting 3.7 m map of cropland probabilities within
+a segmentation algorithm to delineate crop field boundaries. Based on an
+independent map reference sample (n=1,207), the cropland probability and
+field boundary maps have respective overall accuracies of 88\% and
+86.7\%, user's accuracies for the cropland class of 61.2\% and 78.9\%,
+and producer's accuracies of 67.3\% and 58.2\%. Using the map reference
+sample to calculate an unbiased area estimate from the field boundary
+map, we found that cropland covers 17.1\% (15.4-18.9\%) of Ghana. Using
+the most accurately digitized labels to calculate and correct for biases
+in the segmented field boundaries map, we further estimated the average
+size (1.73 ha) and total number (1,662,281) of crop fields in Ghana. Our
+results demonstrate an adaptable and transferrable approach for mapping
+the characteristics of croplands on an annual basis and over national
+extents, with several features that effectively mitigate the errors
+inherent in remote sensing of smallholder-dominated agriculture.
 \end{abstract}
@@
 will have on societies, economies, and the environment (Searchinger et
-al. 2019). Significant efforts are being made to address the various
-aspects of this challenge, including work on diagnosing and closing
-yield gaps (Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al.
-2012), expanding and commercializing production (Morris and Byerlee
-2009), and to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017,
-Davis et al. 2020) and mitigate (Estes et al. 2016b) agriculture's
-ecological impacts. Answering many of the questions these efforts seek
-to address depends on reliable data that describes the location and
-characteristics of cropland (Fritz et al. 2015), and how these are
-changing over time, ideally on an annual basis, given the rapid pace of
-change (Gibbs et al. 2010, Bullock et al. 2021). Unfortunately, the data
-that do exist are in many places inaccurate and infrequently updated.
-Existing estimates of how much global cropland there is tend to vary
-widely, and they often disagree about where cropland is located (e.g.
-Fritz et al. 2011, 2013). Such errors can propagate in subsequent
-analyses that rely on cropland data as inputs, resulting in potentially
-misleading answers (Estes et al. 2018). Beyond cropland distributions,
-few data are available on key cropland characteristics such as field
-size, an important variable needed to estimate yield and other key food
-security variables (Carletto et al. 2015), and as an indicator of farm
-size (Levin 2006, Samberg et al. 2016), a critical component of rural
-livelihoods given increasing population densities and longstanding
-debates about the relationship between farm size and productivity (Feder
-1985, Carletto et al. 2013, Desiere and Jolliffe 2018).
-
-These informational inadequacies are due to the fact that cropland data
-in much of the world are derived from remotely sensed landcover maps,
-which can be notoriously high in error, particularly over regions such
-as Africa (Fritz et al. 2010, Estes et al. 2018), where agricultural
-changes will be largest and the need for accurate baseline data is thus
-greatest (Searchinger et al. 2015, Estes et al. 2016b, Bullock et al.
-2021). Cropland mapping over Africa is difficult for several reasons.
-The primary reason relates to the characteristics of the continent's
-smallholder-dominated croplands, where half of all fields are smaller
-than 1 ha (Lesiv et al. 2019). This size is small relative to the 30-250
-m resolution of the sensors typically used in many landcover mapping
-efforts (e.g. Chen et al. 2015, Sulla-Menashe et al. 2019), which
-results in errors due to mixed pixels and aspects of the modifiable area
-unit problem (Openshaw and Taylor 1979, Boschetti et al. 2004){]}. In
-the latter case, the pixel's shape may be poorly matched to that of
-cropland, and is too coarse to aggregate to approximate that shape at
-the characteristic scales of crop fields (Dark and Bram 2007, Estes et
-al. 2018). On top of the matter of scale is 1) high intra-class
-variability of the cropland class, compounded by the fact that these
-particular croplands can be heavily intergraded with surrounding
-vegetation (Debats et al. 2016, Estes et al. 2016a), and 2) the
-substantial temporal variability within croplands, both within and
-between seasons. These latter two aspects pose challenges for the
+al. 2019). A number of efforts are underway to address various aspects
+of this challenge, including work on diagnosing and closing yield gaps
+(Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al. 2012),
+expanding and commercializing production (Morris and Byerlee 2009), and
+to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017, Davis et al.
+2020) and mitigate (Estes et al. 2016b) agriculture's ecological
+impacts. The success of these efforts depends heavily on data that
+accurately describes the location and characteristics of croplands
+(Fritz et al. 2015), and, given the rapid pace of agricultural change
+(Gibbs et al. 2010, Zeng et al. 2018, Bullock et al. 2021), how these
+are changing from one year to the next. Unfortunately, for many regions,
+existing cropland datasets are inaccurate, and are usually created as
+once-off or infrequently update products. As such, estimates of global
+cropland area tend to vary widely, often disagree about where croplands
+are located (e.g. Fritz et al. 2011, 2013), and become rapidly outdated.
+Errors in these maps can propagate in subsequent analyses that use
+cropland data as inputs, resulting in potentially misleading answers
+(Estes et al. 2018). Beyond distributions, few data are available on key
+cropland characteristics such as field size, an important variable
+needed to estimate yield and other key food security variables (Carletto
+et al. 2015), and as an indicator of farm size (Levin 2006, Samberg et
+al. 2016), a critical component of rural livelihoods given increasing
+population densities and longstanding debates about the relationship
+between farm size and productivity (Feder 1985, Carletto et al. 2013,
+Desiere and Jolliffe 2018).
+
+The deficit of information is due to the fact that in many regions the
+only source of cropland data are remotely sensed land cover maps, which
+are prone to error. This is particularly true in Africa (Fritz et al.
+2010, Estes et al. 2018), where agricultural changes will be largest and
+the need for accurate baseline data is thus greatest (Searchinger et al.
+2015, Estes et al. 2016b, Bullock et al. 2021), and where the
+characteristics of croplands exacerbate the error inherent in remote
+sensing analyses. Half of all fields in Africa's smallholder-dominated
+agricultural systems are smaller than 1 ha (Lesiv et al. 2019). This
+size is small relative to the 30-250 m resolution of the sensors
+typically used in many landcover mapping efforts (e.g. Chen et al. 2015,
+Sulla-Menashe et al. 2019), which results in errors due to mixed pixels
+and aspects of the modifiable area unit problem (Openshaw and Taylor
+1979, Boschetti et al. 2004), wherein the pixel's shape does not match
+that of crop fields, and is too coarse to aggregate into an
+approximation of that shape (Dark and Bram 2007, Estes et al. 2018). On
+top of the matter of scale is the high variability within and between
+fields, their tendency to intergrade with surrounding vegetation (Debats
+et al. 2016, Estes et al. 2016a), and the high temporal variability
+within croplands. These last three aspects pose challenges for the
 classification algorithms that are applied to the imagery.
@@
 high (\textless5 m) to near-high (10 m) resolution imagery at sub-weekly
-intervals (Drusch et al. 2012, McCabe et al. 2017). This high spatial
-\emph{and} temporal resolution imagery addresses the sensor-field scale
-mismatch, and more effectively captures the intra-seasonal dynamics of
+intervals (Drusch et al. 2012, McCabe et al. 2017). The spatial
+resolution of these imagery addresses the scale mismatch between sensor
+and field, and their high frequency captures the seasonal dynamics of
 cropland, which helps classifiers distinguish cropland from surrounding
 cover types (Debats et al. 2016, Defourny et al. 2019). On top of this,
-advances in cloud computing and the opening of satellite image archives
-(Wulder et al. 2016), and next generation machine learning approaches
-are placing large volumes of these moderate to near-high resolution
-imagery together with the computational and algorithmic resources
-necessary to classify them at scale (Gorelick et al. 2017). These
-capabilities are already being used to create a new generation of higher
-resolution (10-30 m) cropland and landcover maps for Africa and other
-regions {[}ESA (n.d.); Lesiv et al. (2017); Xiong et al. (2017); (Zhang
-et al. 2021){]}. However, the potential of the highest resolution
-(\textless5 m) imagery to map cropland over very large extents
-(e.g.~country scales) has yet to be realized, presumably because these
-data are commercial and relatively expensive, and require significant
-computational resource to process.
+the opening of satellite image archives (Wulder et al. 2016) and
+advances in cloud computing are placing large volumes of moderate to
+near-high resolution imagery together with the computational and
+algorithmic resources necessary to classify them at scale (Gorelick et
+al. 2017). These capabilities have already been used to create a new
+generation of higher resolution (10-30 m) cropland and landcover maps
+for Africa and other regions {[}ESA (n.d.); Lesiv et al. (2017); Xiong
+et al. (2017); (Zhang et al. 2021){]}. However, the potential of the
+highest resolution (\textless5 m) imagery to map cropland over very
+large extents (e.g.~country scales) has yet to be realized, presumably
+because these data are commercial and relatively expensive, and require
+significant computational resource to process.
 
@@
 for large training datasets, which are typically unavailable, hard to
-collect, or contain numerous errors in developing regions (Ma et al.
-2019, Elmes et al. 2020, Burke et al. 2021). To build sufficient samples
-for training cropland mapping models, as well as the reference data
-needed to objectively assess their performance (we refer collectively to
-both sample types as ``labels,'' distinguishing between each type where
+collect, or contain numerous errors (Ma et al. 2019, Elmes et al. 2020,
+Burke et al. 2021). To build sufficient training samples, as well as the
+reference data needed to objectively assess their performance (we refer
+collectively to both types as ``labels,'' distinguishing between each as
 needed), map-makers rely heavily on visual interpretation of high
@@
 collect these data in the field over large areas, particularly on an
-ongoing basis. Several web-based platforms have been developed to
-facilitate such efforts, which provide convenient and highly scalable
-tools for training data collection (Fritz et al. 2012, Estes et al.
-2016a, e.g. Bey et al. 2016). Visually interpreted labels present two
-particular problems. The first is that they inevitably contain errors of
-interpretation that can vary substantially according to the skill of the
-labeller, particularly over complex croplands with small field sizes
+ongoing basis. Consequently, a number of web-based platforms have been
+developed to collect such labels (Fritz et al. 2012, Estes et al. 2016a,
+e.g. Bey et al. 2016). Image-drawn labels present two particular
+problems. The first is that they inevitably contain errors of
+interpretation, which can vary substantially according to the skill of
+the labeller, particularly over complex croplands with small field sizes
 (Estes et al. 2016a, Waldner et al. 2019). The second problem is that
 visual interpretation depends on high resolution imagery (\textless5 m),
-as field boundaries become increasingly difficult to discern as image
-resolution decreases. Typically the only practical source of high
-resolution imagery is ``virtual globe'' basemaps (e.g.~Bing or Google
-Maps), which are composed of mosaics of various high resolution
-satellite and aerial images that typically span 3-5 years of time within
-a single country (Lesiv et al. 2018). This within-mosaic temporal
-variation can set up a temporal mismatch between the imagery being
-interpreted and the imagery being classified, which is usually from a
-different source (e.g.~Landsat, Sentinel; Xiong et al. (2017)). If a
-land change occurs in the interval between the two image sets (e.g.~a
-new field was created), the label, even if accurately drawn, introduces
-error into the classifier. This source of error may be elevated in
-tropical smallholder systems, where swidden agricultural is common (Van
-Vliet et al. 2013), or in rapidly developing agricultural frontiers
-(Zeng et al. 2018). Despite the high potential for label error, they are
-typically not accounted for during model training and map accuracy
-assessment, resulting not only in the potential for maps to be misused
-or misinterpreted in subsequent analyses, but in missed opportunities to
-improve model performance (Estes et al. 2018, Stehman and Foody 2019,
-Elmes et al. 2020).
+as fields are increasingly difficult to discern as image resolution
+decreases. Typically the only available source of high resolution
+imagery is ``virtual globe'' basemaps (e.g.~Bing or Google Maps), which
+present mosaics of high resolution satellite and aerial images collected
+over a span of several years (Lesiv et al. 2018). This within-mosaic
+temporal variation can create a temporal mismatch between the labels and
+the imagery being classified, which is usually from a different source
+(e.g.~Landsat, Sentinel; Xiong et al. (2017)). If a land change occurs
+in the interval between the two image sets (e.g.~a new field was
+created), the label, even if accurately drawn, introduces error into the
+classifier. This source of error may be elevated in croplands where
+swidden agriculture is practiced (Van Vliet et al. 2013), or in rapidly
+developing agricultural frontiers (Zeng et al. 2018). Despite the high
+potential for it, label error is often not considered during model
+training and map accuracy assessment, resulting not only in the
+potential for maps to be misused or misinterpreted, but in missed
+opportunities to improve model performance (Estes et al. 2018, Stehman
+and Foody 2019, Elmes et al. 2020).
 
@@
 al. 2016). Beyond the frequency, PlanetScope's 3.7 m
-resolution--although substantially coarser the 0.5-1 m available in most
-areas covered by virtual globes--is sufficiently resolved for humans to
-discern small fields under many conditions (Fourie 2009, e.g.~see Estes
-et al. 2018). This allows labels to be made using the same imagery that
-is classified, which helps to minimize label error. To further reduce
-label noise, we developed a platform that includes rigorous label
-accuracy assessment protocols and a novel approach for creating
-consensus labels, which helps reduce mistakes made by individual
-labellers (Estes et al. 2016a, Elmes et al. 2020). We couple the
-labelling platform with a machine learning model inside an active
-learning (Cohn et al. 1994, Tuia et al. 2011) framework, in which model
-training is done iteratively and interactively, wherein the model's
-prediction uncertainty over unlabelled areas is assessed after each
-training step, and a sample of the most uncertain sites is selected for
-additional labelling (Cohn et al. 1994, Tuia et al. 2011). This approach
-helps boost the performance of the classifier while reducing the overall
+resolution--although substantially coarser than the 0.5-1 m imagery
+available in most areas covered by virtual globes--is sufficiently
+resolved for humans to discern small fields under many conditions
+(Fourie 2009, e.g.~see Estes et al. 2018). This allows labels to be made
+using the same imagery that is classified, which helps to minimize label
+error. To further reduce label noise, we developed a platform that
+includes rigorous label accuracy assessment protocols and a novel
+approach for creating consensus labels, which helps reduce mistakes made
+by individual labellers (Estes et al. 2016a, Elmes et al. 2020). We
+couple the labelling platform with a machine learning model inside an
+active learning (Cohn et al. 1994, Tuia et al. 2011) framework, in which
+the model is trained interactively, using the model's prediction
+uncertainty over unlabelled areas to select new sites for additional
+labelling (Cohn et al. 1994, Tuia et al. 2011). This approach helps
+boost the performance of the classifier while reducing the overall
 number of labels required to achieve a given level of performance
-(Debats et al. 2017, e.g. Hamrouni et al. 2021). Finally, an
-unsupervised segmentation step is applied to convert pixel-wise cropland
+(Debats et al. 2017, e.g. Hamrouni et al. 2021). An unsupervised
+segmentation step is then applied to convert pixel-wise cropland
 predictions into vectorized maps of individual field boundaries.
@@
 of crop field boundaries in Ghana, a country where smallholder farming
-predominates, and which has a broad mix of climates and agricultural
-systems, including large areas where shifting agriculture is practiced
-(Samberg et al. 2016, Kansanga et al. 2019). The map we develop
-represents a single agricultural year (2018-2019), as opposed to a
-multi-year composite, thereby demonstrating a capacity for annual, high
-resolution maps that can be used to monitor rapidly evolving small-scale
-agricultural systems, including key characteristics such as field size.
-In addition to presenting one of the most spatially extensive
-agricultural applications of CubeSats to date, we provide a new
-technique for converting daily imagery into seasonal composites, and
-show how best practices for model training and label collection can be
-applied to improve map accuracy (Elmes et al. 2020).
+predominates across a broad mix of climate and agricultural systems,
+ranging from primarily grain and vegetable crop production in the
+northern savannas to tree crop-dominated systems in the forested
+southwest, including large areas where shifting agriculture is practiced
+(Samberg et al. 2016, Kansanga et al. 2019). The map represents a single
+agricultural year (2018-2019), as opposed to a multi-year epoch, thereby
+demonstrating a capacity for annual, high resolution maps that can be
+used to monitor rapidly evolving small-scale agricultural systems,
+including key characteristics such as field size. In addition to
+providing valuable new data and insight into Ghana's agriculture, our
+study demonstrates one of the most spatially extensive agricultural
+applications of CubeSats to date, provides a new technique for
+converting daily imagery into seasonal composites, and shows how best
+practices for model training and label collection can be applied to
+improve map accuracy (Elmes et al. 2020).
 
@@
 
-The cropland mapping approach we developed is comprised of four primary
-components (Figure \ref{fig:systemoverview}) that were developed using
-open source software, and is designed to run in a cloud computing
-environment. The first component collects daily PlanetScope imagery for
-a period of several months over a particular mapping geography, and
-converts these into cloud-free seasonal composites. The second component
-is a custom-built platform that a provides a set tools for labelling the
-PlanetScope composites, and built-in procedures for assessing and
-minimizing label error. This platform interacts with the third
-component, a machine learning process running in Apache Spark, which
-includes a routine for 1) selecting new training sites based on their
-prediction uncertainty, and 2) returning these to the labelling platform
-for labeling. This interaction, which represents active learning (Cohn
-et al. 1994, Tuia et al. 2011), repeats until a model performance
-threshold is reached. The fourth and final component is a segmentation
-process that is applied to the seasonal image composites to segment the
-entire image, and then merged with the pixel-wise probability maps
-generated by active learning to identify the segments corresponding to
-croplands, resulting in a final vectorized field boundaries. Each system
-component is available on an open repository (see data and software
-availability section for details).
+The mapping approach we developed is comprised of four open source
+components (Figure \ref{fig:systemoverview}) that are designed to run in
+a cloud computing environment. The first component collects daily
+PlanetScope imagery and converts them into cloud-free seasonal
+composites. The second is a custom-built platform that provides tools
+for labelling the composites, along with procedures to assess and
+minimize label error. This platform interacts with the third component,
+a machine learning process, within an active learning (Cohn et al. 1994,
+Tuia et al. 2011) loop, to produce a map of predicted cropland
+probabilities for each image pixel. The fourth and final component is an
+algorithm that segments the image composites, then filters the resulting
+polygons using the pixel-wise cropland predictions produced by the
+active learning classifier, resulting in a final set of vectorized field
+boundaries.
 
@@
 
-\caption{An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
+\caption{An overview of the primary mapping components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them.
 The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.}\label{fig:systemoverview}
@@
 
+We describe each component in further detail in the following section,
+and how we applied them to map Ghana's annual cropland boundaries,
+excluding tree crops.
+
 \hypertarget{image-compositing}{%
@@
 
-The image processing components of our system were designed to work with
-PlanetScope Analytic surface reflectance data (PlanetTeam 2018).
-PlanetScope provides three visual (red, green, blue) and a near-infrared
-band at 3-4 m resolution at nominal daily frequency. Although these
-images are already pre-preprocessed and corrected for atmospheric
-effects, there are residual errors from inter-sensor differences and the
-radiometric normalization process (Houborg and McCabe 2018), variation
-in the orientation of scene footprints, as well as a high frequency of
-cloud cover over the study region (Wilson and Jetz 2016). To correct for
-these factors, we developed a procedure for creating temporal composites
-representing the primary growing and non-growing seasons within a single
-year.
-
-PlanetScope imagery is accessed via the Planet API (PlanetTeam 2018),
-and an initial order is placed for all imagery falling within the
-mapping geography and the date ranges for the two compositing periods.
-The imagery is collected and transferred directly to a cloud storage
-platform (Amazon Web Services {[}AWS{]} S3).
-
-Individual images are then transformed into analysis ready data (ARD)
-(Dwyer et al. 2018), by subsetting each downloaded image into 0.05
-degree tiles, regardless of cloud cover. Tiles are organized within a
-larger 1 degree resolution grid that covers the entire continent, which
-defines the minimum mapping area of interest (AOI; Figure
-\ref{fig:aois}).
-
-\begin{figure}
-
-{\centering \includegraphics[width=1\linewidth]{figures/figure2}
-
-}
-
-\caption{The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.}\label{fig:aois}
-\end{figure}
-
-The temporal compositing process is applied to the tiled daily images
-for the time period of interest, which in this case was one of two
-multi-month seasons, the primary growing and dry seasons for a single
-agricultural year. Imagery from two seasons helps to improve the
-performance of cropland classifiers (Debats et al. 2016), while having
-the seasons in the same year helps to minimize differences caused by
-land change. For each pixel in each image in each ARD temporal stack for
-a given season, two weights are calculated:
+The image processing component was designed for PlanetScope Analytic
+surface reflectance imagery (PlanetTeam 2018), which provides three
+visual (red, green, blue) and near-infrared bands at 3.7 m resolution at
+nominal daily frequency. The images are provided as ortho-rectified and
+converted to surface reflectance, although there are residual errors
+from inter-sensor differences and the radiometric normalization process
+(Houborg and McCabe 2018), variation in the orientation of scene
+footprints, as well as a high frequency of cloud cover over the study
+region (Wilson and Jetz 2016, Roy et al. 2021) that are not fully
+captured by the provided cloud masks. To minimize the effect of these
+residual errors, we developed a procedure for creating temporal
+composites of the primary growing and non-growing seasons within a
+single 12-month period. For Ghana, we defined the primary growing season
+as May through September, followed by the off (or dry) season from
+November or December through February. We chose these two seasons
+because prior work shows that the contrast between them improves
+cropland classifications (Debats et al. 2016), Furthermore, capturing
+the seasons in this sequence during the same year helps minimize
+differences caused by land change. The wide time intervals we used to
+define each season were necessary for collecting a sufficient number of
+images to make high quality composites, as Ghana's cloud cover renders
+many scenes unusable and therefore unavailable in Planet's catalog, thus
+the effective return interval can be substantially longer than 24 hours
+during the cloudiest months (Roy et al. 2021).
+
+We collected all available scenes intersecting Ghana and falling within
+these two seasons during the 2018 agricultural year (defined here as
+March, 2018-February, 2019) via the Planet API (PlanetTeam 2018), and
+transferred these to cloud storage (Amazon Web Services {[}AWS{]} S3).
+We then converted each scene into analysis ready data (Dwyer et al.
+2018) by cropping each to the boundaries of a 0.05\(^\circ\) grid that
+it intersected (see Figure S1 in Supplemental Information {[}SI{]}),
+which provided the dimensions for making composited image tiles. We
+chose this cell size for tiling because it is slightly narrower than the
+short axis of a PlanetScope scene, which increases the number of
+intersecting scenes that completely cover the tile, thereby helping to
+minimize edge artifacts in the composites.
+
+To create a seasonal composite, we calculated two weights for the time
+series of each pixel within the ARD stack for a given season:
 
@@
 
-Where \emph{t} is a particular date in near-daily time series of
-PlanetScope images, which begins at date 1 for the given compositing
-period and ends on date \emph{i}, \emph{blue} is the blue band, and
-\emph{NIR} the near infrared band. Equation \ref{eq:cloud} assigns lower
-weights to hazy and clouded pixels as the blue band is sensitive to haze
-and cloud pixels (Zhang et al. 2002), while Equation \ref{eq:shadow}
-assigns low weights to pixels in cloud shadow considering the
-significant darkening effect of the cloud shadows in the Near Infrared
-band (Zhu and Woodcock 2012, Qiu et al. 2020)
-
-Once these two weights are calculated, the final composited pixel value
-for each of the four PlanetScope bands is:
+Where \emph{t} is a particular date in the pixel time series, which
+begins at date 1 for the given compositing period and ends on date
+\emph{i}, \emph{blue} is the blue band, and \emph{NIR} the near infrared
+band. Equation \ref{eq:cloud} assigns lower weights to hazy and clouded
+pixels as the blue band is sensitive to these atmospheric features
+(Zhang et al. 2002), while Equation \ref{eq:shadow} assigns low weights
+to pixels in cloud shadow (Zhu and Woodcock 2012, Qiu et al. 2020)
+
+After assigning these two weights, we calculated the final composited
+pixel value:
 
@@
 
-Which is a weighted mean for each pixel for each band \emph{B} for the
-particular compositing period. The composited tiles were then added to
-the S3 store (Figure \ref{fig:systemoverview}), where they are stored as
-cloud-optimized geotiffs, and a ``slippy
+Which is the weighted mean for each pixel for each band \emph{B} for the
+given season.
+
+Each composited seasonal tile was saved as a cloud-optimized geotiff,
+and a ``slippy
 map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}''
-rendering is created for each composite using Raster Foundry (Azavea
-2020). The web-rendered imagery is presented within the training data
-platform (next section).
-
-\hypertarget{labelling-platform}{%
-\subsubsection{Labelling platform}\label{labelling-platform}}
-
-Training and reference data are collected by a custom labelling
-platform, which was originally designed for AWS's Mechanical Turk job
-marketplace (Estes et al. 2016a). The basic structure of the system
-remains the same, but we converted it into a standalone platform that
-allows us to enroll and pay people directly for their labelling, and is
-designed to control and supervise the machine learning process. The
-platform runs on a Linux virtual machine hosted on an AWS EC2 instance
-and is comprised of a database (PostGIS/Postgres), a mapping interface
-(OpenLayers 3), an image server (Raster Foundry), and a set of utilities
-for managing, assessing, and converting digitization work into
-rasterized labels for training a machine learning algorithm. Each
-instance of the platform focuses on a specific AOI (Figure
-\ref{fig:aois}A)
-
-The following sections provide an overview of the labelling platform's
-architecture.
-
-\hypertarget{mapping-workflow}{%
-\paragraph{Mapping workflow}\label{mapping-workflow}}
-
-\hypertarget{selecting-training-and-reference-sites}{%
-\subparagraph{Selecting training and reference
-sites}\label{selecting-training-and-reference-sites}}
-
-The labelling process begins with the random selection of a subset
-(e.g.~500) of cells from a 0.005 degree grid, with the selection itself
-potentially split into a training and validation sample, according to
-predetermined proportions. The grid, which is nested within the tiling
-and larger 1 degree grids (Figure \ref{fig:aois}C) defines the spatial
-unit for a labelling job. The selected cells are placed into a queue
-within the platform's database, and then converted into a mapping
-\emph{task} that has a specified number of \emph{assignments}
-(boundaries drawn by an individual labeller) that must be completed
-before the task is complete.
-
-\hypertarget{mapping-assignments}{%
-\subparagraph{Mapping assignments}\label{mapping-assignments}}
-
-Labellers registered in the system log in to the mapping platform (built
-with Flask) and navigate to the OpenLayers-based field mapping interface
-(Figure \ref{fig:labeller}), where they are presented with a white
-target box representing the randomly selected grid cell, a set of
-digitizing tools, and different image backdrops, including true and
-false color renderings of the growing season and off-growing season
-PlanetScope composites, and several virtual globe basemaps. Following a
-set of pre-defined digitizing rules (see SI), the labeller uses the
-polygon drawing tool to digitize the boundaries of all crop fields
-intersecting the target grid cell that are visible within the
-PlanetScope overlays. To aid with interpretation, the labeller can
-toggle between the PlanetScope renderings and the basemaps to help form
-a judgement about what constitutes a field. The labeller assigns each
-polygon a class category (e.g.~annual cropland), and upon completing all
-fields submits the assignment to the database. In cases where the target
-grid cell does not contain any fields, the labeller simply submits the
-assignment to mark it complete. The labeller is then directed to the
-next available assignment from a different labelling task.
+rendering was created for each composite using Raster Foundry (Azavea
+2020), for display within the labelling platform (next section).
+
+We generated a catalog of 16232 composite tiles (hereafter simply
+``tiles'') for Ghana, consisting of a seasonal pair for each of the 8116
+0.05\(^\circ\) tile grid cells covering Ghana. To assess the quality of
+the resulting composites, 50 tile grid cells were randomly selected, and
+two separate observers graded each corresponding seasonal composite
+using a four category that evaluated the degree of 1) residual cloud and
+2) cloud shadow, 3) the number of visible scene boundary artifacts, and
+4) the proportion of the image with resolution degraded below the 3.7 m
+PlanetScope resolution (e.g.~because of between-date image
+mis-registrations). Each category was qualitatively ranked from 0-3,
+with 0 being the lowest quality, and 3 the highest (see SI for complete
+protocol), making the highest possible score 12. We rescaled scores to
+fall between 0 and 1.
+
+\hypertarget{mapping-cropland-probabilities-with-active-learning}{%
+\subsection{Mapping cropland probabilities with active
+learning}\label{mapping-cropland-probabilities-with-active-learning}}
+
+The first step in creating a country-wide field boundary map of Ghana
+was to create a pixel-wise classification of cropland probabilities
+throughout the country. Given the high resolution of the imagery and the
+need to minimize the computational burden, we divided Ghana into 16
+distinct mapping regions, or Areas of Interest (AOIs). We constructed
+the AOIs by grouping together tile grids into blocks representing the
+larger 1\(^\circ\) cells used to assign tile identifiers (Figure S1A).
+We grouped tile cells from 1\(^\circ\) degree cells that overlapped
+Ghana's boundaries together with those from the nearest 1\(^\circ\) cell
+contained entirely within Ghana (with the exception of AOI 16, which was
+comprised of tile grids from the 1\(^\circ\) cells along Ghana's
+southern coast. The average extent of the resulting AOIs was 15,457
+km\(^2\) (range 12,160-23,535 km\(^2\)).
+
+We used the active learning process to develop a separate cropland
+classification model for each of these AOIs, based on an approach
+described by Debats et al (2017). We initiated the process by training a
+starter model using labels from a set of randomly selected training
+sites drawn from a 0.005\(^\circ\) grid that was nested within the
+tiling grid. This finer grid, which we refer to as the ``primary grid''
+for simplicity, provided the target area for creating labels (section
+2.2.1), as well as the unit for distributing computing jobs (section
+2.2.2). We then assessed the performance of the starter model against a
+separate set of validation labels developed for each AOI, applied the
+model to predict cropland probabilities for pixels in unlabelled primary
+grid cells in each AOI, and calculated an uncertainty criterion (Debats
+et al. 2017):
+
+\begin{equation}
+\mathrm{Q_I = \sum_{I(x, y) \epsilon I} (p(x, y) - 0.5)^2}
+\end{equation}
+
+Where Q is the uncertainty for each unlabelled primary grid cell I,
+calculated from the predicted probability \emph{p} of a randomly
+selected subset of pixels (x, y) drawn from it. Pixels with predicted
+probabilities closer to 0.5 are least certain as to their
+classification, thus the lowest values of \(Q\) represent primary grid
+cells posing the most difficulty for the classifier.
+
+We ranked the unlabelled primary grid cells from least to most certain,
+randomly selected a subset of cells from the top 30\% of the ranking (to
+minimize the risk of spatial autocorrelation), and sent these back to
+the labelling platform. After these new sites were labelled, they were
+added to the starter pool of labels, the model was retrained with the
+larger training set, its performance and prediction uncertainty was
+reassessed, and a new sample of the most uncertain primary grid cells
+was again sent for labelling. This loop repeated until model performance
+gains saturated or reached a pre-defined threshold, after which a final
+map of cropland probabilities was made for the AOI.
+
+In the next two sections, we describe the labelling and machine learning
+components of the active learning process in more detail.
+
+\hypertarget{labelling}{%
+\subsubsection{Labelling}\label{labelling}}
+
+To collect the initial randomized samples for model training, we grouped
+the AOIs (Figure S1A) into three clusters based on approximate
+agro-ecological similarity: the 6 northernmost savanna-zone AOIs
+(Cluster 1), a central to southeastern cluster (Cluster 2) consisting of
+the 3 middle (AOIs 7-9) and 2 southeastern AOIs (12 and 15), and a
+southwestern cluster (Cluster 3) made up of the forest zone AOIs (10,
+11, 13, 14, 16). Within each cluster, we randomly selected and labelled
+500 primary grid cells, which provided relatively large initial training
+samples for these agro-ecologically similar regions, while helping to
+minimize the overall amount of labelling effort. To create validation
+samples, we randomly selected and labelled 100 primary grid cells per
+AOI, and a further 100 cells were labelled in each AOI during each
+active learning iteration.
+
+In addition to training and validation labels, we also collected
+training reference labels and map reference labels (Elmes et al. 2020).
+The former were a set of 98 primary grid cells selected to represent the
+range of cropland types and densities in Ghana, which were labelled by
+expert analysts (the lead researchers on this project). We used these to
+assess the performance of the individual labellers collecting training
+and validation labels. Map reference labels were collected and used to
+assess the accuracy of the final map (see Section 2.4).
+
+We collected all labels using a custom-built platform that we adapted
+from an earlier prototype we developed for crowdsourced labelling (Estes
+et al. 2016a). We enhanced this platform by making several major
+additions, including an independent backend that allowed us to recruit
+and manage our own labelling teams, improved procedures for assessing
+and improving label accuracy, and processes for automating the machine
+learning component. The platform runs on a cloud-hosted Linux virtual
+server (AWS EC2) and is comprised of a database (PostGIS/Postgres), a
+mapping interface (OpenLayers 3), an image server (Raster Foundry), and
+a set of utilities for managing, assessing, and converting digitized
+field boundaries into rasterized labels.
+
+We created a separate labelling instance for each AOI. To create
+training and validation labels, labellers (the co-authors of this paper)
+logged into the website (built with Flask) for a particular AOI and
+navigated to the mapping interface (Figure \ref{fig:labeller}), where
+they were presented with a white target box representing a primary grid
+cell to label, a set of digitizing tools, and several different sources
+of imagery. These included true and false color renderings of the
+growing season and dry season PlanetScope composites, and several
+virtual globe basemaps. They then used the polygon drawing tool to
+digitize the boundaries of all crop fields visible within the
+PlanetScope overlays that intersect the target grid cell. For this
+project, labellers were instructed to digitize active or recently active
+crop fields, avoiding tree crops, and fallow or potentially abandoned
+fields (see SI for digitizing rules) To aid with interpretation,
+labellers toggled between the PlanetScope renderings and the basemaps to
+help form a judgement about what constitutes a field. The labeller
+assigned each digitized polygon a class category (e.g.~annual cropland),
+saved all completed fields to the database, and were then presented with
+the next target to label. If the target grid cell did not contain any
+fields, labellers simply pressed save to go to the next cell.
 
@@
 
-{\centering \includegraphics[width=0.95\linewidth]{figures/figure3}
+{\centering \includegraphics[width=0.95\linewidth]{figures/figure2}
 
@@
 
-\hypertarget{processing-completed-assignments}{%
-\subparagraph{Processing completed
-assignments}\label{processing-completed-assignments}}
-
-All submitted polygons are cleaned to fix topological irregularities
-that arose during digitization (see supporting information {[}SI{]}) and
-stored in a PostGIS table. Each completed assignment represents one of
-two types of tasks: 1) accuracy assessment, or 2) model training or
-validation. For the former type, an accuracy assessment routine is
-invoked that executes a series of comparisons between the labeller's
-results and a training reference dataset, resulting in a assignment
-score:
+The flow of labelling targets presented to each worker was determined by
+the platform's built-in scheduler. Each primary grid cell selected for
+labeling was placed into a queue within the platform's database, and
+converted into a labelling \emph{task} with a specified number of
+\emph{assignments} (the boundaries drawn by an individual labeller) that
+had to be completed in order to finish the task. There were two types of
+tasks, accuracy assessment or model training/validation, with the
+assignments for each indistinguishable to labellers. Upon completing an
+accuracy assessment assignment, the platform invoked a scoring algorithm
+that compared the labeller's digitized boundaries against a set of
+training reference polygons, resulting in a label quality score:
 
@@
 
-Over time, labellers are assessed multiple times across a range of
-accuracy tasks, which are selected to represent the variability of the
-agricultural system being mapped. Each labeller's score history is
-averaged to provide an overall accuracy measure, and this information is
-used for creating labels, the second task.
-
-If the labeller's completed assignment was a training/validation task,
-their maps remain stored in the database until the task's outstanding
-assignments are completed by other labellers. Once complete, another
-routine is invoked, which combines the task's completed assignments into
-a single consensus label using a Bayesian merging approach:
+We configured the platform's scheduler to present workers with accuracy
+assessment assignments at a rate of 1 for every 5 assignments mapped.
+This generated a history of accuracy assessment scores that we used to
+assess label quality and minimize label error.
+
+For training and validation, where there was no reference data to assess
+label accuracy, we set each task to have four assignments, i.e.~each was
+completed by four separate labellers. When all four assignments were
+complete, a Bayesian merging routine was invoked to combine the four
+sets of labels into a single consensus label:
 
@@
 field), \emph{D} is the label assigned to that pixel by a labeller, and
-\(W_i\) is an individual labeller. P(\(\theta\)\textbar D) is therefore
-the probability that the actual cover type is what the labellers who
-mapped it says it is, while P(W\(_i\)\textbar D) is an individual
-labeller's average score over all the accuracy assessment assignments
-they have completed, and P(W\(\theta\)\textbar D, \emph{W}\(_i\)) is the
-labeller's label for that pixel. This approach therefore uses the
-overall accuracy of each labeller to weight their labels when combined
-with those made by other labellers' for the same pixel (see SI for
-further details). As a further measure of confidence in the final
-consensus label, its average Bayesian Risk can be calculated (see SI).
-This measure ranges between 0 and 1, with 0 indicating full agreement
-between labellers for all pixels (n = 40000) in the label, and 1
+\(W_i\) is an individual labeller. P(\(\theta\)\textbar D) is the
+probability that the actual cover type is what the labellers who mapped
+it says it is, while P(W\(_i\)\textbar D) is the average score (ranging
+between 0 and 1) of the accuracy assessment assignments an individual
+labeller completed within the AOI, and P(W\(\theta\)\textbar D,
+\emph{W}\(_i\)) is the labeller's label for that pixel. This approach
+therefore used the average assignment quality score for to weight each
+labeller's label for a given pixel (see SI for further details). Each
+pixel in the target grid cell was merged using this approach (n =
+40000), which helps to minimize individual labellers' errors. We
+estimated a confidence measure for each consensus label by calculating
+its Bayesian Risk (see SI), which ranges between 0 and 1, with 0
+indicating full agreement between labellers for all pixels, and 1
 indicating complete disagreement.
 
-\hypertarget{classification-pipeline}{%
-\subsubsection{Classification pipeline}\label{classification-pipeline}}
-
-Upon completion of a batch of labels, the platform automatically
-launches an ephemeral Elastic Map
-Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf}
-cluster consisting of tens of instances, depending on the size of the
-AOI.
-
-\hypertarget{feature-extraction}{%
-\paragraph{Feature extraction}\label{feature-extraction}}
-
-The first step is the extraction of additional features from each
-seasonal image composite. Previous work showed that a large number of
-simple features that summarize the statistical properties of reflectance
-and vegetation indices in local neighborhoods are highly effective for
-classifying smallholder croplands (Debats et al. 2016). We followed this
-logic in this study, but were constrained to use a smaller feature space
-because the storage and memory requirements for our mapping geographies
-in this case were several orders of magnitude larger. For this
-implemention, we thus extract a set of 16 features, which are the mean
-and standard deviations calculated within an 11X11 and 5X5 moving
-window, respectively (initial tests revealed these two window sizes to
-be most effective), resulting in 24 overall features, including the
-original bands (Table 1).
+\hypertarget{cropland-classification-model}{%
+\subsubsection{Cropland classification
+model}\label{cropland-classification-model}}
+
+Upon completing each batch of labels, the platform automatically
+launched a machine learning cluster (Elastic Map
+Reduce\footnote{https://docs.aws.amazon.com/emr/latest/APIReference/emr-api.pdf})
+comprised of several hundred to a thousand CPUs, depending on the size
+of the AOI.
+
+The first step in the process was to derive a set of features from the
+image composites. Previous work showed that a large number of simple
+features summarizing image reflectance and vegetation indices within
+local neighborhoods were highly effective for classifying smallholder
+croplands (Debats et al. 2016). We followed that logic in this study,
+but used a smaller feature set because the storage and memory required
+for our mapping geographies were several orders of magnitude larger. For
+each seasonal composite, we calculated the mean and standard deviation
+of each band within an 11X11 and 5X5 moving window, respectively
+(initial tests revealed these two window sizes to be most effective).
+This provided an overall set of 24 features, including the unmodified
+bands of both composites (Table 1).
 
@@
 
-Feature extraction and the conversion of image features is handled by a
-combination of
+We used a combination of
 \texttt{GeoTrellis}\footnote{https://github.com/locationtech/geotrellis},
 \texttt{rasterio}\footnote{https://rasterio.readthedocs.io/en/latest/},
-and \texttt{RasterFrames}\footnote{https://rasterframes.io/}. These
-collectively extract subsets of imagery from the PlanetScope temporal
-composites, derive the features, and convert these into Apache Spark
-DataFrames. Features are extracted on the fly for each cell in the
-training and validation sets, a functionality enabled by storing the
-image composites as Cloud-optimized
-Geotiffs\footnote{https://www.cogeo.org/} (COGs).
-
-\hypertarget{classification}{%
-\paragraph{Classification}\label{classification}}
-
-Once the features from the training sites are extracted into
-\texttt{RasterFrames}, these are combined with their corresponding
-labels and passed to the machine learning classifier, a
-\texttt{SparkMLlib} implementation of Random Forests (Breiman 2001). For
-this study, the model was trained with a balanced sample and a tree
-depth of 15 and total tree number of 60, which initial testing showed to
-provide a reasonable balance between computational time and performance.
-
-After fitting, the model is applied to the features of the model
-validation set, and a set of performance metrics is calculated,
-including binary accuracy, the F1 score (the geometric mean of precision
-and recall), and the area under the curve of the Receiver Operating
-Characteristic (Pontius and Si 2014).
-
-\hypertarget{the-active-learning-loop}{%
-\subsubsection{The active learning
-loop}\label{the-active-learning-loop}}
-
-After fitting and model evaluation a second prediction is undertaken to
-enable active learning. The feature extraction process is repeated for
-the rest of the mapping geography that falls outside of the training and
-validation sample, but applied to a subset of randomly drawn pixels from
-each cell in order to reduce computational demand. The fitted model is
-applied to predict the cropland probability for these selected pixels,
-and an uncertainty criterion (Debats et al. 2017) is calculated for each
-grid cell:
-
-\begin{equation}
-\mathrm{Q_I = \sum_{I(x, y) \epsilon I} (p(x, y) - 0.5)^2}
-\end{equation}
-
-Where Q is the uncertainty for grid cell I, calculated from the
-predicted probability \emph{p} of the randomly selected subset of pixels
-(x, y) drawn from it. Pixels with predicted probabilities closer to 0.5
-are least certain as to their classification, thus images with the
-lowest values of \(Q\) represent sites posing the most difficulty for
-the classifier.
-
-After scoring with the uncertainty criterion, the top \emph{N} most
-uncertain grid cells are selected and sent back to the labelling
-platform, which are then digitized by the labellers. The resulting
-consensus labels from the actively selected sample are added to the
-initial randomly selected sample, and a new cluster is launched. The
-model is retrained, assesses uncertainty across the remaining unlabelled
-sites, and selects the next most uncertain sites for labelling. This
-loop repeats until model performance gains against the validation set
-show diminishing returns.
+and \texttt{RasterFrames}\footnote{https://rasterframes.io/} to derive
+the features on the fly (which was enabled by converting the composites
+to Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/}) and
+convert them into Apache Spark DataFrames. The extracted features were
+combined with their corresponding training and validation labels and
+passed to the machine learning classifier, a \texttt{SparkMLlib}
+implementation of Random Forests (Breiman 2001). We trained the model
+with a balanced sample and a tree depth of 15 and total tree number of
+60, which initial testing showed to provide a reasonable balance between
+computational time and performance.
+
+\hypertarget{model-performance}{%
+\subsubsection{Model performance}\label{model-performance}}
+
+To assess performance of the Random Forests classifier, we used the
+validation sample to calculate binary accuracy, the F1 score (the
+geometric mean of precision and recall), and the area under the curve of
+the Receiver Operating Characteristic (Pontius and Si 2014), as well as
+the false positive rate. We calculated these measures each time the
+model was retrained for a given AOI, in order to assess the change in
+classifier performance with each active learning iteration.
+
+To evaluate whether active learning improved model performance relative
+to randomized label selection, we ran an additional test within three
+AOIs (1, 8, and 15), in which we retrained the model with 100 randomly
+selected labels for each iteration. We then compared the differences in
+accuracy, AUC, and F1 between the actively and randomly trained models
+(Debats et al. 2017).
+
+To quantify the potential impact of label error on classification
+results, we conducted two further analyses. We evaluated the performance
+differences between models trained with three different sets of labels:
+1) those from the lowest scoring labeller to map each training site, 2)
+those from the highest scoring labeller, and 3) the consensus labels. We
+also calculated the correlations between the mean Bayesian Risk of
+labels in each AOI and the corresponding model performance metrics
+(Table S3).
 
 \hypertarget{segmentation}{%
-\subsubsection{Segmentation}\label{segmentation}}
-
-After the final iteration, the segmentation algorithm is invoked, which
-entails several steps. In the first step, the meanshift algorithm
-(Yizong Cheng 1995) is applied to the original bands of the dry season
-composite. A Sobel filter is then applied to the green, red, and
-near-infrared mean-shifted bands and the probability map, and a combined
-edge image is computed using the sum of these four edge images for the
-dry season only. A compact watershed algorithm (Neubert and Protzel
-2014) is then run on the weighted edge image, with a high level of
-segmentation specified. In this case, we specified \textbf{6400}
-segments per tile.
-
-Third, a region adjacency graph is constructed for each image tile, in
-which each node represents all pixels within each polygon created in the
-previous step. The edge between two adjacent regions (polygons) is
-calculated as the norm of the difference between the means of normalized
-colors of all bands. Hierarchical merging is then applied, in which the
-most similar pairs of adjacent nodes are merged until there are no edges
-remaining below a predetermined threshold of 0.05.
-
-In the fourth step, the merged polygons are overlaid with the posterior
-probabilities resulting from the final active learning loop, and
-polygons in which the average posterior probability is greater than a
-predetermined threshold (here 0.5, but could vary locally) are retained
-as field polygons.
-
-In the final step, the retained polygons are refined by removing holes
-and smoothing their boundaries using the Visvalingam algorithm
-(Visvalingam and Whyatt 1993). Neighboring polygons that overlap along
-tile boundaries are then merged.
-
-To assess the accuracy of the final segmented boundaries, we used a
-two-step approach. First, we assessed the overall thematic accuracy of
-the resulting classification against our map reference data. Second, to
-assess the quality of the segmentation, we compared the mean area and
-relative frequencies of the segmented polygons within different size
-classes against the same metrics derived from the digitized fields of
-the most accurate worker to create the given map. We selected this
-relatively simple procedure, as opposed to more complex measures of
-object accuracy (Ye et al. 2018), because, on the one hand, both the
-automated segmentation algorithm and labellers are cueing in on the same
-features--abrupt, physically detectable breaks within the imagery. On
-the other hand, no matter how well the intepreted/segmented boundaries
-align with the boundaries of fields in the imagery, it is logistically
-difficult to evaluate performance against real-world boundaries as the
-spectral distinction of field boundaries will vary across different crop
-types and land use arrangements.
-
-\hypertarget{applying-the-system-to-map-ghana}{%
-\subsection{Applying the system to map
-Ghana}\label{applying-the-system-to-map-ghana}}
-
-We applied the system to map Ghana's croplands, excluding areas
-primarily cultivated with tree crops. Ghana has several distinct
-agricultural regions, ranging from the primarily grain and vegetable
-crop producing regions in the northern savannas to tree crop-dominated
-system in the forested southwest, where cocoa and oil palm are among the
-dominant crops. For these latter regions, we did not attempt to classify
-tree crops, and instead mapped clearings that potentially contain field
-crops or newly felled or recently replanted tree crops. We made this
-decision because PlanetScope's resolution is not high enough for
-labellers to distinguish many tree crops from surrounding forest, and
-the boundaries of many tree crops (e.g.~cocoa) are often not visible.
-
-To create the cropland maps, we divided the country into 16 distinct
-AOIs, which were developed by grouping together each one degree cell
-fully contained within Ghana with the tiles belonging to any adjacent
-degree cell that overlapped neighboring countries (Figure
-\ref{fig:aois}A). The exception was AOI 16, which consisted of the four
-degree cells intersecting Ghana's southern coast. The resulting AOIs
-ranged from 12,160 to 23,535 km\(^2\) in extent (average = 15,457
-km\(^2\)), A separate active learning and segmentation process was run
-for each of these AOIs.
-
-To collect the initial randomized samples for model training, we grouped
-the AOIs into three clusters: a northern cluster comprising the 6
-northernmost AOIs (Cluster 1), a central to southeastern cluster
-(Cluster 2) consisting of the 3 middle (AOIs 7-9) and 2 southeastern
-AOIs (12 and 15), and a southwestern cluster (Cluster 3) made up of the
-forest zone AOIs (10, 11, 13, 14, 16). Within each cluster, we randomly
-selected and labelled 500 grid cells, which provided relatively large
-initial training samples for these agro-ecologically similar regions,
-while helping to minimize the overall amount of labelling effort. In
-addition to these samples, we randomly selected and labelled 100 grid
-cells within each AOI to provide a validation sample.
-
-After collecting the initial training and validation samples, we trained
-a starter model for each cluster and applied it to each of the block's
-AOI. For each iteration, 100 samples were actively selected within each
-AOI, and added to the training pool.
-
-During the collection of training and validation samples, labellers were
-tasked to only digitize active or recently active crop fields, avoiding
-tree crops, and fallow or potentially abandoned fields (see SI for the
-digitizing rules).
-
-To evaluate the performance of the system, we performed several analyses
-described in sections 2.3.1-4.
-
-\hypertarget{image-quality}{%
-\subsubsection{Image quality}\label{image-quality}}
-
-We evaluated the overall quality of the resulting seasonal image
-composites by assessing a random selection of 50 tiles. We graded both
-seasonal composites for each tile using a four category quality score,
-which evaluated the degree of 1) residual cloud and 2) cloud shadow, 3)
-the number of visible scene boundary artifacts, and 4) the proportion of
-the image that had its resolution degraded below the typical 3-4 m
-PlanetScope resolution (e.g.~because of between-date image
-mis-registrations). Each category was qualitatively ranked from 0-3,
-with 0 being the lowest quality, and 3 the highest (see SI for complete
-protocol), making the highest possible score 12. We rescaled scores to
-fall between 0 and 1.
-
-\hypertarget{model-gains-per-iteration}{%
-\subsubsection{Model gains per
-iteration}\label{model-gains-per-iteration}}
-
-To assess the gain in model performance due to active learning, we
-measured the change in accuracy, F1, and AUC (see 2.2.3.2) between each
-iteration and between the first and last iterations for each AOI.
-
-To evaluate whether active learning improved model performance relative
-to a purely random approach to selecting new training sites, we ran
-additional tests within a subset of AOIs (1, 8, and 15). We first
-randomly selected and labelled 300-400 sites in each AOI. We then
-progressively added 100 of the randomly selected samples to the relevant
-training pool and retrained the model, repeating the process so that the
-number of iterations and samples matched those from the active learning
-process. We then compared the difference in accuracy, AUC, and F1
-between the randomly trained models and those trained with active
-learning (Debats et al. 2017).
-
-\hypertarget{accounting-for-label-error}{%
-\subsubsection{Accounting for label
-error}\label{accounting-for-label-error}}
-
-To quantify the potential impact of label error on classification
-results, we evaluated the performance differences between models trained
-with three different sets of labels: 1) those from the lowest scoring
-labeller to map each training site, 2) those from the highest scoring
-labeller, and 3) the consensus labels. This assessment follows
-recommended Tier 1 (i.e.~best practice) standards to account for
-training data errors (Elmes et al. 2020).
-
-\hypertarget{accuracy-assessment}{%
-\subsubsection{Accuracy assessment}\label{accuracy-assessment}}
-
-The model performance assessments described above (2.3.2-3) were not
-fully independent because they used the same validation sites over
-multiple iterations (Elmes et al. 2020). To independently assess the
-accuracy of our final map products, we followed recommended guidelines
-(Stehman and Foody 2019) to create a separate map reference sample. We
-used a stratified design, randomly assigning square polygons of
-\textasciitilde0.1 ha extent into cropland and non-cropland strata,
-developed the map of segmented field boundaries. Four classes were used
-for the map reference sample: cropland; non-cropland; unsure but likely
-cropland; unsure but likely non-cropland. The latter two classes were
-used to provide insight into the degree of uncertainty in the map
-reference sample. For efficiency, two separate supervisors evaluated
-separate portions of the reference sample, but both jointly assessed a
-small subset of the sample. We calculated their level of agreement on
-this subset to provide an additional assessment of uncertainty in the
-map reference sample (Stehman and Foody 2019). The SI contains further
-details on the design and collection of the map reference sample.
-
-The map reference polygons were then intersected with both the
-probability images and the segmented field boundaries, and confusion
-matrixes between the map reference labels and the extracted map classes
-were constructed to assess the categorical accuracy of each map product.
-We calculated the overall accuracy for each map, as well as the
-class-wise User's and Producer's accuracy, as well as the 95\%
-confidence intervals for each accuracy measure (Olofsson et al. 2013,
-2014, Stehman and Foody 2019).
-
-To assess the accuracy of the segmented field boundaries, we compared
-the size class distributions of the segmented field boundaries against
-those of the workers' digitized polygons at map validation sites. We
-chose this approach because of existing uncertainties in polygon-based
-accuracy assessment methods (Ye et al. 2018), and because the map's
-ability to represent field sizes was of greatest interest. To undertake
-this comparison, we selected the field polygons from the most accurate
-labeller to digitize each of the 100 validation sites in each AOI, and
-calculated the site-wise average area and number of polygons. We then
-calculated the same statistics from the segmented boundaries that
-intersected each validation grid. We compared the distributions and
-proximity of two measures of central tendency (mean and median)
-calculated from the two datasets for each AOI, and across all AOIs.
-
-\hypertarget{assessing-the-characteristics-of-ghanaian-cropland}{%
-\subsection{Assessing the characteristics of Ghanaian
-cropland}\label{assessing-the-characteristics-of-ghanaian-cropland}}
-
-Using the final mapped results, we calculated the estimated area of
-cropland in Ghana, as well as the average size and total number of
-fields in the different AOIs. We used the map reference sample to
-calculate adjusted area estimates and confidence intervals for each map
-class, and used the differences between labellers' polygons and
-segmented boundaries at validation sites to calculate bias-adjusted
-estimates of mean field sizes and the total number of fields.
+\subsection{Segmentation}\label{segmentation}}
+
+Upon completion of the active learning process, we deployed a five-step
+algorithm to create a segmented map of field boundaries. In the first
+step, we identified edge features within the imagery. To do this, we
+applied the meanshift algorithm (Yizong Cheng 1995) to each dry-season
+composite tile, and then passed a Sobel filter over the mean-shifted
+green, red, and near-infrared bands, and the corresponding map of
+predicted cropland probabilities. We then summed the four resulting edge
+images to produce a combined edge image.
+
+In the second step, we used a compact watershed algorithm (Neubert and
+Protzel 2014) to segment the edge image, specifying a high number of
+segments (6,400) segments per tile, so that the mean segment size
+(\textless0.5 ha) was finer than the expected mean field size
+(\textgreater1 ha).
+
+In the third step, we hierarchically merged the resulting polygons. We
+first constructed a region adjacency graph for each tile, with each node
+representing all image pixels within each polygon. The edge between two
+adjacent regions (polygons) was calculated as the difference between the
+means of the normalized colors of all bands. We then merged the most
+similar pairs of adjacent nodes until there were no edges remaining
+below the predetermined threshold of 0.05.
+
+In the fourth step, we overlaid the merged polygons with the cropland
+probability images, and polygons in which the mean probability was
+greater than 0.5 were retained as crop fields.
+
+In the fifth and final step, we refined the crop field polygons, by
+removing holes and smoothing boundaries using the Visvalingam algorithm
+(Visvalingam and Whyatt 1993). We then merged neighboring polygons that
+overlapped along tile boundaries.
+
+The resulting map represents dry season crop field boundaries, as we did
+not segment growing season images. We made this choice because labels
+were primarily drawn on dry season composites, when boundaries were
+typically more visible.
+
+\hypertarget{map-assessment}{%
+\subsection{Map assessment}\label{map-assessment}}
+
+We followed recommended guidelines (Stehman and Foody 2019) to conduct
+an independent assessment of the categorical accuracy of the final maps,
+using a set of 1207 (487 cropland; 720 non-cropland) point-based, map
+reference labels, which were placed across Ghana using a stratified
+random sample design, and collected through the labelling platform by
+two expert supervisors (see SI for full details on sample design and
+collection). For efficiency, the supervisors labelled separate portions
+of the sample, but overlapped on a small subset (n = 23). We calculated
+the label agreement (87\%) on this subset to estimate uncertainty in the
+map reference sample (Stehman and Foody 2019). In addition to this, the
+sample was labelled with four classes: cropland; non-cropland; unsure
+but likely cropland; unsure but likely non-cropland. The last two
+classes, which constituted 15.7\% of the sample, provided a further
+measure of uncertainty in the map reference sample
+
+We used the sample to calculate the overall accuracy for each map, the
+class-wise User's and Producer's accuracy, and the 95\% confidence
+intervals for each accuracy measure (Olofsson et al. 2013, 2014, Stehman
+and Foody 2019). We calculated these measures across the entire country,
+as well as several different zones, to evaluate regional difference in
+accuracy. We defined two sets of zonations (Figure S4), each containing
+four zones, the first created by grouping 1) the three northern AOIs
+(1-3), 2) the six central AOIs (4-9), 3) the four southwestern AOIs (10,
+11, 13, 14, 16), and 4) the two southeastern zones (13, 15). This
+grouping differs from the three clusters used to collect initial model
+training samples, as we designed these to divide the country more
+finely, and to isolate the less forested southeastern third of Ghana
+from the more forest northwest. The second zonation was developed by
+grouping the country's eight agro-ecological zones into four broader
+clusters (Figure S4B). We applied this zonation only to the per-pixel
+classification, to better understand patterns of error in the model.
+
+To assess how effectively the segmentations captured field
+characteristics, we compared the size class distributions of the
+segmented field boundaries against the field boundaries digitized by the
+labellers' over the 100 validation sites in each AOI. We chose this
+approach because of existing uncertainties in polygon-based accuracy
+assessment methods (Ye et al. 2018), and because the map's ability to
+represent field sizes was of greatest interest. To undertake this
+comparison, we selected the polygons from the most accurate labeller to
+digitize the 100 validation grids in each AOI, and calculated the
+average area and number of polygons in each cell. We then calculated the
+same statistics from the segmented boundaries that intersected each
+validation grid, and compared the two sets of statistics.
+
+We used the final maps to evaluate the characteristics of Ghana's
+croplands. We calculated the estimated area of cropland in Ghana, as
+well as the average size and total number of fields in the different
+AOIs. We used the map reference sample to calculate adjusted area
+estimates and confidence intervals for each map class, and used the
+differences between labellers' polygons and segmented boundaries at
+validation sites to calculate bias-adjusted estimates of mean field
+sizes and the total number of fields.
 
@@
 
-We developed maps of Ghana's cultivated croplands within an area of
-247,299 km\(^2\), which included portions of neighboring countries
-overlapped by images tiles.
-
-\hypertarget{image-catalog-and-quality}{%
-\subsection{Image catalog and quality}\label{image-catalog-and-quality}}
-
-To develop the maps, we first generated an image catalog for the 8,116
-tiles covering Ghana. This entailed processing all PlanetScope imagery
-intersecting these tiles between May-September, 2018 (the growing
-season) and December, 2018 to February, 2019 (the subsequent dry
-season). The longer period was necessary for the growing season because
-of the frequent cloud cover, which substantially limits the number of
-clear scenes for any tile (Figure S3). For the cloudiest regions (AOIs
-10, 11, 13, 14, 16) we started the dry season window in November.
-
-An assessment with two observers (see SI for observer details) found
-that average quality per growing season composite tile was 0.88, with 70
-percent having scores \(\geq\) 0.85, while the average quality of dry
-season composites was 0.92 (74 percent \(\geq\) 0.85). Composite quality
-in both seasons was highest in the northern half of the country and
-lowest in the southwest (Figure \ref{fig:imqual}A), where the
-substantially greater cloud cover resulted in a much lower density of
-available PlanetScope imagery for each time period (Figure S3).
+Our results produced two separate maps of Ghana's annual croplands, over
+a total area of 248,343 km\(^2\) that included portions of the
+neighboring countries overlapped by image tiles.
+
+\hypertarget{image-quality}{%
+\subsection{Image quality}\label{image-quality}}
+
+The assessment of image composites found that their quality in both
+seasons was highest in the northern half of the country and lowest in
+the southwest, (Figure \ref{fig:imqual}A), where the substantially
+greater cloud cover resulted in a much lower density of available
+PlanetScope imagery for each time period (Figure S5). The average
+quality score of growing season composites was 0.88, with 70 percent
+having scores \(\geq\) 0.85 (out of 1; Figure \ref{fig:imqual}B), while
+the mean score of dry season composites was 0.92 (74 percent \(\geq\)
+0.85).
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{figures/figure4}
+{\centering \includegraphics[width=0.7\linewidth]{figures/figure3}
 
@@
 
-\hypertarget{active-learning}{%
-\subsection{Active learning}\label{active-learning}}
-
-\hypertarget{training-data-collection}{%
-\subsubsection{Training data
-collection}\label{training-data-collection}}
-
-After training of models with the initial randomly selected label sets,
-the active learning process was run for 3 iterations for 12 of the 16
-AOIs, resulting in 800 labels per AOI. AOIs 10 and 14 stopped after one
-and two iterations, respectively, as they started with high initial
-validation accuracies (\textgreater83\%) and showed little subsequent
-improvement. The models for these two AOIs were thus trained with 600 -
-700 samples. AOI 15 was run for 4 iterations (900 samples), while AOI 3
-underwent a second active learning cycle because the model produced
-during the first cycle was inaccurate (see SI). In this second run, 300
-initial training sites randomly selected within the AOI were used
-(Figure S4A), followed by 2 subsequent active learning iterations,
-resulting in a training sample of 500. Labels collected during the
-active learning iterations showed distinct patterns in several AOIs,
-which often fell along ecotones, such as the boundaries between
-agroecozones (see Figure S4A). The total number of unique training and
-validation sites across the country were 6,299 and 1,600, respectively.
-
-The distribution of training and validation sample collection effort was
-divided across 20 labellers, with a core group of 13 who mapped more
-than 1,000 sites each. As each training site was mapped by 4 separate
-labellers, 34,014 sets of vector labels were made. Each labeller
-digitized an average of 2,001 (see Figure S5A for more details on
-labelling effort). Labeller accuracy was scored 9,389 times against 98
-unique training reference sites (Figure S4A), with each labeller
-assessed an average of 552 times at a rate of 1 training reference site
-for every 3.62 training site mapped. The mean of each labeller's average
-accuracy score was 0.71 (range 0.6 to 0.85; see Figure S5B for detailed
-score distributions).
-
-After each site was mapped by four labellers, consensus labels were
-generated. The Bayesian Risk (see SI) of each training and validation
-label was calculated as an additional measure of label quality. The
-average risk was 0.122 for training labels and 0.127 for validation
-labels. Risk was highest in the northen AOIs (AOIs 1-6; Figures S6-7),
-falling between 0.157 for training and 0.173 for validation labels
-(Figures S6-7), and lowest in the southwestern AOIs (AOIs 10, 11, 13,
-14, 16; training risk = 0.079; validation risk = 0.065). Label risk in
-the central-southeastern AOIs (AOIs 7-9, 12, 15) was slightly lower
-(training = 0.127; validation = 0.136) than in the north. Labeller
-experience also appeared to reduce risk, which we observed during a
-relabelling of the 500 initial random site in this cluster (see SI); the
-mean risk of the updated labels was 0.055, compared to 0.172 for
-original labels.
+\hypertarget{cropland-probabilities}{%
+\subsection{Cropland probabilities}\label{cropland-probabilities}}
+
+To make the initial maps of cropland probabilities, the active learning
+process ran for 3 iterations in 12 of 16 AOIs, varying from as little as
+1 to as many as 4 iterations across the other 4 AOIs, with the number of
+iterations varying according to the performance of the starter models
+(i.e.~AOIs with higher starting performance stopped after fewer
+iterations, see SI). Each AOI's model was trained by 300-500 randomly
+selected labels (Figure S6A), plus an additional 600 - 900 (typically
+800) within each the AOI that were selected by active learning. Actively
+selected labels showed distinctive patterns in several AOIs (Figure
+S6B), such as concentrating along ecotones or the boundaries of
+agro-ecological zones. A total of 6,299 training and 1,600 validation
+labels were collected by 20 labellers to develop and assess model
+performance (Figure S7).
 
@@
 
-Model performance was calculated for each iteration within each AOI. The
-average accuracy, AUC, and F1 at iteration 0 were 0.786, 0.809, and
-0.464, respectively, increasing to 0.825, 0.818, and 0.507 by iteration
-3 (Figure \ref{fig:alperformance}). These differences represent
-respective gains of 4.9, 1.1, and 9.1 percent for the three metrics. The
-largest gains for each metric occurred on iteration 1, averaging 2.9, 1,
-and 3.8 percent for accuracy, AUC, and F1, while the lowest gains were
-realized on iteration 3, with accuracy, F1, and AUC respectively
-increasing by just 1.2\%, 0.9\%, and 0.3\%. The scores achieved on the
-final iteration varied substantially across AOIs and metrics. Accuracy
-ranged between 0.725 (AOI 15) and 0.948 (AOI 16), while AUC varied from
-0.725 (AOI 4) and 0.93 (AOI 11), and F1 from 0.252 (AOI 13) and 0.636
-(AOI 8).
+The performance of the Random Forest classifier typically improved with
+each active learning iteration. The average accuracy, AUC, and F1 at
+iteration 0 were 0.786, 0.809, and 0.464, respectively, increasing to
+0.825, 0.818, and 0.507 by iteration 3 (Figure \ref{fig:alperformance}).
+These differences represent respective gains of 4.9, 1.1, and 9.1
+percent for the three metrics. The largest gains for each metric
+occurred on iteration 1, averaging 2.9, 1, and 3.8 percent for accuracy,
+AUC, and F1, while the lowest gains were realized on iteration 3, with
+accuracy, F1, and AUC respectively increasing by just 1.2\%, 0.9\%, and
+0.3\%. The scores achieved on the final iteration varied substantially
+across AOIs and metrics. Accuracy ranged between 0.725 (AOI 15) and
+0.948 (AOI 16), while AUC varied from 0.725 (AOI 4) and 0.93 (AOI 11),
+and F1 from 0.252 (AOI 13) and 0.636 (AOI 8).
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/figure5}
+{\centering \includegraphics[width=1\linewidth]{figures/figure4}
 
@@
 
-The comparison of active versus randomized training sample collection
-(in AOIs 1, 8, and 15) showed that the former approach outperformed the
-latter. After three iterations, the accuracy, AUC, and F1 scores
-resulting from active learning were respectively 0.8, 0.6, and 2.3
-percent higher than the scores from a randomly trained model (Figure
-S8). However, there was more variability in earlier iterations, with
-average score differences of -1.7 (accuracy), 0.6 (AUC), and 0.8 percent
-(F1) after iteration 1, and -0.3 (accuracy), 0.4 (AUC), and 1.8 (F1)
-percent after iteration 2. The negative results for accuracy was caused
-by results at AOI 15, where active learning accuracy was 8.37 percent
-lower than random training after iteration 1 (see Figure
-\ref{fig:alperformance}). In comparison, iteration 1 active learning
-accuracies were 2.88 and 0.45 percent higher than random training for
-AOIs 1 and 8, respectively. Accuracy under active learning for AOI 15
-exceeded randomized training after 4 iterations.
-
-\hypertarget{the-impact-of-training-data-error}{%
-\subsubsection{The impact of training data
-error}\label{the-impact-of-training-data-error}}
-
-The potential impact of label errors on map quality was assessed in four
-AOIs (1, 2, 8, and 15). The results of these tests showed that the
+The experiment conducted in three AOIs (in AOIs 1, 8, and 15) showed
+that training models with active learning improved performance compared
+to randomized approaches to label selection. After three iterations, the
+accuracy, AUC, and F1 scores for the actively trained models were
+respectively 0.8, 0.6, and 2.3 percent higher than those for randomly
+trained models (Figure S8). However, there was more variability in
+earlier iterations, with average score differences of -1.7 (accuracy),
+0.6 (AUC), and 0.8 percent (F1) after iteration 1, and -0.3 (accuracy),
+0.4 (AUC), and 1.8 (F1) percent after iteration 2 (see SI for more
+details).
+
+\hypertarget{the-impact-of-label-error-and-uncertainty-on-model-performance}{%
+\subsubsection{The impact of label error and uncertainty on model
+performance}\label{the-impact-of-label-error-and-uncertainty-on-model-performance}}
+
+We used the two measures of label quality calculated by the platform,
+the average quality score of each labeller and Bayesian Risk (or simply
+``label risk''), to assess the potential impacts of label error on model
+performance. The average of each labeller's AOI-specific accuracy score
+was 0.71 (range 0.6 to 0.85; see Figures S4 and S5 for details on label
+scores and number of assignments per labeller). The average Bayesian
+Risk was 0.124, with highest label risk (0.165) in the northern AOIs
+(AOIs 1-6; Figures S6-7), lowest (0.165) in the southwestern AOIs (AOIs
+10, 11, 13, 14, 16), and intermediate (0.131) in the
+central-southeastern AOIs (AOIs 7-9, 12, 15).
+
+Treating each labeller's average label quality scores (Figure S9) as a
+proxy for error, we used these scores to develop training sets to test
+the impact of label error on model performance. The results of these
+tests, which were conducted in AOIs 1, 2, 8, and 15, showed that the
 average accuracy, AUC, and F1 scores for models trained with the
@@
 
-{\centering \includegraphics[width=1\linewidth]{figures/figure6}
+{\centering \includegraphics[width=1\linewidth]{figures/figure5}
 
@@
 
-A second measure of the impact of label error is found within the
-correlations between the mean label risk per AOI and the model
-performance metrics (Table S3). Accuracy and AUC had strong (Spearman's
-Rank Correlation = -0.824 to moderate (r = -0.568) negative correlations
-with label risk, while F1 had a weaker but moderate positive association
-(r = 0.456). The positive sign of the latter relationship is
-counter-intuitive, but is explained by risk's association with
-precision, one of two inputs to F1, which was moderately positive
-(0.629), whereas risk had a negligible correlation with recall (0.206),
-F1's other component. The correlation between risk and the false
-positive rate (0.688), another important performance metric, shows that
-labelling uncertainty may increase model commission error.
+Correlations (Table S3) between the mean label risk per AOI (Figures
+S10-11) and model performance metrics showed strong (Spearman's Rank
+Correlation = -0.824) to moderate (r = -0.568) negative correlations
+between label risk and accuracy and AUC, respectively, while F1 had a
+weaker but moderate positive association (r = 0.456). The positive sign
+of the latter relationship is counter-intuitive, but is explained by
+risk's association with precision, one of two inputs to F1, which was
+moderately positive (r = 0.629), whereas risk had a negligible
+correlation with recall (r = 0.206), F1's other component. The
+correlation between risk and the false positive rate (r = 0.688),
+another important performance metric, shows that labelling uncertainty
+may increase model commission error.
 
@@
 
-We used a map reference sample of 1207 sites (487 cropland; 720
-non-cropland) to evaluate the accuracy of the per-pixel classifications
-(resulting from thresholding the Random Forests probability), as well as
-the segmented field boundary maps. We first evaluated the uncertainty in
-the map reference classes by assessing 1) the overall agreement between
-map reference labels collected by two separate supervisors at 23 sites,
-and 2) the confidence of the labels assigned by the supervisors (see SI
-for details). The first measure showed that the two individual
-supervisors' labels agreed at 87\% of common sites, while the second
-showed that 15.7 of sites were labelled with the two classes that
-indicated a level of uncertainty.
-
-We found that the overall accuracy of the pixel-wise classifications was
+We used the map reference sample to evaluate the accuracy of the
+cropland probability map (after classifying it using a threshold
+probability of 0.5) and the map of segmented field boundary maps. We
+found that the overall accuracy of the pixel-wise classifications was
 88\% against this map reference sample (Table \ref{tab:mapaccuracy}).
-Confining the map reference sample to four distinct zones (Figure S10A)
+Confining the map reference sample to four distinct zones (Figure S4A)
 shows that overall accuracy ranged from 83.3\% in Zone 1 (AOIs 1-3) to
@@
 of the cropland class was 61.7\% across Ghana, ranging from 45.6\% in
-Zone 3 to 67.9\% in Zone 1, while the User's accuracy for was 67.3\%
+Zone 3 to 67.9\% in Zone 1, while the User's accuracy was 67.3\%
 overall, ranging from 59.8\% in Zone 4 to 71.2\% in Zone 1. Both
@@
 
-\hypertarget{segmentation-accuracy}{%
-\subsubsection{Segmentation accuracy}\label{segmentation-accuracy}}
+\hypertarget{segmentation-quality}{%
+\subsubsection{Segmentation quality}\label{segmentation-quality}}
 
@@
 polygons. This discrepancy was primarily caused by results in four AOIs
-(2, 3, 7, and 15; Figure S11), where segments averaged between 7.76 and
+(2, 3, 7, and 15; Figure S13), where segments averaged between 7.76 and
 10.76 ha, compared to 2.18 - 2.77 ha for the corresponding
@@
 site averaged 3.08 (median = 2.66; StDev = 2.9) compared to 4.4 (median
-= 3.38; StDev = 4.52) for digitized polygons (Figure S12).
+= 3.38; StDev = 4.52) for digitized polygons (Figure S14).
 
@@
 third of AOI 1 also has noticeable densities of cropland. Several
-notable areas of low cropland density are also apparent, indicating the
-presence of large protected areas, such as Mole National Park in the
-southeastern corner of AOI 1 and Digya National Park in the northwestern
-corner of AOI 12. In contrast, the relative absence of cropland in AOIs
-13, 14, and 16 does not reflect the scarcity of agriculture in these
-areas, but rather the predominance of tree crops, which we did not map.
+prominent areas of low cropland density indicate the presence of large
+protected areas, such as Mole National Park in the southeastern corner
+of AOI 1 and Digya National Park in the northwestern corner of AOI 12.
+The relative absence of cropland in AOIs 13, 14, and 16 does not reflect
+the scarcity of agriculture in these areas, but rather the predominance
+of tree crops, which we did not map.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{figures/figure7}
+{\centering \includegraphics[width=0.9\linewidth]{figures/figure6}
 
@@
 
-Both the per-pixel and vectorized maps, when combined with the map
-reference sample, enable separate estimates of the total extent of
-croplands in Ghana. The cropland extent estimated from the vectorized
-map is 42,359 km\(^2\) (with a margin of error of 4,395 km\(^2\)), or
-17.1 (15.4-18.9\%) of the mapped area. The estimate based on the per
-pixel map is 43,233 km\(^2\) (margin of error = 4,904 km\(^2\)), or 17.6
-(15.6-19.6\%) of area.
-
-The vectorized map provides additional information on how the
-characteristics of croplands can vary geographically, ranging from
-narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s
-lower right inset) to more densely packed, less distinctly shaped fields
-in AOI 5 (upper right inset in Figure \ref{fig:mainmap}). To explore how
-field characteristics varied geographically, we mapped the average field
-size and total number of fields within each 0.05 degree tile (Figure
-S13). These patterns generally correspond to those seen in the cropland
+Using the map reference sample and each map, we made two separate
+estimates of the total cropland area in Ghana in 2018. The cropland
+extent estimated from the field boundary map was 42,359 km\(^2\) (with a
+margin of error of 4,395 km\(^2\)), or 17.1\% (15.4-18.9\%) of the
+mapped area. The estimate from the per pixel map was 43,233 km\(^2\)
+(margin of error = 4,904 km\(^2\)), or 17.6\% (15.6-19.6\%) of area.
+
+The field boundary map provides additional information on how the
+characteristics of croplands vary across Ghana, ranging from narrow,
+strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower
+right inset) to more densely packed, less distinctly shaped fields in
+AOI 5 (upper right inset in Figure \ref{fig:mainmap}). To explore how
+field characteristics varied geographically, we mapped the average size
+and total number of fields within each 0.05 degree tile grid (Figure
+S15). These patterns generally correspond to those seen in the cropland
 density map (Figure \ref{fig:mainmap}), with larger sizes and field
-counts generally occurring in areas of higher field density, although
-the biases inherent in both measures (Figures S11-12) complicate the
-interpretation of those variations. However, we can use the estimated
-biases to develop adjusted estimates of field sizes and counts for each
-AOI, and for Ghana overall (Table \ref{tab:sizentab}). These adjusted
-estimates show that the typical field size in Ghana is 1.73 ha, ranging
-from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the forest zone
-AOIs (10, 11, 13, 14, 16) generally smaller than those in the northern
-half of the country (Table \ref{tab:sizentab}). The total number of
-fields was estimated to be 1,662,281 overall, or 205 fields per tile on
-average, ranging from 108/tile in AOI 4 to 399/tile in AOI 6.
+counts occurring where field densities are higher, although the biases
+(relative to the validation labels) in both measures (Figures S13-14)
+complicate interpretations of those variations. To minimize this
+complication, we used the calculated biases to develop adjusted
+estimates of field size and count (Table \ref{tab:sizentab}). These
+adjusted estimates show that the typical field size in Ghana is 1.73 ha,
+ranging from 0.96 in AOI 4 to 2.82 ha in AOI 4, with fields in the
+forest zone AOIs (10, 11, 13, 14, 16) generally smaller than those in
+the northern half of the country (Table \ref{tab:sizentab}). The
+estimated total number of fields is 1,662,281, or 205 fields per tile
+grid cells, varying from 108 fields/tile cell in AOI 4 to 399 in AOI 6.
 
@@
 
-These results demonstrate a capability for mapping the characteristics
-of smallholder-dominated cropping systems at high spatial resolution,
+These results demonstrate a capability to map the characteristics of
+smallholder-dominated cropping systems at high spatial resolution,
 annual time steps, and national scales. The resulting maps provide an
@@
 croplands in Ghana, complementing existing national to regional land
-cover maps derived from moderate resolution sensors (Hackman et al.
-2017, Xiong et al. 2017, ESA n.d.). This prior work found that cropland
-covered 19.4 (Xiong et al. 2017) to 32\% (Hackman et al. 2017) of Ghana
-in 2015, whereas our 2018 maps have cropland cover of 16.1-23.2\% (Table
-\ref{tab:mapaccuracy}), and our map reference sample-based estimates
-finds 17.1-17.6\% cover. Our results thus suggest that Ghana's cropland
-is less than previously estimated, but the difference is perhaps
-attributable to our use of a cropland definition that excluded longer
-fallows and abandoned fields, which in some regions can account for over
-half of the total area that could be counted as cropland (Tong et al.
-2020).
-
-In addition to the more detailed update of cropland extent, our maps
-also provide new information on the size and number of fields in Ghana
-(Figures \ref{fig:mainmap}, S11-12). Previous work to estimate such
-agricultural characteristics have often focused on farm, rather than
-field, size using census data (Von Braun 2004, Samberg et al. 2016,
-Jayne et al. 2016, Lowder et al. 2016). Efforts to map field boundaries
-in smallholder-dominated agricultural systems have either used \emph{in
+cover maps derived from moderate resolution imagery (Hackman et al.
+2017, Xiong et al. 2017, ESA n.d.). Those prior studies estimated that
+cropland covers 19.4 (Xiong et al. 2017) to 32\% (Hackman et al. 2017)
+of Ghana in 2015. In contrast, our 2018 maps provide a raw estimate of
+16.1-23.2\% cover (Table \ref{tab:mapaccuracy}), and our map reference
+sample-based estimate was 17.1-17.6\%. Our results thus suggest that
+Ghana's croplands are less extensive than those previous estimates.
+However, this difference may arise from our use of a cropland definition
+that excludes longer fallows and abandoned fields, which in some regions
+may comprise over half of total cropland area (Tong et al. 2020).
+
+In addition to this updated information on Ghana's cropland extent and
+distribution, our results provide new insights into field size and
+number at a national scale (Figures \ref{fig:mainmap}, S11-12). Previous
+efforts to map smallholder field boundaries have either used \emph{in
 situ} data collection (Carletto et al. 2013, 2015) or remote sensing
@@
 2019) or discontiguous (Estes et al. 2016a) areas. The most extensive
-studies to date used crowdsourced volunteers to classify fields into
-broad size classes, based on their interpretations of imagery sampled
-from high resolution virtual globes (Fritz et al. 2015, Lesiv et al.
-2019). Those efforts included country-specific results for Ghana (n =
-263), which can be converted into an average field size estimate of 5.33
-ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided the largest size class.
+studies to date enlisted crowdsourced volunteers to classify fields
+visible within high resolution imagery sampled from virtual globes into
+broad size categories (Fritz et al. 2015, Lesiv et al. 2019). Those
+efforts included country-specific results for Ghana (n = 263), which
+yield an average field size estimate of 5.33
+ha\footnote{Obtained by calculating the weighted mean from the count of the five size classes and the mean of the hectare range provided for the four smallest size classes, and the lower bound of the size range provided for the largest size class.
 Data sourced from Table S3 in Lesiv et al.
@@
 AOIs 1-9, 12, and 15, which is where most of the crowdsourced sample
-appears to have been collected. However, our bias-corrected estimates of
+appears to have been collected. However, our bias-adjusted estimates of
 1.73 (Ghana-wide) and 1.87 (AOIs 1-9, 12, and 15) ha were much smaller.
@@
 
-Although the maps generated by our system provide valuable new
-information, they nevertheless contain substantial errors. The overall
-map accuracies (86.7-88\%, Table \ref{tab:mapaccuracy}) are near the
-boundary of what might be considered \emph{achievable} map accuracy
-(Elmes et al. 2020), given the inherent uncertainty in the map reference
-sample, our best estimate of the ``truth,'' in which we have roughly
-85\% confidence. However, accuracies for the cropland class were much
-lower, falling between 62 (Producer's) to 67 (User's) percent
-country-wide for the per-pixel map (Table \ref{tab:mapaccuracy}),
-meaning the model produced substantial commission and omission errors
-for this class. The segmented boundary maps had fewer omission errors
-(Producer's accuracy = 79\%), but higher false positive errors (User's
-accuracy = 58.2\%). These accuracies are near the middle to upper ranges
-of those reported for the cropland class in other large-area mapping
-studies (Hackman et al. 2017, Xiong et al. 2017, Lesiv et al. 2017).
-
-The patterns of cropland-class accuracies varied by zone. These zones
-largely align, albeit with some discrepancies, with the country's
-agroecozones, thus the accuracy patterns may be partially attributed to
-some regions being harder to map than others. Producer's accuracies for
-both maps were highest in the two northern zones (1 and 2), which are
-primarily savannas (Figure S10), and lowest in zones 3 and 4, which are
-comprised of forest or coastal savannas. User's accuracies followed a
-similar pattern, with the exception of Zone 3, which had the highest
-User's accuracy, albeit from a very small sample. Aligning the reference
-samples more precisely with agroecozone boundaries (Figure S10B)
-provides further insight into error patterns (Table S4). Coastal
-savannas in the southeast had the highest Producer's accuracy but lowest
-User's accuracy for the per-pixel map, presumably because this region's
-numerous areas of high density cropland, combined with low woody cover
-in surrounding uncultivated areas, helped to promote commission error.
-Maps in the two northern savanna agroecozones had the best balance
+Although these maps provide valuable new information, they nevertheless
+contain substantial errors. The overall accuracies (86.7-88\%, Table
+\ref{tab:mapaccuracy}) are near the boundary of what might be considered
+\emph{achievable} map accuracy (Elmes et al. 2020), given that we only
+have \textasciitilde85\% confidence in our map reference sample, which
+is our best estimate of the ``truth.'' However, accuracies for the
+cropland class were much lower, falling between 62 (producer's) to 67
+(user's) percent country-wide for the per-pixel map (Table
+\ref{tab:mapaccuracy}), meaning the model produced substantial
+commission and omission errors for this class. The segmented boundary
+maps had fewer omission errors (producer's accuracy = 79\%), but higher
+false positives (user's accuracy = 58.2\%). These accuracies are near
+the middle to upper ranges of those reported for the cropland class in
+other large-area mapping studies (Hackman et al. 2017, Xiong et al.
+2017, Lesiv et al. 2017).
+
+The patterns of accuracies within the cropland class varied by zone.
+These zones largely align, albeit with some discrepancies, with the
+country's agro-ecological zones (AEZs), thus the accuracy patterns may
+be in part because some regions are simply more difficult to map.
+Producer's accuracy for both maps was highest in the two northern zones
+(1 and 2), which are primarily savannas (Figure S4), and lowest in zones
+3 and 4, which are comprised of forest or coastal savannas. User's
+accuracy followed a similar pattern, with the exception of Zone 3, which
+had the highest user's accuracy, albeit from a very small sample.
+Aligning the reference samples more precisely with agroecozone
+boundaries (Figure S4B) provides further insight into error patterns
+within the per-pixel map's cropland class (Table S4). Coastal savannas
+in the southeast had the highest producer's and lowest user's accuracy,
+perhaps because this region has high density cropland inter-mixed with
+uncultivated areas that have low woody cover, which could help promote
+commission error. Maps in the northern savannas had the best balance
 between omission and commission error, and had the highest overall
-User's accuracy. The transitional (between forest and savanna)
-agroecozone had a very low Producer's accuracy (21\%), which likely
-reflects the fact that it was divided between several AOIs for mapping
-(Figure S4), within which it typically covered a smaller share of area
-relative to the other agroecozones. This likely caused insufficient
-representation of this AEZ in training samples, particularly in AOIs 10
-and 11 (Figure S4B).
+user's accuracy. The transitional zone between forest and savanna had a
+very low Producer's accuracy (21\%), which likely reflects the fact that
+it was divided between several AOIs for mapping (Figure S4), and thus
+was under-represented in the training samples, particularly in AOIs 10
+and 11 (Figure S6B).
 
 Beyond the errors linked to regional differences, several other
-important factors contributed to reducing the accuracies in the cropland
-class. The first of these stems from the overall mapping extent and the
-high resolution of the data. Given the goal of developing a
-country-scale map at high resolution, the attendant data volume required
-us to use a relatively small set of image features and less than the
+important factors contributed to map error. The first of these related
+to the large mapping extent and image resolution. Given the goal of
+developing a high resolution, country-scale map, the large data volume
+constrained us to use a relatively small feature set and less than the
 recommended tree number and depth (Maxwell et al. 2018) in our Random
-Forests implementation, in order to limit computational costs. Previous
-work found that Random Forests achieves much better performance on
+Forests models, in order to limit computational costs. Previous work
+found that Random Forests achieves much better performance on
 small-scale croplands when trained on a much larger number of features
-(Debats et al. 2016). However, applying such a large feature set within
-the extent of our AOIs was intractable, as the computing time would have
-been several-fold larger than the \textasciitilde4-8 hours of runtime on
-800 CPUs required for a single active learning iteration, followed by
-\textasciitilde10-14 hours for prediction. This reduced the skill of the
-model, particularly when it came to differentiating cropland from
-adjacent bare patches or natural vegetation with sparse herbaceous
-cover, which were common in many AOIs.
+(Debats et al. 2016, Lebourgeois et al. 2017). However, applying such a
+large feature set within the extent of our AOIs was not possible, as the
+computing time and costs would have been several times
+larger\footnote{Each active learning iteration ran for $\sim$4-8 hours on 800 CPUs, followed by a final $\sim$10-14 hours for prediction}.
+This reduced the skill of the model, particularly when it came to
+differentiating cropland from bare or sparsely vegetated patches, which
+were common in many AOIs.
 
 The inherent difficulty of the labelling task was another major limiting
-factor. Our system was designed to minimize the error inherent in
-labelling, but determining croplands from non-croplands in these
-agricultural systems can be a difficult task. Labellers have to evaluate
-multiple image sources and to rely heavily on judgement, which
-inevitably leads to errors. Interpretation is particularly hard where
-the background savanna vegetation and croplands have similar reflectance
-during the dry season, which is a particular problem in AOIs 2 and 3.
-Smaller field sizes also complicate labelling, as these become
-increasingly indistinct in the \textasciitilde4 m PlanetScope
-composites. The difficulty of labelling is reflected in the magnitude of
-the Bayesian Risk metrics (Figure S6), and by the average score achieved
-by each labeller against our training reference dataset (71\%; Figure
-S5B). Although prior work (Rodriguez-Galiano et al. 2012, Mellor et al.
-2015) found that Random Forests are robust to label errors, we found
-that they have substantial impact (Figure \ref{fig:trainingimpact}),
-which suggest that simply improving label quality may be one of the
-single most important investments towards improved model accuracy.
+factor. Our platform was designed to minimize label errors, but
+determining croplands from non-croplands in these agricultural systems
+can be difficult. Labellers had to evaluate multiple image sources and
+to rely heavily on their judgment, which inevitably led to errors.
+Interpretation was particularly hard where croplands and surrounding
+landscapes had similar dry season reflectances, which was a particular
+problem in the northernmost savannas. Smaller field sizes also
+complicated labelling, as these become increasingly indistinct in the
+\textasciitilde4 m PlanetScope composites. The difficulty of labelling
+is reflected in the magnitude of the Bayesian Risk metrics (Figures
+S10-11), and by the average assignment quality scores of each labeller
+(71\%; Figure S9). Although prior work (Rodriguez-Galiano et al. 2012,
+Mellor et al. 2015) found that Random Forests are robust to label error,
+we found that it has substantial impact (Figure
+\ref{fig:trainingimpact}), which suggests that improving label quality
+is one of the most important factors in increasing model accuracy. Newer
+models, such as convolutional neural networks, may be less sensitive to
+label error, provided the error is random and the map reference samples
+are of high quality (Burke et al. 2021). However, over many smallholder
+systems training label errors will likely be biased in a particular
+direction (e.g.~towards omission when fields are not easily
+distinguished from the background), and our results show that reference
+labels can have substantial uncertainty.
 
 Image quality was another issue, although primarily in the forested
-AOIs, where frequent cloud cover reduced the number of available images
-in all seasons, resulting in composites with more brightness artifacts
-and blur (Figure \ref{fig:imqual}). This impacted labellers' abilities
-to discern fields, and doubtless affected model predictions. There is
-little to be done to mitigate these errors, short of confining imagery
-to the less cloudy dry season, which may further undermine model
-performance, given the importance of multi-temporal imagery for cropland
-classification (Debats et al. 2016, Defourny et al. 2019). Composite
-quality could be improved by using imagery from the same seasons over
-multiple years, but this would undermine the goal of developing annual
-maps, while the dynamism of the croplands would blur field boundaries
-within the imagery.
+AOIs, where frequent cloud cover and the corresponding lower number of
+available images resulted in lower quality composites (Figure
+\ref{fig:imqual}), with more brightness artifacts and blur. This
+impacted labellers' abilities to discern fields, and doubtless affected
+model predictions. Little can be done to mitigate these errors, short of
+confining imagery to the less cloudy dry season, which could reduce
+model performance by removing the temporal contrast (Debats et al. 2016,
+Defourny et al. 2019), or by adding radar data to the predictor set,
+which would reduce map resolution. Composite quality could be improved
+by using imagery from the same seasons over multiple years, but this
+would undermine the goal of developing annual maps, while the dynamism
+of the croplands would blur field boundaries within the imagery.
 
@@
 vectorized maps had high commission errors caused by uncertainties in
-the Random Forests predictions. Model uncertainty led to many pixels
-over non-crop areas with probabilities straddling the 0.5 classification
-threshold. Segments that intersected such areas were retained as fields
-when the average probability of intersecting pixels exceeded 0.5. A more
-accurate classifier would reduce such errors, or the application of a
-locally varying classification threshold (e.g. Waldner and Diakogiannis
-2020). Over-merging was another source of error in the segmentation
-algorithm, which in some areas led to overestimated field sizes and
-unrealistic shapes, particularly in high density croplands (e.g.~in AOIs
-2 and 8; Figure \ref{fig:mainmap}) where the boundaries between adjacent
-fields are often indistinct in the PlanetScope imagery. Minimizing or
-preventing merging would help in such cases, although could would result
-in the opposite problem, over-segmentation, and thereby underestimate
-field size.
+the Random Forests predictions. Model uncertainty meant that many pixels
+in non-cropland areas had probabilities with values near 0.5. Segments
+in these areas were retained if the average probability of intersected
+pixels exceeded the 0.5 classification threshold. A more accurate
+classifier would reduce such errors, as would a locally varying
+classification threshold (e.g. Waldner and Diakogiannis 2020).
+Over-merging was another source of error in the segmentation algorithm,
+which led to overestimated field sizes and unrealistic shapes in some
+areas, particularly in high density croplands (e.g.~in AOIs 2 and 8;
+Figure \ref{fig:mainmap}) where boundaries between adjacent fields were
+indistinct in the imagery. Preventing merging could help in such cases,
+but potentially lead to over-segmentation, thereby underestimating field
+sizes.
 
@@
 
-Despite the error sources mentioned above, several features of the
-system proved effective in mitigating error, leading to a higher overall
-accuracy than would have otherwise been possible. Label accuracy
-assessment and consensus labelling appeared to be the most effective
-error mitigation tools. Label accuracy measures allowed us to quantify
-the substantial impact of label error on model performance (Figure
-\ref{fig:trainingimpact}), while consensus labels substantially reduced
-individual labelling errors, resulting in maps that were more accurate
-than they would have been had we used individually generated labels.
-Labeller-specific accuracy measures also helped to improve the quality
-of the consensus labels, by placing higher weight on labels more likely
-to be accurate during the merging process, rather than giving equal
-weight to potentially less accurate labels. The ability to select the
-most accurate individual labels for a site also allowed us to develop
-independent estimates of field size to which measures of confidence can
-be attached (Figure S5B), which we were in turn able to correct
-estimates of field sizes and numbers (Table \ref{tab:sizentab}).
-
-The active learning approach helped to improve overall model performance
-relative to randomized training site selection, in line with findings
-from two recent efforts (Debats et al. 2017, Hamrouni et al. 2021).
-Although the performance gains relative to randomized model training
-that we observed were smaller (e.g. Debats et al. (2017) 29\% higher
-model performance after one iteration, and 8\% higher on the final
-iterations), those comparisons were made from lower initial bases, with
-initial training samples that were less than 1/10 the size, in terms of
-pixels, of our initial training sample. Our large initial randomly
-selected sample (500 grid cells) meant that our models were
-substantially trained before they were exposed to actively selected
-labels, thereby diluting their impact on performance. Nevertheless, the
-higher average performance of the active approach across three
-performance metrics demonstrated its effectiveness. Most notable were
-the larger improvements seen in the F1 score (Figure S8), a balanced
-performance metric. Gains in accuracy and AUC were smaller. For
-accuracy, the reduced advantage was primarily due to active learning
-being outperformed by randomized training after the first iteration in
-AOI 15, which proved one of the hardest AOIs to both map and label.
-Active learning likely resulted in the selection of sites that were
-harder to label than randomly selected ones, leading to more label
-error, and thus initially lower model accuracy. However, this deficit
-was overcome by the 5th iteration. The plateau in AUC gains at 0.5\%
-better than randomized training reflects the findings that active
-learning reduced both the false and true positive rates, the two inputs
-to AUC. Although the decline of the false positive rate (30.7\% between
-Iterations 0 and 3) was nearly three times larger than that of the true
-positive rate (10.9\%), AUC should be quite sensitive to the reduction
-in the latter, as it assesses how the tradeoff between the two rates
-varies across a full range of possible classification thresholds
-(Pontius and Si 2014).
-
-The detail, temporal precision, and large extent of these maps was
-enabled by the system's ability to process PlanetScope data, which is
-currently the only source of sub-5 meter imagery with daily coverage
-(McCabe et al. 2017). Daily revisits are important for creating seasonal
-composites within a single year over cloudy areas. The compositing
-technique we developed allowed us to develop a complete image catalog
-for the country representing the two seasons for 2018 agricultural year.
-Although Sentinel-2 is free, has better radiometric quality, and has
-sufficiently high resolution (10 m) to accurately classify small-scale
+Despite these numerous source of errors, our approach was effective in
+mitigating several of these error sources. Label quality assessment and
+consensus labelling were the most effective error mitigation tools.
+Label quality scores allowed us to quantify the impact of label error on
+model performance (Figure \ref{fig:trainingimpact}), while consensus
+labels produced maps that were more accurate than they would have been
+if we had relied on individually generated labels. The quality scores
+also helped to improve the overall accuracy of consensus labels, by
+placing higher weight on the work of more accurate labellers. In
+addition to these benefits, label quality scores (Figure S9) also
+allowed us to select the labels most likely to accurately capture field
+sizes and numbers, which we used to estimated and correct the biases in
+these two measures derived from the segmented field boundaries.
+
+Active learning improved overall model performance relative to
+randomized training site selection, in line with findings from two
+recent efforts (Debats et al. 2017, Hamrouni et al. 2021). Although the
+relative performance gains that we observed were smaller (e.g. Debats et
+al. (2017) 29\% higher model performance after one iteration, and 8\%
+higher on the final iterations), those comparisons were made by starting
+with a training sample that was \textless1/10 the size of ours. Our
+large starter sample meant that the models were substantially trained
+before they were exposed to actively selected labels, thereby diluting
+their impact on performance. Nevertheless, we found higher performance
+from active learning, most notably in the F1 score (Figure S8), a
+balanced performance metric, which further demonstrates its
+effectiveness.
+
+The detail, temporal precision, and large extent of our maps was enabled
+by the ability to process PlanetScope data, which is currently the only
+source of sub-5 meter imagery with daily coverage (McCabe et al. 2017).
+Daily revisits are important for creating seasonal composites within a
+single year over cloudy areas. The compositing technique we developed
+allowed us to develop a complete image catalog for the country
+representing the two seasons for 2018 agricultural year. Although
+Sentinel-2 is free, has better radiometric quality, and has sufficient
+spatial resolution (10 m) to accurately classify small-scale
 agricultural systems (e.g. Defourny et al. 2019, Kerner et al. 2020),
-its 5-day interval may be too infrequent to generate sufficiently
-cloud-free composites during the growing season over southern Ghana.
-Sentinel-1 is not affected by the same problem, but labelling fields in
-more coarsely resolved radar images is challenging.
+its 5-day interval may be too infrequent to generate cloud-free
+composites during the growing season over southern Ghana. Sentinel-1 is
+not affected by the same problem, but labeling fields in radar images is
+more challenging.
 
@@
 Several potential issues not addressed in our assessment merit further
-exploration. One of these is the degree of correspondence between image-
-and ground-collected labels. However, such comparisons may reveal
-unresolvable differences between the two perspectives. The highly
-dynamic nature of many agricultural systems means that relatively narrow
-differences between the dates of ground- and image-based digitizing
-campaigns can lead to substantial disagreement between the resulting
-field boundaries, simply because the fields themselves may have shifted
-during the interval (Elmes et al. 2020). These discrepancies could be
-further exacerbated by differences in the definition of what constitutes
-a field, which might vary on the ground depending on who is being asked,
-or who is doing the collecting. These factors suggest that ground versus
-image label differences would not necessarily indicate how far
-image-based labellers were from the ``truth.'' Nevertheless, a
-comparison against ground data would help to assess how accurately
-image-collected labels capture the typical size of fields, and thus
-merits further investigation for this reason.
+exploration. One of these was the degree of correspondence between
+image- and ground-collected labels. However, such comparisons may reveal
+unresolvable discrepancies between the two perspectives. The highly
+dynamic nature of these agricultural systems means that relatively
+narrow differences between the dates of ground- and image-based label
+collection can lead to substantial disagreement, simply because the
+fields themselves may have shifted during the interval (Elmes et al.
+2020). These discrepancies can be exacerbated by the definition used to
+determine what constitutes a field, which might vary on the ground
+depending on who is being asked, or who is doing the collecting. These
+factors suggest that difference between ground- and image-collected
+labels would not necessarily indicate how far image labellers were from
+the ``truth.'' Nevertheless, a comparison against ground data would help
+to assess how accurately image-collected labels capture the typical size
+of fields, and thus merits further investigation.
 
@@
 
-Another potential issue is the degree to which our assessment of the
-impact of label error on model performance (Figure
-\ref{fig:trainingimpact}) was influenced by the validation labels we
-used, which were generated using the consensus method. This could have
-confounded the assessment, particularly when comparing models trained
-with the most accurate individual label and those trained with consensus
-labels. However, the visual assessment of their resulting probability
-maps confirm the differences in scores: consensus and most accurate
-individual labels produce nearly identical maps with relatively high
-certainty, while low quality labels led to a markedly less certain map
-(Figure S9).
-
-\hypertarget{next-steps}{%
-\subsection{Next steps}\label{next-steps}}
-
-The maps presented here represent a version 1 product that is freely
-available to use, along with its underlying code (see SI for details).
-These data were developed according to the recommended best practices
-for training and assessing error in machine learning models (Elmes et
-al. 2020). In their current form, the maps may be useful for a variety
-of research applications. For example, analyzing the distributions of
-values in the probability maps may provide additional insight into the
-relative extents of active versus fallow croplands (Tong et al. 2020).
-However, use of these data, particularly for decision-making processes
-(e.g.~cropped area estimates), should be careful to account for the
-reported errors (Olofsson et al. 2014, Stehman and Foody 2019).
-
-To facilitate the next step, generating more accurate version 2 maps,
-several improvements will be made. The first is to replace Random
-Forests with a more advanced convolutional neural network (CNN), which
-can generate and learn from a large number of features representing a
-variety of spatial scales (Ma et al. 2019). Recent work suggests that a
-common architecture such as U-Net, when trained to distinguish field
-edges from interiors and combined with a post-hoc segmentation routine,
-is effective in delineating field boundaries (Waldner and Diakogiannis
-2020). Our system can readily incorporate such a model. The labelling
-platform already provides the methods needed to develop and assess the
-quality of labels that include field edges and interior classes, while
-active learning has proven to be effective for optimizing training
-datasets for deep learning models (Liu et al. 2017, Cao et al. 2020).
-Our current framework can be adjusted so that it starts by training a
-CNN from scratch with a large initial random sample, and then uses a
-transfer learning approach (Pan and Yang 2010) to update the model with
-the most informative samples from different AOIs or agroecozones.
-
-\hypertarget{conclusion}{%
-\subsection{Conclusion}\label{conclusion}}
+Another potential issue is the degree to which our assessment of label
+error on model performance (Figure \ref{fig:trainingimpact}) was
+influenced by the validation dataset we used, which was based on
+consensus labels. This could have confounded the analysis, particularly
+when comparing the consensus label-trained models with those trained
+with the most accurate individual labels. However, a visual assessment
+of the resulting probability maps confirms that models trained with the
+consensus and most accurate individual labels were more precise than the
+model trained with lower quality labels (Figure S12).
+
+\hypertarget{broader-applications}{%
+\subsection{Broader applications}\label{broader-applications}}
 
@@
 resolution, annual maps of smallholder-dominated croplands at national
-to regional scales, using a framework that can be readily updated to
-improve map accuracy as technologies improve. Maps that include
-information on field boundaries can help improve remote estimation of
-crop planted area and yield, and provide deeper insights into important
-socioeconomic aspects of agricultural systems, such as the relationships
-between agricultural productivity and livelihoods. Such maps will be
-important for developing an understanding of the rapid agricultural
-change that is currently unfolding throughout much of the continent.
+to regional scales, using an approach that follows recommended best
+practices for training and assessing machine learning models (Elmes et
+al. 2020). This approach can be readily updated to integrate
+improvements, such as newer machine learning models. Beyond providing
+valuable insights into field characteristics, field boundary maps can
+help improve remote estimation of crop areas and yield (e.g. Estes et
+al. 2013), and provide deeper insights into important socioeconomic
+aspects of agricultural systems, such as the relationships between
+agricultural productivity and farm size (Feder 1985, Carletto et al.
+2013, Desiere and Jolliffe 2018). Such maps will be important for
+understanding the rapid agricultural change that is currently occurring
+in Africa.
+
+\hypertarget{data-availability-and-usage}{%
+\subsection{Data availability and
+usage}\label{data-availability-and-usage}}
+
+The maps presented here are a version 1 product that is freely available
+to use, along with its underlying code (see SI). In their current form,
+they may be useful for a variety of research applications. For example,
+analyzing the distributions of values in the probability maps may
+provide additional insight into the relative extents of active versus
+fallow croplands (Tong et al. 2020). However, use of these data,
+particularly for decision-making processes (e.g.~cropped area
+estimates), should be careful to account for the reported errors
+(Olofsson et al. 2014, Stehman and Foody 2019).
 
@@
 
-\leavevmode\hypertarget{ref-caoHyperspectralImageClassification2020}{}%
-Cao, X., J. Yao, Z. Xu, and D. Meng. 2020. Hyperspectral {Image
-Classification With Convolutional Neural Network} and {Active Learning}.
-IEEE Transactions on Geoscience and Remote Sensing 58:4604--4616.
-
 \leavevmode\hypertarget{ref-CarlettoGuesstimatesGPStimatesLand2015}{}%
@@
 its impact on downstream analyses. Global Change Biology 24:322--337.
+
+\leavevmode\hypertarget{ref-EstesProjectedclimateimpacts2013}{}%
+Estes, L. D., H. Beukes, B. A. Bradley, S. R. Debats, M. Oppenheimer, A.
+C. Ruane, R. Schulze, and M. Tadross. 2013. Projected climate impacts to
+{South African} maize and wheat production in 2055: A comparison of
+empirical and mechanistic modeling approaches. Global Change Biology
+19:3762--3774.
 
@@
 
-\leavevmode\hypertarget{ref-jayneAfricaChangingFarm2016}{}%
-Jayne, T. s., J. Chamberlin, L. Traub, N. Sitko, M. Muyanga, F. K.
-Yeboah, W. Anseeuw, A. Chapoto, A. Wineman, C. Nkonde, and R. Kachule.
-2016. Africa's changing farm size distribution patterns: The rise of
-medium-scale farms. Agricultural Economics 47:197--214.
-
 \leavevmode\hypertarget{ref-KansangaTraditionalagriculturetransition2019}{}%
@@
 in {Data Sparse Regions}. arXiv:2006.16866 {[}cs, eess{]}.
+
+\leavevmode\hypertarget{ref-lebourgeoisCombinedRandomForest2017}{}%
+Lebourgeois, V., S. Dupuy, É. Vintrou, M. Ameline, S. Butler, and A.
+Bégué. 2017. A {Combined Random Forest} and {OBIA Classification Scheme}
+for {Mapping Smallholder Agriculture} at {Different Nomenclature Levels
+Using Multisource Data} ({Simulated Sentinel}-2 {Time Series}, {VHRS}
+and {DEM}). Remote Sensing 9:259.
 
@@
 
-\leavevmode\hypertarget{ref-liuActiveDeepLearning2017}{}%
-Liu, P., H. Zhang, and K. B. Eom. 2017. Active {Deep Learning} for
-{Classification} of {Hyperspectral Images}. IEEE Journal of Selected
-Topics in Applied Earth Observations and Remote Sensing 10:712--724.
-
 \leavevmode\hypertarget{ref-lobellCropYieldGaps2009}{}%
@@
 Environment and Resources 34:179--204.
-
-\leavevmode\hypertarget{ref-LowderNumberSizeDistribution2016}{}%
-Lowder, S. K., J. Skoet, and T. Raney. 2016. The {Number}, {Size}, and
-{Distribution} of {Farms}, {Smallholder Farms}, and {Family Farms
-Worldwide}. World Development 87:16--29.
 
@@
 
-\leavevmode\hypertarget{ref-panSurveyTransferLearning2010}{}%
-Pan, S. J., and Q. Yang. 2010. A {Survey} on {Transfer Learning}. IEEE
-Transactions on Knowledge and Data Engineering 22:1345--1359.
-
 \leavevmode\hypertarget{ref-perselloDelineationAgriculturalFields2019}{}%
@@
 
+\leavevmode\hypertarget{ref-royGlobalAnalysisTemporal2021}{}%
+Roy, D. P., H. Huang, R. Houborg, and V. S. Martins. 2021. A global
+analysis of the temporal availability of {PlanetScope} high spatial
+resolution multi-spectral imagery. Remote Sensing of Environment
+264:112586.
+
 \leavevmode\hypertarget{ref-RulliFoodappropriationlarge2014}{}%
@@
 elimination of points. The Cartographic Journal 30:46--51.
-
-\leavevmode\hypertarget{ref-VonBraunSmallscalefarmersliberalized2004}{}%
-Von Braun, J. 2004. Small-scale farmers in a liberalized trade
-environment. Page 21.
 

### 2021-07-23 inst/manuscript/si.tex
Added dataset documentation, manuscript statement (EarthArxiv), several datasets

@@
 
-\hypertarget{assessing-quality-of-composites}{%
-\subsection{Assessing quality of
-composites}\label{assessing-quality-of-composites}}
+\hypertarget{map-reference-system}{%
+\subsection{Map reference system}\label{map-reference-system}}
+
+The map reference system used in our mapping approach had three
+different levels. At the coarsest level, Ghana was divided into 16
+different mapping zones, or areas of interest (AOIS; Figure
+\ref{fig:aois}A), that were used to create AOI-specific mapping models.
+Each AOIs was comprised of 400 to 777 adjacent tiles. These tiles were
+used to create seasonal image composites, and were defined within a 0.05
+degree grid (Figure \ref{fig:aois}B)), with each tile numbered to
+correspond to a larger 1 degree grid cell that it sits within (dotted
+lines in Figure \ref{fig:aois}B). Training and reference labels were
+created within a 0.005 degree grid that nested within each tile (Figure
+\ref{fig:aois}C). Therefore, each tile has 100 grid cells, and there are
+400 tiles per 1X1 degree. The smallest AOI consists of a single 1X1
+degree, which fall in the center of the country (AOIs 5, 8, 11, and 14).
+AOIs falling along Ghana's boundaries were created by tiles from 1X1
+degree cells that straddled Ghana's border with those from the closest
+degree cells that were fully contained within Ghana (e.g.~AOI 1), with
+the exception of AOI 16, which was comprised of tiles in three partial
+1X1 degree cells along Ghana's coast.
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=1\linewidth,]{figures/reference_system}
+
+}
+
+\caption{The reference system used in this mapping approach, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.}\label{fig:aois}
+\end{figure}
+
+\hypertarget{image-compositing}{%
+\subsection{Image compositing}\label{image-compositing}}
+
+\hypertarget{variation-in-lengths-of-temporal-window-used}{%
+\subsubsection{Variation in lengths of temporal window
+used}\label{variation-in-lengths-of-temporal-window-used}}
+
+The typical window for compositing dry season imagery was December, 2018
+to February, 2019, but in the cloudiest regions (AOIs 10, 11, 13, 14,
+16) we started the dry season window in November, to ensure a sufficient
+density of images for compositing.
+
+\hypertarget{assessing-composite-quality}{%
+\subsubsection{Assessing composite
+quality}\label{assessing-composite-quality}}
 
@@
 
-Two supervisors (Ye and Estes) who were not involved in labelling
-developed the map reference sample evaluated the PlanetScope composites
-at each site to determine its class membership. The following set of
-rules were followed in collecting the sample:
+Two supervisors (Ye and Estes), who were not involved in collecting
+training and validation labels, collected the map reference sample by
+evaluating the PlanetScope composites at each site to determine its
+class membership. The following set of rules were followed in collecting
+the sample:
 
@@
 \item
-  After determining the class, if
+  After determining the class, if:
 \end{enumerate}
@@
 agroecozones\footnote{sourced from https://wheregeospatial.com/agro-ecological-zones-ghana/}
+
+\hypertarget{map-reference-label-uncertainty}{%
+\paragraph{Map reference label
+uncertainty}\label{map-reference-label-uncertainty}}
+
+The size of the collected validation sample was 1207, with 1036 samples
+collected and interpreted by one observer (Su Ye) and 171 collected by a
+second observer (Estes). To evaluate the uncertainty inherent in
+defining the map reference labels, the pair mapped 23 common sites,
+showing an overall level of agreement of 87\%, and a Spearman Rank
+Correlation of 0.76.
+
+Although this overlap between observers was limited, the map reference
+classification scheme provided two additional measures of uncertainty,
+which were classes defined as ``unsure but most likely a field'' or
+``unsure but most likely not a field,'' which the reference labeller
+would choose when they could not with high confidence state whether the
+site was either cropland or not cropland. Of the 487 map reference
+samples that were identified as cropland, 22\% fell into the lower
+confidence category, as did 11.5\% of the 720 non-cropland samples.
+Across both classes, the reference labellers had lower confidence in
+15.7 of sites labelled.
+
+\hypertarget{accuracy-assessment-zones}{%
+\paragraph{Accuracy assessment zones}\label{accuracy-assessment-zones}}
+
+The zones used to assess regional variation in map accuracy are
+illustrated in Figure \ref{fig:aoizones}.
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.9\linewidth,]{figures/si_aoi_zones}
+
+}
+
+\caption{Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy.
+In B, the ID number of Ghana's individual agroecozones is shown in grey: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.}\label{fig:aoizones}
+\end{figure}
 
@@
 
-\hypertarget{active-learning-process}{%
-\subsection{Active learning process}\label{active-learning-process}}
+\hypertarget{mapping-cropland-probabilities-with-active-learning}{%
+\subsection{Mapping cropland probabilities with active
+learning}\label{mapping-cropland-probabilities-with-active-learning}}
+
+The number of active learning iterations per AOI was three, for all but
+four AOIs. AOIs 10 and 14 stopped after one and two iterations,
+respectively, as they started with high initial validation accuracies
+(\textgreater83\%) and showed little subsequent improvement. The models
+for these two AOIs were thus trained with 600 - 700 labels. AOI 15 was
+run for 4 iterations (900 samples), while AOI 3 underwent a second
+active learning cycle because the model produced during the first cycle
+was inaccurate (see next section). In this second run, 300 initial
+training sites randomly selected within the AOI were used, followed by 2
+subsequent active learning iterations, resulting in a training sample of
+500.
 
@@
 training sites confined to this AOI (blue points in Figure
-\ref{fig:trainval}A)
+\ref{fig:trainval}A).
 
@@
 
-The distribution of data collection effort and scores from training
-reference sites by labelling team are shown in Figure
-\ref{fig:trainqual}. These results include those from re-labelling the
-initial training in Cluster 2 (AOIs 7-9, 12, 15). This was done because
-we discovered a small spatial offset in the original image composites,
-which we corrected by reprocessing the images. We replaced the image
-overlays in the instance for central-southeast (Figure
-\ref{fig:trainval}A), and the labelling team mapped these sites a second
-time during the production run in late 2019. The reprocessed labels were
-used to initially train the model for AOIs 9, 12, and 15.
-
-\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.5\linewidth,]{figures/si_training_qual_n}
-
-}
-
-\caption{The A) distributions of scores at training reference sites for each labeller (means indicated by X in boxplots), and B) the number of training and training reference sites completed by each labeller.
-Labellers' identities are anonymized.}\label{fig:trainqual}
-\end{figure}
-
-The average Bayesian Risk of each training and validation site is shown
-in Figure \ref{fig:labelrisk}, as well as the distribution of risk
-values per AOI and the initial training clusters (Figure
-\ref{fig:labelriskhist}). The three initial clusters include the second
-mapping of Cluster 2.
-
-\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.8\linewidth,]{figures/si_label_risk_map}
-
-}
-
-\caption{The average Bayes Risk of each training and validation site in Ghana.}\label{fig:labelrisk}
-\end{figure}
-
-\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.4\linewidth,]{figures/si_label_risk_hists}
-
-}
-
-\caption{The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
-The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.}\label{fig:labelriskhist}
-\end{figure}
-
-\hypertarget{validation}{%
-\subsubsection{Validation}\label{validation}}
+The distribution of training and validation sample collection effort was
+divided across 20 labellers, with a core group of 13 who completed more
+than 1,000 assignment each (Figure \ref{fig:assignmentcount}). As each
+training/validation task was undertaken by 4 separate labellers, 34,014
+sets of labels were made. Each labeller digitized an average of 2,001
+training/validation assignments.
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.8\linewidth,]{figures/si_training_n}
+
+}
+
+\caption{The A) number of training/validation and accuracy assessment assignments completed by each labeller, and B) the distributions of quality scores at training reference sites for each labeller (means indicated by X in boxplots).
+Labellers' identities are anonymized.}\label{fig:assignmentcount}
+\end{figure}
+
+These results include those from re-labelling the initial training in
+Cluster 2 (AOIs 7-9, 12, 15). This was done because we discovered a
+small spatial offset in the original image composites, which we
+corrected by reprocessing the images. We replaced the image overlays in
+the instances for this Cluster (Figure \ref{fig:trainval}A), and the
+labelling team mapped these sites a second time during the production
+run in late 2019. The reprocessed labels were used to initially train
+the model for AOIs 9, 12, and 15.
+
+\hypertarget{model-performance}{%
+\subsubsection{Model performance}\label{model-performance}}
 
@@
 differences at iteration 0, when both models were trained with the same
-level set.\\
+level set.
 
@@
 \end{figure}
+
+The lower accuracy of actively versus randomly trained models in earlier
+iterations was caused by results at AOI 15, where active learning
+accuracy was 8.37 percent lower than random training after iteration 1
+(see Figure \ref{fig:randomvactive}). In comparison, iteration 1 active
+learning accuracies were 2.88 and 0.45 percent higher than random
+training for AOIs 1 and 8, respectively. Accuracy under active learning
+for AOI 15 exceeded randomized training after 4 iterations.
 
@@
 error}\label{the-impact-of-training-data-error}}
+
+Two measures of label quality, average quality score per labeller and
+the Bayesian Risk of labels, were calculated to provide proxy measures
+of label error. Labeller quality was scored 9,389 times against 98
+unique training reference sites, with each labeller assessed an average
+of 552 times at a rate of 1 training reference site for every 3.62
+training site mapped. The mean of each labeller's average accuracy score
+was 0.71 (range 0.6 to 0.85; Figure \ref{fig:labelqual}).
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.7\linewidth,]{figures/si_label_quality}
+
+}
+
+\caption{Boxplots showing the distributions of quality scores from the accuracy assessment assignments undertaken by each labeller (means indicated by X in boxplots).
+Labellers' identities are anonymized.}\label{fig:labelqual}
+\end{figure}
+
+The average Bayesian Risk of each training and validation site is shown
+in Figure \ref{fig:labelrisk}, as well as the distribution of risk
+values per AOI and the initial training clusters (Figure
+\ref{fig:labelriskhist}). The three initial clusters include the second
+mapping of Cluster 2 (Figure \ref{fig:trainval}A).
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.8\linewidth,]{figures/si_label_risk_map}
+
+}
+
+\caption{The average Bayes Risk of each training and validation site in Ghana.}\label{fig:labelrisk}
+\end{figure}
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=0.4\linewidth,]{figures/si_label_risk_hists}
+
+}
+
+\caption{The distribution of Bayesian Risk values for training (left column) and validation (right column) sites in each AOI, with the average value indicated by vertical lines.
+The first three rows (panel titles beginning with 'Cl') indicate distributions of Bayes Risk in the initial training clusters, including the first ('Cl2a') and second ('Cl2b') mappings of Cluster 2.}\label{fig:labelriskhist}
+\end{figure}
+
+The average Bayesian risk was 0.122 for training labels and 0.127 for
+validation labels. Risk was highest in the northern AOIs (AOIs 1-6;
+Figures S6-7), falling between 0.157 for training and 0.173 for
+validation labels (Figures S6-7), and lowest in the southwestern AOIs
+(AOIs 10, 11, 13, 14, 16; training risk = 0.079; validation risk =
+0.065). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15)
+was slightly lower (training = 0.127; validation = 0.136) than in the
+north. Labeller experience also appeared to reduce risk, which we
+observed during a relabelling of the 500 initial random site in this
+cluster; the mean risk of the updated labels was 0.055, compared to
+0.172 for original labels.
 
@@
 mean score of each labeller against the training reference sites (Figure
-\ref{fig:trainqual}), as assessed when labelling a given AOI. These
+\ref{fig:labelqual}), as assessed when labelling a given AOI. These
 images were created for a single tile in AOI 1.
@@
 
-\hypertarget{reference-label-uncertainty}{%
-\subsubsection{Reference label
-uncertainty}\label{reference-label-uncertainty}}
-
-The size of the collected validation sample was 1207, with 1036 samples
-collected and interpreted by one observer (Su Ye) and 171 collected by a
-second observer (Estes). To evaluate the uncertainty inherent in
-defining the map reference labels, the pair mapped 23 common sites,
-showing an overall level of agreement of 87\%, and a Spearman Rank
-Correlation of 0.76.
-
-Although this overlap between observers was limited, the map reference
-classification scheme provided two additional measures of uncertainty,
-which were classes defined as ``unsure but most likely a field'' or
-``unsure but most likely not a field,'' which the reference labeller
-would choose when they could not with high confidence state whether the
-site was either cropland or not cropland. Of the 487 map reference
-samples that were identified as cropland, 22\% fell into the lower
-confidence category, as did 11.5\% of the 720 non-cropland samples.
-Across both classes, the reference labellers had lower confidence in
-15.7 of sites labelled.
-
 \hypertarget{categorical-accuracy}{%
@@
 \ref{tab:aezaccuracy}).
-
-\begin{figure}[!ht]
-
-{\centering \includegraphics[width=0.9\linewidth,]{figures/si_aoi_zones}
-
-}
-
-\caption{Four zones used to combine the A) AOIs and B) agroecozones (B), in order to assess sub-national mapping accuracy.
-In B, the ID number of Ghana's individual agroecozones is shown in grey: 1 = Coastal savanna; 2 = Wet evergreen; 3 = Moist evergreen; 4 = Deciduous forest; 5 = Transitional zone; 6 = Guinea savanna; 7 = Sudan savanna; 8 (not shown) = Volta Lake.}\label{fig:aoizones}
-\end{figure}
 

### 2021-07-23 R/data.R

@@ -1,0 +1,116 @@
+#' Model performance metrics per iteration
+#' @name accuracy_reports
+#' @docType data
+#' @description Results of accuracy assessment performed with map reference
+#' sample, including margins of error for overall, producer's, and user's
+#' accuracy, as well as class area estimates, with confidence intervals.
+#' @keywords data
+#' @examples
+#' data(accuracy_reports)
+#'
+NULL
+
+#' Model performance metrics per iteration
+#' @name consensus_high_low
+#' @docType data
+#' @description Performance scores from final iteration for 4 AOIs (1, 2, 8, 15)
+#' where label strategy experiments were run: models trained with consensus
+#' labels versus those trained with most and least accurate individual labels
+#' @keywords data
+#' @examples
+#' data(consensus_high_low)
+NULL
+
+#' Error matrices using map reference sample
+#' @name error_matrices
+#' @docType data
+#' @description Error matrices are calculated for the per pixel and segmented
+#' field boundary map, for particular zones consisted of grouped AOIs and AEZ,
+#' and for all of Ghana.
+#' @keywords data
+#' @examples
+#' data(error_matrices)
+NULL
+
+#' Model performance metrics per iteration
+#' @name image_quality
+#' @docType data
+#' @description Quality scores for the seasonal image composites
+#' @keywords data
+#' @examples
+#' data(image_quality)
+NULL
+
+#' Model performance metrics per iteration
+#' @name iteration_metrics
+#' @docType data
+#' @description Performance scores per iteration for each of the AOIs, including
+#' score changes measures between iterations
+#' @keywords data
+#' @examples
+#' data(iteration_metrics)
+NULL
+
+#' Bayesian Risk metrics per site with summary statistics
+#' @name label_risk
+#' @docType data
+#' @description Provides the mean Bayesian Risk and proportion of risky pixels
+#' for each site per AOI, and summarized per AOI.
+#' @keywords data
+#' @examples
+#' data(label_risk)
+NULL
+
+#' Summary Bayesian Risk metrics along with per AOI model performance metrics
+#' @name label_risk_metrics
+#' @docType data
+#' @description Provides the mean Bayesian Risk per AOI and scores for various
+#' Random Forests performance metrics assessed against validation sites.
+#' @keywords data
+#' @examples
+#' data(label_risk_metrics)
+NULL
+
[…]

### 2021-07-25 inst/manuscript/manuscript.Rmd
Edited cover note to reflect peer review status

@@
 footerdate: no
-note: This pre-print not yet undergone peer review.
+note: This pre-print has not yet completed peer review.
 It has been submitted to \emph{Frontiers in Artificial Intelligence}.

### 2021-10-07 inst/manuscript/figures.Rmd
Revisions for Frontiers in AI round 1 reviews

@@
 ### Figure 1: System architecture
-Created in keynote
+Created in Keynote
 
@@
 
-### Figure S3: Map reference sample
+### Figure S3: Segmentation workflow 
+Created in Keynote.
+
+### Figure S4: Map reference sample
 ```{r, eval = FALSE}
@@
 
-### Figure S4: AOI Zones
+### Figure S5: AOI Zones
 
@@
 # read in field segments
-flds_1 <- read_sf(segment_files$file[1])
-flds_5 <- read_sf(segment_files$file[5])
-flds_11 <- read_sf(segment_files$file[11])
-flds_15 <- read_sf(segment_files$file[15])
+flds_1 <- read_sf(segment_files$file[1]) %>% st_make_valid()
+flds_5 <- read_sf(segment_files$file[5]) %>% st_make_valid()
+flds_11 <- read_sf(segment_files$file[11]) %>% st_make_valid()
+flds_15 <- read_sf(segment_files$file[15]) %>% st_make_valid()
 
@@
 vpal <- viridis::viridis(20)
-inset_cols <- c("red", "orange", "pink", "brown")
+# inset_cols <- c("red", "orange", "pink", "brown")
+inset_cols <- rep("red", 4)
 fld_tiles <- list(flds_tile1, flds_tile2, flds_tile3, flds_tile4)
@@
 
+# create lines to point to insets
+tile_ss_xy <- st_coordinates(st_centroid(tile_ss))
+lns <- st_sfc(
+  list(
+    st_linestring(
+      rbind(tile_ss_xy[1, ], c(maprange[1] * 0.9, maprange[4] * 0.885))
+    ), 
+    st_linestring(
+      rbind(tile_ss_xy[2, ], c(maprange[2] * 0.9, maprange[4] * 0.8517))
+    ),
+    st_linestring(
+      rbind(tile_ss_xy[3, ], c(maprange[1] * 0.9, maprange[4] * 0.619))
+    ),
+    st_linestring(
+      rbind(
+        tile_ss_xy[4, ], c(maprange[2] * 0.9, maprange[4] * 0.5681))
+    )
+  ), crs = 4326
+)
+
 # cropland fractions
 p <- ggplot(africa) + 
-  geom_sf(fill = "gray40", color = "grey", size = 0.25) +  
+  geom_sf(fill = "gray60", color = "grey", size = 0.25) +  
   geom_stars(data = cropfrac) +
@@
   geom_sf(data = tile_ss, fill = inset_cols, col = inset_cols, size = 0.1) +
-  # geom_sf(data = tile_ss, fill = "red", col = "red") +
+  geom_sf(data = lns, col = inset_cols) + 
+  geom_sf(data = tile_ss, fill = "red", col = "red") +
   scale_fill_viridis_c(name = "%", na.value = "transparent") +
   coord_sf(xlim = maprange[1:2], ylim = maprange[3:4]) + 
+  ggsn::north(symbol = 16, scale = 1.2, x.min = 0.5, x.max = 1, 
+              y.min = 10.5, y.max = 11) +
+  ggsn::scalebar(ghana, dist = 100, dist_unit = "km", location = "bottomleft", 
+                 transform = TRUE, model = "WGS84", st.size = 3, st.dist = 0.01, 
+                 anchor = c(x = -0.7, y = 4.8), border.size = 0.1) + 
   theme_void() + 
-  theme(legend.position = c(0.64, 0.08), legend.direction = "horizontal", 
+  theme(legend.position = c(0.675, 0.1), legend.direction = "horizontal",
         legend.key.height = unit(0.3, "cm"),
+        legend.key.width = unit(0.67, "cm"),
         legend.text = element_text(colour = "black"),
-        legend.title = element_text(size = 10, colour = "black"))
+        legend.text.align = 0.5,
+        legend.title.align = 1,
+        legend.spacing.y = unit(0.05, "cm"),
+        legend.title = element_text(
+          vjust = 1, size = 11, colour = "black"
+        )
+  )
+
 # with insets added
@@
 
-### Figure S5: PlanetScope image density
+### Figure S6: PlanetScope image density
 The density of imagery available for download from the image catalog.
@@
 
-### Figure S6: Initial training site distribution
+### Figure S7: Initial training site distribution
 Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
@@
 
-### Figure S7: Assignment count
+### Figure S8: Assignment count
 ```{r assignmentcount, eval=FALSE}
@@
 
-### Figure S9: Label quality
+### Figure S10: Label quality
 ```{r labelquality, eval=FALSE}
@@
 
-### Figures S10 and S11: Consensus label (Bayesian) risk
+### Figures S11 and S12: Consensus label (Bayesian) risk
 
@@
 
-### Figure S12: Training data error impact on probability images
+### Figure S13: Training data error impact on probability images
 ```{r, eval = FALSE}
@@
 
-### Figures S13 and S14: Segmentation statistics (grid scale)
+### Figures S14 and S15: Segmentation statistics (grid scale)
 
@@
 
-### Figure S15: Average fields sizes
+### Figure S16: Average field sizes
 

### 2021-10-07 inst/manuscript/si.Rmd
Revisions for Frontiers in AI round 1 reviews

@@
 Using equation \ref{eq:pixelrisk}, the corresponding risk associated with this particular pixel's label is thus `r rj * (1 - 1) + (1 - rj) * 1` (i.e., 1 - `r rj <- round((wtA * 1 + wtB * 0) / (wtA + wtB), 3); rj`).
+
+### Segmentation
+An overview of the two inputs to and outputs from the segmentation algorithm are shown in Figure \ref{fig:segfig}.
+
+```{r segfig, echo = FALSE, fig.align='center', out.width="100%", message=FALSE, warning=FALSE, fig.cap="An overview of the inputs and key outputs from the five-step segmentation algorithm."}
+knitr::include_graphics('figures/si_segmentation_workflow.png')
+```
 

### 2021-10-07 inst/manuscript/si.tex
Revisions for Frontiers in AI round 1 reviews

@@
 
+\hypertarget{segmentation}{%
+\subsubsection{Segmentation}\label{segmentation}}
+
+An overview of the two inputs to and outputs from the segmentation
+algorithm are shown in Figure \ref{fig:segfig}.
+
+\begin{figure}[!ht]
+
+{\centering \includegraphics[width=1\linewidth,]{figures/si_segmentation_workflow}
+
+}
+
+\caption{An overview of the inputs and key outputs from the five-step segmentation algorithm.}\label{fig:segfig}
+\end{figure}
+
 \hypertarget{accuracy-assessment}{%
@@
 The average Bayesian Risk of each training and validation site is shown
-in Figure \ref{fig:labelrisk}, as well as the distribution of risk
-values per AOI and the initial training clusters (Figure
-\ref{fig:labelriskhist}). The three initial clusters include the second
-mapping of Cluster 2 (Figure \ref{fig:trainval}A).
+in Figure \ref{fig:labelrisk}, and the distribution of risk values per
+AOI and the initial training clusters in Figure \ref{fig:labelriskhist}.
+The three initial clusters include the second mapping of Cluster 2
+(Figure \ref{fig:trainval}A).
 
@@
 validation labels. Risk was highest in the northern AOIs (AOIs 1-6;
-Figures S6-7), falling between 0.157 for training and 0.173 for
-validation labels (Figures S6-7), and lowest in the southwestern AOIs
-(AOIs 10, 11, 13, 14, 16; training risk = 0.079; validation risk =
-0.065). Label risk in the central-southeastern AOIs (AOIs 7-9, 12, 15)
-was slightly lower (training = 0.127; validation = 0.136) than in the
-north. Labeller experience also appeared to reduce risk, which we
-observed during a relabelling of the 500 initial random site in this
-cluster; the mean risk of the updated labels was 0.055, compared to
-0.172 for original labels.
+Figures \ref{fig:labelrisk}-\ref{fig:labelriskhist}), falling between
+0.157 for training and 0.173 for validation labels, and lowest in the
+southwestern AOIs (AOIs 10, 11, 13, 14, 16; training risk = 0.079;
+validation risk = 0.065). Label risk in the central-southeastern AOIs
+(AOIs 7-9, 12, 15) was slightly lower (training = 0.127; validation =
+0.136) than in the north. Labeller experience also appeared to reduce
+risk, which we observed during a relabelling of the 500 initial random
+site in this cluster; the mean risk of the updated labels was 0.055,
+compared to 0.172 for original labels.
 
@@
   Source code for the segmentation algorithm. The repository
-  \footnote{github.com/agroimpacts/segmenter} is currently private, but
-  will be opened after completing a security audit.
+  \footnote{github.com/agroimpacts/segmenter} is open.
 \end{itemize}
@@
 \leavevmode\hypertarget{ref-hijmansRasterGeographicData2020}{}%
-Hijmans, R. J. 2020. Raster: {Geographic} data analysis and modeling.
+Hijmans, R. J. 2020. Raster: Geographic data analysis and modeling.
 Manual.
@@
 \leavevmode\hypertarget{ref-pebesmaSimpleFeaturesStandardized2018}{}%
-Pebesma, E. 2018. Simple features for r: {Standardized} support for
+Pebesma, E. 2018. Simple features for r: Standardized support for
 spatial vector data. The R Journal 10:439--446.

### 2021-10-07 R/consensus_high_low.R

@@ -1,57 +1,0 @@
-# Iteration metrics from models trained with consensus and highest and lowest
-# accuracy individual labels
-
-library(dplyr)
-library(aws.s3)
-library(glue)
-data("iteration_metrics")
-
-fnames <- unlist(lapply(paste0(c('low', 'high'), ".csv"), function(x) {
-  paste0(glue("planet/incoming_metrics_{c(1, 2, 8, 15)}_"), x)
-}))
-high_low_metrics <- lapply(fnames, function(x) {
-  dat <- s3read_using(FUN = readr::read_csv, object = x,
-                      bucket = "activemapper")
-  highlow <- gsub("planet/incoming_metrics_[0-9]+_|.csv", "", x)
-  dat %>% mutate(aoi = paste0(aoi, "_", highlow)) %>% rename(AOI = aoi) %>%
-    select(AOI, !!names(.))
-}) %>% do.call(rbind, .)
-
-# Process and reshape
-metrics <- c("Accuracy", "AUC", "F1", "TSS", "Precision", "Recall", "TPR",
-             "FPR")
-high_low_metrics <- high_low_metrics %>%
-  mutate(iteration = iteration - 1) %>% # rebase iteration counter on 0
-  mutate(F1 = 2 * (precision * recall) / (precision + recall)) %>%
-  rename(Iteration = iteration, Accuracy = accuracy, TSS = tss,
-         Precision = precision, Recall = recall, TPR = tpr,
-         FPR = fpr) %>%
-  select(AOI, Iteration, !!metrics) %>%
-  tidyr::pivot_longer(cols = Accuracy:FPR, names_to = "Metric",
-                      values_to = "Score")
-
-# save out for keeping
-# usethis::use_data(high_low_metrics, overwrite = TRUE)
-
-# bring in iteration metrics and join with consensus values from those
-consensus <- iteration_metrics %>%
-  filter(AOI %in% paste0("labeller", c(1, 2, 8, 15))) %>%
-  group_by(AOI) %>%
-  filter(Iteration == max(Iteration) & !grepl("change", Metric)) %>%
-  ungroup() %>%
-  mutate(Strategy = "Consensus") %>%
-  mutate(AOI = gsub("labeller", "", AOI))
-
-# join
-consensus_high_low <- high_low_metrics %>%
-  mutate(Strategy = tools::toTitleCase(gsub("[0-9]+_", "", AOI)),
-         AOI = gsub("_.*", "", AOI)) %>%
[…]

### 2021-10-08 inst/manuscript/manuscript.Rmd
MS version without EarthArxiv note

--- 2021-07-23 inst/manuscript/manuscript.Rmd
@@
 ---
-title: "High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"
+title: "High resolution, annual maps of field boundaries for  smallholder-dominated croplands at national scales"
 
@@
 geometry: margin=1in
-linenumbers: no
+linenumbers: TRUE
 spacing: doublespacing
 footerdate: no
-note: This pre-print not yet undergone peer review.
-It has been submitted to \emph{Frontiers in Artificial Intelligence}.
-This version will be updated as it is revised, and the final published version will be accessible through its DOI link.
-abstract: Mapping the changing characteristics of Africa's smallholder-dominated agricultural systems, including the sizes and numbers of fields, is crucial for understanding food security and a range of other socioeconomic and environmental concerns.
-However, accurately mapping these systems is difficult because of 1) the spatial and temporal mismatch between satellite sensors and smallholder fields, and 2) the lack of high-quality labels needed to train and assess machine learning classifiers.
-We developed an approach designed to address these two problems, which we used to map Ghana’s annual croplands for the year 2018.
-To overcome the first problem, we converted daily, high resolution CubeSat (PlanetScope) imagery into two cloud-free seasonal composites covering a single agricultural year.
-To address the second problem, we created a labelling platform that rigorously assesses and minimizes label error, and used it to iteratively train a Random Forests classifier with active learning, which identifies the most informative training sample based on prediction uncertainty.
+
+abstract: Mapping the characteristics of Africa's smallholder-dominated croplands, including the sizes and numbers of fields, can provide critical insights into food security and a range of other socioeconomic and environmental concerns.
+However, accurately mapping these systems is difficult because there is 1) a spatial and temporal mismatch between satellite sensors and smallholder fields, and 2) a lack of high-quality labels needed to train and assess machine learning classifiers.
+We developed an approach designed to address these two problems, and used it to map Ghana’s croplands.
+To overcome the spatio-temporal mismatch, we converted daily, high resolution imagery into two cloud-free composites (the primary growing season and subsequent dry season) covering the 2018 agricultural year, providing a seasonal contrast that helps to improve classification accuracy.
+To address the problem of label availability, we created a platform that rigorously assesses and minimizes label error, and used it to iteratively train a Random Forests classifier with active learning, which identifies the most informative training sample based on prediction uncertainty.
 Minimizing label errors improved model F1 scores by up to 25%.
@@
 We used the resulting 3.7 m map of cropland probabilities within a segmentation algorithm to delineate crop field boundaries.
-Based on an independent map reference sample (n=1,207), the cropland probability and field boundary maps have respective overall accuracies of 88% and 86.7%, user’s accuracies for the cropland class of 61.2% and 78.9%, and producer’s accuracies of 67.3% and 58.2%.
-Using the map reference sample to calculate an unbiased area estimate from the field boundary map, we found that cropland covers 17.1% (15.4-18.9%) of Ghana.
-Using the most accurately digitized labels to calculate and correct for biases in the segmented field boundaries map, we further estimated the average size (1.73 ha) and total number (1,662,281) of crop fields in Ghana.
-Our results demonstrate an adaptable and transferrable approach for mapping the characteristics of croplands on an annual basis and over national extents, with several features that effectively mitigate the errors inherent in remote sensing of smallholder-dominated agriculture.
+Using an independent map reference sample (n=1,207), we found that the cropland probability and field boundary maps had respective overall accuracies of 88% and 86.7%, user’s accuracies for the cropland class of 61.2% and 78.9%, and producer’s accuracies of 67.3% and 58.2%.
+An unbiased area estimate calculated from the map reference sample indicates that cropland covers 17.1% (15.4-18.9%) of Ghana.
+Using the most accurate validation labels to correct for biases in the segmented field boundaries map, we estimated that the average size and total number of field in Ghana are 1.73 ha and 1,662,281, respectively.
+Our results demonstrate an adaptable and transferable approach for developing annual, country-scale maps of crop field boundaries, with several features that effectively mitigate the errors inherent in remote sensing of smallholder-dominated agriculture.
 
@@
 The success of these efforts depends heavily on data that accurately describes the location and characteristics of croplands [@FritzMappingglobalcropland2015], and, given the rapid pace of agricultural change [@bullockThreeDecadesLand2021; @GibbsTropicalforestswere2010; @ZengHighlandcroplandexpansion2018], how these are changing from one year to the next.
-Unfortunately, for many regions, existing cropland datasets are inaccurate, and are usually created as once-off or infrequently update products.
+Unfortunately, for many regions, existing cropland datasets are inaccurate, and are usually created as once-off or infrequently updated products.
 As such, estimates of global cropland area tend to vary widely, often disagree about where croplands are located [e.g.
@@
 We generated a catalog of `r as.integer(nrow(tile_key) * 2)` composite tiles (hereafter simply "tiles") for Ghana, consisting of a seasonal pair for each of the `r nrow(tile_key)` 0.05$^\circ$ tile grid cells covering Ghana.
-To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using a four category that evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image with resolution degraded below the 3.7 m PlanetScope resolution (e.g.
+To assess the quality of the resulting composites, `r tilen` tile grid cells were randomly selected, and two separate observers graded each corresponding seasonal composite using four categories that evaluated the degree of 1) residual cloud and 2) cloud shadow, 3) the number of visible scene boundary artifacts, and 4) the proportion of the image with resolution degraded below the 3.7 m PlanetScope resolution (e.g.
 because of between-date image mis-registrations).
@@
 After these new sites were labelled, they were added to the starter pool of labels, the model was retrained with the larger training set, its performance and prediction uncertainty was reassessed, and a new sample of the most uncertain primary grid cells was again sent for labelling.
-This loop repeated until model performance gains saturated or reached a pre-defined threshold, after which a final map of cropland probabilities was made for the AOI.
+This loop was typically repeated for 3 iterations, after which a final map of cropland probabilities was made.
 
@@
 They then used the polygon drawing tool to digitize the boundaries of all crop fields visible within the PlanetScope overlays that intersect the target grid cell.
-For this project, labellers were instructed to digitize active or recently active crop fields, avoiding tree crops, and fallow or potentially abandoned fields (see SI for digitizing rules) To aid with interpretation, labellers toggled between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
+For this project, labellers were instructed to digitize active or recently active crop fields, avoiding tree crops, and fallow or potentially abandoned fields (see SI for digitizing rules).
+To aid with interpretation, labellers toggled between the PlanetScope renderings and the basemaps to help form a judgement about what constitutes a field.
 The labeller assigned each digitized polygon a class category (e.g.
@@
 P($\theta$\|D) is the probability that the actual cover type is what the labellers who mapped it says it is, while P(W$_i$\|D) is the average score (ranging between 0 and 1) of the accuracy assessment assignments an individual labeller completed within the AOI, and P(W$\theta$\|D, *W*$_i$) is the labeller's label for that pixel.
-This approach therefore used the average assignment quality score for to weight each labeller's label for a given pixel (see SI for further details).
+This approach therefore used the average assignment quality score to weight each labeller's label for a given pixel (see SI for further details).
 Each pixel in the target grid cell was merged using this approach (n = 40000), which helps to minimize individual labellers' errors.
@@
 We used a combination of `GeoTrellis`\footnote{https://github.com/locationtech/geotrellis}, `rasterio`\footnote{https://rasterio.readthedocs.io/en/latest/}, and `RasterFrames`\footnote{https://rasterframes.io/} to derive the features on the fly (which was enabled by converting the composites to Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/}) and convert them into Apache Spark DataFrames.
+
 The extracted features were combined with their corresponding training and validation labels and passed to the machine learning classifier, a `SparkMLlib` implementation of Random Forests [@BreimanRandomForests2001].
-We trained the model with a balanced sample and a tree depth of 15 and total tree number of 60, which initial testing showed to provide a reasonable balance between computational time and performance.
+We trained the model with a balanced sample and a tree depth of 15 and total tree number of 60.
+Initial testing showed that model performance saturated with increasing values of these parameters (cluster failures occurred when tree depths and numbers were simultaneously $\geq16$ and $\geq50$, respectively), and that model stability was satisfactory with these settings, as there was $\leq$ 0.01 difference in accuracy for separate models trained on the same labels.
 
@@
 ## Segmentation
-Upon completion of the active learning process, we deployed a five-step algorithm to create a segmented map of field boundaries.
+Upon completion of the active learning process, we deployed a five-step algorithm to create a segmented map of field boundaries (see Figure S3 for illustration of the steps).
 In the first step, we identified edge features within the imagery.
@@
 
-In the second step, we used a compact watershed algorithm [@neubertCompactWatershedPreemptive2014] to segment the edge image, specifying a high number of segments (6,400) segments per tile, so that the mean segment size (<0.5 ha) was finer than the expected mean field size (>1 ha).
+In the second step, we used a compact watershed algorithm [@neubertCompactWatershedPreemptive2014] to segment the edge image, specifying a high number of segments (6,400) per tile, so that the mean segment size (<0.5 ha) was finer than the expected mean field size (>1 ha).
 
@@
 We calculated these measures across the entire country, as well as several different zones, to evaluate regional difference in accuracy.
-We defined two sets of zonations (Figure S4), each containing four zones, the first created by grouping 1) the three northern AOIs (1-3), 2) the six central AOIs (4-9), 3) the four southwestern AOIs (10, 11, 13, 14, 16), and 4) the two southeastern zones (13, 15).
+We defined two sets of zonations (Figure S5), each containing four zones, the first created by grouping 1) the three northern AOIs (1-3), 2) the six central AOIs (4-9), 3) the four southwestern AOIs (10, 11, 13, 14, 16), and 4) the two southeastern zones (13, 15).
 This grouping differs from the three clusters used to collect initial model training samples, as we designed these to divide the country more finely, and to isolate the less forested southeastern third of Ghana from the more forest northwest.
-The second zonation was developed by grouping the country's eight agro-ecological zones into four broader clusters (Figure S4B).
+The second zonation was developed by grouping the country's eight agro-ecological zones into four broader clusters (Figure S5B).
 We applied this zonation only to the per-pixel classification, to better understand patterns of error in the model.
 
-To assess how effectively the segmentations captured field characteristics, we compared the size class distributions of the segmented field boundaries against the field boundaries digitized by the labellers' over the 100 validation sites in each AOI.
+To assess how effectively the segmentations captured field characteristics, we compared the size class distributions of the segmented field boundaries against those calculated from the field boundaries digitized by the labellers within the 100 validation sites from each AOI.
 We chose this approach because of existing uncertainties in polygon-based accuracy assessment methods [@yeReviewAccuracyAssessment2018], and because the map's ability to represent field sizes was of greatest interest.
@@
 ```
-The assessment of image composites found that their quality in both seasons was highest in the northern half of the country and lowest in the southwest, (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S5).
+The assessment of image composites found that their quality in both seasons was highest in the northern half of the country and lowest in the southwest, (Figure \ref{fig:imqual}A), where the substantially greater cloud cover resulted in a much lower density of available PlanetScope imagery for each time period (Figure S6).
 The average quality score of growing season composites was `r tqmu["Growing"]`, with `r tqct["Growing"]` percent having scores $\geq$ 0.85 (out of 1; Figure \ref{fig:imqual}B), while the mean score of dry season composites was `r tqmu["Dry"]` (`r tqct["Dry"]` percent $\geq$ 0.85).
@@
 AOIs with higher starting performance stopped after fewer iterations, see SI).
-Each AOI's model was trained by 300-500 randomly selected labels (Figure S6A), plus an additional `r aoi10n` - `r aoi15n` (typically `r typicaln`) within each the AOI that were selected by active learning.
-Actively selected labels showed distinctive patterns in several AOIs (Figure S6B), such as concentrating along ecotones or the boundaries of agro-ecological zones.
-A total of `r ttrain` training and `r tval` validation labels were collected by 20 labellers to develop and assess model performance (Figure S7).
+Each AOI's model was trained by 300-500 randomly selected labels (Figure S7A), plus an additional `r aoi10n` - `r aoi15n` (typically `r typicaln`) labels within each AOI that were selected by active learning.
+Actively selected labels showed distinctive patterns in several AOIs (Figure S7B), such as concentrating along ecotones or the boundaries of agro-ecological zones.
+A total of `r ttrain` training and `r tval` validation labels were collected by 20 labellers to develop and assess model performance (Figure S8).
 
@@
 The experiment conducted in three AOIs (in AOIs 1, 8, and 15) showed that training models with active learning improved performance compared to randomized approaches to label selection.
-After three iterations, the accuracy, AUC, and F1 scores for the actively trained models were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than those for randomly trained models (Figure S8).
+After three iterations, the accuracy, AUC, and F1 scores for the actively trained models were respectively `r u(ixl$III["Accuracy"])`, `r u(ixl$III["AUC"])`, and `r u(ixl$III["F1"])` percent higher than those for randomly trained models (Figure S9).
 However, there was more variability in earlier iterations, with average score differences of `r u(ixl$I["Accuracy"])` (accuracy), `r u(ixl$I["AUC"])` (AUC), and `r u(ixl$I["F1"])` percent (F1) after iteration 1, and `r u(ixl$II["Accuracy"])` (accuracy), `r u(ixl$II["AUC"])` (AUC), and `r u(ixl$II["F1"])` (F1) percent after iteration 2 (see SI for more details).
@@
 We used the two measures of label quality calculated by the platform, the average quality score of each labeller and Bayesian Risk (or simply "label risk"), to assess the potential impacts of label error on model performance.
-The average of each labeller's AOI-specific accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figures S4 and S5 for details on label scores and number of assignments per labeller).
-The average Bayesian Risk was `r murisk$Risk`, with highest label risk (`r cl1$Risk`) in the northern AOIs (AOIs 1-6; Figures S6-7), lowest (`r cl1$Risk`) in the southwestern AOIs (AOIs 10, 11, 13, 14, 16), and intermediate (`r cl2$Risk`) in the central-southeastern AOIs (AOIs 7-9, 12, 15).
+The average of each labeller's AOI-specific accuracy score was `r round(mean(label_summary$score_stats$Mean), 2)` (range `r round(min(label_summary$score_stats$Mean), 2)` to `r round(max(label_summary$score_stats$Mean), 2)`; see Figures S7 and S8 for details on label scores and number of assignments per labeller).
+The average Bayesian Risk was `r murisk$Risk`, with highest label risk (`r cl1$Risk`) in the northern AOIs (AOIs 1-6; Figures S10-11), lowest (`r cl1$Risk`) in the southwestern AOIs (AOIs 10, 11, 13, 14, 16), and intermediate (`r cl2$Risk`) in the central-southeastern AOIs (AOIs 7-9, 12, 15).
 
@@
 
-Treating each labeller's average label quality scores (Figure S9) as a proxy for error, we used these scores to develop training sets to test the impact of label error on model performance.
+Treating each labeller's average label quality scores (Figure S10) as a proxy for error, we used these scores to develop training sets to test the impact of label error on model performance.
 The results of these tests, which were conducted in AOIs 1, 2, 8, and 15, showed that the average accuracy, AUC, and F1 scores for models trained with the consensus labels were respectively `r metdiffs$Accuracy$cons`, `r metdiffs$AUC$cons`, and `r metdiffs$F1$cons` (Figure \ref{fig:trainingimpact}).
@@
 
-Correlations (Table S3) between the mean label risk per AOI (Figures S10-11) and model performance metrics showed strong (Spearman's Rank Correlation = `r risk_corr[, "Accuracy"]`) to moderate (r = `r risk_corr[, "AUC"]`) negative correlations between label risk and accuracy and AUC, respectively, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
+Correlations (Table S3) between the mean label risk per AOI (Figures S11-12) and model performance metrics showed strong (Spearman's Rank Correlation = `r risk_corr[, "Accuracy"]`) to moderate (r = `r risk_corr[, "AUC"]`) negative correlations between label risk and accuracy and AUC, respectively, while F1 had a weaker but moderate positive association (r = `r risk_corr[, "F1"]`).
 The positive sign of the latter relationship is counter-intuitive, but is explained by risk's association with precision, one of two inputs to F1, which was moderately positive (r = `r risk_corr[, "Precision"]`), whereas risk had a negligible correlation with recall (r = `r risk_corr[, "Recall"]`), F1's other component.
@@
 The comparisons of digitized versus segmented field boundaries showed that the mean field size across all validation sites averaged `r round(segha, 2)` ha (Median = `r round(seghamed, 2)`; StDev = `r round(seghasd, 2)`), which was `r round((segha - valha) / valha * 100) / 100` times larger than the `r round(valha, 2)` ha (Median = `r round(valhamed, 2)`; StDev = `r round(valhasd, 2)`) mean area of labeller-digitized polygons.
-This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S13), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
-The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S14).
+This discrepancy was primarily caused by results in four AOIs (2, 3, 7, and 15; Figure S14), where segments averaged between `r round(lrgst_seg$rng[1], 2)` and `r round(lrgst_seg$rng[2], 2)` ha, compared to `r round(lrgst_val$rng[1], 2)` - `r round(lrgst_val$rng[2], 2)` ha for the corresponding hand-digitized polygons.
+The number of segmented fields per validation site averaged `r segn` (median = `r segnmed`; StDev = `r segnsd`) compared to `r valn` (median = `r valnmed`; StDev = `r valnsd`) for digitized polygons (Figure S15).
 
@@
 The field boundary map provides additional information on how the characteristics of croplands vary across Ghana, ranging from narrow, strip-like fields in parts of AOI 15 (Figure \ref{fig:mainmap}'s lower right inset) to more densely packed, less distinctly shaped fields in AOI 5 (upper right inset in Figure \ref{fig:mainmap}).
-To explore how field characteristics varied geographically, we mapped the average size and total number of fields within each 0.05 degree tile grid (Figure S15).
-These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts occurring where field densities are higher, although the biases (relative to the validation labels) in both measures (Figures S13-14) complicate interpretations of those variations.
+To explore how field characteristics varied geographically, we mapped the average size and total number of fields within each 0.05 degree tile grid (Figure S16).
+These patterns generally correspond to those seen in the cropland density map (Figure \ref{fig:mainmap}), with larger sizes and field counts occurring where field densities were higher, although the biases (relative to the validation labels) in both measures (Figures S14-15) complicate interpretations of those variations.
 To minimize this complication, we used the calculated biases to develop adjusted estimates of field size and count (Table \ref{tab:sizentab}).
@@
 ## Map accuracy and key sources of error
-Although these maps provide valuable new information, they nevertheless contain substantial errors.
+Although these maps provide valuable new information, they nevertheless contain substantial errors that can impact "downstream" uses (e.g.
+estimating crop production estimates) and decisions based on these maps in unpredictable ways [@Esteslargeareaspatiallycontinuous2018].
 The overall accuracies (86.7-88%, Table \ref{tab:mapaccuracy}) are near the boundary of what might be considered *achievable* map accuracy [@ElmesAccountingtrainingdata2020], given that we only have ~85% confidence in our map reference sample, which is our best estimate of the "truth".
@@
 Maps in the northern savannas had the best balance between omission and commission error, and had the highest overall user's accuracy.
-The transitional zone between forest and savanna had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S4), and thus was under-represented in the training samples, particularly in AOIs 10 and 11 (Figure S6B).
+The transitional zone between forest and savanna had a very low Producer's accuracy (21%), which likely reflects the fact that it was divided between several AOIs for mapping (Figure S4), and thus was under-represented in the training samples, particularly in AOIs 10 and 11 (Figure S7B).
 
 Beyond the errors linked to regional differences, several other important factors contributed to map error.
-The first of these related to the large mapping extent and image resolution.
+The first of these related to the mapping extent and image resolution.
 Given the goal of developing a high resolution, country-scale map, the large data volume constrained us to use a relatively small feature set and less than the recommended tree number and depth [@MaxwellImplementationmachinelearningclassification2018] in our Random Forests models, in order to limit computational costs.
@@
 Smaller field sizes also complicated labelling, as these become increasingly indistinct in the \~4 m PlanetScope composites.
-The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figures S10-11), and by the average assignment quality scores of each labeller (71%; Figure S9).
+The difficulty of labelling is reflected in the magnitude of the Bayesian Risk metrics (Figures S11-12), and by the average assignment quality scores of each labeller (71%; Figure S10).
 Although prior work [@mellorExploringIssuesTraining2015; @rodriguez-galianoAssessmentEffectivenessRandom2012] found that Random Forests are robust to label error, we found that it has substantial impact (Figure \ref{fig:trainingimpact}), which suggests that improving label quality is one of the most important factors in increasing model accuracy.
@@
 This impacted labellers' abilities to discern fields, and doubtless affected model predictions.
-Little can be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which could reduce model performance by removing the temporal contrast [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019], or by adding radar data to the predictor set, which would reduce map resolution.
+Little can be done to mitigate these errors, short of confining imagery to the less cloudy dry season, which could reduce model performance by removing the temporal contrast [@Debatsgeneralizedcomputervision2016; @Defournyrealtimeagriculturemonitoring2019], or by adding radar data (e.g.
+Sentinel-1) to the predictor set, which would reduce map resolution.
 Composite quality could be improved by using imagery from the same seasons over multiple years, but this would undermine the goal of developing annual maps, while the dynamism of the croplands would blur field boundaries within the imagery.
@@
 ## Error mitigation features
-Despite these numerous source of errors, our approach was effective in mitigating several of these error sources.
+Despite these numerous sources of errors, our approach was effective in mitigating several of these error sources.
 Label quality assessment and consensus labelling were the most effective error mitigation tools.
@@
 The quality scores also helped to improve the overall accuracy of consensus labels, by placing higher weight on the work of more accurate labellers.
-In addition to these benefits, label quality scores (Figure S9) also allowed us to select the labels most likely to accurately capture field sizes and numbers, which we used to estimated and correct the biases in these two measures derived from the segmented field boundaries.
+In addition to these benefits, label quality scores (Figure S10) also allowed us to select the labels most likely to accurately capture field sizes and numbers, which we used to estimated and correct the biases in these two measures derived from the segmented field boundaries.
 
@@
 Our large starter sample meant that the models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance.
-Nevertheless, we found higher performance from active learning, most notably in the F1 score (Figure S8), a balanced performance metric, which further demonstrates its effectiveness.
-
-The detail, temporal precision, and large extent of our maps was enabled by the ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
-Daily revisits are important for creating seasonal composites within a single year over cloudy areas.
-The compositing technique we developed allowed us to develop a complete image catalog for the country representing the two seasons for 2018 agricultural year.
-Although Sentinel-2 is free, has better radiometric quality, and has sufficient spatial resolution (10 m) to accurately classify small-scale agricultural systems [e.g.
-@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval may be too infrequent to generate cloud-free composites during the growing season over southern Ghana.
-Sentinel-1 is not affected by the same problem, but labeling fields in radar images is more challenging.
+Nevertheless, we found higher performance from active learning, most notably in the F1 score (Figure S9), a balanced performance metric, which further demonstrates its effectiveness.
+
+The detail, temporal precision, and large extent of our maps was enabled by our use of PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
+These spatial-temporal characteristics, together with the compositing technique we developed, allowed us to develop a complete image catalog for Ghana covering the two major seasons in the 2018 agricultural year.
+Daily revisits were key to this capability, as they increased the number of cloud-free observations that could be collected in each season.
+Over rainy tropical regions, such as southern Ghana, the odds of obtaining a single clear PlanetScope observation within a 1-2 week period is often less than 50% [@royGlobalAnalysisTemporal2021].
+Although Sentinel-2 is free and has sufficient spatial resolution to effectively classify small-scale croplands [e.g.
+@Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is likely too infrequent to generate adequate seasonal composites over much of Ghana.
+The smaller number of clear observations that Sentinel-2 can provide compared to a daily acquisition schedule would result in greater reflectance discontinuities and residual cloud cover in the resulting composites.
+Given Ghana's persistent cloudiness, such artifacts were present in a number of our PlanetScope composites.
+However, these were not large enough to have an appreciable impact on the resulting maps, as the mapping algorithms appeared to be relatively insensitive to these anomalies and any discontinuities along tile boundaries.
 
@@
 This could have confounded the analysis, particularly when comparing the consensus label-trained models with those trained with the most accurate individual labels.
-However, a visual assessment of the resulting probability maps confirms that models trained with the consensus and most accurate individual labels were more precise than the model trained with lower quality labels (Figure S12).
+However, a visual assessment of the resulting probability maps confirms that models trained with the consensus and most accurate individual labels were more precise than the model trained with lower quality labels (Figure S13).
 
@@
 For example, analyzing the distributions of values in the probability maps may provide additional insight into the relative extents of active versus fallow croplands [@Tongforgottenlanduse2020].
-However, use of these data, particularly for decision-making processes (e.g.
-cropped area estimates), should be careful to account for the reported errors [@OlofssonGoodpracticesestimating2014; @StehmanKeyissuesrigorous2019].
+However, use of these data for inventory estimates, to develop other map products, or to guide decision-making should be made with caution and account for the reported errors [@OlofssonGoodpracticesestimating2014; @StehmanKeyissuesrigorous2019; @Esteslargeareaspatiallycontinuous2018].
 

### 2021-10-08 inst/manuscript/manuscript.tex
MS version without EarthArxiv note

@@
             pdfauthor={},
-            pdftitle={High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales},
+            pdftitle={High resolution, annual maps of field boundaries for smallholder-dominated croplands at national scales},
             colorlinks=true,
@@
 
-\title{High resolution, annual maps of the characteristics of
+\title{High resolution, annual maps of field boundaries for
 smallholder-dominated croplands at national scales}
@@
 \begin{center}
-\huge High resolution, annual maps of the characteristics of
+\huge High resolution, annual maps of field boundaries for
 smallholder-dominated croplands at national scales
@@
 }
-
-%\begin{center}
-This pre-print not yet undergone peer review. It has been submitted to
-\emph{Frontiers in Artificial Intelligence}. This version will be
-updated as it is revised, and the final published version will be
-accessible through its DOI link.
-%\end{center}
-
 \end{justify}
@@
 \begin{abstract}
-Mapping the changing characteristics of Africa's smallholder-dominated
-agricultural systems, including the sizes and numbers of fields, is
-crucial for understanding food security and a range of other
-socioeconomic and environmental concerns. However, accurately mapping
-these systems is difficult because of 1) the spatial and temporal
-mismatch between satellite sensors and smallholder fields, and 2) the
-lack of high-quality labels needed to train and assess machine learning
-classifiers. We developed an approach designed to address these two
-problems, which we used to map Ghana's annual croplands for the year
-2018. To overcome the first problem, we converted daily, high resolution
-CubeSat (PlanetScope) imagery into two cloud-free seasonal composites
-covering a single agricultural year. To address the second problem, we
-created a labelling platform that rigorously assesses and minimizes
-label error, and used it to iteratively train a Random Forests
-classifier with active learning, which identifies the most informative
-training sample based on prediction uncertainty. Minimizing label errors
-improved model F1 scores by up to 25\%. Active learning increased F1
-scores by an average of 9.1\% between first and last training
-iterations, and 2.3\% more than models trained with randomly selected
-labels. We used the resulting 3.7 m map of cropland probabilities within
-a segmentation algorithm to delineate crop field boundaries. Based on an
-independent map reference sample (n=1,207), the cropland probability and
-field boundary maps have respective overall accuracies of 88\% and
-86.7\%, user's accuracies for the cropland class of 61.2\% and 78.9\%,
-and producer's accuracies of 67.3\% and 58.2\%. Using the map reference
-sample to calculate an unbiased area estimate from the field boundary
-map, we found that cropland covers 17.1\% (15.4-18.9\%) of Ghana. Using
-the most accurately digitized labels to calculate and correct for biases
-in the segmented field boundaries map, we further estimated the average
-size (1.73 ha) and total number (1,662,281) of crop fields in Ghana. Our
-results demonstrate an adaptable and transferrable approach for mapping
-the characteristics of croplands on an annual basis and over national
-extents, with several features that effectively mitigate the errors
-inherent in remote sensing of smallholder-dominated agriculture.
+Mapping the characteristics of Africa's smallholder-dominated croplands,
+including the sizes and numbers of fields, can provide critical insights
+into food security and a range of other socioeconomic and environmental
+concerns. However, accurately mapping these systems is difficult because
+there is 1) a spatial and temporal mismatch between satellite sensors
+and smallholder fields, and 2) a lack of high-quality labels needed to
+train and assess machine learning classifiers. We developed an approach
+designed to address these two problems, and used it to map Ghana's
+croplands. To overcome the spatio-temporal mismatch, we converted daily,
+high resolution imagery into two cloud-free composites (the primary
+growing season and subsequent dry season) covering the 2018 agricultural
+year, providing a seasonal contrast that helps to improve classification
+accuracy. To address the problem of label availability, we created a
+platform that rigorously assesses and minimizes label error, and used it
+to iteratively train a Random Forests classifier with active learning,
+which identifies the most informative training sample based on
+prediction uncertainty. Minimizing label errors improved model F1 scores
+by up to 25\%. Active learning increased F1 scores by an average of
+9.1\% between first and last training iterations, and 2.3\% more than
+models trained with randomly selected labels. We used the resulting 3.7
+m map of cropland probabilities within a segmentation algorithm to
+delineate crop field boundaries. Using an independent map reference
+sample (n=1,207), we found that the cropland probability and field
+boundary maps had respective overall accuracies of 88\% and 86.7\%,
+user's accuracies for the cropland class of 61.2\% and 78.9\%, and
+producer's accuracies of 67.3\% and 58.2\%. An unbiased area estimate
+calculated from the map reference sample indicates that cropland covers
+17.1\% (15.4-18.9\%) of Ghana. Using the most accurate validation labels
+to correct for biases in the segmented field boundaries map, we
+estimated that the average size and total number of field in Ghana are
+1.73 ha and 1,662,281, respectively. Our results demonstrate an
+adaptable and transferable approach for developing annual, country-scale
+maps of crop field boundaries, with several features that effectively
+mitigate the errors inherent in remote sensing of smallholder-dominated
+agriculture.
 \end{abstract}
@@
 existing cropland datasets are inaccurate, and are usually created as
-once-off or infrequently update products. As such, estimates of global
+once-off or infrequently updated products. As such, estimates of global
 cropland area tend to vary widely, often disagree about where croplands
@@
 two separate observers graded each corresponding seasonal composite
-using a four category that evaluated the degree of 1) residual cloud and
+using four categories that evaluated the degree of 1) residual cloud and
 2) cloud shadow, 3) the number of visible scene boundary artifacts, and
@@
 reassessed, and a new sample of the most uncertain primary grid cells
-was again sent for labelling. This loop repeated until model performance
-gains saturated or reached a pre-defined threshold, after which a final
-map of cropland probabilities was made for the AOI.
+was again sent for labelling. This loop was typically repeated for 3
+iterations, after which a final map of cropland probabilities was made.
 
@@
 crop fields, avoiding tree crops, and fallow or potentially abandoned
-fields (see SI for digitizing rules) To aid with interpretation,
+fields (see SI for digitizing rules). To aid with interpretation,
 labellers toggled between the PlanetScope renderings and the basemaps to
@@
 \emph{W}\(_i\)) is the labeller's label for that pixel. This approach
-therefore used the average assignment quality score for to weight each
+therefore used the average assignment quality score to weight each
 labeller's label for a given pixel (see SI for further details). Each
@@
 to Cloud-optimized Geotiffs\footnote{https://www.cogeo.org/}) and
-convert them into Apache Spark DataFrames. The extracted features were
-combined with their corresponding training and validation labels and
-passed to the machine learning classifier, a \texttt{SparkMLlib}
-implementation of Random Forests (Breiman 2001). We trained the model
-with a balanced sample and a tree depth of 15 and total tree number of
-60, which initial testing showed to provide a reasonable balance between
-computational time and performance.
+convert them into Apache Spark DataFrames.
+
+The extracted features were combined with their corresponding training
+and validation labels and passed to the machine learning classifier, a
+\texttt{SparkMLlib} implementation of Random Forests (Breiman 2001). We
+trained the model with a balanced sample and a tree depth of 15 and
+total tree number of 60. Initial testing showed that model performance
+saturated with increasing values of these parameters (cluster failures
+occurred when tree depths and numbers were simultaneously \(\geq16\) and
+\(\geq50\), respectively), and that model stability was satisfactory
+with these settings, as there was \(\leq\) 0.01 difference in accuracy
+for separate models trained on the same labels.
 
@@
 Upon completion of the active learning process, we deployed a five-step
-algorithm to create a segmented map of field boundaries. In the first
-step, we identified edge features within the imagery. To do this, we
-applied the meanshift algorithm (Yizong Cheng 1995) to each dry-season
-composite tile, and then passed a Sobel filter over the mean-shifted
-green, red, and near-infrared bands, and the corresponding map of
-predicted cropland probabilities. We then summed the four resulting edge
-images to produce a combined edge image.
+algorithm to create a segmented map of field boundaries (see Figure S3
+for illustration of the steps). In the first step, we identified edge
+features within the imagery. To do this, we applied the meanshift
+algorithm (Yizong Cheng 1995) to each dry-season composite tile, and
+then passed a Sobel filter over the mean-shifted green, red, and
+near-infrared bands, and the corresponding map of predicted cropland
+probabilities. We then summed the four resulting edge images to produce
+a combined edge image.
 
@@
 Protzel 2014) to segment the edge image, specifying a high number of
-segments (6,400) segments per tile, so that the mean segment size
-(\textless0.5 ha) was finer than the expected mean field size
-(\textgreater1 ha).
+segments (6,400) per tile, so that the mean segment size (\textless0.5
+ha) was finer than the expected mean field size (\textgreater1 ha).
 
@@
 class-wise User's and Producer's accuracy, and the 95\% confidence
-intervals for each accuracy measure (Olofsson et al. 2013, 2014, Stehman
-and Foody 2019). We calculated these measures across the entire country,
-as well as several different zones, to evaluate regional difference in
-accuracy. We defined two sets of zonations (Figure S4), each containing
-four zones, the first created by grouping 1) the three northern AOIs
-(1-3), 2) the six central AOIs (4-9), 3) the four southwestern AOIs (10,
-11, 13, 14, 16), and 4) the two southeastern zones (13, 15). This
-grouping differs from the three clusters used to collect initial model
-training samples, as we designed these to divide the country more
-finely, and to isolate the less forested southeastern third of Ghana
-from the more forest northwest. The second zonation was developed by
-grouping the country's eight agro-ecological zones into four broader
-clusters (Figure S4B). We applied this zonation only to the per-pixel
-classification, to better understand patterns of error in the model.
+intervals for each accuracy measure (Olofsson et al. 2013, Olofsson et
+al. 2014, Stehman and Foody 2019). We calculated these measures across
+the entire country, as well as several different zones, to evaluate
+regional difference in accuracy. We defined two sets of zonations
+(Figure S5), each containing four zones, the first created by grouping
+1) the three northern AOIs (1-3), 2) the six central AOIs (4-9), 3) the
+four southwestern AOIs (10, 11, 13, 14, 16), and 4) the two southeastern
+zones (13, 15). This grouping differs from the three clusters used to
+collect initial model training samples, as we designed these to divide
+the country more finely, and to isolate the less forested southeastern
+third of Ghana from the more forest northwest. The second zonation was
+developed by grouping the country's eight agro-ecological zones into
+four broader clusters (Figure S5B). We applied this zonation only to the
+per-pixel classification, to better understand patterns of error in the
+model.
 
@@
 characteristics, we compared the size class distributions of the
-segmented field boundaries against the field boundaries digitized by the
-labellers' over the 100 validation sites in each AOI. We chose this
-approach because of existing uncertainties in polygon-based accuracy
-assessment methods (Ye et al. 2018), and because the map's ability to
-represent field sizes was of greatest interest. To undertake this
-comparison, we selected the polygons from the most accurate labeller to
-digitize the 100 validation grids in each AOI, and calculated the
-average area and number of polygons in each cell. We then calculated the
-same statistics from the segmented boundaries that intersected each
-validation grid, and compared the two sets of statistics.
+segmented field boundaries against those calculated from the field
+boundaries digitized by the labellers within the 100 validation sites
+from each AOI. We chose this approach because of existing uncertainties
+in polygon-based accuracy assessment methods (Ye et al. 2018), and
+because the map's ability to represent field sizes was of greatest
+interest. To undertake this comparison, we selected the polygons from
+the most accurate labeller to digitize the 100 validation grids in each
+AOI, and calculated the average area and number of polygons in each
+cell. We then calculated the same statistics from the segmented
+boundaries that intersected each validation grid, and compared the two
+sets of statistics.
 
@@
 greater cloud cover resulted in a much lower density of available
-PlanetScope imagery for each time period (Figure S5). The average
+PlanetScope imagery for each time period (Figure S6). The average
 quality score of growing season composites was 0.88, with 70 percent
@@
 iterations, see SI). Each AOI's model was trained by 300-500 randomly
-selected labels (Figure S6A), plus an additional 600 - 900 (typically
-800) within each the AOI that were selected by active learning. Actively
-selected labels showed distinctive patterns in several AOIs (Figure
-S6B), such as concentrating along ecotones or the boundaries of
+selected labels (Figure S7A), plus an additional 600 - 900 (typically
+800) labels within each AOI that were selected by active learning.
+Actively selected labels showed distinctive patterns in several AOIs
+(Figure S7B), such as concentrating along ecotones or the boundaries of
 agro-ecological zones. A total of 6,299 training and 1,600 validation
 labels were collected by 20 labellers to develop and assess model
-performance (Figure S7).
+performance (Figure S8).
 
@@
 respectively 0.8, 0.6, and 2.3 percent higher than those for randomly
-trained models (Figure S8). However, there was more variability in
+trained models (Figure S9). However, there was more variability in
 earlier iterations, with average score differences of -1.7 (accuracy),
@@
 performance. The average of each labeller's AOI-specific accuracy score
-was 0.71 (range 0.6 to 0.85; see Figures S4 and S5 for details on label
+was 0.71 (range 0.6 to 0.85; see Figures S7 and S8 for details on label
 scores and number of assignments per labeller). The average Bayesian
 Risk was 0.124, with highest label risk (0.165) in the northern AOIs
-(AOIs 1-6; Figures S6-7), lowest (0.165) in the southwestern AOIs (AOIs
-10, 11, 13, 14, 16), and intermediate (0.131) in the
+(AOIs 1-6; Figures S10-11), lowest (0.165) in the southwestern AOIs
+(AOIs 10, 11, 13, 14, 16), and intermediate (0.131) in the
 central-southeastern AOIs (AOIs 7-9, 12, 15).
 
-Treating each labeller's average label quality scores (Figure S9) as a
+Treating each labeller's average label quality scores (Figure S10) as a
 proxy for error, we used these scores to develop training sets to test
@@
 Correlations (Table S3) between the mean label risk per AOI (Figures
-S10-11) and model performance metrics showed strong (Spearman's Rank
+S11-12) and model performance metrics showed strong (Spearman's Rank
 Correlation = -0.824) to moderate (r = -0.568) negative correlations
@@
 polygons. This discrepancy was primarily caused by results in four AOIs
-(2, 3, 7, and 15; Figure S13), where segments averaged between 7.76 and
+(2, 3, 7, and 15; Figure S14), where segments averaged between 7.76 and
 10.76 ha, compared to 2.18 - 2.77 ha for the corresponding
@@
 site averaged 3.08 (median = 2.66; StDev = 2.9) compared to 4.4 (median
-= 3.38; StDev = 4.52) for digitized polygons (Figure S14).
+= 3.38; StDev = 4.52) for digitized polygons (Figure S15).
 
@@
 and total number of fields within each 0.05 degree tile grid (Figure
-S15). These patterns generally correspond to those seen in the cropland
+S16). These patterns generally correspond to those seen in the cropland
 density map (Figure \ref{fig:mainmap}), with larger sizes and field
-counts occurring where field densities are higher, although the biases
-(relative to the validation labels) in both measures (Figures S13-14)
+counts occurring where field densities were higher, although the biases
+(relative to the validation labels) in both measures (Figures S14-15)
 complicate interpretations of those variations. To minimize this
@@
 Although these maps provide valuable new information, they nevertheless
-contain substantial errors. The overall accuracies (86.7-88\%, Table
-\ref{tab:mapaccuracy}) are near the boundary of what might be considered
-\emph{achievable} map accuracy (Elmes et al. 2020), given that we only
-have \textasciitilde85\% confidence in our map reference sample, which
-is our best estimate of the ``truth.'' However, accuracies for the
-cropland class were much lower, falling between 62 (producer's) to 67
-(user's) percent country-wide for the per-pixel map (Table
-\ref{tab:mapaccuracy}), meaning the model produced substantial
+contain substantial errors that can impact ``downstream'' uses
+(e.g.~estimating crop production estimates) and decisions based on these
+maps in unpredictable ways (Estes et al. 2018). The overall accuracies
+(86.7-88\%, Table \ref{tab:mapaccuracy}) are near the boundary of what
+might be considered \emph{achievable} map accuracy (Elmes et al. 2020),
+given that we only have \textasciitilde85\% confidence in our map
+reference sample, which is our best estimate of the ``truth.'' However,
+accuracies for the cropland class were much lower, falling between 62
+(producer's) to 67 (user's) percent country-wide for the per-pixel map
+(Table \ref{tab:mapaccuracy}), meaning the model produced substantial
 commission and omission errors for this class. The segmented boundary
@@
 was under-represented in the training samples, particularly in AOIs 10
-and 11 (Figure S6B).
+and 11 (Figure S7B).
 
@@
 important factors contributed to map error. The first of these related
-to the large mapping extent and image resolution. Given the goal of
-developing a high resolution, country-scale map, the large data volume
-constrained us to use a relatively small feature set and less than the
-recommended tree number and depth (Maxwell et al. 2018) in our Random
-Forests models, in order to limit computational costs. Previous work
-found that Random Forests achieves much better performance on
-small-scale croplands when trained on a much larger number of features
-(Debats et al. 2016, Lebourgeois et al. 2017). However, applying such a
-large feature set within the extent of our AOIs was not possible, as the
-computing time and costs would have been several times
+to the mapping extent and image resolution. Given the goal of developing
+a high resolution, country-scale map, the large data volume constrained
+us to use a relatively small feature set and less than the recommended
+tree number and depth (Maxwell et al. 2018) in our Random Forests
+models, in order to limit computational costs. Previous work found that
+Random Forests achieves much better performance on small-scale croplands
+when trained on a much larger number of features (Debats et al. 2016,
+Lebourgeois et al. 2017). However, applying such a large feature set
+within the extent of our AOIs was not possible, as the computing time
+and costs would have been several times
 larger\footnote{Each active learning iteration ran for $\sim$4-8 hours on 800 CPUs, followed by a final $\sim$10-14 hours for prediction}.
@@
 is reflected in the magnitude of the Bayesian Risk metrics (Figures
-S10-11), and by the average assignment quality scores of each labeller
-(71\%; Figure S9). Although prior work (Rodriguez-Galiano et al. 2012,
+S11-12), and by the average assignment quality scores of each labeller
+(71\%; Figure S10). Although prior work (Rodriguez-Galiano et al. 2012,
 Mellor et al. 2015) found that Random Forests are robust to label error,
@@
 model performance by removing the temporal contrast (Debats et al. 2016,
-Defourny et al. 2019), or by adding radar data to the predictor set,
-which would reduce map resolution. Composite quality could be improved
-by using imagery from the same seasons over multiple years, but this
-would undermine the goal of developing annual maps, while the dynamism
-of the croplands would blur field boundaries within the imagery.
+Defourny et al. 2019), or by adding radar data (e.g.~Sentinel-1) to the
+predictor set, which would reduce map resolution. Composite quality
+could be improved by using imagery from the same seasons over multiple
+years, but this would undermine the goal of developing annual maps,
+while the dynamism of the croplands would blur field boundaries within
+the imagery.
 
@@
 
-Despite these numerous source of errors, our approach was effective in
+Despite these numerous sources of errors, our approach was effective in
 mitigating several of these error sources. Label quality assessment and
@@
 placing higher weight on the work of more accurate labellers. In
-addition to these benefits, label quality scores (Figure S9) also
+addition to these benefits, label quality scores (Figure S10) also
 allowed us to select the labels most likely to accurately capture field
@@
 their impact on performance. Nevertheless, we found higher performance
-from active learning, most notably in the F1 score (Figure S8), a
+from active learning, most notably in the F1 score (Figure S9), a
 balanced performance metric, which further demonstrates its
@@
 The detail, temporal precision, and large extent of our maps was enabled
-by the ability to process PlanetScope data, which is currently the only
-source of sub-5 meter imagery with daily coverage (McCabe et al. 2017).
-Daily revisits are important for creating seasonal composites within a
-single year over cloudy areas. The compositing technique we developed
-allowed us to develop a complete image catalog for the country
-representing the two seasons for 2018 agricultural year. Although
-Sentinel-2 is free, has better radiometric quality, and has sufficient
-spatial resolution (10 m) to accurately classify small-scale
-agricultural systems (e.g. Defourny et al. 2019, Kerner et al. 2020),
-its 5-day interval may be too infrequent to generate cloud-free
-composites during the growing season over southern Ghana. Sentinel-1 is
-not affected by the same problem, but labeling fields in radar images is
-more challenging.
+by our use of PlanetScope data, which is currently the only source of
+sub-5 meter imagery with daily coverage (McCabe et al. 2017). These
+spatial-temporal characteristics, together with the compositing
+technique we developed, allowed us to develop a complete image catalog
+for Ghana covering the two major seasons in the 2018 agricultural year.
+Daily revisits were key to this capability, as they increased the number
+of cloud-free observations that could be collected in each season. Over
+rainy tropical regions, such as southern Ghana, the odds of obtaining a
+single clear PlanetScope observation within a 1-2 week period is often
+less than 50\% (Roy et al. 2021). Although Sentinel-2 is free and has
+sufficient spatial resolution to effectively classify small-scale
+croplands (e.g. Defourny et al. 2019, Kerner et al. 2020), its 5-day
+interval is likely too infrequent to generate adequate seasonal
+composites over much of Ghana. The smaller number of clear observations
+that Sentinel-2 can provide compared to a daily acquisition schedule
+would result in greater reflectance discontinuities and residual cloud
+cover in the resulting composites. Given Ghana's persistent cloudiness,
+such artifacts were present in a number of our PlanetScope composites.
+However, these were not large enough to have an appreciable impact on
+the resulting maps, as the mapping algorithms appeared to be relatively
+insensitive to these anomalies and any discontinuities along tile
+boundaries.
 
@@
 consensus and most accurate individual labels were more precise than the
-model trained with lower quality labels (Figure S12).
+model trained with lower quality labels (Figure S13).
 
@@
 provide additional insight into the relative extents of active versus
-fallow croplands (Tong et al. 2020). However, use of these data,
-particularly for decision-making processes (e.g.~cropped area
-estimates), should be careful to account for the reported errors
-(Olofsson et al. 2014, Stehman and Foody 2019).
+fallow croplands (Tong et al. 2020). However, use of these data for
+inventory estimates, to develop other map products, or to guide
+decision-making should be made with caution and account for the reported
+errors (Olofsson et al. 2014, Estes et al. 2018, Stehman and Foody
+2019).
 
@@
 Turia, G. Gamoga, H. Abe, E. Kaidong, and G. Miceli. 2016. Collect
-{Earth}: {Land Use} and {Land Cover Assessment} through {Augmented
+{Earth}: Land {Use} and {Land Cover Assessment} through {Augmented
 Visual Interpretation}. Remote Sensing 8:807.
@@
 conflict between omission and commission in low spatial resolution
-dichotomic thematic products: {The Pareto Boundary}. Remote Sensing of
+dichotomic thematic products: The {Pareto Boundary}. Remote Sensing of
 Environment 91:280--292.
@@
 Carletto, C., S. Gourlay, and P. Winters. 2015. From {Guesstimates} to
-{GPStimates}: {Land Area Measurement} and {Implications} for
+{GPStimates}: Land {Area Measurement} and {Implications} for
 {Agricultural Analysis}. Journal of African Economies 24:593--628.
@@
 \leavevmode\hypertarget{ref-CarlettoFactartifactimpact2013}{}%
-Carletto, C., S. Savastano, and A. Zezza. 2013. Fact or artifact: {The}
+Carletto, C., S. Savastano, and A. Zezza. 2013. Fact or artifact: The
 impact of measurement errors on the farm size{}productivity
@@
 Peng, M. Lu, W. Zhang, X. Tong, and J. Mills. 2015. Global land cover
-mapping at 30 m resolution: {A POK}-based operational approach. ISPRS
+mapping at 30 m resolution: A {POK}-based operational approach. ISPRS
 Journal of Photogrammetry and Remote Sensing 103:7--27.
@@
 Koetz. 2019. Near real-time agriculture monitoring at national scale at
-parcel resolution: {Performance} assessment of the {Sen2}-{Agri}
-automated system in various cropping systems around the world. Remote
-Sensing of Environment 221:551--568.
+parcel resolution: Performance assessment of the {Sen2}-{Agri} automated
+system in various cropping systems around the world. Remote Sensing of
+Environment 221:551--568.
 
 \leavevmode\hypertarget{ref-desiereLandProductivityPlot2018}{}%
-Desiere, S., and D. Jolliffe. 2018. Land productivity and plot size:
-{Is} measurement error driving the inverse relationship? Journal of
+Desiere, S., and D. Jolliffe. 2018. Land productivity and plot size: Is
+measurement error driving the inverse relationship? Journal of
 Development Economics 130:84--98.
@@
 Dwyer, J. L., D. P. Roy, B. Sauer, C. B. Jenkerson, H. K. Zhang, and L.
-Lymburner. 2018. Analysis {Ready Data}: {Enabling Analysis} of the
+Lymburner. 2018. Analysis {Ready Data}: Enabling {Analysis} of the
 {Landsat Archive}. Remote Sensing 10:1363.
@@
 Feder, G. 1985. The relation between farm size and farm productivity:
-{The} role of family labor, supervision and credit constraints. Journal
-of Development Economics 18:297--313.
+The role of family labor, supervision and credit constraints. Journal of
+Development Economics 18:297--313.
 
@@
 Fritz, S., I. McCallum, C. Schill, C. Perger, L. See, D. Schepaschenko,
-M. van der Velde, F. Kraxner, and M. Obersteiner. 2012. Geo-{Wiki}: {An}
+M. van der Velde, F. Kraxner, and M. Obersteiner. 2012. Geo-{Wiki}: An
 online platform for improving global land cover. Environmental Modelling
@@
 Gorelick, N., M. Hancher, M. Dixon, S. Ilyushchenko, D. Thau, and R.
-Moore. 2017. Google {Earth Engine}: {Planetary}-scale geospatial
-analysis for everyone. Remote Sensing of Environment 202:18--27.
+Moore. 2017. Google {Earth Engine}: Planetary-scale geospatial analysis
+for everyone. Remote Sensing of Environment 202:18--27.
 
@@
 Hamrouni, Y., E. Paillassa, V. Chéret, C. Monteil, and D. Sheeren. 2021.
-From local to global: {A} transfer learning-based approach for mapping
+From local to global: A transfer learning-based approach for mapping
 poplar plantations at national scale using {Sentinel}-2. ISPRS Journal
@@
 Lobell, D. B., K. G. Cassman, and C. B. Field. 2009. Crop {Yield Gaps}:
-{Their Importance}, {Magnitudes}, and {Causes}. Annual Review of
+Their {Importance}, {Magnitudes}, and {Causes}. Annual Review of
 Environment and Resources 34:179--204.
@@
 Ma, L., Y. Liu, X. Zhang, Y. Ye, G. Yin, and B. A. Johnson. 2019. Deep
-learning in remote sensing applications: {A} meta-analysis and review.
+learning in remote sensing applications: A meta-analysis and review.
 ISPRS Journal of Photogrammetry and Remote Sensing 152:166--177.
@@
 Neubert, P., and P. Protzel. 2014. Compact {Watershed} and {Preemptive
-SLIC}: {On Improving Trade}-offs of {Superpixel Segmentation
+SLIC}: On {Improving Trade}-offs of {Superpixel Segmentation
 Algorithms}. Pages 996--1001 2014 22nd {International Conference} on
@@
 Olofsson, P., G. M. Foody, S. V. Stehman, and C. E. Woodcock. 2013.
-Making better use of accuracy data in land change studies: {Estimating}
+Making better use of accuracy data in land change studies: Estimating
 accuracy and area and quantifying uncertainty using stratified
@@
 Qiu, S., Z. Zhu, and C. E. Woodcock. 2020. Cirrus clouds that adversely
-affect {Landsat} 8 images: {What} are they and how to detect them?
-Remote Sensing of Environment 246:111884.
+affect {Landsat} 8 images: What are they and how to detect them? Remote
+Sensing of Environment 246:111884.
 
@@
 Searchinger, T., R. Waite, C. Hanson, J. Ranganathan, P. Dumas, E.
-Matthews, and C. Klirs. 2019. Creating a sustainable food future: {A}
-menu of solutions to feed nearly 10 billion people by 2050. {Final}
-report. {WRI}.
+Matthews, and C. Klirs. 2019. Creating a sustainable food future: A menu
+of solutions to feed nearly 10 billion people by 2050. {Final} report.
+{WRI}.
 
@@
 2019. Hierarchical mapping of annual global land cover 2001 to present:
-{The MODIS Collection} 6 {Land Cover} product. Remote Sensing of
+The {MODIS Collection} 6 {Land Cover} product. Remote Sensing of
 Environment 222:183--194.
@@
 Rasmussen, F. Tian, T. Tagesson, W. Zhang, and R. Fensholt. 2020. The
-forgotten land use class: {Mapping} of fallow fields across the {Sahel}
+forgotten land use class: Mapping of fallow fields across the {Sahel}
 using {Sentinel}-2. Remote Sensing of Environment 239:111598.
@@
 Waldner, F., and F. I. Diakogiannis. 2020. Deep learning on edge:
-{Extracting} field boundaries from satellite images with a convolutional
+Extracting field boundaries from satellite images with a convolutional
 neural network. Remote Sensing of Environment 245:111741.
@@
 Belward, W. B. Cohen, E. A. Fosnight, J. Shaw, J. G. Masek, and D. P.
-Roy. 2016. The global {Landsat} archive: {Status}, consolidation, and
+Roy. 2016. The global {Landsat} archive: Status, consolidation, and
 direction. Remote Sensing of Environment 185:271--283.
@@
 Ye, S., R. G. Pontius, and R. Rakshit. 2018. A review of accuracy
-assessment for object-based image analysis: {From} per-pixel to
+assessment for object-based image analysis: From per-pixel to
 per-polygon approaches. ISPRS Journal of Photogrammetry and Remote

### 2021-10-08 inst/manuscript/manuscript.Rmd
EarthArxiv statement added back

@@
 footerdate: no
+
+note: This pre-print is a revised version of a manuscript that has undergone one round of peer review in \emph{Frontiers in Artificial Intelligence}, where it has been resubmitted.
+This version will be updated as it is further revised, and the final published version will be accessible through its DOI link.
 

### 2021-10-08 inst/manuscript/manuscript.tex
EarthArxiv statement added back

@@
 }
+
+%\begin{center}
+This pre-print is a revised version of a manuscript that has undergone
+one round of peer review in \emph{Frontiers in Artificial Intelligence},
+where it has been resubmitted. This version will be updated as it is
+further revised, and the final published version will be accessible
+through its DOI link.
+%\end{center}
+
 \end{justify}

### 2021-10-16 inst/manuscript/manuscript.Rmd
Frontiers second revision

--- 2021-10-08 inst/manuscript/manuscript.Rmd
@@
 The detail, temporal precision, and large extent of our maps was enabled by our use of PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017].
-These spatial-temporal characteristics, together with the compositing technique we developed, allowed us to develop a complete image catalog for Ghana covering the two major seasons in the 2018 agricultural year.
+These spatial-temporal characteristics, together with the compositing technique we developed, allowed us to create a complete image catalog for Ghana covering the two major seasons in the 2018 agricultural year.
 Daily revisits were key to this capability, as they increased the number of cloud-free observations that could be collected in each season.
@@
 
-## Broader applications
-This work demonstrates a proof of concept for developing high resolution, annual maps of smallholder-dominated croplands at national to regional scales, using an approach that follows recommended best practices for training and assessing machine learning models [@ElmesAccountingtrainingdata2020].
-This approach can be readily updated to integrate improvements, such as newer machine learning models.
-Beyond providing valuable insights into field characteristics, field boundary maps can help improve remote estimation of crop areas and yield [e.g.
+## Conclusion
+This work demonstrates a proof of concept for developing high resolution, annual maps of field boundaries in smallholder-dominated croplands at national to regional scales.
+The approach we used to develop these maps includes a novel procedure for compositing daily PlanetScope imagery, and follows recommended best practices for training and assessing machine learning models [@ElmesAccountingtrainingdata2020].
+Using PlanetScope data enabled us to make a country-wide, high-resolution, two-season catalog of imagery for a single year, even over areas with frequent cloud cover.
+Methods to assess labeller accuracy and to create consensus labels helped to  minimize training label errors, which in turn substantially improved the accuracy of the mapping model.
+Model performance was further enhanced by using an active learning framework, which selects the most informative training sample.
+This mapping approach can be readily updated to integrate improvements, such as newer machine learning models.
+
+The resulting maps provide valuable information on Ghana's agricultural systems, including an updated estimate of cropped area, and new data on the characteristics of Ghana's cropping systems, such as the sizes and numbers of fields.
+Beyond the  information about field characteristics, field boundary maps can help improve remote estimation of crop areas and yield [e.g.
 @EstesProjectedclimateimpacts2013], and provide deeper insights into important socioeconomic aspects of agricultural systems, such as the relationships between agricultural productivity and farm size [@CarlettoFactartifactimpact2013; @desiereLandProductivityPlot2018; @federRelationFarmSize1985].
@@
 
-## Data availability and usage
+# Data availability and usage
 The maps presented here are a version 1 product that is freely available to use, along with its underlying code (see SI).

### 2021-10-16 inst/manuscript/manuscript.tex
Frontiers second revision, latex

--- 2021-10-08 inst/manuscript/manuscript.tex
@@
 spatial-temporal characteristics, together with the compositing
-technique we developed, allowed us to develop a complete image catalog
+technique we developed, allowed us to create a complete image catalog
 for Ghana covering the two major seasons in the 2018 agricultural year.
@@
 
-\hypertarget{broader-applications}{%
-\subsection{Broader applications}\label{broader-applications}}
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
 
 This work demonstrates a proof of concept for developing high
-resolution, annual maps of smallholder-dominated croplands at national
-to regional scales, using an approach that follows recommended best
-practices for training and assessing machine learning models (Elmes et
-al. 2020). This approach can be readily updated to integrate
-improvements, such as newer machine learning models. Beyond providing
-valuable insights into field characteristics, field boundary maps can
-help improve remote estimation of crop areas and yield (e.g. Estes et
-al. 2013), and provide deeper insights into important socioeconomic
-aspects of agricultural systems, such as the relationships between
-agricultural productivity and farm size (Feder 1985, Carletto et al.
-2013, Desiere and Jolliffe 2018). Such maps will be important for
-understanding the rapid agricultural change that is currently occurring
-in Africa.
+resolution, annual maps of field boundaries in smallholder-dominated
+croplands at national to regional scales. The approach we used to
+develop these maps includes a novel procedure for compositing daily
+PlanetScope imagery, and follows recommended best practices for training
+and assessing machine learning models (Elmes et al. 2020). Using
+PlanetScope data enabled us to make a country-wide, high-resolution,
+two-season catalog of imagery for a single year, even over areas with
+frequent cloud cover. Methods to assess labeller accuracy and to create
+consensus labels helped to minimize training label errors, which in turn
+substantially improved the accuracy of the mapping model. Model
+performance was further enhanced by using an active learning framework,
+which selects the most informative training sample. This mapping
+approach can be readily updated to integrate improvements, such as newer
+machine learning models.
+
+The resulting maps provide valuable information on Ghana's agricultural
+systems, including an updated estimate of cropped area, and new data on
+the characteristics of Ghana's cropping systems, such as the sizes and
+numbers of fields. Beyond the information about field characteristics,
+field boundary maps can help improve remote estimation of crop areas and
+yield (e.g. Estes et al. 2013), and provide deeper insights into
+important socioeconomic aspects of agricultural systems, such as the
+relationships between agricultural productivity and farm size (Feder
+1985, Carletto et al. 2013, Desiere and Jolliffe 2018). Such maps will
+be important for understanding the rapid agricultural change that is
+currently occurring in Africa.
 
 \hypertarget{data-availability-and-usage}{%
-\subsection{Data availability and
+\section{Data availability and
 usage}\label{data-availability-and-usage}}

### 2021-10-16 inst/manuscript/manuscript.Rmd
Proper section headers added for conclusion

@@
 
-## Conclusion
+# Conclusion
 This work demonstrates a proof of concept for developing high resolution, annual maps of field boundaries in smallholder-dominated croplands at national to regional scales.

### 2022-02-06 R/accuracy_report.R
Commit a number of scripts and datasets

@@ -18,7 +18,6 @@
 #' areas <- c(22353, 1122543, 610228)
 #' acc_mat <- accuracy(err_mat, areas)
 #' accuracy_report(err_mat, acc_mat, classes = paste0("Class", 1:3))
-
 accuracy_report <- function(err_mat, acc_mat, classes, digits = 1, z = 1.96) {
 
   rnd <- function(tab, d = digits) sprintf(glue("%0.{d}f"), tab)

### 2022-02-06 R/consensus_map_functions.R

@@ -19,7 +19,7 @@ get_db_tables <- function(connection, dbtables = "default") {
   return(dbtbls)
 }
 
-#' Collection assignment history for a given worker
+#' Collect assignment history for a given worker
 #' @param assignment_data Assignment data table
 #' @param accuracy_data Accuracy data table
 #' @param workerid Integer ID of worker
@@ -46,7 +46,7 @@ assignment_history <- function(assignment_data, accuracy_data, workerid) {
   return(measurements)
 }
 
-#' Collection qualification assignment history for a given worker
+#' Collect qualification assignment history for a given worker
 #' @param qual_assignment_data Assignment data table
 #' @param qual_accuracy_data Qualification accuracy table
 #' @param workerid Integer ID of worker
@@ -186,7 +186,8 @@ get_user_maps <- function(assignmentid, categories, category, grid_poly,
 #' @param crs Coordinate reference system, current 4326, otherwise proj4string
 #' @return List containing rasters (label, heat map, risk map) and Bayesian
 #' measures of uncertainty.
-#' @details This
+#' @details This function is a modification of the function in rmapaccuracy,
+#' which is designed to allowed consensus labels to be made as a batch.
 #' @import dplyr
 #' @importFrom sf st_sf st_sfc st_is_empty st_union st_bbox
 #' @importFrom data.table as.data.table

### 2022-02-06 R/data.R

@@ -93,6 +93,17 @@ NULL
 #' data(mapped_areas)
 NULL
 
+#' Summary statistics for digitized versus machine-segmented fields
+#' @name segment_quality_stats
+#' @docType data
+#' @description Mean, median areas and average counts of labeller-digitized
+#' fields versus machine-segmented fields at the grid, tile, and AOI scales
+#' @keywords data
+#' @examples
+#' data(segment_quality_stats)
+NULL
+
+
 #' Training and validation sites for AOIs 1-16
 #' @name tile_key
 #' @docType data

### 2022-02-06 data-raw/cropland_fractions.R
Additional data files and scripts added

@@ -1,4 +1,4 @@
-# Create probability map (grid scale) from RF 3 m probability images
+# Create grid scale fractional map from RF 3 m probability images
 # Ran on rstudio instance on EC2, to avoid download into S3
 
 library(sf)
@@ -65,7 +65,9 @@ cropfrac_005 <- do.call(terra::mosaic, cropfrac_aoi)
 # probs_005[probs_005 > 100] <- NA
 
 
-save(cropfrac_aoi, file = here::here("inst/extdata/aoi_probs.rda"))
+# save(cropfrac_aoi, file = here::here("inst/extdata/aoi_probs.rda"))
+save(cropfrac_aoi, file = here::here("inst/extdata/cropfrac_aoi.rda"))
 writeRaster(cropfrac_005,
-            filename = here::here("inst/extdata/probs_005.tif"))
+            # filename = here::here("inst/extdata/probs_005.tif"))
+            filename = here::here("inst/extdata/cropfrac_005.tif"))
 

### 2022-02-06 inst/manuscript/manuscript.Rmd

@@
 typicaln <- raster::modal(initial$n) + raster::modal(active$n)
+typicala <- raster::modal(active$n)
+aoi10a <- active %>% filter(aoi == "labeller10") %>% pull(n)
 aoi10n <- (initial %>% filter(Cluster %in% 3) %>% pull(n)) + 
@@
   (active %>% filter(aoi == "labeller15") %>% pull(n))
+aoi15a <- active %>% filter(aoi == "labeller15") %>% pull(n)
 aoi3init <- initial %>% filter(!Cluster %in% 1:3) %>% pull(n)
@@
 AOIs with higher starting performance stopped after fewer iterations, see SI).
-Each AOI's model was trained by 300-500 randomly selected labels (Figure S7A), plus an additional `r aoi10n` - `r aoi15n` (typically `r typicaln`) labels within each AOI that were selected by active learning.
+Each AOI's model was trained by 300-500 randomly selected labels (Figure S7A), plus an additional `r aoi10a` - `r aoi15a` (typically `r typicala`) labels within each AOI that were selected by active learning.
 Actively selected labels showed distinctive patterns in several AOIs (Figure S7B), such as concentrating along ecotones or the boundaries of agro-ecological zones.

### 2022-02-06 inst/manuscript/si.Rmd

@@
 ---
-title: "High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales"
-
+title: "High resolution, annual maps of field boundaries for  smallholder-dominated croplands at national scales"
 subtitle: "Supporting Information"
@@
 spacing: doublespacing
-# abstract: The abstract can go either here or below
 keywords: rmarkdown, reproducible science
@@
 \bleft
+
+# Code and Data Availability
+The repositories containing the data, code, and manuscript source files used to produce this paper are listed below:
+
+- [`activemapper`](https://github.com/agroimpacts/activemapper): Repository\footnote{github.com/agroimpacts/activemapper} containing code and derived data used to write this paper. 
+- [`imager`](https:github.com/agroimpacts/imager): Repository\footnote{github.com/agroimpacts/imager} containing code for the image processing software.
+- [`labeller`](https:github.com/agroimpacts/labeller):  Repository\footnote{github.com/agroimpacts/labeller} containing code for the labelling platform. 
+- [`learner`](https:github.com/agroimpacts/learner): Repository\footnote{github.com/agroimpacts/learner} containing code for the active learning component.
+- [`segmenter`](https:github.com/agroimpacts/segmenter): Repository\footnote{github.com/agroimpacts/segmenter} containing code for the segmentation algorithm.
+- [`mappingafrica.io`](https://mappingafrica.io): Project website\footnote{https://mappingafrica.io} linking to web map displaying field boundary maps.
+The web map is currently undergoing redevelopment, and will be hosted on a new URL and links download maps.
+In the interim, map data are available on request.
 
@@
 
-# Code and Data Availability
-The repositories containing the data, code, and manuscript source files used to produce this paper are listed below:
+\clearpage
 
-- [`activemapper`](https://github.com/agroimpacts/activemapper): Contains the source code and derived data for this paper.
-The repository\footnote{github.com/agroimpacts/activemapper} will be opened following publication in a peer-reviewed journal.
-- [`croplandmaps`](https://github.com/agroimpacts/croplandmaps_V2): Source code for web map displaying image composites and cropland probability maps.
-The repository\footnote{github.com/agroimpacts/croplandmaps\_V2} is open, and the field boundary datasets are available for download.
-- [`imager`](https:github.com/agroimpacts/imager): Source code for the image processing software.
-The repository\footnote{github.com/agroimpacts/imager} is currently private, but will be opened after completing a security audit.
-- [`labeller`](https:github.com/agroimpacts/labeller): Source code for the labelling platform.
-The repository\footnote{github.com/agroimpacts/labeller} is currently private, but will be opened after completing a security audit.
-- [`learner`](https:github.com/agroimpacts/learner): Source code for the active learning component. The repository\footnote{github.com/agroimpacts/learner} is open.
-- [`segmenter`](https:github.com/agroimpacts/segmenter): Source code for the segmentation algorithm. The repository \footnote{github.com/agroimpacts/segmenter} is open.
-
-\clearpage
 # References

### 2022-02-06 inst/manuscript/si.tex

@@
 \hypersetup{
-  pdftitle={High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales},
+  pdftitle={High resolution, annual maps of field boundaries for smallholder-dominated croplands at national scales},
   pdfkeywords={rmarkdown, reproducible science},
@@
 \bleft
+
+\hypertarget{code-and-data-availability}{%
+\section{Code and Data Availability}\label{code-and-data-availability}}
+
+The repositories containing the data, code, and manuscript source files
+used to produce this paper are listed below:
+
+\begin{itemize}
+\tightlist
+\item
+  \href{https://github.com/agroimpacts/activemapper}{\texttt{activemapper}}:
+  Repository\footnote{github.com/agroimpacts/activemapper} containing
+  code and derived data used to write this paper.
+\item
+  \href{https:github.com/agroimpacts/imager}{\texttt{imager}}:
+  Repository\footnote{github.com/agroimpacts/imager} containing code for
+  the image processing software.
+\item
+  \href{https:github.com/agroimpacts/labeller}{\texttt{labeller}}:
+  Repository\footnote{github.com/agroimpacts/labeller} containing code
+  for the labelling platform.
+\item
+  \href{https:github.com/agroimpacts/learner}{\texttt{learner}}:
+  Repository\footnote{github.com/agroimpacts/learner} containing code
+  for the active learning component.
+\item
+  \href{https:github.com/agroimpacts/segmenter}{\texttt{segmenter}}:
+  Repository\footnote{github.com/agroimpacts/segmenter} containing code
+  for the segmentation algorithm.
+\item
+  \href{https://mappingafrica.io}{\texttt{mappingafrica.io}}: Project
+  website\footnote{https://mappingafrica.io} linking to web map
+  displaying field boundary maps. The web map is currently undergoing
+  redevelopment, and will be hosted on a new URL and links download
+  maps. In the interim, map data are available on request.
+\end{itemize}
 
@@
 
-\hypertarget{code-and-data-availability}{%
-\section{Code and Data Availability}\label{code-and-data-availability}}
-
-The repositories containing the data, code, and manuscript source files
-used to produce this paper are listed below:
-
-\begin{itemize}
-\tightlist
-\item
-  \href{https://github.com/agroimpacts/activemapper}{\texttt{activemapper}}:
-  Contains the source code and derived data for this paper. The
-  repository\footnote{github.com/agroimpacts/activemapper} will be
-  opened following publication in a peer-reviewed journal.
-\item
-  \href{https://github.com/agroimpacts/croplandmaps_V2}{\texttt{croplandmaps}}:
-  Source code for web map displaying image composites and cropland
-  probability maps. The
-  repository\footnote{github.com/agroimpacts/croplandmaps\_V2} is open,
-  and the field boundary datasets are available for download.
-\item
-  \href{https:github.com/agroimpacts/imager}{\texttt{imager}}: Source
-  code for the image processing software. The
-  repository\footnote{github.com/agroimpacts/imager} is currently
-  private, but will be opened after completing a security audit.
-\item
-  \href{https:github.com/agroimpacts/labeller}{\texttt{labeller}}:
-  Source code for the labelling platform. The
-  repository\footnote{github.com/agroimpacts/labeller} is currently
-  private, but will be opened after completing a security audit.
-\item
-  \href{https:github.com/agroimpacts/learner}{\texttt{learner}}: Source
-  code for the active learning component. The
-  repository\footnote{github.com/agroimpacts/learner} is open.
-\item
-  \href{https:github.com/agroimpacts/segmenter}{\texttt{segmenter}}:
-  Source code for the segmentation algorithm. The repository
-  \footnote{github.com/agroimpacts/segmenter} is open.
-\end{itemize}
-
 \clearpage
